### 2017-05-15 Notes/detrendify_spingarn.tex
Initial commit

%
% A simple LaTeX template for Books
%  (c) Aleksander Morgado <aleksander@es.gnu.org>
%  Released into public domain
%

\documentclass{article}
\usepackage[a4paper, top=3cm, bottom=3cm]{geometry}
\usepackage[latin1]{inputenc}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{tocloft}
\usepackage[usenames]{color}
\RequirePackage[OT1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{graphicx}
\usepackage{booktabs}

%% For special lists like inparaenum, compactenum, compactitem
\usepackage{paralist}
%% For \boldsymbol
\usepackage{amsbsy}

%% For \mathscr
\usepackage[mathscr]{eucal}

%% For \bm (bold math)
\usepackage{bm}

%% For proper spacing in text macros
\usepackage{xspace}

%% For \subfloat command
\usepackage{subfig}

%% Algorithm
\usepackage{algorithm,algorithmic}
%\usepackage{algpseudocode}

%% Useful shorthand for converting equations to be inline
\newenvironment{inlinemath}{$}{$}

%% Define hypen
\mathchardef\mhyphen="2D

%% For hyperlinks. Should always be the last package. Omit for DVI version.
\ifpdf
\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\else
\newcommand{\href}[2]{{#2}}
\usepackage{url}
\fi

%% ----------------------------------------------------------------------
%% Hyper-linked References
%% ----------------------------------------------------------------------

\ifpdf
\newcommand{\Sec}[1]{\hyperref[sec:#1]{\S\ref*{sec:#1}}} %section
\newcommand{\App}[1]{\hyperref[sec:#1]{Appendix~\ref*{sec:#1}}} %appendix
\newcommand{\Eqn}[1]{\hyperref[eq:#1]{{\rm (\ref*{eq:#1})}}} %equation
\newcommand{\Part}[1]{\hyperref[part:#1]{(\ref*{part:#1})}} %part of theorem
\newcommand{\Fig}[1]{\hyperref[fig:#1]{Figure~\ref*{fig:#1}}} %figure
\newcommand{\Tab}[1]{\hyperref[tab:#1]{Table~\ref*{tab:#1}}} %table
\newcommand{\Thm}[1]{\hyperref[thm:#1]{Theorem~\ref*{thm:#1}}} %theorem
\newcommand{\Lem}[1]{\hyperref[lem:#1]{Lemma~\ref*{lem:#1}}} %lemma
\newcommand{\Prop}[1]{\hyperref[prop:#1]{Proposition~\ref*{prop:#1}}} %proposition
\newcommand{\Cor}[1]{\hyperref[cor:#1]{Corollary~\ref*{cor:#1}}} %corollary
\newcommand{\Def}[1]{\hyperref[def:#1]{Definition~\ref*{def:#1}}} %definition
\newcommand{\Alg}[1]{\hyperref[alg:#1]{Algorithm~\ref*{alg:#1}}} %algorithm
\newcommand{\Ex}[1]{\hyperref[ex:#1]{Example~\ref*{ex:#1}}} %example
\newcommand{\As}[1]{\hyperref[as:#1]{Assumption~{\rm\ref*{as:#1}}}} %assumption
\newcommand{\Reg}[1]{\hyperref[as:#1]{Condition~\ref*{reg:#1}}} %regularity condition
\newcommand{\AlgLine}[2]{\hyperref[alg:#1]{line~\ref*{line:#2} of Algorithm~\ref*{alg:#1}}}
\newcommand{\AlgLines}[3]{\hyperref[alg:#1]{lines~\ref*{line:#2}--\ref*{line:#3} of Algorithm~\ref*{alg:#1}}}
\else
\newcommand{\Sec}[1]{{\S\ref{sec:#1}}} %section
\newcommand{\App}[1]{{Appendix~\ref{sec:#1}}} %appendix
\newcommand{\Eqn}[1]{{(\ref{eq:#1})}} %equation
\newcommand{\Part}[1]{{(\ref{part:#1})}} %part of proof
\newcommand{\Fig}[1]{{Figure~\ref{fig:#1}}} %figure
\newcommand{\Tab}[1]{{Table~\ref{tab:#1}}} %table
\newcommand{\Thm}[1]{{Theorem~\ref{thm:#1}}} %theorem
\newcommand{\Lem}[1]{{Lemma~\ref{lem:#1}}} %lemma
\newcommand{\Prop}[1]{{Proposition~\ref{prop:#1}}} %proposition
\newcommand{\Cor}[1]{{Corollary~\ref{cor:#1}}} %corollary
\newcommand{\Def}[1]{{Definition~\ref{def:#1}}} %definition
\newcommand{\Alg}[1]{{Algorithm~\ref{alg:#1}}} %algorithm
\newcommand{\Ex}[1]{{Example~\ref{ex:#1}}} %example
\newcommand{\Reg}[1]{{R~\ref*{reg:#1}}} %regularity condition
\fi

%% ----------------------------------------------------------------------
%% Definitions
%% ----------------------------------------------------------------------

%% Shorthand for Real
\newcommand{\Real}{\mathbb{R}}
\newcommand{\dom}{{\bf dom}\,}

%% Math notation
\newcommand{\Tra}{^{\sf T}} % Transpose
\newcommand{\Inv}{^{-1}} % Inverse
\newcommand{\tr}{\operatorname{tr}} % Trace
\newcommand{\Ker}{\operatorname{Ker}} % Kernel
\def\vec{\mathop{\rm vec}\nolimits}
\def\mtcz{\mathop{\rm mat}\nolimits} % Matricize operation
\newcommand{\amp}{\mathop{\:\:\,}\nolimits}
\newcommand{\bic}{\text{BIC}}

%% Vectors
\newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
\newcommand{\Vbar}[1]{{\bm{\M{A}r \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vn}[2]{\V{#1}^{(#2)}} % n-th vector
\newcommand{\VnE}[3]{{#1}^{(#2)}_{#3}} % n-th vector
\newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}} % vector element

%% Matrices
\newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} % matrix element
\newcommand{\MC}[2]{\V{#1}_{#2}}
\newcommand{\Mr}[2]{\V{#1}_{#2}} % matrix row
\newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\MhatC}[2]{\Vhat{#1}_{#2}} % matrix column
\newcommand{\Mbar}[1]{{\bm{\M{A}r \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\MbarC}[2]{\Vbar{#1}_{#2}} % matrix column
\newcommand{\Mn}[2]{\M{#1}^{(#2)}} % n-th matrix
\newcommand{\Mbarn}[2]{\Mbar{#1}^{(#2)}} % n-th matrix
\newcommand{\Mtilden}[2]{\Mtilde{#1}^{(#2)}} % n-th matrix
\newcommand{\MnTra}[2]{\M{#1}^{(#2){{\sf T}}}} % n-th matrix transpose
\newcommand{\MnE}[3]{\MakeLowercase{#1}^{(#2)}_{#3}} % n-th matrix element
\newcommand{\MnC}[3]{\V{#1}^{(#2)}_{#3}} % n-th matrix column
\newcommand{\MbarnC}[3]{\Vbar{#1}^{(#2)}_{#3}} % n-th matrix column
\newcommand{\MnCTra}[3]{\V{#1}^{(#2){{\sf T}}}_{#3}} % n-th matrix column transpose
\newcommand{\MCTra}[2]{\V{#1}^{{\sf T}}_{#2}} % n-th matrix column transpose

%% Tensors
\newcommand{\T}[1]{\boldsymbol{\mathscr{\MakeUppercase{#1}}}} %tensor
\newcommand{\Tbar}[1]{\boldsymbol{\M{A}r \mathscr{\MakeUppercase{#1}}}} %tensor
\newcommand{\That}[1]{\boldsymbol{\hat \mathscr{\MakeUppercase{#1}}}} %tensor
\newcommand{\Ttilde}[1]{\boldsymbol{\tilde \mathscr{\MakeUppercase{#1}}}} %tensor
\newcommand{\TE}[2]{\MakeLowercase{#1}_{\MI{#2}}} % tensor element with multi-index

%% Matrix-matrix operations
\newcommand{\Kron}{\otimes} %Kronecker
\newcommand{\Khat}{\odot} %Khatri-Rao
\newcommand{\Hada}{\ast} %Hadamard
\newcommand{\Divide}{\varoslash}

% Matricize
\newcommand{\Mz}[2]{\M{#1}_{(#2)}} % n-th mode matricize
\newcommand{\Mzsets}[3]{\TM{#1}{#2\times#3}} % matricze with just the sets specified
\newcommand{\Mzall}[4]{\TM{#1}{#2\times#3\,:\,#4}} % full matricize specification

%% K-Tensor
\newcommand{\KT}[1]{\left\llbracket #1 \right\rrbracket} % kruskal operator
\newcommand{\KTsmall}[1]{\llbracket #1 \rrbracket} % small kruskal operator
\newcommand{\KG}[1]{\langle #1 \rangle} % kruskal "group"

%% text with quads around it
\newcommand{\qtext}[1]{\quad\text{#1}\quad}

%% Derivative
\newcommand{\FD}[2]{\frac{\partial #1}{\partial #2}}

%% Prox-operator
\newcommand{\prox}[2]{\operatorname{prox}_{#1}({#2})}

%% Sign-operator
\newcommand{\sign}[1]{\operatorname{sign}({#1})}

%% convex hull
\newcommand{\conv}[1]{\operatorname{conv}({#1})}

%% For comments to each other
\newcommand{\Remark}[2]{{\color{red}Remark from #1: #2}\xspace}
\newcommand{\Note}[1]{{{\color{red}{#1}}}\xspace}

%% ----------------------------------------------------------------------
%% Environments
%% ----------------------------------------------------------------------

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
%% Assumptions environment
\newtheorem{assumption}[theorem]{Assumption}

\begin{document}
\title{Quantile Trend Filtering via Spingarn's Method}
\noindent
\textbf{Quantile Trend Filtering via Spingarn's Method}\\
\textbf{Date: \today}\\

%% ----------------------------------------------------------------------
%% Introduction
%% ----------------------------------------------------------------------
\section{Introduction}

We introduce a linear time algorithm for removing trends in time series.

Examples:
\begin{itemize}
\item Air quality 
\item ECG \cite{Sanyal2012}
\item Chromatogram baseline estimation  \cite{Ning2014}
\end{itemize}

Things to do:
\begin{itemize}
	\item Convergence of the algorithm
	\item Convergence rate?
	\item Timing experiments of LP versus Spingarn
	\item Make an R package - detrendr
	\item Compare on synthetic data quality of solution with existing methods
	\item Do comparisons on real data examples
\end{itemize}

%% ----------------------------------------------------------------------
%% Quantile Regression
%% ----------------------------------------------------------------------
\section{Quantile Regression}

The classic least squares regression is notoriously sensitive to outliers. One remedy to blunt the influence of outliers is to compute the least absolute deviations (LAD) solution in place of the least squares one. Given a design matrix $\M{X} \in \Real^{n \times p}$ and continuous responses $\V{y} \in \Real^n$, we estimate a regression vector $\V{\theta} \in \Real^p$ so that $\M{X}\V{\theta}$ is a good approximation of $\V{y}$. The LAD estimator is a solution to the problem
\begin{eqnarray*}
\underset{\V{\theta}}{\min}\; \frac{1}{n}\lVert \V{y} - \M{X}\V{\theta} \rVert_1.
\end{eqnarray*}
The above optimization problem generalizes the notion of the median of a collection of numbers. A median $\mu$ of $n$ reals $\VE{y}{1}, \ldots, \VE{y}{n}$ is the minimizer of the function
\begin{eqnarray*}
f(u) & = & \frac{1}{n}\sum_{i=1}^n \lvert \VE{y}{i} - \theta \rvert.
\end{eqnarray*}
Recall that the median is the 50th percentile or 0.5-quantile, namely half of the $\VE{y}{i}$ are less than or equal to $\mu$ and the other half is greater than or equal to $\mu$. The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ as follows. First define the so-called ``check function"
\begin{eqnarray*}
\rho_\tau(\Delta) & = & \begin{cases}
\tau \Delta & \text{$\Delta \geq 0$} \\
-(1-\tau)\Delta & \text{$\Delta < 0$}. \\
\end{cases}
\end{eqnarray*}
Then the $\tau$th quantile of the $\VE{y}{i}$ is a minimizer of the function
\begin{eqnarray*}
f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \theta).
\end{eqnarray*}
Returning to the regression context, we can generalize LAD regression to quantile regression, namely computing the minimizer of the function
\begin{eqnarray*}
f_\tau(\V{\theta}) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \langle \V{x}_i \mid \V{\theta} \rangle),
\end{eqnarray*}
where $\V{x}_i \in \Real^p$ denotes the $i$th row of $\M{X}$.

%% ----------------------------------------------------------------------
%% Trend Filtering
%% ----------------------------------------------------------------------
\section{Trend Filtering}

In the trend filtering problem \cite{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $\V{y} \in \Real^n$ by solving the following convex problem.
\begin{eqnarray*}
\underset{\V{\theta}}{\arg\min}\; \frac{1}{2n} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
\end{eqnarray*}
where $\lambda \geq 0$ is a regularization parameter that trades off the emphasis on the data fidelity term and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$. To understand the purpose of penalizing $\Mn{D}{k+1}$ consider the difference operator when $k = 0$.
\begin{eqnarray*}
\Mn{D}{1} = \begin{pmatrix}
-1 & 1 & 0 & \cdots & 0 & 0 \\
0 & -1 & 1 & \cdots & 0 & 0 \\
\vdots & & & & & \\
0 & 0 & 0 & \cdots & -1 & 1 \\
\end{pmatrix}
\end{eqnarray*}
Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \VE{\theta}{i} - \VE{\theta}{i+1} \rvert$ which is just total variation denoising in one dimension. The penalty incentivizes solutions which are piece-wise constant. For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
\begin{eqnarray*}
\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
\end{eqnarray*}
By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.

%% ----------------------------------------------------------------------
%% Quantile Trend Filtering
%% ----------------------------------------------------------------------
\section{Quantile Trend Filtering}

We combine the ideas of quantile regression and trend filtering, namely consider the signal approximation problem, where the design $\M{X}$ is the identity matrix.

The estimation of the quantile trend filtering model can be posed as the following optimization problem.
\begin{eqnarray}
\label{eq:quantile_trend}
	\underset{\V{\theta}}{\min}\; \frac{1}{n} \sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) + \lambda \lVert \Mn{D}{k} \V{\theta} \rVert_1,
\end{eqnarray}
where $\lambda$ is a nonnegative tuning parameter. As with the classic quantile regression, the quantile trend filtering problem can be solved by a linear program. We argue that it is better solved by Spingarn's method of partial inverses.

%% ----------------------------------------------------------------------
%% Spingarn's method of partial inverses
%% ----------------------------------------------------------------------
\section{Spingarn's method of partial inverses}

We first review Spingarn's method \cite{Spingarn1985}, which solves the following equality constrained convex problem:
\begin{equation}
\label{eq:spingarn}
\begin{split}
\text{minimize}\; & \psi(\V{x}) \\
\text{subject to}\; & \V{x} \in V,
\end{split}
\end{equation}
where $V$ is a subspace. The problem \Eqn{spingarn} can be expressed as the unconstrained optimization problem
\begin{equation}
\label{eq:spingarn2}
\text{minimize}\; \psi(\V{x}) + \iota_V(\V{x}),
\end{equation}
where $\iota_V$ is the indicator function of the set $V$. Spingarn's method applies Douglas-Rachford splitting to the problem \Eqn{spingarn2} to give the following updates.

\begin{eqnarray*}
\V{x}^+ & = & \prox{t\psi}{\V{z}} \\
\V{y}^+ & = & P_V(2\V{x}^+ - \V{z}) \\
\V{z}^+ & = & \V{z} + \V{y}^+ - \V{x}^+.
\end{eqnarray*}
The parameter $t$ is a step-size which can be fixed at 1 to guarantee convergence. The mapping $P_V$ is the orthogonal projection onto the set $V$.

%% ----------------------------------------------------------------------
%% Applying Spingarn's Method to Quantile Trend Filtering
%% ----------------------------------------------------------------------
\section{Applying Spingarn's Method to Quantile Trend Filtering}

To simplify the notation we suppress the order $k$ and write $\Mn{D}{k}$ as $\M{D}$.
We can reformulate our optimization problem \Eqn{quantile_trend} as the following equality constrained convex optimization problem.

\begin{eqnarray*}
\text{minimize}\; && f_1(\V{\theta}) + f_2(\V{\eta}) \\
\text{subject to}\; && \V{\eta} = \M{D}\V{\theta}
\end{eqnarray*}
where
\begin{eqnarray*}
f_1(\V{\theta}) = \sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) \quad\quad \text{and} \quad\quad
f_2(\V{\eta}) = \lVert \V{\eta} \rVert_1.
\end{eqnarray*}

If we set $\psi(\V{\theta}, \V{\eta}) = f_1(\V{\theta}) + f_2(\V{\eta})$ and $V = \{ \V{z}\Tra = (\V{\theta}\Tra, \V{\eta}\Tra) : \V{\eta} = \M{D}\V{\theta} \}$, then we can apply Spingarn's method. Note that
\begin{eqnarray*}
\prox{th}{\V{\theta}, \V{\eta}} & = & (\prox{tf_1}{\V{\theta}}, \prox{tf_2}{\V{\eta}}) \\
P_V(\V{\theta}, \V{\eta}) & = & \begin{pmatrix}
\M{I} \\ \M{D}
\end{pmatrix}
\begin{pmatrix}
\M{I} + \M{D}\Tra\M{D}
\end{pmatrix}\Inv
\begin{pmatrix}
\V{\theta} + \M{D}\Tra\V{\eta}
\end{pmatrix},
\end{eqnarray*}
where the projection $P_V$ requires a banded linear system solve, with bandwidth $k + 1$.
This linear solve can be accomplished in $\mathcal{O}(n(k + 1)^2)$. The first solve using a banded Cholesky decomposition requires $\mathcal{O}(n(k + 1)^2)$. Subsequent solves require
$\mathcal{O}(n(k+1))$.

We should also consider trying the reparameterization for ADMM trend filtering proposed in \cite{Ramdas2016}.

%% ----------------------------------------------------------------------
%% Proximal mappings
%% ----------------------------------------------------------------------
\subsection*{Proximal mappings}

We need the proximal mappings for $tf_1$ and $tf_2$.

\begin{eqnarray*}
\left [\prox{tf_1}{\V{\theta}} \right ]_i & = & \prox{(t/n)\rho_\tau}{\VE{y}{i} - \VE{\theta}{i}}, \\
\left [\prox{tf_2}{\V{\eta}} \right ]_j & = & S(\VE{\eta}{j}, t\lambda).
\end{eqnarray*}

The proximal mapping for $tf_2$ is the element-wise softhresholding operator. We now derive the  proximal mapping of $\rho_\tau(\Delta)$, which can be evaluated in closed form. We need to find the minimizer of the following univariate function
\begin{eqnarray*}
g_\tau(\Delta) & = & \Delta[\tau - I(\Delta < 0)] +  \frac{n}{2t} (\Delta - w)^2, \\
\end{eqnarray*}
where $w \in \Real$ is given and $I(\Delta < 0)$ is 0 when $\Delta < 0$ and 1 otherwise.

The subgradient of $\rho_\tau(\Delta) = \Delta[\tau - I(\Delta < 0)]$ is given by
\begin{eqnarray*}
\partial \rho_\tau(\Delta) & = & \begin{cases}
\tau & \text{if $\Delta > 0$} \\
\tau - 1 & \text{if $\Delta < 0$} \\
[\tau - 1, \tau] & \text{if $\Delta = 0$}.
\end{cases}
\end{eqnarray*}
The stationary condition is
\begin{eqnarray*}
\frac{n}{t}[w - \Delta] & \in & \partial \rho_\tau(\Delta).
\end{eqnarray*}
Therefore, the proximal mapping is given by 
\begin{eqnarray*}
\prox{(t/n) \rho_\tau}{w} & = & \begin{cases}
w - \tau \frac{t}{n}& \text{if $w > \tau \frac{t}{n}$} \\
w + (1-\tau)\frac{t}{n} & \text{if $w < - (1-\tau)\frac{t}{n}$} \\
0 & \text{if $-(1- \tau)\frac{t}{n} \leq w \leq \tau \frac{t}{n}$.}
\end{cases}
\end{eqnarray*}

\subsection*{Computational Costs}

\subsubsection*{Precomputation}

The following calculations need only be done once.

\begin{itemize}
\item $\mathcal{O}(n(k + 1)^2)$ to compute the banded Cholesky factorization of $\M{I} + [\Mn{D}{k}]\Tra[\Mn{D}{k}]$
\end{itemize}

banded linear system solve, with bandwidth $k + 1$.
This linear solve can be accomplished in $\mathcal{O}(n(k + 1)^2)$. The first solve using a banded Cholesky decomposition requires $\mathcal{O}(n(k + 1)^2)$. Subsequent solves require
$\mathcal{O}(n(k+1))$.

\subsubsection*{Per-Iteration}

The following calculations will be done every iteration.

\begin{itemize}
\item $\mathcal{O}(n)$ to compute $\prox{t\psi}{\V{\theta}, \V{\eta}}$
\item $\mathcal{O}((k+1)(n-k+1))$ to compute $\V{\theta} + [\Mn{D}{k}]\Tra\V{\eta}$
\item $\mathcal{O}(n(k+1))$ to compute $\V{\phi} = (\M{I} +[\Mn{D}{k}]\Tra[\Mn{D}{k}] )\Inv(\V{\theta} + \M{D}\Tra\V{\eta})$
\item $\mathcal{O}((k+1)(n-k+1))$ to compute $\begin{pmatrix}
\M{I} \\ \Mn{D}{k} \end{pmatrix}\V{\phi}$
\end{itemize}

The total cost when $n \gg p$ is $\mathcal{O}(nk)$.

\subsection{Summary}

\begin{itemize}
\item The overall computational complexity is essentially linear $\mathcal{O}(nk^2)$ for the initial banded Cholesky decomposition and the per-iteration complexity is $\mathcal{O}(nk)$.
\item One could also apply Anderson acceleration to reduce the number of Spingarin updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
\end{itemize}

%\section{Problem}
%
%\begin{eqnarray*}
%\text{minimize}\; \left [\sum_{i=1}^n w_i h\left (\langle \V{x}_i, \V{\beta} \rangle \right) \right ] + \lambda \lVert \V{\beta} \rVert_1,
%\end{eqnarray*}
%where
%\begin{eqnarray*}
%h(r) & = & \left (1 - r \right )_+ \\
%\partial h(r) & = & \begin{cases}
%-1 & r < 1 \\
%[-1,0] & r = 1 \\
%0 & r > 1.
%\end{cases}
%\end{eqnarray*}

%\section*{ADMM formulation}
%
%\begin{eqnarray*}
%\underset{\V{\beta},\V{z}}{\text{minimize}}\;  & & \left [\sum_{i=1}^n w_i h\left (\VE{z}{i} \right) \right ] + \lambda \lVert \V{\beta} \rVert_1 \\
%\text{subject to}\; & & \V{z} - \M{X}\V{\beta} = \V{0}.
%\end{eqnarray*}
%
%ADMM Updates
%
%\begin{eqnarray*}
%\Vn{z}{m+1} & = & \underset{\V{z}}{\arg\min}\;  \left [\sum_{i=1}^n w_i h\left (\VE{z}{i} \right) \right ] + \frac{\rho}{2} \lVert \V{z} - \M{X}\Vn{\beta}{m} + \Vn{u}{m} \rVert_2^2 \\
%\VnE{z}{m+1}{i} & = & \prox{(w_i/\rho) h}{\langle \V{x}_i, \Vn{\beta}{m} \rangle - \VnE{u}{m}{i}} \\
%\Vn{\beta}{m+1} & = & \underset{\V{\beta}}{\arg\min}\;  \frac{\lambda}{\rho} \lVert \V{\beta} \rVert_1 + \frac{1}{2} \lVert \Vn{z}{m+1} - \M{X}\V{\beta} + \Vn{u}{m} \rVert_2^2 \\
%\Vn{u}{m+1} & = & \Vn{u}{m} + \Vn{z}{m+1} - \M{X}\Vn{\beta}{m+1}.
%\end{eqnarray*}
%
%
%\begin{eqnarray*}
%\text{minimize}\; && f_1(\V{\beta}) + f_2(\M{X}\V{\beta}) \\
%\end{eqnarray*}

\bibliographystyle{plain}
\bibliography{detrendify}

\end{document}

### 2017-05-15 Notes/detrendify_spingarn.tex
Small edits to clean up.
typos
Added some references; fixed some typos.
Added some notes on detrending versus denoising and detrending.

@@
 \item ECG \cite{Sanyal2012}
-\item Chromatogram baseline estimation  \cite{Ning2014}
+\item Chromatogram baseline estimation  \cite{Ning2014, Ilewicz2015}
+\item Galaxy spectrum baseline estimation \cite{Ilewicz2015,Bacher2016}
 \end{itemize}
+
+Need to decide between detrending versus detrending + denoising.
+The BEADS \cite{Ning2014} method does both.
+We may wish to focus on detrending and then use wavelet SURE denoising as a postprocessing step, i.e.
+do a two-stage procedure, both of which can be done in linear time.
+On paper this should be faster than the BEADS procedure.
+Or we may just want to stick with detrending.
+There's Matlab code for BEADS, and the BEADS paper also points to two other popular methods in chromotography.
 
@@
 	\item Convergence of the algorithm
-	\item Convergence rate?
+	\item Convergence rate \cite{He2012, He2015, Davis2017}
 	\item Timing experiments of LP versus Spingarn
@@
 \begin{eqnarray*}
-f_1(\V{\theta}) = \sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) \quad\quad \text{and} \quad\quad
-f_2(\V{\eta}) = \lVert \V{\eta} \rVert_1.
+f_1(\V{\theta}) = \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) \quad\quad \text{and} \quad\quad
+f_2(\V{\eta}) = \lambda\lVert \V{\eta} \rVert_1.
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-\left [\prox{tf_1}{\V{\theta}} \right ]_i & = & \prox{(t/n)\rho_\tau}{\VE{y}{i} - \VE{\theta}{i}}, \\
+\left [\prox{tf_1}{\V{\theta}} \right ]_i & = & \VE{y}{i} - \prox{(t/n)\rho_\tau}{\VE{y}{i} - \VE{\theta}{i}}, \\
 \left [\prox{tf_2}{\V{\eta}} \right ]_j & = & S(\VE{\eta}{j}, t\lambda).
@@
 
-banded linear system solve, with bandwidth $k + 1$.
-This linear solve can be accomplished in $\mathcal{O}(n(k + 1)^2)$. The first solve using a banded Cholesky decomposition requires $\mathcal{O}(n(k + 1)^2)$. Subsequent solves require
-$\mathcal{O}(n(k+1))$.
-
 \subsubsection*{Per-Iteration}
@@
 
-The total cost when $n \gg p$ is $\mathcal{O}(nk)$.
+The total cost is $\mathcal{O}(nk)$.
 
@@
 
-%\section{Problem}
-%
-%\begin{eqnarray*}
-%\text{minimize}\; \left [\sum_{i=1}^n w_i h\left (\langle \V{x}_i, \V{\beta} \rangle \right) \right ] + \lambda \lVert \V{\beta} \rVert_1,
-%\end{eqnarray*}
-%where
-%\begin{eqnarray*}
-%h(r) & = & \left (1 - r \right )_+ \\
-%\partial h(r) & = & \begin{cases}
-%-1 & r < 1 \\
-%[-1,0] & r = 1 \\
-%0 & r > 1.
-%\end{cases}
-%\end{eqnarray*}
-
-%\section*{ADMM formulation}
-%
-%\begin{eqnarray*}
-%\underset{\V{\beta},\V{z}}{\text{minimize}}\;  & & \left [\sum_{i=1}^n w_i h\left (\VE{z}{i} \right) \right ] + \lambda \lVert \V{\beta} \rVert_1 \\
-%\text{subject to}\; & & \V{z} - \M{X}\V{\beta} = \V{0}.
-%\end{eqnarray*}
-%
-%ADMM Updates
-%
-%\begin{eqnarray*}
-%\Vn{z}{m+1} & = & \underset{\V{z}}{\arg\min}\;  \left [\sum_{i=1}^n w_i h\left (\VE{z}{i} \right) \right ] + \frac{\rho}{2} \lVert \V{z} - \M{X}\Vn{\beta}{m} + \Vn{u}{m} \rVert_2^2 \\
-%\VnE{z}{m+1}{i} & = & \prox{(w_i/\rho) h}{\langle \V{x}_i, \Vn{\beta}{m} \rangle - \VnE{u}{m}{i}} \\
-%\Vn{\beta}{m+1} & = & \underset{\V{\beta}}{\arg\min}\;  \frac{\lambda}{\rho} \lVert \V{\beta} \rVert_1 + \frac{1}{2} \lVert \Vn{z}{m+1} - \M{X}\V{\beta} + \Vn{u}{m} \rVert_2^2 \\
-%\Vn{u}{m+1} & = & \Vn{u}{m} + \Vn{z}{m+1} - \M{X}\Vn{\beta}{m+1}.
-%\end{eqnarray*}
-%
-%
-%\begin{eqnarray*}
-%\text{minimize}\; && f_1(\V{\beta}) + f_2(\M{X}\V{\beta}) \\
-%\end{eqnarray*}
-
 \bibliographystyle{plain}

### 2017-05-30 Notes/detrendify_spingarn.tex
Some small edits to the document.

@@
 \V{y}^+ & = & P_V(2\V{x}^+ - \V{z}) \\
-\V{z}^+ & = & \V{z} + \V{y}^+ - \V{x}^+.
+\V{z}^+ & = & \V{z} + \lambda(\V{y}^+ - \V{x}^+).
 \end{eqnarray*}
-The parameter $t$ is a step-size which can be fixed at 1 to guarantee convergence. The mapping $P_V$ is the orthogonal projection onto the set $V$.
+The parameter $t$ is a step-size and $\lambda_n$ is Karolinski-Mann.
+We require that $\lambda_n \in ]0, 2[$ and that $\sum_n \lambda_n(1-\lambda_n) = \infty$.
+The mapping $P_V$ is the orthogonal projection onto the set $V$.
 
@@
 \item The overall computational complexity is essentially linear $\mathcal{O}(nk^2)$ for the initial banded Cholesky decomposition and the per-iteration complexity is $\mathcal{O}(nk)$.
-\item One could also apply Anderson acceleration to reduce the number of Spingarin updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
+\item One could also apply Anderson acceleration to reduce the number of Spingarn updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
 \end{itemize}

### 2017-06-01 Notes/detrendify_spingarn.tex
Added some more references and version 2 of the algorithm.

--- 2017-05-15 Notes/detrendify_spingarn.tex
@@
 \item Galaxy spectrum baseline estimation \cite{Ilewicz2015,Bacher2016}
+\item Identifying absorption dips from black body radiation
 \end{itemize}
@@
 \begin{eqnarray*}
-\V{x}^+ & = & \prox{t\psi}{\V{z}} \\
-\V{y}^+ & = & P_V(2\V{x}^+ - \V{z}) \\
-\V{z}^+ & = & \V{z} + \V{y}^+ - \V{x}^+.
+\Vn{x}{k+1} & = & \prox{t\psi}{\Vn{z}{k}} \\
+\Vn{y}{k+1} & = & P_V\left(2\Vn{x}{k+1} - \Vn{z}{k} \right) \\
+\Vn{z}{k+1} & = & \Vn{z}{k} + \lambda^{(k)}\left(\Vn{y}{k+1} - \Vn{x}{k+1} \right).
 \end{eqnarray*}
-The parameter $t$ is a step-size which can be fixed at 1 to guarantee convergence. The mapping $P_V$ is the orthogonal projection onto the set $V$.
+The parameter $t$ is a step-size and $\lambda^{(k)}$ is Krasnosel'ski{\u i}-Mann iteration (Need citation).
+We require that $\lambda^{(k)} \in ]0, 2[$ and that $\sum_n \lambda^{(k)} (2-\lambda^{(k)}) = \infty$.
+The mapping $P_V$ is the orthogonal projection onto the set $V$.
+Note that the algorithm iterates $\Vn{x}{k}$ will converge to a solution to problem $\Eqn{spingarn}$ \cite{Combettes2005}.
+We need to experiment with different over / under-relaxation parameters $\lambda^{(k)}$ and step sizes $t$.
+It will converge if we take $\lambda^{(k)} = 1$ and $t = 1$, but we may be able to converge faster in practice by taking non-trivial values.
 
@@
 This linear solve can be accomplished in $\mathcal{O}(n(k + 1)^2)$. The first solve using a banded Cholesky decomposition requires $\mathcal{O}(n(k + 1)^2)$. Subsequent solves require
-$\mathcal{O}(n(k+1))$.
-
-We should also consider trying the reparameterization for ADMM trend filtering proposed in \cite{Ramdas2016}.
+$\mathcal{O}(n(k+1))$. We can use RccpArmadillo to do banded Cholesky \cite{Eddelbuettel2014}.
 
@@
 \item The overall computational complexity is essentially linear $\mathcal{O}(nk^2)$ for the initial banded Cholesky decomposition and the per-iteration complexity is $\mathcal{O}(nk)$.
-\item One could also apply Anderson acceleration to reduce the number of Spingarin updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
+\item One could also apply Anderson acceleration \cite{Walker2011} to reduce the number of Spingarn updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
+\end{itemize}
+
+%% ----------------------------------------------------------------------
+%% Applying Spingarn's Method to Quantile Trend Filtering version 2
+%% ----------------------------------------------------------------------
+\section{Applying Spingarn's Method to Quantile Trend Filtering version 2}
+
+Following the specialized ADMM algorithm for trend filtering \cite{Ramdas2016}, we can also take advantage of fast exact solvers of the one-dimensional fussed lasso problem \cite{davies2001, Johnson2013}.
+The first method is based on taut strings, and the second is based on dynamic programming.
+Both of these methods run in linear time.
+A third exact method with worst case quadratic penalty but linear time in typical cases was proposed in \cite{Condat2013}.
+C code is available for all three methods.
+We should compare them.
+
+What is the reparameterization?
+
+\begin{eqnarray*}
+\text{minimize}\; && f_1(\V{\theta}) + f_2(\V{\eta}) \\
+\text{subject to}\; && \V{\eta} = \M{D}\V{\theta}
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+f_1(\V{\theta}) = \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) \quad\quad \text{and} \quad\quad
+f_2(\V{\eta}) = \lambda\lVert \Mn{D}{1}\V{\eta} \rVert_1.
+\end{eqnarray*}
+
+Difference with version 1: If we are interested in a 3rd order penalization, then $\M{D} = \Mn{D}{2}$ in version 2 as opposed to $\M{D} = \Mn{D}{3}$ in version 1.
+The proximal mapping for $f_2$ is solved using one of the exact solvers \cite{davies2001, Johnson2013, Condat2013}.
+Show some timing results between the different versions.
+Point readers to \cite{Ramdas2016} for a discussion of why this minor reparameterization may lead to speed up.
+
+\section{Other Practical Issues}
+
+\begin{itemize}
+	\item Read compressed data
+	\item May want to read data directly in C and not pull into R; make R just an interface
+	\item Use historical data to choose $\tau$ and $\lambda$
+	\item 
 \end{itemize}

### 2017-06-02 Notes/detrendify_spingarn.tex
built pdf

@@
 Recall that the median is the 50th percentile or 0.5-quantile, namely half of the $\VE{y}{i}$ are less than or equal to $\mu$ and the other half is greater than or equal to $\mu$.
-The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ as follows.
+The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ to give us quantile regression \cite{Koenker1978}.
+
 First define the so-called ``check function"
@@
 	\item Use historical data to choose $\tau$ and $\lambda$
-	\item 
+\end{itemize}
+
+\section{To Do}
+
+\begin{itemize}
+	\item Find out if we can use the EPA data. Add co-authors?
+	\item Make version 2 with the three fast fused lasso solvers
+	\item Add homotopy / warm start
+	\item Add model selection
+	\item Do experiments to evaluate different choices of $t$ and $\lambda^{(k)}$
+	\item See if Anderson acceleration helps. Talk to Tim Kelley?
+	\item Write vignette
+	\item Do comparisons with BEAD
+	\item Do wind polar plots with and without removing trends
+	\item Compare with Splines
 \end{itemize}

### 2017-06-07 Notes/timing_and_convergence_examples.Rmd
Added timing and convergence examples to notes

---
title: "detrendr examples"
output: pdf_document
---

## Some examples of baseline fitting with detrendr

```{r load, include=FALSE, echo=FALSE, cache=FALSE}
library(Matrix)
library(devtools)
library(roxygen2)
library(Rcpp)
load_all("../Code/detrendr")

```

```{r simulate}

simulate <- function(n, k){
  x <- seq(1/n, 1, length.out=n)
  f <- 2*(x + 2)^2 + 3*cos(3*pi*x)
  tau <- 5e4
  g1 <- 40*exp(-tau*(x-0.3)^2)
  g2 <- 30*exp(-tau/2*(x-0.58)^2)
  g4 <- 45*exp(-tau/7*(x-0.6)^2)
  g3 <- 37*exp(-tau*(x-0.63)^2)
  y <- f + g1 + g2 + g3 + g4 + rnorm(n)
  D <- get_Dk(n, k)
  M <- diag(n) + crossprod(D)
  cholM <- as.matrix(chol(M))
  return(list(y=y, f=f, x=x, D=D, M=M, cholM=cholM))
}
```

## Current timing experiments
```{r timing, include=TRUE, echo=TRUE, cache=TRUE}
nSeq <- seq(100, 1500, 200)
timing <- data.frame(n=nSeq, R = NA, C=NA)
tau <- 0.01
lambda <- 10
step <- 1

for (i in 1:length(nSeq)){
  simData <- simulate(nSeq[i], 3)
  y <- simData$y
  theta <- simData$y
  D <- simData$D
  eta <- matrix(D%*%theta)
  M <- simData$M
  cholM <- simData$cholM
  timing$R[i] <- system.time( 
      multi_step_R <- spingarn_multi_step_R(theta, eta, y, D, M, lambda, 
                                            tau, step, 100))[3]
  timing$C[i] <- system.time(
    multi_step <- spingarn_multi_step(theta, eta, y, D, cholM,
                                     lambda, tau, step, 100))[3]
}

plot(C~n, timing, type="l", col="blue", ylab="Elapsed Time (s) per 100 iterations", main = "R code is in black, C in blue")
lines(R~n, timing, type="l")
```

## Convergence when n=100, lambda = 10
```{r convergence, include=TRUE, echo=TRUE, cache=TRUE}

n <- 500
simData <- simulate(n, 3)

y <- simData$y
theta <- simData$y
D <- simData$D
eta <- matrix(D%*%theta)
M <- simData$M
cholM <- simData$cholM
tau <- 0.05
lambda <- 10
step <- 1

max_iter <- 70000
rerr <- double(max_iter-1)
THx <- matrix(NA, n, max_iter)
for (iter in 1:max_iter) {
  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
  theta <- one_step[[1]]
  eta <- one_step[[2]]
  THx[,iter] <- prox_f1(theta, y, tau)
  if (iter > 1){
    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
  }
}
theta_last <- prox_f1(theta, y, tau)

par(mfrow=c(2,1), mar=c(3,4,1,1))
plot(log(rerr), type="l")

plot(simData$x,simData$f,type='l',col='blue',
ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
points(simData$x,y, cex=.6)
lines(simData$x,theta_last,col='red', lwd=3)

```

## Convergence when n=1000, lambda = 100
```{r convergence2, include=TRUE, echo=FALSE, cache=TRUE}

n <- 1000
simData <- simulate(n, 3)

y <- simData$y
theta <- simData$y
D <- simData$D
eta <- matrix(D%*%theta)
M <- simData$M
cholM <- simData$cholM
tau <- 0.05
lambda <- 100
step <- 1

max_iter <- 70000
rerr <- double(max_iter-1)
THx <- matrix(NA, n, max_iter)
for (iter in 1:max_iter) {
  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
  theta <- one_step[[1]]
  eta <- one_step[[2]]
  THx[,iter] <- prox_f1(theta, y, tau)
  if (iter > 1){
    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
  }
}
theta_last <- prox_f1(theta, y, tau)

par(mfrow=c(2,1), mar=c(3,4,1,1))
plot(log(rerr), type="l")

plot(simData$x,simData$f,type='l',col='blue',
ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
points(simData$x,y, cex=.6)
lines(simData$x,theta_last,col='red', lwd=3)

```

## Convergence when n=2000, lambda = 100
```{r convergence3, include=TRUE, echo=FALSE, cache=TRUE}

n <- 2000
simData <- simulate(n, 3)

y <- simData$y
theta <- simData$y
D <- simData$D
eta <- matrix(D%*%theta)
M <- simData$M
cholM <- simData$cholM
tau <- 0.05
lambda <- 100
step <- 1

max_iter <- 70000
rerr <- double(max_iter-1)
THx <- matrix(NA, n, max_iter)
for (iter in 1:max_iter) {
  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
  theta <- one_step[[1]]
  eta <- one_step[[2]]
  THx[,iter] <- prox_f1(theta, y, tau)
  if (iter > 1){
    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
  }
}
theta_last <- prox_f1(theta, y, tau)

par(mfrow=c(2,1), mar=c(3,4,1,1))
plot(log(rerr), type="l")

plot(simData$x,simData$f,type='l',col='blue',
ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
points(simData$x,y, cex=.6)
lines(simData$x,theta_last,col='red', lwd=3)

```

### 2017-06-07 Code/detrendr/R/RcppExports.R

@@ -1,32 +1,6 @@
 # Generated by using Rcpp::compileAttributes() -> do not edit by hand
 # Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
-#' @title 
-#' 
-#' \code{get_D1} computes in the discrete derivative matrix.
-#' 
-#' @param n length of input
-NULL
-
-#' @title
-#' kth order difference matrix
-#' 
-#' \code{get_Dkn} computes the discrete kth derivative matrix
-#' 
-#' @param n length of input
-#' @param k order of the derivative
-NULL
-
-#' @title 
-#' Project onto subspace
-#' 
-#' \code{project_V} projects (theta, eta) onto the subspace eta = D%*%theta
-#' 
-#' @param theta first input
-#' @param eta second input
-#' @param D differencing matrix
-NULL
-
 #' @useDynLib detrendr
 #' @importFrom Rcpp evalCpp
 #' 
@@ -93,7 +67,8 @@ prox_f2 <- function(eta, lambda, step = 1) {
 #' @title
 #' Proximal mapping
 #' 
-#' \code{prox} computes the block separable proximal mapping. 
+#' \code{prox} computes the block separable proximal mapping, changes theta
+#' and eta in place
 #' @param theta input
 #' @param eta input
 #' @param y response
@@ -102,22 +77,122 @@ prox_f2 <- function(eta, lambda, step = 1) {
 #' @param step step-size
 #' @export 
 prox <- function(theta, eta, y, lambda, tau = 0.05, step = 1.0) {
-    .Call('detrendr_prox', PACKAGE = 'detrendr', theta, eta, y, lambda, tau, step)
+    invisible(.Call('detrendr_prox', PACKAGE = 'detrendr', theta, eta, y, lambda, tau, step))
 }
 
 #' @title
-#' Function to test differencing matrix, returns single element of Dk(n)
+#' Proximal Mapping Test
+#' 
+#' \code{prox_test} computes the block separable proximal mapping.
+#' Returns values of theta and eta. 
+#' @param theta input
+#' @param eta input
+#' @param y response
+#' @param lambda regularization parameter
+#' @param tau quantile parameter
+#' @param step step-size
+#' @export 
[…]

### 2017-06-07 Code/detrendr/src/RcppExports.cpp

@@ -47,8 +47,23 @@ BEGIN_RCPP
 END_RCPP
 }
 // prox
-Rcpp::List prox(arma::vec theta, arma::vec eta, arma::vec y, double lambda, double tau, double step);
+void prox(arma::vec& theta, arma::vec& eta, arma::vec y, double lambda, double tau, double step);
 RcppExport SEXP detrendr_prox(SEXP thetaSEXP, SEXP etaSEXP, SEXP ySEXP, SEXP lambdaSEXP, SEXP tauSEXP, SEXP stepSEXP) {
+BEGIN_RCPP
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< arma::vec& >::type theta(thetaSEXP);
+    Rcpp::traits::input_parameter< arma::vec& >::type eta(etaSEXP);
+    Rcpp::traits::input_parameter< arma::vec >::type y(ySEXP);
+    Rcpp::traits::input_parameter< double >::type lambda(lambdaSEXP);
+    Rcpp::traits::input_parameter< double >::type tau(tauSEXP);
+    Rcpp::traits::input_parameter< double >::type step(stepSEXP);
+    prox(theta, eta, y, lambda, tau, step);
+    return R_NilValue;
+END_RCPP
+}
+// prox_test
+Rcpp::List prox_test(arma::vec theta, arma::vec eta, arma::vec y, double lambda, double tau, double step);
+RcppExport SEXP detrendr_prox_test(SEXP thetaSEXP, SEXP etaSEXP, SEXP ySEXP, SEXP lambdaSEXP, SEXP tauSEXP, SEXP stepSEXP) {
 BEGIN_RCPP
     Rcpp::RObject rcpp_result_gen;
     Rcpp::RNGScope rcpp_rngScope_gen;
@@ -58,35 +73,79 @@ BEGIN_RCPP
     Rcpp::traits::input_parameter< double >::type lambda(lambdaSEXP);
     Rcpp::traits::input_parameter< double >::type tau(tauSEXP);
     Rcpp::traits::input_parameter< double >::type step(stepSEXP);
-    rcpp_result_gen = Rcpp::wrap(prox(theta, eta, y, lambda, tau, step));
+    rcpp_result_gen = Rcpp::wrap(prox_test(theta, eta, y, lambda, tau, step));
     return rcpp_result_gen;
 END_RCPP
 }
-// test_Dk
[…]

### 2017-06-07 Code/detrendr/src/prox.cpp

@@ -119,7 +119,8 @@ void prox(arma::vec& theta,
 //' @title
 //' Proximal Mapping Test
 //' 
-//' \code{prox_test} computes the block separable proximal mapping. 
+//' \code{prox_test} computes the block separable proximal mapping.
+//' Returns values of theta and eta. 
 //' @param theta input
 //' @param eta input
 //' @param y response
@@ -138,35 +139,35 @@ Rcpp::List prox_test(arma::vec theta,
   return Rcpp::List::create(theta = theta, eta = eta);
 }
 
-//' @title 
+//' @title Discrete derivative matrix
 //' 
-//' \code{get_D1} computes in the discrete derivative matrix.
+//' \code{get_D1} computes the sparse discrete derivative matrix.
 //' 
 //' @param n length of input
+//' @examples
+//' n <- 5
+//' D1 <- get_D1(n)
+//' @export
+//[[Rcpp::export]]
 arma::sp_mat get_D1(int n){
   int numberNonZero = 2*(n-1);
   arma::vec values = ones<vec>(numberNonZero);
   values.subvec(n-1, 2*(n-1)-1) = -1*values.subvec(n-1, 2*(n-1)-1);
-  
   arma::umat locs = repmat(linspace<urowvec>(0,n-2,n-1),2,2);
   locs.submat(1, n-1, 1, numberNonZero-1) = locs.submat(1, n-1, 1, 
               numberNonZero-1) + 1;
-  
   arma::sp_mat D1 = arma::sp_mat(locs, values);
-
-  //Rcout << "D1" << std::endl << D1 << std::endl;
-  
   return D1;
 }
 
 //' @title
-//' kth order difference matrix
+//' kth order sparse difference matrix
 //' 
-//' \code{get_Dkn} computes the discrete kth derivative matrix
+//' \code{get_Dkn} computes the sparse discrete kth derivative matrix
 //' 
 //' @param n length of input
 //' @param k order of the derivative
-//' 
+//' @export
 //[[Rcpp::export]]
 arma::sp_mat get_Dk(int n, 
                      int k){
@@ -187,7 +188,8 @@ arma::sp_mat get_Dk(int n,
 //' @param theta first input
 //' @param eta second input
 //' @param D differencing matrix
-//' @param cholM upper triangular cholesky of  I + DtD
+//' @param cholM upper triangular cholesky decomposition of  I + DtD
+//' @export
 //[[Rcpp::export]]
 void project_V(arma::vec& theta,
                      arma::vec& eta,
[…]

### 2017-06-20 Notes/timing_and_convergence_examples.Rmd
documentation updated, all tests currently pass

@@
   M <- diag(n) + crossprod(D)
-  cholM <- as.matrix(chol(M))
+  cholM <- chol(M)
   return(list(y=y, f=f, x=x, D=D, M=M, cholM=cholM))
@@
 ```{r timing, include=TRUE, echo=TRUE, cache=TRUE}
-nSeq <- seq(100, 1500, 200)
+nSeq <- seq(500, 4500, 500)
 timing <- data.frame(n=nSeq, R = NA, C=NA)
 tau <- 0.01
-lambda <- 10
+lambda0 <- .000001
 step <- 1
+k <- 4
 
 for (i in 1:length(nSeq)){
-  simData <- simulate(nSeq[i], 3)
+  simData <- simulate(nSeq[i], k)
+  lambda <- lambda0*nSeq[i]^k/factorial(k)
   y <- simData$y
@@
   timing$R[i] <- system.time( 
-      multi_step_R <- spingarn_multi_step_R(theta, eta, y, D, M, lambda, 
-                                            tau, step, 100))[3]
+      multi_step_R <- spingarn_multi_step_R(theta, eta, y, D, M, lambda, tau, step, 100))[3]
   timing$C[i] <- system.time(
     multi_step <- spingarn_multi_step(theta, eta, y, D, cholM,
-                                     lambda, tau, step, 100))[3]
+                                     lambda, tau, step, 100, k))[3]
 }
 
-plot(C~n, timing, type="l", col="blue", ylab="Elapsed Time (s) per 100 iterations", main = "R code is in black, C in blue")
+plot(C~n, timing, type="l", col="blue", ylab="Elapsed Time (s) per 100 iterations", main = "R code is in black, C in blue", ylim = c(0, 1))
 lines(R~n, timing, type="l")
@@
 ```{r convergence3, include=TRUE, echo=FALSE, cache=TRUE}
-
-n <- 2000
-simData <- simulate(n, 3)
+lambda0 <- 2*10^(-6)
+n <- 1000
+k <- 3
+simData <- simulate(n, k)
 
@@
 tau <- 0.05
-lambda <- 100
+lambda <- lambda0*n^k/factorial(k)
 step <- 1
 
-max_iter <- 70000
-rerr <- double(max_iter-1)
-THx <- matrix(NA, n, max_iter)
-for (iter in 1:max_iter) {
-  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
-  theta <- one_step[[1]]
-  eta <- one_step[[2]]
-  THx[,iter] <- prox_f1(theta, y, tau)
-  if (iter > 1){
-    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
-                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
-  }
-}
+max_iter <- 100000
+
+multi_step <- spingarn_multi_step(theta, eta, y, D, cholM,
+                                 lambda, tau, step, max_iter, k)
+
+theta <- multi_step[[1]]
+eta <- multi_step[[2]]
 theta_last <- prox_f1(theta, y, tau)
-
-par(mfrow=c(2,1), mar=c(3,4,1,1))
-plot(log(rerr), type="l")
 
@@
 points(simData$x,y, cex=.6)
-lines(simData$x,theta_last,col='red', lwd=3)
+lines(simData$x,theta_last,col='purple', lwd=3)
 

### 2017-06-20 Code/detrendr/R/RcppExports.R

@@ -3,22 +3,26 @@
 
 #' @useDynLib detrendr
 #' @importFrom Rcpp evalCpp
+NULL
+
+#' Banded Cholesky Solve
 #' 
-#' @title 
 #' \code{chol_solve} Solves a linear system cholM%*%x=b 
 #' when cholM is a sparse banded cholesky.
 #' 
-#' @param cholM sparse banded cholseky matrix
+#' @param cholM sparse banded cholseky decomposition of discrete derivative 
+#' matrix of order k
 #' @param b dense solution vector
-#' @param k order of trend filtering
-#' @param upper indicates wheter cholM is upper or lower triangular
+#' @param k order of discrete derivative matrix
+#' @param upper boolean indicator of whether cholM is upper or lower triangular
 #' @export
 #' 
 chol_solve <- function(cholM, b, k, upper = TRUE) {
     .Call('detrendr_chol_solve', PACKAGE = 'detrendr', cholM, b, k, upper)
 }
 
-#' @title
+#' Proximal Mapping
+#' 
 #' \code{prox_quantile} computes the proximal mapping of the check function.
 #'
 #' @param w input
@@ -42,7 +46,6 @@ prox_quantile <- function(w, tau, alpha) {
     .Call('detrendr_prox_quantile', PACKAGE = 'detrendr', w, tau, alpha)
 }
 
-#' @title
 #' Proximal mapping of f_1
 #' 
 #' \code{prox_f1} computes the proximal mapping of the average quantile loss
@@ -51,13 +54,11 @@ prox_quantile <- function(w, tau, alpha) {
 #' @param y response
 #' @param tau quantile parameter
 #' @param step step-size
-#' @examples
 #' @export
 prox_f1 <- function(theta, y, tau = 0.05, step = 1.0) {
     .Call('detrendr_prox_f1', PACKAGE = 'detrendr', theta, y, tau, step)
 }
 
-#' @title
 #' Proximal mapping of f_2
 #' 
 #' \code{prox_f2} computes the proximal mapping of the L1 penalty
@@ -78,11 +79,11 @@ prox_f2 <- function(eta, lambda, step = 1) {
     .Call('detrendr_prox_f2', PACKAGE = 'detrendr', eta, lambda, step)
 }
 
-#' @title
 #' Proximal mapping
 #' 
 #' \code{prox} computes the block separable proximal mapping, changes theta
 #' and eta in place
+#' 
[…]

### 2017-06-20 Code/detrendr/src/RcppExports.cpp

@@ -166,3 +166,23 @@ BEGIN_RCPP
     return rcpp_result_gen;
 END_RCPP
 }
+
+static const R_CallMethodDef CallEntries[] = {
+    {"detrendr_chol_solve", (DL_FUNC) &detrendr_chol_solve, 4},
+    {"detrendr_prox_quantile", (DL_FUNC) &detrendr_prox_quantile, 3},
+    {"detrendr_prox_f1", (DL_FUNC) &detrendr_prox_f1, 4},
+    {"detrendr_prox_f2", (DL_FUNC) &detrendr_prox_f2, 3},
+    {"detrendr_prox", (DL_FUNC) &detrendr_prox, 6},
+    {"detrendr_prox_test", (DL_FUNC) &detrendr_prox_test, 6},
+    {"detrendr_get_D1", (DL_FUNC) &detrendr_get_D1, 1},
+    {"detrendr_get_Dk", (DL_FUNC) &detrendr_get_Dk, 2},
+    {"detrendr_project_V", (DL_FUNC) &detrendr_project_V, 5},
+    {"detrendr_spingarn_one_step", (DL_FUNC) &detrendr_spingarn_one_step, 9},
+    {"detrendr_spingarn_multi_step", (DL_FUNC) &detrendr_spingarn_multi_step, 10},
+    {NULL, NULL, 0}
+};
+
+RcppExport void R_init_detrendr(DllInfo *dll) {
+    R_registerRoutines(dll, NULL, CallEntries, NULL, NULL);
+    R_useDynamicSymbols(dll, FALSE);
+}

### 2017-06-20 Notes/timing_and_convergence_examples.Rmd
updates to timing comparisons

@@
   M <- diag(n) + crossprod(D)
-  cholM <- chol(M)
+  cholM <- Matrix::chol(M)
   return(list(y=y, f=f, x=x, D=D, M=M, cholM=cholM))
 }
-```
 
-## Current timing experiments
-```{r timing, include=TRUE, echo=TRUE, cache=TRUE}
-nSeq <- seq(500, 4500, 500)
-timing <- data.frame(n=nSeq, R = NA, C=NA)
-tau <- 0.01
-lambda0 <- .000001
-step <- 1
-k <- 4
-
-for (i in 1:length(nSeq)){
-  simData <- simulate(nSeq[i], k)
-  lambda <- lambda0*nSeq[i]^k/factorial(k)
+compareTimeR <- function(n){
+  simData <- simulate(n, k)
+  lambda <- lambda0*n^k/factorial(k)
   y <- simData$y
@@
   cholM <- simData$cholM
-  timing$R[i] <- system.time( 
-      multi_step_R <- spingarn_multi_step_R(theta, eta, y, D, M, lambda, tau, step, 100))[3]
-  timing$C[i] <- system.time(
-    multi_step <- spingarn_multi_step(theta, eta, y, D, cholM,
-                                     lambda, tau, step, 100, k))[3]
+  spingarn_multi_step_R(theta, eta, y, D, M, lambda, tau, step, 500)
 }
 
-plot(C~n, timing, type="l", col="blue", ylab="Elapsed Time (s) per 100 iterations", main = "R code is in black, C in blue", ylim = c(0, 1))
-lines(R~n, timing, type="l")
+compareTimeC <- function(n){
+  simData <- simulate(n, k)
+  lambda <- lambda0*n^k/factorial(k)
+  y <- simData$y
+  theta <- simData$y
+  D <- simData$D
+  eta <- matrix(D%*%theta)
+  M <- simData$M
+  cholM <- simData$cholM
+  spingarn_multi_step(theta, eta, y, D, M, lambda, tau, step, 500)
+}
+```
+
+## Current timing experiments
+```{r timing, include=TRUE, echo=TRUE, cache=TRUE}
+nSeq <- seq(200, 3000, 200)
+timing <- data.frame(n=nSeq, R_mean = NA, R_lq = NA, R_uq = NA,
+                             C_mean=NA, C_lq=NA, C_uq = NA)
+tau <- 0.01
+lambda0 <- 1e-8
+step <- 1
+k <- 3
+
+for (i in 1:length(nSeq)){
+
+  set.seed(38912498)
+  Rbench <- microbenchmark(compareTimeR(nSeq[i]), times=100)
+  set.seed(38912498)
+  Cbench <- microbenchmark(compareTimeC(nSeq[i]), times=100)
+
+  timing$R_mean[i] <- median(Rbench$time*1e-9)
+  timing$R_lq[i] <- quantile(Rbench$time*1e-9, .25)
+  timing$R_uq[i] <- quantile(Rbench$time*1e-9, .75)
+  timing$C_mean[i] <- median(Cbench$time*1e-9)
+  timing$C_lq[i] <- quantile(Cbench$time*1e-9, .25)
+  timing$C_uq[i] <- quantile(Cbench$time*1e-9, .75)
+  print(i)
+}
+
+library(microbenchmark)
+
+plot(C_mean~n, timing, type="l", col="blue", ylab="Elapsed Time (s)' per 500 iterations", main = "R code is in black, C in blue")
+lines(C_lq~n, timing, col="blue")
+lines(C_uq~n, timing, col="blue")
+lines(R_mean~n, timing, type="l")
 ```

### 2017-06-20 Code/detrendr/src/prox.cpp

@@ -331,10 +331,11 @@ Rcpp::List spingarn_multi_step(arma::vec theta,
                              double step = 1,
                              double numberIter=1, 
                              int k=3){
-
+  arma::vec theta_cp = theta;
+  arma::vec eta_cp = eta;
   for (int i = 0; i < numberIter; i++){
-    spingarn_one_step(theta, eta, y, D, cholM, lambda, tau, step, k);
+    spingarn_one_step(theta_cp, eta_cp, y, D, cholM, lambda, tau, step, k);
   }
 
-  return Rcpp::List::create(theta=theta, eta=eta);
+  return Rcpp::List::create(theta=theta_cp, eta=eta_cp);
 }

### 2017-06-23 Notes/timing_and_convergence_examples.Rmd
added RcppEigen functions and some timing comparisons

--- 2017-06-07 Notes/timing_and_convergence_examples.Rmd
@@
 ---
-
-## Some examples of baseline fitting with detrendr
 
@@
 library(Rcpp)
+library(microbenchmark)
 load_all("../Code/detrendr")
@@
 
-```{r simulate}
+# Compare cholesky decompositions and solvers
+```{r cholesky_comparison}
+nSeq <- seq(500, 20000, 500) 
+k <- 4
+
+timing <- data.frame(n=nSeq, Matrix_chol = NA, Matrix_solve = NA, 
+                     Eigen_chol = NA, chol_solve = NA)
+i <- 1
+for (n in nSeq){
+  D <- get_Dk(n, k)
+  M <-  as(Diagonal(n) + Matrix::crossprod(D), "dgCMatrix")
+  theta <- rnorm(n)
+
+  timing$Matrix_chol[i] <- median(microbenchmark(cholM <- chol(M))$time*1e-3)
+  timing$Eigen_chol[i] <- median(microbenchmark(cholM2 <- chol_eigen(M))$time*1e-3)
+  timing$Matrix_solve[i] <- median(microbenchmark(
+    x1 <- as.numeric(Matrix::solve(cholM, Matrix::solve(t(cholM), theta))))$time*1e-3)
+  timing$chol_solve[i] <- median(microbenchmark(
+    x2 <- chol_solve(cholM, chol_solve(t(cholM), theta, k, FALSE), 
+                   k, TRUE))$time*1e-3)
+  i <- i+1
+}
+
+plot(Matrix_chol~n, timing, ylab="microseconds", 
+     main="Matrix package sparse cholesky decomposition")
+plot(chol_solve~n, timing, ylab ="microseconds", 
+     main="RcppEigen sparse cholesky decomposition")
+plot(chol_solve~n, timing, ylab = "microseconds", 
+     main="My banded cholesky solver (black) vs. Matrix solver (blue)")
+points(Matrix_solve~n, timing, col="blue")
+
+```
+
+## Some examples of baseline fitting with detrendr
+
+```{r simulate, include=FALSE, echo=FALSE}
 
@@
 
-## Current timing experiments
-```{r timing, include=TRUE, echo=TRUE, cache=TRUE}
-nSeq <- seq(100, 1500, 200)
-timing <- data.frame(n=nSeq, R = NA, C=NA)
-tau <- 0.01
-lambda <- 10
-step <- 1
+<!-- ## Current timing experiments -->
+<!-- ```{r timing, include=FALSE, echo=FALSE, cache=TRUE, } -->
+<!-- nSeq <- seq(100, 1500, 200) -->
+<!-- timing <- data.frame(n=nSeq, R = NA, C=NA) -->
+<!-- tau <- 0.01 -->
+<!-- lambda <- 10 -->
+<!-- step <- 1 -->
 
-for (i in 1:length(nSeq)){
-  simData <- simulate(nSeq[i], 3)
-  y <- simData$y
-  theta <- simData$y
-  D <- simData$D
-  eta <- matrix(D%*%theta)
-  M <- simData$M
-  cholM <- simData$cholM
-  timing$R[i] <- system.time( 
-      multi_step_R <- spingarn_multi_step_R(theta, eta, y, D, M, lambda, 
-                                            tau, step, 100))[3]
-  timing$C[i] <- system.time(
-    multi_step <- spingarn_multi_step(theta, eta, y, D, cholM,
-                                     lambda, tau, step, 100))[3]
-}
+<!-- for (i in 1:length(nSeq)){ -->
+<!--   simData <- simulate(nSeq[i], 3) -->
+<!--   y <- simData$y -->
+<!--   theta <- simData$y -->
+<!--   D <- simData$D -->
+<!--   eta <- matrix(D%*%theta) -->
+<!--   M <- simData$M -->
+<!--   cholM <- simData$cholM -->
+<!--   timing$R[i] <- system.time(  -->
+<!--       multi_step_R <- spingarn_multi_step_R(theta, eta, y, D, M, lambda,  -->
+<!--                                             tau, step, 100))[3] -->
+<!--   timing$C[i] <- system.time( -->
+<!--     multi_step <- spingarn_multi_step(theta, eta, y, D, cholM, -->
+<!--                                      lambda, tau, step, 100))[3] -->
+<!-- } -->
 
-plot(C~n, timing, type="l", col="blue", ylab="Elapsed Time (s) per 100 iterations", main = "R code is in black, C in blue")
-lines(R~n, timing, type="l")
-```
+<!-- plot(C~n, timing, type="l", col="blue", ylab="Elapsed Time (s) per 100 iterations", main = "R code is in black, C in blue") -->
+<!-- lines(R~n, timing, type="l") -->
+<!-- ``` -->
 
-## Convergence when n=100, lambda = 10
-```{r convergence, include=TRUE, echo=TRUE, cache=TRUE}
+<!-- ## Convergence when n=100, lambda = 10 -->
+<!-- ```{r convergence, include=TRUE, echo=TRUE, cache=TRUE} -->
 
-n <- 500
-simData <- simulate(n, 3)
+<!-- n <- 500 -->
+<!-- simData <- simulate(n, 3) -->
 
-y <- simData$y
-theta <- simData$y
-D <- simData$D
-eta <- matrix(D%*%theta)
-M <- simData$M
-cholM <- simData$cholM
-tau <- 0.05
-lambda <- 10
-step <- 1
+<!-- y <- simData$y -->
+<!-- theta <- simData$y -->
+<!-- D <- simData$D -->
+<!-- eta <- matrix(D%*%theta) -->
+<!-- M <- simData$M -->
+<!-- cholM <- simData$cholM -->
+<!-- tau <- 0.05 -->
+<!-- lambda <- 10 -->
+<!-- step <- 1 -->
 
-max_iter <- 70000
-rerr <- double(max_iter-1)
-THx <- matrix(NA, n, max_iter)
-for (iter in 1:max_iter) {
-  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
-  theta <- one_step[[1]]
-  eta <- one_step[[2]]
-  THx[,iter] <- prox_f1(theta, y, tau)
-  if (iter > 1){
-    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
-                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
-  }
-}
-theta_last <- prox_f1(theta, y, tau)
+<!-- max_iter <- 70000 -->
+<!-- rerr <- double(max_iter-1) -->
+<!-- THx <- matrix(NA, n, max_iter) -->
+<!-- for (iter in 1:max_iter) { -->
+<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
+<!--   theta <- one_step[[1]] -->
+<!--   eta <- one_step[[2]] -->
+<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
+<!--   if (iter > 1){ -->
+<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
+<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
+<!--   } -->
+<!-- } -->
+<!-- theta_last <- prox_f1(theta, y, tau) -->
 
-par(mfrow=c(2,1), mar=c(3,4,1,1))
-plot(log(rerr), type="l")
+<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
+<!-- plot(log(rerr), type="l") -->
 
-plot(simData$x,simData$f,type='l',col='blue',
-ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
-points(simData$x,y, cex=.6)
-lines(simData$x,theta_last,col='red', lwd=3)
+<!-- plot(simData$x,simData$f,type='l',col='blue', -->
+<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
+<!-- points(simData$x,y, cex=.6) -->
+<!-- lines(simData$x,theta_last,col='red', lwd=3) -->
 
-```
+<!-- ``` -->
 
-## Convergence when n=1000, lambda = 100
-```{r convergence2, include=TRUE, echo=FALSE, cache=TRUE}
+<!-- ## Convergence when n=1000, lambda = 100 -->
+<!-- ```{r convergence2, include=TRUE, echo=FALSE, cache=TRUE} -->
 
-n <- 1000
-simData <- simulate(n, 3)
+<!-- n <- 1000 -->
+<!-- simData <- simulate(n, 3) -->
 
-y <- simData$y
-theta <- simData$y
-D <- simData$D
-eta <- matrix(D%*%theta)
-M <- simData$M
-cholM <- simData$cholM
-tau <- 0.05
-lambda <- 100
-step <- 1
+<!-- y <- simData$y -->
+<!-- theta <- simData$y -->
+<!-- D <- simData$D -->
+<!-- eta <- matrix(D%*%theta) -->
+<!-- M <- simData$M -->
+<!-- cholM <- simData$cholM -->
+<!-- tau <- 0.05 -->
+<!-- lambda <- 100 -->
+<!-- step <- 1 -->
 
-max_iter <- 70000
-rerr <- double(max_iter-1)
-THx <- matrix(NA, n, max_iter)
-for (iter in 1:max_iter) {
-  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
-  theta <- one_step[[1]]
-  eta <- one_step[[2]]
-  THx[,iter] <- prox_f1(theta, y, tau)
-  if (iter > 1){
-    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
-                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
-  }
-}
-theta_last <- prox_f1(theta, y, tau)
+<!-- max_iter <- 70000 -->
+<!-- rerr <- double(max_iter-1) -->
+<!-- THx <- matrix(NA, n, max_iter) -->
+<!-- for (iter in 1:max_iter) { -->
+<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
+<!--   theta <- one_step[[1]] -->
+<!--   eta <- one_step[[2]] -->
+<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
+<!--   if (iter > 1){ -->
+<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
+<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
+<!--   } -->
+<!-- } -->
+<!-- theta_last <- prox_f1(theta, y, tau) -->
 
-par(mfrow=c(2,1), mar=c(3,4,1,1))
-plot(log(rerr), type="l")
+<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
+<!-- plot(log(rerr), type="l") -->
 
-plot(simData$x,simData$f,type='l',col='blue',
-ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
-points(simData$x,y, cex=.6)
-lines(simData$x,theta_last,col='red', lwd=3)
+<!-- plot(simData$x,simData$f,type='l',col='blue', -->
+<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
+<!-- points(simData$x,y, cex=.6) -->
+<!-- lines(simData$x,theta_last,col='red', lwd=3) -->
 
-```
+<!-- ``` -->
 
-## Convergence when n=2000, lambda = 100
-```{r convergence3, include=TRUE, echo=FALSE, cache=TRUE}
+<!-- ## Convergence when n=2000, lambda = 100 -->
+<!-- ```{r convergence3, include=TRUE, echo=FALSE, cache=TRUE} -->
 
-n <- 2000
-simData <- simulate(n, 3)
+<!-- n <- 2000 -->
+<!-- simData <- simulate(n, 3) -->
 
-y <- simData$y
-theta <- simData$y
-D <- simData$D
-eta <- matrix(D%*%theta)
-M <- simData$M
-cholM <- simData$cholM
-tau <- 0.05
-lambda <- 100
-step <- 1
+<!-- y <- simData$y -->
+<!-- theta <- simData$y -->
+<!-- D <- simData$D -->
+<!-- eta <- matrix(D%*%theta) -->
+<!-- M <- simData$M -->
+<!-- cholM <- simData$cholM -->
+<!-- tau <- 0.05 -->
+<!-- lambda <- 100 -->
+<!-- step <- 1 -->
 
-max_iter <- 70000
-rerr <- double(max_iter-1)
-THx <- matrix(NA, n, max_iter)
-for (iter in 1:max_iter) {
-  one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step)
-  theta <- one_step[[1]]
-  eta <- one_step[[2]]
-  THx[,iter] <- prox_f1(theta, y, tau)
-  if (iter > 1){
-    rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/
-                    (1 + norm(THx[,iter,drop=FALSE], 'f'))
-  }
-}
-theta_last <- prox_f1(theta, y, tau)
+<!-- max_iter <- 70000 -->
+<!-- rerr <- double(max_iter-1) -->
+<!-- THx <- matrix(NA, n, max_iter) -->
+<!-- for (iter in 1:max_iter) { -->
+<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
+<!--   theta <- one_step[[1]] -->
+<!--   eta <- one_step[[2]] -->
+<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
+<!--   if (iter > 1){ -->
+<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
+<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
+<!--   } -->
+<!-- } -->
+<!-- theta_last <- prox_f1(theta, y, tau) -->
 
-par(mfrow=c(2,1), mar=c(3,4,1,1))
-plot(log(rerr), type="l")
+<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
+<!-- plot(log(rerr), type="l") -->
 
-plot(simData$x,simData$f,type='l',col='blue',
-ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="")
-points(simData$x,y, cex=.6)
-lines(simData$x,theta_last,col='red', lwd=3)
+<!-- plot(simData$x,simData$f,type='l',col='blue', -->
+<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
+<!-- points(simData$x,y, cex=.6) -->
+<!-- lines(simData$x,theta_last,col='red', lwd=3) -->
 
-```
+<!-- ``` -->

### 2017-06-23 Code/detrendr/R/RcppExports.R

@@ -210,3 +210,16 @@ spingarn_multi_step <- function(theta, eta, y, D, cholM, lambda, tau = 0.05, ste
     .Call('detrendr_spingarn_multi_step', PACKAGE = 'detrendr', theta, eta, y, D, cholM, lambda, tau, step, numberIter, k)
 }
 
+#' @title
+#' Perform cholesky decomposition and solve linear system
+#' @param X sparse matrix
+#' @param yy vector 
+#' @export
+chol_solve_eigen <- function(X, yy) {
+    .Call('detrendr_chol_solve_eigen', PACKAGE = 'detrendr', X, yy)
+}
+
+chol_eigen <- function(X) {
+    .Call('detrendr_chol_eigen', PACKAGE = 'detrendr', X)
+}
+

### 2017-06-23 Code/detrendr/src/RcppExports.cpp

@@ -2,6 +2,7 @@
 // Generator token: 10BE3573-1514-4C36-9D1C-5A225CD40393
 
 #include <RcppArmadillo.h>
+#include <RcppEigen.h>
 #include <Rcpp.h>
 
 using namespace Rcpp;
@@ -165,4 +166,49 @@ BEGIN_RCPP
     rcpp_result_gen = Rcpp::wrap(spingarn_multi_step(theta, eta, y, D, cholM, lambda, tau, step, numberIter, k));
     return rcpp_result_gen;
 END_RCPP
+}
+// chol_solve_eigen
+Eigen::VectorXd chol_solve_eigen(Eigen::MappedSparseMatrix<double> X, Eigen::VectorXd yy);
+RcppExport SEXP detrendr_chol_solve_eigen(SEXP XSEXP, SEXP yySEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< Eigen::MappedSparseMatrix<double> >::type X(XSEXP);
+    Rcpp::traits::input_parameter< Eigen::VectorXd >::type yy(yySEXP);
+    rcpp_result_gen = Rcpp::wrap(chol_solve_eigen(X, yy));
+    return rcpp_result_gen;
+END_RCPP
+}
+// chol_eigen
+SEXP chol_eigen(Eigen::MappedSparseMatrix<double> X);
+RcppExport SEXP detrendr_chol_eigen(SEXP XSEXP) {
+BEGIN_RCPP
+    Rcpp::RObject rcpp_result_gen;
+    Rcpp::RNGScope rcpp_rngScope_gen;
+    Rcpp::traits::input_parameter< Eigen::MappedSparseMatrix<double> >::type X(XSEXP);
+    rcpp_result_gen = Rcpp::wrap(chol_eigen(X));
+    return rcpp_result_gen;
+END_RCPP
+}
+
+static const R_CallMethodDef CallEntries[] = {
+    {"detrendr_chol_solve", (DL_FUNC) &detrendr_chol_solve, 4},
+    {"detrendr_prox_quantile", (DL_FUNC) &detrendr_prox_quantile, 3},
+    {"detrendr_prox_f1", (DL_FUNC) &detrendr_prox_f1, 4},
[…]

### 2017-06-23 Code/detrendr/src/prox.cpp

@@ -17,7 +17,7 @@ using namespace arma;
 //' @param upper indicates wheter cholM is upper or lower triangular
 //' @export
 //' 
-//[[Rcpp::export]]
+// [[Rcpp::export]]
 arma::vec chol_solve(arma::sp_mat cholM, 
                      arma::vec b, 
                      int k,
@@ -147,7 +147,7 @@ arma::vec prox_f2(arma::vec eta,
 //' @param tau quantile parameter
 //' @param step step-size
 //' @export 
-//[[Rcpp::export]]
+// [[Rcpp::export]]
 void prox(arma::vec& theta, 
                 arma::vec& eta, 
                 arma::vec y, 
@@ -170,7 +170,7 @@ void prox(arma::vec& theta,
 //' @param tau quantile parameter
 //' @param step step-size
 //' @export 
-//[[Rcpp::export]]
+// [[Rcpp::export]]
 Rcpp::List prox_test(arma::vec theta, 
                      arma::vec eta, 
                      arma::vec y, 
@@ -190,7 +190,7 @@ Rcpp::List prox_test(arma::vec theta,
 //' n <- 5
 //' D1 <- get_D1(n)
 //' @export
-//[[Rcpp::export]]
+// [[Rcpp::export]]
 arma::sp_mat get_D1(int n){
   int numberNonZero = 2*(n-1);
   arma::vec values = ones<vec>(numberNonZero);
@@ -210,7 +210,7 @@ arma::sp_mat get_D1(int n){
 //' @param n length of input
 //' @param k order of the derivative
 //' @export
-//[[Rcpp::export]]
+// [[Rcpp::export]]
 arma::sp_mat get_Dk(int n, 
                      int k){
   arma::sp_mat D = get_D1(n);
@@ -232,7 +232,7 @@ arma::sp_mat get_Dk(int n,
 //' @param D differencing matrix
 //' @param cholM upper triangular cholesky decomposition of  I + DtD
 //' @export
-//[[Rcpp::export]]
+// [[Rcpp::export]]
 void project_V(arma::vec& theta,
                      arma::vec& eta,
                      arma::sp_mat D, 
@@ -258,7 +258,7 @@ void project_V(arma::vec& theta,
 //' @param tau quantile parameter
 //' @param step step-size
 //' @export
-//[[Rcpp::export]]
+// [[Rcpp::export]]
 void spingarn_one_step(arma::vec& theta, 
                              arma::vec& eta, 
[…]

### 2017-07-05 Notes/detrendify_spingarn.tex
Added reference on air quality.
Notes and timing comparison updates

@@
 \usepackage{booktabs}
+\graphicspath{{./Figures/}}
 
@@
 
-We introduce a linear time algorithm for removing trends in time series.
+\subsection{Background}
+
+\textbf{Oh, Lee, and Nychka 2011 - Fast Nonparametric Quantile Regression With Arbitrary Smoothing Method} Unified framework for non-parametric quantile regression by approximating check loss function with quadratic loss near zero.\\
+\textbf{Kim, Koh, Boyd, and Gorinevsky 2009} - Introduce the concept of l1 trend filtering.\\
+\textbf{Tibshirani, 2014 - Adaptive piecewise polynomial estimation via trend filtering} Describes properties of trend filtering using quadratic loss.
+Shows that trend filtering estimates adapt to the local level of smoothness much better than smoothing splines, and exhibit
+a remarkable similarity to locally adaptive regression splines. Prove that (with the right choice of tuning parameter) the trend filtering estimate
+converges to the true underlying function at the minimax rate for functions whose kth derivative is of bounded variation.\\
+\textbf{Ning, Selesnick, Duval 2014 - Chromatogram baseline estimation and denoising using sparsity (BEADS)} Address problem of estimating a smooth baseline in noisy data with drift. \\
+We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate that can be computes using a linear time algorithm for removing trends in time series.
+Can be extended to model multiple quantiles and ensure non-crossing.
+
+\subsection{Application}
+
+	 \begin{figure}[!h] 
+	 	\caption{Raw Data - three collocated sensors}
+		\includegraphics[width = 5in]{spodRaw.png}
+	\end{figure}
+	 \begin{figure}[!h] 
+	 	\caption{Detrended data - three collocated sensors}
+	\includegraphics[width = 5in]{spodDetrended.png}
+	\end{figure}
 
@@
 \begin{itemize}
-\item Air quality 
+\item Air quality \cite{Apte2017}
 \item ECG \cite{Sanyal2012}
@@
 We argue that it is better solved by Spingarn's method of partial inverses.
+
+\section{Related Work}
+
+\begin{itemize}
+	\item Quantile splines \cite{Oh2011}
+	\item BEADS 
+\end{itemize}
 
@@
 
+\section{Numerical Studies}
+
+We compare our detrending method with BEADS and the more general nonparametric quantile method introduced in  \cite{Oh2011}.
+
 \bibliographystyle{plain}

### 2017-07-05 Notes/timing_and_convergence_examples.Rmd
Notes and timing comparison updates

--- 2017-06-23 Notes/timing_and_convergence_examples.Rmd
@@
 
-# Compare cholesky decompositions and solvers
-```{r cholesky_comparison}
-nSeq <- seq(500, 20000, 500) 
-k <- 4
+# Compare cholesky decompositions and solvers as n increases
+```{r cholesky_comparison, cache=TRUE}
+nSeq <- seq(500, 50000, 500) 
+k <- 3
 
@@
 for (n in nSeq){
+  D <- get_Dk(n, k)
+  M <-  as(Diagonal(n) + Matrix::crossprod(D), "dgCMatrix")
+  theta <- rnorm(n)
+
+  cholM <- chol(M)
+  timing$Matrix_solve[i] <- median(microbenchmark(
+    x1 <- as.numeric(Matrix::solve(cholM, Matrix::solve(t(cholM), theta))))$time*1e-3)
+  timing$chol_solve[i] <- median(microbenchmark(
+    x2 <- chol_solve(cholM, chol_solve(t(cholM), theta, k, FALSE), 
+                   k, TRUE))$time*1e-3)
+  i <- i+1
+}
+
+t1 <- M%*%x1
+plot(as.numeric(t1)~theta)
+
+plot(chol_solve~n, timing, ylab = "microseconds", 
+     main="My banded cholesky solver (black) vs. Matrix solver (blue)")
+points(Matrix_solve~n, timing, col="blue")
+abline(0, .12)
+
+```
+
+# Compare cholesky decompositions and solvers as k increases
+```{r cholesky_comparison_k, cache=TRUE}
+n <- 40
+kSeq <- seq(2, 10, 1)
+
+timing <- data.frame(k=kSeq, Matrix_chol = NA, Matrix_solve = NA, 
+                     Eigen_chol = NA, chol_solve = NA)
+i <- 1
+for (k in kSeq){
   D <- get_Dk(n, k)
@@
 
-plot(Matrix_chol~n, timing, ylab="microseconds", 
+t1 <- M%*%x1
+plot(as.numeric(t1)~theta)
+
+plot(Matrix_chol~k, timing, ylab="microseconds", 
      main="Matrix package sparse cholesky decomposition")
-plot(chol_solve~n, timing, ylab ="microseconds", 
+plot(Eigen_chol~k, timing, ylab ="microseconds", 
      main="RcppEigen sparse cholesky decomposition")
-plot(chol_solve~n, timing, ylab = "microseconds", 
+plot(chol_solve~k, timing, ylab = "microseconds", 
      main="My banded cholesky solver (black) vs. Matrix solver (blue)")
-points(Matrix_solve~n, timing, col="blue")
+points(Matrix_solve~k, timing, col="blue")
 
 ```
+# Algorithm order
+```{r algorithm order, cache=TRUE}
+nSeq <- seq(500, 15000, 500) 
+k <- 3
 
-## Some examples of baseline fitting with detrendr
+timing <- data.frame(n=nSeq, multistep = NA)
+i <- 1
+for (n in nSeq){
+  D <- get_Dk(n, k)
+  M <-  as(Diagonal(n) + Matrix::crossprod(D), "dgCMatrix")
+  theta <- rnorm(n)
+  lambda0 <- 2*10^(-6)
+  lambda <- lambda0*n^(k-1)/factorial(k-1)
+  y <- rnorm(n)
+  eta <- matrix(D%*%theta)
+  tau <- 0.05
+  cholM <- chol(M)
+  timing$multistep[i] <- median(microbenchmark(
+      multi_step <- spingarn_multi_step(theta, eta, y, D, cholM,
+                                 lambda, tau, 1, 100, k)
+  )$time*1e-9)
 
-```{r simulate, include=FALSE, echo=FALSE}
+  i <- i+1
+  print(i)
+}
 
-simulate <- function(n, k){
-  x <- seq(1/n, 1, length.out=n)
-  f <- 2*(x + 2)^2 + 3*cos(3*pi*x)
-  tau <- 5e4
-  g1 <- 40*exp(-tau*(x-0.3)^2)
-  g2 <- 30*exp(-tau/2*(x-0.58)^2)
-  g4 <- 45*exp(-tau/7*(x-0.6)^2)
-  g3 <- 37*exp(-tau*(x-0.63)^2)
-  y <- f + g1 + g2 + g3 + g4 + rnorm(n)
-  D <- get_Dk(n, k)
-  M <- diag(n) + crossprod(D)
-  cholM <- as.matrix(chol(M))
-  return(list(y=y, f=f, x=x, D=D, M=M, cholM=cholM))
-}
+plot(multistep~n, timing, ylab="Seconds", xlab = "n", main = "100 steps of algorithm")
+abline(0, 2.95*1e-6)
+
 ```
-
-<!-- ## Current timing experiments -->
-<!-- ```{r timing, include=FALSE, echo=FALSE, cache=TRUE, } -->
-<!-- nSeq <- seq(100, 1500, 200) -->
-<!-- timing <- data.frame(n=nSeq, R = NA, C=NA) -->
-<!-- tau <- 0.01 -->
-<!-- lambda <- 10 -->
-<!-- step <- 1 -->
-
-<!-- for (i in 1:length(nSeq)){ -->
-<!--   simData <- simulate(nSeq[i], 3) -->
-<!--   y <- simData$y -->
-<!--   theta <- simData$y -->
-<!--   D <- simData$D -->
-<!--   eta <- matrix(D%*%theta) -->
-<!--   M <- simData$M -->
-<!--   cholM <- simData$cholM -->
-<!--   timing$R[i] <- system.time(  -->
-<!--       multi_step_R <- spingarn_multi_step_R(theta, eta, y, D, M, lambda,  -->
-<!--                                             tau, step, 100))[3] -->
-<!--   timing$C[i] <- system.time( -->
-<!--     multi_step <- spingarn_multi_step(theta, eta, y, D, cholM, -->
-<!--                                      lambda, tau, step, 100))[3] -->
-<!-- } -->
-
-<!-- plot(C~n, timing, type="l", col="blue", ylab="Elapsed Time (s) per 100 iterations", main = "R code is in black, C in blue") -->
-<!-- lines(R~n, timing, type="l") -->
-<!-- ``` -->
-
-<!-- ## Convergence when n=100, lambda = 10 -->
-<!-- ```{r convergence, include=TRUE, echo=TRUE, cache=TRUE} -->
-
-<!-- n <- 500 -->
-<!-- simData <- simulate(n, 3) -->
-
-<!-- y <- simData$y -->
-<!-- theta <- simData$y -->
-<!-- D <- simData$D -->
-<!-- eta <- matrix(D%*%theta) -->
-<!-- M <- simData$M -->
-<!-- cholM <- simData$cholM -->
-<!-- tau <- 0.05 -->
-<!-- lambda <- 10 -->
-<!-- step <- 1 -->
-
-<!-- max_iter <- 70000 -->
-<!-- rerr <- double(max_iter-1) -->
-<!-- THx <- matrix(NA, n, max_iter) -->
-<!-- for (iter in 1:max_iter) { -->
-<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
-<!--   theta <- one_step[[1]] -->
-<!--   eta <- one_step[[2]] -->
-<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
-<!--   if (iter > 1){ -->
-<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
-<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
-<!--   } -->
-<!-- } -->
-<!-- theta_last <- prox_f1(theta, y, tau) -->
-
-<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
-<!-- plot(log(rerr), type="l") -->
-
-<!-- plot(simData$x,simData$f,type='l',col='blue', -->
-<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
-<!-- points(simData$x,y, cex=.6) -->
-<!-- lines(simData$x,theta_last,col='red', lwd=3) -->
-
-<!-- ``` -->
-
-<!-- ## Convergence when n=1000, lambda = 100 -->
-<!-- ```{r convergence2, include=TRUE, echo=FALSE, cache=TRUE} -->
-
-<!-- n <- 1000 -->
-<!-- simData <- simulate(n, 3) -->
-
-<!-- y <- simData$y -->
-<!-- theta <- simData$y -->
-<!-- D <- simData$D -->
-<!-- eta <- matrix(D%*%theta) -->
-<!-- M <- simData$M -->
-<!-- cholM <- simData$cholM -->
-<!-- tau <- 0.05 -->
-<!-- lambda <- 100 -->
-<!-- step <- 1 -->
-
-<!-- max_iter <- 70000 -->
-<!-- rerr <- double(max_iter-1) -->
-<!-- THx <- matrix(NA, n, max_iter) -->
-<!-- for (iter in 1:max_iter) { -->
-<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
-<!--   theta <- one_step[[1]] -->
-<!--   eta <- one_step[[2]] -->
-<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
-<!--   if (iter > 1){ -->
-<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
-<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
-<!--   } -->
-<!-- } -->
-<!-- theta_last <- prox_f1(theta, y, tau) -->
-
-<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
-<!-- plot(log(rerr), type="l") -->
-
-<!-- plot(simData$x,simData$f,type='l',col='blue', -->
-<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
-<!-- points(simData$x,y, cex=.6) -->
-<!-- lines(simData$x,theta_last,col='red', lwd=3) -->
-
-<!-- ``` -->
-
-<!-- ## Convergence when n=2000, lambda = 100 -->
-<!-- ```{r convergence3, include=TRUE, echo=FALSE, cache=TRUE} -->
-
-<!-- n <- 2000 -->
-<!-- simData <- simulate(n, 3) -->
-
-<!-- y <- simData$y -->
-<!-- theta <- simData$y -->
-<!-- D <- simData$D -->
-<!-- eta <- matrix(D%*%theta) -->
-<!-- M <- simData$M -->
-<!-- cholM <- simData$cholM -->
-<!-- tau <- 0.05 -->
-<!-- lambda <- 100 -->
-<!-- step <- 1 -->
-
-<!-- max_iter <- 70000 -->
-<!-- rerr <- double(max_iter-1) -->
-<!-- THx <- matrix(NA, n, max_iter) -->
-<!-- for (iter in 1:max_iter) { -->
-<!--   one_step <- spingarn_one_step_R(theta, eta, y, D, M, lambda, tau, step) -->
-<!--   theta <- one_step[[1]] -->
-<!--   eta <- one_step[[2]] -->
-<!--   THx[,iter] <- prox_f1(theta, y, tau) -->
-<!--   if (iter > 1){ -->
-<!--     rerr[iter] <- norm(as.matrix(THx[,iter] - THx[,iter-1]),'f')/ -->
-<!--                     (1 + norm(THx[,iter,drop=FALSE], 'f')) -->
-<!--   } -->
-<!-- } -->
-<!-- theta_last <- prox_f1(theta, y, tau) -->
-
-<!-- par(mfrow=c(2,1), mar=c(3,4,1,1)) -->
-<!-- plot(log(rerr), type="l") -->
-
-<!-- plot(simData$x,simData$f,type='l',col='blue', -->
-<!-- ylim=c(min(y),max(y)), lwd=3,ylab="", xlab="") -->
-<!-- points(simData$x,y, cex=.6) -->
-<!-- lines(simData$x,theta_last,col='red', lwd=3) -->
-
-<!-- ``` -->

### 2017-07-05 Code/detrendr/src/prox.cpp

@@ -332,10 +332,21 @@ Rcpp::List spingarn_multi_step(arma::vec theta,
                              double numberIter=1, 
                              int k=3){
   arma::vec theta_cp = theta;
+  arma::vec theta_cur = theta_cp;
   arma::vec eta_cp = eta;
+  double rerr = 1;
   for (int i = 0; i < numberIter; i++){
     spingarn_one_step(theta_cp, eta_cp, y, D, cholM, lambda, tau, step, k);
+    if (i % 10 == 0){
+      Rcpp::checkUserInterrupt();
+      rerr = norm(theta_cp - theta_cur,2)/
+        (1+norm(theta_cp,2));
+      if (rerr < pow(10, -6)){
+        break;
+      }
+      theta_cur = theta_cp;
+    }
   }
 
-  return Rcpp::List::create(theta=theta_cp, eta=eta_cp);
+  return Rcpp::List::create(theta_cp, eta_cp, rerr);
 }

### 2017-07-12 Code/detrendr/R/warmStart.R
warmStart added

@@ -1,0 +1,34 @@
+#' @title 
+#' Warm start for Spingarn algorithm 
+#' @description 
+#' Get approximation of initial values by aggregating data and fitting on 
+#' smaller dataset
+#' @param y original data to fit baseline
+#' @param k order of differencing matrix
+#' @param lambda regularization parameter
+#' @param tau quantile level
+#' @param reduction number of observations to aggregate
+#' @export
+warmStart <- function(y, k, lambda, tau, reduction){
+  x <- seq(1, length(y), 1)
+  df <- data.frame(y=y, x=x)
+  df$xBreaks <- cut(df$x, seq(1, length(x), reduction))
+  dfAgg <- ddply(df, .(xBreaks), summarize, 
+                 yMean = mean(y))
+  
+  y2 <- dfAgg$yMean
+  theta <- y2
+  n <- length(y2)
+  D <- get_Dk(n, k)
+  M <- Diagonal(n) + Matrix::crossprod(D)
+  cholM <- Matrix::chol(M)
+  eta <- matrix(D%*%theta)
+  multi_step <- spingarn_multi_step(theta, eta, y2, D, cholM,
+                                    lambda, tau, 1, 10000, k)
+  dfAgg$theta <- prox_f1(multi_step[[1]], y2, tau)
+  # plot(y2, type="l")
+  # lines(dfAgg$theta, col="red")
+  
+  dfAll <- merge(df, dfAgg)
+  return(dfAll$theta)
+}

### 2018-01-10 Manuscript/detrendify_JCGS.tex

--- 2017-06-02 Notes/detrendify_spingarn.tex
@@
-%
-% A simple LaTeX template for Books
-%  (c) Aleksander Morgado <aleksander@es.gnu.org>
-%  Released into public domain
-%
+\documentclass[12pt]{article}
+\RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
+\usepackage{bm,graphicx,psfrag,epsf}
+\usepackage{enumerate}
+\usepackage{natbib}
+%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+%\RequirePackage{hypernat}
+\usepackage{paralist, booktabs, multirow}
+%\usepackage{url}
 
-\documentclass{article}
-\usepackage[a4paper, top=3cm, bottom=3cm]{geometry}
-\usepackage[latin1]{inputenc}
-\usepackage{setspace}
-\usepackage{fancyhdr}
-\usepackage{tocloft}
+\newtheorem{lemma}{Lemma}
+\newcommand{\blind}{0}
+\newcommand{\trace}{\text{Trace}}
+\numberwithin{equation}{section}
+\theoremstyle{plain}
+%\newtheorem{thm}{Theorem}[section]
+%\newtheorem{mydef}{Definition}
+%\newtheorem{remark}{Remark}
+
+%% ----------------------------------------------------------------------
+%% Packages
+%% ----------------------------------------------------------------------
+%\usepackage{amsmath,amsfonts,amssymb}
+
+\usepackage{amssymb}
 \usepackage[usenames]{color}
-\RequirePackage[OT1]{fontenc}
-\usepackage{amsmath,amsfonts,amssymb,amsthm}
-\usepackage{graphicx}
-\usepackage{booktabs}
-
-%% For special lists like inparaenum, compactenum, compactitem
-\usepackage{paralist}
-%% For \boldsymbol
-\usepackage{amsbsy}
+%\usepackage{amsbsy,amsthm,amsmath,amssymb}
 
@@
 
-%% For \bm (bold math)
-\usepackage{bm}
+%% For \llbracket and \rrbracket, \varoast, \varoslash
+\usepackage{stmaryrd}
+
+%% For \boldsymbol
+%\usepackage{amsbsy}
+
+%%% For \bm (bold math)
+%\usepackage{bm}
 
@@
 
+%% For special lists like inparaenum, compactenum, compactitem
+\usepackage{paralist}
+
+%% Turn on/off options for PDF
+\usepackage{ifpdf}
+
+%% Graphics
+\usepackage{graphicx}
+\ifpdf
+\usepackage{epstopdf}
+% REMOVE NEXT LINE TO REVERT TO PNG FILES
+\DeclareGraphicsExtensions{.eps,.pdf}
+\else
+\usepackage{epsfig}
+\fi
+
 %% For \subfloat command
 \usepackage{subfig}
-
-%% Algorithm
-\usepackage{algorithm,algorithmic}
-%\usepackage{algpseudocode}
-
-%% Useful shorthand for converting equations to be inline
-\newenvironment{inlinemath}{$}{$}
-
-%% Define hypen
-\mathchardef\mhyphen="2D
 
@@
 
+%% For tabulary
+\usepackage{tabulary}
+
+%% Show labels
+%\usepackage[notref,notcite]{showkeys}
+
+%% Algorithm
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+
 %% ----------------------------------------------------------------------
@@
 \ifpdf
-\newcommand{\Sec}[1]{\hyperref[sec:#1]{\S\ref*{sec:#1}}} %section
+\newcommand{\Sec}[1]{\hyperref[sec:#1]{Section~\ref*{sec:#1}}} %section
 \newcommand{\App}[1]{\hyperref[sec:#1]{Appendix~\ref*{sec:#1}}} %appendix
@@
 \else
-\newcommand{\Sec}[1]{{\S\ref{sec:#1}}} %section
+\newcommand{\Sec}[1]{{Section~\ref{sec:#1}}} %section
 \newcommand{\App}[1]{{Appendix~\ref{sec:#1}}} %appendix
@@
 \newcommand{\Lem}[1]{{Lemma~\ref{lem:#1}}} %lemma
-\newcommand{\Prop}[1]{{Proposition~\ref{prop:#1}}} %proposition
+\newcommand{\Prop}[1]{{Property~\ref{prop:#1}}} %property
 \newcommand{\Cor}[1]{{Corollary~\ref{cor:#1}}} %corollary
@@
 \newcommand{\Ex}[1]{{Example~\ref{ex:#1}}} %example
-\newcommand{\Reg}[1]{{R~\ref*{reg:#1}}} %regularity condition
+\newcommand{\As}[1]{{Assumption~\ref{as:#1}}} %assumption
+\newcommand{\Reg}[1]{{R~\ref{reg:#1}}} %regularity condition
+\newcommand{\AlgLine}[2]{{line~\ref{line:#2} of Algorithm~\ref{alg:#1}}}
+\newcommand{\AlgLines}[3]{{lines~\ref{line:#2}--\ref{line:#3} of Algorithm~\ref{alg:#1}}}
 \fi
+
+\usepackage{paralist, booktabs, multirow}
+
+%% ----------------------------------------------------------------------
+%% Environments
+%% ----------------------------------------------------------------------
+
+%% Assumptions environments
+\newtheorem{assumption}{Assumption}[section]
+%% Special proof where box is inside final equation
+% The function \proof and \endproof are defined inside the SIAM class file.
+% Be sure to use \myproofend inside the final equation to produce the
+% special box that signifies the end of the proof.
+\newenvironment{myproof}{\proof}{}
+\newcommand{\myproofend}{\quad\endproof}
+\newtheorem{theorem}{Theorem}[section]
+
+%Added by Arvind
+\usepackage[normalem]{ulem}
+\newcommand{\arvind}[1]{{\color{red} #1}}
+
+%%%%%
+
+%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+%
+%\RequirePackage[OT1]{fontenc}
+%\RequirePackage{amsthm,amsmath}
+%\RequirePackage [round,authoryear]{natbib}
+%\RequirePackage{hypernat}
+%
+%\usepackage{amsbsy,amsthm,amsmath,amssymb}
+%\usepackage{graphicx,booktabs}
+%%\usepackage{subfig}
+\usepackage{mathtools}
+\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{tikz}
+\usetikzlibrary{arrows}
+%\setlength{\oddsidemargin}{.75in}
+%\setlength{\evensidemargin}{.75in}
+%\setlength{\textwidth}{5in}
+\newtheorem{proposition}{Proposition}[section]
+%\newtheorem{corollary}{Corollary}[section]
+%\newtheorem{example}{Example}[section]
+%\newtheorem{definition}{Definition}[section]
 
@@
 
+%\startlocaldefs
+%\numberwithin{equation}{section}
+%\theoremstyle{plain}
+%\newtheorem{thm}{Theorem}[section]
+%\endlocaldefs
+
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+\pdfminorversion=4
+
 %% ----------------------------------------------------------------------
-%% Environments
+%% Main Document
 %% ----------------------------------------------------------------------
+\begin{document}
 
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{definition}[theorem]{Definition}
-%% Assumptions environment
-\newtheorem{assumption}[theorem]{Assumption}
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+{#1}\small\normalsize} \spacingset{1}
 
-\begin{document}
-\title{Quantile Trend Filtering via Spingarn's Method}
-\noindent
-\textbf{Quantile Trend Filtering via Spingarn's Method}\\
-\textbf{Date: \today}\\
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\if0\blind
+{
+  \title{\bf Quantile Trend Filtering via Spingarn's Method}
+  \author{Halley Brantley\thanks{
+    Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: hlbrantl@ncsu.edu)} \,
+         Joseph Guinness\thanks{
+    Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: jsguinne@ncsu.edu)} \,
+    and
+%    and
+    Eric C. Chi\thanks{Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: eric$\_$chi@ncsu.edu).}    \\}
+    \date{}
+  \maketitle
+} \fi
+
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\LARGE\bf Title}
+\end{center}
+  \medskip
+} \fi
+
+\bigskip
+\begin{abstract}
+Abstract
+\end{abstract}
+
+\noindent%
+{\it Keywords:}  Key words
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
 %% ----------------------------------------------------------------------
@@
 
-We introduce a linear time algorithm for removing trends in time series.
+\subsection{Background}
+
+\begin{itemize}
+\item \cite{Oh2011} unified framework for non-parametric quantile regression by approximating check loss function with quadratic loss near zero.
+\item \cite{KoenkerNgPortnoy1994} introduce quantile smoothing splines. We need to discuss how our approach is different.
+\item \cite{Kim2009} introduce the concept of $\ell_1$-trend filtering.
+\item \cite{Tib2014} describes properties of trend filtering using quadratic loss.
+Shows that trend filtering estimates adapt to the local level of smoothness much better than smoothing splines, and exhibit a remarkable similarity to locally adaptive regression splines.
+Prove that (with the right choice of tuning parameter) the trend filtering estimate
+converges to the true underlying function at the minimax rate for functions whose kth derivative is of bounded variation.
+\item \cite{Ning2014} address problem of estimating a smooth baseline in noisy data with drift.
+\end{itemize}
+We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate that can be computes using a linear time algorithm for removing trends in time series.
+The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
+Moreover we extend the basic framework to model multiple quantiles and ensure non-crossing.
+
+\subsection{Application}
+
+	 \begin{figure}[!h] 
+	 	\caption{Raw Data - three collocated sensors}
+		\includegraphics[width = 5in]{Figures/spodRaw.png}
+	\end{figure}
+	 \begin{figure}[!h] 
+	 	\caption{Detrended data - three collocated sensors}
+	\includegraphics[width = 5in]{Figures/spodDetrended.png}
+	\end{figure}
 
@@
 \begin{itemize}
-\item Air quality 
-\item ECG \cite{Sanyal2012}
-\item Chromatogram baseline estimation  \cite{Ning2014, Ilewicz2015}
-\item Galaxy spectrum baseline estimation \cite{Ilewicz2015,Bacher2016}
+\item Air quality \citep{Apte2017}
+\item ECG \citep{Sanyal2012}
+\item Chromatogram baseline estimation  \citep{Ning2014, Ilewicz2015}
+\item Galaxy spectrum baseline estimation \citep{Ilewicz2015,Bacher2016}
 \item Identifying absorption dips from black body radiation
@@
 Need to decide between detrending versus detrending + denoising.
-The BEADS \cite{Ning2014} method does both.
+The BEADS \citep{Ning2014} method does both.
 We may wish to focus on detrending and then use wavelet SURE denoising as a postprocessing step, i.e.
@@
 	\item Convergence of the algorithm
-	\item Convergence rate \cite{He2012, He2015, Davis2017}
+	\item Convergence rate \citep{He2012, He2015, Davis2017}
 	\item Timing experiments of LP versus Spingarn
@@
 Recall that the median is the 50th percentile or 0.5-quantile, namely half of the $\VE{y}{i}$ are less than or equal to $\mu$ and the other half is greater than or equal to $\mu$.
-The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ to give us quantile regression \cite{Koenker1978}.
+The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ to give us quantile regression \citep{Koenker1978}.
 
@@
 
-In the trend filtering problem \cite{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $\V{y} \in \Real^n$ by solving the following convex problem.
+In the trend filtering problem \citep{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $\V{y} \in \Real^n$ by solving the following convex problem.
 \begin{eqnarray*}
@@
 
+\section{Related Work}
+
+\begin{itemize}
+	\item Quantile splines \citep{Oh2011}
+	\item BEADS 
+\end{itemize}
+
 %% ----------------------------------------------------------------------
@@
 
-We first review Spingarn's method \cite{Spingarn1985}, which solves the following equality constrained convex problem:
+We first review Spingarn's method \citep{Spingarn1985}, which solves the following equality constrained convex problem:
 \begin{equation}
@@
 The mapping $P_V$ is the orthogonal projection onto the set $V$.
-Note that the algorithm iterates $\Vn{x}{k}$ will converge to a solution to problem $\Eqn{spingarn}$ \cite{Combettes2005}.
+Note that the algorithm iterates $\Vn{x}{k}$ will converge to a solution to problem $\Eqn{spingarn}$ \citep{Combettes2005}.
 We need to experiment with different over / under-relaxation parameters $\lambda^{(k)}$ and step sizes $t$.
@@
 This linear solve can be accomplished in $\mathcal{O}(n(k + 1)^2)$. The first solve using a banded Cholesky decomposition requires $\mathcal{O}(n(k + 1)^2)$. Subsequent solves require
-$\mathcal{O}(n(k+1))$. We can use RccpArmadillo to do banded Cholesky \cite{Eddelbuettel2014}.
+$\mathcal{O}(n(k+1))$. We can use RccpArmadillo to do banded Cholesky \citep{Eddelbuettel2014}.
 
@@
 \item The overall computational complexity is essentially linear $\mathcal{O}(nk^2)$ for the initial banded Cholesky decomposition and the per-iteration complexity is $\mathcal{O}(nk)$.
-\item One could also apply Anderson acceleration \cite{Walker2011} to reduce the number of Spingarn updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
+\item One could also apply Anderson acceleration \citep{Walker2011} to reduce the number of Spingarn updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
 \end{itemize}
@@
 
-Following the specialized ADMM algorithm for trend filtering \cite{Ramdas2016}, we can also take advantage of fast exact solvers of the one-dimensional fussed lasso problem \cite{davies2001, Johnson2013}.
+Following the specialized ADMM algorithm for trend filtering \citep{Ramdas2016}, we can also take advantage of fast exact solvers of the one-dimensional fussed lasso problem \citep{davies2001, Johnson2013}.
 The first method is based on taut strings, and the second is based on dynamic programming.
 Both of these methods run in linear time.
-A third exact method with worst case quadratic penalty but linear time in typical cases was proposed in \cite{Condat2013}.
+A third exact method with worst case quadratic penalty but linear time in typical cases was proposed by \cite{Condat2013}.
 C code is available for all three methods.
@@
 Difference with version 1: If we are interested in a 3rd order penalization, then $\M{D} = \Mn{D}{2}$ in version 2 as opposed to $\M{D} = \Mn{D}{3}$ in version 1.
-The proximal mapping for $f_2$ is solved using one of the exact solvers \cite{davies2001, Johnson2013, Condat2013}.
+The proximal mapping for $f_2$ is solved using one of the exact solvers \citep{davies2001, Johnson2013, Condat2013}.
 Show some timing results between the different versions.
@@
 
-\bibliographystyle{plain}
+\section{Numerical Studies}
+
+We compare our detrending method with BEADS and the more general nonparametric quantile method introduced by \cite{Oh2011}.
+
+\bibliographystyle{asa}
 \bibliography{detrendify}

### 2018-10-01 Notes/100118_meeting_notes.Rmd
reorganizing

---
title: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(devtools)
require(Matrix)
require(ggplot2)
load_all("../detrendify/Code/detrendr")
```

## R Markdown

```{r dataimport}
dataDir <- "../../Desktop/SPod_Data/TestRange_Dec2017"
datafiles <- dir(dataDir, pattern=".csv", full.names=TRUE)
spod <- read.csv(datafiles[3], header=TRUE,  na.strings = "N/A")
spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
                                 format= "%m/%d/%Y %H:%M:%S"))

node <- "f"
pidCol <- paste(node, "SPOD.PID..V.", sep=".")
spodNode <- spod[, c("time", pidCol)]
names(spodNode)[2] <- c("pid")
spodNode <- subset(spodNode, !is.na(pid))
plot(pid~time, spodNode, type="l", main = "raw")

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-10-01 Code/detrendr/R/qr_utils.R

@@ -1,3 +1,6 @@
+#' @title 
+#' Envelope method 
+#' @export
 quantreg_trendfilter = function(y, q=0.5, k=3L, lambda = 1, rel_tol = 1e-4, max_it = 100) {
   # fits a nonparametric quantile regression model via polynomial trend filtering
   # y: ordered data points, assumed to be on a regularly spaced grid
@@ -24,7 +27,7 @@ quantreg_trendfilter = function(y, q=0.5, k=3L, lambda = 1, rel_tol = 1e-4, max_
   while(!converged && it_counter <= max_it) {
     # Compute the weights and pseudo-data in Taylor approximation of log likelihood
 
-    weights = sign(residual)/(residual + 1e-10) 
+    weights = sign(residual)/(residual + 1e-10) + 1e-11
     weights[is.na(weights)]
     z = y - kappa/weights
     tfk = glmgen::trendfilter(z, weights=weights, k = k, 
@@ -36,8 +39,6 @@ quantreg_trendfilter = function(y, q=0.5, k=3L, lambda = 1, rel_tol = 1e-4, max_
     rel_norm[it_counter+1] = travel/(sum(new_residual^2) + rel_tol)
     converged =  rel_norm[it_counter+1] < rel_tol
     it_counter = it_counter + 1
-    #points(weights)
-    lines(beta_hat, col="red")
   }
   list(beta = beta_hat, tfk=tfk, rel_norm = rel_norm)
 }

### 2018-10-26 Notes/Gurobi_notes.Rmd
added smooth relaxation functions, notes on Gurobi optimization

---
title: "detrendr Meeting notes"
output: html_document
---

```{r setup, include=FALSE}
library(gurobi)
library(plyr)
library(devtools)
library(microbenchmark)
library(trustOptim)
library(Rglpk)
library(tidyverse)
library(parallel)
library(lubridate)
library(lbfgs)
load_all("../Code/detrendr")
```

```{r load, include = FALSE}
dataDir <- "~/Desktop/EPA/SPod_Data/TestRange_Dec2017"
datafiles <- dir(dataDir, pattern=".csv", full.names=TRUE)
spod <- read.csv(datafiles[3], header=TRUE,  na.strings = "N/A")
spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
                                 format= "%m/%d/%Y %H:%M:%S"))

node <- "h"
pidCol <- paste(node, "SPOD.PID..V.", sep=".")
spodNode <- spod[, c("time", pidCol)]
names(spodNode)[2] <- c("pid")
spodNode <- subset(spodNode, !is.na(pid))
spodNode$pid <- as.numeric(scale(spodNode$pid, center = TRUE))
```

## Gurobi vs Rglpk Timing Experiments

```{r, cache = TRUE}
tau <- 0.05
k <- 3

m_gurobi <- microbenchmark(
   gurobi_trend(spodNode$pid[1:200], tau, 200, k),
   gurobi_trend(spodNode$pid[1:400], tau, 400, k),
   gurobi_trend(spodNode$pid[1:1000], tau, 1000, k),
   gurobi_trend(spodNode$pid[1:2000], tau, 2000, k),
   times = 10
)

m_Rglpk <- microbenchmark(
  lpglpk_trendfilter(spodNode$pid[1:200], tau, 200, k),
  lpglpk_trendfilter(spodNode$pid[1:400], tau, 400, k),
  lpglpk_trendfilter(spodNode$pid[1:1000], tau, 1000, k),
  lpglpk_trendfilter(spodNode$pid[1:2000], tau, 2000, k),
  times = 10
)

knitr::kable(data.frame(n = c(200, 400, 1000, 2000), gurobi = summary(m_gurobi)$mean, Rglpk = summary(m_Rglpk)$mean), digits = 0)
```

### Approximation to objective function
Trend filtering objective:
min $\sum_i\rho_{\tau}(y_i - \theta_i) + ||D\theta||_1$

Approximate objective: 
Define $\tilde{f}(x, \tau) = \sum_i \frac{1}{2*100\pi}log[cosh(100\pi x_i)] + (\tau - \frac{1}{2}) x_i$

min $\tilde{f}(y - \theta, \tau) + \tilde{f}(D\theta, 0.5)$

Both gradient and Hessian are easy to compute. Hessian is banded and sparse.

```{r, cache = TRUE}
lambda <- 400/2
y <- spodNode$pid[1:400]
theta_g <- gurobi_trend(y, tau, 400, k)
D <- get_Dk(length(y), k)

phi0 <- y - quantile(y, tau)

m2 <- microbenchmark(
lbfgs_sol <- lbfgs(obj, grad_obj, vars = phi0,  y=y, tau=tau, D=D, 
                lambda = lambda, invisible = 1),
trust_sol <- trust.optim(phi0, obj, grad_obj, hess_obj, 
                      method = "Sparse", 
                      control = list(report.level = 0), 
                      y=y, tau=tau, 
                      D=D, lambda=lambda),
times = 1
)

theta_lbfgs <- y-lbfgs_sol$par
theta_trust <- y-trust_sol$solution

knitr::kable(data.frame(method = c("LBFGS", "Trust Region"), 
                        time = summary(m2)$mean), digits = 1)
```

```{r}
plot(y, type="l", col="grey")
lines(theta_g, col="red")
lines(theta_lbfgs, col="blue")
lines(theta_trust, col = "darkgreen")
```

## Cross validation to choose lambda - simulated data

```{r, cache = TRUE, echo = TRUE, message = FALSE, results=FALSE}
n <- 400
x <- seq(1, n, 1)
y <- sin(x*2*pi/n) + rnorm(n, 0, .5)

k <- 3
tau <- c(0.05, 0.5)

lambdaSeq <- seq(50, 1600, 50)
loss_cv <- lambda_cv(y, tau, k, lambdaSeq,
                     numFolds = 5, parallel = TRUE)
plot(rowSums(loss_cv$loss)~lambdaSeq, 
     ylab = "Check loss of validation data", 
     xlab = "Lambda")

theta_df <- get_windows(y, x, k, tau, loss_cv$lambda, n, 0)
plot.df <- left_join(theta_df, 
                     data.frame(time=x, y=y, 
                                true = sin(x*2*pi/n))) 
```

```{r, echo = FALSE}
# Chosen lambda
loss_cv$lambda

ggplot(plot.df, aes(x = time, y = y)) +
  geom_line(alpha = 0.2) +
  geom_line(aes(y=true), col="blue")+
  geom_line(aes(y=theta, col = factor(window), linetype=tau)) +
  theme_bw() + 
  guides(col = "none")

```

## Cross-validation on real data

```{r, cache = TRUE}
k <- 3
tau <- c(0.05, 0.5)
df <- spodNode
y <- df$pid[1:3600]
n <- length(y)
lambdaSeq <- c(5*n, 10*n, 11*n, 12*n, 13*n, 14*n, 15*n)

spod_cv <- lambda_cv(y, tau, k, lambdaSeq,
          numFolds = 5, parallel = TRUE)
plot(rowSums(spod_cv$loss)~lambdaSeq, 
     ylab = "Check loss of validation data", 
     xlab = "Lambda")
spod_cv$lambda
```

```{r, include=FALSE}
theta_df <- get_windows(y, df$time[1:n], k, tau, loss_cv$lambda, n, 0)
plot.df <- left_join(theta_df, spodNode) 
```

```{r, echo = FALSE, message = FALSE}
ggplot(plot.df, aes(x = time, y = pid)) +
  geom_line(alpha = 0.2) +
  geom_line(aes(y=theta, col = factor(window), linetype=tau)) +
  theme_bw() + 
  guides(col = "none")
```

# Complete day using hour long windows 
```{r, eval= TRUE, cache = TRUE, include = FALSE}
theta_df <- get_windows(df$pid, df$time, k, tau, spod_cv$lambda, 
                        length(y), 600)
plot.df <- left_join(theta_df, spodNode) 
```

```{r, message = FALSE}
ggplot(plot.df, aes(x = time, y = pid)) +
    geom_line(alpha = 0.2) +
    geom_line(aes(y=theta, col = factor(window), linetype=tau)) +
    theme_bw()
```

```{r, message=FALSE}
ggplot(plot.df, aes(x = time, y = pid)) +
    geom_line(alpha = 0.2) +
    geom_line(aes(y=theta, col = factor(window), linetype=tau)) +
    xlim(c(ymd_hms("2017-11-30 9:00:00", tz = "EST"), 
           ymd_hms("2017-11-30 11:30:00", tz = "EST")))+
    theme_bw()
```

# Compare all 3 nodes

```{r, message = FALSE}
spod_pid <- spod %>%
  select(time, contains("SPOD.PID..V.")) %>%
  mutate_at(vars(-time), function(x) as.numeric(scale(x, center=TRUE))) %>%
  gather("node", "pid", -time) %>%
  mutate(node_name = substr(node, 1, 1))

ggplot(spod_pid, aes(x=time, y=pid, col=node_name)) + 
  geom_line(alpha = 0.2) + 
  scale_color_brewer(palette = "Set1") + 
  theme_bw()
```

## Use selected lambda to detrend

```{r, cache = TRUE, message = FALSE}
nodes <- c("f", "g", "h")
plot_df <- data.frame(time = spodNode$time)
tau <- 0.05
for (node in nodes){
  pidCol <- paste(node, "SPOD.PID..V.", sep=".")
  spodNode <- spod[, c("time", pidCol)]
  names(spodNode)[2] <- c("pid")
  spodNode$pid <- as.numeric(scale(spodNode$pid, center = TRUE))
  theta_df <- get_windows(spodNode$pid, spodNode$time, k, tau,
                          43200, 3600, 600)
  theta_comb <- theta_df %>% 
    group_by(time) %>% 
    summarise(
      theta = mean(theta)
      )

  spodNode[,sprintf("pidC_%s", node)] <- spodNode$pid - theta_comb$theta
  plot_df <- left_join(plot_df, spodNode[,c(1,3)]) 
}

plot_long <- plot_df %>% gather("node", "pid", -time)

ggplot(plot_long, aes(x=time, y=pid, col=node)) + 
  geom_line(alpha = 0.2) + 
  scale_color_brewer(palette = "Set1") + 
  theme_bw()

ggplot(plot_long, aes(x=time, y=pid, col=node)) + 
  geom_line(alpha = 0.3) + 
  scale_color_brewer(palette = "Set1") + 
  xlim(c(ymd_hms("2017-11-30 9:30:00", tz = "EST"), 
         ymd_hms("2017-11-30 11:00:00", tz = "EST")))+
  theme_bw()
```

### 2018-10-26 Code/detrendr/R/get_windows.R

@@ -1,0 +1,47 @@
+#' Estimate trend using rolling windows
+#'
+#' \code{get_windows} Estimates trend in rolling windows
+#'
+#' @param y observed data
+#' @param time timestamp for observations
+#' @param k order of differencing
+#' @param tau quantile levels at which to evaluate trend
+#' @param lambda smoothing penalty parameter
+#' @param window_size integer length of window
+#' @param overlap integer length of overlap between windows
+get_windows <- function(y, time, k, tau, lambda, 
+                        window_size, overlap){
+  y_n <- length(y)
+  window_end <- window_size
+  window_start <- 1
+  
+  theta_df <- {}
+  i <- 1
+  pb <- txtProgressBar(min = 0, max = round(y_n/(window_size-overlap)),
+                       style = 3)
+  
+  while(window_end < y_n){
+    y0 <- y[window_start:window_end]
+    theta_gurobi <- as.data.frame(gurobi_trend(y0, tau, lambda, k))
+    names(theta_gurobi) <- paste0("tau_", tau)
+    theta_gurobi$time <- time[window_start:window_end]
+    theta <- theta_gurobi %>% gather(tau, theta, -time)
+    theta$window <- i
+    theta_df <- bind_rows(theta_df, theta)
+    window_start <- window_start + window_size - overlap
+    window_end <- window_start + window_size - 1
+    i <- i+1
+    setTxtProgressBar(pb, i)
+  }
+  
+  window_end <- y_n
+  y0 <- y[window_start:window_end]
+  theta_gurobi <- as.data.frame(gurobi_trend(y0, tau, lambda, k))
+  names(theta_gurobi) <- paste0("tau_", tau)
+  theta_gurobi$time <- time[window_start:window_end]
+  theta <- theta_gurobi %>% gather(tau, theta, -time)
+  theta$window <- i
+  theta_df <- bind_rows(theta_df, theta)
+  
+  return(theta_df)
+}

### 2018-10-26 Code/detrendr/R/gurobi_trend.R

@@ -7,6 +7,12 @@
 #' @param lambda penalty paramter controlling smoothness
 #' @param k order of differencing
 gurobi_trend <- function(y, tau, lambda, k){
+
+  if(tau >= 1 || tau <= 0){
+    stop("tau must be between 0 and 1.")
+  }
+
+  tau <- sort(tau)
   D <- as.matrix(get_Dk(length(y), k))
   n <- length(y)
   m <- nrow(D)
@@ -56,8 +62,6 @@ gurobi_trend <- function(y, tau, lambda, k){
   # Type of variables (binary)
   params <- list(OutputFlag=0)
   result <- gurobi(model, params)
-  print('Solution:')
-  print(result$objval)
 
   theta <- matrix(0, nrow=n, ncol = nT)
   for (i in 1:nT){
@@ -66,3 +70,6 @@ gurobi_trend <- function(y, tau, lambda, k){
   }
   return(theta)
 }
+
+
+

### 2018-10-26 Code/detrendr/R/smooth_relaxation.R

@@ -1,0 +1,36 @@
+
+approx_checkLoss <- function(x, tau){
+  x <- as.numeric(x)
+  f <- (log(cosh(x*1e4*pi))/(2*1e4*pi) + (tau - .5)*x)
+  f[is.infinite(f) & x > 0] <- tau*x[is.infinite(f) & x > 0]
+  f[is.infinite(f) & x < 0] <- (tau-1)*x[is.infinite(f) & x < 0]
+  return(f)
+}
+
+hess_checkLoss <- function(x){
+  x <- as.numeric(x)
+  return(1e4*pi*cosh(1e4*pi*x)^(-2)/2)
+}
+
+grad_checkLoss <- function(x, tau){
+  x <- as.numeric(x)
+  return(tanh(x*1e4*pi)/2 - 1/2 + tau)
+}
+
+obj <- function(phi, y, tau, D, lambda){
+  return(
+    sum(approx_checkLoss(phi, tau),
+    lambda*approx_checkLoss(D%*%(y-phi), 0.5))
+    )
+}
+
+grad_obj <- function(phi, y, tau, D, lambda){
+  as.numeric(grad_checkLoss(phi, tau) -
+    lambda*Matrix::crossprod(D,grad_checkLoss(D%*%(y-phi), 0.5)))
+}
+
+hess_obj <- function(phi, y, tau, D, lambda){
+  diag(hess_checkLoss(phi)) +
+    lambda*Matrix::crossprod(D, diag(hess_checkLoss(D%*%(y-phi))))%*%D
+}
+

### 2018-10-26 Notes/Gurobi_notes.Rmd
Meeting notes updated

@@
 Approximate objective: 
-Define $\tilde{f}(x, \tau) = \sum_i \frac{1}{2*100\pi}log[cosh(100\pi x_i)] + (\tau - \frac{1}{2}) x_i$
+Define $\hat{f}(x, \tau) = \sum_i \frac{1}{2*10^4\pi}log[cosh(10^4\pi x_i)] + (\tau - \frac{1}{2}) x_i$
 
-min $\tilde{f}(y - \theta, \tau) + \tilde{f}(D\theta, 0.5)$
+min $\hat{f}(y - \theta, \tau) + \hat{f}(D\theta, 0.5)$
 
@@
 
+## Compare times on n=200
 ```{r, cache = TRUE}
-lambda <- 400/2
-y <- spodNode$pid[1:400]
-theta_g <- gurobi_trend(y, tau, 400, k)
+lambda <- 200/2
+y <- spodNode$pid[1:200]
+theta_g <- gurobi_trend(y, tau, 200, k)
 D <- get_Dk(length(y), k)
@@
 m2 <- microbenchmark(
+  theta_g <- gurobi_trend(y, tau, 200, k),
   lbfgs_sol <- lbfgs(obj, grad_obj, vars = phi0,  y=y, tau=tau, D=D, 
@@
                         D=D, lambda=lambda),
-times = 1
+  times = 5
 )
@@
 
-knitr::kable(data.frame(method = c("LBFGS", "Trust Region"), 
+knitr::kable(data.frame(method = c("Gurobi", "LBFGS", "Trust Region"), 
                         time = summary(m2)$mean), digits = 1)
 ```
+## Compare results
 
@@
 n <- length(y)
-lambdaSeq <- c(5*n, 10*n, 11*n, 12*n, 13*n, 14*n, 15*n)
+lambdaSeq <- c(10*n, 12*n, 14*n, 16*n)
 
@@
   theta_df <- get_windows(spodNode$pid, spodNode$time, k, tau,
-                          43200, 3600, 600)
+                          spod_cv$lambda, 3600, 600)
   theta_comb <- theta_df %>% 

### 2018-11-05 Notes/consensus_notes.Rmd
ADMM working

---
title: "detrendr Meeting notes"
output: html_document
---

## Consensus ADMM algorithm
minimize $\sum f_i(x_i)$
subject to $x_i - z = 0$

$x_i^{k+1} = \arg\min_{x_i}\left(f_i(x_i) + w_i^{kT}(x_i - \bar{x}^k) + \frac{\rho}{2}||x_i - \bar{x}^k||_2^2\right)$

$w_i^{k+1} = w_i^k + \rho(x_i^{k+1}-\bar{x}^{k+1})$

Use $f_i(\theta_i) = \rho_\tau(y_i-\theta_i) + ||D\theta_i||_1$ where $y_i$ is the window.

![Windows](../Code/windows_fit.png)

![Consensus](../Code/consensus_fig.png)

Single window size 1000, quadratic program takes 1427 ms and linear program takes 621 ms.

## Paper innovations

* Trend filtering - compared to smoothing splines, splines with fixed knots, splines with adaptive knots
* Quantile trend over time
* Cross validation for smoothing parameter
* Parallelizable algorithm for smooth across smaller windows
* Non-crossing quantiles for ensuring valid estimates
* Application to low cost air quality sensors

### 2018-11-05 Code/detrendr/R/consensus_ADMM.R

@@ -1,130 +1,3 @@
-
-#' Consensus ADMM for overlapping windows
-#'
-#' \code{consensus ADMM} Uses ADMM to smooth overlapping windows
-#'
-#' @param y observed data
-#' @param tau quantile levels at which to evaluate trend
-#' @param lambda smoothing penalty parameter
-#' @param k order of differencing
-#' @param rho step size for ADMM
-#' @param overlap integer length of overlap between windows
-#' @param max_iter Maximum number of iterations
-#' @examples
-#' n <- 100
-#' x <- seq(1, n, 1)
-#' y <- sin(x*2*pi/n) + rnorm(n, 0, .4)
-#' lambda <- 10
-#' k <- 3
-#' tau <- c(0.05)
-#' overlap <- 20
-#' rho <- 1
-#' result <- consensus_ADMM(y, tau, lambda, k, rho, overlap, 300)
-#' y_n <- length(y)
-#' window_size <- floor((y_n+overlap)/2)
-#' plot(result$phiBar~x, type="l")
-#' points(result$phi1~x[1:window_size])
-#' points(result$phi2~x[(y_n-window_size+1):y_n], col="blue")
-#' plot(y~x)
-#' lines(result$theta)
-#' plot(result$primal_norm)
-#' plot(result$dual_norm)
-consensus_ADMM2 <- function(y, tau, lambda, k, rho, overlap, max_iter, eps = 0.01){
-  y_n <- length(y)
-  nT <- length(tau)
-  if( (y_n + overlap) %% 2 != 0 ){
-    stop("length of y plus overlap must be even")
-  }
-
-  window_size <- (y_n+overlap)/2
-
-  Ind1 <- (window_size-overlap+1):(window_size)
-  Ind2 <- 1:overlap
-
-  y1 <- y[1:window_size]
-  y2 <- y[(y_n-window_size+1):y_n]
-
-  # Initial values
-  w1 <- numeric(length(y1)*length(tau))
-  w2 <- numeric(length(y2)*length(tau))
-  phiBar <- numeric(length(y1)*length(tau))
-
-  # x-update
-  phi1 <- quad_update(y1, tau, lambda, k, w1, phiBar, 0, first = TRUE)
-  phi2 <- quad_update(y2, tau, lambda, k, w2, phiBar, 0, first = FALSE)
-
-  # Average update
-  phiBar <- matrix(NA, nrow = y_n, ncol = length(tau))
-  phiBar[1:window_size,] <- phi1
[…]

### 2018-11-05 Code/parallel_experiments.R

@@ -1,0 +1,46 @@
+phiBar_list <- list()
+for (i in 1:n_windows){
+  phiBar_list[[i]] <- as.numeric(phiBar[windows[,i],])
+}
+
+# Window Update
+update_loop <- function(y_list, w_list, phiBar_list, tau, lambda, k, rho){
+  phi_list <- vector("list",length(y_list))
+  
+  for (i in 1:n_windows){
+    phi_list[[i]] <- quad_update(y_list[[i]], tau, lambda, k, w_list[[i]],
+                                 phiBar_list[[i]], rho)
+  }
+  return(phi_list)
+}
+
+update_mapply <- function(y_list, w_list, phiBar_list, tau, lambda, k, rho){
+  
+  phi_list <- mapply(quad_update, y_list, w_list, phiBar_list,
+                     MoreArgs = list(tau=tau, lambda=lambda, k=k,
+                                     rho = rho),
+                     SIMPLIFY = FALSE)
+  return(phi_list)
+}
+
+library(parallel)
+update_mcapply <- function(y_list, w_list, phiBar_list, tau, lambda, k, rho){
+  
+  phi_list <- mcmapply(quad_update, y_list, w_list, phiBar_list,
+                       MoreArgs = list(tau=tau, lambda=lambda, k=k,
+                                       rho = rho),
+                       SIMPLIFY = FALSE)
+  return(phi_list)
+}
+
+
+
+microbenchmark(
+  #update_loop(y_list, w_list, phiBar_list, tau, lambda, k, rho),
+  update_mapply(y_list, w_list, phiBar_list, tau, lambda, k, rho),
+  update_mcapply(y_list, w_list, phiBar_list, tau, lambda, k, rho),
+  update_foreach(y_list, w_list, phiBar_list, tau, lambda, k, rho),
+  times = 5
+)
+
+stopCluster(cl)

### 2018-11-13 Manuscript/detrendify_JCGS.tex
Added code from Oh and Nychka, updated references, worked on simulations

@@
 {
-  \title{\bf Quantile Trend Filtering via Spingarn's Method}
+  \title{\bf Drift Removal for Time Series Data Using Quantile Trend Filtering}
   \author{Halley Brantley\thanks{
@@
 \item \cite{Ning2014} address problem of estimating a smooth baseline in noisy data with drift.
+\item \cite{Takeuchi2006nonparametric} Nonparametric quantile regression using SVM with Gaussian RBF kernels and check (pinball) loss. 
+\item \cite{yuan2006gacv} Comparison of cross-validation methods for quantile smoothing splines.  
 \end{itemize}

### 2018-11-13 Code/detrendr/R/lambda_cv.R

@@ -12,17 +12,15 @@
 lambda_cv <- function(y, tau, k,
                       lambdaSeq = c(length(y)/10, length(y), length(y)*10),
                       numFolds = 5,
-                      parallel = TRUE){
+                      parallel = TRUE, 
+                      cl = NULL){
   
   tau <- sort(tau)
   folds <- sample(1:numFolds, length(y), replace = TRUE)
-  no_cores <- detectCores() - 1
-  
+
   if (parallel){
-    cl <- makeCluster(no_cores, type="FORK")
     loss <- parLapply(cl, lambdaSeq, valid_checkLoss, folds,
                       y=y, tau=tau, k=k)
-    stopCluster(cl)
   } else {
     loss <- lapply(lambdaSeq, valid_checkLoss, folds,
                    y=y, tau=tau, k=k)

### 2018-11-13 Code/ucgs_a_10712118_sm0001/Rcodes/run_jcgs.R

@@ -1,0 +1,83 @@
+library(fields)
+library(quantreg)
+library(stats)
+library(wavethresh)
+source("function_jcgs.R")
+# it is for 1-d simulation (asymmetric noise case) in Section 4.1
+n <- 2000
+i <- c(1:n)
+x <- i/n
+ss <- 1
+set.seed(167)
+noise <- rnorm(n,0,0.07)
+ss <- 1
+noise<-rgamma(n,3,1)
+y <- sin(10*x)+noise
+data <- cbind(x,y)
+rx <- sort(data[,1])
+ry <- data[,2][order(data[,1])]
+k <- 100
+
+tau <- 0.10 #(0.5,0.9)
+finit <- as.vector(int.fit.fig(k,n,tau,ry))
+fit1.two <- aresult(x,(2000-100+1),0.04,hry1(100,2000,x),hry2(100,tau,2000,y)) #Yu's method
+fit1.pro2 <- qreq1d.sreg.fig(x,y,fit1.two,tau,0.01) # the proposed method
+temp <- rqss(y~qss(x,constraint="N"),tau=tau,lambda=0.1) 
+fit1.rqss <- temp$coef[1]+temp$coef[-1] # Koenker's method 
+tr1 <- sin(10*x)+qgamma(tau,3,1)
+
+plot(x,tr1,type="l")
+lines(x,fit1.two,col=2)
+lines(x,fit1.pro2,col=3)
+lines(x[-1],fit1.rqss,col=4)
+
+mean((tr1-fit1.pro2)^2)
+mean((tr1-fit1.two)^2)
+mean((tr1[-1]-fit1.rqss)^2)
+
+# it is for 2-d simulation in Section 5.1
+n <- 10
+i <- c(1:n)
+x <- i/n
+y <- x
+z <- sin(3*pi*x)%*%t(cos(pi*y))
+m <- ncol(z)
+noise <- matrix(rnorm(n*m),ncol=m) # for normal noise
+z.noise <- z+noise/3
+
+tau <- 0.5
+tr <- z+matrix(rep(qnorm(tau),n*m),ncol=m)/3 # true quantile with normal noise
+
+k <- 3
+zinit <- d2int.fit.sy(x,y,z.noise,k,tau) # it is for a initial fit
+grid <- NULL
+for(i in 1:n){
+    tmp <- cbind(rep(x[i],length(y)),y)
+    grid <- rbind(grid,tmp)
+}
[…]

### 2018-11-15 Manuscript/detrendify_JCGS.tex
wrote a lot of ADMM and CV notes

@@
 
+\section{ADMM for Smoothing Windows}
+
+To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
+We propose an alternating direction method of multipliers (ADMM) algorithm for solving large problems in a piecewise fashion.
+The ADMM algorithm, described by \cite{boyd2011distributed}, relies on the idea of dual ascent.
+If we consider the equality optimization problem
+\begin{align}
+&\text{minimize } f(x)\\
+&\text{subject to } Ax = b
+\end{align}
+with variable $x\in \mathbf{R}^n$, where $A \in \mathbf{R}^{mxn}$ and $f: \mathbf{R}^n \rightarrow \mathbf{R}$ is convex. The corresponding Lagrangian function is 
+\begin{equation}
+L(x,\omega) = f(x) + \omega^T(Ax-b)
+\end{equation}
+and the corresponding dual problem  is 
+\begin{equation}
+ \text{maximize } \inf_x L(x,\omega)
+\end{equation}
+Assuming that strong duality holds (Slater's constraint: there exists a solution in the interior of the domain), the optimal values of the primal and dual variables are the same.
+The dual ascent method, uses gradient ascent to solve the dual problem and thus the primal problem using the iterates
+\begin{align}
+&x^{k+1} = \arg\min L(x, \omega^k)\\
+&\omega^{k+1} = \omega^k + \alpha^k(Ax^{k+1} - b)
+\end{align}
+where $\alpha^k > 0$ is a step size and $k$ is the iteration counter.
+In many cases (including our problem), the $x$ update fails because the problem is unbounded below.
+To remedy this issue the augmented Lagrangian was introduced:
+\begin{equation}
+L_{\gamma}(x,\omega) = f(x)+\omega^T(Ax-b) + \frac{\gamma}{2}||Ax-b||_2^2
+\end{equation}
+where $\gamma > 0$ is called the penalty parameter. Applying dual ascent to the modified problem results in the method of multipliers algorithm which converges under more general conditions:
+\begin{align}
+&x^{k+1} = \arg\min L_\gamma(x, \omega^k)\\
+&\omega^{k+1} = \omega^k + \gamma(Ax^{k+1} - b)
+\end{align}
+Consider a new optimization problem with a separable objective function. 
+\begin{align}
+&\text{minimize } f_1(x_1) + f_2(x_2)\\
+&\text{subject to } A_1x_1 + A_2x_2 = b
+\end{align}
+where $x = (x_1, x_2)$, $i \in \{1,2\}$, and  $x_i \in \mathbf{R}^{n_i}$. The dual ascent updates can be expressed as 
+\begin{align}
+&x_i^{k+1} = \arg\min L_i(x_i, \omega^k)\\
+&\omega^{k+1} = \omega^k + \alpha^k(Ax^{k+1} - b)
+\end{align}
+however we lose the separability if the augmented Lagrangian is used instead, i.e.
+the augmented Lagrangian can not be separated into functions of $x_i$.
+The Alternating Direction Method of Multipliers (ADMM) addresses this problem and maintains separability while using the augmented Lagrangian to improve convergence through the following updates
+\begin{align}
+&x_1^{k+1} = \arg\min_{x_1} L_{\gamma}(x_1, x_2^k, \omega^k)\\
+&x_2^{k+1} = \arg\min_{x_2} L_{\gamma}(x_1^{k+1}, x_2, \omega^k)\\
+&\omega^{k+1} = \omega^k + \alpha^k(A_1x_1^{k+1} + A_2x_2^{k+1} - b)
+\end{align}
+
+We now consider the quantile regression trend filtering problem,  
+\begin{equation}
+\text{minimize } \rho_\tau(y-\theta) + \lambda||D^{(k)}\theta||_1\\
+\end{equation}
+where $y,\theta \in \mathbf{R}^n$, $y$ is the observed data, $\rho_{\tau}(u) = \sum_i(\tau - I(u_i < 0))u_i$ is the check loss function corresponding to quantile level $\tau$ and $D^{(k)}$ is the discrete differencing matrix of order $k$ and $\lambda$ is a tuning parameter that controls smoothness.
+We first re-parameterize $\phi = y-\theta$ so the problem is
+\begin{equation}
+\text{minimize } \rho_\tau(\phi) + \lambda||D^{(k)}(y-\phi)||_1\\
+\end{equation}
+We further divide $\phi$ order to solve smaller problems: Defining
+\begin{align}
+&\phi_1 = (\phi_{11}, \phi_{12})\\
+&\phi_2 = (\phi_{21}, \phi_{22}, \phi_{23})\\
+&\phi_3 = (\phi_{31}, \phi_{32})\\
+&\phi = (\phi_{11}, \phi_{12}=\phi_{21}, \phi_{22}, \phi_{23}=\phi_{31}, \phi_{32}) \\
+\end{align}
+Dividing $y$ similarly, the problem then becomes 
+\begin{align}
+&\text{minimize } \sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1\\
+&\text{subject to: } \phi_{12}=\phi_{21}, ~~ \phi_{23}=\phi_{31}\\
+\end{align}
+We can further simplify by defining 
+\begin{align}
+&\overline{\phi} = (\phi_{11}, \frac{\phi_{12}+\phi_{21}}{2}, \phi_{22}, \frac{\phi_{23}+\phi_{31}}{2}, \phi_{32}) \\
+&\overline{\phi_1} = (\phi_{11}, \frac{\phi_{12}+\phi_{21}}{2})\\
+&\overline{\phi_2} = ( \frac{\phi_{12}+\phi_{21}}{2}, \phi_{22}, \frac{\phi_{23}+\phi_{31}}{2})\\
+&\overline{\phi_3} = (\frac{\phi_{23}+\phi_{31}}{2}, \phi_{32})
+\end{align}
+so the problem becomes
+\begin{align}
+&\text{minimize } \sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1\\
+&\text{subject to: } \phi_{i}=\overline{\phi_i}\\
+\end{align}
+The augmented Lagrangian for this problem is 
+\begin{align}
+&L_\gamma(\phi_1,\phi_2, \phi_3, \overline{\phi_1}, \overline{\phi_2}, \overline{\phi_3}, \omega) = \\
+&\sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1 + \omega_i^T(\phi_i - \overline{\phi_i}) + \frac{\gamma}{2}||\phi_i - \overline{\phi_i}||_2^2
+\end{align}
+The ADMM updates are then given by
+\begin{align}
+&\phi_i^{k+1} = \arg\min_{\phi_i}\rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1 + \omega_i^{kT}(\phi_i - \overline{\phi_i}^k) + \frac{\gamma}{2}||\phi_i - \overline{\phi_i}^k||_2^2\\
+&\omega_i^{k+1} = \omega_i^{k} + \gamma(\phi_i^{k+1} - \overline{\phi_i}^{k+1})
+\end{align}
+The $\phi_i$ updates can be obtained using a quadratic program solver such as Gurobi and can be obtained in parallel.
+
+\section {Cross Validation Methods}
+Our method can easily handle missing data by changing the check loss function to output 0 for missing values.
+This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data.
+A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.
+The traditional smoothing spline for $y \in \mathbf{R}^n$ as a function of a predictor $x \in \mathbf{R}^n$ using the squared loss is estimated by minimizing
+
+\begin{equation}
+\frac{1}{n}\sum_i (y_i - f(x_i))^2 + \lambda \int(f''(x))^2 dx
+\end{equation}
+
+It has been shown that the solution to this problem is a cubic spline with a knot at each of the observed $x_i$.
+This problem can then be reformulated using a spline basis matrix $B \in \mathbf{R}^{nxn}$:
+\begin{equation}
+(y-B\beta)^T(y-B\beta) + n\lambda\beta^T\Omega\beta
+\end{equation}  
+where $\Omega_{jk} = \int B_j''(x)B_k''(x)dx$.
+Since each spline basis function is a defined as a cubic polynomial with non-zero second derivatives on a finite region, the integrals are finite.
+The solution is then given by
+$$f(x) = B\widehat{\beta} = B(B^TB + n\lambda\Omega)^{-1}B^Ty$$
+Because the solution is a linear combination of the observed $y$, the leave-one-out cross validation (LOOCV) statistics can be calculated without refitting the model.
+Defining $W = B\widehat{\beta} = B(B^TB + n\lambda\Omega)^{-1}B$ the LOOCV residual is given by
+\begin{equation}
+y_i - \widehat{y_i}^{(-i)} = \frac{y_i - \widehat{y_i}}{1-W_{[i,i]}}
+\end{equation}
+where $\widehat{y_i}$ is the prediction using the full dataset and $\widehat{y_i}^{(-i)}$ is the prediction from the model fit without $y_i$.
+In practice, generalized cross validation (GCV) has also been found to be an effective metric for choosing the smoothing parameter, with the validation MSE calculated as
+\begin{equation}
+\frac{\frac{1}{n}\sum(y_i-\widehat{y_i})^2}{(1-\frac{1}{n}tr(W))^2}
+\end{equation}
+
+Returning to the quantile regression case, \cite{nychka1995nonparametric} proposed a pseudo-data algorithm for a quantile spline estimator of the form 
+\begin{equation}
+ \sum_i \rho_{\tau}(y_i - g(x_i)) + \lambda\int(g''(x))^2dx.
+\end{equation}
+
+If $\rho_{\tau}(\cdot)$ were differentiable, the solution to this equation would take a form similar to that of the squared loss smoothing spline with weights equal to $\frac{\rho'_{\tau}(y_i-g(x_i))}{2(y_i - g(x_i))}$.
+Relying on this idea, Nychka proposed to solve the problem by iteratively solving the weighted smoothing spline.
+To address the non-differentiability they propose an approximation
+\begin{equation}
+\rho_{\tau, \delta}(u) = [\tau I(u>0) + (1-\alpha)I(u<0)]u^2/\delta
+\end{equation}
+They propose a corresponding approximate cross validation metric for choosing the tuning parameter:
+\begin{equation}
+ACV(\lambda) = \frac{1}{n}\sum_i \rho_{\tau}\left(\frac{y_i - \widehat{g}_\lambda(x_i)}{1-h_{ii}}\right)
+\end{equation}
+where $h_{ii} = \frac{\partial\widehat{g}_{\tau, \lambda}(x_i)}{\partial y_i}$.
+However \cite{yuan2006gacv} showed through simulation that this is not a good approximation to the robust cross validation score
+\begin{equation}
+RCV(\lambda) = \frac{1}{n}\sum\rho_{\tau}(y_i - g_\lambda^{(-i)}(x_i))
+\end{equation}
+and propose the GACV as an alternative 
+\begin{equation}
+GACV(\lambda) = \sum_i \rho_{\tau}\left(\frac{y_i - \widehat{g}_\lambda(x_i)}{n-tr(H)}\right)
+\end{equation}
+where $H_{ij} = \frac{\partial\widehat{g}_{\tau, \lambda}(x_i)}{\partial y_j}$
+
+\cite{KoenkerNgPortnoy1994} proposed a quantile smoothing spline using the check loss function and 1-norm rather than 2-norm:
+\begin{equation}
+R_{\tau,\lambda} = \rho_{\tau}(y - g(x)) + \lambda\int_0^1|g''(x)|dx
+\end{equation}
+with $0 = x_0 < ...
+< x_n = 1$.
+They prove that the resulting function $\widehat{g(x)}$ is a linear spline with knots at the observed values of $x$.
+They also argue that the solutions $\widehat{g}_{\tau,\lambda}(\cdot)$ are piecewise constant in $\lambda$ and that there exists a mesh $0= \lambda_0 < \lambda_1 < ...
+< \lambda_J$ such that $\widehat{g}_{\tau,\lambda}(\cdot)$ is the solution for all $\lambda \in [\lambda_{i-1}, \lambda_i]$.
+Relating $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
+\begin{equation}
+SIC(p_{\lambda}) = log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}log n
+\end{equation}
+
 \bibliographystyle{asa}

### 2018-11-15 Code/simulations.R

@@ -4,7 +4,7 @@ library(quantreg)
 library(devtools)
 library(qrsvm)
 load_all("detrendr")
-source("function_jcgs.R")
+#source("function_jcgs.R")
 rm(list=ls())
 
 x <- (seq(1,200,1)-1)/200
@@ -13,6 +13,28 @@ y <- sin(10*x) + (x + .25)*rnorm(200, 0, .07)/.1
 
 fit_svm <- qrsvm(as.matrix(x), y, tau = .1, sigma = 100)
 lines(fit_svm$fitted~x, col="blue")
+
+tau <- .1
+fit_qsreg <- qsreg(x, y, alpha=tau, lam=1)
+f <- predict(fit_qsreg)
+resid <- y-f
+w <- rep(tau, length(resid))
+w[resid < 0] <- tau-1
+w <- w/(2*resid)
+w <- w/sum(w)
+
+fit_sq <- sreg(x, y, weights = w)
+fit_sq2 <- smooth.spline(x,y, w)
+A <- fit_sq$diagA
+f2 <- predict(fit_sq)
+resid <- y-f2
+
+LOOCV_mse <- mean(resid^2/(1-A)^2)
+# GCV MSE with weights
+GCV_mse <- crossprod(resid, diag(w))%*%resid/(1-mean(A))^2/n
+
+summary(fit_sq)
+
 
 # Fit Koenker Smoothing Splines
 rmse_qss <- function(x, y, tau, lambda, f_true){

### 2018-11-19 Code/detrendr/R/lambda_SIC.R
Description of ADMM algorthim for windows, and smoothing splines, function for lambda selection using SIC and validation dataset.

@@ -1,0 +1,34 @@
+#' Choose smoothing parameter using SIC
+#'
+#' \code{lambda_SIC} Selects smoothing parameter using Schwarz Information
+#'
+#' @param y observed data
+#' @param tau quantile levels at which to evaluate trend
+#' @param k order of differencing
+#' @param lambdaSeq smoothing penalty parameter options to compare
+#' @export
+lambda_SIC <- function(y, tau, k,
+                       lambdaSeq = seq(length(y)/10, length(y)*10, length(y)/10), 
+                       df_tol = 1e-9){
+  
+  df_trend <- c()
+  SIC_trend <- c()
+  
+  for (i in 1:length(lambdaSeq)){
+    lam <- lambdaSeq[i]
+    f_trend <- gurobi_trend(y, tau, lam, k)
+    resid_trend <- checkloss(y-f_trend, tau)
+    df_trend[i] <- sum(resid_trend<df_tol)
+    SIC_trend[i] <- log(mean(resid_trend)) + log(n)*df_trend[i]/(2*n)
+  }
+  
+  plot(SIC_trend~lambdaSeq, type="l", col="red")
+  abline(v=lambdaSeq[which.min(SIC_trend)], col="red")
+  abline(h=min(SIC_trend), col="red")
+  lam <- lambdaSeq[which.min(SIC_trend)]
+
+  
+  return(list(lambda = lam, 
+              SIC = SIC_trend, 
+              df = df_trend))
+}

### 2018-11-19 Code/simulations.R

@@ -7,14 +7,42 @@ load_all("detrendr")
 #source("function_jcgs.R")
 rm(list=ls())
 
-x <- (seq(1,200,1)-1)/200
-n <- length(x)
-y <- sin(10*x) + (x + .25)*rnorm(200, 0, .07)/.1
-
-fit_svm <- qrsvm(as.matrix(x), y, tau = .1, sigma = 100)
-lines(fit_svm$fitted~x, col="blue")
-
+n <- 500
+x <- (seq(1,n,1)-1)
+y <- sin(10*x/n) + (x/n + .25)*rnorm(n, 0, .1)/.1
 tau <- .1
+f_true <- sin(10*x/n) + (x/n + .25)*qnorm(tau, 0, .1)/.1
+
+fit_qsreg <- qsreg(x, y, maxit.cv = 50, alpha=tau, hmin = -9)
+f_qsreg <- predict(fit_qsreg)
+
+lam_SIC <- lambda_SIC(y, tau, 3)
+lam_SIC2 <- lambda_SIC(y, tau, 2)
+lam_valid <- lambda_valid(y, tau, 3, 5)
+
+fit <- rqss(y ~ qss(x, lambda = 2*lam_SIC2$lambda), tau = tau)
+fhat <- predict(fit, data.frame(x=x))
+f_trend <- gurobi_trend(y, tau, lam_SIC2$lambda, k=2)
+f_trend_SIC <- gurobi_trend(y, tau, lam_SIC$lambda, k=3)
+f_trend_valid <- gurobi_trend(y, tau, lam_valid$lambda, k=3)
+
+plot(f_true~x, type="l")
+lines(fhat~x, col="red")
+lines(f_trend_SIC~x, col="blue")
+lines(f_trend_valid~x, col="purple")
+lines(f_qsreg~x, col="darkgreen")
+
+mean((f_true-f_trend)^2)
+mean((f_true - f_trend_SIC)^2)
+mean((f_true - f_trend_valid)^2)
+mean((f_true - f_qsreg)^2)
+
+
+# Need to figure out how to choose sigma (kernel parameter)
+fit_svm <- qrsvm(as.matrix(x/n), y, tau = tau, sigma = 10)
+plot(fit_svm$fitted~x, col="cyan")
+
+
 fit_qsreg <- qsreg(x, y, alpha=tau, lam=1)
 f <- predict(fit_qsreg)
 resid <- y-f

### 2018-12-04 Manuscript/simulation_notes.tex
simulation notes and figures

\documentclass[]{article}
\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{placeins}
\usepackage{float}
\usepackage{caption}
\graphicspath{{../Figures/}}
\usepackage{gensymb}
\doublespacing
\usepackage{amsmath, amssymb} % Math packages
\usepackage{amsthm}
\usepackage{natbib}
\setlength{\parindent}{0cm}

\begin{document}

\section{Simulation Study}

We compare the performance of our quantile trend filtering model with the models of 3 others using designs proposed by \cite{Racine2017}. The models compared are

\begin{itemize}
	\item \texttt{npqw}: \cite{Racine2017} constrain the response to follow a smooth location scale model of the form $Y_i = a(X_i) + b(X_i)\epsilon_i$. They estimate the $\tau$\textsubscript{th} conditional quantile given $X_i = x$ using a kernel estimator
	\begin{equation}
		q_{\tau}(x) = \frac{\sum_{i=1}^n \Phi^{-1}_{(Y_i, b(X_i))}(\delta_0)K_h(X_i,x)}{\sum_{i=1}^nK_h(X_i, x)}
	\end{equation}
	defining $\Phi^{-1}_{(Y_i, b(X_i))}(\delta_0)$ as the quantile function of the Normal distribution with mean $Y_i$ and standard deviation $b(X_i)$ evaluated at $\tau$, $\delta_0$ is a function of $\tau$ and chosen empirically, $h$ is a tuning parameter and $K$ is a kernel function. Code was obtained from the author for the \texttt{quantile-ll} method.

	\item \texttt{qsreg}: \cite{Oh2011} proposed a pseudo-data algorithm for a quantile spline estimator of the form 
	\begin{equation}
	\sum_i \rho_{\tau}(y_i - g(x_i)) + \lambda\int(g''(x))^2dx.
	\end{equation} 
	If $\rho_{\tau}(\cdot)$ were differentiable, the solution to this equation would take a form similar to that of the squared loss smoothing spline with weights equal to $\frac{\rho'_{\tau}(y_i-g(x_i))}{2(y_i - g(x_i))}$. Relying on this idea, Nychka proposed to solve the problem by iteratively solving the weighted smoothing spline. To address the non-differentiability they propose an approximation 
	\begin{equation}
	\rho_{\tau, \delta}(u) = [\tau I(u>0) + (1-\alpha)I(u<0)]u^2/\delta
	\end{equation}
	The function \texttt{qsreg} in the texttt{fields} R package was used.

	\item \texttt{rqss}: \cite{KoenkerNgPortnoy1994} Koenker proposed smoothing splines using trend filtering with the second order differencing matrix which results in linear splines. The function \texttt{rqss} in the \texttt{quantreg} package implements this method. The smoothing parameter $\lambda$ is chosen using a grid search and minimizing 
	\begin{equation}
	SIC(p_{\lambda}) = log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}log n
	\end{equation}
	where $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots.

	\item \texttt{detrendr\_SIC}: Our method where we minimize
	$\sum_i\rho_{\tau}(y_i - \theta_i) + \lambda||D\theta||_1$ and $\lambda$ is chosen using SIC from above. 
	\item \texttt{detrendr\_valid}: Our method where lambda is chosen by leaving out every 5th observation as a validation data set and evaluating the check loss function on the validation data. 
\end{itemize}

Three simulation designs from \cite{Racine2017} were considered. For all designs $X_i$ was generated as a uniformly space sequence in $[0,1]$ and the mean function was given as 
$$Y_i = sin(2\pi X_i) + \epsilon_i(X_i)$$
The three error distributions considered were 
\begin{itemize}
 \item Gaussian: $\epsilon_i \sim N\left(0, \left(\frac{1+x_i^2}{4}\right)^2\right)$
 \item Beta: $\epsilon_i \sim Beta(1, 11-10x_i)$
\item Mixed normal: $\epsilon_i$ is simulated from a mixture of $N(-1,1)$ and  $N(1,1)$ with mixing probability $x_i$.
\end{itemize}
\begin{figure}
\caption{Simulated data with true quantiles $\tau \in \{0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99\}$}	
\includegraphics[width=.3\linewidth]{Figures/gaus.png}
\includegraphics[width=.3\linewidth]{Figures/shapebeta.png}
\includegraphics[width=.3\linewidth]{Figures/mixednorm.png}
\end{figure}

100 datasets were generated of sizes 300, 500 and 1000. The MSE was calculated as $\frac{1}{n}\sum_i (\hat{q}_{\tau}(x_i) - q_\tau(x_i))^2$.

\includegraphics[width=\linewidth]{Figures/gaus_mse.png}

\includegraphics[width=\linewidth]{Figures/shapebeta_mse.png}

\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
\bibliographystyle{asa}
\bibliography{detrendify}

\end{document}

### 2018-12-04 Code/simulation_figs.R

@@ -2,34 +2,45 @@ library(tidyverse)
 rm(list=ls())
 source("trueQuantile.R")
 
-tau <- c(0.05, 0.1, 0.5, .9, 0.95)
+tau <- c(0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99)
 n <- 300
 nSim <- 100
 simDesigns <- c("gaus", "mixednorm", "shapebeta")
-x <- seq(0.5, n, 1)/n
 methods <- c("detrend_SIC", "detrend_valid", "npqw", "qsreg", "rqss") #, "rqss")
 MSEs <- as.data.frame(matrix(NA, nrow = nSim*length(methods)*length(simDesigns), 
-                             ncol = length(tau)+3))
-colnames(MSEs) <- c("Design", "Sim", "Method", paste0("tau_", tau))
+                             ncol = length(tau)+4))
+colnames(MSEs) <- c("Design", "Sim", "Method", "n", paste0("tau_", tau))
 k <- 1
 for(simDesign in simDesigns){  
-  Q <- trueQuantile(simDesign, x, tau)
+  for (n in c(300,500,1000)){
   for (i in 1:nSim){
+    load(sprintf("../SimData/%s_n_%i_sim%03.0f.RData", simDesign, n, i))
+    Q <- trueQuantile(simDesign, df$x, tau)
     for (method in methods){
       load(sprintf("../SimResults/%s/%s_n_%i_sim%03.0f.RData", 
               method, simDesign, n, i))
       MSEs[k,1] <- simDesign
       MSEs[k,2] <- i
       MSEs[k,3] <- method
-      MSEs[k,4:ncol(MSEs)] <- colMeans((trend - Q)^2)
+      MSEs[k,4] <- n
+      MSEs[k,5:ncol(MSEs)] <- colMeans((trend - Q)^2)
       k <- k+1
     }
   }
+  }
 }
 
-MSEs_long <- MSEs %>% gather("tau", "MSE", -c("Design", "Sim", "Method")) 
+tmp <- MSEs %>% filter(Design == "gaus", n==1000)
+
+hist(tmp[tmp$Method == "detrend_SIC", "tau_0.01"])
+mean(tmp[tmp$Method == "rqss", "tau_0.01"])
+
+which(tmp[tmp$Method == "detrend_SIC", "tau_0.01"]>.04)
+
+
[…]

### 2018-12-04 Manuscript/simulation_notes.tex

@@
 
-We compare the performance of our quantile trend filtering model with the models of 3 others using designs proposed by \cite{Racine2017}. The models compared are
+We compare the performance of our quantile trend filtering method with the three previously published methods using designs proposed by \cite{Racine2017}. The methods compared are
 
@@
 	\end{equation}
-	defining $\Phi^{-1}_{(Y_i, b(X_i))}(\delta_0)$ as the quantile function of the Normal distribution with mean $Y_i$ and standard deviation $b(X_i)$ evaluated at $\tau$, $\delta_0$ is a function of $\tau$ and chosen empirically, $h$ is a tuning parameter and $K$ is a kernel function.
+	defining $\Phi^{-1}_{(Y_i, b(X_i))}(\delta_0)$ as the quantile function of the Normal distribution with mean $Y_i$ and standard deviation $b(X_i)$ evaluated at $\tau$.
+$\delta_0$ is a function of $\tau$ and chosen empirically, $h$ is a tuning parameter and $K$ is a kernel function.
 Code was obtained from the author for the \texttt{quantile-ll} method.
@@
 	\end{equation}
-	The function \texttt{qsreg} in the texttt{fields} R package was used.
+	The function \texttt{qsreg} in the \texttt{fields} R package was used. The smoothing parameter is chosen automatically using generalized cross validation on the pseudo data.
 
@@
 	$\sum_i\rho_{\tau}(y_i - \theta_i) + \lambda||D\theta||_1$ and $\lambda$ is chosen using SIC from above.
+A single value of $\lambda$ was chosen by scaling and summing SIC values across all quantiles.
 	\item \texttt{detrendr\_valid}: Our method where lambda is chosen by leaving out every 5th observation as a validation data set and evaluating the check loss function on the validation data. 
@@
 
-Three simulation designs from \cite{Racine2017} were considered. For all designs $X_i$ was generated as a uniformly space sequence in $[0,1]$ and the mean function was given as 
-$$Y_i = sin(2\pi X_i) + \epsilon_i(X_i)$$
+Three simulation designs from \cite{Racine2017} were considered. For all designs $X_i$ was generated as a uniformly spaced sequence in $[0,1]$ and the response $Y$ was generated as 
+$$Y_i = sin(2\pi x_i) + \epsilon_i(x_i)$$
 The three error distributions considered were 
 \begin{itemize}
- \item Gaussian: $\epsilon_i \sim N\left(0, \left(\frac{1+x_i^2}{4}\right)^2\right)$
+ \item Gaussian: $\epsilon_i(x_i) \sim N\left(0, \left(\frac{1+x_i^2}{4}\right)^2\right)$
  \item Beta: $\epsilon_i \sim Beta(1, 11-10x_i)$
@@
 
-100 datasets were generated of sizes 300, 500 and 1000. The MSE was calculated as $\frac{1}{n}\sum_i (\hat{q}_{\tau}(x_i) - q_\tau(x_i))^2$.
+100 datasets were generated of sizes 300, 500 and 1000.
+The MSE was calculated as $\frac{1}{n}\sum_i (\hat{q}_{\tau}(x_i) - q_\tau(x_i))^2$.
+The plots below show the mean MSE $\pm$ twice the standard error by method, quantile level and sample size.
 

### 2018-12-05 Manuscript/simulation_notes.tex
peak figures and extended BIC function

@@
 \includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
+
+\section{Peaks Simulation}
+We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
+\begin{equation}
+Y(t) = s(t) + b(t) + \epsilon
+\end{equation} 
+where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon \sim N(0, \sigma^2)$ is an error component.
+We assume $t$ is a uniformly spaced sequence between 0 and 1.
+We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from 2 to 10 with equal probability, and coefficients drawn from a normal distribution with mean and variance equal to 1.
+The true signal function is assumed to be zero with Gaussian peaks.
+The number of peaks is sampled from 5 to 15 with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $2/n$ and $2/n + .01$.
+One hundred datasets were generated for $n=500$ and $n=1000$.
+
+\begin{figure}
+	\caption{Example of simulated peaks, baseline, and observed measurements.}
+	\includegraphics[width = \linewidth]{Figures/ex_peaks.png}
+\end{figure}
+
+We compare the methods ability to estimate the baseline using a low quantile, $\tau \in \{0.05, 0.1\}$ and calculate the MSE using the simulated baseline value as the standard.
+The npqw method performs significantly worse than the other methods and is not included in the figures.
+
+\begin{figure}[h!]
+	\caption{MSEs compared to the simulated baseline function.}
+	\includegraphics[width = \linewidth]{Figures/peaks_mse.png}
+\end{figure}
+
+\begin{figure}[h!]
+	\caption{Example baseline fit.}
+	\includegraphics[width = \linewidth]{Figures/ex_baseline.png}
+\end{figure}
+
+\pagebreak
+
 \bibliographystyle{asa}
 \bibliography{detrendify}
-
 \end{document}

### 2018-12-05 Code/design_figures.R

@@ -48,3 +48,16 @@ ggplot(df, aes(x=x, y=y)) +
                      labels = tau) +
   labs(title = "Mixed Normal")
 ggsave("../Manuscript/Figures/mixednorm.png", width = 3, height = 3) 
+
+
+n <- 1000
+simDesign <- "peaks"
+load(sprintf("../SimData/%s_n_%i_sim%03.0f.RData", simDesign, n, i))
+ggplot(df, aes(x=x, y=y, col = "data")) + 
+  geom_line() +
+  geom_line(aes(y=baseline, col = "baseline")) +
+  geom_line(aes(y=peaks+baseline, col = "signal + baseline")) + 
+  theme_bw() + 
+  scale_color_manual(values = c("red", "grey", "blue")) + 
+  labs(x = "t", col = "")
+ggsave("../Manuscript/Figures/ex_peaks.png", width = 7, height = 2.5)

### 2018-12-07 Manuscript/simulation_notes.tex
updated eBIC function to have scale parameter

@@
 	\item \texttt{detrendr\_valid}: Our method where lambda is chosen by leaving out every 5th observation as a validation data set and evaluating the check loss function on the validation data.
+	\item \texttt{detrendr\_eBIC}:   The traditional BIC is given by 
+	\begin{equation}
+	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n 
+	\end{equation}	
+	where $\theta(s)$ is the parameter $\theta$ with those components outside $s$ being set to 0, and $\nu(s)$ is the number of components in $s$.
+If we assume an asymmetric Laplace likelihood $L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_i\rho_\tau(\frac{y_i - \theta_i}{\sigma})\right\}$ and the number of non-zero elements of $D\theta$ as $df$
+	\begin{equation}
+	\mbox{BIC}(df) = 2\sum_i\frac{1}{\sigma}\rho_{\tau}(y_i-\theta_i) + df\log n
+	\end{equation} 
+	We can choose and $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates. \cite{chen2008} proposed the extended BIC for large parameter spaces 
+	\begin{equation}
+	BIC_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j}
+	\end{equation}
+	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model.
+
 \end{itemize}

### 2018-12-07 Manuscript/simulation_notes.tex
update simulation notes and figures

@@
 
+\subsection*{Methods}
 We compare the performance of our quantile trend filtering method with the three previously published methods using designs proposed by \cite{Racine2017}. The methods compared are
@@
 	\begin{equation}
-	BIC_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j}
+	BIC_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j,}~~\gamma \in [0,1]
 	\end{equation}
 	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model.
+We used this criteria with $\gamma = 1$, $P=n-k$ where $k$ is the order of the differencing matrix and $j = \nu(s)$ is the number of non-zero entries in $D^{(k)}\theta$.
 
@@
 
+\subsection*{BIC examples}
+
+We used a single dataset to illustrate the difference between the scaled, unscaled and extended BIC criteria.
+
+\begin{figure}[h!]
+	\caption{Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.} 
+	\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
+	\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
+	\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
+\end{figure}
+
+\subsection*{Design}
 Three simulation designs from \cite{Racine2017} were considered. For all designs $X_i$ was generated as a uniformly spaced sequence in $[0,1]$ and the response $Y$ was generated as 
@@
 
-\includegraphics[width=\linewidth]{Figures/gaus_mse.png}
+\includegraphics[width=.6\linewidth]{Figures/gaus_mse.png}
 
-\includegraphics[width=\linewidth]{Figures/shapebeta_mse.png}
+\includegraphics[width=.6\linewidth]{Figures/shapebeta_mse.png}
 
-\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
+\includegraphics[width=.6\linewidth]{Figures/mixednorm_mse.png}
 
@@
 The number of peaks is sampled from 5 to 15 with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $2/n$ and $2/n + .01$.
-One hundred datasets were generated for $n=500$ and $n=1000$.
+One hundred datasets were generated for $n=\{300,500,1000\}$.
 

### 2018-12-07 Code/BIC_example_figs.R

@@ -1,0 +1,93 @@
+library(tidyverse)
+library(devtools)
+load_all("detrendr")
+rm(list=ls())
+
+set.seed(9856882)
+n <- 500
+x <- seq(0.5, n, 1)/n
+f <- sin(6*pi*x) 
+y <- f + ((1+x^2)/4)*rnorm(n)
+df <- data.frame(y=y, x=x, f=f)
+
+y <- df$y
+tau <- 0.1
+k <- 3
+lambdaSeq = length(y)^seq(0, 2, length.out=100)
+df_trend <- matrix(NA, nrow=length(lambdaSeq), ncol=length(tau))
+SIC <- matrix(NA, nrow=length(lambdaSeq), ncol=length(tau))
+SIC2 <- matrix(NA, nrow=length(lambdaSeq), ncol=length(tau))
+SIC3 <- matrix(NA, nrow=length(lambdaSeq), ncol=length(tau))
+SIC4 <- matrix(NA, nrow=length(lambdaSeq), ncol=length(tau))
+tau_min <- 0.5 - abs(0.5-tau)
+D <- get_Dk(length(y), k)
+gamma <- 1
+df_tol <- 1e-9
+for (i in 1:length(lambdaSeq)){
+  lam <- lambdaSeq[i]
+  suppressMessages(f_trend <- gurobi_trend(y, tau, lam, k))
+  resid_trend <- checkloss(y-f_trend, tau)
+  discr_diff <- abs(D%*%f_trend)
+  df_trend[i,] <- Matrix::colSums(discr_diff > df_tol) #colSums(abs(y-f_trend)<df_tol)
+  SIC[i] <- 2*colSums(resid_trend)/tau_min + log(n)*df_trend[i,] +
+    2*gamma*log(choose(n-k, df_trend[i,]))
+  SIC2[i] <- 2*colSums(resid_trend) + log(n)*df_trend[i,] 
+  SIC3[i] <- 2*colSums(resid_trend)/tau_min + log(n)*df_trend[i,] 
+  SIC4[i] <- log(colMeans(resid_trend)) + log(n)*df_trend[i,]/(2*n) 
+}
+
+plot.df <- data.frame(df = df_trend[,1])
+plot.df$eBIC <- SIC
+plot.df$BIC_noScale <- SIC2
+plot.df$BIC_scale <- SIC3
+plot.df$SIC <- n*(SIC4+5)
+plot.df$lambda <- lambdaSeq
+
+
[…]

### 2018-12-07 Code/detrendr/R/lambda_SIC.R

@@ -24,7 +24,7 @@ lambda_SIC <- function(y, tau, k,
     SIC_trend[i,] <- log(colMeans(resid_trend)) + log(n)*df_trend[i,]/(2*n)
   }
   
-  SIC_scale <- as.data.frame(scale(SIC_trend))
+  SIC_scale <- as.data.frame(scale(SIC_trend, scale = FALSE))
   SIC_scale$mean <- rowMeans(SIC_scale)
   
   if (single_lambda){

### 2018-12-12 Manuscript/simulation_notes.tex
new simulation folders, cleaning and organizing package code

@@
 
-\includegraphics[width=.6\linewidth]{Figures/gaus_mse.png}
+\includegraphics[width=\linewidth]{Figures/gaus_mse.png}
 
-\includegraphics[width=.6\linewidth]{Figures/shapebeta_mse.png}
+\includegraphics[width=\linewidth]{Figures/shapebeta_mse.png}
 
-\includegraphics[width=.6\linewidth]{Figures/mixednorm_mse.png}
+\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
 
@@
 We assume $t$ is a uniformly spaced sequence between 0 and 1.
-We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from 2 to 10 with equal probability, and coefficients drawn from a normal distribution with mean and variance equal to 1.
+We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from $n/50$ to $n/25$ with equal probability, and coefficients drawn from an exponential distribution with rate 1.
 The true signal function is assumed to be zero with Gaussian peaks.
-The number of peaks is sampled from 5 to 15 with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $2/n$ and $2/n + .01$.
-One hundred datasets were generated for $n=\{300,500,1000\}$.
+The number of peaks is sampled from $n/100$ to $n/50$ with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $1/n$ and $5/n$ and areas uniformly distributed between 0 and $20/n$.
+One hundred datasets were generated for $n=\{300,500,1000, 5000\}$.
+We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE.
 
@@
 
-We compare the methods ability to estimate the baseline using a low quantile, $\tau \in \{0.05, 0.1\}$ and calculate the MSE using the simulated baseline value as the standard.
-The npqw method performs significantly worse than the other methods and is not included in the figures.
-
 \begin{figure}[h!]
-	\caption{MSEs compared to the simulated baseline function.}
+	\caption{RMSEs compared to the simulated baseline function.}
 	\includegraphics[width = \linewidth]{Figures/peaks_mse.png}
-\end{figure}
-
-\begin{figure}[h!]
-	\caption{Example baseline fit.}
-	\includegraphics[width = \linewidth]{Figures/ex_baseline.png}
 \end{figure}

### 2018-12-12 Code/design_figures.R

@@ -50,12 +50,13 @@ ggplot(df, aes(x=x, y=y)) +
 ggsave("../Manuscript/Figures/mixednorm.png", width = 3, height = 3) 
 
 n <- 1000
+i <- 5
 simDesign <- "peaks"
 load(sprintf("../SimData/%s_n_%i_sim%03.0f.RData", simDesign, n, i))
 ggplot(df, aes(x=x, y=y, col = "data")) + 
   geom_line() +
+  geom_line(aes(y=peaks+baseline, col = "signal + baseline")) +
   geom_line(aes(y=baseline, col = "baseline")) +
-  geom_line(aes(y=peaks+baseline, col = "signal + baseline")) + 
   theme_bw() + 
   scale_color_manual(values = c("red", "grey", "blue")) + 
   labs(x = "t", col = "")

### 2018-12-12 Code/detrendr/R/get_model_data.R

@@ -1,0 +1,95 @@
+#' Get constraint matrix for linear program 
+#'
+#' \code{get_constraint_mat} Returns the constraint matrix
+#'
+#' @param tau quantiles being estimated
+#' @param lambda smoothing parameters, should be same length as tau
+#' @param n length of data
+#' @param m n-k where k is order of differencing matrix
+#' @param missInd output of which(is.na(y))
+#' @export
+get_obj <- function(tau, lambda, n, m, missInd){
+  obj <- c()
+  nT <- length(tau)
+  for (i in 1:nT){
+    obj <- c(obj, rep(tau[i], n), rep((1-tau[i]), n), rep(lambda[i], 2*m))
+  }
+  obj[missInd] <- 0
+  obj[missInd + n] <- 0
+  return(obj)
+}
+
+#' Get constraint matrix for linear program 
+#'
+#' \code{get_constraint_mat} Returns the constraint matrix
+#'
+#' @param D discrete difference matrix
+#' @param nT number of quantiles being estimated
+#' @export
+get_constraint_mat <- function(D, nT){
+  n <- ncol(D)
+  m <- nrow(D)
+  np <- 2*n + 2*m
+  # Constraint Matrix
+  if (nT == 1){
+    A  <- cbind(D, -D, Diagonal(m), -Diagonal(m))
+  } else {
+    A <- Matrix(0, nrow =  m*nT + n*(nT-1), ncol= np*nT, sparse=TRUE)
+    for (i in 1:nT){
+      # D%*%theta = eta constraint
+      A[(1+m*(i-1)):(m*i), (1+np*(i-1)):(np*i)] <-
+        cbind(D, -D, Diagonal(m), -Diagonal(m))
+      
+      # Non-crossing theta(tau) constrains
+      if (i < nT){
+        A[(m*nT+1+n*(i-1)):(m*nT+n*(i)), (1+(i-1)*np):(2*n + (i-1)*np)] <-
+          cbind(Diagonal(n), -Diagonal(n))
+        
+        A[(m*nT+1+n*(i-1)):(m*nT+n*(i)),
+          (1+i*np):(2*n + i*np)] <-
+          cbind(-Diagonal(n), Diagonal(n))
+      }
+    }
+  }
+  return(A)
+}
+
+#' Get linear program arguments for gurobi or glpk solver
+#'
+#' \code{get_model_data} Returns the linear program solver arguments
+#'
[…]

### 2019-01-03 Manuscript/detrendify_JCGS.tex
Description of ADMM algorthim for windows, and smoothing splines, function for lambda selection using SIC and validation dataset.
updated and organized manuscript

@@
 \RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
-\usepackage{bm,graphicx,psfrag,epsf}
-\usepackage{enumerate}
+\usepackage{graphicx}
+%\usepackage{enumerate}
 \usepackage{natbib}
-%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-%\RequirePackage{hypernat}
-\usepackage{paralist, booktabs, multirow}
-%\usepackage{url}
+\usepackage{url} % not crucial - just used below for the URL
 
-\newtheorem{lemma}{Lemma}
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
 \newcommand{\blind}{0}
-\newcommand{\trace}{\text{Trace}}
-\numberwithin{equation}{section}
-\theoremstyle{plain}
-%\newtheorem{thm}{Theorem}[section]
-%\newtheorem{mydef}{Definition}
-%\newtheorem{remark}{Remark}
 
-%% ----------------------------------------------------------------------
-%% Packages
-%% ----------------------------------------------------------------------
-%\usepackage{amsmath,amsfonts,amssymb}
-
-\usepackage{amssymb}
-\usepackage[usenames]{color}
-%\usepackage{amsbsy,amsthm,amsmath,amssymb}
-
-%% For \mathscr
-\usepackage[mathscr]{eucal}
-
-%% For \llbracket and \rrbracket, \varoast, \varoslash
-\usepackage{stmaryrd}
-
-%% For \boldsymbol
-%\usepackage{amsbsy}
-
-%%% For \bm (bold math)
-%\usepackage{bm}
-
-%% For proper spacing in text macros
-\usepackage{xspace}
-
-%% For special lists like inparaenum, compactenum, compactitem
-\usepackage{paralist}
-
-%% Turn on/off options for PDF
-\usepackage{ifpdf}
-
-%% Graphics
-\usepackage{graphicx}
-\ifpdf
-\usepackage{epstopdf}
-% REMOVE NEXT LINE TO REVERT TO PNG FILES
-\DeclareGraphicsExtensions{.eps,.pdf}
-\else
-\usepackage{epsfig}
-\fi
-
-%% For \subfloat command
-\usepackage{subfig}
-
-%% For hyperlinks. Should always be the last package. Omit for DVI version.
-\ifpdf
-\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
-\else
-\newcommand{\href}[2]{{#2}}
-\usepackage{url}
-\fi
-
-%% For tabulary
-\usepackage{tabulary}
-
-%% Show labels
-%\usepackage[notref,notcite]{showkeys}
-
-%% Algorithm
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-
-%% ----------------------------------------------------------------------
-%% Hyper-linked References
-%% ----------------------------------------------------------------------
-
-\ifpdf
-\newcommand{\Sec}[1]{\hyperref[sec:#1]{Section~\ref*{sec:#1}}} %section
-\newcommand{\App}[1]{\hyperref[sec:#1]{Appendix~\ref*{sec:#1}}} %appendix
-\newcommand{\Eqn}[1]{\hyperref[eq:#1]{{\rm (\ref*{eq:#1})}}} %equation
-\newcommand{\Part}[1]{\hyperref[part:#1]{(\ref*{part:#1})}} %part of theorem
-\newcommand{\Fig}[1]{\hyperref[fig:#1]{Figure~\ref*{fig:#1}}} %figure
-\newcommand{\Tab}[1]{\hyperref[tab:#1]{Table~\ref*{tab:#1}}} %table
-\newcommand{\Thm}[1]{\hyperref[thm:#1]{Theorem~\ref*{thm:#1}}} %theorem
-\newcommand{\Lem}[1]{\hyperref[lem:#1]{Lemma~\ref*{lem:#1}}} %lemma
-\newcommand{\Prop}[1]{\hyperref[prop:#1]{Proposition~\ref*{prop:#1}}} %proposition
-\newcommand{\Cor}[1]{\hyperref[cor:#1]{Corollary~\ref*{cor:#1}}} %corollary
-\newcommand{\Def}[1]{\hyperref[def:#1]{Definition~\ref*{def:#1}}} %definition
-\newcommand{\Alg}[1]{\hyperref[alg:#1]{Algorithm~\ref*{alg:#1}}} %algorithm
-\newcommand{\Ex}[1]{\hyperref[ex:#1]{Example~\ref*{ex:#1}}} %example
-\newcommand{\As}[1]{\hyperref[as:#1]{Assumption~{\rm\ref*{as:#1}}}} %assumption
-\newcommand{\Reg}[1]{\hyperref[as:#1]{Condition~\ref*{reg:#1}}} %regularity condition
-\newcommand{\AlgLine}[2]{\hyperref[alg:#1]{line~\ref*{line:#2} of Algorithm~\ref*{alg:#1}}}
-\newcommand{\AlgLines}[3]{\hyperref[alg:#1]{lines~\ref*{line:#2}--\ref*{line:#3} of Algorithm~\ref*{alg:#1}}}
-\else
-\newcommand{\Sec}[1]{{Section~\ref{sec:#1}}} %section
-\newcommand{\App}[1]{{Appendix~\ref{sec:#1}}} %appendix
-\newcommand{\Eqn}[1]{{(\ref{eq:#1})}} %equation
-\newcommand{\Part}[1]{{(\ref{part:#1})}} %part of proof
-\newcommand{\Fig}[1]{{Figure~\ref{fig:#1}}} %figure
-\newcommand{\Tab}[1]{{Table~\ref{tab:#1}}} %table
-\newcommand{\Thm}[1]{{Theorem~\ref{thm:#1}}} %theorem
-\newcommand{\Lem}[1]{{Lemma~\ref{lem:#1}}} %lemma
-\newcommand{\Prop}[1]{{Property~\ref{prop:#1}}} %property
-\newcommand{\Cor}[1]{{Corollary~\ref{cor:#1}}} %corollary
-\newcommand{\Def}[1]{{Definition~\ref{def:#1}}} %definition
-\newcommand{\Alg}[1]{{Algorithm~\ref{alg:#1}}} %algorithm
-\newcommand{\Ex}[1]{{Example~\ref{ex:#1}}} %example
-\newcommand{\As}[1]{{Assumption~\ref{as:#1}}} %assumption
-\newcommand{\Reg}[1]{{R~\ref{reg:#1}}} %regularity condition
-\newcommand{\AlgLine}[2]{{line~\ref{line:#2} of Algorithm~\ref{alg:#1}}}
-\newcommand{\AlgLines}[3]{{lines~\ref{line:#2}--\ref{line:#3} of Algorithm~\ref{alg:#1}}}
-\fi
-
-\usepackage{paralist, booktabs, multirow}
-
-%% ----------------------------------------------------------------------
-%% Environments
-%% ----------------------------------------------------------------------
-
-%% Assumptions environments
-\newtheorem{assumption}{Assumption}[section]
-%% Special proof where box is inside final equation
-% The function \proof and \endproof are defined inside the SIAM class file.
-% Be sure to use \myproofend inside the final equation to produce the
-% special box that signifies the end of the proof.
-\newenvironment{myproof}{\proof}{}
-\newcommand{\myproofend}{\quad\endproof}
-\newtheorem{theorem}{Theorem}[section]
-
-%Added by Arvind
-\usepackage[normalem]{ulem}
-\newcommand{\arvind}[1]{{\color{red} #1}}
-
-%%%%%
-
-%\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
-%
-%\RequirePackage[OT1]{fontenc}
-%\RequirePackage{amsthm,amsmath}
-%\RequirePackage [round,authoryear]{natbib}
-%\RequirePackage{hypernat}
-%
-%\usepackage{amsbsy,amsthm,amsmath,amssymb}
-%\usepackage{graphicx,booktabs}
-%%\usepackage{subfig}
-\usepackage{mathtools}
-\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-%\setlength{\oddsidemargin}{.75in}
-%\setlength{\evensidemargin}{.75in}
-%\setlength{\textwidth}{5in}
-\newtheorem{proposition}{Proposition}[section]
-%\newtheorem{corollary}{Corollary}[section]
-%\newtheorem{example}{Example}[section]
-%\newtheorem{definition}{Definition}[section]
-
-%% ----------------------------------------------------------------------
-%% Definitions
-%% ----------------------------------------------------------------------
-
-%% Shorthand for Real
-\newcommand{\Real}{\mathbb{R}}
-\newcommand{\dom}{{\bf dom}\,}
-
-%% Math notation
-\newcommand{\Tra}{^{\sf T}} % Transpose
-\newcommand{\Inv}{^{-1}} % Inverse
-\newcommand{\tr}{\operatorname{tr}} % Trace
-\newcommand{\Ker}{\operatorname{Ker}} % Kernel
-\def\vec{\mathop{\rm vec}\nolimits}
-\def\mtcz{\mathop{\rm mat}\nolimits} % Matricize operation
-\newcommand{\amp}{\mathop{\:\:\,}\nolimits}
-\newcommand{\bic}{\text{BIC}}
-
-%% Vectors
-\newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} % vector
-\newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
-\newcommand{\Vbar}[1]{{\bm{\M{A}r \mathbf{\MakeLowercase{#1}}}}} % vector
-\newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
-\newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}} % vector
-\newcommand{\Vn}[2]{\V{#1}^{(#2)}} % n-th vector
-\newcommand{\VnE}[3]{{#1}^{(#2)}_{#3}} % n-th vector
-\newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}} % vector element
-
-%% Matrices
-\newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}} % matrix
-\newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} % matrix element
-\newcommand{\MC}[2]{\V{#1}_{#2}}
-\newcommand{\Mr}[2]{\V{#1}_{#2}} % matrix row
-\newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} % matrix
-\newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} % matrix
-\newcommand{\MhatC}[2]{\Vhat{#1}_{#2}} % matrix column
-\newcommand{\Mbar}[1]{{\bm{\M{A}r \mathbf{\MakeUppercase{#1}}}}} % matrix
-\newcommand{\MbarC}[2]{\Vbar{#1}_{#2}} % matrix column
-\newcommand{\Mn}[2]{\M{#1}^{(#2)}} % n-th matrix
-\newcommand{\Mbarn}[2]{\Mbar{#1}^{(#2)}} % n-th matrix
-\newcommand{\Mtilden}[2]{\Mtilde{#1}^{(#2)}} % n-th matrix
-\newcommand{\MnTra}[2]{\M{#1}^{(#2){{\sf T}}}} % n-th matrix transpose
-\newcommand{\MnE}[3]{\MakeLowercase{#1}^{(#2)}_{#3}} % n-th matrix element
-\newcommand{\MnC}[3]{\V{#1}^{(#2)}_{#3}} % n-th matrix column
-\newcommand{\MbarnC}[3]{\Vbar{#1}^{(#2)}_{#3}} % n-th matrix column
-\newcommand{\MnCTra}[3]{\V{#1}^{(#2){{\sf T}}}_{#3}} % n-th matrix column transpose
-\newcommand{\MCTra}[2]{\V{#1}^{{\sf T}}_{#2}} % n-th matrix column transpose
-
-%% Tensors
-\newcommand{\T}[1]{\boldsymbol{\mathscr{\MakeUppercase{#1}}}} %tensor
-\newcommand{\Tbar}[1]{\boldsymbol{\M{A}r \mathscr{\MakeUppercase{#1}}}} %tensor
-\newcommand{\That}[1]{\boldsymbol{\hat \mathscr{\MakeUppercase{#1}}}} %tensor
-\newcommand{\Ttilde}[1]{\boldsymbol{\tilde \mathscr{\MakeUppercase{#1}}}} %tensor
-\newcommand{\TE}[2]{\MakeLowercase{#1}_{\MI{#2}}} % tensor element with multi-index
-
-%% Matrix-matrix operations
-\newcommand{\Kron}{\otimes} %Kronecker
-\newcommand{\Khat}{\odot} %Khatri-Rao
-\newcommand{\Hada}{\ast} %Hadamard
-\newcommand{\Divide}{\varoslash}
-
-% Matricize
-\newcommand{\Mz}[2]{\M{#1}_{(#2)}} % n-th mode matricize
-\newcommand{\Mzsets}[3]{\TM{#1}{#2\times#3}} % matricze with just the sets specified
-\newcommand{\Mzall}[4]{\TM{#1}{#2\times#3\,:\,#4}} % full matricize specification
-
-%% K-Tensor
-\newcommand{\KT}[1]{\left\llbracket #1 \right\rrbracket} % kruskal operator
-\newcommand{\KTsmall}[1]{\llbracket #1 \rrbracket} % small kruskal operator
-\newcommand{\KG}[1]{\langle #1 \rangle} % kruskal "group"
-
-%% text with quads around it
-\newcommand{\qtext}[1]{\quad\text{#1}\quad}
-
-%% Derivative
-\newcommand{\FD}[2]{\frac{\partial #1}{\partial #2}}
-
-%% Prox-operator
-\newcommand{\prox}[2]{\operatorname{prox}_{#1}({#2})}
-
-%% Sign-operator
-\newcommand{\sign}[1]{\operatorname{sign}({#1})}
-
-%% convex hull
-\newcommand{\conv}[1]{\operatorname{conv}({#1})}
-
-%% For comments to each other
-\newcommand{\Remark}[2]{{\color{red}Remark from #1: #2}\xspace}
-\newcommand{\Note}[1]{{{\color{red}{#1}}}\xspace}
-
-%\startlocaldefs
-%\numberwithin{equation}{section}
-%\theoremstyle{plain}
-%\newtheorem{thm}{Theorem}[section]
-%\endlocaldefs
-
+% DON'T change margins - should be 1 inch all around.
 \addtolength{\oddsidemargin}{-.5in}%
@@
 
-\pdfminorversion=4
+\begin{document}
 
-%% ----------------------------------------------------------------------
-%% Main Document
-%% ----------------------------------------------------------------------
-\begin{document}
+	%\bibliographystyle{natbib}
 
@@
 		\author{Halley Brantley\thanks{
-    Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: hlbrantl@ncsu.edu)} \,
+				Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: hlbrantl@ncsu.edu)} \,
 			Joseph Guinness\thanks{
-    Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: jsguinne@ncsu.edu)} \,
+				Department of Biological Statistics and Computational Biology, Cornell University, Ithaca, NY 14853 (E-mail: guinness@cornell.edu )} \,
 			and
 			%    and
-    Eric C. Chi\thanks{Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: eric$\_$chi@ncsu.edu).}    \\}
+			Eric C. Chi\thanks{Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: eric$\_$chi@ncsu.edu).}    \\}
 		\date{}
@@
 	\begin{abstract}
-Abstract
+		The text of your abstract.  200 or fewer words.
 	\end{abstract}
@@
 	\noindent%
-{\it Keywords:}  Key words
+	{\it Keywords:}  3 to 6 keywords, that do not appear in the title
 	\vfill
@@
 	\newpage
-\spacingset{1.45} % DON'T change the spacing!
-%% ----------------------------------------------------------------------
-%% Introduction
-%% ----------------------------------------------------------------------
+	\spacingset{1.5} % DON'T change the spacing!
 	\section{Introduction}
+	\label{sec:intro}
 
-\subsection{Background}
+	We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate for removing trends in time series.
+The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
+Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles.
+We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and proposed a modified criteria for choosing the smoothing parameter.
+We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.
 
-\begin{itemize}
-\item \cite{Oh2011} unified framework for non-parametric quantile regression by approximating check loss function with quadratic loss near zero.
-\item \cite{KoenkerNgPortnoy1994} introduce quantile smoothing splines. We need to discuss how our approach is different.
-\item \cite{Kim2009} introduce the concept of $\ell_1$-trend filtering.
-\item \cite{Tib2014} describes properties of trend filtering using quadratic loss.
-Shows that trend filtering estimates adapt to the local level of smoothness much better than smoothing splines, and exhibit a remarkable similarity to locally adaptive regression splines.
-Prove that (with the right choice of tuning parameter) the trend filtering estimate
-converges to the true underlying function at the minimax rate for functions whose kth derivative is of bounded variation.
-\item \cite{Ning2014} address problem of estimating a smooth baseline in noisy data with drift.
-\item \cite{Takeuchi2006nonparametric} Nonparametric quantile regression using SVM with Gaussian RBF kernels and check (pinball) loss. 
-\item \cite{yuan2006gacv} Comparison of cross-validation methods for quantile smoothing splines.  
-\end{itemize}
-We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate that can be computes using a linear time algorithm for removing trends in time series.
-The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
-Moreover we extend the basic framework to model multiple quantiles and ensure non-crossing.
-
-\subsection{Application}
-
-	 \begin{figure}[!h] 
-	 	\caption{Raw Data - three collocated sensors}
-		\includegraphics[width = 5in]{Figures/spodRaw.png}
-	\end{figure}
-	 \begin{figure}[!h] 
-	 	\caption{Detrended data - three collocated sensors}
-	\includegraphics[width = 5in]{Figures/spodDetrended.png}
-	\end{figure}
-
-Examples:
-\begin{itemize}
-\item Air quality \citep{Apte2017}
-\item ECG \citep{Sanyal2012}
-\item Chromatogram baseline estimation  \citep{Ning2014, Ilewicz2015}
-\item Galaxy spectrum baseline estimation \citep{Ilewicz2015,Bacher2016}
-\item Identifying absorption dips from black body radiation
-\end{itemize}
-
-Need to decide between detrending versus detrending + denoising.
-The BEADS \citep{Ning2014} method does both.
-We may wish to focus on detrending and then use wavelet SURE denoising as a postprocessing step, i.e.
-do a two-stage procedure, both of which can be done in linear time.
-On paper this should be faster than the BEADS procedure.
-Or we may just want to stick with detrending.
-There's Matlab code for BEADS, and the BEADS paper also points to two other popular methods in chromotography.
-
-Things to do:
-\begin{itemize}
-	\item Convergence of the algorithm
-	\item Convergence rate \citep{He2012, He2015, Davis2017}
-	\item Timing experiments of LP versus Spingarn
-	\item Make an R package - detrendr
-	\item Compare on synthetic data quality of solution with existing methods
-	\item Do comparisons on real data examples
-\end{itemize}
-
-%% ----------------------------------------------------------------------
-%% Quantile Regression
-%% ----------------------------------------------------------------------
-\section{Quantile Regression}
+	\subsection{Quantile Regression}
 
@@
 One remedy to blunt the influence of outliers is to compute the least absolute deviations (LAD) solution in place of the least squares one.
-Given a design matrix $\M{X} \in \Real^{n \times p}$ and continuous responses $\V{y} \in \Real^n$, we estimate a regression vector $\V{\theta} \in \Real^p$ so that $\M{X}\V{\theta}$ is a good approximation of $\V{y}$.
+Given a design matrix $\mathbf{X} \in \mathbb{R}^{n \times p}$ and continuous responses $y \in \mathbb{R}^n$, we estimate a regression vector $\theta \in \mathbb{R}^p$ so that $\mathbf{X}\theta$ is a good approximation of $y$.
 The LAD estimator is a solution to the problem
-\begin{eqnarray*}
-\underset{\V{\theta}}{\min}\; \frac{1}{n}\lVert \V{y} - \M{X}\V{\theta} \rVert_1.
-\end{eqnarray*}
-The above optimization problem generalizes the notion of the median of a collection of numbers. A median $\mu$ of $n$ reals $\VE{y}{1}, \ldots, \VE{y}{n}$ is the minimizer of the function
-\begin{eqnarray*}
-f(u) & = & \frac{1}{n}\sum_{i=1}^n \lvert \VE{y}{i} - \theta \rvert.
-\end{eqnarray*}
-Recall that the median is the 50th percentile or 0.5-quantile, namely half of the $\VE{y}{i}$ are less than or equal to $\mu$ and the other half is greater than or equal to $\mu$.
+	\begin{eqnarray}
+		\underset{\theta}{\min}\; \frac{1}{n}\lVert y - \mathbf{X}\theta \rVert_1.
+	\end{eqnarray}
+	The above optimization problem generalizes the notion of the median of a collection of numbers. A median $\mu$ of $n$ reals $y_1, \ldots, y_n$ is the minimizer of the function
+	\begin{eqnarray}
+		f(u) & = & \frac{1}{n}\sum_{i=1}^n \lvert y_i - \theta \rvert.
+	\end{eqnarray}
+	Recall that the median is the 50th percentile or 0.5-quantile, namely half of the $y_i$ are less than or equal to $\mu$ and the other half is greater than or equal to $\mu$.
 The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ to give us quantile regression \citep{Koenker1978}.
 
-First define the so-called ``check function"
-\begin{eqnarray*}
+	First define the so-called "check function"
+	\begin{eqnarray}
 		\rho_\tau(\Delta) & = & \begin{cases}
-\tau \Delta & \text{$\Delta \geq 0$} \\
--(1-\tau)\Delta & \text{$\Delta < 0$}. \\
+			\tau \Delta & \Delta \geq 0 \\
+			-(1-\tau)\Delta & \Delta < 0 \\
 		\end{cases}
-\end{eqnarray*}
-Then the $\tau$th quantile of the $\VE{y}{i}$ is a minimizer of the function
-\begin{eqnarray*}
-f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \theta).
-\end{eqnarray*}
+	\end{eqnarray}
+	Then the $\tau$th quantile of the $y_i$ is a minimizer of the function
+	\begin{eqnarray}
+		f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(y_i - \theta).
+	\end{eqnarray}
 	Returning to the regression context, we can generalize LAD regression to quantile regression, namely computing the minimizer of the function
-\begin{eqnarray*}
-f_\tau(\V{\theta}) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \langle \V{x}_i \mid \V{\theta} \rangle),
-\end{eqnarray*}
-where $\V{x}_i \in \Real^p$ denotes the $i$th row of $\M{X}$.
+	\begin{eqnarray}
+		f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(y_i - \langle x_i \mid \theta \rangle),
+	\end{eqnarray}
+	where $x_i \in \mathbb{R}^p$ denotes the $i$th row of $\mathbf{X}$.
 
@@
 	%% ----------------------------------------------------------------------
-\section{Trend Filtering}
+	\subsection{Trend Filtering}
 
-In the trend filtering problem \citep{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $\V{y} \in \Real^n$ by solving the following convex problem.
-\begin{eqnarray*}
-\underset{\V{\theta}}{\arg\min}\; \frac{1}{2n} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
-\end{eqnarray*}
-where $\lambda \geq 0$ is a regularization parameter that trades off the emphasis on the data fidelity term and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
-To understand the purpose of penalizing $\Mn{D}{k+1}$ consider the difference operator when $k = 0$.
-\begin{eqnarray*}
-\Mn{D}{1} = \begin{pmatrix}
+	In the trend filtering problem \citep{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $y \in \mathbb{R}^n$ by solving the following convex problem.
+	\begin{eqnarray}
+		\underset{\theta}{\arg\min}\; \frac{1}{2n} \lVert y - \theta \rVert_2^2 + \lambda \lVert \mathbf{D}^{(k+1)}\theta \rVert_1,
+	\end{eqnarray}
+	where $\lambda \geq 0$ is a regularization parameter that trades off the emphasis on the data fidelity term and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
+To understand the purpose of penalizing $\mathbf{D}^{(k+1)}$ consider the difference operator when $k = 0$.
+	\begin{eqnarray}
+		\mathbf{D}^{(1)} = \begin{pmatrix}
 			-1 & 1 & 0 & \cdots & 0 & 0 \\
@@
 		\end{pmatrix}
-\end{eqnarray*}
-Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \VE{\theta}{i} - \VE{\theta}{i+1} \rvert$ which is just total variation denoising in one dimension.
+	\end{eqnarray}
+	Thus, $\lVert \mathbf{D}^{(1)}\theta \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
 The penalty incentivizes solutions which are piece-wise constant.
-For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
-\begin{eqnarray*}
-\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
-\end{eqnarray*}
+For $k \geq 1$, the difference operator $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n-k-1) \times n}$ is defined recursively as follows
+	\begin{eqnarray}
+		\mathbf{D}^{(k+1)} & = & \mathbf{D}^{(1)}\mathbf{D}^{(k)}.
+	\end{eqnarray}
 	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
@@
 	%% ----------------------------------------------------------------------
-\section{Quantile Trend Filtering}
+	\section{Methods}
 
-We combine the ideas of quantile regression and trend filtering, namely consider the signal approximation problem, where the design $\M{X}$ is the identity matrix.
+	\subsection{Quantile Trend Filtering}
+
+	We combine the ideas of quantile regression and trend filtering, namely consider the signal approximation problem, where the design $\mathbf{X}$ is the identity matrix.
 
@@
 	\label{eq:quantile_trend}
-	\underset{\V{\theta}}{\min}\; \frac{1}{n} \sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) + \lambda \lVert \Mn{D}{k} \V{\theta} \rVert_1,
+	\underset{\theta}{\min}\; \frac{1}{n} \sum_{i=1}^n \rho_\tau(y_i - \theta_i) + \lambda \lVert \mathbf{D}^{(k)} \theta \rVert_1,
 	\end{eqnarray}
-where $\lambda$ is a nonnegative tuning parameter.
-As with the classic quantile regression, the quantile trend filtering problem can be solved by a linear program.
-We argue that it is better solved by Spingarn's method of partial inverses.
+	where $\lambda$ is a non-negative tuning parameter. As with the classic quantile regression, the quantile trend filtering problem can be solved by a linear program.
 
-\section{Related Work}
+	\subsection{Smoothing parameter choice}
+	Our method can easily handle missing data by changing the check loss function to output 0 for missing values.
+This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data.
+A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.
+\cite{KoenkerNgPortnoy1994} relate $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
+	\begin{equation}
+	SIC(p_{\lambda}) = log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}log n
+	\end{equation}
 
-\begin{itemize}
-	\item Quantile splines \citep{Oh2011}
-	\item BEADS 
-\end{itemize}
+	The traditional Bayesian Information Criterion (BIC) is given by 
+	\begin{equation}
+	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n 
+	\end{equation}	
+	where $\theta(s)$ is the parameter $\theta$ with those components outside $s$ being set to 0, and $\nu(s)$ is the number of components in $s$.
+If we assume an asymmetric Laplace likelihood $L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_i\rho_\tau(\frac{y_i - \theta_i}{\sigma})\right\}$ and the number of non-zero elements of $D\theta$ as $df$
+	\begin{equation}
+	\mbox{BIC}(df) = 2\sum_i\frac{1}{\sigma}\rho_{\tau}(y_i-\theta_i) + df\log n
+	\end{equation} 
+	We can choose and $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates. \cite{chen2008} proposed the extended BIC for large parameter spaces 
+	\begin{equation}
+	BIC_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j,}~~\gamma \in [0,1]
+	\end{equation}
+	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model.
+We used this criteria with $\gamma = 1$, $P=n-k$ where $k$ is the order of the differencing matrix and $j = \nu(s)$ is the number of non-zero entries in $D^{(k)}\theta$.
 
-%% ----------------------------------------------------------------------
-%% Spingarn's method of partial inverses
-%% ----------------------------------------------------------------------
-\section{Spingarn's method of partial inverses}
+	We used a single dataset to illustrate the difference between the scaled, unscaled and extended BIC criteria.
 
-We first review Spingarn's method \citep{Spingarn1985}, which solves the following equality constrained convex problem:
-\begin{equation}
-\label{eq:spingarn}
-\begin{split}
-\text{minimize}\; & \psi(\V{x}) \\
-\text{subject to}\; & \V{x} \in V,
-\end{split}
-\end{equation}
-where $V$ is a subspace. The problem \Eqn{spingarn} can be expressed as the unconstrained optimization problem
-\begin{equation}
-\label{eq:spingarn2}
-\text{minimize}\; \psi(\V{x}) + \iota_V(\V{x}),
-\end{equation}
-where $\iota_V$ is the indicator function of the set $V$. Spingarn's method applies Douglas-Rachford splitting to the problem \Eqn{spingarn2} to give the following updates.
+	\begin{figure}[h!]
+	\caption{Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.} 
+	\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
+	\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
+	\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
+	\end{figure}
 
-\begin{eqnarray*}
-\Vn{x}{k+1} & = & \prox{t\psi}{\Vn{z}{k}} \\
-\Vn{y}{k+1} & = & P_V\left(2\Vn{x}{k+1} - \Vn{z}{k} \right) \\
-\Vn{z}{k+1} & = & \Vn{z}{k} + \lambda^{(k)}\left(\Vn{y}{k+1} - \Vn{x}{k+1} \right).
-\end{eqnarray*}
-The parameter $t$ is a step-size and $\lambda^{(k)}$ is Krasnosel'ski{\u i}-Mann iteration (Need citation).
-We require that $\lambda^{(k)} \in ]0, 2[$ and that $\sum_n \lambda^{(k)} (2-\lambda^{(k)}) = \infty$.
-The mapping $P_V$ is the orthogonal projection onto the set $V$.
-Note that the algorithm iterates $\Vn{x}{k}$ will converge to a solution to problem $\Eqn{spingarn}$ \citep{Combettes2005}.
-We need to experiment with different over / under-relaxation parameters $\lambda^{(k)}$ and step sizes $t$.
-It will converge if we take $\lambda^{(k)} = 1$ and $t = 1$, but we may be able to converge faster in practice by taking non-trivial values.
-
-%% ----------------------------------------------------------------------
-%% Applying Spingarn's Method to Quantile Trend Filtering
-%% ----------------------------------------------------------------------
-\section{Applying Spingarn's Method to Quantile Trend Filtering}
-
-To simplify the notation we suppress the order $k$ and write $\Mn{D}{k}$ as $\M{D}$.
-We can reformulate our optimization problem \Eqn{quantile_trend} as the following equality constrained convex optimization problem.
-
-\begin{eqnarray*}
-\text{minimize}\; && f_1(\V{\theta}) + f_2(\V{\eta}) \\
-\text{subject to}\; && \V{\eta} = \M{D}\V{\theta}
-\end{eqnarray*}
-where
-\begin{eqnarray*}
-f_1(\V{\theta}) = \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) \quad\quad \text{and} \quad\quad
-f_2(\V{\eta}) = \lambda\lVert \V{\eta} \rVert_1.
-\end{eqnarray*}
-
-If we set $\psi(\V{\theta}, \V{\eta}) = f_1(\V{\theta}) + f_2(\V{\eta})$ and $V = \{ \V{z}\Tra = (\V{\theta}\Tra, \V{\eta}\Tra) : \V{\eta} = \M{D}\V{\theta} \}$, then we can apply Spingarn's method.
-Note that
-\begin{eqnarray*}
-\prox{th}{\V{\theta}, \V{\eta}} & = & (\prox{tf_1}{\V{\theta}}, \prox{tf_2}{\V{\eta}}) \\
-P_V(\V{\theta}, \V{\eta}) & = & \begin{pmatrix}
-\M{I} \\ \M{D}
-\end{pmatrix}
-\begin{pmatrix}
-\M{I} + \M{D}\Tra\M{D}
-\end{pmatrix}\Inv
-\begin{pmatrix}
-\V{\theta} + \M{D}\Tra\V{\eta}
-\end{pmatrix},
-\end{eqnarray*}
-where the projection $P_V$ requires a banded linear system solve, with bandwidth $k + 1$.
-This linear solve can be accomplished in $\mathcal{O}(n(k + 1)^2)$. The first solve using a banded Cholesky decomposition requires $\mathcal{O}(n(k + 1)^2)$. Subsequent solves require
-$\mathcal{O}(n(k+1))$. We can use RccpArmadillo to do banded Cholesky \citep{Eddelbuettel2014}.
-
-%% ----------------------------------------------------------------------
-%% Proximal mappings
-%% ----------------------------------------------------------------------
-\subsection*{Proximal mappings}
-
-We need the proximal mappings for $tf_1$ and $tf_2$.
-
-\begin{eqnarray*}
-\left [\prox{tf_1}{\V{\theta}} \right ]_i & = & \VE{y}{i} - \prox{(t/n)\rho_\tau}{\VE{y}{i} - \VE{\theta}{i}}, \\
-\left [\prox{tf_2}{\V{\eta}} \right ]_j & = & S(\VE{\eta}{j}, t\lambda).
-\end{eqnarray*}
-
-The proximal mapping for $tf_2$ is the element-wise softhresholding operator.
-We now derive the  proximal mapping of $\rho_\tau(\Delta)$, which can be evaluated in closed form.
-We need to find the minimizer of the following univariate function
-\begin{eqnarray*}
-g_\tau(\Delta) & = & \Delta[\tau - I(\Delta < 0)] +  \frac{n}{2t} (\Delta - w)^2, \\
-\end{eqnarray*}
-where $w \in \Real$ is given and $I(\Delta < 0)$ is 0 when $\Delta < 0$ and 1 otherwise.
-
-The subgradient of $\rho_\tau(\Delta) = \Delta[\tau - I(\Delta < 0)]$ is given by
-\begin{eqnarray*}
-\partial \rho_\tau(\Delta) & = & \begin{cases}
-\tau & \text{if $\Delta > 0$} \\
-\tau - 1 & \text{if $\Delta < 0$} \\
-[\tau - 1, \tau] & \text{if $\Delta = 0$}.
-\end{cases}
-\end{eqnarray*}
-The stationary condition is
-\begin{eqnarray*}
-\frac{n}{t}[w - \Delta] & \in & \partial \rho_\tau(\Delta).
-\end{eqnarray*}
-Therefore, the proximal mapping is given by 
-\begin{eqnarray*}
-\prox{(t/n) \rho_\tau}{w} & = & \begin{cases}
-w - \tau \frac{t}{n}& \text{if $w > \tau \frac{t}{n}$} \\
-w + (1-\tau)\frac{t}{n} & \text{if $w < - (1-\tau)\frac{t}{n}$} \\
-0 & \text{if $-(1- \tau)\frac{t}{n} \leq w \leq \tau \frac{t}{n}$.}
-\end{cases}
-\end{eqnarray*}
-
-\subsection*{Computational Costs}
-
-\subsubsection*{Precomputation}
-
-The following calculations need only be done once.
-
-\begin{itemize}
-\item $\mathcal{O}(n(k + 1)^2)$ to compute the banded Cholesky factorization of $\M{I} + [\Mn{D}{k}]\Tra[\Mn{D}{k}]$
-\end{itemize}
-
-\subsubsection*{Per-Iteration}
-
-The following calculations will be done every iteration.
-
-\begin{itemize}
-\item $\mathcal{O}(n)$ to compute $\prox{t\psi}{\V{\theta}, \V{\eta}}$
-\item $\mathcal{O}((k+1)(n-k+1))$ to compute $\V{\theta} + [\Mn{D}{k}]\Tra\V{\eta}$
-\item $\mathcal{O}(n(k+1))$ to compute $\V{\phi} = (\M{I} +[\Mn{D}{k}]\Tra[\Mn{D}{k}] )\Inv(\V{\theta} + \M{D}\Tra\V{\eta})$
-\item $\mathcal{O}((k+1)(n-k+1))$ to compute $\begin{pmatrix}
-\M{I} \\ \Mn{D}{k} \end{pmatrix}\V{\phi}$
-\end{itemize}
-
-The total cost is $\mathcal{O}(nk)$.
-
-\subsection{Summary}
-
-\begin{itemize}
-\item The overall computational complexity is essentially linear $\mathcal{O}(nk^2)$ for the initial banded Cholesky decomposition and the per-iteration complexity is $\mathcal{O}(nk)$.
-\item One could also apply Anderson acceleration \citep{Walker2011} to reduce the number of Spingarn updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
-\end{itemize}
-
-%% ----------------------------------------------------------------------
-%% Applying Spingarn's Method to Quantile Trend Filtering version 2
-%% ----------------------------------------------------------------------
-\section{Applying Spingarn's Method to Quantile Trend Filtering version 2}
-
-Following the specialized ADMM algorithm for trend filtering \citep{Ramdas2016}, we can also take advantage of fast exact solvers of the one-dimensional fussed lasso problem \citep{davies2001, Johnson2013}.
-The first method is based on taut strings, and the second is based on dynamic programming.
-Both of these methods run in linear time.
-A third exact method with worst case quadratic penalty but linear time in typical cases was proposed by \cite{Condat2013}.
-C code is available for all three methods.
-We should compare them.
-
-What is the reparameterization?
-
-\begin{eqnarray*}
-\text{minimize}\; && f_1(\V{\theta}) + f_2(\V{\eta}) \\
-\text{subject to}\; && \V{\eta} = \M{D}\V{\theta}
-\end{eqnarray*}
-where
-\begin{eqnarray*}
-f_1(\V{\theta}) = \frac{1}{n}\sum_{i=1}^n \rho_\tau(\VE{y}{i} - \VE{\theta}{i}) \quad\quad \text{and} \quad\quad
-f_2(\V{\eta}) = \lambda\lVert \Mn{D}{1}\V{\eta} \rVert_1.
-\end{eqnarray*}
-
-Difference with version 1: If we are interested in a 3rd order penalization, then $\M{D} = \Mn{D}{2}$ in version 2 as opposed to $\M{D} = \Mn{D}{3}$ in version 1.
-The proximal mapping for $f_2$ is solved using one of the exact solvers \citep{davies2001, Johnson2013, Condat2013}.
-Show some timing results between the different versions.
-Point readers to \cite{Ramdas2016} for a discussion of why this minor reparameterization may lead to speed up.
-
-\section{Other Practical Issues}
-
-\begin{itemize}
-	\item Read compressed data
-	\item May want to read data directly in C and not pull into R; make R just an interface
-	\item Use historical data to choose $\tau$ and $\lambda$
-\end{itemize}
-
-\section{To Do}
-
-\begin{itemize}
-	\item Find out if we can use the EPA data. Add co-authors?
-	\item Make version 2 with the three fast fused lasso solvers
-	\item Add homotopy / warm start
-	\item Add model selection
-	\item Do experiments to evaluate different choices of $t$ and $\lambda^{(k)}$
-	\item See if Anderson acceleration helps. Talk to Tim Kelley?
-	\item Write vignette
-	\item Do comparisons with BEAD
-	\item Do wind polar plots with and without removing trends
-	\item Compare with Splines
-\end{itemize}
-
-\section{Numerical Studies}
-
-We compare our detrending method with BEADS and the more general nonparametric quantile method introduced by \cite{Oh2011}.
-
-\section{ADMM for Smoothing Windows}
+	\subsection{ADMM for Smoothing Windows}
 
@@
 
-\section {Cross Validation Methods}
-Our method can easily handle missing data by changing the check loss function to output 0 for missing values.
-This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data.
-A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.
-The traditional smoothing spline for $y \in \mathbf{R}^n$ as a function of a predictor $x \in \mathbf{R}^n$ using the squared loss is estimated by minimizing
+	\pagebreak
 
+	\begin{figure}[!h] 
+		\caption{Windows fit separately compared to simultaneous fit, no signal present.}
+		\includegraphics[width = \linewidth]{Figures/overlapping_windows.png}
+	\end{figure}
+
+	\begin{figure}[!h] 
+		\caption{Windows fit with 150 iterations of ADMM, no signal present.}
+		\includegraphics[width = \linewidth]{Figures/admm_windows.png}
+	\end{figure}
+
+	\begin{figure}[!h] 
+		\caption{Windows fit separately compared to simultaneous fit, signal present.}
+		\includegraphics[width = \linewidth]{Figures/overlapping_windows2.png}
+	\end{figure}
+
+	\begin{figure}[!h] 
+		\caption{Windows fit with 300 iterations of ADMM, signal present.}
+		\includegraphics[width = \linewidth]{Figures/admm_windows2.png}
+	\end{figure}
+
+	\section{Simulation Study}
+
+	We compare the performance of our quantile trend filtering method with the three previously published methods using designs proposed by \cite{Racine2017}.
+The methods compared are: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
+The smoothing parameter $\lambda$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria as described in \cite{KoenkerNgPortnoy1994}.
+We further compare three criteria for choosing the smoothing parameter for our detrend method: \texttt{detrendr\_SIC}: Our method where we minimize $\sum_i\rho_{\tau}(y_i - \theta_i) + \lambda||D\theta||_1$ and $\lambda$ is chosen using SIC \citep{KoenkerNgPortnoy1994}.
+\texttt{detrendr\_valid}: Our method where lambda is chosen by leaving out every 5th observation as a validation data set and minimizing the evaluating the check loss function evaluated at the validation data.
+\texttt{detrendr\_eBIC}:  the new criteria we have proposed based on the extended BIC proposed by \cite{chen2008}.
+
+	Three simulation designs from \cite{Racine2017} were considered. For all designs $X_i$ was generated as a uniformly spaced sequence in $[0,1]$ and the response $Y$ was generated as 
+	$$Y_i = sin(2\pi x_i) + \epsilon_i(x_i)$$
+	The three error distributions considered were 
+	\begin{itemize}
+		\item Gaussian: $\epsilon_i(x_i) \sim N\left(0, \left(\frac{1+x_i^2}{4}\right)^2\right)$
+		\item Beta: $\epsilon_i \sim Beta(1, 11-10x_i)$
+		\item Mixed normal: $\epsilon_i$ is simulated from a mixture of $N(-1,1)$ and  $N(1,1)$ with mixing probability $x_i$.
+	\end{itemize}
+	\begin{figure}
+		\caption{Simulated data with true quantiles $\tau \in \{0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99\}$}	
+		\includegraphics[width=.3\linewidth]{Figures/gaus.png}
+		\includegraphics[width=.3\linewidth]{Figures/shapebeta.png}
+		\includegraphics[width=.3\linewidth]{Figures/mixednorm.png}
+	\end{figure}
+
+	100 datasets were generated of sizes 300, 500 and 1000.
+The MSE was calculated as $\frac{1}{n}\sum_i (\hat{q}_{\tau}(x_i) - q_\tau(x_i))^2$.
+The plots below show the mean MSE $\pm$ twice the standard error by method, quantile level and sample size.
+
+	\includegraphics[width=\linewidth]{Figures/gaus_mse.png}
+
+	\includegraphics[width=\linewidth]{Figures/shapebeta_mse.png}
+
+	\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
+
+	In all of the designs the proposed detrend methods are either better than or comparable to existing methods.
+The npqw method performs particularly poorly in the mixed normal design, due to the fact that it assumes the data comes from a scale-location model which is violated in this case.
+
+	\section{Peaks Simulation}
+	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
 	\begin{equation}
-\frac{1}{n}\sum_i (y_i - f(x_i))^2 + \lambda \int(f''(x))^2 dx
+	Y(t) = s(t) + b(t) + \epsilon
 	\end{equation} 
+	where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon \sim N(0, \sigma^2)$ is an error component.
+We assume $t$ is a uniformly spaced sequence between 0 and 1.
+We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from $n/50$ to $n/25$ with equal probability, and coefficients drawn from an exponential distribution with rate 1.
+The true signal function is assumed to be zero with Gaussian peaks.
+The number of peaks is sampled from $n/100$ to $n/50$ with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $1/n$ and $5/n$ and areas uniformly distributed between 0 and $20/n$.
+One hundred datasets were generated for $n=\{300,500,1000, 5000\}$.
+We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE.
 
-It has been shown that the solution to this problem is a cubic spline with a knot at each of the observed $x_i$.
-This problem can then be reformulated using a spline basis matrix $B \in \mathbf{R}^{nxn}$:
-\begin{equation}
-(y-B\beta)^T(y-B\beta) + n\lambda\beta^T\Omega\beta
-\end{equation}  
-where $\Omega_{jk} = \int B_j''(x)B_k''(x)dx$.
-Since each spline basis function is a defined as a cubic polynomial with non-zero second derivatives on a finite region, the integrals are finite.
-The solution is then given by
-$$f(x) = B\widehat{\beta} = B(B^TB + n\lambda\Omega)^{-1}B^Ty$$
-Because the solution is a linear combination of the observed $y$, the leave-one-out cross validation (LOOCV) statistics can be calculated without refitting the model.
-Defining $W = B\widehat{\beta} = B(B^TB + n\lambda\Omega)^{-1}B$ the LOOCV residual is given by
-\begin{equation}
-y_i - \widehat{y_i}^{(-i)} = \frac{y_i - \widehat{y_i}}{1-W_{[i,i]}}
-\end{equation}
-where $\widehat{y_i}$ is the prediction using the full dataset and $\widehat{y_i}^{(-i)}$ is the prediction from the model fit without $y_i$.
-In practice, generalized cross validation (GCV) has also been found to be an effective metric for choosing the smoothing parameter, with the validation MSE calculated as
-\begin{equation}
-\frac{\frac{1}{n}\sum(y_i-\widehat{y_i})^2}{(1-\frac{1}{n}tr(W))^2}
-\end{equation}
+	\begin{figure}
+		\caption{Example of simulated peaks, baseline, and observed measurements.}
+		\includegraphics[width = \linewidth]{Figures/ex_peaks.png}
+	\end{figure}
 
-Returning to the quantile regression case, \cite{nychka1995nonparametric} proposed a pseudo-data algorithm for a quantile spline estimator of the form 
-\begin{equation}
- \sum_i \rho_{\tau}(y_i - g(x_i)) + \lambda\int(g''(x))^2dx.
-\end{equation}
+	\begin{figure}[h!]
+		\caption{RMSEs compared to the simulated baseline function.}
+		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}
+	\end{figure}
 
-If $\rho_{\tau}(\cdot)$ were differentiable, the solution to this equation would take a form similar to that of the squared loss smoothing spline with weights equal to $\frac{\rho'_{\tau}(y_i-g(x_i))}{2(y_i - g(x_i))}$.
-Relying on this idea, Nychka proposed to solve the problem by iteratively solving the weighted smoothing spline.
-To address the non-differentiability they propose an approximation
-\begin{equation}
-\rho_{\tau, \delta}(u) = [\tau I(u>0) + (1-\alpha)I(u<0)]u^2/\delta
-\end{equation}
-They propose a corresponding approximate cross validation metric for choosing the tuning parameter:
-\begin{equation}
-ACV(\lambda) = \frac{1}{n}\sum_i \rho_{\tau}\left(\frac{y_i - \widehat{g}_\lambda(x_i)}{1-h_{ii}}\right)
-\end{equation}
-where $h_{ii} = \frac{\partial\widehat{g}_{\tau, \lambda}(x_i)}{\partial y_i}$.
-However \cite{yuan2006gacv} showed through simulation that this is not a good approximation to the robust cross validation score
-\begin{equation}
-RCV(\lambda) = \frac{1}{n}\sum\rho_{\tau}(y_i - g_\lambda^{(-i)}(x_i))
-\end{equation}
-and propose the GACV as an alternative 
-\begin{equation}
-GACV(\lambda) = \sum_i \rho_{\tau}\left(\frac{y_i - \widehat{g}_\lambda(x_i)}{n-tr(H)}\right)
-\end{equation}
-where $H_{ij} = \frac{\partial\widehat{g}_{\tau, \lambda}(x_i)}{\partial y_j}$
+	\begin{figure}[h!]
+		\caption{Example signal classification using threshold. Red indicates true signal > 0.1, blue indicates classified as signal after baseline removal using eBIC detrendr and a threshold of 1.}
+		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
+	\end{figure}
 
-\cite{KoenkerNgPortnoy1994} proposed a quantile smoothing spline using the check loss function and 1-norm rather than 2-norm:
-\begin{equation}
-R_{\tau,\lambda} = \rho_{\tau}(y - g(x)) + \lambda\int_0^1|g''(x)|dx
-\end{equation}
-with $0 = x_0 < ...
-< x_n = 1$.
-They prove that the resulting function $\widehat{g(x)}$ is a linear spline with knots at the observed values of $x$.
-They also argue that the solutions $\widehat{g}_{\tau,\lambda}(\cdot)$ are piecewise constant in $\lambda$ and that there exists a mesh $0= \lambda_0 < \lambda_1 < ...
-< \lambda_J$ such that $\widehat{g}_{\tau,\lambda}(\cdot)$ is the solution for all $\lambda \in [\lambda_{i-1}, \lambda_i]$.
-Relating $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
-\begin{equation}
-SIC(p_{\lambda}) = log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}log n
-\end{equation}
+	\begin{figure}[h!]
+		\caption{Miss-classification rates by threshold and method, values above the upper limit (npqw) not shown.}
+		\includegraphics[width = \linewidth]{Figures/peaks_missclass.png}
+	\end{figure}
+
+	\section{Application}
+
+	\begin{figure}
+		\caption{Example of simulated peaks, baseline, and observed measurements.}
+		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
+		\includegraphics[width = \linewidth]{Figures/corrected_data.png}
+		\includegraphics[width = \linewidth]{Figures/corrected_zoom_data.png}
+	\end{figure}
+
+	\section{Conclusion}
+	\label{sec:conc}
+
+	\bigskip
+	\begin{center}
+		{\large\bf SUPPLEMENTARY MATERIAL}
+	\end{center}
+
+	\begin{description}
+
+		\item[R-package for detrend routine:] R-package detrendr containing code to perform the diagnostic methods described in the article.
+The package also contains all datasets used as examples in the article.
+(GNU zipped tar file)
+
+	\end{description}
+
+	\section{References}
 

### 2019-01-03 Manuscript/extra_notes.tex
updated and organized manuscript

--- 2017-06-02 Notes/detrendify_spingarn.tex
@@
 
-%% ----------------------------------------------------------------------
-%% Environments
-%% ----------------------------------------------------------------------
+%\startlocaldefs
+%\numberwithin{equation}{section}
+%\theoremstyle{plain}
+%\newtheorem{thm}{Theorem}[section]
+%\endlocaldefs
 
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{definition}[theorem]{Definition}
-%% Assumptions environment
-\newtheorem{assumption}[theorem]{Assumption}
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+
+\pdfminorversion=4
 
 \begin{document}
-\title{Quantile Trend Filtering via Spingarn's Method}
-\noindent
-\textbf{Quantile Trend Filtering via Spingarn's Method}\\
-\textbf{Date: \today}\\
+\def\spacingset#1{\renewcommand{\baselinestretch}%
+	{#1}\small\normalsize} \spacingset{1}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\if0\blind
+{
+	\title{\bf Drift Removal for Time Series Data Using Quantile Trend Filtering}
+	\author{Halley Brantley\thanks{
+			Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: hlbrantl@ncsu.edu)} \,
+		Joseph Guinness\thanks{
+			Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: jsguinne@ncsu.edu)} \,
+		and
+		%    and
+		Eric C. Chi\thanks{Department of Statistics, North Carolina State University, Raleigh, NC 27695-8203 (E-mail: eric$\_$chi@ncsu.edu).}    \\}
+	\date{}
+	\maketitle
+} \fi
+
+\if1\blind
+{
+	\bigskip
+	\bigskip
+	\bigskip
+	\begin{center}
+		{\LARGE\bf Title}
+	\end{center}
+	\medskip
+} \fi
+
+\bigskip
+\begin{abstract}
+	Abstract
+\end{abstract}
+
+\noindent%
+{\it Keywords:}  Key words
+\vfill
+
+\newpage
+\spacingset{1.45} % DON'T change the spacing!
 %% ----------------------------------------------------------------------
@@
 
-We introduce a linear time algorithm for removing trends in time series.
+\subsection{Background}
+
+\begin{itemize}
+	\item \cite{Oh2011} unified framework for non-parametric quantile regression by approximating check loss function with quadratic loss near zero.
+	\item \cite{KoenkerNgPortnoy1994} introduce quantile smoothing splines. We need to discuss how our approach is different.
+	\item \cite{Kim2009} introduce the concept of $\ell_1$-trend filtering.
+	\item \cite{Tib2014} describes properties of trend filtering using quadratic loss.
+Shows that trend filtering estimates adapt to the local level of smoothness much better than smoothing splines, and exhibit a remarkable similarity to locally adaptive regression splines.
+Prove that (with the right choice of tuning parameter) the trend filtering estimate
+	converges to the true underlying function at the minimax rate for functions whose kth derivative is of bounded variation.
+	\item \cite{Ning2014} address problem of estimating a smooth baseline in noisy data with drift.
+	\item \cite{Takeuchi2006nonparametric} Nonparametric quantile regression using SVM with Gaussian RBF kernels and check (pinball) loss. 
+	\item \cite{yuan2006gacv} Comparison of cross-validation methods for quantile smoothing splines.  
+\end{itemize}
+We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate that can be computes using a linear time algorithm for removing trends in time series.
+The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
+Moreover we extend the basic framework to model multiple quantiles and ensure non-crossing.
+
+\subsection{Application}
+
+\begin{figure}[!h] 
+	\caption{Raw Data - three collocated sensors}
+	\includegraphics[width = 5in]{Figures/spodRaw.png}
+\end{figure}
+\begin{figure}[!h] 
+	\caption{Detrended data - three collocated sensors}
+	\includegraphics[width = 5in]{Figures/spodDetrended.png}
+\end{figure}
 
@@
 \begin{itemize}
-\item Air quality 
-\item ECG \cite{Sanyal2012}
-\item Chromatogram baseline estimation  \cite{Ning2014, Ilewicz2015}
-\item Galaxy spectrum baseline estimation \cite{Ilewicz2015,Bacher2016}
+	\item Air quality \citep{Apte2017}
+	\item ECG \citep{Sanyal2012}
+	\item Chromatogram baseline estimation  \citep{Ning2014, Ilewicz2015}
+	\item Galaxy spectrum baseline estimation \citep{Ilewicz2015,Bacher2016}
 	\item Identifying absorption dips from black body radiation
@@
 	\item The overall computational complexity is essentially linear $\mathcal{O}(nk^2)$ for the initial banded Cholesky decomposition and the per-iteration complexity is $\mathcal{O}(nk)$.
-\item One could also apply Anderson acceleration \cite{Walker2011} to reduce the number of Spingarn updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
+	\item One could also apply Anderson acceleration \citep{Walker2011} to reduce the number of Spingarn updates, since the Douglas-Rachford algorithm is a fixed point algorithm.
 \end{itemize}
@@
 
-\bibliographystyle{plain}
+\section{Numerical Studies}
+
+We compare our detrending method with BEADS and the more general nonparametric quantile method introduced by \cite{Oh2011}.
+
+\section{ADMM for Smoothing Windows}
+
+To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
+We propose an alternating direction method of multipliers (ADMM) algorithm for solving large problems in a piecewise fashion.
+The ADMM algorithm, described by \cite{boyd2011distributed}, relies on the idea of dual ascent.
+If we consider the equality optimization problem
+\begin{align}
+&\text{minimize } f(x)\\
+&\text{subject to } Ax = b
+\end{align}
+with variable $x\in \mathbf{R}^n$, where $A \in \mathbf{R}^{mxn}$ and $f: \mathbf{R}^n \rightarrow \mathbf{R}$ is convex. The corresponding Lagrangian function is 
+\begin{equation}
+L(x,\omega) = f(x) + \omega^T(Ax-b)
+\end{equation}
+and the corresponding dual problem  is 
+\begin{equation}
+\text{maximize } \inf_x L(x,\omega)
+\end{equation}
+Assuming that strong duality holds (Slater's constraint: there exists a solution in the interior of the domain), the optimal values of the primal and dual variables are the same.
+The dual ascent method, uses gradient ascent to solve the dual problem and thus the primal problem using the iterates
+\begin{align}
+&x^{k+1} = \arg\min L(x, \omega^k)\\
+&\omega^{k+1} = \omega^k + \alpha^k(Ax^{k+1} - b)
+\end{align}
+where $\alpha^k > 0$ is a step size and $k$ is the iteration counter.
+In many cases (including our problem), the $x$ update fails because the problem is unbounded below.
+To remedy this issue the augmented Lagrangian was introduced:
+\begin{equation}
+L_{\gamma}(x,\omega) = f(x)+\omega^T(Ax-b) + \frac{\gamma}{2}||Ax-b||_2^2
+\end{equation}
+where $\gamma > 0$ is called the penalty parameter. Applying dual ascent to the modified problem results in the method of multipliers algorithm which converges under more general conditions:
+\begin{align}
+&x^{k+1} = \arg\min L_\gamma(x, \omega^k)\\
+&\omega^{k+1} = \omega^k + \gamma(Ax^{k+1} - b)
+\end{align}
+Consider a new optimization problem with a separable objective function. 
+\begin{align}
+&\text{minimize } f_1(x_1) + f_2(x_2)\\
+&\text{subject to } A_1x_1 + A_2x_2 = b
+\end{align}
+where $x = (x_1, x_2)$, $i \in \{1,2\}$, and  $x_i \in \mathbf{R}^{n_i}$. The dual ascent updates can be expressed as 
+\begin{align}
+&x_i^{k+1} = \arg\min L_i(x_i, \omega^k)\\
+&\omega^{k+1} = \omega^k + \alpha^k(Ax^{k+1} - b)
+\end{align}
+however we lose the separability if the augmented Lagrangian is used instead, i.e.
+the augmented Lagrangian can not be separated into functions of $x_i$.
+The Alternating Direction Method of Multipliers (ADMM) addresses this problem and maintains separability while using the augmented Lagrangian to improve convergence through the following updates
+\begin{align}
+&x_1^{k+1} = \arg\min_{x_1} L_{\gamma}(x_1, x_2^k, \omega^k)\\
+&x_2^{k+1} = \arg\min_{x_2} L_{\gamma}(x_1^{k+1}, x_2, \omega^k)\\
+&\omega^{k+1} = \omega^k + \alpha^k(A_1x_1^{k+1} + A_2x_2^{k+1} - b)
+\end{align}
+
+We now consider the quantile regression trend filtering problem,  
+\begin{equation}
+\text{minimize } \rho_\tau(y-\theta) + \lambda||D^{(k)}\theta||_1\\
+\end{equation}
+where $y,\theta \in \mathbf{R}^n$, $y$ is the observed data, $\rho_{\tau}(u) = \sum_i(\tau - I(u_i < 0))u_i$ is the check loss function corresponding to quantile level $\tau$ and $D^{(k)}$ is the discrete differencing matrix of order $k$ and $\lambda$ is a tuning parameter that controls smoothness.
+We first re-parameterize $\phi = y-\theta$ so the problem is
+\begin{equation}
+\text{minimize } \rho_\tau(\phi) + \lambda||D^{(k)}(y-\phi)||_1\\
+\end{equation}
+We further divide $\phi$ order to solve smaller problems: Defining
+\begin{align}
+&\phi_1 = (\phi_{11}, \phi_{12})\\
+&\phi_2 = (\phi_{21}, \phi_{22}, \phi_{23})\\
+&\phi_3 = (\phi_{31}, \phi_{32})\\
+&\phi = (\phi_{11}, \phi_{12}=\phi_{21}, \phi_{22}, \phi_{23}=\phi_{31}, \phi_{32}) \\
+\end{align}
+Dividing $y$ similarly, the problem then becomes 
+\begin{align}
+&\text{minimize } \sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1\\
+&\text{subject to: } \phi_{12}=\phi_{21}, ~~ \phi_{23}=\phi_{31}\\
+\end{align}
+We can further simplify by defining 
+\begin{align}
+&\overline{\phi} = (\phi_{11}, \frac{\phi_{12}+\phi_{21}}{2}, \phi_{22}, \frac{\phi_{23}+\phi_{31}}{2}, \phi_{32}) \\
+&\overline{\phi_1} = (\phi_{11}, \frac{\phi_{12}+\phi_{21}}{2})\\
+&\overline{\phi_2} = ( \frac{\phi_{12}+\phi_{21}}{2}, \phi_{22}, \frac{\phi_{23}+\phi_{31}}{2})\\
+&\overline{\phi_3} = (\frac{\phi_{23}+\phi_{31}}{2}, \phi_{32})
+\end{align}
+so the problem becomes
+\begin{align}
+&\text{minimize } \sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1\\
+&\text{subject to: } \phi_{i}=\overline{\phi_i}\\
+\end{align}
+The augmented Lagrangian for this problem is 
+\begin{align}
+&L_\gamma(\phi_1,\phi_2, \phi_3, \overline{\phi_1}, \overline{\phi_2}, \overline{\phi_3}, \omega) = \\
+&\sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1 + \omega_i^T(\phi_i - \overline{\phi_i}) + \frac{\gamma}{2}||\phi_i - \overline{\phi_i}||_2^2
+\end{align}
+The ADMM updates are then given by
+\begin{align}
+&\phi_i^{k+1} = \arg\min_{\phi_i}\rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1 + \omega_i^{kT}(\phi_i - \overline{\phi_i}^k) + \frac{\gamma}{2}||\phi_i - \overline{\phi_i}^k||_2^2\\
+&\omega_i^{k+1} = \omega_i^{k} + \gamma(\phi_i^{k+1} - \overline{\phi_i}^{k+1})
+\end{align}
+The $\phi_i$ updates can be obtained using a quadratic program solver such as Gurobi and can be obtained in parallel.
+
+\pagebreak
+
+\begin{figure}[!h] 
+	\caption{Windows fit separately compared to simultaneous fit, no signal present.}
+	\includegraphics[width = \linewidth]{Figures/overlapping_windows.png}
+\end{figure}
+
+\begin{figure}[!h] 
+	\caption{Windows fit with 150 iterations of ADMM, no signal present.}
+	\includegraphics[width = \linewidth]{Figures/admm_windows.png}
+\end{figure}
+
+\begin{figure}[!h] 
+	\caption{Windows fit separately compared to simultaneous fit, signal present.}
+	\includegraphics[width = \linewidth]{Figures/overlapping_windows2.png}
+\end{figure}
+
+\begin{figure}[!h] 
+	\caption{Windows fit with 300 iterations of ADMM, signal present.}
+	\includegraphics[width = \linewidth]{Figures/admm_windows2.png}
+\end{figure}
+
+\pagebreak
+
+\section {Cross Validation Methods}
+Our method can easily handle missing data by changing the check loss function to output 0 for missing values.
+This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data.
+A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.
+The traditional smoothing spline for $y \in \mathbf{R}^n$ as a function of a predictor $x \in \mathbf{R}^n$ using the squared loss is estimated by minimizing
+
+\begin{equation}
+\frac{1}{n}\sum_i (y_i - f(x_i))^2 + \lambda \int(f''(x))^2 dx
+\end{equation}
+
+It has been shown that the solution to this problem is a cubic spline with a knot at each of the observed $x_i$.
+This problem can then be reformulated using a spline basis matrix $B \in \mathbf{R}^{nxn}$:
+\begin{equation}
+(y-B\beta)^T(y-B\beta) + n\lambda\beta^T\Omega\beta
+\end{equation}  
+where $\Omega_{jk} = \int B_j''(x)B_k''(x)dx$.
+Since each spline basis function is a defined as a cubic polynomial with non-zero second derivatives on a finite region, the integrals are finite.
+The solution is then given by
+$$f(x) = B\widehat{\beta} = B(B^TB + n\lambda\Omega)^{-1}B^Ty$$
+Because the solution is a linear combination of the observed $y$, the leave-one-out cross validation (LOOCV) statistics can be calculated without refitting the model.
+Defining $W = B\widehat{\beta} = B(B^TB + n\lambda\Omega)^{-1}B$ the LOOCV residual is given by
+\begin{equation}
+y_i - \widehat{y_i}^{(-i)} = \frac{y_i - \widehat{y_i}}{1-W_{[i,i]}}
+\end{equation}
+where $\widehat{y_i}$ is the prediction using the full dataset and $\widehat{y_i}^{(-i)}$ is the prediction from the model fit without $y_i$.
+In practice, generalized cross validation (GCV) has also been found to be an effective metric for choosing the smoothing parameter, with the validation MSE calculated as
+\begin{equation}
+\frac{\frac{1}{n}\sum(y_i-\widehat{y_i})^2}{(1-\frac{1}{n}tr(W))^2}
+\end{equation}
+
+Returning to the quantile regression case, \cite{nychka1995nonparametric} proposed a pseudo-data algorithm for a quantile spline estimator of the form 
+\begin{equation}
+\sum_i \rho_{\tau}(y_i - g(x_i)) + \lambda\int(g''(x))^2dx.
+\end{equation}
+
+If $\rho_{\tau}(\cdot)$ were differentiable, the solution to this equation would take a form similar to that of the squared loss smoothing spline with weights equal to $\frac{\rho'_{\tau}(y_i-g(x_i))}{2(y_i - g(x_i))}$.
+Relying on this idea, Nychka proposed to solve the problem by iteratively solving the weighted smoothing spline.
+To address the non-differentiability they propose an approximation
+\begin{equation}
+\rho_{\tau, \delta}(u) = [\tau I(u>0) + (1-\alpha)I(u<0)]u^2/\delta
+\end{equation}
+They propose a corresponding approximate cross validation metric for choosing the tuning parameter:
+\begin{equation}
+ACV(\lambda) = \frac{1}{n}\sum_i \rho_{\tau}\left(\frac{y_i - \widehat{g}_\lambda(x_i)}{1-h_{ii}}\right)
+\end{equation}
+where $h_{ii} = \frac{\partial\widehat{g}_{\tau, \lambda}(x_i)}{\partial y_i}$.
+However \cite{yuan2006gacv} showed through simulation that this is not a good approximation to the robust cross validation score
+\begin{equation}
+RCV(\lambda) = \frac{1}{n}\sum\rho_{\tau}(y_i - g_\lambda^{(-i)}(x_i))
+\end{equation}
+and propose the GACV as an alternative 
+\begin{equation}
+GACV(\lambda) = \sum_i \rho_{\tau}\left(\frac{y_i - \widehat{g}_\lambda(x_i)}{n-tr(H)}\right)
+\end{equation}
+where $H_{ij} = \frac{\partial\widehat{g}_{\tau, \lambda}(x_i)}{\partial y_j}$
+
+\cite{KoenkerNgPortnoy1994} proposed a quantile smoothing spline using the check loss function and 1-norm rather than 2-norm:
+\begin{equation}
+R_{\tau,\lambda} = \rho_{\tau}(y - g(x)) + \lambda\int_0^1|g''(x)|dx
+\end{equation}
+with $0 = x_0 < ...
+< x_n = 1$.
+They prove that the resulting function $\widehat{g(x)}$ is a linear spline with knots at the observed values of $x$.
+They also argue that the solutions $\widehat{g}_{\tau,\lambda}(\cdot)$ are piecewise constant in $\lambda$ and that there exists a mesh $0= \lambda_0 < \lambda_1 < ...
+< \lambda_J$ such that $\widehat{g}_{\tau,\lambda}(\cdot)$ is the solution for all $\lambda \in [\lambda_{i-1}, \lambda_i]$.
+Relating $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
+\begin{equation}
+SIC(p_{\lambda}) = log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}log n
+\end{equation}
+
+\section{Simulation Study}
+
+\subsection*{Methods}
+We compare the performance of our quantile trend filtering method with the three previously published methods using designs proposed by \cite{Racine2017}. The methods compared are
+
+\begin{itemize}
+	\item \texttt{npqw}: \cite{Racine2017} constrain the response to follow a smooth location scale model of the form $Y_i = a(X_i) + b(X_i)\epsilon_i$.
+They estimate the $\tau$\textsubscript{th} conditional quantile given $X_i = x$ using a kernel estimator
+	\begin{equation}
+		q_{\tau}(x) = \frac{\sum_{i=1}^n \Phi^{-1}_{(Y_i, b(X_i))}(\delta_0)K_h(X_i,x)}{\sum_{i=1}^nK_h(X_i, x)}
+	\end{equation}
+	defining $\Phi^{-1}_{(Y_i, b(X_i))}(\delta_0)$ as the quantile function of the Normal distribution with mean $Y_i$ and standard deviation $b(X_i)$ evaluated at $\tau$.
+$\delta_0$ is a function of $\tau$ and chosen empirically, $h$ is a tuning parameter and $K$ is a kernel function.
+Code was obtained from the author for the \texttt{quantile-ll} method.
+
+	\item \texttt{qsreg}: \cite{Oh2011} proposed a pseudo-data algorithm for a quantile spline estimator of the form 
+	\begin{equation}
+	\sum_i \rho_{\tau}(y_i - g(x_i)) + \lambda\int(g''(x))^2dx.
+	\end{equation} 
+	If $\rho_{\tau}(\cdot)$ were differentiable, the solution to this equation would take a form similar to that of the squared loss smoothing spline with weights equal to $\frac{\rho'_{\tau}(y_i-g(x_i))}{2(y_i - g(x_i))}$.
+Relying on this idea, Nychka proposed to solve the problem by iteratively solving the weighted smoothing spline.
+To address the non-differentiability they propose an approximation
+	\begin{equation}
+	\rho_{\tau, \delta}(u) = [\tau I(u>0) + (1-\alpha)I(u<0)]u^2/\delta
+	\end{equation}
+	The function \texttt{qsreg} in the \texttt{fields} R package was used. The smoothing parameter is chosen automatically using generalized cross validation on the pseudo data.
+
+	\item \texttt{rqss}: \cite{KoenkerNgPortnoy1994} Koenker proposed smoothing splines using trend filtering with the second order differencing matrix which results in linear splines.
+The function \texttt{rqss} in the \texttt{quantreg} package implements this method.
+The smoothing parameter $\lambda$ is chosen using a grid search and minimizing
+	\begin{equation}
+	SIC(p_{\lambda}) = log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}log n
+	\end{equation}
+	where $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots.
+
+	\item \texttt{detrendr\_SIC}: Our method where we minimize
+	$\sum_i\rho_{\tau}(y_i - \theta_i) + \lambda||D\theta||_1$ and $\lambda$ is chosen using SIC from above.
+A single value of $\lambda$ was chosen by scaling and summing SIC values across all quantiles.
+	\item \texttt{detrendr\_valid}: Our method where lambda is chosen by leaving out every 5th observation as a validation data set and evaluating the check loss function on the validation data.
+	\item \texttt{detrendr\_eBIC}:   The traditional BIC is given by 
+	\begin{equation}
+	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n 
+	\end{equation}	
+	where $\theta(s)$ is the parameter $\theta$ with those components outside $s$ being set to 0, and $\nu(s)$ is the number of components in $s$.
+If we assume an asymmetric Laplace likelihood $L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_i\rho_\tau(\frac{y_i - \theta_i}{\sigma})\right\}$ and the number of non-zero elements of $D\theta$ as $df$
+	\begin{equation}
+	\mbox{BIC}(df) = 2\sum_i\frac{1}{\sigma}\rho_{\tau}(y_i-\theta_i) + df\log n
+	\end{equation} 
+	We can choose and $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates. \cite{chen2008} proposed the extended BIC for large parameter spaces 
+	\begin{equation}
+	BIC_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j,}~~\gamma \in [0,1]
+	\end{equation}
+	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model.
+We used this criteria with $\gamma = 1$, $P=n-k$ where $k$ is the order of the differencing matrix and $j = \nu(s)$ is the number of non-zero entries in $D^{(k)}\theta$.
+
+\end{itemize}
+
+\subsection*{BIC examples}
+
+We used a single dataset to illustrate the difference between the scaled, unscaled and extended BIC criteria.
+
+\begin{figure}[h!]
+	\caption{Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.} 
+	\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
+	\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
+	\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
+\end{figure}
+
+\subsection*{Design}
+Three simulation designs from \cite{Racine2017} were considered. For all designs $X_i$ was generated as a uniformly spaced sequence in $[0,1]$ and the response $Y$ was generated as 
+$$Y_i = sin(2\pi x_i) + \epsilon_i(x_i)$$
+The three error distributions considered were 
+\begin{itemize}
+ \item Gaussian: $\epsilon_i(x_i) \sim N\left(0, \left(\frac{1+x_i^2}{4}\right)^2\right)$
+ \item Beta: $\epsilon_i \sim Beta(1, 11-10x_i)$
+\item Mixed normal: $\epsilon_i$ is simulated from a mixture of $N(-1,1)$ and  $N(1,1)$ with mixing probability $x_i$.
+\end{itemize}
+\begin{figure}
+\caption{Simulated data with true quantiles $\tau \in \{0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99\}$}	
+\includegraphics[width=.3\linewidth]{Figures/gaus.png}
+\includegraphics[width=.3\linewidth]{Figures/shapebeta.png}
+\includegraphics[width=.3\linewidth]{Figures/mixednorm.png}
+\end{figure}
+
+100 datasets were generated of sizes 300, 500 and 1000.
+The MSE was calculated as $\frac{1}{n}\sum_i (\hat{q}_{\tau}(x_i) - q_\tau(x_i))^2$.
+The plots below show the mean MSE $\pm$ twice the standard error by method, quantile level and sample size.
+
+\includegraphics[width=\linewidth]{Figures/gaus_mse.png}
+
+\includegraphics[width=\linewidth]{Figures/shapebeta_mse.png}
+
+\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
+
+\section{Peaks Simulation}
+We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
+\begin{equation}
+Y(t) = s(t) + b(t) + \epsilon
+\end{equation} 
+where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon \sim N(0, \sigma^2)$ is an error component.
+We assume $t$ is a uniformly spaced sequence between 0 and 1.
+We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from $n/50$ to $n/25$ with equal probability, and coefficients drawn from an exponential distribution with rate 1.
+The true signal function is assumed to be zero with Gaussian peaks.
+The number of peaks is sampled from $n/100$ to $n/50$ with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $1/n$ and $5/n$ and areas uniformly distributed between 0 and $20/n$.
+One hundred datasets were generated for $n=\{300,500,1000, 5000\}$.
+We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE.
+
+\begin{figure}
+	\caption{Example of simulated peaks, baseline, and observed measurements.}
+	\includegraphics[width = \linewidth]{Figures/ex_peaks.png}
+\end{figure}
+
+\begin{figure}[h!]
+	\caption{RMSEs compared to the simulated baseline function.}
+	\includegraphics[width = \linewidth]{Figures/peaks_mse.png}
+\end{figure}
+
+\begin{figure}[h!]
+	\caption{Example signal classification using threshold. Red indicates true signal > 0.1, blue indicates classified as signal after baseline removal using eBIC detrendr and a threshold of 1.}
+	\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
+\end{figure}
+
+\begin{figure}[h!]
+	\caption{Miss-classification rates by threshold and method, values above the upper limit (npqw) not shown.}
+	\includegraphics[width = \linewidth]{Figures/peaks_missclass.png}
+\end{figure}
+
+\section{Application}
+
+\begin{figure}
+	\caption{Example of simulated peaks, baseline, and observed measurements.}
+	\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
+	\includegraphics[width = \linewidth]{Figures/corrected_data.png}
+	\includegraphics[width = \linewidth]{Figures/corrected_zoom_data.png}
+\end{figure}
+
+\pagebreak
+
+\bibliographystyle{asa}
 \bibliography{detrendify}
-
 \end{document}

### 2019-01-03 Code/SimulateData.R

@@ -1,6 +1,6 @@
 library(splines)
 rm(list=ls())
-source("data_generating_functions.R")
+source("sim_generating_functions.R")
 
 set.seed(39207491)
 for(n in c(300,500,1000)){
@@ -11,7 +11,5 @@ for(n in c(300,500,1000)){
     save(df, file = sprintf("../SimData/gaus_n_%i_sim%03.0f.RData", n, i))
     df <- generate_shapebeta(n)
     save(df, file = sprintf("../SimData/shapebeta_n_%i_sim%03.0f.RData", n, i))
-    df <- generate_peaks_design(n)
-    save(df, file = sprintf("../SimData/peaks_n_%i_sim%03.0f.RData", n, i))
   }
 }

### 2019-01-03 Code/detrendr/R/get_trend_windows.R

@@ -34,7 +34,7 @@
 #' @export
 get_trend_windows <- function(y, tau, lambda, k, rho=1, window_size,
                            overlap, max_iter, update=10, 
-                           quad = FALSE){
+                           quad = TRUE){
   if(!("gurobi" %in% installed.packages()[,"Package"])){
     stop("Must have gurobi package installed to run this function.")
   } else {

### 2019-01-03 Code/detrendr/R/solve_functions.R

@@ -53,8 +53,8 @@ solve_gurobi <- function(model){
   iter <- 1
   while(result$status != "OPTIMAL"){
     print("Problem not solved, adding 0.01 to lambda.")
-    lambda <- model$obj[length(model_list$obj)]
-    model_list$obj[model$obj == lambda] <- lambda + 0.01
+    lambda <- model$obj[length(model$obj)]
+    model$obj[model$obj == lambda] <- lambda + 0.01
     iter <- iter+1
     if(iter == 10){
       stop("Problem not solved.")

### 2019-01-03 Code/sim_generating_functions.R

@@ -73,3 +73,21 @@ generate_mixednorm <- function(n){
   df <- data.frame(y=y, x=x, f=f)
   return(df)
 }
+
+trueQuantile <- function(simDesign, x, tau){
+  Q <- matrix(NA, length(x), length(tau))
+  for (i in 1:length(tau)){
+    if (simDesign == "gaus"){
+      Q[,i] <- sin(2*pi*x) + ((1+x^2)/4)*qnorm(tau[i])
+    } else if (simDesign == "shapebeta") {
+      Q[,i] <- sin(2*pi*x) + qbeta(tau[i], 1, 11-10*x)
+    } else if (simDesign == "mixednorm") {
+      cdf <- function(e, x, tau){
+        x*pnorm(e, 1, sqrt(.25)) + (1-x)*pnorm(e, -1, sqrt(.25)) - tau
+      }
+      q <- sapply(x, function(x) uniroot(cdf, c(-5, 5), x=x, tau = tau[i])$root)
+      Q[,i] <- sin(2*pi*x) + q
+    }
+  }
+  return(Q)
+}

### 2019-01-03 Code/simulations.R

@@ -1,197 +1,0 @@
-# Simulations
-library(fields)
-library(quantreg)
-library(devtools)
-# library(qrsvm)
-library(np)
-library(microbenchmark)
-load_all("detrendr")
-library(tidyverse)
-#source("function_jcgs.R")
-rm(list=ls())
-nSim <- 100
-n <- 500
-x <- (seq(1,n,1)-1)
-tau <- .1
-set.seed(986532)
-f_true <- sin(10*x/n) + (x/n + .25)*qnorm(tau, 0, .1)/.1
-loss_df <- data.frame(SimNum = 1:nSim, 
-                      trend_SIC = NA, trend_valid = NA, rqss = NA, 
-                      qsreg = NA, npqreg = NA)
-
-for (i in 1:nSim){
-  y <- sin(10*x/n) + (x/n + .25)*rnorm(n, 0, .1)/.1
-  lam_SIC <- lambda_SIC(y, tau, 3)
-  lam_valid <- lambda_valid(y, tau, 3, 5)
-  lam_SIC2 <- lambda_SIC(y, tau, 2, 
-                         lambdaSeq = seq(.5, length(y)/10, .2))
-  
-  f_trend_SIC <- gurobi_trend(y, tau, lam_SIC$lambda, k=3)
-  loss_df$trend_SIC[i] <- mean((f_trend_SIC-f_true)^2)
-  
-  f_trend_valid <- gurobi_trend(y, tau, lam_valid$lambda, k=3)
-  loss_df$trend_valid[i] <- mean((f_trend_valid-f_true)^2)
-  
-  fit_rqss <- rqss(y ~ qss(x, lambda = 2*lam_SIC2$lambda), tau = tau)
-  f_rqss <- predict(fit_rqss, data.frame(x=x))
-  loss_df$rqss[i] <- mean((f_rqss-f_true)^2)
-  
-  fit_qsreg <- qsreg(x, y, maxit.cv = 50, alpha=tau, hmin = -9)
-  f_qsreg <- predict(fit_qsreg)
-  loss_df$qsreg[i] <- mean((f_qsreg-f_true)^2)
-  
-  bw <- npcdistbw(y~x)
-  sink("NUL")
-  fit_npqreg <- npqreg(bws=bw, tau=tau)
-  sink()
-  f_npqreg <- fitted(fit_npqreg)
[…]

### 2019-01-03 Code/trueQuantile.R

@@ -1,21 +1,0 @@
-trueQuantile <- function(simDesign, x, tau){
-  Q <- matrix(NA, length(x), length(tau))
-  for (i in 1:length(tau)){
-    if (simDesign == "gaus"){
-      Q[,i] <- sin(2*pi*x) + ((1+x^2)/4)*qnorm(tau[i])
-    } else if (simDesign == "shapebeta") {
-      Q[,i] <- sin(2*pi*x) + qbeta(tau[i], 1, 11-10*x)
-    } else if (simDesign == "mixednorm") {
-      cdf <- function(e, x, tau){
-        x*pnorm(e, 1, sqrt(.25)) + (1-x)*pnorm(e, -1, sqrt(.25)) - tau
-      }
-      q <- sapply(x, function(x) uniroot(cdf, c(-5, 5), x=x, tau = tau[i])$root)
-      Q[,i] <- sin(2*pi*x) + q
-    }
-  }
-  return(Q)
-}
-
-truePeaks <- function(df){
-  
-}

### 2019-01-04 Manuscript/detrendify_JCGS.tex
manuscript updates, application metrics

@@
 \usepackage{url} % not crucial - just used below for the URL 
+\usepackage{placeins}
 
@@
 The number of peaks is sampled from $n/100$ to $n/50$ with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $1/n$ and $5/n$ and areas uniformly distributed between 0 and $20/n$.
-One hundred datasets were generated for $n=\{300,500,1000, 5000\}$.
+One hundred datasets were generated for $n=\{300, 500, 1000, 5000\}$.
 We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE.
 
-	\begin{figure}
+	\begin{figure}[h]
 		\caption{Example of simulated peaks, baseline, and observed measurements.}
@@
 	\end{figure}
+
+	In addition to the RMSE we also calculated the signal miss-classification rate.
+We first classified the "true" peaks function into signal or not based on a threshold of 0.1.
+We then classified the detrended data using three different thresholds and calculated the fraction miss-classified.
+An illustration of the observations classified as signal after detrending compared to the "true signal" is shown in Fig.
+9.
 
@@
 
+	Our detrend\_BIC method performs the best overall in terms of both RMSE and miss-classification rate.
+The lowest miss-classification rates were obtained using the detrend\_eBIC method and a threshold of 1 for all data sizes.
+While qsreg was competitive with our method in some cases, both the RMSE and miss-classification rate increased substantially with the size of the dataset.
+
 	\begin{figure}[h!]
-		\caption{Example signal classification using threshold. Red indicates true signal > 0.1, blue indicates classified as signal after baseline removal using eBIC detrendr and a threshold of 1.}
+		\caption{Example signal classification using threshold. Red indicates true signal $>0.1$, blue indicates classified as signal after baseline removal using eBIC detrendr and a threshold of 1.}
 		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
@@
 	\begin{figure}[h!]
-		\caption{Miss-classification rates by threshold and method, values above the upper limit (npqw) not shown.}
+		\caption{Miss-classification rates by threshold, data size, and method, values above the upper limit (npqw) not shown.}
 		\includegraphics[width = \linewidth]{Figures/peaks_missclass.png}
@@
 
+	\FloatBarrier
+
 	\section{Application}
 
+	Our windowed detrend method was used to removed the baseline drift from low cost air quality sensors so that the signal could be categorized using a simple threshold.
+The measurements were first standardized to have mean zero and variance 1.
+Three quantile levels for estimating the baseline trend were compared.
+The total dataset consisted of 52,322 observations per node.
+	The signal thresholds were set using the first 15,000 observations where it was known no signal was present.
+The thresholds were set as 3 times the standard deviation plus the mean of observations in this time period.
+The total number of seconds of signal for each node as well as the number of seconds where multiple nodes both reported signal is shown in Table 1.
+Table 2 shows the fraction of observations with different signal classifications by node combination.
+
 	\begin{figure}
-		\caption{Example of simulated peaks, baseline, and observed measurements.}
+		\caption{Low cost sensor data before drift removal (top), after drift removal using windowed detrend with eBIC (middle), 
+			and zoomed in on signal area (bottom). Horizontal lines represent signal thresholds.}
 		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
@@
 	\end{figure}
+
+	\input{application_missclass.tex}	
+	\input{application_signal.tex}
 

### 2019-01-04 Code/applications.R

@@ -1,28 +1,23 @@
+################################################################################
+# Calculate mis-classification rates in peaks simulation
+# Halley Brantley
+################################################################################
+library(tidyverse)
 library(devtools)
-library(gurobi)
+library(Hmisc)
 load_all("detrendr")
 rm(list=ls())
-
-dataDir <- "~/Desktop/EPA/SPod_Data/TestRange_Dec2017"
-datafiles <- dir(dataDir, pattern=".csv", full.names=TRUE)
-
-spod <- read.csv(datafiles[3], header=TRUE,  na.strings = "N/A")
+spod <- read.csv("../SPod/fhrdata_2017-11-30.csv", 
+                 header=TRUE,  na.strings = "N/A")
 spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
                                  format= "%m/%d/%Y %H:%M:%S")) 
 
-node <- "h"
-pidCol <- paste(node, "SPOD.PID..V.", sep=".")
-spodNode <- spod[, c("time", pidCol)]
-names(spodNode)[2] <- c("pid")
-spodNode <- subset(spodNode, !is.na(pid))
-spodNode$pid <- as.numeric(scale(spodNode$pid, center = TRUE))
 window_size <- 5000
 overlap <- 500
 max_iter <- 3
-lambda <- 2*window_size
-tau <- 0.1
+tau <- c(0.1, 0.2, 0.3)
 k <- 3
-spod_trends <- data.frame(time = spod$time, f=NA, g = NA, h = NA)
+spod_trends <- data.frame(time = spod$time)
 
 for (node in c("f", "g", "h")){
   pidCol <- paste(node, "SPOD.PID..V.", sep=".")
@@ -30,22 +25,31 @@ for (node in c("f", "g", "h")){
   names(spodNode)[2] <- c("pid")
   spodNode$pid <- as.numeric(scale(spodNode$pid, center = TRUE))
   result <- get_windows_BIC(spodNode$pid, tau, k, window_size, overlap,
-                          lambdaSeq = window_size^seq(1.2, 1.5, length.out=5), 
+                          lambdaSeq = window_size^seq(1.1, 1.5, length.out=10), 
                           df_tol = 1e-9, 
                           gamma = 1,
                           plot_lambda = FALSE, 
[…]

### 2019-01-04 Code/peaks_tests.R

@@ -1,68 +1,0 @@
-library(splines)
-library(fields)
-n <- 21000
-tau <- c(0.01, .05, .1)
-x <- seq(0.5, n, 1)/n
-df <- sample(5:20, 1)
-splineBasis <- ns(x, df=df)
-theta <- rexp(ncol(splineBasis), .5)
-baseline <- splineBasis%*%theta
-
-plot(splineBasis[,1], type="l")
-for(i in 1:ncol(splineBasis)){
-  lines(splineBasis[,i])
-}
-plot(baseline, type="l")
-
-numberOfPeaks <- sample(10:25, 1)
-peakCenters <- runif(numberOfPeaks)*.8+.1
-peakArea <- runif(numberOfPeaks)*.05
-peakWidths <- runif(numberOfPeaks)*.01 + 100/n
-
-peaks <- rep(0, length(baseline))
-if (numberOfPeaks > 0){
-  for (i in 1:numberOfPeaks){
-    peaks <- peaks + 
-      peakArea[i]*dnorm(x, mean=peakCenters[i], sd = peakWidths[i])
-  }
-}
-noise <- 0.25*rnorm(n)
-y <- peaks + baseline + noise
-plot(y~x, type="l", col="grey")
-lines((peaks+baseline)~x)
-lines((baseline+qnorm(.1, sd = .25))~x, col="red")
-
-df <- data.frame(y=y, x=x, baseline=baseline, peaks = peaks)
-
-
-trend <- matrix(NA, n, length(tau))
-for (j in 1:length(tau)){
-  fit_qsreg <- qsreg(df$x, df$y, maxit.cv = 50, alpha=tau[j], 
-                     hmin = -30)
-  trend[,j] <- predict(fit_qsreg)    
-}
-
-plot(y~x, type="l", col="grey")
-lines((baseline+qnorm(.1, sd = .25))~x, col="red")
-for (i in 1:length(tau)){
-  lines(trend[,i]~df$x, col="blue")
-}
-
-
-lam <- lambda_eBIC(df$y, tau, 3,  gamma = 1,
-                   lambdaSeq = n^seq(1, 1.4, length.out=15),
-                   plot_lambda = FALSE, single_lambda = FALSE)
-trend2 <- gurobi_trend(df$y, tau, lam$lambda, k=3)
-
-for (i in 1:length(tau)){
[…]

### 2019-01-04 Code/sim_misclass.R

@@ -1,3 +1,8 @@
+################################################################################
+# Calculate mis-classification rates in peaks simulation
+# Halley Brantley
+################################################################################
+
 rm(list=ls())
 simDesign <- "peaks"
 tau <- c(0.01, 0.05, 0.1)

### 2019-01-08 Manuscript/detrendify_JCGS.tex
introduction updates, timing experiments

@@
 	\label{sec:intro}
+
+	There have been a number of occasions when researchers have sought to estimate the trend in the quantiles of a random variable over time or another co-variate given noisy observations of that variable.
+\cite{nychka1995nonparametric} estimated smooth quantile curves of friction profiles of syringes in order to automate their quality assessment procedures.
+\cite{Ning2014} proposed a method for separating chromatogram peaks from the smooth baseline and noise present in the measurements.
+\cite{brantley2014mobile} examined air pollutant measurements collected using a monitoring vehicle and proposed using a smooth baseline trend as an estimate of regional background pollution.
+We are interested in the problem posed by \cite{thoma2016south} concerning low-cost air pollutant sensors.
+Researchers at the U.S.
+Environmental Protection Agency have built prototypes of low-cost sensors that provide an un-calibrated measure of volatile organic compounds (VOCs) and hazardous air pollutants (HAPs) at a 1 second time resolution.
+To reduce cost and power consumption, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled.
+A result the output signal exhibits a slowly varying baseline drift (minutes to hours).
+The purpose of these sensors is to protect the fenceline of industrial facilities by detecting whether VOC emissions from a local source are present and providing wind and concentration data to help identify the source and speed repairs.
 

### 2019-01-12 Manuscript/detrendify_JCGS.tex
figure updates, new classification metrics

@@
 
-	We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate for removing trends in time series.
-The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
-Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles.
-We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and proposed a modified criteria for choosing the smoothing parameter.
-We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.
-
-	\subsection{Quantile Regression}
-
-	The classic least squares regression is notoriously sensitive to outliers.
-One remedy to blunt the influence of outliers is to compute the least absolute deviations (LAD) solution in place of the least squares one.
-Given a design matrix $\mathbf{X} \in \mathbb{R}^{n \times p}$ and continuous responses $y \in \mathbb{R}^n$, we estimate a regression vector $\theta \in \mathbb{R}^p$ so that $\mathbf{X}\theta$ is a good approximation of $y$.
-The LAD estimator is a solution to the problem
-	\begin{eqnarray}
-		\underset{\theta}{\min}\; \frac{1}{n}\lVert y - \mathbf{X}\theta \rVert_1.
-	\end{eqnarray}
-	The above optimization problem generalizes the notion of the median of a collection of numbers. A median $\mu$ of $n$ reals $y_1, \ldots, y_n$ is the minimizer of the function
-	\begin{eqnarray}
-		f(u) & = & \frac{1}{n}\sum_{i=1}^n \lvert y_i - \theta \rvert.
-	\end{eqnarray}
-	Recall that the median is the 50th percentile or 0.5-quantile, namely half of the $y_i$ are less than or equal to $\mu$ and the other half is greater than or equal to $\mu$.
-The median can be generalized to arbitrary $\tau$-quantiles for $\tau \in (0,1)$ to give us quantile regression \citep{Koenker1978}.
-
-	First define the so-called "check function"
-	\begin{eqnarray}
-		\rho_\tau(\Delta) & = & \begin{cases}
-			\tau \Delta & \Delta \geq 0 \\
-			-(1-\tau)\Delta & \Delta < 0 \\
-		\end{cases}
-	\end{eqnarray}
-	Then the $\tau$th quantile of the $y_i$ is a minimizer of the function
-	\begin{eqnarray}
-		f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(y_i - \theta).
-	\end{eqnarray}
-	Returning to the regression context, we can generalize LAD regression to quantile regression, namely computing the minimizer of the function
-	\begin{eqnarray}
-		f_\tau(\theta) & = & \frac{1}{n}\sum_{i=1}^n \rho_\tau(y_i - \langle x_i \mid \theta \rangle),
-	\end{eqnarray}
-	where $x_i \in \mathbb{R}^p$ denotes the $i$th row of $\mathbf{X}$.
-
-	%% ----------------------------------------------------------------------
-	%% Trend Filtering
-	%% ----------------------------------------------------------------------
-	\subsection{Trend Filtering}
+	A variety of approaches have been proposed for estimating quantile trends.
+\cite{Koenker1978} was the first to propose the use of the check loss function to estimate conditional quantiles.
+Given a design matrix $\mathbf{X} \in \mathbb{R}^{n \times p}$ and continuous responses $y \in \mathbb{R}^n$, we estimate a regression vector $\theta \in \mathbb{R}^p$ so that $\mathbf{X}\theta$ is a good approximation of the $\tau$th quantile of $y$ conditional on $x$.
+The regression vector $\theta$ is estimated by finding the minimizer of $\rho_{\tau}(y-X\theta)$ where $\rho_{\tau}(y) = \sum_{i=1}^n y_i(1-\mathbf{I}(y_i<0))$ is the check-loss function and $\mathbf{I}$ is the indicator function.
 
@@
 
-	%% ----------------------------------------------------------------------
-	%% Quantile Trend Filtering
-	%% ----------------------------------------------------------------------
+	We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate for removing trends in time series.
+The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
+Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles.
+We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and proposed a modified criteria for choosing the smoothing parameter.
+We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.
+
 	\section{Methods}
@@
 We propose an alternating direction method of multipliers (ADMM) algorithm for solving large problems in a piecewise fashion.
-The ADMM algorithm, described by \cite{boyd2011distributed}, relies on the idea of dual ascent.
-If we consider the equality optimization problem
-	\begin{align}
-	&\text{minimize } f(x)\\
-	&\text{subject to } Ax = b
-	\end{align}
-	with variable $x\in \mathbf{R}^n$, where $A \in \mathbf{R}^{mxn}$ and $f: \mathbf{R}^n \rightarrow \mathbf{R}$ is convex. The corresponding Lagrangian function is 
-	\begin{equation}
-	L(x,\omega) = f(x) + \omega^T(Ax-b)
-	\end{equation}
-	and the corresponding dual problem  is 
-	\begin{equation}
-	\text{maximize } \inf_x L(x,\omega)
-	\end{equation}
-	Assuming that strong duality holds (Slater's constraint: there exists a solution in the interior of the domain), the optimal values of the primal and dual variables are the same.
-The dual ascent method, uses gradient ascent to solve the dual problem and thus the primal problem using the iterates
-	\begin{align}
-	&x^{k+1} = \arg\min L(x, \omega^k)\\
-	&\omega^{k+1} = \omega^k + \alpha^k(Ax^{k+1} - b)
-	\end{align}
-	where $\alpha^k > 0$ is a step size and $k$ is the iteration counter.
-In many cases (including our problem), the $x$ update fails because the problem is unbounded below.
-To remedy this issue the augmented Lagrangian was introduced:
-	\begin{equation}
-	L_{\gamma}(x,\omega) = f(x)+\omega^T(Ax-b) + \frac{\gamma}{2}||Ax-b||_2^2
-	\end{equation}
-	where $\gamma > 0$ is called the penalty parameter. Applying dual ascent to the modified problem results in the method of multipliers algorithm which converges under more general conditions:
-	\begin{align}
-	&x^{k+1} = \arg\min L_\gamma(x, \omega^k)\\
-	&\omega^{k+1} = \omega^k + \gamma(Ax^{k+1} - b)
-	\end{align}
-	Consider a new optimization problem with a separable objective function. 
-	\begin{align}
-	&\text{minimize } f_1(x_1) + f_2(x_2)\\
-	&\text{subject to } A_1x_1 + A_2x_2 = b
-	\end{align}
-	where $x = (x_1, x_2)$, $i \in \{1,2\}$, and  $x_i \in \mathbf{R}^{n_i}$. The dual ascent updates can be expressed as 
-	\begin{align}
-	&x_i^{k+1} = \arg\min L_i(x_i, \omega^k)\\
-	&\omega^{k+1} = \omega^k + \alpha^k(Ax^{k+1} - b)
-	\end{align}
-	however we lose the separability if the augmented Lagrangian is used instead, i.e.
-the augmented Lagrangian can not be separated into functions of $x_i$.
-The Alternating Direction Method of Multipliers (ADMM) addresses this problem and maintains separability while using the augmented Lagrangian to improve convergence through the following updates
-	\begin{align}
-	&x_1^{k+1} = \arg\min_{x_1} L_{\gamma}(x_1, x_2^k, \omega^k)\\
-	&x_2^{k+1} = \arg\min_{x_2} L_{\gamma}(x_1^{k+1}, x_2, \omega^k)\\
-	&\omega^{k+1} = \omega^k + \alpha^k(A_1x_1^{k+1} + A_2x_2^{k+1} - b)
-	\end{align}
-
-	We now consider the quantile regression trend filtering problem,  
+The ADMM algorithm is fully described by \cite{boyd2011distributed}.
+We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem,
 	\begin{equation}
@@
 
-	\begin{figure}[!h] 
-		\caption{Windows fit separately compared to simultaneous fit, signal present.}
-		\includegraphics[width = \linewidth]{Figures/overlapping_windows2.png}
-	\end{figure}
+	We conducted timing experiments to show the advantages of using our ADMM algorithm on large datasets.
+We use 2 windows for each data size with an overlap of 2000 and two iterations of the algorithm after the initial estimates.
 
 	\begin{figure}[!h] 
-		\caption{Windows fit with 300 iterations of ADMM, signal present.}
-		\includegraphics[width = \linewidth]{Figures/admm_windows2.png}
+		\caption{Timing experiments comparing quantile trend filtering on a single window versus 2 windows by data size.}
+		\includegraphics[width = \linewidth]{Figures/Fig_timing_experiment.png}
 	\end{figure}
@@
 
-	\begin{figure}[h!]
-		\caption{RMSEs compared to the simulated baseline function.}
-		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}
-	\end{figure}
-
 	Our detrend\_BIC method performs the best overall in terms of both RMSE and miss-classification rate.
@@
 		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
+	\end{figure}
+
+	\begin{figure}[h!]
+		\caption{Class averaged accuracy by threshold, data size, and method (1 is best 0.5 is worst).}
+		\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
+	\end{figure}
+
+	\begin{figure}[h!]
+		\caption{F1 score by threshold, data size, and method (1 is best 0 is worst).}
+		\includegraphics[width = \linewidth]{Figures/peaks_F1.png}
+	\end{figure}
+
+	\begin{figure}[h!]
+		\caption{Precision by threshold, data size, and method (true positive over true positives + false positives).}
+		\includegraphics[width = \linewidth]{Figures/peaks_precision.png}
+	\end{figure}
+
+	\begin{figure}[h!]
+		\caption{Recall by threshold, data size, and method (true positive over true positives + false negatives).}
+		\includegraphics[width = \linewidth]{Figures/peaks_recall.png}
 	\end{figure}
@@
 
-	\input{application_missclass.tex}	
-	\input{application_signal.tex}
+	\input{application_metrics.tex}
+
+	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
+	\includegraphics[width = \linewidth]{Figures/signal_plot.png}
 

### 2019-01-12 Code/fig_timing.R

@@ -1,17 +1,29 @@
 rm(list=ls())
 load("../TimingData/window_times.RData")
+all.times0 <- all.times
+# load("../TimingData/window_times_2.RData")
+# all.times <- rbind(all.times, all.times0)
 windows.df <- data.frame(n = data_lengths,
                       median = apply(all.times, 2, median)*1e-9,
                       mean = apply(all.times, 2, mean)*1e-9)
 
 load("../TimingData/single_window_times.RData")
+all.times0 <- all.times
+load("../TimingData/single_window_times_2.RData")
+all.times0 <- rbind(all.times, all.times0)
+# load("../TimingData/single_window_times_3.RData")
+# all.times <- rbind(all.times, all.times0)
+
 data_lengths <- seq(2000, 50000, 2000)
 time.df <- data.frame(n = data_lengths, 
-                      median = apply(all.times, 2, median)*1e-9,
-                      mean = apply(all.times, 2, mean)*1e-9)
+                      median = apply(all.times, 2, median, na.rm=T)*1e-9,
+                      mean = apply(all.times, 2, mean, na.rm=T)*1e-9)
+
 ggplot(time.df, aes(x=n, y=median)) + 
   geom_line() + 
   geom_point() +
   geom_line(data = windows.df, col="red") +
-  geom_point(data = windows.df, col="red") 
-
+  geom_point(data = windows.df, col="red") +
+  theme_bw() +
+  labs(y = "Time (s)", x = "Data Size")
+ggsave("../Manuscript/Figures/Fig_timing_experiment.png", width = 7, height = 2)

### 2019-01-12 Code/simulation_figs.R

@@ -4,14 +4,13 @@ library(jcolors)
 rm(list=ls())
 source("sim_generating_functions.R")
 load_all("detrendr")
+colPal <- rev(c('#762a83','#9970ab','#c2a5cf',
+                '#a6dba0','#5aae61','#1b7837'))
 
 tau <- c(0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99)
-n <- 300
 nSim <- 100
 simDesigns <- c( "mixednorm", "shapebeta", "gaus")
-methods <- c("detrend_eBIC", "detrend_SIC", "detrend_valid", "npqw", "qsreg", "rqss")
-#methods <- "npqw"
-#simDesigns <- "mixednorm"
+methods <- c("detrend_eBIC", "detrend_SIC", "detrend_valid", "qsreg", "rqss", "npqw")
 MSEs <- as.data.frame(matrix(NA, nrow = nSim*length(methods)*length(simDesigns), 
                              ncol = length(tau)+4))
 colnames(MSEs) <- c("Design", "Sim", "Method", "n", paste0("tau_", tau))
@@ -35,12 +34,6 @@ for(simDesign in simDesigns){
   }
 }
 
-tmp <- MSEs %>% filter(Design == "gaus", n==300)
-
-hist(tmp[tmp$Method == "qsreg", "tau_0.5"])
-median(tmp[tmp$Method == "npqw", "tau_0.5"])
-
-which(tmp[tmp$Method == "npqw", "tau_0.5"] >.1)
 
 
 MSEs_long <- MSEs %>% gather("tau", "MSE", -c("Design", "Sim", "Method", "n")) 
@@ -56,7 +49,8 @@ summary_stats <-
   ) %>%
   ungroup() %>%
   mutate(tau_fac = tau, 
-         tau = as.numeric(substr(tau_fac, 5, 10)))
+         tau = as.numeric(substr(tau_fac, 5, 10)), 
+         Method = factor(Method, levels = methods))
 
 
 
@@ -70,7 +64,7 @@ summary_stats %>%
                  position = position_dodge(width = 0.5))+
   facet_grid(.~factor(tau), scales = "free")+
   theme_bw() +
-  scale_color_brewer(palette = "Paired") +
+  scale_color_manual(values=colPal, breaks = methods) +
   labs(x = "n", y="RMSE", title = "Gaussian")+
   ylim(c(0,.153))
 ggsave("../Manuscript/Figures/gaus_mse.png", width = 10, height = 3)
[…]

### 2019-01-17 Manuscript/detrendify_JCGS.tex
manuscript updates, replaced application metrics with contingency table

@@
 
-	There have been a number of occasions when researchers have sought to estimate the trend in the quantiles of a random variable over time or another co-variate given noisy observations of that variable.
-\cite{nychka1995nonparametric} estimated smooth quantile curves of friction profiles of syringes in order to automate their quality assessment procedures.
-\cite{Ning2014} proposed a method for separating chromatogram peaks from the smooth baseline and noise present in the measurements.
-\cite{brantley2014mobile} examined air pollutant measurements collected using a monitoring vehicle and proposed using a smooth baseline trend as an estimate of regional background pollution.
 	We are interested in the problem posed by \cite{thoma2016south} concerning low-cost air pollutant sensors.
@@
 The purpose of these sensors is to protect the fenceline of industrial facilities by detecting whether VOC emissions from a local source are present and providing wind and concentration data to help identify the source and speed repairs.
+
+	There have been a number of occasions when researchers have sought to estimate the trend in the quantiles of a random variable over time or another co-variate given noisy observations of that variable.
+\cite{nychka1995nonparametric} estimated smooth quantile curves of friction profiles of syringes in order to automate their quality assessment procedures.
+\cite{Ning2014} proposed a method for separating chromatogram peaks from the smooth baseline and noise present in the measurements.
+\cite{brantley2014mobile} examined air pollutant measurements collected using a monitoring vehicle and proposed using a smooth baseline trend as an estimate of regional background pollution.
 
@@
 	where $\lambda$ is a non-negative tuning parameter. As with the classic quantile regression, the quantile trend filtering problem can be solved by a linear program.
-
-	\subsection{Smoothing parameter choice}
-	Our method can easily handle missing data by changing the check loss function to output 0 for missing values.
-This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data.
-A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.
-\cite{KoenkerNgPortnoy1994} relate $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
-	\begin{equation}
-	SIC(p_{\lambda}) = log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}log n
-	\end{equation}
-
-	The traditional Bayesian Information Criterion (BIC) is given by 
-	\begin{equation}
-	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n 
-	\end{equation}	
-	where $\theta(s)$ is the parameter $\theta$ with those components outside $s$ being set to 0, and $\nu(s)$ is the number of components in $s$.
-If we assume an asymmetric Laplace likelihood $L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_i\rho_\tau(\frac{y_i - \theta_i}{\sigma})\right\}$ and the number of non-zero elements of $D\theta$ as $df$
-	\begin{equation}
-	\mbox{BIC}(df) = 2\sum_i\frac{1}{\sigma}\rho_{\tau}(y_i-\theta_i) + df\log n
-	\end{equation} 
-	We can choose and $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates. \cite{chen2008} proposed the extended BIC for large parameter spaces 
-	\begin{equation}
-	BIC_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j,}~~\gamma \in [0,1]
-	\end{equation}
-	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model.
-We used this criteria with $\gamma = 1$, $P=n-k$ where $k$ is the order of the differencing matrix and $j = \nu(s)$ is the number of non-zero entries in $D^{(k)}\theta$.
-
-	We used a single dataset to illustrate the difference between the scaled, unscaled and extended BIC criteria.
-
-	\begin{figure}[h!]
-	\caption{Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.} 
-	\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
-	\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
-	\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
-	\end{figure}
 
@@
 
-	\section{Simulation Study}
+	\subsection{Smoothing parameter choice}
+	Our method can easily handle missing data by changing the check loss function to output 0 for missing values.
+This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data.
+A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.
+\cite{KoenkerNgPortnoy1994} relate $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
+	\begin{equation}
+	\mbox{SIC}(p_{\lambda}) = \log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}\log n
+	\end{equation}
+
+	The traditional Bayesian Information Criterion (BIC) is given by 
+	\begin{equation}
+	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n 
+	\end{equation}	
+	where $\theta(s)$ is the parameter $\theta$ with those components outside $s$ being set to 0, and $\nu(s)$ is the number of components in $s$.
+If we assume an asymmetric Laplace likelihood $L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_i\rho_\tau(\frac{y_i - \theta_i}{\sigma})\right\}$ and the number of non-zero elements of $D\theta$ as $df$
+	\begin{equation}
+	\mbox{BIC}(df) = 2\sum_i\frac{1}{\sigma}\rho_{\tau}(y_i-\theta_i) + \mbox{df}\log n
+	\end{equation} 
+	We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates. \cite{chen2008} proposed the extended BIC for large parameter spaces 
+	\begin{equation}
+	\mbox{BIC}_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j,}~~\gamma \in [0,1]
+	\end{equation}
+	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model.
+We used this criteria with $\gamma = 1$, $P=n-k$ where $k$ is the order of the differencing matrix and $j = \nu(s)$ is the number of non-zero entries in $D^{(k)}\theta$.
+
+	We used a single dataset to illustrate the difference between the scaled, unscaled and extended BIC criteria.
+
+	\begin{figure}[h!]
+		\caption{Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.} 
+		\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
+		\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
+		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
+	\end{figure}
+
+	\section{Simulation Studies}
+
+	\subsection{Estimating Quantiles}
 
@@
 		\caption{Simulated data with true quantiles $\tau \in \{0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99\}$}	
-		\includegraphics[width=.3\linewidth]{Figures/gaus.png}
+		\includegraphics[width=.28\linewidth]{Figures/gaus.png}
 		\includegraphics[width=.3\linewidth]{Figures/shapebeta.png}
-		\includegraphics[width=.3\linewidth]{Figures/mixednorm.png}
+		\includegraphics[width=.45\linewidth]{Figures/mixednorm.png}
 	\end{figure}
@@
 
+	\begin{figure}
+		\caption{RMSE by design, method, quantile and data size.}
 		\includegraphics[width=\linewidth]{Figures/gaus_mse.png}	
-
 		\includegraphics[width=\linewidth]{Figures/shapebeta_mse.png}
-
 		\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
-
+	\end{figure}
 	In all of the designs the proposed detrend methods are either better than or comparable to existing methods.
@@
 
-	\section{Peaks Simulation}
+	\subsection{Peak Detection}
 	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
 	\begin{equation}
-	Y(t) = s(t) + b(t) + \epsilon
+	Y(t) = s(t) + b(t) + \epsilon(t)
 	\end{equation} 
-	where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon \sim N(0, \sigma^2)$ is an error component.
-We assume $t$ is a uniformly spaced sequence between 0 and 1.
+	where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, \sigma^2)$ is an error component.
+We assume $t$ is a uniformly spaced sequence of time points between 0 and 1.
 We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from $n/50$ to $n/25$ with equal probability, and coefficients drawn from an exponential distribution with rate 1.
@@
 One hundred datasets were generated for $n=\{300, 500, 1000, 5000\}$.
-We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE.
+We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the root mean squared error (RMSE).
 
@@
 	In addition to the RMSE we also calculated the signal miss-classification rate.
-We first classified the "true" peaks function into signal or not based on a threshold of 0.1.
+We first classified the ``true" peaks function into signal or not based on a threshold of 0.1.
 We then classified the detrended data using three different thresholds and calculated the fraction miss-classified.
-An illustration of the observations classified as signal after detrending compared to the "true signal" is shown in Fig.
+An illustration of the observations classified as signal after detrending compared to the ``true signal" is shown in Fig.
 9.
@@
 		\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
-	\end{figure}
-
-	\begin{figure}[h!]
-		\caption{F1 score by threshold, data size, and method (1 is best 0 is worst).}
-		\includegraphics[width = \linewidth]{Figures/peaks_F1.png}
-	\end{figure}
-
-	\begin{figure}[h!]
-		\caption{Precision by threshold, data size, and method (true positive over true positives + false positives).}
-		\includegraphics[width = \linewidth]{Figures/peaks_precision.png}
-	\end{figure}
-
-	\begin{figure}[h!]
-		\caption{Recall by threshold, data size, and method (true positive over true positives + false negatives).}
-		\includegraphics[width = \linewidth]{Figures/peaks_recall.png}
-	\end{figure}
-
-	\begin{figure}[h!]
-		\caption{Miss-classification rates by threshold, data size, and method, values above the upper limit (npqw) not shown.}
-		\includegraphics[width = \linewidth]{Figures/peaks_missclass.png}
 	\end{figure}
@@
 
-	\input{application_metrics.tex}
+	\input{complete_confusion.tex}
 
 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
-	\includegraphics[width = \linewidth]{Figures/signal_plot.png}
 

### 2019-01-17 Code/Fig_sim_misclass.R

@@ -18,34 +18,34 @@ thresh <- c(.9, 1, 1.1, 1.2)
 metric_types <- c("missclass", "precision", "recall", "CAA", "F1")
 k <- 1
 
-for (n in c(500,1000,2000,4000)){
-  for (i in 1:nSim){
-    load(sprintf("../SimData/%s_n_%i_sim%03.0f.RData", "peaks", n, i))
-    df$signal <- as.numeric(df$peaks > 0.5)
-
-    for (method in methods){
-      load(sprintf("../SimResults/%s/%s_n_%i_sim%03.0f.RData",
-                   method, "peaks", n, i))
-      for (threshold in thresh){
-        for (metric in metric_types){
-          for (j in 1:length(tau)){
-            metrics$sim[k] <- i
-            metrics$method[k] <- method
-            metrics$n[k] <- n
-            metrics$threshold[k] <- threshold
-            metrics$tau[k] <- tau[j]
-            metrics$metric_type[k] <- metric
-            metrics$metric[k] <- get_metric(trend[,j], df$y, df$signal,
-                                         threshold, metric)
-            k <- k+1
-            metrics[k, ] <- NA
-          }
-        }
-      }
-    }
-    save(metrics, file = "../SimResults/peak_metrics.RData")
-  }
-}
+# for (n in c(500,1000,2000,4000)){
+#   for (i in 1:nSim){
+#     load(sprintf("../SimData/%s_n_%i_sim%03.0f.RData", "peaks", n, i))
+#     df$signal <- as.numeric(df$peaks > 0.5)
+# 
+#     for (method in methods){
+#       load(sprintf("../SimResults/%s/%s_n_%i_sim%03.0f.RData",
+#                    method, "peaks", n, i))
+#       for (threshold in thresh){
+#         for (metric in metric_types){
+#           for (j in 1:length(tau)){
+#             metrics$sim[k] <- i
+#             metrics$method[k] <- method
+#             metrics$n[k] <- n
+#             metrics$threshold[k] <- threshold
+#             metrics$tau[k] <- tau[j]
+#             metrics$metric_type[k] <- metric
+#             metrics$metric[k] <- get_metric(trend[,j], df$y, df$signal,
+#                                          threshold, metric)
+#             k <- k+1
+#             metrics[k, ] <- NA
+#           }
+#         }
+#       }
+#     }
[…]

### 2019-01-17 Code/application.R

@@ -5,6 +5,7 @@
 library(tidyverse)
 library(devtools)
 library(Hmisc)
+library(caret)
 load_all("detrendr")
 rm(list=ls())
 spod <- read.csv("../SPod/fhrdata_2017-11-30.csv", 
@@ -15,7 +16,7 @@ spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp),
 window_size <- 5000
 overlap <- 500
 max_iter <- 3
-tau <- c(0.1, 0.2, 0.3)
+tau <- c(0.05, 0.1, 0.2)
 k <- 3
 # spod_trends <- data.frame(time = spod$time)
 
@@ -48,10 +49,12 @@ spodPeaks <- spodPIDs - select(spod_trends, contains("0.1"))
 spodPeaks$time <- spod_trends$time  
 spodPIDs$time <- spod_trends$time
 
-spodRaw <- spodPIDs %>% gather("node", "PID", -time)
+spodRaw <- spod[, c("time", paste(nodes, "SPOD.PID..V.", sep="."))] %>% 
+  gather("node", "PID", -time)
 ggplot(spodRaw, aes(x=time, y=PID, col=node)) + 
   geom_line(alpha=0.5) + 
-  theme_bw() 
+  theme_bw() + 
+  scale_color_brewer(palette = "Set1", labels = nodes)
 ggsave("../Manuscript/Figures/uncorrected_data.png", width = 7, height = 3)
 
 spodLong <- spodPeaks %>% gather("node","PID", -time)
@@ -62,7 +65,8 @@ PID_thresh <- data.frame(node = names(thresholds), thresh = thresholds)
 ggplot(spodLong, aes(x=time, y=PID, col=node)) + 
   geom_line(alpha=0.5) + 
   geom_hline(data=PID_thresh, aes(yintercept=thresh, col=node)) +
-  theme_bw() 
+  theme_bw() + 
+  scale_color_brewer(palette = "Set1", labels = nodes)
 ggsave("../Manuscript/Figures/corrected_data.png", width = 7, height = 3)
 
 ################################################################################
@@ -77,51 +81,26 @@ get_spod_signal <- function(tau, spod_trends, spodPIDs, thresholds){
 }
 
 spod_signal <- get_spod_signal(0.1, spod_trends, spodPIDs, PID_thresh$thresh)
-get_metric(trend[,j], df$y, df$signal,
-           threshold, metric)
[…]

### 2019-01-17 Code/fig_timing.R

@@ -14,6 +14,15 @@ all.times <- as.data.frame(t(all.times))
 all.times$n <- data_lengths
 times.long <- all.times %>% gather("rep", "time", -n)
 times.long$time <- times.long$time*1e-9
+load("../TimingData/single_window_times.RData")
+all.times <- as.data.frame(t(all.times))
+all.times$n <- data_lengths
+times.long2 <- all.times %>% gather("rep", "time", -n)
+times.long2$time <- times.long2$time*1e-9
+
+time.df <- bind_rows(times.long, times.long2) %>% 
+  group_by(n) %>%
+  summarise(time = mean(time, na.rm=T))
 
 time.df <- data.frame(n = data_lengths, 
                       median = apply(all.times, 2, median, na.rm=T)*1e-9,

### 2019-01-18 Code/detrendr/R/get_trend_windows.R
manuscript updates, application comparison

@@ -43,9 +43,8 @@ get_trend_windows <- function(y, tau, lambda, k, rho=1, window_size,
     # First estimate uses LP not QP
     solver <- "lpSolve"
   }
-  if (!is.integer(window_size)){
-    stop("window_size must be an integer")
-  }
+  
+  window_size <- round(window_size)
   
   y_n <- length(y)
   tau <- sort(tau)

### 2019-01-18 Code/detrendr/R/window_updates.R

@@ -171,6 +171,7 @@ get_model_abs <- function(model, z, rho, nT){
 get_eta <- function(phi, y, k){
   D <- get_Dk(length(y), k)
   eta <- D%*%y - D%*%phi
+  eta[is.na(eta)] <- 0
   return(as.matrix(eta))
 }
 

### 2019-01-18 Code/timing.R

@@ -5,29 +5,46 @@ library(microbenchmark)
 load_all("detrendr")
 rm(list=ls())
 source("sim_generating_functions.R")
-set.seed(12345678)
 tau <- c(0.05, 0.1)
 simDesign <- "peaks"
-df <- generate_peaks_design(100000)
+
 
 plot(y~x, df, type="l")
 lines((peaks+baseline)~x, df, col="blue")
 lines((baseline)~x, df, col="red")
 
 
-times <- 5
-data_lengths <- seq(55000, 65000, 5000)
-all.times <- matrix(NA, ncol = length(data_lengths), nrow = times)
+times <- 3
+overlap <- 1000
+data_lengths <- seq(50000, 65000, 5000)
+single.times <- matrix(NA, ncol = length(data_lengths), nrow = times)
+window2.times <- matrix(NA, ncol = length(data_lengths), nrow = times)
+window3.times <- matrix(NA, ncol = length(data_lengths), nrow = times)
 i <- 1
 for (n in data_lengths){
   print(sprintf("n = %i", n))
-  gc()
-  all.times[,i] <- microbenchmark(
-    get_trend(df$y[1:n], tau, lambda = c(n,n), k=3), 
+  df <- generate_peaks_design(n)
+  single.times[,i] <- microbenchmark(
+    trend <- get_trend(df$y, tau, lambda = c(n,n), k=3), 
+    times = times)$time
+  window2.times[,i] <- microbenchmark(
+    trend_w <- get_trend_windows(df$y[1:n], tau, lambda = c(n,n), k=3, 
+                                 window_size = as.integer(n/2+overlap/2), 
+                                 overlap = overlap, max_iter = 4, quad = TRUE, 
+                                 use_gurobi = TRUE), 
+    times = times)$time
+  window3.times[,i] <- microbenchmark(
+    trend_w3 <- get_trend_windows(df$y[1:n], tau, lambda = c(n,n), k=3, 
+                                 window_size = round(n/3+2*overlap/3) + 1, 
+                                 overlap = overlap, max_iter = 4, quad = TRUE, 
+                                 use_gurobi = TRUE), 
     times = times)$time
   i <- i+1
-  save(all.times, data_lengths, 
-       file="../TimingData/single_window_times.RData")
+  
+  save(df, trend, trend_w, trend_w3, 
+       file = sprintf("../TimingData/trends_%i.RData", n))
+  save(single.times, window2.times, window3.times, data_lengths, 
[…]

### 2019-01-18 Code/timing_windows.R

@@ -16,7 +16,7 @@ lines((baseline)~x, df, col="red")
 
 
 times <- 1
-data_lengths <- seq(15000, 55000, 10000)
+data_lengths <- seq(60000, 65000, 5000)
 all.times <- matrix(NA, ncol = length(data_lengths), nrow = times)
 trend_diff <- list()
 overlap <- 2000
@@ -35,7 +35,7 @@ for (n in data_lengths){
   trend_diff[[i]] <- trend_w - trend
   i <- i+1
   save(all.times, data_lengths, trend_diff,
-       file="../TimingData/stopping_crit_times.RData")
+       file="../TimingData/stopping_crit_times_2.RData")
 }
 
 # load("../TimingData/window_times.RData")

### 2019-01-21 Manuscript/detrendify_JCGS.tex

@@
 
-	We combine the ideas of quantile regression and trend filtering, namely consider the signal approximation problem, where the design $\mathbf{X}$ is the identity matrix.
-
-	The estimation of the quantile trend filtering model can be posed as the following optimization problem.
+	We combine the ideas of quantile regression and trend filtering, namely consider the case where the design $\mathbf{X}$ is the identity matrix.
+For a single quantile level $\tau$ the estimation of the quantile trend filtering model can be posed as the following optimization problem.
 	\begin{eqnarray}
@@
 	\end{eqnarray}
-	where $\lambda$ is a non-negative tuning parameter. As with the classic quantile regression, the quantile trend filtering problem can be solved by a linear program.
+	where $\lambda$ is a non-negative tuning parameter.
+As with the classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of free or commercial solvers.
+In many cases, including ours, we are interested in estimating multiple quantiles simultaneously.
+We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$.
+Given quantiles $\{\tau_1, ..., \tau_J\}$ such that $\tau_1 < \tau_2 < ...
+< \tau_J$, the optimization problem becomes
 
-	\subsection{ADMM for Smoothing Windows}
+	\begin{eqnarray}
+	\label{eq:quantile_trend}
+	\underset{\theta}{\min}\; \sum_{j=1}^J \left [\frac{1}{n} \sum_{i=1}^n \rho_{\tau_j}(y_i - \theta_{ji}) + 
+	\lambda_j \lVert \mathbf{D}^{(k)} \theta_j \rVert_1 \right ] \\
+	 \text{subject to: }\; \theta_{1i} \le \theta_{2i} \le ... \le \theta_{Ji} \text{ for all } i,
+	\end{eqnarray}
+
+	where $\theta = (\theta_1, ..., \theta_J)$ and $\theta_j \in \mathcal{R}^n$.
+The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
+
+	The number of parameters to be estimates is equal to the number of observations multiplied by the number of quantiles of interest.
+As the size of the data and the number of quantiles grows, all solvers will eventually break.
+
+	\subsection{ADMM for Big Data}
 
@@
 We propose an alternating direction method of multipliers (ADMM) algorithm for solving large problems in a piecewise fashion.
-The ADMM algorithm is fully described by \cite{boyd2011distributed}.
+The ADMM algorithm is fully described by \cite{boyd2011distributed, gabay1975dual, glowinski1975approximation}.
 We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem,

### 2019-01-22 Code/Fig_app_short.R
application and timing edits

@@ -6,49 +6,27 @@ library(caret)
 library(zoo)
 library(aricode)
 load_all("detrendr")
+source("application_functions.R")
 rm(list=ls())
 load("../SPod/trends_short.RData")
 
 spodPeaks <- select(spodPIDs, -time) - select(detrendr_trends,
-                                              contains(paste(0.05)))
-plot(spodPeaks$f, type="l")
+                                              contains(paste(0.15)))
+plot(spodPeaks$h, type="l")
 thresholds <- apply(spodPeaks, 2, 
-                      function(x) median(x) + 2.5*mean(abs(x-median(x))))
-abline(h=thresholds[1], col="red")
+                      function(x) median(x, na.rm=T) + 
+                      4*median(abs(x-median(x, na.rm=T)), na.rm=T))
+abline(h=thresholds[3], col="red")
+plot(g~f, spodPeaks)
 
 
 ################################################################################
-get_spod_signal <- function(tau, spod_trends, spodPIDs){
-  spodPeaks <- select(spodPIDs, -time) - select(spod_trends, contains(paste(tau)))
-  thresholds <- apply(spodPeaks, 2, 
-                  function(x) median(x) + 3*mean(abs(x-median(x))))
-  spodSignal <- spodPeaks
-  for (i in 1:length(thresholds)){
-    spodSignal[,i] <- as.numeric(spodPeaks[,i]>thresholds[i])
-  }
-  return(spodSignal)
-}
-
-get_confusion <- function(spod_signal){
-  spod_signal <- na.omit(spod_signal)
-  for (i in 1:ncol(spod_signal)){
-    spod_signal[,i] <- factor(spod_signal[,i])
-  }
-  mat_h1 <- confusionMatrix(spod_signal$f[spod_signal$h==1],
-                            spod_signal$g[spod_signal$h==1])
-
-  mat_h0 <- confusionMatrix(spod_signal$f[spod_signal$h==0],
-                            spod_signal$g[spod_signal$h==0])
-
-  conf_out <- cbind(mat_h0$table, mat_h1$table)
-  return(conf_out)
-}
-
+crit <- 5
 NMI <- {}
 conf <- {}
 for (j in 1:length(tau)){
-  detrendr_signal <- get_spod_signal(tau[j], detrendr_trends, spodPIDs)
-  qsreg_signal <- get_spod_signal(tau[j], qsreg_trends, spodPIDs)
[…]

### 2019-01-22 Code/Fig_application.R

@@ -7,6 +7,7 @@ library(devtools)
 library(Hmisc)
 library(caret)
 load_all("detrendr")
+source("application_functions.R")
 rm(list=ls())
 spod <- read.csv("../SPod/fhrdata_2017-11-30.csv", 
                  header=TRUE,  na.strings = "N/A")
@@ -14,6 +15,7 @@ spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp),
                                  format= "%m/%d/%Y %H:%M:%S")) 
 
 load("../SPod/spod_trends.RData")
+
 nodes <- c("f", "g", "h")
 spodPIDs <- as.data.frame(spod[, paste(nodes, "SPOD.PID..V.", sep=".")]/1000)
 names(spodPIDs) <- nodes
@@ -21,54 +23,102 @@ spodPeaks <- spodPIDs - select(spod_trends, contains("0.1"))
 spodPeaks$time <- spod_trends$time
 spodPIDs$time <- spod_trends$time
 
-spodRaw <- spod[, c("time", paste(nodes, "SPOD.PID..V.", sep="."))] %>%
+spodRaw <- spodPIDs %>%
   gather("node", "PID", -time)
 ggplot(spodRaw, aes(x=time, y=PID, col=node)) +
   geom_line(alpha=0.5) +
   theme_bw() +
   scale_color_brewer(palette = "Set1", labels = nodes)
-ggsave("../Manuscript/Figures/uncorrected_data.png", width = 7, height = 3)
+ggsave("../Manuscript/Figures/uncorrected_data.png", width = 7, height = 2.5)
 
 spodLong <- spodPeaks %>% gather("node","PID", -time)
-thresholds <- 4*apply(spodPeaks[1:10000,1:3], 2, sd) +
-  apply(spodPeaks[1:10000,1:3], 2, mean)
-PID_thresh <- data.frame(node = names(thresholds), thresh = thresholds)
 
 ggplot(spodLong, aes(x=time, y=PID, col=node)) +
   geom_line(alpha=0.5) +
-  geom_hline(data=PID_thresh, aes(yintercept=thresh, col=node)) +
   theme_bw() +
   scale_color_brewer(palette = "Set1", labels = nodes)
-ggsave("../Manuscript/Figures/corrected_data.png", width = 7, height = 3)
+ggsave("../Manuscript/Figures/corrected_data.png", width = 7, height = 2.5)
 
 # ################################################################################
 
-get_spod_signal <- function(tau, spod_trends, spodPIDs, thresholds){
[…]

### 2019-01-22 Code/application_short.R

@@ -17,16 +17,16 @@ spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp),
 
 load("../SPod/spod_trends.RData")
 nodes <- c("f", "g", "h")
-spodPIDs <- as.data.frame(spod[, paste(nodes, "SPOD.PID..V.", sep=".")]/1000)
-names(spodPIDs) <- nodes
-spodPIDs$time <- spod$time
+spodPID <- as.data.frame(spod[, paste(nodes, "SPOD.PID..V.", sep=".")]/1000)
+names(spodPID) <- nodes
+spodPID$time <- spod$time
 
-spodPIDs <- spodPIDs %>% 
-  filter(time > as.POSIXct("2017-11-30 10:15:00"),
-         time <= as.POSIXct("2017-11-30 10:15:00")+8000)
+spodPIDs <- spodPID %>% 
+  filter(time > as.POSIXct("2017-11-30 9:30:00"),
+         time <= as.POSIXct("2017-11-30 9:30:00")+5000)
 plot(spodPIDs$h, type="l")
 
-tau <- c(0.05, 0.1)
+tau <- c(0.05, 0.1, 0.15)
 k <- 3
 detrendr_trends <- data.frame(time = spodPIDs$time)
 qsreg_trends <- data.frame(time = spodPIDs$time)
@@ -35,7 +35,7 @@ x <- seq(1, nrow(spodPIDs), 1)
 for (node in c("f", "g", "h")){
   spodPIDs[, node] <- na.locf(spodPIDs[, node])
   result <- get_trend_BIC(spodPIDs[, node], tau, k, 
-                            lambdaSeq = n^seq(.5, 1.2, length.out=10),
+                            lambdaSeq = n^seq(.8, 1.7, length.out=10),
                             df_tol = 1e-9,
                             gamma = 1,
                             plot_lambda = TRUE,

### 2019-01-22 Code/timing.R

@@ -55,22 +55,6 @@ for (n in data_lengths){
     }
   }
 }
-times_df <- times_df[-i,]    
-diff_df <- diff_df[-j,]  
 
 
-times_df %>% group_by(n, windows) %>%
-  summarise(time_mean = mean(time)) %>%
-  ggplot(aes(x=n, y=time_mean, group=factor(windows), col=factor(windows))) + 
-  geom_line()
-
-
-
-# load("../TimingData/single_window_times_3.RData")
-# time.df <- data.frame(n = data_lengths, 
-#            median = apply(all.times, 2, median)*1e-9,
-#            mean = apply(all.times, 2, mean)*1e-9)
-# 
-# ggplot(time.df, aes(x=n, y=median)) + geom_line() + geom_point()
-# 
 

### 2019-01-22 Manuscript/detrendify_JCGS.tex

@@
 \cite{brantley2014mobile} examined air pollutant measurements collected using a monitoring vehicle and proposed using a smooth baseline trend as an estimate of regional background pollution.
+
+	\begin{figure}
+		\caption{Example of 3 co-located SPod PID sensor readings.}
+		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
+	\end{figure}
 
@@
 
+	We compared our detrend method with the eBIC criteria with the qsreg method on a subset of the real data since the qsreg method cannot handle all 15 hours simultaneously.
+We estimate the baseline trend using both the 5th and 10th quantiles and then classify the signal as any observation above the median plus 2.5 times the median absolute deviation after the baseline trend removal.
+We report both the confusion matrices and the normalized mutual information (NMI) score.
+
+	\input{short_confusion_detrend.tex}
+
+	\begin{figure}
+		\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 10th quantile.}
+		\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
+	\end{figure}
+
+	\input{NMI.tex}
+
 	Our windowed detrend method was used to removed the baseline drift from low cost air quality sensors so that the signal could be categorized using a simple threshold.
@@
 			and zoomed in on signal area (bottom). Horizontal lines represent signal thresholds.}
-		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
 		\includegraphics[width = \linewidth]{Figures/corrected_data.png}
@@
 	\input{complete_confusion.tex}
-
-	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
 

### 2019-01-24 Manuscript/application_short.tex
manuscript and figure edits

--- 2019-01-17 Manuscript/detrendify_JCGS.tex
@@
 
-	%\bibliographystyle{natbib}
-
-	\def\spacingset#1{\renewcommand{\baselinestretch}%
-		{#1}\small\normalsize} \spacingset{1}
-
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-	\if0\blind
-	{
-		\title{\bf Drift Removal for Time Series Data Using Quantile Trend Filtering}
-		\author{Halley Brantley\thanks{
-				Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: hlbrantl@ncsu.edu)} \,
-			Joseph Guinness\thanks{
-				Department of Biological Statistics and Computational Biology, Cornell University, Ithaca, NY 14853 (E-mail: guinness@cornell.edu )} \,
-			and
-			%    and
-			Eric C. Chi\thanks{Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: eric$\_$chi@ncsu.edu).}    \\}
-		\date{}
-		\maketitle
-	} \fi
-
-	\if1\blind
-	{
-		\bigskip
-		\bigskip
-		\bigskip
-		\begin{center}
-			{\LARGE\bf Title}
-		\end{center}
-		\medskip
-	} \fi
-
-	\bigskip
-	\begin{abstract}
-		The text of your abstract.  200 or fewer words.
-	\end{abstract}
-
-	\noindent%
-	{\it Keywords:}  3 to 6 keywords, that do not appear in the title
-	\vfill
-
-	\newpage
-	\spacingset{1.5} % DON'T change the spacing!
-	\section{Introduction}
-	\label{sec:intro}
-
-	We are interested in the problem posed by \cite{thoma2016south} concerning low-cost air pollutant sensors.
-Researchers at the U.S.
-Environmental Protection Agency have built prototypes of low-cost sensors that provide an un-calibrated measure of volatile organic compounds (VOCs) and hazardous air pollutants (HAPs) at a 1 second time resolution.
-To reduce cost and power consumption, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled.
-A result the output signal exhibits a slowly varying baseline drift (minutes to hours).
-The purpose of these sensors is to protect the fenceline of industrial facilities by detecting whether VOC emissions from a local source are present and providing wind and concentration data to help identify the source and speed repairs.
-
-	There have been a number of occasions when researchers have sought to estimate the trend in the quantiles of a random variable over time or another co-variate given noisy observations of that variable.
-\cite{nychka1995nonparametric} estimated smooth quantile curves of friction profiles of syringes in order to automate their quality assessment procedures.
-\cite{Ning2014} proposed a method for separating chromatogram peaks from the smooth baseline and noise present in the measurements.
-\cite{brantley2014mobile} examined air pollutant measurements collected using a monitoring vehicle and proposed using a smooth baseline trend as an estimate of regional background pollution.
-
-	A variety of approaches have been proposed for estimating quantile trends.
-\cite{Koenker1978} was the first to propose the use of the check loss function to estimate conditional quantiles.
-Given a design matrix $\mathbf{X} \in \mathbb{R}^{n \times p}$ and continuous responses $y \in \mathbb{R}^n$, we estimate a regression vector $\theta \in \mathbb{R}^p$ so that $\mathbf{X}\theta$ is a good approximation of the $\tau$th quantile of $y$ conditional on $x$.
-The regression vector $\theta$ is estimated by finding the minimizer of $\rho_{\tau}(y-X\theta)$ where $\rho_{\tau}(y) = \sum_{i=1}^n y_i(1-\mathbf{I}(y_i<0))$ is the check-loss function and $\mathbf{I}$ is the indicator function.
-
-	In the trend filtering problem \citep{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $y \in \mathbb{R}^n$ by solving the following convex problem.
-	\begin{eqnarray}
-	\underset{\theta}{\arg\min}\; \frac{1}{2n} \lVert y - \theta \rVert_2^2 + \lambda \lVert \mathbf{D}^{(k+1)}\theta \rVert_1,
-	\end{eqnarray}
-	where $\lambda \geq 0$ is a regularization parameter that trades off the emphasis on the data fidelity term and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
-To understand the purpose of penalizing $\mathbf{D}^{(k+1)}$ consider the difference operator when $k = 0$.
-	\begin{eqnarray}
-	\mathbf{D}^{(1)} = \begin{pmatrix}
-	-1 & 1 & 0 & \cdots & 0 & 0 \\
-	0 & -1 & 1 & \cdots & 0 & 0 \\
-	\vdots & & & & & \\
-	0 & 0 & 0 & \cdots & -1 & 1 \\
-	\end{pmatrix}
-	\end{eqnarray}
-	Thus, $\lVert \mathbf{D}^{(1)}\theta \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
-The penalty incentivizes solutions which are piece-wise constant.
-For $k \geq 1$, the difference operator $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n-k-1) \times n}$ is defined recursively as follows
-	\begin{eqnarray}
-	\mathbf{D}^{(k+1)} & = & \mathbf{D}^{(1)}\mathbf{D}^{(k)}.
-	\end{eqnarray}
-	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
-
-	We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate for removing trends in time series.
-The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
-Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles.
-We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and proposed a modified criteria for choosing the smoothing parameter.
-We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.
-
-	\section{Methods}
-
-	\subsection{Quantile Trend Filtering}
-
-	We combine the ideas of quantile regression and trend filtering, namely consider the signal approximation problem, where the design $\mathbf{X}$ is the identity matrix.
-
-	The estimation of the quantile trend filtering model can be posed as the following optimization problem.
-	\begin{eqnarray}
-	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \frac{1}{n} \sum_{i=1}^n \rho_\tau(y_i - \theta_i) + \lambda \lVert \mathbf{D}^{(k)} \theta \rVert_1,
-	\end{eqnarray}
-	where $\lambda$ is a non-negative tuning parameter. As with the classic quantile regression, the quantile trend filtering problem can be solved by a linear program.
-
-	\subsection{ADMM for Smoothing Windows}
-
-	To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
-We propose an alternating direction method of multipliers (ADMM) algorithm for solving large problems in a piecewise fashion.
-The ADMM algorithm is fully described by \cite{boyd2011distributed}.
-We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem,
-	\begin{equation}
-	\text{minimize } \rho_\tau(y-\theta) + \lambda||D^{(k)}\theta||_1\\
-	\end{equation}
-	where $y,\theta \in \mathbf{R}^n$, $y$ is the observed data, $\rho_{\tau}(u) = \sum_i(\tau - I(u_i < 0))u_i$ is the check loss function corresponding to quantile level $\tau$ and $D^{(k)}$ is the discrete differencing matrix of order $k$ and $\lambda$ is a tuning parameter that controls smoothness.
-We first re-parameterize $\phi = y-\theta$ so the problem is
-	\begin{equation}
-	\text{minimize } \rho_\tau(\phi) + \lambda||D^{(k)}(y-\phi)||_1\\
-	\end{equation}
-	We further divide $\phi$ order to solve smaller problems: Defining
-	\begin{align}
-	&\phi_1 = (\phi_{11}, \phi_{12})\\
-	&\phi_2 = (\phi_{21}, \phi_{22}, \phi_{23})\\
-	&\phi_3 = (\phi_{31}, \phi_{32})\\
-	&\phi = (\phi_{11}, \phi_{12}=\phi_{21}, \phi_{22}, \phi_{23}=\phi_{31}, \phi_{32}) \\
-	\end{align}
-	Dividing $y$ similarly, the problem then becomes 
-	\begin{align}
-	&\text{minimize } \sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1\\
-	&\text{subject to: } \phi_{12}=\phi_{21}, ~~ \phi_{23}=\phi_{31}\\
-	\end{align}
-	We can further simplify by defining 
-	\begin{align}
-	&\overline{\phi} = (\phi_{11}, \frac{\phi_{12}+\phi_{21}}{2}, \phi_{22}, \frac{\phi_{23}+\phi_{31}}{2}, \phi_{32}) \\
-	&\overline{\phi_1} = (\phi_{11}, \frac{\phi_{12}+\phi_{21}}{2})\\
-	&\overline{\phi_2} = ( \frac{\phi_{12}+\phi_{21}}{2}, \phi_{22}, \frac{\phi_{23}+\phi_{31}}{2})\\
-	&\overline{\phi_3} = (\frac{\phi_{23}+\phi_{31}}{2}, \phi_{32})
-	\end{align}
-	so the problem becomes
-	\begin{align}
-	&\text{minimize } \sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1\\
-	&\text{subject to: } \phi_{i}=\overline{\phi_i}\\
-	\end{align}
-	The augmented Lagrangian for this problem is 
-	\begin{align}
-	&L_\gamma(\phi_1,\phi_2, \phi_3, \overline{\phi_1}, \overline{\phi_2}, \overline{\phi_3}, \omega) = \\
-	&\sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1 + \omega_i^T(\phi_i - \overline{\phi_i}) + \frac{\gamma}{2}||\phi_i - \overline{\phi_i}||_2^2
-	\end{align}
-	The ADMM updates are then given by
-	\begin{align}
-	&\phi_i^{k+1} = \arg\min_{\phi_i}\rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1 + \omega_i^{kT}(\phi_i - \overline{\phi_i}^k) + \frac{\gamma}{2}||\phi_i - \overline{\phi_i}^k||_2^2\\
-	&\omega_i^{k+1} = \omega_i^{k} + \gamma(\phi_i^{k+1} - \overline{\phi_i}^{k+1})
-	\end{align}
-	The $\phi_i$ updates can be obtained using a quadratic program solver such as Gurobi and can be obtained in parallel.
-
-	\pagebreak
-
+	\section{Method comparison application}
 	\begin{figure}[!h]
-		\caption{Windows fit separately compared to simultaneous fit, no signal present.}
-		\includegraphics[width = \linewidth]{Figures/overlapping_windows.png}
+		\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile and a threshold of the median plus 3MAD.}
+		\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
 	\end{figure}
@@
 	\begin{figure}[!h]
-		\caption{Windows fit with 150 iterations of ADMM, no signal present.}
-		\includegraphics[width = \linewidth]{Figures/admm_windows.png}
+		\centering
+		\caption{Variation of Information (VI) between sensor nodes after trend removal by quantile and method.
+The rows in the figure represent the three thresholds used: 3MAD, 4MAD, and 5MAD.
+Lower VI represents more similar classifications.
+}
+		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
 	\end{figure}
 
-	We conducted timing experiments to show the advantages of using our ADMM algorithm on large datasets.
-We use 2 windows for each data size with an overlap of 2000 and two iterations of the algorithm after the initial estimates.
-
 		\begin{figure}[!h]
-		\caption{Timing experiments comparing quantile trend filtering on a single window versus 2 windows by data size.}
-		\includegraphics[width = \linewidth]{Figures/Fig_timing_experiment.png}
+		\centering
+		\caption{Normalized Mutual Information (NMI) between sensor nodes after trend removal by quantile and method.
+The rows in the figure represent the three thresholds used: 3MAD, 4MAD, and 5MAD.
+Higher NMI represents more similar classifications.
+}
+		\includegraphics[width = .9\linewidth]{Figures/NMI_app_short.png}
 	\end{figure}
 
-	\subsection{Smoothing parameter choice}
-	Our method can easily handle missing data by changing the check loss function to output 0 for missing values.
-This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data.
-A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.
-\cite{KoenkerNgPortnoy1994} relate $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
-	\begin{equation}
-	\mbox{SIC}(p_{\lambda}) = \log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}\log n
-	\end{equation}
-
-	The traditional Bayesian Information Criterion (BIC) is given by 
-	\begin{equation}
-	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n 
-	\end{equation}	
-	where $\theta(s)$ is the parameter $\theta$ with those components outside $s$ being set to 0, and $\nu(s)$ is the number of components in $s$.
-If we assume an asymmetric Laplace likelihood $L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_i\rho_\tau(\frac{y_i - \theta_i}{\sigma})\right\}$ and the number of non-zero elements of $D\theta$ as $df$
-	\begin{equation}
-	\mbox{BIC}(df) = 2\sum_i\frac{1}{\sigma}\rho_{\tau}(y_i-\theta_i) + \mbox{df}\log n
-	\end{equation} 
-	We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates. \cite{chen2008} proposed the extended BIC for large parameter spaces 
-	\begin{equation}
-	\mbox{BIC}_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j,}~~\gamma \in [0,1]
-	\end{equation}
-	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model.
-We used this criteria with $\gamma = 1$, $P=n-k$ where $k$ is the order of the differencing matrix and $j = \nu(s)$ is the number of non-zero entries in $D^{(k)}\theta$.
-
-	We used a single dataset to illustrate the difference between the scaled, unscaled and extended BIC criteria.
-
-	\begin{figure}[h!]
-		\caption{Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.} 
-		\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
-		\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
-		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
-	\end{figure}
-
-	\section{Simulation Studies}
-
-	\subsection{Estimating Quantiles}
-
-	We compare the performance of our quantile trend filtering method with the three previously published methods using designs proposed by \cite{Racine2017}.
-The methods compared are: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
-The smoothing parameter $\lambda$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria as described in \cite{KoenkerNgPortnoy1994}.
-We further compare three criteria for choosing the smoothing parameter for our detrend method: \texttt{detrendr\_SIC}: Our method where we minimize $\sum_i\rho_{\tau}(y_i - \theta_i) + \lambda||D\theta||_1$ and $\lambda$ is chosen using SIC \citep{KoenkerNgPortnoy1994}.
-\texttt{detrendr\_valid}: Our method where lambda is chosen by leaving out every 5th observation as a validation data set and minimizing the evaluating the check loss function evaluated at the validation data.
-\texttt{detrendr\_eBIC}:  the new criteria we have proposed based on the extended BIC proposed by \cite{chen2008}.
-
-	Three simulation designs from \cite{Racine2017} were considered. For all designs $X_i$ was generated as a uniformly spaced sequence in $[0,1]$ and the response $Y$ was generated as 
-	$$Y_i = sin(2\pi x_i) + \epsilon_i(x_i)$$
-	The three error distributions considered were 
-	\begin{itemize}
-		\item Gaussian: $\epsilon_i(x_i) \sim N\left(0, \left(\frac{1+x_i^2}{4}\right)^2\right)$
-		\item Beta: $\epsilon_i \sim Beta(1, 11-10x_i)$
-		\item Mixed normal: $\epsilon_i$ is simulated from a mixture of $N(-1,1)$ and  $N(1,1)$ with mixing probability $x_i$.
-	\end{itemize}
-	\begin{figure}
-		\caption{Simulated data with true quantiles $\tau \in \{0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99\}$}	
-		\includegraphics[width=.28\linewidth]{Figures/gaus.png}
-		\includegraphics[width=.3\linewidth]{Figures/shapebeta.png}
-		\includegraphics[width=.45\linewidth]{Figures/mixednorm.png}
-	\end{figure}
-
-	100 datasets were generated of sizes 300, 500 and 1000.
-The MSE was calculated as $\frac{1}{n}\sum_i (\hat{q}_{\tau}(x_i) - q_\tau(x_i))^2$.
-The plots below show the mean MSE $\pm$ twice the standard error by method, quantile level and sample size.
-
-	\begin{figure}
-		\caption{RMSE by design, method, quantile and data size.}
-		\includegraphics[width=\linewidth]{Figures/gaus_mse.png}	
-		\includegraphics[width=\linewidth]{Figures/shapebeta_mse.png}
-		\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
-	\end{figure}
-	In all of the designs the proposed detrend methods are either better than or comparable to existing methods.
-The npqw method performs particularly poorly in the mixed normal design, due to the fact that it assumes the data comes from a scale-location model which is violated in this case.
-
-	\subsection{Peak Detection}
-	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
-	\begin{equation}
-	Y(t) = s(t) + b(t) + \epsilon(t)
-	\end{equation} 
-	where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, \sigma^2)$ is an error component.
-We assume $t$ is a uniformly spaced sequence of time points between 0 and 1.
-We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from $n/50$ to $n/25$ with equal probability, and coefficients drawn from an exponential distribution with rate 1.
-The true signal function is assumed to be zero with Gaussian peaks.
-The number of peaks is sampled from $n/100$ to $n/50$ with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $1/n$ and $5/n$ and areas uniformly distributed between 0 and $20/n$.
-One hundred datasets were generated for $n=\{300, 500, 1000, 5000\}$.
-We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the root mean squared error (RMSE).
-
-	\begin{figure}[h]
-		\caption{Example of simulated peaks, baseline, and observed measurements.}
-		\includegraphics[width = \linewidth]{Figures/ex_peaks.png}
-	\end{figure}
-
-	In addition to the RMSE we also calculated the signal miss-classification rate.
-We first classified the ``true" peaks function into signal or not based on a threshold of 0.1.
-We then classified the detrended data using three different thresholds and calculated the fraction miss-classified.
-An illustration of the observations classified as signal after detrending compared to the ``true signal" is shown in Fig.
-9.
-
-	Our detrend\_BIC method performs the best overall in terms of both RMSE and miss-classification rate.
-The lowest miss-classification rates were obtained using the detrend\_eBIC method and a threshold of 1 for all data sizes.
-While qsreg was competitive with our method in some cases, both the RMSE and miss-classification rate increased substantially with the size of the dataset.
-
-	\begin{figure}[h!]
-		\caption{Example signal classification using threshold. Red indicates true signal $>0.1$, blue indicates classified as signal after baseline removal using eBIC detrendr and a threshold of 1.}
-		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
-	\end{figure}
-
-	\begin{figure}[h!]
-		\caption{Class averaged accuracy by threshold, data size, and method (1 is best 0.5 is worst).}
-		\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
-	\end{figure}
-
-	\FloatBarrier
-
-	\section{Application}
-
-	Our windowed detrend method was used to removed the baseline drift from low cost air quality sensors so that the signal could be categorized using a simple threshold.
-The measurements were first standardized to have mean zero and variance 1.
-Three quantile levels for estimating the baseline trend were compared.
-The total dataset consisted of 52,322 observations per node.
-	The signal thresholds were set using the first 15,000 observations where it was known no signal was present.
-The thresholds were set as 3 times the standard deviation plus the mean of observations in this time period.
-The total number of seconds of signal for each node as well as the number of seconds where multiple nodes both reported signal is shown in Table 1.
-Table 2 shows the fraction of observations with different signal classifications by node combination.
-
-	\begin{figure}
-		\caption{Low cost sensor data before drift removal (top), after drift removal using windowed detrend with eBIC (middle), 
-			and zoomed in on signal area (bottom). Horizontal lines represent signal thresholds.}
-		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
-		\includegraphics[width = \linewidth]{Figures/corrected_data.png}
-		\includegraphics[width = \linewidth]{Figures/corrected_zoom_data.png}
-	\end{figure}
-
-	\input{complete_confusion.tex}
-
-	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
-
-	\section{Conclusion}
-	\label{sec:conc}
-
-	\bigskip
-	\begin{center}
-		{\large\bf SUPPLEMENTARY MATERIAL}
-	\end{center}
-
-	\begin{description}
-
-		\item[R-package for detrend routine:] R-package detrendr containing code to perform the diagnostic methods described in the article.
-The package also contains all datasets used as examples in the article.
-(GNU zipped tar file)
-
-	\end{description}
-
-	\section{References}
-
-	\bibliographystyle{asa}
-	\bibliography{detrendify}
+	\newpage
+	\input{short_confusion_detrend_MAD3.tex} 
+	\input{short_confusion_detrend_MAD4.tex} 
+	\input{short_confusion_detrend_MAD5.tex}
 

### 2019-01-24 Manuscript/detrendify_JCGS.tex
manuscript and figure edits

@@
 \usepackage{placeins}
-
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 %\pdfminorversion=4
@@
 
-	We are interested in the problem posed by \cite{thoma2016south} concerning low-cost air pollutant sensors.
-Researchers at the U.S.
-Environmental Protection Agency have built prototypes of low-cost sensors that provide an un-calibrated measure of volatile organic compounds (VOCs) and hazardous air pollutants (HAPs) at a 1 second time resolution.
-To reduce cost and power consumption, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled.
-A result the output signal exhibits a slowly varying baseline drift (minutes to hours).
-The purpose of these sensors is to protect the fenceline of industrial facilities by detecting whether VOC emissions from a local source are present and providing wind and concentration data to help identify the source and speed repairs.
+	The use of low-cost, portable, air quality sensors has increased dramatically in the last decade.
+These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
+The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
+Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
+To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours (Fig.
+\ref{fig:raw_spod}).
 
-	There have been a number of occasions when researchers have sought to estimate the trend in the quantiles of a random variable over time or another co-variate given noisy observations of that variable.
+	\begin{figure}[b!]
+		\label{fig:raw_spod}
+		\caption{Example of 3 co-located SPod PID sensor readings.}
+		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
+	\end{figure}
+
+	There have been a number of other occasions when researchers have sought to estimate the trend in the quantiles of a random variable over time or another co-variate given noisy observations of that variable.
 \cite{nychka1995nonparametric} estimated smooth quantile curves of friction profiles of syringes in order to automate their quality assessment procedures.
@@
 
-	\begin{figure}
-		\caption{Example of 3 co-located SPod PID sensor readings.}
-		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
-	\end{figure}
-
 	A variety of approaches have been proposed for estimating quantile trends.
@@
 Given a design matrix $\mathbf{X} \in \mathbb{R}^{n \times p}$ and continuous responses $y \in \mathbb{R}^n$, we estimate a regression vector $\theta \in \mathbb{R}^p$ so that $\mathbf{X}\theta$ is a good approximation of the $\tau$th quantile of $y$ conditional on $x$.
-The regression vector $\theta$ is estimated by finding the minimizer of $\rho_{\tau}(y-X\theta)$ where $\rho_{\tau}(y) = \sum_{i=1}^n y_i(1-\mathbf{I}(y_i<0))$ is the check-loss function and $\mathbf{I}$ is the indicator function.
+The regression vector $\theta$ is estimated by finding the minimizer of $\rho_{\tau}(y-X\theta)$ where $\rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0))$ is the check-loss function and $\mathbf{I}$ is the indicator function.
 
@@
 	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \frac{1}{n} \sum_{i=1}^n \rho_\tau(y_i - \theta_i) + \lambda \lVert \mathbf{D}^{(k)} \theta \rVert_1,
+	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \mathbf{D}^{(k)} \theta \rVert_1,
 	\end{eqnarray}
@@
 	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \sum_{j=1}^J \left [\frac{1}{n} \sum_{i=1}^n \rho_{\tau_j}(y_i - \theta_{ji}) + 
+	\underset{\theta}{\min}\; \sum_{j=1}^J \left [\frac{1}{n} \rho_{\tau_j}(y - \theta_{j}) + 
 	\lambda_j \lVert \mathbf{D}^{(k)} \theta_j \rVert_1 \right ] \\
@@
 The ADMM algorithm is fully described by \cite{boyd2011distributed, gabay1975dual, glowinski1975approximation}.
-We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem,
-	\begin{equation}
-	\text{minimize } \rho_\tau(y-\theta) + \lambda||D^{(k)}\theta||_1\\
-	\end{equation}
-	where $y,\theta \in \mathbf{R}^n$, $y$ is the observed data, $\rho_{\tau}(u) = \sum_i(\tau - I(u_i < 0))u_i$ is the check loss function corresponding to quantile level $\tau$ and $D^{(k)}$ is the discrete differencing matrix of order $k$ and $\lambda$ is a tuning parameter that controls smoothness.
-We first re-parameterize $\phi = y-\theta$ so the problem is
-	\begin{equation}
-	\text{minimize } \rho_\tau(\phi) + \lambda||D^{(k)}(y-\phi)||_1\\
-	\end{equation}
-	We further divide $\phi$ order to solve smaller problems: Defining
-	\begin{align}
-	&\phi_1 = (\phi_{11}, \phi_{12})\\
-	&\phi_2 = (\phi_{21}, \phi_{22}, \phi_{23})\\
-	&\phi_3 = (\phi_{31}, \phi_{32})\\
-	&\phi = (\phi_{11}, \phi_{12}=\phi_{21}, \phi_{22}, \phi_{23}=\phi_{31}, \phi_{32}) \\
-	\end{align}
-	Dividing $y$ similarly, the problem then becomes 
-	\begin{align}
-	&\text{minimize } \sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1\\
-	&\text{subject to: } \phi_{12}=\phi_{21}, ~~ \phi_{23}=\phi_{31}\\
-	\end{align}
-	We can further simplify by defining 
-	\begin{align}
-	&\overline{\phi} = (\phi_{11}, \frac{\phi_{12}+\phi_{21}}{2}, \phi_{22}, \frac{\phi_{23}+\phi_{31}}{2}, \phi_{32}) \\
-	&\overline{\phi_1} = (\phi_{11}, \frac{\phi_{12}+\phi_{21}}{2})\\
-	&\overline{\phi_2} = ( \frac{\phi_{12}+\phi_{21}}{2}, \phi_{22}, \frac{\phi_{23}+\phi_{31}}{2})\\
-	&\overline{\phi_3} = (\frac{\phi_{23}+\phi_{31}}{2}, \phi_{32})
-	\end{align}
-	so the problem becomes
-	\begin{align}
-	&\text{minimize } \sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1\\
-	&\text{subject to: } \phi_{i}=\overline{\phi_i}\\
-	\end{align}
-	The augmented Lagrangian for this problem is 
-	\begin{align}
-	&L_\gamma(\phi_1,\phi_2, \phi_3, \overline{\phi_1}, \overline{\phi_2}, \overline{\phi_3}, \omega) = \\
-	&\sum_{i=1}^3 \rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1 + \omega_i^T(\phi_i - \overline{\phi_i}) + \frac{\gamma}{2}||\phi_i - \overline{\phi_i}||_2^2
-	\end{align}
-	The ADMM updates are then given by
-	\begin{align}
-	&\phi_i^{k+1} = \arg\min_{\phi_i}\rho_\tau(\phi_i) + \lambda||D^{(k)}(y_i-\phi_i)||_1 + \omega_i^{kT}(\phi_i - \overline{\phi_i}^k) + \frac{\gamma}{2}||\phi_i - \overline{\phi_i}^k||_2^2\\
-	&\omega_i^{k+1} = \omega_i^{k} + \gamma(\phi_i^{k+1} - \overline{\phi_i}^{k+1})
-	\end{align}
-	The $\phi_i$ updates can be obtained using a quadratic program solver such as Gurobi and can be obtained in parallel.
+We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem given in Eq.
+\ref{eq:quantile_trend}, by dividing our observed series $y$ into overlapping windows $\{y_1, ..., y_M\}$, we then estimate the trend separately in each window while constraining the overlapping pieces of the trends to be equal.
 
-	\pagebreak
+	\begin{algorithm}
+		\caption{ADMM algorithm for quantile trend filtering with windows}\label{euclid}
+		\begin{algorithmic}
+			\State{Given $y$ divided into overlapping windows $\{y_1, ..., y_M\}$ and quantiles $\{\tau_1 < ... < \tau_J\}$}
+			\State {\textbf{initialize:} \\
+				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m} < ...<\theta_{J,m}$. \\
+				 $\bar{\theta}_{j,m}^{(0)}$ is the mean of $\theta_{j,m}^{(0)}$ and all overlapping windows.}\\
+			 	 $\omega_{j,m}^{(0)} = \gamma(\theta_{j,m}^0 - \bar{\theta}_{j,m}^0)$	
+			\Repeat{}
+			 $$
+		\theta_{j,m}^{(q)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1 +  \omega_{j,m}^{(q-1)T}(\theta_{j,m} - \bar{\theta}_{j,m}^{(q-1)}) + 
+			\frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}^{(q-1)}||_2^2
+			$$
+			\qquad \qquad subject to $\theta_{1,m} < ...<\theta_{J,m}$
+
+			\State  $\bar{\theta}_{j,m}^{(q)}$ is the mean of $\theta_{j,m}^{(q)}$ and all overlapping windows.
+			\State $\omega_{j,m}^{(q)} = \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)})$	
+			\Until {convergence}
+			\State \textbf{return} Non-overlapping sequence of $\bar{\theta}_{j,m}^{(q)}$ for all $j$, $m$.
+
+		\end{algorithmic}
+	\end{algorithm}
+
+	We used the stopping criteria described by \cite{boyd2011distributed}.
+The criteria are based on the primal and dual residuals which represent the residuals for the primal and dual feasibility, respectively.
+The primal and dual residuals are defined as
+
+	\begin{eqnarray}
+	r_p^{(q)} = \sqrt{\sum_{m=1}^M \left(\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\right)^2}\\
+	r_d^{(q)} = \gamma\sqrt{\sum_{m=1}^M \left(\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\right)^2}
+	\end{eqnarray}
 
 	\begin{figure}[!h] 
-		\caption{Windows fit separately compared to simultaneous fit, no signal present.}
-		\includegraphics[width = \linewidth]{Figures/overlapping_windows.png}
+		\centering
+		\caption{(Top) Windows fit separately compared to simultaneous fit, no signal present. (Bottom) Windows fit with our ADMM algorithm which converged in 7 iterations.}
+		\includegraphics[width = 0.7\linewidth]{Figures/overlapping_windows.png}
+		\includegraphics[width = 0.7\linewidth]{Figures/admm_windows.png}
 	\end{figure}
 
-	\begin{figure}[!h] 
-		\caption{Windows fit with 150 iterations of ADMM, no signal present.}
-		\includegraphics[width = \linewidth]{Figures/admm_windows.png}
-	\end{figure}
+	Given $N$ equal to the total number of elements in $\bar{\theta}$ and values for $\epsilon_{abs}$ and $\epsilon_{rel}$, the algorithm is stopped when
 
-	We conducted timing experiments to show the advantages of using our ADMM algorithm on large datasets.
-We use 2 windows for each data size with an overlap of 2000 and two iterations of the algorithm after the initial estimates.
+	\begin{eqnarray}
+		r_p^{(q)} < \epsilon_{abs}\sqrt{N} + \epsilon_{rel}\max(\lVert \theta_{j,m}^{(q)} \rVert_2, \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2 )\\
+		r_d^{(q)} < \epsilon_{abs}\sqrt{N} + \epsilon_{rel}\sqrt{\sum_{m=1}^M \left(\omega_{j,m}^{(q)}\right)^2}
+	\end{eqnarray}
+
+	Timing experiments illustrate the advantages of using our ADMM algorithm even on datasets where solving the problem simultaneously is possible.
+We use from one to four windows for each data size with an overlap of 500.
+The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
+For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below and trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using a $\lambda = n/5$.
 
 	\begin{figure}[!h] 
-		\caption{Timing experiments comparing quantile trend filtering on a single window versus 2 windows by data size.}
-		\includegraphics[width = \linewidth]{Figures/Fig_timing_experiment.png}
+		\centering
+		\caption{Timing experiments comparing quantile trend filtering with varying numbers of windows by data size.}
+		\includegraphics[width = 0.7\linewidth]{Figures/Fig_timing_experiment.png}
 	\end{figure}
@@
 
-	\input{short_confusion_detrend.tex}
+		\begin{figure}
+		\centering
+		\caption{Variation of Information between sensor nodes after trend removal by quantile and method and thresholding by factor of MAD.}
+		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
+	\end{figure}
 
@@
 	\end{figure}
-
-	\input{NMI.tex}
 
@@
 	\begin{figure}
-		\caption{Low cost sensor data before drift removal (top), after drift removal using windowed detrend with eBIC (middle), 
-			and zoomed in on signal area (bottom). Horizontal lines represent signal thresholds.}
+		\caption{Low cost sensor data after drift removal using windowed detrend with eBIC.}
 		\includegraphics[width = \linewidth]{Figures/corrected_data.png}
-		\includegraphics[width = \linewidth]{Figures/corrected_zoom_data.png}
 	\end{figure}

### 2019-01-24 Code/Fig_app_short.R

@@ -5,72 +5,132 @@ library(Hmisc)
 library(caret)
 library(zoo)
 library(aricode)
+library(mcclust)
 load_all("detrendr")
-source("application_functions.R")
 rm(list=ls())
 load("../SPod/trends_short.RData")
+source("application_functions.R")
+colPal <- c('#1b7837', '#c2a5cf')
+            '#762a83','#9970ab',,
+                '#a6dba0','#5aae61',)
 
 spodPeaks <- select(spodPIDs, -time) - select(detrendr_trends,
                                               contains(paste(0.15)))
-plot(spodPeaks$h, type="l")
+plot(spodPeaks$f, type="l")
 thresholds <- apply(spodPeaks, 2, 
                       function(x) median(x, na.rm=T) + 
-                      4*median(abs(x-median(x, na.rm=T)), na.rm=T))
+                      3*median(abs(x-median(x, na.rm=T)), na.rm=T))
 abline(h=thresholds[3], col="red")
 plot(g~f, spodPeaks)
 
+cor(spodPeaks$g, spodPeaks$f, method = "spearman")
 
 ################################################################################
-crit <- 5
-NMI <- {}
-conf <- {}
-for (j in 1:length(tau)){
-  detrendr_signal <- get_spod_signal(tau[j], detrendr_trends, spodPIDs, crit)
-  qsreg_signal <- get_spod_signal(tau[j], qsreg_trends, spodPIDs, crit)
-  conf <- rbind(conf, cbind(get_confusion(detrendr_signal), 
-                            get_confusion(qsreg_signal)))
-  NMI <- rbind(NMI, 
-               cbind("fg", round(NMI(detrendr_signal$f, detrendr_signal$g, 
-                               variant="sqrt"),2),
-                     round(NMI(qsreg_signal$f, qsreg_signal$g, 
-                               variant="sqrt"), 2)), 
-               cbind("fh", round(NMI(detrendr_signal$f, detrendr_signal$h, 
-                               variant="sqrt"),2),
-                     round(NMI(qsreg_signal$f, qsreg_signal$h, 
-                               variant="sqrt"), 2)), 
-               cbind("gh", round(NMI(detrendr_signal$g, detrendr_signal$h, 
-                               variant="sqrt"),2),
-                     round(NMI(qsreg_signal$g, qsreg_signal$h, 
[…]

### 2019-01-24 Code/Fig_consensus.R

@@ -4,15 +4,15 @@ load_all("detrendr")
 
 rm(list=ls())
 source("sim_generating_functions.R")
-set.seed(987656)
-overlap <- 100
+set.seed(987651)
+overlap <- 150
 window_size <- 500
 n <- window_size*3 - overlap*2
 df.data <- generate_peaks_design(n)
 y <- df.data$y
 x <- df.data$x
 k <- 3
-tau <- c(0.05, 0.2)
+tau <- c(0.05, 0.1)
 
 lambda <- length(y)
 max_iter <- 25
@@ -34,139 +34,138 @@ ggplot(df.data, aes(x=x, y=y)) +
   scale_color_brewer(palette = "Set1")+
   labs(col="Quantile", linetype = "", x = "") + 
   theme_bw() 
-ggsave("../Manuscript/Figures/overlapping_windows.png", width=7, height=1.5)
+ggsave("../Manuscript/Figures/overlapping_windows.png", width=7, height=2.5)
 
-result <- get_trend_windows(y, tau, lambda, k, rho, window_size, overlap, 
-                            max_iter, update = 1)
-
-
-df_no <- rbind(data.frame(time=x, method = "Single Fit", result0), 
-               data.frame(time=x, method = "Windows", result))
+result <- get_trend_windows(y, tau, lambda, k, window_size, overlap, 
+                            max_iter, update = 1, eps_abs = .01)
 
 
+df_no <- rbind(data.frame(x=x, method = "Single Fit", result0), 
+               data.frame(x=x, method = "Windows", result))
 
-ggplot(df.data, aes(x=time, y=y)) +
+
+ggplot(df.data, aes(x=x, y=y)) +
   geom_line(col="grey") +
   geom_line(data = df_no, aes(y=X1, col = "0.05", linetype = method))+
   geom_line(data = df_no, aes(y=X2, col = "0.10", linetype = method))+
   scale_color_brewer(palette = "Set1")+
   labs(col="Quantile", linetype = "", x = "") + 
   theme_bw() 
-ggsave("../Manuscript/Figures/admm_windows.png", width=7, height=1.5)
-
+ggsave("../Manuscript/Figures/admm_windows.png", width=7, height=2.5)
+max(abs(result0-result))
 
 ################################################################################
 # Application
-dataDir <- "~/Desktop/EPA/SPod_Data/TestRange_Dec2017"
[…]

### 2019-01-25 Manuscript/detrendify_JCGS.tex
Corrected stopping criteria, improved notation

@@
 
+	There have been a number of occasions when researchers have sought to estimate the trends over time given noisy measurements.
+\cite{nychka1995nonparametric} estimated smooth quantile curves of friction profiles of syringes in order to automate their quality assessment procedures.
+\cite{Ning2014} proposed a method for separating chromatogram peaks from the smooth baseline and noise present in the measurements.
+\cite{brantley2014mobile} examined air pollutant measurements collected using a monitoring vehicle and proposed using a smooth baseline trend as an estimate of regional background pollution.
+
 	The use of low-cost, portable, air quality sensors has increased dramatically in the last decade.
@@
 
-	There have been a number of other occasions when researchers have sought to estimate the trend in the quantiles of a random variable over time or another co-variate given noisy observations of that variable.
-\cite{nychka1995nonparametric} estimated smooth quantile curves of friction profiles of syringes in order to automate their quality assessment procedures.
-\cite{Ning2014} proposed a method for separating chromatogram peaks from the smooth baseline and noise present in the measurements.
-\cite{brantley2014mobile} examined air pollutant measurements collected using a monitoring vehicle and proposed using a smooth baseline trend as an estimate of regional background pollution.
-
-	A variety of approaches have been proposed for estimating quantile trends.
-\cite{Koenker1978} was the first to propose the use of the check loss function to estimate conditional quantiles.
+	A variety of approaches have been proposed for estimating quantile trends.	 \cite{Koenker1978} was the first to propose the use of the check loss function to estimate conditional quantiles.
 Given a design matrix $\mathbf{X} \in \mathbb{R}^{n \times p}$ and continuous responses $y \in \mathbb{R}^n$, we estimate a regression vector $\theta \in \mathbb{R}^p$ so that $\mathbf{X}\theta$ is a good approximation of the $\tau$th quantile of $y$ conditional on $x$.
 The regression vector $\theta$ is estimated by finding the minimizer of $\rho_{\tau}(y-X\theta)$ where $\rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0))$ is the check-loss function and $\mathbf{I}$ is the indicator function.
+\cite{nychka1995nonparametric} estimated quantile splines by combining the smoothing spline penalty commonly used in non-parametric regression with the check-loss function used in qunatile regression:
+
+	\begin{equation*}
+	\rho_{\tau}(Y - f) + \lambda\int (f''(t))^2 dt, 
+ 	\end{equation*}
+
+	where $Y(t)$ is a response variable observed at time $t$, $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing.
 
@@
 	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
+
+	\begin{figure}
+		\centering
+		\caption{Examples of trend filtering solutions.}
+		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
+		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}
+	\end{figure}
 
@@
 	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \sum_{j=1}^J \left [\frac{1}{n} \rho_{\tau_j}(y - \theta_{j}) + 
+	\underset{\theta_1, ..., \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
 	\lambda_j \lVert \mathbf{D}^{(k)} \theta_j \rVert_1 \right ] \\
@@
 
-	where $\theta = (\theta_1, ..., \theta_J)$ and $\theta_j \in \mathcal{R}^n$.
-The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
+	where $\theta_j \in \mathcal{R}^n$. The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
 
@@
 We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem given in Eq.
-\ref{eq:quantile_trend}, by dividing our observed series $y$ into overlapping windows $\{y_1, ..., y_M\}$, we then estimate the trend separately in each window while constraining the overlapping pieces of the trends to be equal.
+\ref{eq:quantile_trend}, by dividing our observed series $y(t)$ with $t = \{1, ..., N\}$ into overlapping windows
+
+	\begin{align*}
+	\begin{cases}
+	y_1(t) = y(t) & \mbox{if~~} 1 \le t \le u_{1}\\
+	y_2(t) = y(t) & \mbox{if~~} l_{2} \le t \le u_{2} \\
+	y_3(t) = y(t) & \mbox{if~~} l_{3} \le t \le u_{3} \\
+	\cdots & \\
+	y_M(t) = y(t) & \mbox{if~~} l_{M} \le t \le  N\\
+	\end{cases}
+	\end{align*}
+
+	 with breakpoints $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< N$.
+Given quantiles $\tau_1 < ...
+< \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$.
+In order to write out the constraint the that overlapping sections must be equal we define a consensus variable
+
+	 \begin{align}
+		 \bar{\theta}_{j,m} =&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
+		 =& \begin{cases} 
+			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
+			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
+			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}  
+			 \end{cases},
+	\end{align}
+	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
+	 \begin{align}
+		 \label{eq:quantile_windows}
+		 &\sum_{m=1}^M\underset{\theta_{1,m}, ..., \theta_{J,m}}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
+		 \lambda_j \lVert \mathbf{D}^{(k)} \theta_{j,m} \rVert_1 \right ] \\
+		 &\text{subject to: }\; \theta_{1,m}(t) \le \theta_{2,m}(t) \le ... \le \theta_{J,m}(t) \text{ for all } m,t \\
+		 &\text{subject to: }\; \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
+	 \end{align}
+	 Given values for the Lagrange multiplier $\omega_{j,m}$ and the consensus variable $\bar{\theta}_{j,m}$ we can write the augmented Lagrangian for finding the trends in window $m$:
+	 \begin{equation}
+	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \mathbf{D}^{(k)}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
+	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
+	 \end{equation}
+
+	 We then estimate the trend separately in each window while constraining the overlapping pieces of the trends to be equal according the algorithm 1.
 
@@
 		\begin{algorithmic}		
-			\State{Given $y$ divided into overlapping windows $\{y_1, ..., y_M\}$ and quantiles $\{\tau_1 < ... < \tau_J\}$}
-			\State {\textbf{initialize:} \\
-				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m} < ...<\theta_{J,m}$. \\
-				 $\bar{\theta}_{j,m}^{(0)}$ is the mean of $\theta_{j,m}^{(0)}$ and all overlapping windows.}\\
-			 	 $\omega_{j,m}^{(0)} = \gamma(\theta_{j,m}^0 - \bar{\theta}_{j,m}^0)$	
+			\State \textbf{initialize:} \\
+				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < ...<\theta_{J,m}(t)$ for all $t$. \\		
+			 	 $\omega_{j,m}^{(0)} = \mathbf{0}$	
 			\Repeat{}
-			 $$
-		\theta_{j,m}^{(q)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1 +  \omega_{j,m}^{(q-1)T}(\theta_{j,m} - \bar{\theta}_{j,m}^{(q-1)}) + 
-			\frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}^{(q-1)}||_2^2
-			$$
-			\qquad \qquad subject to $\theta_{1,m} < ...<\theta_{J,m}$
-
-			\State  $\bar{\theta}_{j,m}^{(q)}$ is the mean of $\theta_{j,m}^{(q)}$ and all overlapping windows.
-			\State $\omega_{j,m}^{(q)} = \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)})$	
+			\State  		
+			$\bar{\theta}_{j,m}^{(q)} = g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
+			\State 
+			$\omega_{j,m}^{(q)} = \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \bar{\theta}_{j,m}^{(q)})$	
+			\State
+				$\theta_{j,m}^{(q)} = \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
+			 subject to $\theta_{1,m}(t) < ...<\theta_{J,m}(t)$ for all $t$.
 			\Until {convergence}
@@
 
-	\begin{eqnarray}
-	r_p^{(q)} = \sqrt{\sum_{m=1}^M \left(\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\right)^2}\\
-	r_d^{(q)} = \gamma\sqrt{\sum_{m=1}^M \left(\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\right)^2}
-	\end{eqnarray}
+	\begin{align}
+	&r_p^{(q)} = \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2}\\
+	&r_d^{(q)} = \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2}.
+	\end{align}
+
+	The primal residual, $r_p$, represents the difference between the trend values in the windows and the consensus trend value while the dual residual, $r_d$s represents the change in the consensus variable from one iterate to the next.
+The algorithm is stopped when
+
+	\begin{align}
+		&r_p^{(q)} < \epsilon_{abs}\sqrt{NJ} + \epsilon_{rel}\underset{m}{\max}\left[\max 
+		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
+		&r_d^{(q)} < \epsilon_{abs}\sqrt{NJ} + \epsilon_{rel}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
+	\end{align}
 
@@
 	\end{figure}
-
-	Given $N$ equal to the total number of elements in $\bar{\theta}$ and values for $\epsilon_{abs}$ and $\epsilon_{rel}$, the algorithm is stopped when
-
-	\begin{eqnarray}
-		r_p^{(q)} < \epsilon_{abs}\sqrt{N} + \epsilon_{rel}\max(\lVert \theta_{j,m}^{(q)} \rVert_2, \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2 )\\
-		r_d^{(q)} < \epsilon_{abs}\sqrt{N} + \epsilon_{rel}\sqrt{\sum_{m=1}^M \left(\omega_{j,m}^{(q)}\right)^2}
-	\end{eqnarray}
 
@@
 	\begin{equation}
-	\mbox{SIC}(p_{\lambda}) = \log[n^{-1}\sum\rho_{\tau}(y_i - \widehat{g}(x_i))] + \frac{1}{2n}p_{\lambda}\log n
+	\mbox{SIC}(p_{\lambda}) = \log[\frac{1}{n}\rho_{\tau}(y - \widehat{g}(x))] + \frac{1}{2n}p_{\lambda}\log n
 	\end{equation}
@@
 	\begin{equation}
-	\mbox{BIC}(df) = 2\sum_i\frac{1}{\sigma}\rho_{\tau}(y_i-\theta_i) + \mbox{df}\log n
+	\mbox{BIC}(df) = 2\frac{1}{\sigma}\rho_{\tau}(y-\theta) + \mbox{df}\log n
 	\end{equation} 
@@
 	We compared our detrend method with the eBIC criteria with the qsreg method on a subset of the real data since the qsreg method cannot handle all 15 hours simultaneously.
-We estimate the baseline trend using both the 5th and 10th quantiles and then classify the signal as any observation above the median plus 2.5 times the median absolute deviation after the baseline trend removal.
-We report both the confusion matrices and the normalized mutual information (NMI) score.
+We estimate the baseline trend using both the 5\textsuperscript{th}, 10\textsuperscript{th}, and 15\textsuperscript{th} quantiles.
+We compare three thresholds for separating signal from not signal, the thresholds are calculated using the median and a factor multiple of the median absolute deviation of the detrended series.
+We report both the confusion matrices and the variation of information (VI) distances.
 
@@
 Table 2 shows the fraction of observations with different signal classifications by node combination.
+The BVI scores for the full dataset were 0.14, 0.13, and 0.14 for nodes f and g, f and h, and g and h, respectively.
 
@@
 
-	\input{complete_confusion.tex}
+	\input{full_confusion_detrend.tex}
 

### 2019-01-25 Code/Fig_trendfilter.R

@@ -1,0 +1,27 @@
+library(devtools)
+install_github("hadley/l1tf")
+install_github("glmgen/genlasso")
+library(l1tf)
+library(genlasso)
+source("sim_generating_functions.R")
+set.seed(1122334455)
+n <- 500
+df <- generate_mixednorm(n)
+trend <- trendfilter(df$y, ord = 3)
+df$trend <- coef(trend, lambda = 5000)$beta
+df2 <- generate_peaks_design(n)
+trend2 <- trendfilter(df2$y, ord = 3)
+df2$trend <-  coef(trend2, lambda = 5000)$beta
+
+ggplot(df, aes(x=x, y=y)) + 
+  geom_line(col = "grey") + 
+  geom_line(aes(y=trend), col="red") + 
+  theme_bw()
+ggsave("../Manuscript/Figures/trend_filter_eg1.png", width = 3, height = 2)
+
+ggplot(df2, aes(x=x, y=y)) + 
+  geom_line(col = "grey") + 
+  geom_line(aes(y=trend), col="red") + 
+  theme_bw()
+ggsave("../Manuscript/Figures/trend_filter_eg2.png", width = 3, height = 2)
+

### 2019-01-25 Code/application_full.R

@@ -26,16 +26,13 @@ for (node in c("f", "g", "h")){
   names(spodNode)[2] <- c("pid")
   spodNode$pid <- spodNode$pid/1000
   result <- get_windows_BIC(spodNode$pid, tau, k, window_size, overlap,
-                          lambdaSeq = window_size^seq(0.8, 1.5, length.out=10),
+                          lambdaSeq = window_size^seq(1.1, 1.8, length.out=10),
                           df_tol = 1e-9,
                           gamma = 1,
                           plot_lambda = FALSE,
                           solver = NULL,
                           criteria = "eBIC", 
-                          max_iter = max_iter, 
-                          eps_abs = 0.01, 
-                          eps_rel = 0.002, 
-                          update = 1)
+                          max_iter = max_iter)
   save(result, file=sprintf("../SPod/node_%s_trend.RData", node))
   spod_trends <- cbind(spod_trends, as.data.frame(result$trend))
   names(spod_trends)[(ncol(spod_trends)-length(tau)+1):ncol(spod_trends)] <-

### 2019-01-28 Manuscript/detrendify_JCGS.tex
Introduction edits

@@
 
-	There have been a number of occasions when researchers have sought to estimate the trends over time given noisy measurements.
-\cite{nychka1995nonparametric} estimated smooth quantile curves of friction profiles of syringes in order to automate their quality assessment procedures.
-\cite{Ning2014} proposed a method for separating chromatogram peaks from the smooth baseline and noise present in the measurements.
-\cite{brantley2014mobile} examined air pollutant measurements collected using a monitoring vehicle and proposed using a smooth baseline trend as an estimate of regional background pollution.
-
-	The use of low-cost, portable, air quality sensors has increased dramatically in the last decade.
-These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
-The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
-Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
-To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours (Fig.
-\ref{fig:raw_spod}).
-
-	\begin{figure}[b!]
-		\label{fig:raw_spod}
-		\caption{Example of 3 co-located SPod PID sensor readings.}
-		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
-	\end{figure}
-
-	A variety of approaches have been proposed for estimating quantile trends.	 \cite{Koenker1978} was the first to propose the use of the check loss function to estimate conditional quantiles.
-Given a design matrix $\mathbf{X} \in \mathbb{R}^{n \times p}$ and continuous responses $y \in \mathbb{R}^n$, we estimate a regression vector $\theta \in \mathbb{R}^p$ so that $\mathbf{X}\theta$ is a good approximation of the $\tau$th quantile of $y$ conditional on $x$.
-The regression vector $\theta$ is estimated by finding the minimizer of $\rho_{\tau}(y-X\theta)$ where $\rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0))$ is the check-loss function and $\mathbf{I}$ is the indicator function.
-\cite{nychka1995nonparametric} estimated quantile splines by combining the smoothing spline penalty commonly used in non-parametric regression with the check-loss function used in qunatile regression:
-
-	\begin{equation*}
-	\rho_{\tau}(Y - f) + \lambda\int (f''(t))^2 dt, 
- 	\end{equation*}
-
-	where $Y(t)$ is a response variable observed at time $t$, $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing.
-
-	In the trend filtering problem \citep{Kim2009, Tib2014}, one is interested in finding an adaptive polynomial approximation to noisy data $y \in \mathbb{R}^n$ by solving the following convex problem.
+	In many applications spanning the fields of chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend other more rapidly varying components.
+Given observations $y(t)$, with $t=1, ..., N$, \cite{Kim2009} proposed using \textit{$\ell_1$ trend filtering} to estimate trends that are smooth in the sense of being piecewise linear or piecewise polynomial.
+\cite{Tib2014} then discovered that empirically the trend filtering estimates adapt to the local level of smoothness much better than the more common smoothing splines.
+In the trend filtering problem \citep{Kim2009, Tib2014}, the trend, $\theta \in \mathbb{R}^n$, is estimated by solving the following convex problem.
 	\begin{eqnarray}
-	\underset{\theta}{\arg\min}\; \frac{1}{2n} \lVert y - \theta \rVert_2^2 + \lambda \lVert \mathbf{D}^{(k+1)}\theta \rVert_1,
+	\underset{\theta}{\arg\min}\; \frac{1}{2} \lVert y - \theta \rVert_2^2 + \lambda \lVert \mathbf{D}^{(k+1)}\theta \rVert_1,
 	\end{eqnarray}
-	where $\lambda \geq 0$ is a regularization parameter that trades off the emphasis on the data fidelity term and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
+	where $\lambda \geq 0$ is a regularization parameter and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
 To understand the purpose of penalizing $\mathbf{D}^{(k+1)}$ consider the difference operator when $k = 0$.
@@
 	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
+Trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise (Fig.
+\ref{fig:trendfilter}).
+In some cases, time series may also include a rapidly varying signal component in addition to the trend and noise, in these cases the trend filtering estimate over-estimate the trend in the places where signal is present (Fig.
+\ref{fig:trendfilter}).
 
 	\begin{figure}
+		\label{fig:trendfilter}
 		\centering
-		\caption{Examples of trend filtering solutions.}
+		\caption{Examples of trend filtering solutions. True trend is shown in black, estimated trend is shown in red.}
 		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
@@
 	\end{figure}
+
+	One application in which the observed time series consists of a slowly varying trend, non-negative signal, and rapidly varying noise is the ouput of low cost air quality sensors.
+The use of low-cost, portable, air quality sensors has increased dramatically in the last decade.
+These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
+The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
+Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
+To reduce cost and power consumption of the SPod, the relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours (Fig.
+\ref{fig:raw_spod}).
+
+	\begin{figure}[b!]
+		\label{fig:raw_spod}
+		\caption{Example of 3 co-located SPod PID sensor readings.}
+		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
+	\end{figure}
+
+	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which we can think of as the trend in a low quantile of the data.
+\cite{Koenker1978} was the first to propose substituting the squared loss used in regression with the check loss function  (Eq.
+\ref{eq1}) to estimate a conditional quantile instead of the conditional mean.
+
+	\begin{equation}
+	\label{eq1}
+	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0))
+	\end{equation}
+
+	A variety of approaches have been proposed for estimating quantile trends.
+The regression vector $\theta$ is estimated by finding the minimizer of $\rho_{\tau}(y-X\theta)$ where is the check-loss function and $\mathbf{I}$ is the indicator function.
+\cite{nychka1995nonparametric} estimated quantile splines by combining the smoothing spline penalty commonly used in non-parametric regression with the check-loss function used in qunatile regression:
+
+	\begin{equation*}
+	\rho_{\tau}(Y - f) + \lambda\int (f''(t))^2 dt, 
+ 	\end{equation*}
+
+	where $Y(t)$ is a response variable observed at time $t$, $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing.
 

### 2019-01-28 Manuscript/detrendify_JCGS.tex
added annotations to windows figure

@@
 		\centering
-		\caption{Examples of trend filtering solutions. True trend is shown in black, estimated trend is shown in red.}
+		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}
 		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
@@
 
-	One application in which the observed time series consists of a slowly varying trend, non-negative signal, and rapidly varying noise is the ouput of low cost air quality sensors.
+	One application in which the observed time series consists of a slowly varying trend, non-negative signal, and rapidly varying noise is the output of low cost air quality sensors.
 The use of low-cost, portable, air quality sensors has increased dramatically in the last decade.
@@
 	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \mathbf{D}^{(k)} \theta \rVert_1,
+	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \mathbf{D}^{(k+1)} \theta \rVert_1,
 	\end{eqnarray}
-	where $\lambda$ is a non-negative tuning parameter.
+	where $\lambda$ is a non-negative regularization parameter.
+We address the problem of choosing $\lambda$ in Section \ref{sec:lambda_choice}.
 As with the classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of free or commercial solvers.
@@
 	\begin{eqnarray}
-	\label{eq:quantile_trend}
+	\label{eq:noncross_trend}
 	\underset{\theta_1, ..., \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
-	\lambda_j \lVert \mathbf{D}^{(k)} \theta_j \rVert_1 \right ] \\
+	\lambda_j \lVert \mathbf{D}^{(k+1)} \theta_j \rVert_1 \right ] \\
 	 \text{subject to: }\; \theta_{1i} \le \theta_{2i} \le ... \le \theta_{Ji} \text{ for all } i,
@@
 
-	where $\theta_j \in \mathcal{R}^n$. The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
+	where $\theta_j \in \mathcal{R}^n$.
+The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
+In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi}.
 
@@
 
-	 with breakpoints $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< N$.
+	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< N$.
 Given quantiles $\tau_1 < ...
@@
 		 &\sum_{m=1}^M\underset{\theta_{1,m}, ..., \theta_{J,m}}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
-		 \lambda_j \lVert \mathbf{D}^{(k)} \theta_{j,m} \rVert_1 \right ] \\
+		 \lambda_j \lVert \mathbf{D}^{(k+1)} \theta_{j,m} \rVert_1 \right ] \\
 		 &\text{subject to: }\; \theta_{1,m}(t) \le \theta_{2,m}(t) \le ... \le \theta_{J,m}(t) \text{ for all } m,t \\
@@
 	 \begin{equation}
-	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \mathbf{D}^{(k)}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
+	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \mathbf{D}^{(k+1)}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
 	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
@@
 
+	\begin{figure}[!h] 
+		\centering
+		\caption{Window boundaries and trends fit separately in each window.}
+		\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
+	\end{figure}
+
 	\begin{algorithm}
@@
 		\begin{algorithmic}		
+			\State Define $D = D^{(k+1)}$. 
 			\State \textbf{initialize:} \\
@@
 
-	\begin{figure}[!h] 
+	\begin{figure}
 		\centering
-		\caption{(Top) Windows fit separately compared to simultaneous fit, no signal present. (Bottom) Windows fit with our ADMM algorithm which converged in 7 iterations.}
-		\includegraphics[width = 0.7\linewidth]{Figures/overlapping_windows.png}
-		\includegraphics[width = 0.7\linewidth]{Figures/admm_windows.png}
+		\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.}
+		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
 	\end{figure}
@@
 
-	\subsection{Smoothing parameter choice}
+	\subsection{Regularization Parameter Choice}
+	\label{sec:lambda_choice}
 	Our method can easily handle missing data by changing the check loss function to output 0 for missing values.

### 2019-01-28 Code/Fig_trendfilter.R

@@ -3,24 +3,27 @@ install_github("hadley/l1tf")
 install_github("glmgen/genlasso")
 library(l1tf)
 library(genlasso)
+library(ggplot2)
 source("sim_generating_functions.R")
-set.seed(1122334455)
+set.seed(1122334456)
 n <- 500
-df <- generate_mixednorm(n)
+df <- generate_gaus(n)
 trend <- trendfilter(df$y, ord = 3)
-df$trend <- coef(trend, lambda = 5000)$beta
+df$trend <- coef(trend, lambda = 8000)$beta
 df2 <- generate_peaks_design(n)
 trend2 <- trendfilter(df2$y, ord = 3)
-df2$trend <-  coef(trend2, lambda = 5000)$beta
+df2$trend <-  coef(trend2, lambda = 8000)$beta
 
 ggplot(df, aes(x=x, y=y)) + 
   geom_line(col = "grey") + 
+  geom_line(aes(y=f), col = "black")+
   geom_line(aes(y=trend), col="red") + 
   theme_bw()
 ggsave("../Manuscript/Figures/trend_filter_eg1.png", width = 3, height = 2)
 
 ggplot(df2, aes(x=x, y=y)) + 
   geom_line(col = "grey") + 
+  geom_line(aes(y=baseline), col = "black")+
   geom_line(aes(y=trend), col="red") + 
   theme_bw()
 ggsave("../Manuscript/Figures/trend_filter_eg2.png", width = 3, height = 2)

### 2019-01-28 Code/fig_timing.R

@@ -1,12 +1,16 @@
 library(tidyverse)
 rm(list=ls())
 
-load("../TimingData/timing_comparison.RData")
-times_df <- times_df[-nrow(times_df),]    
-diff_df <- diff_df[-nrow(diff_df),]  
+times_all <- {}
+diff_all <- {}
+for (i in c("", "2", "3", "4")){
+  load(sprintf("../TimingData/timing_comparison%s.RData", i))
+  times_all <- bind_rows(times_all, times_df[-nrow(times_df),])  
+  diff_all <- bind_rows(diff_all, diff_df[-nrow(diff_df),])  
+}
 
 
-times_df %>% group_by(n, windows) %>%
+times_all %>% group_by(n, windows) %>%
   summarise(time_mean = mean(time), 
             se_time = sd(time)/sqrt(n())) %>%
   ggplot(aes(x=n, y=time_mean, group=factor(windows), col=factor(windows))) + 

### 2019-01-28 Manuscript/detrendify_JCGS.tex
Updated simulation study text

@@
 	\begin{figure}
-		\label{fig:trendfilter}
 		\centering
-		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}
+		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the .estimated trend is shown in red.}
 		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
 		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}
+		\label{fig:trendfilter}
 	\end{figure}
@@
 	\begin{figure}[b!]
-		\label{fig:raw_spod}
 		\caption{Example of 3 co-located SPod PID sensor readings.}
 		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
+		\label{fig:raw_spod}
 	\end{figure}
@@
 	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which we can think of as the trend in a low quantile of the data.
-\cite{Koenker1978} was the first to propose substituting the squared loss used in regression with the check loss function  (Eq.
+\cite{Koenker1978} were the first to propose substituting the squared loss used in regression with the check loss function  (Eq.
 \ref{eq1}) to estimate a conditional quantile instead of the conditional mean.
@@
 	\label{eq1}
-	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0))
+	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0)).
 	\end{equation}
@@
 
-	The number of parameters to be estimates is equal to the number of observations multiplied by the number of quantiles of interest.
+	The number of parameters to be estimated is equal to the number of observations multiplied by the number of quantiles of interest.
 As the size of the data and the number of quantiles grows, all solvers will eventually break.
@@
 	\subsection{Peak Detection}
-	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
-	\begin{equation}
-	Y(t) = s(t) + b(t) + \epsilon(t)
-	\end{equation} 
-	where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, \sigma^2)$ is an error component.
-We assume $t$ is a uniformly spaced sequence of time points between 0 and 1.
-We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from $n/50$ to $n/25$ with equal probability, and coefficients drawn from an exponential distribution with rate 1.
-The true signal function is assumed to be zero with Gaussian peaks.
-The number of peaks is sampled from $n/100$ to $n/50$ with equal probability with centers uniformly distributed between 0.1 and 0.9 and bandwidths uniformly distributed between $1/n$ and $5/n$ and areas uniformly distributed between 0 and $20/n$.
-One hundred datasets were generated for $n=\{300, 500, 1000, 5000\}$.
-We compare the methods ability to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the root mean squared error (RMSE).
-
 	\begin{figure}[h]
@@
 
-	In addition to the RMSE we also calculated the signal miss-classification rate.
-We first classified the ``true" peaks function into signal or not based on a threshold of 0.1.
-We then classified the detrended data using three different thresholds and calculated the fraction miss-classified.
+	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
+	\begin{equation}
+	Y(t) = s(t) + b(t) + \epsilon(t)
+	\end{equation} 
+	with $t = 1, ..., n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, \sigma^2)$ is an error component.
+We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from a Poisson distribution with mean parameter equal to $n/100$,  and coefficients drawn from an exponential distribution with rate 1.
+The true signal function is assumed to be zero with peaks generated using the Gaussian density function.
+The number of peaks is sampled from a binomial distribution with size equal to $n$ and probability equal to $0.005$ with location parameters uniformly distributed between $1$ and $n-1$ and bandwidths uniformly distributed between $2$ and $12$.
+The peaks were then multiplied by a factor that was randomly drawn from a normal distribution with mean 20 and standard deviation of 4.
+One hundred datasets were generated for $n=\{500, 1000, 2000, 4000\}$.
+We compare the ability of the methods to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (Fig.
+\ref{fig:peaks_rmse}).
+	\begin{figure}
+		\caption{RMSE by method, quantile and data size for peaks design.}
+		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}	
+		\label{fig:peaks_rmse}
+	\end{figure}
+
+	In our application, we want to accurately classify the observations into signal and no signal using a threshold.
+To evaluate the accuracy of our method compared to other methods we define true signal as any time when the simulated peak value is greater than 0.5.
+We compared three different trends for the baseline estimation and four different thresholds for separating the estimated signal after subtracting the estimated baseline from the observations.
 An illustration of the observations classified as signal after detrending compared to the ``true signal" is shown in Fig.
-9.
+\ref{fig:peaks_class_eg}.
+To compare the resulting signal classifications we calculate the class averaged accuracy (CAA).
+Defining $s(t) \in \{0,1\}$ as the true vector of signal classification and $\hat{s}(t) \in \{0,1\}$ as the estimated signal classification, the CAA is defined as
+	\begin{equation}
+	\mbox{CAA} = \frac{1}{2}\left(\frac{\sum_{t=1}^n \mathbf{I}[s(t) = 1 \cap \hat{s}(t)=1]}{\sum_{t=1}^n \mathbf{I}[s(t) = 1]} + \frac{\sum_{t=1}^n \mathbf{I}[s(t) = 0 \cap \hat{s}(t)=0]}{\sum_{t=1}^n \mathbf{I}[s(t) = 0]}\right).
+	\end{equation}  
+	We use this metric because our classes tend to be very un-balanced in this case with many more $0$s than $1$s.
+The CAA metric will always give a score of 0.5 for random guessing and also for trivial classifiers such as $\hat{s}(t) = 0$ for all $t$.
 
@@
 	\begin{figure}[h!]
-		\caption{Example signal classification using threshold. Red indicates true signal $>0.1$, blue indicates classified as signal after baseline removal using eBIC detrendr and a threshold of 1.}
+		\caption{Example signal classification using threshold.
+Red indicates true signal $>0.5$, blue indicates observations classified as signal after baseline removal using eBIC detrendr and a threshold of 1.2.}
 		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
+		\label{fig:peaks_class_eg}
 	\end{figure}

### 2019-01-28 Code/Fig_peaks_metrics.R

@@ -39,7 +39,8 @@ ggsave("../Manuscript/Figures/ex_baseline.png", width = 7, height = 4)
 simDesign <- "peaks"
 tau <- c(0.01, 0.05, 0.1)
 nSim <- 100
-
+colPal <- rev(c('#762a83','#9970ab','#c2a5cf',
+                '#a6dba0','#5aae61','#1b7837'))
 methods <- c("detrend_eBIC", "detrend_SIC", "detrend_valid", "qsreg", 
              "rqss", "npqw") 
 MSEs <- as.data.frame(matrix(NA, nrow = nSim*length(methods), 
@@ -77,7 +78,8 @@ summary_peaks <-
   ) %>%
   ungroup() %>%
   mutate(tau_fac = tau, 
-         tau = as.numeric(substr(tau_fac, 5, 10)))
+         tau = as.numeric(substr(tau_fac, 5, 10)), 
+         Method = factor(Method, levels = methods))
 
 summary_peaks %>% 
   ggplot( aes(x = factor(n), y = mean_mse, col = Method)) + 
@@ -86,7 +88,7 @@ summary_peaks %>%
                  position = position_dodge(width = 0.5))+
   facet_grid(.~factor(tau), scales = "free")+
   theme_bw() +
-  scale_color_brewer(palette = "Paired") +
+  scale_color_manual(values=colPal, breaks = methods) +
   labs(x = "", y="RMSE",  col = "Method") +
   ylim(c(0,1.2))
 ggsave("../Manuscript/Figures/peaks_mse.png", width = 10, height = 3)

### 2019-01-30 Manuscript/detrendify_JCGS.tex
manuscript updates

@@
 	{
-		\title{\bf Drift Removal for Time Series Data Using Quantile Trend Filtering}
+		\title{\bf Baseline Drift Estimation for Time Series Data Using Quantile Trend Filtering}
 		\author{Halley Brantley\thanks{
@@
 	\begin{abstract}
-		The text of your abstract.  200 or fewer words.
+		We address the problem of estimating smoothly varying baseline trends in time series data using quantile trend filtering.
+We extend the basic framework to ensure non-crossing while estimating multiple trends simultaneously and implement an alternating direction method of moments (ADMM) algorithm for series that cannot be processed simultaneously.
+Timing experiments demonstrate the usefulness of our ADMM algorithm even in cases where the series can be processed simultaneously.
+We also address smoothing parameter selection and propose a modified criteria based on the extended Bayesian Information Criteria.
+We demonstrate both through simulation studies and an application to low cost air quality sensor data that our model provides better or comparable estimates than existing methods of non-parametric quantile trends and improves signal classification of low-cost air quality sensor output.
 	\end{abstract}
@@
 	\noindent%
-	{\it Keywords:}  3 to 6 keywords, that do not appear in the title
+	{\it Keywords:} quantile regression, non-parametric, trend estimation, smoothing splines
 	\vfill
@@
 
-	In many applications spanning the fields of chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend other more rapidly varying components.
-Given observations $y(t)$, with $t=1, ..., N$, \cite{Kim2009} proposed using \textit{$\ell_1$ trend filtering} to estimate trends that are smooth in the sense of being piecewise linear or piecewise polynomial.
-\cite{Tib2014} then discovered that empirically the trend filtering estimates adapt to the local level of smoothness much better than the more common smoothing splines.
-In the trend filtering problem \citep{Kim2009, Tib2014}, the trend, $\theta \in \mathbb{R}^n$, is estimated by solving the following convex problem.
+	In applications spanning the fields of chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components.
+\cite{Kim2009} addressed the case when the series consisted of just a trend and a random noise component and proposed the use of \textit{$\ell_1$ trend filtering} to estimate trends that piecewise linear or piecewise polynomial.
+\cite{Tib2014} later demonstrated that empirically the trend filtering estimates adapt to the local level of smoothness better than the more common smoothing splines.
+In the trend filtering problem, given observations $y(t)$ with $t=1,...,N$, the trend, $\theta(t)$, is estimated by solving the following convex problem.
 	\begin{eqnarray}
@@
 	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
+
 	Trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise (Fig.
 \ref{fig:trendfilter}).
-In some cases, time series may also include a rapidly varying signal component in addition to the trend and noise, in these cases the trend filtering estimate over-estimate the trend in the places where signal is present (Fig.
+In some applications, however, time series may also include a rapidly varying signal component in addition to the trend and noise, in these cases the trend filtering estimate over-estimates the trend in the places where signal is present (Fig.
 \ref{fig:trendfilter}).
@@
 		\centering
-		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the .estimated trend is shown in red.}
+		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}
 		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
@@
 
-	One application in which the observed time series consists of a slowly varying trend, non-negative signal, and rapidly varying noise is the output of low cost air quality sensors.
+	One application in which the observed time series consists of a slowly varying baseline, non-negative signal, and rapidly varying noise is the output of low cost air quality sensors.
 The use of low-cost, portable, air quality sensors has increased dramatically in the last decade.
@@
 Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
-To reduce cost and power consumption of the SPod, the relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours (Fig.
-\ref{fig:raw_spod}).
+To reduce cost and power consumption of the SPod, the relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
+Fig.
+\ref{fig:raw_spod} provides an example of 3 co-located SPod sensor outputs during a proximate release on a test range.
+While all of the sensors response to the pollutant signal, the amount of baseline drift varies from one node to another.
 
@@
 
-	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which we can think of as the trend in a low quantile of the data.
-\cite{Koenker1978} were the first to propose substituting the squared loss used in regression with the check loss function  (Eq.
-\ref{eq1}) to estimate a conditional quantile instead of the conditional mean.
+	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which can be thought of as a trend in a low quantile of the data.
+\cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check loss function  (Eq.
+\ref{eq:checkloss}) to estimate a conditional quantile instead of the conditional mean.
 
 	\begin{equation}
-	\label{eq1}
+	\label{eq:checkloss}
 	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0)).
@@
 
-	A variety of approaches have been proposed for estimating quantile trends.
-The regression vector $\theta$ is estimated by finding the minimizer of $\rho_{\tau}(y-X\theta)$ where is the check-loss function and $\mathbf{I}$ is the indicator function.
-\cite{nychka1995nonparametric} estimated quantile splines by combining the smoothing spline penalty commonly used in non-parametric regression with the check-loss function used in qunatile regression:
+	Later, \cite{KoenkerNgPortnoy1994} addressed the issue of estimated quantile curves by combining the check loss penalty with trend filtering penalty with $k = 2$ which produces quantile trends that are piecewise linear.
+Other approaches to estimating quantile trends have also been proposed.
+Rather than using the trend filtering $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm solving the optimization problem in Eq.
+\ref{eq:smoothingspline}.
 
 	\begin{equation*}
-	\rho_{\tau}(Y - f) + \lambda\int (f''(t))^2 dt, 
+	\label{eq:smoothingspline}
+	\rho_{\tau}(y - f) + \lambda\int (f''(t))^2 dt, 
  	\end{equation*}
 
-	where $Y(t)$ is a response variable observed at time $t$, $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing.
+	where $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing.
+\cite{Oh2011} proposed an algorithm for solving this problem based on approximating the checkloss function with a similar differentiable function.
 
-	We propose to use the trend filtering penalty with the check loss function to produce a non-parametric quantile regression estimate for removing trends in time series.
+	\cite{Racine2017} take an entirely different approach to estimating quantile trends.
+They constrain the response to follow a smooth location scale model of the form $y(t) = a(t) + b(t)\epsilon_i$ and estimate the $\tau$\textsubscript{th} conditional quantile given $t = t_0$ using a kernel estimator and local linear approach.
+
+	We propose to use the trend filtering penalty combined with the check loss function to produce quantile trends that are piecewise quadratic.
 The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
 Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles.
-We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and proposed a modified criteria for choosing the smoothing parameter.
+We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and propose a modified criteria for choosing the smoothing parameter.
 We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.
@@
 
-	We combine the ideas of quantile regression and trend filtering, namely consider the case where the design $\mathbf{X}$ is the identity matrix.
+	We combine the ideas of quantile regression and trend filtering.
 For a single quantile level $\tau$ the estimation of the quantile trend filtering model can be posed as the following optimization problem.
@@
 We address the problem of choosing $\lambda$ in Section \ref{sec:lambda_choice}.
-As with the classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of free or commercial solvers.
+As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of free or commercial solvers.
 In many cases, including ours, we are interested in estimating multiple quantiles simultaneously.
@@
 	\lambda_j \lVert \mathbf{D}^{(k+1)} \theta_j \rVert_1 \right ] \\
-	 \text{subject to: }\; \theta_{1i} \le \theta_{2i} \le ... \le \theta_{Ji} \text{ for all } i,
+	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le ... \le \theta_{J}(t) \text{ for all } t,
 	\end{eqnarray}
@@
 The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
-In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi}.
-
-	The number of parameters to be estimated is equal to the number of observations multiplied by the number of quantiles of interest.
+In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation, but it can easily be substituted for a free solver such as the Rglpk package by \cite{rglpk}.
+The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
 As the size of the data and the number of quantiles grows, all solvers will eventually break.
@@
 We propose an alternating direction method of multipliers (ADMM) algorithm for solving large problems in a piecewise fashion.
-The ADMM algorithm is fully described by \cite{boyd2011distributed, gabay1975dual, glowinski1975approximation}.
+The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is fully described by \cite{boyd2011distributed}.
 We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem given in Eq.
-\ref{eq:quantile_trend}, by dividing our observed series $y(t)$ with $t = \{1, ..., N\}$ into overlapping windows
+\ref{eq:quantile_trend}, by dividing our observed series $y(t)$ with $t = \{1, ..., n\}$ into overlapping windows
 
 	\begin{align*}
-	\begin{cases}
-	y_1(t) = y(t) & \mbox{if~~} 1 \le t \le u_{1}\\
-	y_2(t) = y(t) & \mbox{if~~} l_{2} \le t \le u_{2} \\
-	y_3(t) = y(t) & \mbox{if~~} l_{3} \le t \le u_{3} \\
+	y_1(t) = y(t) & \mbox{~~if~~} 1 \le t \le u_{1}\\
+	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
 	\cdots & \\
-	y_M(t) = y(t) & \mbox{if~~} l_{M} \le t \le  N\\
-	\end{cases}
+	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n\\
 	\end{align*}
 
-	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< N$.
+	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< n$.
+An illustration is given in Fig.
+\ref{fig:windows}.
 Given quantiles $\tau_1 < ...
 < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$.
-In order to write out the constraint the that overlapping sections must be equal we define a consensus variable
+In order to write out the constraint that the overlapping sections must be equal we define a consensus variable
 
-	 \begin{align}
+	 \begin{align*}
 		 \bar{\theta}_{j,m} =&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
@@
 			 \end{cases},
-	\end{align}
+	\end{align*}
 	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
-	 \begin{align}
+	 \begin{align*}
 		 \label{eq:quantile_windows}
@@
 		 \lambda_j \lVert \mathbf{D}^{(k+1)} \theta_{j,m} \rVert_1 \right ] \\
-		 &\text{subject to: }\; \theta_{1,m}(t) \le \theta_{2,m}(t) \le ... \le \theta_{J,m}(t) \text{ for all } m,t \\
-		 &\text{subject to: }\; \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
-	 \end{align}
-	 Given values for the Lagrange multiplier $\omega_{j,m}$ and the consensus variable $\bar{\theta}_{j,m}$ we can write the augmented Lagrangian for finding the trends in window $m$:
-	 \begin{equation}
+		 &\text{subject to: }\; \\
+		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le ... \le \theta_{J,m}(t) \text{ for all } m,t \\
+		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
+	 \end{align*}
+	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$ we can write the augmented Lagrangian for finding the trends in window $m$:
+	 \begin{equation*}
 	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \mathbf{D}^{(k+1)}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
 	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
-	 \end{equation}
-
-	 We then estimate the trend separately in each window while constraining the overlapping pieces of the trends to be equal according the algorithm 1.
+	 \end{equation*}
+	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in algorithm 1.
 
@@
 		\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
+		\label{fig:windows}
 	\end{figure}
@@
 
-	We used the stopping criteria described by \cite{boyd2011distributed}.
+	We measure convergence use the stopping criteria described by \cite{boyd2011distributed}.
 The criteria are based on the primal and dual residuals which represent the residuals for the primal and dual feasibility, respectively.
-The primal and dual residuals are defined as
-
+The primal residual,
+	\begin{equation}
+	r_p^{(q)} = \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
+	\end{equation}
+	represents the difference between the trend values in the windows and the consensus trend value while the dual residual,
+	\begin{equation*}
+	r_d^{(q)} = \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
+	\end{equation*}
+	represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when 
 	\begin{align}
-	&r_p^{(q)} = \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2}\\
-	&r_d^{(q)} = \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2}.
-	\end{align}
-
-	The primal residual, $r_p$, represents the difference between the trend values in the windows and the consensus trend value while the dual residual, $r_d$s represents the change in the consensus variable from one iterate to the next.
-The algorithm is stopped when
-
-	\begin{align}
-		&r_p^{(q)} < \epsilon_{abs}\sqrt{NJ} + \epsilon_{rel}\underset{m}{\max}\left[\max 
+		&r_p^{(q)} < \epsilon_{abs}\sqrt{nJ} + \epsilon_{rel}\underset{m}{\max}\left[\max 
 		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
-		&r_d^{(q)} < \epsilon_{abs}\sqrt{NJ} + \epsilon_{rel}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
+		&r_d^{(q)} < \epsilon_{abs}\sqrt{nJ} + \epsilon_{rel}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
 	\end{align}
@@
 	Timing experiments illustrate the advantages of using our ADMM algorithm even on datasets where solving the problem simultaneously is possible.
+For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below and trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using a $\lambda = n/5$.
 We use from one to four windows for each data size with an overlap of 500.
 The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
-For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below and trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using a $\lambda = n/5$.
 
@@
 	\label{sec:lambda_choice}
-	Our method can easily handle missing data by changing the check loss function to output 0 for missing values.
-This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$ and to compare method performance on real data.
-A number of methods have been proposed for selecting the quantile regression smoothing spline tuning parameter \cite{yuan2006gacv}.
-\cite{KoenkerNgPortnoy1994} relate $\lambda$ to the number of interpolated points $p_{\lambda} = \sum I(y_i = \widehat{g}_i(x_i))$, which can be thought of as active knots, they propose the Schwarz criterion for the selection of $\lambda$
+	An important problem in trend estimation is the choice of regularization parameter or degree of smoothness.
+Our method can easily handle missing data by defining the check loss function to output 0 for missing values.
+This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$.
+However, the use of an information criteria metric can result in a better choice of regularization parameter than the validation method.
+\cite{KoenkerNgPortnoy1994} addressed the choice of regularization parameter by proposing the Schwarz criterion for the selection of $\lambda$
 	\begin{equation}
-	\mbox{SIC}(p_{\lambda}) = \log[\frac{1}{n}\rho_{\tau}(y - \widehat{g}(x))] + \frac{1}{2n}p_{\lambda}\log n
+	\label{eq:SIC}
+	\mbox{SIC}(p_{\lambda}) = \log\left[\frac{1}{n}\rho_{\tau}(y - \theta)\right] + \frac{1}{2n}p_{\lambda}\log n.
 	\end{equation}
+	where $p_{\lambda} = \sum_t I(y(t) = \widehat{\theta}(t))$ is the number of interpolated points, which can be thought of as active knots.
+The SIC is based on the traditional Bayesian Information Criterion (BIC) which is given by
+	\begin{equation}
+	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}\}) + \nu\log n 
+	\end{equation}	
+	where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\hat{\theta}$.
+If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and assume that minimizing the checkloss function corresponds to maximizing the asymmetric Laplace likelihood, \begin{equation}
+	L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_t\rho_\tau\left(\frac{y(t) - \theta(t)}{\sigma}\right)\right\},
+	\end{equation} 
+	we can compute the BIC as 
+	\begin{equation}
+	\mbox{BIC}(df) = 2\frac{1}{\sigma}\rho_{\tau}(y-\hat{\theta}) + \mbox{df}\log n
+	\end{equation} 
+	where $df$ is the number of non-zero elements of $D^{(k+1)}\hat{\theta}$. We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates.
 
-	The traditional Bayesian Information Criterion (BIC) is given by 
+	Another criteria, the extended Bayesian Information Criteria (eBIC), specifically designed for large parameter spaces was proposed by \cite{chen2008}. 
 	\begin{equation}
-	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n 
+	\label{eq:eBIC}
+	\mbox{eBIC}_{\gamma}(s) = -2\log(L\{\hat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
 	\end{equation}
-	where $\theta(s)$ is the parameter $\theta$ with those components outside $s$ being set to 0, and $\nu(s)$ is the number of components in $s$.
-If we assume an asymmetric Laplace likelihood $L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_i\rho_\tau(\frac{y_i - \theta_i}{\sigma})\right\}$ and the number of non-zero elements of $D\theta$ as $df$
-	\begin{equation}
-	\mbox{BIC}(df) = 2\frac{1}{\sigma}\rho_{\tau}(y-\theta) + \mbox{df}\log n
-	\end{equation} 
-	We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates. \cite{chen2008} proposed the extended BIC for large parameter spaces 
-	\begin{equation}
-	\mbox{BIC}_{\gamma}(s) = -2\log(L\{\hat{\theta}(s)\}) + \nu(s)\log n  + 2\gamma\log{P\choose j,}~~\gamma \in [0,1]
-	\end{equation}
-	where $P$ is the total number of possible parameters and $j$ is the number of parameters included in given model.
-We used this criteria with $\gamma = 1$, $P=n-k$ where $k$ is the order of the differencing matrix and $j = \nu(s)$ is the number of non-zero entries in $D^{(k)}\theta$.
-
-	We used a single dataset to illustrate the difference between the scaled, unscaled and extended BIC criteria.
+	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given model.
+We used this criteria with $\gamma = 1$, and $P=n-k-1$.
+We use a single dataset from our simulation study to illustrate the difference between the scaled, unscaled ($\sigma = 1$) and scaled extended BIC criteria in Fig.
+\ref{fig:BIC}.
 
 	\begin{figure}[h!]
-		\caption{Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.} 
+		\caption{(Left) Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.
+(Middle) Estimated 10th quantile trend with regularization parameter chosen using various criterion.
+(Right) Criterion values by $\log(\lambda)$ vertical lines indicate locations of minima.}
 		\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
@@
 		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png} 
+		\label{fig:BIC}
 	\end{figure}
@@
 
+	We conduct two simulation studies to compare the performance of our quantile trend filtering method and regularization parameter selection criteria with previously published methods.
+The first study compares the method's ability to estimate quantiles when the only components of the observed series are a smooth trend and a random component.
+The second study is based on our application and compares the method's ability to estimate baseline trends and enable peak detection when the time series contains a non-negative signal component in addition to the trend and random component.
+
+	We compare the performance of our quantile trend filtering method with the three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011, nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
+The regularization parameter $\lambda$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria as described in \cite{KoenkerNgPortnoy1994}, the regularization parameter for \texttt{qsreg} was chosen using generalized cross-validation based on the quantile criterion \cite{Oh2011}.
+
+	We also compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC (Eq.
+\ref{eq:SIC}) \citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria (Eq.
+\ref{eq:eBIC}).
+
 	\subsection{Estimating Quantiles}
-
-	We compare the performance of our quantile trend filtering method with the three previously published methods using designs proposed by \cite{Racine2017}.
-The methods compared are: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
-The smoothing parameter $\lambda$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria as described in \cite{KoenkerNgPortnoy1994}.
-We further compare three criteria for choosing the smoothing parameter for our detrend method: \texttt{detrendr\_SIC}: Our method where we minimize $\sum_i\rho_{\tau}(y_i - \theta_i) + \lambda||D\theta||_1$ and $\lambda$ is chosen using SIC \citep{KoenkerNgPortnoy1994}.
-\texttt{detrendr\_valid}: Our method where lambda is chosen by leaving out every 5th observation as a validation data set and minimizing the evaluating the check loss function evaluated at the validation data.
-\texttt{detrendr\_eBIC}:  the new criteria we have proposed based on the extended BIC proposed by \cite{chen2008}.
-
-	Three simulation designs from \cite{Racine2017} were considered. For all designs $X_i$ was generated as a uniformly spaced sequence in $[0,1]$ and the response $Y$ was generated as 
-	$$Y_i = sin(2\pi x_i) + \epsilon_i(x_i)$$
-	The three error distributions considered were 
-	\begin{itemize}
-		\item Gaussian: $\epsilon_i(x_i) \sim N\left(0, \left(\frac{1+x_i^2}{4}\right)^2\right)$
-		\item Beta: $\epsilon_i \sim Beta(1, 11-10x_i)$
-		\item Mixed normal: $\epsilon_i$ is simulated from a mixture of $N(-1,1)$ and  $N(1,1)$ with mixing probability $x_i$.
-	\end{itemize}
 	\begin{figure}
-		\caption{Simulated data with true quantiles $\tau \in \{0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99\}$}	
+		\caption{Simulated data with true quantile trends.}	
 		\includegraphics[width=.28\linewidth]{Figures/gaus.png}
@@
 
-	100 datasets were generated of sizes 300, 500 and 1000.
-The MSE was calculated as $\frac{1}{n}\sum_i (\hat{q}_{\tau}(x_i) - q_\tau(x_i))^2$.
-The plots below show the mean MSE $\pm$ twice the standard error by method, quantile level and sample size.
+	To compare performance in estimating quantile trends, three simulation designs from \cite{Racine2017} were considered.
+For all designs $t = 1, ..., n$,  $x(t) = t/n$, and the response $y$ was generated as
+	$$y(t) = sin(2\pi x(t)) + \epsilon(x(t))$$
+	The three error distributions considered were 
+	\begin{itemize}
+		\item Gaussian: $\epsilon(x(t)) \sim N\left(0, \left(\frac{1+x(t)^2}{4}\right)^2\right)$
+		\item Beta: $\epsilon(x(t)) \sim Beta(1, 11-10x(t))$
+		\item Mixed normal: $\epsilon(x(t))$ is simulated from a mixture of $N(-1,1)$ and  $N(1,1)$ with mixing probability $x(t)$.
+	\end{itemize}
+
+	One hundred datasets were generated of sizes 300, 500 and 1000.
+For each method quantile trends were estimated for $\tau = \{0.05, 0.25, 0.5, 0.75, 0.95\}$.
+Only our detrend methods guarantee non-crossing quantiles.
+For each quantile trend and method the root mean squared error was calculated as $\mbox{RMSE} = \sqrt{\frac{1}{n}\sum_t (\hat{q}_{\tau}(t) - q_\tau(t))^2}$.
+The mean RMSE $\pm$ twice the standard error for each method, quantile level and sample size is shown in Figure \ref{fig:quantile_mse}.
+	In all three designs the proposed detrend methods are either better than or comparable to existing methods.
+Overall the \texttt{detrend\_eBIC} performs best, and especially in the mixed normal design our methods have lower RMSEs for the 5\textsuperscript{th} and 95\textsubscript{th} quantiles.
+The \texttt{npqw} method performs particularly poorly in the mixed normal design due to the fact that it assumes the data comes from a scale-location model which is violated in this case.
 
 	\begin{figure}
-		\caption{RMSE by design, method, quantile and data size.}
+		\caption{RMSE by design, method, quantile and data size. Points and error bars represent mean RMSE $\pm$ twice the standard error.}
 		\includegraphics[width=\linewidth]{Figures/gaus_mse.png}	
@@
 		\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
+		\label{fig:quantile_mse}
 	\end{figure}
-	In all of the designs the proposed detrend methods are either better than or comparable to existing methods.
-The npqw method performs particularly poorly in the mixed normal design, due to the fact that it assumes the data comes from a scale-location model which is violated in this case.
 
 	\subsection{Peak Detection}
-	\begin{figure}[h]
-		\caption{Example of simulated peaks, baseline, and observed measurements.}
-		\includegraphics[width = \linewidth]{Figures/ex_peaks.png}
-	\end{figure}
-
+%	\begin{figure}[h]
+%		\caption{Example of simulated peaks, baseline, and observed measurements.}
+%		\includegraphics[width = \linewidth]{Figures/ex_peaks.png}
+%	\end{figure}
 	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
@@
 	\end{equation} 
-	with $t = 1, ..., n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, \sigma^2)$ is an error component.
-We generate $b(t)$ using a cubic natural spline basis function with degrees of freedom sampled from a Poisson distribution with mean parameter equal to $n/100$,  and coefficients drawn from an exponential distribution with rate 1.
+	with $t = 1, ..., n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, 0.25^2)$ is an error component.
+We generate $b(t)$ using cubic natural spline basis functions with degrees of freedom sampled from a Poisson distribution with mean parameter equal to $n/100$,  and coefficients drawn from an exponential distribution with rate 1.
 The true signal function is assumed to be zero with peaks generated using the Gaussian density function.
@@
 \ref{fig:peaks_rmse}).
+In this case our \texttt{detrend\_eBIC} method outperforms the others substantially, the \texttt{qsreg} method is comparable on the smaller datasets but its performance deteriorates as the data size grows.
+The \texttt{npqw} and \texttt{detrend\_valid} methods both perform poorly on this design.
+
 	\begin{figure}
@@
 To evaluate the accuracy of our method compared to other methods we define true signal as any time when the simulated peak value is greater than 0.5.
-We compared three different trends for the baseline estimation and four different thresholds for separating the estimated signal after subtracting the estimated baseline from the observations.
-An illustration of the observations classified as signal after detrending compared to the ``true signal" is shown in Fig.
+We compared three different quantiles for the baseline estimation and four different thresholds for separating the estimated signal after subtracting the estimated baseline from the observations.
+An illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal" is shown in Fig.
 \ref{fig:peaks_class_eg}.
 To compare the resulting signal classifications we calculate the class averaged accuracy (CAA).
-Defining $s(t) \in \{0,1\}$ as the true vector of signal classification and $\hat{s}(t) \in \{0,1\}$ as the estimated signal classification, the CAA is defined as
+Defining $s(t) \in \{0,1\}$ as the vector of true signal classification and $\hat{s}(t) \in \{0,1\}$ as the estimated signal classification, the CAA is defined as
 	\begin{equation}
@@
 	\end{equation}  
-	We use this metric because our classes tend to be very un-balanced in this case with many more $0$s than $1$s.
+	We use this metric because our classes tend to be very un-balanced with many more $0$s than $1$s.
 The CAA metric will always give a score of 0.5 for random guessing and also for trivial classifiers such as $\hat{s}(t) = 0$ for all $t$.
 
-	Our detrend\_BIC method performs the best overall in terms of both RMSE and miss-classification rate.
-The lowest miss-classification rates were obtained using the detrend\_eBIC method and a threshold of 1 for all data sizes.
-While qsreg was competitive with our method in some cases, both the RMSE and miss-classification rate increased substantially with the size of the dataset.
+	Our detrend\_BIC method performs the best overall in terms of both RMSE and CAA.
+While \texttt{qsreg} was competitive with our method in some cases, in the majority of cases the largest CAA values for each threshold were produced using the \texttt{detrend\_eBIC} method with the 1\textsuperscript{st} or 5\textsuperscript{th} quantiles.
 
@@
 
-	We compared our detrend method with the eBIC criteria with the qsreg method on a subset of the real data since the qsreg method cannot handle all 15 hours simultaneously.
-We estimate the baseline trend using both the 5\textsuperscript{th}, 10\textsuperscript{th}, and 15\textsuperscript{th} quantiles.
-We compare three thresholds for separating signal from not signal, the thresholds are calculated using the median and a factor multiple of the median absolute deviation of the detrended series.
-We report both the confusion matrices and the variation of information (VI) distances.
+	The low-cost ``SPod" sensors that are being investigated by researchers at the U.S.
+Environmental Protection Agency output a time series that includes a slowly varying baseline, high frequency random noise, and the sensor response to VOCs proximity.
+A potential use for these sensors is to monitor pollutant concentrations at fenceline facilities and detect time points when high concentrations are present.
+Ideally 3 co-located and time aligned sensors (as shown in Fig.
+\ref{fig:raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
+
+	We compare our \texttt{detrend\_eBIC} method with the \texttt{qsreg} method on a subset of the real data (n=5000) since the \texttt{qsreg} method cannot handle all 15 hours simultaneously.
+We estimate the baseline trend using the 5\textsuperscript{th}, 10\textsuperscript{th}, and 15\textsuperscript{th} quantiles and compare three thresholds for classifying signal.
+The thresholds are calculated using the median plus a multiple of the median absolute deviation (Eq.
+\ref{eq:MAD}) of the detrended series.
+	\begin{equation}
+	\label{eq:MAD}
+	\mbox{MAD} = \mbox{median}(y-\tilde{y})
+	\end{equation}
+	where $\tilde{y}$ is the median of $y$.
+Given a method, quantile level and MAD multiple, we estimate the quantile trend for each of the sensor nodes and subtract it from the observations.
+We then calculate the threshold using the median plus the multiple times MAD of the corrected series and classify the corrected series based on the threshold.
+Fig.
+\ref{fig:rugplot} shows the time series used after estimating the baseline using the \texttt{detrend\_eBIC} estimate of the 10\textsuperscript{th} quantile and a threshold with an MAD multiple of 2.
+
+ 	\begin{figure}
+	 	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 10th quantile.}
+	 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
+	 	\label{fig:rugplot}
+	 \end{figure}
+
+	Given the signal classifications for node g, $s_g(t) \in \{0,1\}$ and node f, $s_f(t)\in\{0,1\}$ we want to compare the similarity between the two classifications.
+One metric for evaluating the distance between two classifications is the variation of information (VI):
+	\begin{align}
+		r_{ij} =& \frac{1}{n}\sum_t \mathbf{I}(s_f(t) = i  \cap s_g(t) = j)\\
+		VI(s_f, s_g) =& -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_f(t) = i)}\right) + 
+		\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_g(t) = j)}\right) \right]
+	\end{align}	
+	The VI is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
+The VIs by method, quantile and trend are shown in Fig.
+\ref{fig:vi}.
+Using the smallest threshold the methods perform similarly although the lowest VIs are observed for the 15\textsuperscript{th} quantile using the \texttt{detrend\_eBIC} method.
+For the larger thresholds the \texttt{detrend\_eBIC} method produces substantially lower VI distances between nodes than the \texttt{qsreg} method.
 
@@
 		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
+		\label{fig:vi}
 	\end{figure}
 
-	\begin{figure}
-		\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 10th quantile.}
-		\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
-	\end{figure}
-
-	Our windowed detrend method was used to removed the baseline drift from low cost air quality sensors so that the signal could be categorized using a simple threshold.
-The measurements were first standardized to have mean zero and variance 1.
-Three quantile levels for estimating the baseline trend were compared.
-The total dataset consisted of 52,322 observations per node.
-	The signal thresholds were set using the first 15,000 observations where it was known no signal was present.
-The thresholds were set as 3 times the standard deviation plus the mean of observations in this time period.
-The total number of seconds of signal for each node as well as the number of seconds where multiple nodes both reported signal is shown in Table 1.
-Table 2 shows the fraction of observations with different signal classifications by node combination.
-The BVI scores for the full dataset were 0.14, 0.13, and 0.14 for nodes f and g, f and h, and g and h, respectively.
+	Our windowed \texttt{detrend\_eBIC} method was used to removed the baseline drift from the total dataset consisting of 52,322 observations per node.
+The VI scores for the full dataset were 0.14, 0.13, and 0.14 for nodes f and g, f and h, and g and h, respectively.
+The complete confusion matrices for the nodes using a MAD multiple of 3 for the threshold is given in Table 1.
 
@@
 
-	\section{Conclusion}
+	\section{Conclusion and Discussion}
+	We have expanded the ideas of quantile trend filtering and demonstrated it's utility in both simulations and applied settings.
+Our ADMM algorithm for large series both reduces the computing time and allows trends to be estimated on series that cannot be estimated simultaneously while our scaled extended BIC criteria was shown to provided better estimated of quantile trends in series with and without a signal component.
+We have also shown that the baseline drift in low cost air quality sensors can be removed through estimating quantile trends.
+	In the future quantile trend filtering could be extended to observations measured at non-uniform spacing by incorporating the distance in covariate spacing into the differencing matrix.
+It could further be extended to estimate smooth spatial trends by adjusting the differencing matrix accordingly.
+
 	\label{sec:conc}
@@
 
-		\item[R-package for detrend routine:] R-package detrendr containing code to perform the diagnostic methods described in the article.
-The package also contains all datasets used as examples in the article.
-(GNU zipped tar file)
+		\item[R-package for detrend routine:] R-package detrendr containing code to perform the diagnostic methods described in the article. (GNU zipped tar file)
 
 	\end{description}
-
-	\section{References}
 

### 2019-01-30 Code/Fig_peaks_metrics.R

@@ -81,6 +81,8 @@ summary_peaks <-
          tau = as.numeric(substr(tau_fac, 5, 10)), 
          Method = factor(Method, levels = methods))
 
+text_size <- 14
+
 summary_peaks %>% 
   ggplot( aes(x = factor(n), y = mean_mse, col = Method)) + 
   geom_point(position = position_dodge(width = 0.5)) +
@@ -89,9 +91,12 @@ summary_peaks %>%
   facet_grid(.~factor(tau), scales = "free")+
   theme_bw() +
   scale_color_manual(values=colPal, breaks = methods) +
+  theme(text = element_text(size=text_size), 
+        axis.text.x = element_text(size=(text_size-5)),
+        plot.title = element_text(size = text_size)) +
   labs(x = "", y="RMSE",  col = "Method") +
   ylim(c(0,1.2))
-ggsave("../Manuscript/Figures/peaks_mse.png", width = 10, height = 3)
+ggsave("../Manuscript/Figures/peaks_mse.png", width = 7, height = 2.5)
 
 
 

### 2019-01-30 Code/application_full.R

@@ -12,12 +12,28 @@ spod <- read.csv("../SPod/fhrdata_2017-11-30.csv",
 spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
                                  format= "%m/%d/%Y %H:%M:%S")) 
 
-window_size <- 5000
-overlap <- 500
-max_iter <- 20
+window_size <- 10000
+overlap <- 1000
+max_iter <- 30
 tau <- c(0.05, 0.1, 0.15)
 k <- 3
 spod_trends <- data.frame(time = spod$time)
+
+max_iter <- 100
+result <- get_trend_windows(spodNode$pid[10001:28000], tau, 
+                          lambda = 2e5,
+                          k, window_size, overlap,
+                          max_iter = max_iter, 
+                          update = 1,
+                          eps_abs = 0.005,
+                          rho = 20)
+
+result0 <- get_trend(spodNode$pid[30001:33100], tau, 
+                     lambda = 1e4,
+                     k)
+plot(spodNode$pid[30001:33100], type="l")
+lines(result0[,3])
+lines(result[,3], col="red")
 
 
 for (node in c("f", "g", "h")){
@@ -25,15 +41,16 @@ for (node in c("f", "g", "h")){
   spodNode <- spod[, c("time", pidCol)]
   names(spodNode)[2] <- c("pid")
   spodNode$pid <- spodNode$pid/1000
-  result <- get_windows_BIC(spodNode$pid, tau, k, window_size, overlap,
-                          lambdaSeq = window_size^seq(1.1, 1.8, length.out=10)[10],
+  result <- get_windows_BIC(spodNode$pid[30000:40000], tau, k, window_size, overlap,
+                          lambdaSeq = window_size^seq(1.8, 2.3, length.out=4)[2],
                           df_tol = 1e-9,
                           gamma = 1,
-                          plot_lambda = FALSE,
+                          plot_lambda = TRUE,
                           solver = NULL,
                           criteria = "eBIC", 
                           max_iter = max_iter, 
-                          rho = 10)
+                          eps_abs = 0.01,
+                          rho = 15)
   save(result, file=sprintf("../SPod/node_%s_trend.RData", node))
   spod_trends <- cbind(spod_trends, as.data.frame(result$trend))
   names(spod_trends)[(ncol(spod_trends)-length(tau)+1):ncol(spod_trends)] <-

### 2019-01-30 Code/application_short.R

@@ -23,10 +23,10 @@ spodPID$time <- spod$time
 
 spodPIDs <- spodPID %>% 
   filter(time > as.POSIXct("2017-11-30 9:30:00"),
-         time <= as.POSIXct("2017-11-30 9:30:00")+5000)
+         time <= as.POSIXct("2017-11-30 9:30:00")+10000)
 plot(spodPIDs$h, type="l")
 
-tau <- c(0.05, 0.1, 0.15)
+tau <- c(0.15)
 k <- 3
 detrendr_trends <- data.frame(time = spodPIDs$time)
 qsreg_trends <- data.frame(time = spodPIDs$time)
@@ -35,7 +35,7 @@ x <- seq(1, nrow(spodPIDs), 1)
 for (node in c("f", "g", "h")){
   spodPIDs[, node] <- na.locf(spodPIDs[, node])
   result <- get_trend_BIC(spodPIDs[, node], tau, k, 
-                            lambdaSeq = n^seq(1.1, 1.8, length.out=10),
+                            lambdaSeq = exp(seq(5,12,.5)),
                             df_tol = 1e-9,
                             gamma = 1,
                             plot_lambda = TRUE,
@@ -57,6 +57,6 @@ for (node in c("f", "g", "h")){
     paste(node, tau, sep = "_")
 }
 
-save(spodPIDs, detrendr_trends, qsreg_trends, tau,
-     file = "../SPod/trends_short.RData")
+# save(spodPIDs, detrendr_trends, qsreg_trends, tau,
+#      file = "../SPod/trends_short.RData")
 

### 2019-01-30 Code/detrendr/R/update_windows.R

@@ -18,6 +18,9 @@ update_windows <- function(w_list, phiBar_list, etaBar_list,
                        rho=rho, nT=nT, SIMPLIFY = FALSE)
 
   phi_list <- lapply(model_list, solve_model, solver=solver, trend = FALSE)
+  # If solver didn't succeed use value of phiBar
+  phi_list0 <- mapply(function(x,y) {if(!any(x != 0)){return(y)}else{return(x)}}, 
+                     phi_list, phiBar_list)
   return(phi_list)
 }
 
@@ -30,8 +33,8 @@ update_windows <- function(w_list, phiBar_list, etaBar_list,
 #' @param overlapInd Indices of overlap between windows
 #' @export
 update_consensus <- function(phi_list, windows, overlapInd){
-
-  phiBar <- matrix(0, nrow = nrow(windows), ncol = ncol(phi_list[[1]]))
+  nT <- dim(phi_list[[1]])[2]
+  phiBar <- matrix(0, nrow = nrow(windows), ncol = nT)
   n_windows <- length(phi_list)
   
   for (i in 1:n_windows){

### 2019-01-30 Code/fig_timing.R

@@ -12,7 +12,8 @@ for (i in c("", "2", "3", "4")){
 
 times_all %>% group_by(n, windows) %>%
   summarise(time_mean = mean(time), 
-            se_time = sd(time)/sqrt(n())) %>%
+            se_time = sd(time)/sqrt(n()), 
+            ct = n()) %>%
   ggplot(aes(x=n, y=time_mean, group=factor(windows), col=factor(windows))) + 
   geom_line() + 
   geom_linerange(aes(ymin = time_mean - 2*se_time, ymax = time_mean + 2*se_time))+

### 2019-01-30 Code/timing.R

@@ -41,7 +41,7 @@ for (n in data_lengths){
                                  overlap = overlap, max_iter = 20, 
                                  update = 10, rho = 3, eps_abs = .01, 
                                  eps_rel = 1e-3, 
-                                 quad = TRUE, use_gurobi = TRUE), 
+                                 use_gurobi = TRUE), 
             times = 1)$time*1e-9
         diff_df$n[j] <- n
         diff_df$t[j] <- t

### 2019-02-01 Manuscript/detrendify_JCGS.tex
new application figures

@@
 Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
-To reduce cost and power consumption of the SPod, the relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
+To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
 Fig.
-\ref{fig:raw_spod} provides an example of 3 co-located SPod sensor outputs during a proximate release on a test range.
+\ref{fig:raw_spod} provides an example of 3 co-located SPod sensor outputs from near an industrial facility.
 While all of the sensors response to the pollutant signal, the amount of baseline drift varies from one node to another.
@@
 The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
+As is shown in Fig.
+\ref{fig:timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
+The timing experiments were conducted on
 
@@
 		\includegraphics[width = 0.7\linewidth]{Figures/Fig_timing_experiment.png}
+		\label{fig:timing}
 	\end{figure}
@@
 
-	We compare the performance of our quantile trend filtering method with the three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011, nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
+	We compare the performance of our quantile trend filtering method with three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011, nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
 The regularization parameter $\lambda$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria as described in \cite{KoenkerNgPortnoy1994}, the regularization parameter for \texttt{qsreg} was chosen using generalized cross-validation based on the quantile criterion \cite{Oh2011}.
@@
 The number of peaks is sampled from a binomial distribution with size equal to $n$ and probability equal to $0.005$ with location parameters uniformly distributed between $1$ and $n-1$ and bandwidths uniformly distributed between $2$ and $12$.
-The peaks were then multiplied by a factor that was randomly drawn from a normal distribution with mean 20 and standard deviation of 4.
-One hundred datasets were generated for $n=\{500, 1000, 2000, 4000\}$.
-We compare the ability of the methods to estimate the true quantiles of $b(t) + \epsilon$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (Fig.
+The simulated peaks were multiplied by a factor that was randomly drawn from a normal distribution with mean 20 and standard deviation of 4.
+One hundred datasets were generated for each $n=\{500, 1000, 2000, 4000\}$.
+We compare the ability of the methods to estimate the true quantiles of $Y(t)-s(t)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (Fig.
 \ref{fig:peaks_rmse}).
-In this case our \texttt{detrend\_eBIC} method outperforms the others substantially, the \texttt{qsreg} method is comparable on the smaller datasets but its performance deteriorates as the data size grows.
+In this simulation study our \texttt{detrend\_eBIC} method outperforms the others substantially.
+The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets but its performance deteriorates as the data size grows.
 The \texttt{npqw} and \texttt{detrend\_valid} methods both perform poorly on this design.
@@
 	In our application, we want to accurately classify the observations into signal and no signal using a threshold.
-To evaluate the accuracy of our method compared to other methods we define true signal as any time when the simulated peak value is greater than 0.5.
-We compared three different quantiles for the baseline estimation and four different thresholds for separating the estimated signal after subtracting the estimated baseline from the observations.
+To evaluate the accuracy of our method compared to other methods we define true signal as any time point when the simulated peak value is greater than 0.5.
+We compare three different quantiles for the baseline estimation and four different thresholds for classifying the signal after subtracting the estimated baseline from the observations.
 An illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal" is shown in Fig.
@@
 
-	Our detrend\_BIC method performs the best overall in terms of both RMSE and CAA.
+	Our \texttt{detrend\_BIC} method performs the best overall in terms of both RMSE and CAA.
 While \texttt{qsreg} was competitive with our method in some cases, in the majority of cases the largest CAA values for each threshold were produced using the \texttt{detrend\_eBIC} method with the 1\textsuperscript{st} or 5\textsuperscript{th} quantiles.
@@
 		\caption{Example signal classification using threshold.
-Red indicates true signal $>0.5$, blue indicates observations classified as signal after baseline removal using eBIC detrendr and a threshold of 1.2.}
+Red indicates true signal $>0.5$, blue indicates observations classified as signal after baseline removal using \texttt{detrend\_eBIC} and a threshold of 1.2.}
 		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
@@
 
-	The low-cost ``SPod" sensors that are being investigated by researchers at the U.S.
-Environmental Protection Agency output a time series that includes a slowly varying baseline, high frequency random noise, and the sensor response to VOCs proximity.
-A potential use for these sensors is to monitor pollutant concentrations at fenceline facilities and detect time points when high concentrations are present.
-Ideally 3 co-located and time aligned sensors (as shown in Fig.
+	The low-cost ``SPod" air quality sensors output a time series that includes a slowly varying baseline, high frequency random noise, and the sensor response to pollutants.
+A potential use for these sensors is to monitor pollutant concentrations at the fence lines of industrial facilities and detect time points when high concentrations are present.
+Ideally, three co-located and time aligned sensors (as shown in Fig.
 \ref{fig:raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
 
-	We compare our \texttt{detrend\_eBIC} method with the \texttt{qsreg} method on a subset of the real data (n=5000) since the \texttt{qsreg} method cannot handle all 15 hours simultaneously.
-We estimate the baseline trend using the 5\textsuperscript{th}, 10\textsuperscript{th}, and 15\textsuperscript{th} quantiles and compare three thresholds for classifying signal.
+	\begin{figure}
+		\caption{Estimated 15\textsuperscript{th} quantile trends on subset of the data using \texttt{qsreg} and \texttt{detrend\_eBIC}.} 
+		\includegraphics[width = \linewidth]{Figures/short_trends.png}
+		\label{fig:short-trends}
+	\end{figure}
+
+	We compare our \texttt{detrend\_eBIC} method with the \texttt{qsreg} method on a subset of the SPod data (n=6000) since the \texttt{qsreg} method cannot handle all 24 hours simultaneously.
+We estimate the baseline trend using the 10\textsuperscript{th} and 15\textsuperscript{th} quantiles and compare three thresholds for classifying signal.
 The thresholds are calculated using the median plus a multiple of the median absolute deviation (Eq.
@@
 	\label{eq:MAD}
-	\mbox{MAD} = \mbox{median}(y-\tilde{y})
+	\mbox{MAD} = \mbox{median}(y-\tilde{y}),
 	\end{equation}
 	where $\tilde{y}$ is the median of $y$.
-Given a method, quantile level and MAD multiple, we estimate the quantile trend for each of the sensor nodes and subtract it from the observations.
-We then calculate the threshold using the median plus the multiple times MAD of the corrected series and classify the corrected series based on the threshold.
-Fig.
-\ref{fig:rugplot} shows the time series used after estimating the baseline using the \texttt{detrend\_eBIC} estimate of the 10\textsuperscript{th} quantile and a threshold with an MAD multiple of 2.
+Given a method, quantile level, and MAD multiple, we estimate the quantile trend for each of the sensor nodes and subtract it from the observations.
+We then calculate the threshold using the median plus the MAD multiple of the corrected series and classify the corrected series based on the threshold.
+An example of the estimated baseline fit for each method is shown in Fig.
+\ref{fig:short-trends}, while Fig.
+\ref{fig:rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 3.
 
  	\begin{figure}
-	 	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 10th quantile.}
+	 	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile. Horizontal dashed lines represent the thresholds.}
 	 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
@@
 
-	Given the signal classifications for node g, $s_g(t) \in \{0,1\}$ and node f, $s_f(t)\in\{0,1\}$ we want to compare the similarity between the two classifications.
+	Given the signal classifications for node 1, $s_1(t) \in \{0,1\}$ and node 2, $s_2(t)\in\{0,1\}$ we want to compare the similarity between the two classifications.
 One metric for evaluating the distance between two classifications is the variation of information (VI):
-	\begin{align}
-		r_{ij} =& \frac{1}{n}\sum_t \mathbf{I}(s_f(t) = i  \cap s_g(t) = j)\\
-		VI(s_f, s_g) =& -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_f(t) = i)}\right) + 
-		\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_g(t) = j)}\right) \right]
-	\end{align}	
+	\begin{align*}
+		r_{ij} =& \frac{1}{n}\sum_t \mathbf{I}(s_1(t) = i  \cap s_2(t) = j)\\
+		VI(s_1, s_2) =& -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_1(t) = i)}\right) + 
+		\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_2(t) = j)}\right) \right]
+	\end{align*}	
 	The VI is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
@@
 \ref{fig:vi}.
-Using the smallest threshold the methods perform similarly although the lowest VIs are observed for the 15\textsuperscript{th} quantile using the \texttt{detrend\_eBIC} method.
-For the larger thresholds the \texttt{detrend\_eBIC} method produces substantially lower VI distances between nodes than the \texttt{qsreg} method.
+In all cases our \texttt{detrend} method results in classifications that are more similar than those from the \texttt{qsreg} method.
+As is illustrated in Fig.
+\ref{fig:short-trends} and Fig.
+\ref{fig:rugplot}, our method results in a smoother baseline estimate which improves signal classification.
 
@@
 
-	Our windowed \texttt{detrend\_eBIC} method was used to removed the baseline drift from the total dataset consisting of 52,322 observations per node.
-The VI scores for the full dataset were 0.14, 0.13, and 0.14 for nodes f and g, f and h, and g and h, respectively.
-The complete confusion matrices for the nodes using a MAD multiple of 3 for the threshold is given in Table 1.
+	Our windowed \texttt{detrend\_eBIC} method was used to removed the baseline drift from the total dataset consisting of 86,401 observations per node.
+The VI scores for the full dataset were 0.36, 0.24, and 0.41 for nodes c and d, c and e, and d and e, respectively.
+The complete confusion matrices for the nodes using a MAD multiple of 5 for the threshold is given in Table 1.
 
@@
 	\section{Conclusion and Discussion}
-	We have expanded the ideas of quantile trend filtering and demonstrated it's utility in both simulations and applied settings.
+	We have expanded the quantile trend filtering method by implementing a non-crossing constraint and a new algorithm for processing large series, and proposing a modified criteria for smoothing parameter selection.
+Furthermore we have demonstrated the utility of quantile trend filtering in both simulations and applied settings.
 Our ADMM algorithm for large series both reduces the computing time and allows trends to be estimated on series that cannot be estimated simultaneously while our scaled extended BIC criteria was shown to provided better estimated of quantile trends in series with and without a signal component.
 We have also shown that the baseline drift in low cost air quality sensors can be removed through estimating quantile trends.
-	In the future quantile trend filtering could be extended to observations measured at non-uniform spacing by incorporating the distance in covariate spacing into the differencing matrix.
-It could further be extended to estimate smooth spatial trends by adjusting the differencing matrix accordingly.
+
+	In the future, quantile trend filtering could be extended to observations measured at non-uniform spacing by incorporating the distance in covariate spacing into the differencing matrix.
+It could also be extended to estimate smooth spatial trends by a similar adjustment to the differencing matrix based on spatial distances between observations.
 

### 2019-02-01 Code/Fig_app_full.R

@@ -10,7 +10,7 @@ load_all("detrendr")
 rm(list=ls())
 source("application_functions.R")
 load("../SPod/spodPIDs.RData")
-
+nodes <- c("c", "d", "e")
 spodRaw <- spodPIDs %>%
   gather("node", "PID", -time)
 
@@ -19,14 +19,13 @@ ggplot(spodRaw, aes(x=time, y=PID, col=node)) +
   theme_bw() +
   #facet_grid(node~., scales = "free")+
   scale_color_brewer(palette = "Set1", labels = nodes)+
-  xlim(c(as.POSIXct("2017-04-13 12:30:01 EST"), 
-         as.POSIXct("2017-04-13 13:30:01 EST")))
+  # xlim(c(as.POSIXct("2017-04-13 12:30:01 EST"), 
+  #        as.POSIXct("2017-04-13 13:30:01 EST")))
 ggsave("../Manuscript/Figures/uncorrected_data.png", width = 7, height = 2.5)
 
 load("../SPod/spod_trends.RData")
-spodPeaks <- spodPIDs - select(spod_trends, contains("0.15"))
+spodPeaks <- select(spodPIDs, -time) - select(spod_trends, contains("0.15"))
 spodPeaks$time <- spod_trends$time
-spodPIDs$time <- spod_trends$time
 spodLong <- spodPeaks %>% gather("node","PID", -time)
 
 ggplot(spodLong, aes(x=time, y=PID, col=node)) +
@@ -41,6 +40,7 @@ methods <- c("detrendr")
 metric_df <- tibble(metric=NA, method=NA, tau=NA, crit=NA, metric_type=NA)
 i <- 1
 metrics <- c("confusion", "NMI", "VI")
+nodes <- c("c", "d", "e")
 
 for (method in methods){
   trends <- get(paste(method, "trends", sep = "_"))
@@ -49,11 +49,11 @@ for (method in methods){
       signal <- get_spod_signal(tau[j], trends, spodPIDs, crit)
       for (metric in metrics){
         if (metric == "confusion"){
-          metric_df$metric[i] <- I(list(get_confusion(signal)))
+          metric_df$metric[i] <- I(list(get_confusion(signal, nodes)))
         } else if (metric == "NMI"){
-          metric_df$metric[i] <- I(list(get_NMI(signal)))
+          metric_df$metric[i] <- I(list(get_NMI(signal, nodes)))
         } else if (metric == "VI") {
[…]

### 2019-02-01 Code/Fig_peaks_metrics.R

@@ -3,7 +3,7 @@
 # Halley Brantley
 ################################################################################
 library(tidyverse)
-
+rm(list=ls())
 i <- 10
 n <- 4000
 tau <- c(0.01, 0.05, 0.1)
@@ -41,8 +41,8 @@ tau <- c(0.01, 0.05, 0.1)
 nSim <- 100
 colPal <- rev(c('#762a83','#9970ab','#c2a5cf',
                 '#a6dba0','#5aae61','#1b7837'))
-methods <- c("detrend_eBIC", "detrend_SIC", "detrend_valid", "qsreg", 
-             "rqss", "npqw") 
+methods <- c("detrend_eBIC", "detrend_SIC", "detrend_valid", 
+             "rqss", "npqw", "qsreg") 
 MSEs <- as.data.frame(matrix(NA, nrow = nSim*length(methods), 
                              ncol = length(tau)+3))
 colnames(MSEs) <- c("Sim", "Method", "n", paste0("tau_", tau))

### 2019-02-05 Manuscript/detrendify_JCGS.tex
incorporated Joe's comments on sections 1 and 2

@@
 			Joseph Guinness\thanks{
-				Department of Biological Statistics and Computational Biology, Cornell University, Ithaca, NY 14853 (E-mail: guinness@cornell.edu )} \,
+				Department of Statistics and Data Science, Cornell University, Ithaca, NY 14853 (E-mail: guinness@cornell.edu )} \,
 			and
@@
 		We address the problem of estimating smoothly varying baseline trends in time series data using quantile trend filtering.
-We extend the basic framework to ensure non-crossing while estimating multiple trends simultaneously and implement an alternating direction method of moments (ADMM) algorithm for series that cannot be processed simultaneously.
-Timing experiments demonstrate the usefulness of our ADMM algorithm even in cases where the series can be processed simultaneously.
-We also address smoothing parameter selection and propose a modified criteria based on the extended Bayesian Information Criteria.
-We demonstrate both through simulation studies and an application to low cost air quality sensor data that our model provides better or comparable estimates than existing methods of non-parametric quantile trends and improves signal classification of low-cost air quality sensor output.
+We first extend the basic framework to ensure non-crossing while estimating multiple quantile trends simultaneously.
+We also implement a parallelizable alternating direction method of moments (ADMM) algorithm for estimating trends in longer series.
+The ADMM algorthim enables the estimation of trends in a piecewise manner, both reducing the computation time and extending the limits of the method to larger data sizes.
+We also address smoothing parameter selection and propose a modified criterion based on the extended Bayesian Information Criterion.
+Through simulation studies and an application to low cost air quality sensor data, we demonstrate that our model provides better quantile trend estimates than existing methods and improves signal classification of low-cost air quality sensor output.
 	\end{abstract}
@@
 
-	In applications spanning the fields of chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components.
-\cite{Kim2009} addressed the case when the series consisted of just a trend and a random noise component and proposed the use of \textit{$\ell_1$ trend filtering} to estimate trends that piecewise linear or piecewise polynomial.
-\cite{Tib2014} later demonstrated that empirically the trend filtering estimates adapt to the local level of smoothness better than the more common smoothing splines.
-In the trend filtering problem, given observations $y(t)$ with $t=1,...,N$, the trend, $\theta(t)$, is estimated by solving the following convex problem.
+	In applications spanning chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components.
+\cite{Kim2009} proposed \textit{$\ell_1$ trend filtering} to estimate trends that are piece-wise linear or polynomial, assuming that the series consists of a trend plus uncorrelated noise.
+\cite{Tib2014} later demonstrated empirically that the trend filtering estimates adapt to the local level of smoothness better than smoothing splines.
+In the trend filtering problem, given observations $y(t)$ with $t=1,...,N$, the trend, $\theta \in \mathbb{R}^n$, is estimated by solving the following convex problem:
 	\begin{eqnarray}
@@
 	\end{eqnarray}
-	where $\lambda \geq 0$ is a regularization parameter and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
+	where $y \in \mathbb{R}^n$ is the vector of observations, $\lambda \geq 0$ is a regularization parameter, and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
 To understand the purpose of penalizing $\mathbf{D}^{(k+1)}$ consider the difference operator when $k = 0$.
@@
 	\end{eqnarray}
-	Thus, $\lVert \mathbf{D}^{(1)}\theta \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
-The penalty incentivizes solutions which are piece-wise constant.
+	Thus, $\lVert \mathbf{D}^{(1)}\theta \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_{i+1} - \theta_{i} \rvert$ which is just total variation in one dimension.
+The penalty incentivizes solutions that are piece-wise constant.
 For $k \geq 1$, the difference operator $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n-k-1) \times n}$ is defined recursively as follows
@@
 	\end{eqnarray}
-	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
+	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piece-wise polynomials of order $k$.
 
-	Trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise (Fig.
+	\cite{Tib2014} showed that trend filtering is both fast and locally adaptive when the time series consists of only the trend and random noise.
+However, in some applications, such as the air quality monitoring problem considered in this paper, the data contain a rapidly varying signal in addition to the slowly varying trend and noise.
+Standard trend filtering struggles to distinguish between the slowly varying trend and the rapidly-varying signal.
+We propose quantile trend filtering as a more robust method for detrending in these situations (Fig.
 \ref{fig:trendfilter}).
-In some applications, however, time series may also include a rapidly varying signal component in addition to the trend and noise, in these cases the trend filtering estimate over-estimates the trend in the places where signal is present (Fig.
-\ref{fig:trendfilter}).
+To estimate the trend in the $\tau$\textsuperscript{th} quantile, we minimize the criterion
+	\begin{eqnarray}
+		\label{eq:quantile_trend}
+		\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \mathbf{D}^{(k+1)} \theta \rVert_1,
+	\end{eqnarray}	
+	where $\rho_\tau(z)$ is the check-loss function: 
+	\begin{equation}
+		\label{eq:checkloss}
+		\rho_{\tau}(z) = \sum_{i=1}^n z_i(\tau-\mathbf{I}(z_i<0)),
+	\end{equation}
+	 and $\mathbf{I}(\cdot) $ is the indicator function.
+This formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
+Here, we extend the basic framework to model multiple quantiles simultaneously.
+We also implement non-crossing constraints to ensure valid trend estimates.
+To reduce computation time and extend the method to larger series, we develop and a parallelizable ADMM algorithm.
+Finally, we propose a modified criterion for choosing the smoothing parameter.
 
@@
 		\centering
-		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}
 		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
 		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}
+		\caption{Examples of trend filtering solutions (red) and 15\textsuperscript{th} quantile trending filtering solution (blue).}
 		\label{fig:trendfilter}
@@
 
-	One application in which the observed time series consists of a slowly varying baseline, non-negative signal, and rapidly varying noise is the output of low cost air quality sensors.
-The use of low-cost, portable, air quality sensors has increased dramatically in the last decade.
+	Our research is motivated by the analysis of low cost and portable air quality sensors, whose use has increased dramatically in the last decade.
 These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
@@
 To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
-Fig.
-\ref{fig:raw_spod} provides an example of 3 co-located SPod sensor outputs from near an industrial facility.
-While all of the sensors response to the pollutant signal, the amount of baseline drift varies from one node to another.
-
-	\begin{figure}[b!]
+Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensor nodes co-located at the border of an industrial facility.
+All of the sensors respond to the pollutant signal, but the amount of baseline drift varies from one node to another.
+	\begin{figure}[t!]
+		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}		
 		\caption{Example of 3 co-located SPod PID sensor readings.}
-		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
 		\label{fig:raw_spod}
@@
 
-	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which can be thought of as a trend in a low quantile of the data.
-\cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check loss function  (Eq.
+	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, the goal is to estimate the trend in the baseline not the mean.
+We can define the trend in the baseline as the trend in a low quantile of the data.
+A variety of methods for estimating quantile trends have already been proposed.
+\cite{Koenker1978} were the first to propose substituting the squared-loss penalty with the check-loss penalty (Eq.
 \ref{eq:checkloss}) to estimate a conditional quantile instead of the conditional mean.
-
-	\begin{equation}
-	\label{eq:checkloss}
-	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0)).
-	\end{equation}
-
-	Later, \cite{KoenkerNgPortnoy1994} addressed the issue of estimated quantile curves by combining the check loss penalty with trend filtering penalty with $k = 2$ which produces quantile trends that are piecewise linear.
-Other approaches to estimating quantile trends have also been proposed.
-Rather than using the trend filtering $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm solving the optimization problem in Eq.
-\ref{eq:smoothingspline}.
+Later, \cite{KoenkerNgPortnoy1994} proposed quantile trend filtering with $k = 2$ producing quantile trends that are piecewise linear, but did not extent it to higher orders.
+Rather than using the $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm:
 
@@
  	\end{equation*}	
+	where $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degree of smoothing.
+\cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check-loss function with a similar differentiable function.
+\cite{Racine2017} take a different approach and constrain the response to follow a smooth location scale model of the form $y(t) = a(t) + b(t)\epsilon_i$ and estimate the $\tau$\textsuperscript{th} conditional quantile using a kernel estimator and local linear approach.
 
-	where $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing.
-\cite{Oh2011} proposed an algorithm for solving this problem based on approximating the checkloss function with a similar differentiable function.
-
-	\cite{Racine2017} take an entirely different approach to estimating quantile trends.
-They constrain the response to follow a smooth location scale model of the form $y(t) = a(t) + b(t)\epsilon_i$ and estimate the $\tau$\textsubscript{th} conditional quantile given $t = t_0$ using a kernel estimator and local linear approach.
-
-	We propose to use the trend filtering penalty combined with the check loss function to produce quantile trends that are piecewise quadratic.
-The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
-Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles.
-We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and propose a modified criteria for choosing the smoothing parameter.
-We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods and is a more effective method of drift removal for low-cost air quality sensors.
+	In the following sections we describe in detail our quantile trend filtering algorithms and regularization parameter criterion (Section \ref{sec:methods}).
+We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods (Section \ref{sec:simluation}).
+We further show that quantile trend filtering is a more effective method of drift removal for low-cost air quality sensors and results in improved signal classification compared to quantile smoothing splines (Section \ref{sec:application}).
+Finally, we discuss potential extensions of quantile trend filtering (Section \ref{sec:discussion}).
 
 	\section{Methods}
-
+	\label{sec:methods}
 	\subsection{Quantile Trend Filtering}
@@
 	We combine the ideas of quantile regression and trend filtering.
-For a single quantile level $\tau$ the estimation of the quantile trend filtering model can be posed as the following optimization problem.
-	\begin{eqnarray}
-	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \mathbf{D}^{(k+1)} \theta \rVert_1,
-	\end{eqnarray}
-	where $\lambda$ is a non-negative regularization parameter.
-We address the problem of choosing $\lambda$ in Section \ref{sec:lambda_choice}.
-As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of free or commercial solvers.
+For a single quantile level $\tau$, the quantile trend filtering problem is given in Eq.
+\ref{eq:quantile_trend}.
+As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of methods.
 In many cases, including ours, we are interested in estimating multiple quantiles simultaneously.
@@
 	\end{eqnarray}	
-
 	where $\theta_j \in \mathcal{R}^n$.
-The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
-In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation, but it can easily be substituted for a free solver such as the Rglpk package by \cite{rglpk}.
-The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
-As the size of the data and the number of quantiles grows, all solvers will eventually break.
+The additional constraints are linear in the parameters, so the non-crossing quantile trends can still be estimated by a number of available solvers.
+In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation.
+However, we could easily substitute a free solver such as the Rglpk package by \cite{rglpk}.
 
@@
 
+	The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
+As the size of the data and the number of quantiles grows, all solvers will eventually break.
 To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
-We propose an alternating direction method of multipliers (ADMM) algorithm for solving large problems in a piecewise fashion.
+We propose an ADMM algorithm for solving large problems in a piece-wise fashion.
 The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is fully described by \cite{boyd2011distributed}.
-We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem given in Eq.
-\ref{eq:quantile_trend}, by dividing our observed series $y(t)$ with $t = \{1, ..., n\}$ into overlapping windows
+We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in Eq.
+(\ref{eq:quantile_trend}), by dividing our observed series $y(t)$ with $t = \{1, ..., n\}$ into overlapping windows
 
@@
 	\end{align*}
-
 	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< n$.
@@
 Given quantiles $\tau_1 < ...
-< \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$.
-In order to write out the constraint that the overlapping sections must be equal we define a consensus variable
+< \tau_J$, we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$ at time $t$.
+In order to constrain the overlapping sections to be equal we define a consensus variable
 
 	 \begin{align*}
-		 \bar{\theta}_{j,m} =&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
-		 =& \begin{cases} 
+	\bar{\theta}_{j, m}	=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
+	\bar{\theta}_{j, m}(t)	 =& \begin{cases} 
 			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
@@
 		 \label{eq:quantile_windows}
-		 &\sum_{m=1}^M\underset{\theta_{1,m}, ..., \theta_{J,m}}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
+		 &\underset{\theta_{1,m}, ..., \theta_{J,m}}{\min}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
 		 \lambda_j \lVert \mathbf{D}^{(k+1)} \theta_{j,m} \rVert_1 \right ] \\
@@
 	 \end{align*}
-	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$ we can write the augmented Lagrangian for finding the trends in window $m$:
+	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$, the augmented Lagrangian for finding the trends in window $m$ is
 	 \begin{equation*}
 	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \mathbf{D}^{(k+1)}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
-	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
+	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2.
 	 \end{equation*}
-	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in algorithm 1.
+	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in Algorithm 1.
+The use of the Augmented Lagrangian converts the problem from a linear program into a quadratic program.
+The \texttt{gurobi} R package \citep{gurobi} can solve quadratic programs in addition to linear programs, but we can also use the free R package \texttt{quadprog} \citep{quadprog}.
 
@@
 		\centering
+		\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
 		\caption{Window boundaries and trends fit separately in each window.}
-		\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
 		\label{fig:windows}
@@
 
-	We measure convergence use the stopping criteria described by \cite{boyd2011distributed}.
-The criteria are based on the primal and dual residuals which represent the residuals for the primal and dual feasibility, respectively.
+	We measure convergence using the stopping criteria described by \cite{boyd2011distributed}.
+The criteria are based on the primal and dual residuals, which represent the residuals for primal and dual feasibility, respectively.
 The primal residual,
@@
 	\end{equation}
-	represents the difference between the trend values in the windows and the consensus trend value while the dual residual,
+	represents the difference between the trend values in the windows and the consensus trend value. The dual residual,
 	\begin{equation*}
@@
 		\centering
+		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
 		\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.}
-		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
 	\end{figure}
 
-	Timing experiments illustrate the advantages of using our ADMM algorithm even on datasets where solving the problem simultaneously is possible.
-For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below and trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using a $\lambda = n/5$.
+	Timing experiments illustrate the advantages of using our ADMM algorithm when the trend can be estimated with a single window.
+For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below.
+Trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using $\lambda = n/5$.
 We use from one to four windows for each data size with an overlap of 500.
@@
 \ref{fig:timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
-The timing experiments were conducted on
+The timing experiments were conducted on an Intel Xeon based Linux cluster using two processor cores.
 
@@
 		\centering
+		\includegraphics[width = 0.7\linewidth]{Figures/Fig_timing_experiment.png}
 		\caption{Timing experiments comparing quantile trend filtering with varying numbers of windows by data size.}
-		\includegraphics[width = 0.7\linewidth]{Figures/Fig_timing_experiment.png}
 		\label{fig:timing}
@@
 	\label{sec:lambda_choice}
-	An important problem in trend estimation is the choice of regularization parameter or degree of smoothness.
-Our method can easily handle missing data by defining the check loss function to output 0 for missing values.
-This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$.
-However, the use of an information criteria metric can result in a better choice of regularization parameter than the validation method.
+	An important problem in trend estimation is the choice of regularization parameter, or degree of smoothness.
+Our method can easily handle missing data by defining the check loss function to output 0 for missing values, allowing us to implement cross-validation.
+We also explore three information criteria.
 \cite{KoenkerNgPortnoy1994} addressed the choice of regularization parameter by proposing the Schwarz criterion for the selection of $\lambda$
@@
 	where $p_{\lambda} = \sum_t I(y(t) = \widehat{\theta}(t))$ is the number of interpolated points, which can be thought of as active knots.
-The SIC is based on the traditional Bayesian Information Criterion (BIC) which is given by
+The SIC is based on the traditional Bayesian Information Criterion,
 	\begin{equation}
-	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}\}) + \nu\log n 
+	\mbox{BIC}(\nu) = -2\log(L\{\hat{\theta}\}) + \nu\log n 
 	\end{equation}	
 	where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\hat{\theta}$.
-If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and assume that minimizing the checkloss function corresponds to maximizing the asymmetric Laplace likelihood, \begin{equation}
+If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and view minimizing the checkloss function as maximizing the asymmetric Laplace likelihood, \begin{equation}
 	L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_t\rho_\tau\left(\frac{y(t) - \theta(t)}{\sigma}\right)\right\},
@@
 	\begin{equation}
-	\mbox{BIC}(df) = 2\frac{1}{\sigma}\rho_{\tau}(y-\hat{\theta}) + \mbox{df}\log n
+	\mbox{BIC}(\nu) = 2\frac{1}{\sigma}\rho_{\tau}(y-\hat{\theta}) + \nu\log n
 	\end{equation} 
-	where $df$ is the number of non-zero elements of $D^{(k+1)}\hat{\theta}$. We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates.
+	where $\nu$ is the number of non-zero elements of $\mathbf{D}^{(k+1)}\hat{\theta}$.
+We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates.
 
@@
 \ref{fig:BIC}.
+In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
 
 	\begin{figure}[h!]
+		\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
+		\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
+		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png} 
 		\caption{(Left) Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.
@@
 (Right) Criterion values by $\log(\lambda)$ vertical lines indicate locations of minima.}
-		\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
-		\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
-		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png} 
 		\label{fig:BIC}
@@
 	\section{Simulation Studies}
-
+	\label{sec:simluation}
 	We conduct two simulation studies to compare the performance of our quantile trend filtering method and regularization parameter selection criteria with previously published methods.
@@
 	\begin{figure}
-		\caption{Simulated data with true quantile trends.}	
 		\includegraphics[width=.28\linewidth]{Figures/gaus.png}
@@
 		\includegraphics[width=.45\linewidth]{Figures/mixednorm.png}
+		\caption{Simulated data with true quantile trends.}	
 	\end{figure}
@@
 	\begin{figure}
-		\caption{RMSE by design, method, quantile and data size. Points and error bars represent mean RMSE $\pm$ twice the standard error.}
 		\includegraphics[width=\linewidth]{Figures/gaus_mse.png}	
@@
 		\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
+		\caption{RMSE by design, method, quantile and data size. Points and error bars represent mean RMSE $\pm$ twice the standard error.}
 		\label{fig:quantile_mse}
@@
 	\begin{figure}
+		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}	
 		\caption{RMSE by method, quantile and data size for peaks design.}
-		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}	
 		\label{fig:peaks_rmse}
@@
 	\begin{figure}[h!]
+		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
 		\caption{Example signal classification using threshold.
 Red indicates true signal $>0.5$, blue indicates observations classified as signal after baseline removal using \texttt{detrend\_eBIC} and a threshold of 1.2.}
-		\includegraphics[width = \linewidth]{Figures/peaks_eg_class.png}
 		\label{fig:peaks_class_eg}
@@
 	\begin{figure}[h!]
+		\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
 		\caption{Class averaged accuracy by threshold, data size, and method (1 is best 0.5 is worst).}
-		\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
 	\end{figure}
@@
 	\section{Application}
-
+	\label{sec:application}
 	The low-cost ``SPod" air quality sensors output a time series that includes a slowly varying baseline, high frequency random noise, and the sensor response to pollutants.
@@
 	\begin{figure}
+		\includegraphics[width = \linewidth]{Figures/short_trends.png}
 		\caption{Estimated 15\textsuperscript{th} quantile trends on subset of the data using \texttt{qsreg} and \texttt{detrend\_eBIC}.} 
-		\includegraphics[width = \linewidth]{Figures/short_trends.png}
 		\label{fig:short-trends}
@@
  	\begin{figure}
+	 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
 	 	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile. Horizontal dashed lines represent the thresholds.}
-	 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
 	 	\label{fig:rugplot}
@@
 		\centering
+		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
 		\caption{Variation of Information between sensor nodes after trend removal by quantile and method and thresholding by factor of MAD.}
-		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
 		\label{fig:vi}
@@
 	\begin{figure}
+		\includegraphics[width = \linewidth]{Figures/corrected_data.png}
 		\caption{Low cost sensor data after drift removal using windowed detrend with eBIC.}
-		\includegraphics[width = \linewidth]{Figures/corrected_data.png}
 	\end{figure}
@@
 	\section{Conclusion and Discussion}
+	\label{sec:discussion}
 	We have expanded the quantile trend filtering method by implementing a non-crossing constraint and a new algorithm for processing large series, and proposing a modified criteria for smoothing parameter selection.

### 2019-02-05 Code/Fig_trendfilter.R

@@ -1,4 +1,5 @@
 library(devtools)
+load_all("detrendr")
 install_github("hadley/l1tf")
 install_github("glmgen/genlasso")
 library(l1tf)

### 2019-02-08 Manuscript/detrendify_JCGS.tex
Some edits to sections 1 and 2.

--- 2019-02-01 Manuscript/detrendify_JCGS.tex
@@
 \RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
-\usepackage{graphicx}
+\usepackage{bm,graphicx,psfrag,epsf}
 %\usepackage{enumerate}
@@
 \usepackage{algpseudocode}
+\usepackage{subfig} %% For \subfloat command
+
 %\pdfminorversion=4
@@
 \addtolength{\topmargin}{-.8in}%
+
+\usepackage[usenames]{color}
+\usepackage{xspace}
+
+%% For comments to each other
+\definecolor{blue}{rgb}{0.2,0.5,0.7}
+\definecolor{green}{rgb}{0.3,0.68,0.29}
+\definecolor{purple}{rgb}{0.6,0.31,0.64}
+\newcommand{\Halley}[2]{{\bf {\color{purple}@Halley from #1: #2}}\xspace}
+\newcommand{\Joe}[2]{{\bf {\color{blue}@Joe from #1: #2}}\xspace}
+\newcommand{\Eric}[2]{{\bf {\color{green}@Eric from #1: #2}}\xspace}
+
+%% Turn on/off options for PDF
+\usepackage{ifpdf}
+
+%% For hyperlinks. Should always be the last package. Omit for DVI version.
+\ifpdf
+\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
+\else
+\newcommand{\href}[2]{{#2}}
+\usepackage{url}
+\fi
+
+%% ----------------------------------------------------------------------
+%% Hyper-linked References
+%% ----------------------------------------------------------------------
+
+\ifpdf
+\newcommand{\Sec}[1]{\hyperref[sec:#1]{Section~\ref*{sec:#1}}} %section
+\newcommand{\App}[1]{\hyperref[sec:#1]{Appendix~\ref*{sec:#1}}} %appendix
+\newcommand{\Eqn}[1]{\hyperref[eq:#1]{{\rm (\ref*{eq:#1})}}} %equation
+\newcommand{\Part}[1]{\hyperref[part:#1]{(\ref*{part:#1})}} %part of theorem
+\newcommand{\Fig}[1]{\hyperref[fig:#1]{Figure~\ref*{fig:#1}}} %figure
+\newcommand{\Tab}[1]{\hyperref[tab:#1]{Table~\ref*{tab:#1}}} %table
+\newcommand{\Thm}[1]{\hyperref[thm:#1]{Theorem~\ref*{thm:#1}}} %theorem
+\newcommand{\Lem}[1]{\hyperref[lem:#1]{Lemma~\ref*{lem:#1}}} %lemma
+\newcommand{\Prop}[1]{\hyperref[prop:#1]{Proposition~\ref*{prop:#1}}} %proposition
+\newcommand{\Cor}[1]{\hyperref[cor:#1]{Corollary~\ref*{cor:#1}}} %corollary
+\newcommand{\Def}[1]{\hyperref[def:#1]{Definition~\ref*{def:#1}}} %definition
+\newcommand{\Alg}[1]{\hyperref[alg:#1]{Algorithm~\ref*{alg:#1}}} %algorithm
+\newcommand{\Ex}[1]{\hyperref[ex:#1]{Example~\ref*{ex:#1}}} %example
+\newcommand{\As}[1]{\hyperref[as:#1]{Assumption~{\rm\ref*{as:#1}}}} %assumption
+\newcommand{\Reg}[1]{\hyperref[as:#1]{Condition~\ref*{reg:#1}}} %regularity condition
+\newcommand{\AlgLine}[2]{\hyperref[alg:#1]{line~\ref*{line:#2} of Algorithm~\ref*{alg:#1}}}
+\newcommand{\AlgLines}[3]{\hyperref[alg:#1]{lines~\ref*{line:#2}--\ref*{line:#3} of Algorithm~\ref*{alg:#1}}}
+\else
+\newcommand{\Sec}[1]{{Section~\ref{sec:#1}}} %section
+\newcommand{\App}[1]{{Appendix~\ref{sec:#1}}} %appendix
+\newcommand{\Eqn}[1]{{(\ref{eq:#1})}} %equation
+\newcommand{\Part}[1]{{(\ref{part:#1})}} %part of proof
+\newcommand{\Fig}[1]{{Figure~\ref{fig:#1}}} %figure
+\newcommand{\Tab}[1]{{Table~\ref{tab:#1}}} %table
+\newcommand{\Thm}[1]{{Theorem~\ref{thm:#1}}} %theorem
+\newcommand{\Lem}[1]{{Lemma~\ref{lem:#1}}} %lemma
+\newcommand{\Prop}[1]{{Property~\ref{prop:#1}}} %property
+\newcommand{\Cor}[1]{{Corollary~\ref{cor:#1}}} %corollary
+\newcommand{\Def}[1]{{Definition~\ref{def:#1}}} %definition
+\newcommand{\Alg}[1]{{Algorithm~\ref{alg:#1}}} %algorithm
+\newcommand{\Ex}[1]{{Example~\ref{ex:#1}}} %example
+\newcommand{\As}[1]{{Assumption~\ref{as:#1}}} %assumption
+\newcommand{\Reg}[1]{{R~\ref{reg:#1}}} %regularity condition
+\newcommand{\AlgLine}[2]{{line~\ref{line:#2} of Algorithm~\ref{alg:#1}}}
+\newcommand{\AlgLines}[3]{{lines~\ref{line:#2}--\ref{line:#3} of Algorithm~\ref{alg:#1}}}
+\fi
+
+%% ----------------------------------------------------------------------
+%% Definitions
+%% ----------------------------------------------------------------------
+
+%% Shorthand for Real
+\newcommand{\Real}{\mathbb{R}}
+
+%% Math notation
+\newcommand{\Tra}{^{\sf T}} % Transpose
+\newcommand{\Inv}{^{-1}} % Inverse
+\newcommand{\tr}{\operatorname{tr}} % Trace
+\newcommand{\Ker}{\operatorname{Ker}} % Kernel
+\def\vec{\mathop{\rm vec}\nolimits}
+\newcommand{\amp}{\mathop{\:\:\,}\nolimits}
+\newcommand{\bic}{\text{BIC}}
+
+%% Vectors
+\newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
+\newcommand{\Vbar}[1]{{\bm{\M{A}r \mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\Vn}[2]{\V{#1}^{(#2)}} % n-th vector
+\newcommand{\VnE}[3]{{#1}^{(#2)}_{#3}} % n-th vector
+\newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}} % vector element
+
+%% Matrices
+\newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} % matrix element
+\newcommand{\MC}[2]{\V{#1}_{#2}}
+\newcommand{\Mr}[2]{\V{#1}_{#2}} % matrix row
+\newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\MhatC}[2]{\Vhat{#1}_{#2}} % matrix column
+\newcommand{\Mbar}[1]{{\bm{\M{A}r \mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\MbarC}[2]{\Vbar{#1}_{#2}} % matrix column
+\newcommand{\Mn}[2]{\M{#1}^{(#2)}} % n-th matrix
+\newcommand{\Mbarn}[2]{\Mbar{#1}^{(#2)}} % n-th matrix
+\newcommand{\Mtilden}[2]{\Mtilde{#1}^{(#2)}} % n-th matrix
+\newcommand{\MnTra}[2]{\M{#1}^{(#2){{\sf T}}}} % n-th matrix transpose
+\newcommand{\MnE}[3]{\MakeLowercase{#1}^{(#2)}_{#3}} % n-th matrix element
+\newcommand{\MnC}[3]{\V{#1}^{(#2)}_{#3}} % n-th matrix column
+\newcommand{\MbarnC}[3]{\Vbar{#1}^{(#2)}_{#3}} % n-th matrix column
+\newcommand{\MnCTra}[3]{\V{#1}^{(#2){{\sf T}}}_{#3}} % n-th matrix column transpose
+\newcommand{\MCTra}[2]{\V{#1}^{{\sf T}}_{#2}} % n-th matrix column transpose
 
@@
 	\noindent%
-	{\it Keywords:} quantile regression, non-parametric, trend estimation, smoothing splines
+	{\it Keywords:} quantile regression, non-parametric regression, trend estimation, smoothing splines
 	\vfill
@@
 	In applications spanning the fields of chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components.
-\cite{Kim2009} addressed the case when the series consisted of just a trend and a random noise component and proposed the use of \textit{$\ell_1$ trend filtering} to estimate trends that piecewise linear or piecewise polynomial.
-\cite{Tib2014} later demonstrated that empirically the trend filtering estimates adapt to the local level of smoothness better than the more common smoothing splines.
-In the trend filtering problem, given observations $y(t)$ with $t=1,...,N$, the trend, $\theta(t)$, is estimated by solving the following convex problem.
+\cite{Kim2009} addressed the case when the series consisted of just a trend and a random noise component and proposed the use of \textit{$\ell_1$-trend filtering} to estimate trends that piecewise linear or piecewise polynomial.
+\cite{Tib2014} later demonstrated that empirically the trend filtering estimates adapt to the local level of smoothness better than the more common smoothing splines \Halley{Eric}{If he proved theorems, minimax stuff, we should state that.}.
+In the trend filtering problem, given observations $y(t)$ with $t=1,\ldots,n$, the trend, $\theta(t)$, is estimated by solving the following convex optimization problem.
 	\begin{eqnarray}
-	\underset{\theta}{\arg\min}\; \frac{1}{2} \lVert y - \theta \rVert_2^2 + \lambda \lVert \mathbf{D}^{(k+1)}\theta \rVert_1,
+	\underset{\V{\theta}}{\text{minimize}}\; \frac{1}{2} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
 	\end{eqnarray}
-	where $\lambda \geq 0$ is a regularization parameter and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
-To understand the purpose of penalizing $\mathbf{D}^{(k+1)}$ consider the difference operator when $k = 0$.
+	where $\lambda \geq 0$ is a regularization parameter and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
+To understand the purpose of penalizing $\Mn{D}{k+1}\V{\theta}$ consider the difference operator when $k = 0$.
 	\begin{eqnarray}
-	\mathbf{D}^{(1)} = \begin{pmatrix}
+	\Mn{D}{1} & = & \begin{pmatrix}
 	-1 & 1 & 0 & \cdots & 0 & 0 \\
@@
 	0 & 0 & 0 & \cdots & -1 & 1 \\
-	\end{pmatrix}
+	\end{pmatrix}.
 	\end{eqnarray}
-	Thus, $\lVert \mathbf{D}^{(1)}\theta \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
-The penalty incentivizes solutions which are piece-wise constant.
-For $k \geq 1$, the difference operator $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n-k-1) \times n}$ is defined recursively as follows
+	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
+The penalty term incentivizes solutions which are piece-wise constant.
+For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
 	\begin{eqnarray}
-	\mathbf{D}^{(k+1)} & = & \mathbf{D}^{(1)}\mathbf{D}^{(k)}.
+	\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
 	\end{eqnarray}
@@
 
-	Trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise (Fig.
-\ref{fig:trendfilter}).
-In some applications, however, time series may also include a rapidly varying signal component in addition to the trend and noise, in these cases the trend filtering estimate over-estimates the trend in the places where signal is present (Fig.
-\ref{fig:trendfilter}).
+	\Fig{tpn} shows that trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise.
+In some applications, however, time series may also include a rapidly varying signal component in addition to the trend and noise.
+\Fig{tpspn} shows that in these cases the trend filtering estimate over-estimates the trend in the places where signal is present.
 
-	\begin{figure}
+	\begin{figure}[t]
 		\centering
-		\caption{Examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}
-		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
-		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}
+		\subfloat[Slow Trend + Noise]{\label{fig:tpn}
+		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}}
+		\subfloat[Slow Trend + Spikes + Noise]{\label{fig:tpspn}
+		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}}
+		\caption{Two examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}		
 		\label{fig:trendfilter}
@@
 To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
-Fig.
-\ref{fig:raw_spod} provides an example of 3 co-located SPod sensor outputs from near an industrial facility.
+\Fig{raw_spod} provides an example of 3 co-located SPod sensor outputs from near an industrial facility.
 While all of the sensors response to the pollutant signal, the amount of baseline drift varies from one node to another.
@@
 	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which can be thought of as a trend in a low quantile of the data.
-\cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check loss function  (Eq.
-\ref{eq:checkloss}) to estimate a conditional quantile instead of the conditional mean.
+\cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check loss function  \Eqn{checkloss} to estimate a conditional quantile instead of the conditional mean.
 
@@
 Other approaches to estimating quantile trends have also been proposed.
-Rather than using the trend filtering $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm solving the optimization problem in Eq.
-\ref{eq:smoothingspline}.
+Rather than using the trend filtering $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm solving the optimization problem in \Eqn{smoothingspline}.
 
@@
 	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \mathbf{D}^{(k+1)} \theta \rVert_1,
+	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \Mn{D}{k+1} \theta \rVert_1,
 	\end{eqnarray}
@@
 We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$.
-Given quantiles $\{\tau_1, ..., \tau_J\}$ such that $\tau_1 < \tau_2 < ...
-< \tau_J$, the optimization problem becomes
+Given quantiles $\{\tau_1, \ldots, \tau_J\}$ such that $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
 
@@
 	\label{eq:noncross_trend}
-	\underset{\theta_1, ..., \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
-	\lambda_j \lVert \mathbf{D}^{(k+1)} \theta_j \rVert_1 \right ] \\
-	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le ... \le \theta_{J}(t) \text{ for all } t,
+	\underset{\theta_1, \ldots, \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
+	\lambda_j \lVert \Mn{D}{k+1} \theta_j \rVert_1 \right ] \\
+	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le \ldots \le \theta_{J}(t) \text{ for all } t,
 	\end{eqnarray}
@@
 The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is fully described by \cite{boyd2011distributed}.
-We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem given in Eq.
-\ref{eq:quantile_trend}, by dividing our observed series $y(t)$ with $t = \{1, ..., n\}$ into overlapping windows
+We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem given in \Eqn{quantile_trend}, by dividing our observed series $y(t)$ with $t = \{1, \ldots, n\}$ into overlapping windows
 
@@
 
-	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< n$.
-An illustration is given in Fig.
-\ref{fig:windows}.
-Given quantiles $\tau_1 < ...
-< \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$.
+	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \ldots< n$.
+An illustration is given in \Fig{windows}.
+Given quantiles $\tau_1 < \ldots < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$.
 In order to write out the constraint that the overlapping sections must be equal we define a consensus variable
@@
 		 \label{eq:quantile_windows}
-		 &\sum_{m=1}^M\underset{\theta_{1,m}, ..., \theta_{J,m}}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
-		 \lambda_j \lVert \mathbf{D}^{(k+1)} \theta_{j,m} \rVert_1 \right ] \\
+		 &\sum_{m=1}^M\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
+		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
 		 &\text{subject to: }\; \\
-		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le ... \le \theta_{J,m}(t) \text{ for all } m,t \\
+		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
 		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
@@
 	 \begin{equation*}
-	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \mathbf{D}^{(k+1)}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
+	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
 	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
@@
 			\State \textbf{initialize:} \\
-				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < ...<\theta_{J,m}(t)$ for all $t$. \\		
+				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$. \\		
 			 	 $\omega_{j,m}^{(0)} = \mathbf{0}$	
@@
 				$\theta_{j,m}^{(q)} = \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
-			 subject to $\theta_{1,m}(t) < ...<\theta_{J,m}(t)$ for all $t$.
+			 subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
 			\Until {convergence}
@@
 The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
-As is shown in Fig.
-\ref{fig:timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
+As is shown in \Fig{timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
 The timing experiments were conducted on
@@
 We used this criteria with $\gamma = 1$, and $P=n-k-1$.
-We use a single dataset from our simulation study to illustrate the difference between the scaled, unscaled ($\sigma = 1$) and scaled extended BIC criteria in Fig.
-\ref{fig:BIC}.
+We use a single dataset from our simulation study to illustrate the difference between the scaled, unscaled ($\sigma = 1$) and scaled extended BIC criteria in \Fig{BIC}.
 
@@
 
-	We also compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC (Eq.
-\ref{eq:SIC}) \citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria (Eq.
-\ref{eq:eBIC}).
+	We also compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC \Eqn{SIC}\citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria \Eqn{eBIC}.
 
@@
 	To compare performance in estimating quantile trends, three simulation designs from \cite{Racine2017} were considered.
-For all designs $t = 1, ..., n$,  $x(t) = t/n$, and the response $y$ was generated as
+For all designs $t = 1, \ldots, n$,  $x(t) = t/n$, and the response $y$ was generated as
 	$$y(t) = sin(2\pi x(t)) + \epsilon(x(t))$$
@@
 	\end{equation} 
-	with $t = 1, ..., n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, 0.25^2)$ is an error component.
+	with $t = 1, \ldots, n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, 0.25^2)$ is an error component.
 We generate $b(t)$ using cubic natural spline basis functions with degrees of freedom sampled from a Poisson distribution with mean parameter equal to $n/100$,  and coefficients drawn from an exponential distribution with rate 1.
@@
 One hundred datasets were generated for each $n=\{500, 1000, 2000, 4000\}$.
-We compare the ability of the methods to estimate the true quantiles of $Y(t)-s(t)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (Fig.
-\ref{fig:peaks_rmse}).
+We compare the ability of the methods to estimate the true quantiles of $Y(t)-s(t)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}).
 In this simulation study our \texttt{detrend\_eBIC} method outperforms the others substantially.
@@
 We compare three different quantiles for the baseline estimation and four different thresholds for classifying the signal after subtracting the estimated baseline from the observations.
-An illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal" is shown in Fig.
-\ref{fig:peaks_class_eg}.
+An illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal" is shown in \Fig{peaks_class_eg}.
 To compare the resulting signal classifications we calculate the class averaged accuracy (CAA).
@@
 A potential use for these sensors is to monitor pollutant concentrations at the fence lines of industrial facilities and detect time points when high concentrations are present.
-Ideally, three co-located and time aligned sensors (as shown in Fig.
-\ref{fig:raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
+Ideally, three co-located and time aligned sensors (as shown in \Fig{raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
 
@@
 We estimate the baseline trend using the 10\textsuperscript{th} and 15\textsuperscript{th} quantiles and compare three thresholds for classifying signal.
-The thresholds are calculated using the median plus a multiple of the median absolute deviation (Eq.
-\ref{eq:MAD}) of the detrended series.
+The thresholds are calculated using the median plus a multiple of the median absolute deviation \Eqn{MAD} of the detrended series.
 	\begin{equation}
@@
 We then calculate the threshold using the median plus the MAD multiple of the corrected series and classify the corrected series based on the threshold.
-An example of the estimated baseline fit for each method is shown in Fig.
-\ref{fig:short-trends}, while Fig.
-\ref{fig:rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 3.
+An example of the estimated baseline fit for each method is shown in \Fig{short-trends}, while \Fig{rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 3.
 
@@
 	The VI is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
-The VIs by method, quantile and trend are shown in Fig.
-\ref{fig:vi}.
+The VIs by method, quantile and trend are shown in \Fig{vi}.
 In all cases our \texttt{detrend} method results in classifications that are more similar than those from the \texttt{qsreg} method.
-As is illustrated in Fig.
-\ref{fig:short-trends} and Fig.
-\ref{fig:rugplot}, our method results in a smoother baseline estimate which improves signal classification.
+As is illustrated in \Fig{short-trends} and \Fig{rugplot}, our method results in a smoother baseline estimate which improves signal classification.
 

### 2019-02-08 Manuscript/detrendify_JCGS.tex
Eric's edits.

@@
 \RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
+\usepackage{bbm}
 \usepackage{bm,graphicx,psfrag,epsf}
@@
 \newcommand{\bic}{\text{BIC}}
+\newcommand{\VI}{\text{VI}}
+\newcommand{\One}{\mathbbm{1}}
 
@@
 We also address smoothing parameter selection and propose a modified criteria based on the extended Bayesian Information Criteria.
-We demonstrate both through simulation studies and an application to low cost air quality sensor data that our model provides better or comparable estimates than existing methods of non-parametric quantile trends and improves signal classification of low-cost air quality sensor output.
+We demonstrate both through simulation studies and an application to lost-cost air quality sensor data that our model provides better or comparable estimates than existing methods of non-parametric quantile trends and improves signal classification of low-cost air quality sensor output.
 	\end{abstract}
@@
 In the trend filtering problem, given observations $y(t)$ with $t=1,\ldots,n$, the trend, $\theta(t)$, is estimated by solving the following convex optimization problem.
-	\begin{eqnarray}
+\Halley{Eric}{We maybe should define two objects $y(t)$ which is a function over the reals and $\V{y} \in \Real^n$, where $\VE{y}{i} = y(t_i)$.}
+	\begin{eqnarray*}
 	\underset{\V{\theta}}{\text{minimize}}\; \frac{1}{2} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
-	\end{eqnarray}
+	\end{eqnarray*}
 	where $\lambda \geq 0$ is a regularization parameter and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
 To understand the purpose of penalizing $\Mn{D}{k+1}\V{\theta}$ consider the difference operator when $k = 0$.
-	\begin{eqnarray}
+	\begin{eqnarray*}
 	\Mn{D}{1} & = & \begin{pmatrix}
@@
 	\end{pmatrix}.
-	\end{eqnarray}
+	\end{eqnarray*}
 	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
@@
 For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
-	\begin{eqnarray}
+	\begin{eqnarray*}
 	\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
-	\end{eqnarray}
+	\end{eqnarray*}
 	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
@@
 	\Fig{tpn} shows that trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise.
-In some applications, however, time series may also include a rapidly varying signal component in addition to the trend and noise.
+In some applications, however, time series may also include, in addition to the trend and noise, a signal component that varies more rapidly than the trend.
 	\Fig{tpspn} shows that in these cases the trend filtering estimate over-estimates the trend in the places where signal is present.
@@
 		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}}
-		\caption{Two examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}		
+		\caption{Two examples of trend filtering solutions.
+The true trend is shown in black, while the estimated trend is shown in red.
+\Halley{Eric}{Could you explain to the reader what they should be seeing in the two panels, especially the one on the right?
+Don't leave to chance that a busy reviewer may miss the point.}}
 		\label{fig:trendfilter}
@@
 
-	One application in which the observed time series consists of a slowly varying baseline, non-negative signal, and rapidly varying noise is the output of low cost air quality sensors.
+	One application in which the observed time series consists of a slowly varying baseline, non-negative signal, and rapidly varying noise is the output of lost-cost air quality sensors.
 The use of low-cost, portable, air quality sensors has increased dramatically in the last decade.
@@
 	\begin{figure}[b!]
+		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
 		\caption{Example of 3 co-located SPod PID sensor readings.}		
-		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
 		\label{fig:raw_spod}
@@
 	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which can be thought of as a trend in a low quantile of the data.
-\cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check loss function  \Eqn{checkloss} to estimate a conditional quantile instead of the conditional mean.
+\cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check-loss function  \Eqn{check-loss} to estimate a conditional quantile instead of the conditional mean.
+	\begin{eqnarray}
+	\label{eq:check-loss}
+	 \rho_{\tau}(\V{z}) & = & \sum_{i=1}^n z_i(1-\One(z_i<0)),
+	\end{eqnarray}
+	where $\One(A)$ is 1 if its input $A$ is true and 0 otherwise.
 
-	\begin{equation}
-	\label{eq:checkloss}
-	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0)).
-	\end{equation}
-
-	Later, \cite{KoenkerNgPortnoy1994} addressed the issue of estimated quantile curves by combining the check loss penalty with trend filtering penalty with $k = 2$ which produces quantile trends that are piecewise linear.
+	Later, \cite{KoenkerNgPortnoy1994} addressed the issue of estimated quantile curves by combining the check-loss penalty with trend filtering penalty with $k = 2$ which produces quantile trends that are piecewise linear.
 Other approaches to estimating quantile trends have also been proposed.
@@
  	\end{equation*}
-
 where $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degrees of smoothing.
-\cite{Oh2011} proposed an algorithm for solving this problem based on approximating the checkloss function with a similar differentiable function.
+\Halley{Eric}{What is $\rho_{\tau}(\cdot)$ when its input is a function over the reals?
+It's defined above as a function of a vector input.} \cite{Oh2011} proposed an algorithm for solving this problem based on approximating the check-loss function with a similar differentiable function.
 
@@
 
-	We propose to use the trend filtering penalty combined with the check loss function to produce quantile trends that are piecewise quadratic.
+	We propose to use the trend filtering penalty combined with the check-loss function to produce quantile trends that are piecewise quadratic.
 The formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
-Moreover we extend the basic framework to ensure non-crossing while modeling multiple quantiles.
+Moreover, we extend the basic framework to ensure non-crossing while modeling multiple quantiles.
 We also implement a parallel ADMM algorithm for series that are too large to be computed simultaneously and propose a modified criteria for choosing the smoothing parameter.
@@
 	We combine the ideas of quantile regression and trend filtering.
-For a single quantile level $\tau$ the estimation of the quantile trend filtering model can be posed as the following optimization problem.
+For a single quantile level $\tau$ the estimation of the quantile trend filtering model can be posed as the following convex optimization problem.
 	\begin{eqnarray}
 	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \Mn{D}{k+1} \theta \rVert_1,
+	\underset{\V{\theta}}{\text{minimize}}\; \rho_\tau(\V{y} - \V{\theta}) + \lambda \lVert \Mn{D}{k+1} \V{\theta} \rVert_1,
 	\end{eqnarray}
@@
 We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$.
-Given quantiles $\{\tau_1, \ldots, \tau_J\}$ such that $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
+\Halley{Eric}{Please define $Q$.
+The reader can infer what it is, but it's better not to put that on the reader.} Given quantiles $\{\tau_1, \ldots, \tau_J\}$ such that $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
 
@@
 	\label{eq:noncross_trend}
-	\underset{\theta_1, \ldots, \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
-	\lambda_j \lVert \Mn{D}{k+1} \theta_j \rVert_1 \right ] \\
-	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le \ldots \le \theta_{J}(t) \text{ for all } t,
+	\underset{\V{\theta}_1, \ldots, \V{\theta}_J}{\text{minimize}}\; \sum_{j=1}^J \left [\rho_{\tau_j}(\V{y} - \V{\theta}_{j}) + 
+	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
+	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le \ldots \le \theta_{J}(t) \text{ for all } t, \nonumber
 	\end{eqnarray}
-
-	where $\theta_j \in \mathcal{R}^n$.
+	where $\V{\theta}_j \in \Real^n$.
 The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
 In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation, but it can easily be substituted for a free solver such as the Rglpk package by \cite{rglpk}.
-The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
+The number of parameters to be estimated in this problem is equal to the product of the number of observations $n$ and the number of quantiles of interest $J$.
 As the size of the data and the number of quantiles grows, all solvers will eventually break.
+\Halley{Eric}{Can we be more specific about what we mean when we say ``break?"}
 
@@
 	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
-	\cdots & \\
-	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n\\
+	\vdots & \\
+	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n,
 	\end{align*}
-
-	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \ldots< n$.
+with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots< n$.
 An illustration is given in \Fig{windows}.
 Given quantiles $\tau_1 < \ldots < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$.
-In order to write out the constraint that the overlapping sections must be equal we define a consensus variable
-
+In order to write out the constraint that the overlapping sections must be equal, we define a consensus variable
 	 \begin{align*}
@@
 		 \label{eq:quantile_windows}
-		 &\sum_{m=1}^M\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
+		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\min}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
 		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
@@
 	 \begin{equation*}
-	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
+	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \bar{\theta}_{j,m}) +
 	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
@@
 		\begin{algorithmic}		
-			\State Define $D = D^{(k+1)}$. 
+			\State Define $\M{D} = \Mn{D}{k+1}$. 
 			\State \textbf{initialize:} \\
-				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$. \\		
-			 	 $\omega_{j,m}^{(0)} = \mathbf{0}$	
+				 $\Vn{\theta}{0}_{j,m} \gets \arg \min \sum_{j=1}^J\rho_{\tau_j}(\V{y}_m - \V{\theta}_{j,m})+\lambda \lVert \M{D}\V{\theta}_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
+\\
+			 	 $\Vn{\omega}{0}_{j,m} \gets \V{0}$	
 			\Repeat{}
 			\State  		
-			$\bar{\theta}_{j,m}^{(q)} = g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
+			$\bar{\theta}_{j,m}^{(q)} \gets g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
 			\State 
-			$\omega_{j,m}^{(q)} = \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \bar{\theta}_{j,m}^{(q)})$	
+			$\omega_{j,m}^{(q)} \gets \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \bar{\theta}_{j,m}^{(q)})$	
 			\State
-				$\theta_{j,m}^{(q)} = \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
+				$\theta_{j,m}^{(q)} \gets \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
 			 subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
@@
 The primal residual,
-	\begin{equation}
-	r_p^{(q)} = \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
-	\end{equation}
+	\begin{eqnarray*}
+	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
+	\end{eqnarray*}
 	represents the difference between the trend values in the windows and the consensus trend value while the dual residual,
-	\begin{equation*}
-	r_d^{(q)} = \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
-	\end{equation*}
+	\begin{eqnarray*}
+	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
+	\end{eqnarray*}
 	represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when 
-	\begin{align}
-		&r_p^{(q)} < \epsilon_{abs}\sqrt{nJ} + \epsilon_{rel}\underset{m}{\max}\left[\max 
+	\begin{eqnarray*}
+		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{m}{\max}\left[\max 
 		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
-		&r_d^{(q)} < \epsilon_{abs}\sqrt{nJ} + \epsilon_{rel}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
-	\end{align}
+		r_d^{(q)} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
+	\end{eqnarray*}
 
@@
 		\centering
+		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
 		\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.}
-		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
 	\end{figure}
@@
 We use from one to four windows for each data size with an overlap of 500.
-The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
-As is shown in \Fig{timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
+The windows algorithm was run until the stopping criteria were met using $\epsilon_{\text{abs}} = 0.01$ and $\epsilon_{\text{rel}} = 0.001$.
+\Fig{timing} shows that using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
 The timing experiments were conducted on
@@
 	An important problem in trend estimation is the choice of regularization parameter or degree of smoothness.
-Our method can easily handle missing data by defining the check loss function to output 0 for missing values.
+Our method can easily handle missing data by defining the check-loss function to output 0 for missing values.
 This allows us to leave out validation observations that can be used to select the tuning parameter $\lambda$.
@@
 \cite{KoenkerNgPortnoy1994} addressed the choice of regularization parameter by proposing the Schwarz criterion for the selection of $\lambda$
-	\begin{equation}
+	\begin{eqnarray*}
 	\label{eq:SIC}
-	\mbox{SIC}(p_{\lambda}) = \log\left[\frac{1}{n}\rho_{\tau}(y - \theta)\right] + \frac{1}{2n}p_{\lambda}\log n.
-	\end{equation}
-	where $p_{\lambda} = \sum_t I(y(t) = \widehat{\theta}(t))$ is the number of interpolated points, which can be thought of as active knots.
+	\mbox{SIC}(p_{\lambda}) & = & \log\left[\frac{1}{n}\rho_{\tau}(y - \theta)\right] + \frac{1}{2n}p_{\lambda}\log n.
+	\end{eqnarray*}
+	where $p_{\lambda} = \sum_t \One(y(t) = \widehat{\theta}(t))$ is the number of interpolated points, which can be thought of as active knots.
 The SIC is based on the traditional Bayesian Information Criterion (BIC) which is given by
-	\begin{equation}
-	\mbox{BIC}(s) = -2\log(L\{\hat{\theta}\}) + \nu\log n 
-	\end{equation}	
+	\begin{eqnarray*}
+	\mbox{BIC}(s) & = & -2\log(L\{\hat{\theta}\}) + \nu\log n 
+	\end{eqnarray*}	
 	where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\hat{\theta}$.
-If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and assume that minimizing the checkloss function corresponds to maximizing the asymmetric Laplace likelihood, \begin{equation}
-	L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_t\rho_\tau\left(\frac{y(t) - \theta(t)}{\sigma}\right)\right\},
-	\end{equation} 
+If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and assume that minimizing the check-loss function corresponds to maximizing the asymmetric Laplace likelihood,
+	\begin{eqnarray*}
+	L(y \mid \theta) & = & \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_t\rho_\tau\left(\frac{y(t) - \theta(t)}{\sigma}\right)\right\},
+	\end{eqnarray*} 
 	we can compute the BIC as 
-	\begin{equation}
-	\mbox{BIC}(df) = 2\frac{1}{\sigma}\rho_{\tau}(y-\hat{\theta}) + \mbox{df}\log n
-	\end{equation} 
-	where $df$ is the number of non-zero elements of $D^{(k+1)}\hat{\theta}$. We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates.
+	\begin{eqnarray*}
+	\mbox{BIC}(\mbox{df}) & = & 2\frac{1}{\sigma}\rho_{\tau}(y-\hat{\theta}) + \mbox{df}\log n
+	\end{eqnarray*} 
+	where $df$ is the number of non-zero elements of $\Mn{D}{k+1}\Vhat{\theta}$.
+We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-\lvert 1-2\tau\rvert}{2}$ produces stable estimates.
 
 	Another criteria, the extended Bayesian Information Criteria (eBIC), specifically designed for large parameter spaces was proposed by \cite{chen2008}. 
-	\begin{equation}
+	\begin{eqnarray*}
 	\label{eq:eBIC}
-	\mbox{eBIC}_{\gamma}(s) = -2\log(L\{\hat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
-	\end{equation}
+	\mbox{eBIC}_{\gamma}(s) & = & -2\log(L\{\hat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
+	\end{eqnarray*}
 	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given model.
@@
 	\begin{figure}[h!]
+		\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
+		\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
+		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png} 
 		\caption{(Left) Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.
@@
 (Right) Criterion values by $\log(\lambda)$ vertical lines indicate locations of minima.}
-		\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
-		\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
-		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png} 
 		\label{fig:BIC}		
@@
 
-	We also compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC \Eqn{SIC}\citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria \Eqn{eBIC}.
+	We also compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC \Eqn{SIC}\citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check-loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria \Eqn{eBIC}.
 
@@
 		\includegraphics[width=.45\linewidth]{Figures/mixednorm.png}
+		\caption{Simulated data with true quantile trends.}			
 	\end{figure}
@@
 		\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
+		\caption{RMSE by design, method, quantile and data size. Points and error bars represent mean RMSE $\pm$ twice the standard error.}		
 		\label{fig:quantile_mse}
@@
 	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
-	\begin{equation}
-	Y(t) = s(t) + b(t) + \epsilon(t)
-	\end{equation} 
+	\begin{eqnarray*}
+	Y(t) & = & s(t) + b(t) + \epsilon(t)
+	\end{eqnarray*} 
 	with $t = 1, \ldots, n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, 0.25^2)$ is an error component.
@@
 	\begin{figure}
+		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}
 		\caption{RMSE by method, quantile and data size for peaks design.}			
-		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}	
 		\label{fig:peaks_rmse}
@@
 Defining $s(t) \in \{0,1\}$ as the vector of true signal classification and $\hat{s}(t) \in \{0,1\}$ as the estimated signal classification, the CAA is defined as
-	\begin{equation}
-	\mbox{CAA} = \frac{1}{2}\left(\frac{\sum_{t=1}^n \mathbf{I}[s(t) = 1 \cap \hat{s}(t)=1]}{\sum_{t=1}^n \mathbf{I}[s(t) = 1]} + \frac{\sum_{t=1}^n \mathbf{I}[s(t) = 0 \cap \hat{s}(t)=0]}{\sum_{t=1}^n \mathbf{I}[s(t) = 0]}\right).
-	\end{equation}  
+	\begin{eqnarray*}
+	\mbox{CAA} & = & \frac{1}{2}\left(\frac{\sum_{t=1}^n \One[s(t) = 1 \cap \hat{s}(t)=1]}{\sum_{t=1}^n \One[s(t) = 1]} + \frac{\sum_{t=1}^n \One[s(t) = 0 \cap \hat{s}(t)=0]}{\sum_{t=1}^n \One[s(t) = 0]}\right).
+	\end{eqnarray*}  
 	We use this metric because our classes tend to be very un-balanced with many more $0$s than $1$s.
@@
 	\begin{figure}[h!]
+		\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
 		\caption{Class averaged accuracy by threshold, data size, and method (1 is best 0.5 is worst).}		
-		\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
 	\end{figure}
@@
 The thresholds are calculated using the median plus a multiple of the median absolute deviation \Eqn{MAD} of the detrended series.
-	\begin{equation}
+	\begin{eqnarray}
 	\label{eq:MAD}
-	\mbox{MAD} = \mbox{median}(y-\tilde{y}),
-	\end{equation}
+	\mbox{MAD} & = & \mbox{median}(y-\tilde{y}),
+	\end{eqnarray}
 	where $\tilde{y}$ is the median of $y$.
@@
  	\begin{figure}
+	 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
 	 	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile. Horizontal dashed lines represent the thresholds.}		
-	 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
 	 	\label{fig:rugplot}
@@
 One metric for evaluating the distance between two classifications is the variation of information (VI):
-	\begin{align*}
-		r_{ij} =& \frac{1}{n}\sum_t \mathbf{I}(s_1(t) = i  \cap s_2(t) = j)\\
-		VI(s_1, s_2) =& -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_1(t) = i)}\right) + 
-		\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_2(t) = j)}\right) \right]
-	\end{align*}	
-	The VI is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
+	\begin{eqnarray*}
+		r_{ij} & = & \frac{1}{n}\sum_t \One(s_1(t) = i  \cap s_2(t) = j)\\
+		\VI(s_1, s_2) & = & -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \One(s_1(t) = i)}\right) + 
+		\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \One(s_2(t) = j)}\right) \right]
+	\end{eqnarray*}	
+	The $\VI$ is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
 The VIs by method, quantile and trend are shown in \Fig{vi}.
@@
 		\centering
+		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
 		\caption{Variation of Information between sensor nodes after trend removal by quantile and method and thresholding by factor of MAD.}		
-		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
 		\label{fig:vi}
@@
 	\begin{figure}
+		\includegraphics[width = \linewidth]{Figures/corrected_data.png}
 		\caption{Low cost sensor data after drift removal using windowed detrend with eBIC.}		
-		\includegraphics[width = \linewidth]{Figures/corrected_data.png}
 	\end{figure}
@@
 Our ADMM algorithm for large series both reduces the computing time and allows trends to be estimated on series that cannot be estimated simultaneously while our scaled extended BIC criteria was shown to provided better estimated of quantile trends in series with and without a signal component.
-We have also shown that the baseline drift in low cost air quality sensors can be removed through estimating quantile trends.
+We have also shown that the baseline drift in lost-cost air quality sensors can be removed through estimating quantile trends.
 

### 2019-02-08 Manuscript/detrendify_JCGS.tex
More of Eric's edits.

@@
 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots< n$.
-An illustration is given in \Fig{windows}.
+\Fig{windows} illustrates how we break the samples into overlapping windows.
 Given quantiles $\tau_1 < \ldots < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$.
@@
 		 \label{eq:quantile_windows}
-		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\min}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
+		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\text{minimize}}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
 		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
@@
 		\centering
+		\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
 		\caption{Window boundaries and trends fit separately in each window.}		
-		\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
 		\label{fig:windows}
@@
 	where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\hat{\theta}$.
-If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and assume that minimizing the check-loss function corresponds to maximizing the asymmetric Laplace likelihood,
+\Halley{Eric}{What is $s$?} If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and assume that minimizing the check-loss function corresponds to maximizing the asymmetric Laplace likelihood,
 	\begin{eqnarray*}
@@
 	where $df$ is the number of non-zero elements of $\Mn{D}{k+1}\Vhat{\theta}$.
-We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-\lvert 1-2\tau\rvert}{2}$ produces stable estimates.
+\Halley{Eric}{What is $\hat{\theta}$?
+This is the first time you used this notation.} We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-\lvert 1-2\tau\rvert}{2}$ produces stable estimates.
 

### 2019-02-08 Manuscript/detrendify_JCGS.tex
updated draft in responds to Joe's comments on 3-5

--- 2019-02-08 Manuscript/detrendify_JCGS.tex
@@
 \RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
-\usepackage{bm,graphicx,psfrag,epsf}
+\usepackage{graphicx}
 %\usepackage{enumerate}
@@
 \usepackage{algpseudocode}
-\usepackage{subfig} %% For \subfloat command
-
 %\pdfminorversion=4
@@
 \addtolength{\topmargin}{-.8in}%
-
-\usepackage[usenames]{color}
-\usepackage{xspace}
-
-%% For comments to each other
-\definecolor{blue}{rgb}{0.2,0.5,0.7}
-\definecolor{green}{rgb}{0.3,0.68,0.29}
-\definecolor{purple}{rgb}{0.6,0.31,0.64}
-\newcommand{\Halley}[2]{{\bf {\color{purple}@Halley from #1: #2}}\xspace}
-\newcommand{\Joe}[2]{{\bf {\color{blue}@Joe from #1: #2}}\xspace}
-\newcommand{\Eric}[2]{{\bf {\color{green}@Eric from #1: #2}}\xspace}
-
-%% Turn on/off options for PDF
-\usepackage{ifpdf}
-
-%% For hyperlinks. Should always be the last package. Omit for DVI version.
-\ifpdf
-\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
-\else
-\newcommand{\href}[2]{{#2}}
-\usepackage{url}
-\fi
-
-%% ----------------------------------------------------------------------
-%% Hyper-linked References
-%% ----------------------------------------------------------------------
-
-\ifpdf
-\newcommand{\Sec}[1]{\hyperref[sec:#1]{Section~\ref*{sec:#1}}} %section
-\newcommand{\App}[1]{\hyperref[sec:#1]{Appendix~\ref*{sec:#1}}} %appendix
-\newcommand{\Eqn}[1]{\hyperref[eq:#1]{{\rm (\ref*{eq:#1})}}} %equation
-\newcommand{\Part}[1]{\hyperref[part:#1]{(\ref*{part:#1})}} %part of theorem
-\newcommand{\Fig}[1]{\hyperref[fig:#1]{Figure~\ref*{fig:#1}}} %figure
-\newcommand{\Tab}[1]{\hyperref[tab:#1]{Table~\ref*{tab:#1}}} %table
-\newcommand{\Thm}[1]{\hyperref[thm:#1]{Theorem~\ref*{thm:#1}}} %theorem
-\newcommand{\Lem}[1]{\hyperref[lem:#1]{Lemma~\ref*{lem:#1}}} %lemma
-\newcommand{\Prop}[1]{\hyperref[prop:#1]{Proposition~\ref*{prop:#1}}} %proposition
-\newcommand{\Cor}[1]{\hyperref[cor:#1]{Corollary~\ref*{cor:#1}}} %corollary
-\newcommand{\Def}[1]{\hyperref[def:#1]{Definition~\ref*{def:#1}}} %definition
-\newcommand{\Alg}[1]{\hyperref[alg:#1]{Algorithm~\ref*{alg:#1}}} %algorithm
-\newcommand{\Ex}[1]{\hyperref[ex:#1]{Example~\ref*{ex:#1}}} %example
-\newcommand{\As}[1]{\hyperref[as:#1]{Assumption~{\rm\ref*{as:#1}}}} %assumption
-\newcommand{\Reg}[1]{\hyperref[as:#1]{Condition~\ref*{reg:#1}}} %regularity condition
-\newcommand{\AlgLine}[2]{\hyperref[alg:#1]{line~\ref*{line:#2} of Algorithm~\ref*{alg:#1}}}
-\newcommand{\AlgLines}[3]{\hyperref[alg:#1]{lines~\ref*{line:#2}--\ref*{line:#3} of Algorithm~\ref*{alg:#1}}}
-\else
-\newcommand{\Sec}[1]{{Section~\ref{sec:#1}}} %section
-\newcommand{\App}[1]{{Appendix~\ref{sec:#1}}} %appendix
-\newcommand{\Eqn}[1]{{(\ref{eq:#1})}} %equation
-\newcommand{\Part}[1]{{(\ref{part:#1})}} %part of proof
-\newcommand{\Fig}[1]{{Figure~\ref{fig:#1}}} %figure
-\newcommand{\Tab}[1]{{Table~\ref{tab:#1}}} %table
-\newcommand{\Thm}[1]{{Theorem~\ref{thm:#1}}} %theorem
-\newcommand{\Lem}[1]{{Lemma~\ref{lem:#1}}} %lemma
-\newcommand{\Prop}[1]{{Property~\ref{prop:#1}}} %property
-\newcommand{\Cor}[1]{{Corollary~\ref{cor:#1}}} %corollary
-\newcommand{\Def}[1]{{Definition~\ref{def:#1}}} %definition
-\newcommand{\Alg}[1]{{Algorithm~\ref{alg:#1}}} %algorithm
-\newcommand{\Ex}[1]{{Example~\ref{ex:#1}}} %example
-\newcommand{\As}[1]{{Assumption~\ref{as:#1}}} %assumption
-\newcommand{\Reg}[1]{{R~\ref{reg:#1}}} %regularity condition
-\newcommand{\AlgLine}[2]{{line~\ref{line:#2} of Algorithm~\ref{alg:#1}}}
-\newcommand{\AlgLines}[3]{{lines~\ref{line:#2}--\ref{line:#3} of Algorithm~\ref{alg:#1}}}
-\fi
-
-%% ----------------------------------------------------------------------
-%% Definitions
-%% ----------------------------------------------------------------------
-
-%% Shorthand for Real
-\newcommand{\Real}{\mathbb{R}}
-
-%% Math notation
-\newcommand{\Tra}{^{\sf T}} % Transpose
-\newcommand{\Inv}{^{-1}} % Inverse
-\newcommand{\tr}{\operatorname{tr}} % Trace
-\newcommand{\Ker}{\operatorname{Ker}} % Kernel
-\def\vec{\mathop{\rm vec}\nolimits}
-\newcommand{\amp}{\mathop{\:\:\,}\nolimits}
-\newcommand{\bic}{\text{BIC}}
-
-%% Vectors
-\newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} % vector
-\newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
-\newcommand{\Vbar}[1]{{\bm{\M{A}r \mathbf{\MakeLowercase{#1}}}}} % vector
-\newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
-\newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}} % vector
-\newcommand{\Vn}[2]{\V{#1}^{(#2)}} % n-th vector
-\newcommand{\VnE}[3]{{#1}^{(#2)}_{#3}} % n-th vector
-\newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}} % vector element
-
-%% Matrices
-\newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}} % matrix
-\newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} % matrix element
-\newcommand{\MC}[2]{\V{#1}_{#2}}
-\newcommand{\Mr}[2]{\V{#1}_{#2}} % matrix row
-\newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} % matrix
-\newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} % matrix
-\newcommand{\MhatC}[2]{\Vhat{#1}_{#2}} % matrix column
-\newcommand{\Mbar}[1]{{\bm{\M{A}r \mathbf{\MakeUppercase{#1}}}}} % matrix
-\newcommand{\MbarC}[2]{\Vbar{#1}_{#2}} % matrix column
-\newcommand{\Mn}[2]{\M{#1}^{(#2)}} % n-th matrix
-\newcommand{\Mbarn}[2]{\Mbar{#1}^{(#2)}} % n-th matrix
-\newcommand{\Mtilden}[2]{\Mtilde{#1}^{(#2)}} % n-th matrix
-\newcommand{\MnTra}[2]{\M{#1}^{(#2){{\sf T}}}} % n-th matrix transpose
-\newcommand{\MnE}[3]{\MakeLowercase{#1}^{(#2)}_{#3}} % n-th matrix element
-\newcommand{\MnC}[3]{\V{#1}^{(#2)}_{#3}} % n-th matrix column
-\newcommand{\MbarnC}[3]{\Vbar{#1}^{(#2)}_{#3}} % n-th matrix column
-\newcommand{\MnCTra}[3]{\V{#1}^{(#2){{\sf T}}}_{#3}} % n-th matrix column transpose
-\newcommand{\MCTra}[2]{\V{#1}^{{\sf T}}_{#2}} % n-th matrix column transpose
 
@@
 	\noindent%
-	{\it Keywords:} quantile regression, non-parametric regression, trend estimation, smoothing splines
+	{\it Keywords:} quantile regression, non-parametric, trend estimation, smoothing splines
 	\vfill
@@
 
-	In applications spanning the fields of chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components.
-\cite{Kim2009} addressed the case when the series consisted of just a trend and a random noise component and proposed the use of \textit{$\ell_1$-trend filtering} to estimate trends that piecewise linear or piecewise polynomial.
-\cite{Tib2014} later demonstrated that empirically the trend filtering estimates adapt to the local level of smoothness better than the more common smoothing splines \Halley{Eric}{If he proved theorems, minimax stuff, we should state that.}.
-In the trend filtering problem, given observations $y(t)$ with $t=1,\ldots,n$, the trend, $\theta(t)$, is estimated by solving the following convex optimization problem.
+	In applications spanning chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components.
+\cite{Kim2009} proposed \textit{$\ell_1$ trend filtering} to estimate trends that are piece-wise linear or polynomial, assuming that the series consists of a trend plus uncorrelated noise.
+\cite{Tib2014} later demonstrated empirically that the trend filtering estimates adapt to the local level of smoothness better than smoothing splines.
+In the trend filtering problem, given observations $y(t)$ with $t=1,...,N$, the trend, $\theta \in \mathbb{R}^n$, is estimated by solving the following convex problem:
 	\begin{eqnarray}
-	\underset{\V{\theta}}{\text{minimize}}\; \frac{1}{2} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
+	\underset{\theta}{\arg\min}\; \frac{1}{2} \lVert y - \theta \rVert_2^2 + \lambda \lVert \mathbf{D}^{(k+1)}\theta \rVert_1,
 	\end{eqnarray}
-	where $\lambda \geq 0$ is a regularization parameter and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
-To understand the purpose of penalizing $\Mn{D}{k+1}\V{\theta}$ consider the difference operator when $k = 0$.
+	where $y \in \mathbb{R}^n$ is the vector of observations, $\lambda \geq 0$ is a regularization parameter, and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
+To understand the purpose of penalizing $\mathbf{D}^{(k+1)}$ consider the difference operator when $k = 0$.
 	\begin{eqnarray}
-	\Mn{D}{1} & = & \begin{pmatrix}
+	\mathbf{D}^{(1)} = \begin{pmatrix}
 	-1 & 1 & 0 & \cdots & 0 & 0 \\
@@
 	0 & 0 & 0 & \cdots & -1 & 1 \\
-	\end{pmatrix}.
+	\end{pmatrix}
 	\end{eqnarray}
-	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
-The penalty term incentivizes solutions which are piece-wise constant.
-For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
+	Thus, $\lVert \mathbf{D}^{(1)}\theta \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_{i+1} - \theta_{i} \rvert$ which is just total variation in one dimension.
+The penalty incentivizes solutions that are piece-wise constant.
+For $k \geq 1$, the difference operator $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n-k-1) \times n}$ is defined recursively as follows
 	\begin{eqnarray}
-	\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
+	\mathbf{D}^{(k+1)} & = & \mathbf{D}^{(1)}\mathbf{D}^{(k)}.
 	\end{eqnarray}
-	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
+	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piece-wise polynomials of order $k$.
 
-	\Fig{tpn} shows that trend filtering provides excellent estimates of trends when the only components of the time series are the trend and random noise.
-In some applications, however, time series may also include a rapidly varying signal component in addition to the trend and noise.
-\Fig{tpspn} shows that in these cases the trend filtering estimate over-estimates the trend in the places where signal is present.
+	\cite{Tib2014} showed that trend filtering is both fast and locally adaptive when the time series consists of only the trend and random noise.
+However, in some applications, such as the air quality monitoring problem considered in this paper, the data contain a rapidly varying signal in addition to the slowly varying trend and noise.
+Standard trend filtering struggles to distinguish between the slowly varying trend and the rapidly-varying signal.
+We propose quantile trend filtering as a more robust method for detrending in these situations (Fig.
+\ref{fig:trendfilter}).
+To estimate the trend in the $\tau$\textsuperscript{th} quantile, we minimize the criterion
+	\begin{eqnarray}
+		\label{eq:quantile_trend}
+		\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \mathbf{D}^{(k+1)} \theta \rVert_1,
+	\end{eqnarray}	
+	where $\rho_\tau(z)$ is the check-loss function: 
+	\begin{equation}
+		\label{eq:checkloss}
+		\rho_{\tau}(z) = \sum_{i=1}^n z_i(\tau-\mathbf{I}(z_i<0)),
+	\end{equation}
+	 and $\mathbf{I}(\cdot) $ is the indicator function.
+This formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
+Here, we extend the basic framework to model multiple quantiles simultaneously.
+We also implement non-crossing constraints to ensure valid trend estimates.
+To reduce computation time and extend the method to larger series, we develop and a parallelizable ADMM algorithm.
+Finally, we propose a modified criterion for choosing the smoothing parameter.
 
-	\begin{figure}[t]
+	\begin{figure}
 		\centering
-		\subfloat[Slow Trend + Noise]{\label{fig:tpn}
-		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}}
-		\subfloat[Slow Trend + Spikes + Noise]{\label{fig:tpspn}
-		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}}
-		\caption{Two examples of trend filtering solutions. The true trend is shown in black, while the estimated trend is shown in red.}		
+		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
+		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}
+		\caption{Examples of trend filtering solutions (red) and 15\textsuperscript{th} quantile trending filtering solution (blue).}
 		\label{fig:trendfilter}
@@
 To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
-\Fig{raw_spod} provides an example of 3 co-located SPod sensor outputs from near an industrial facility.
-While all of the sensors response to the pollutant signal, the amount of baseline drift varies from one node to another.
-
-	\begin{figure}[b!]
+Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
+All of the sensors respond to the pollutant signal, but the amount of baseline drift varies from one to another.
+	\begin{figure}[t!]
+		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}		
 		\caption{Example of 3 co-located SPod PID sensor readings.}
-		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
 		\label{fig:raw_spod}
@@
 
-	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, it is not the mean trend that is desired but rather the baseline trend, which can be thought of as a trend in a low quantile of the data.
-\cite{Koenker1978} were the first to propose substituting the squared loss typically used in regression with the check loss function  \Eqn{checkloss} to estimate a conditional quantile instead of the conditional mean.
-
-	\begin{equation}
-	\label{eq:checkloss}
-	 \rho_{\tau}(z) = \sum_{i=1}^n z_i(1-\mathbf{I}(z_i<0)).
-	\end{equation}
-
-	Later, \cite{KoenkerNgPortnoy1994} addressed the issue of estimated quantile curves by combining the check loss penalty with trend filtering penalty with $k = 2$ which produces quantile trends that are piecewise linear.
-Other approaches to estimating quantile trends have also been proposed.
-Rather than using the trend filtering $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm solving the optimization problem in \Eqn{smoothingspline}.
+	In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, the goal is to estimate the trend in the baseline not the mean.
+We can define the trend in the baseline as the trend in a low quantile of the data.
+A variety of methods for estimating quantile trends have already been proposed.
+\cite{Koenker1978} were the first to propose substituting the squared-loss penalty with the check-loss penalty (Eq.
+\ref{eq:checkloss}) to estimate a conditional quantile instead of the conditional mean.
+Later, \cite{KoenkerNgPortnoy1994} proposed quantile trend filtering with $k = 2$ producing quantile trends that are piecewise linear, but did not extent it to higher orders.
+Rather than using the $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm:
 
@@
 	We combine the ideas of quantile regression and trend filtering.
-For a single quantile level $\tau$ the estimation of the quantile trend filtering model can be posed as the following optimization problem.
-	\begin{eqnarray}
-	\label{eq:quantile_trend}
-	\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \Mn{D}{k+1} \theta \rVert_1,
-	\end{eqnarray}
-	where $\lambda$ is a non-negative regularization parameter.
-We address the problem of choosing $\lambda$ in Section \ref{sec:lambda_choice}.
-As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of free or commercial solvers.
+For a single quantile level $\tau$, the quantile trend filtering problem is given in Eq.
+\ref{eq:quantile_trend}.
+As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of methods.
 In many cases, including ours, we are interested in estimating multiple quantiles simultaneously.
 We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$.
-Given quantiles $\{\tau_1, \ldots, \tau_J\}$ such that $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
+Given quantiles $\{\tau_1, ..., \tau_J\}$ such that $\tau_1 < \tau_2 < ...
+< \tau_J$, the optimization problem becomes
 
@@
 	\label{eq:noncross_trend}
-	\underset{\theta_1, \ldots, \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
-	\lambda_j \lVert \Mn{D}{k+1} \theta_j \rVert_1 \right ] \\
-	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le \ldots \le \theta_{J}(t) \text{ for all } t,
+	\underset{\theta_1, ..., \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
+	\lambda_j \lVert \mathbf{D}^{(k+1)} \theta_j \rVert_1 \right ] \\
+	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le ... \le \theta_{J}(t) \text{ for all } t,
 	\end{eqnarray}	
-
 	where $\theta_j \in \mathcal{R}^n$.
-The additional constraints are linear in the parameters so the non-crossing quantile trends can still be estimated by a number of available solvers.
-In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation, but it can easily be substituted for a free solver such as the Rglpk package by \cite{rglpk}.
-The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
-As the size of the data and the number of quantiles grows, all solvers will eventually break.
+The additional constraints are linear in the parameters, so the non-crossing quantile trends can still be estimated by a number of available solvers.
+In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation.
+However, we could easily substitute a free solver such as the Rglpk package by \cite{rglpk}.
 
@@
 
+	The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
+As the size of the data and the number of quantiles grows, all solvers will eventually break.
 To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
-We propose an alternating direction method of multipliers (ADMM) algorithm for solving large problems in a piecewise fashion.
+We propose an ADMM algorithm for solving large problems in a piece-wise fashion.
 The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is fully described by \cite{boyd2011distributed}.
-We apply the consensus ADMM algorithm to the the quantile regression trend filtering problem given in \Eqn{quantile_trend}, by dividing our observed series $y(t)$ with $t = \{1, \ldots, n\}$ into overlapping windows
+We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in Eq.
+(\ref{eq:quantile_trend}), by dividing our observed series $y(t)$ with $t = \{1, ..., n\}$ into overlapping windows
 
@@
 	\end{align*}
-
-	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \ldots< n$.
-An illustration is given in \Fig{windows}.
-Given quantiles $\tau_1 < \ldots < \tau_J$ to be estimated we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $M$ at time point $t$.
-In order to write out the constraint that the overlapping sections must be equal we define a consensus variable
+	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< n$.
+An illustration is given in Fig.
+\ref{fig:windows}.
+Given quantiles $\tau_1 < ...
+< \tau_J$, we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$ at time $t$.
+In order to constrain the overlapping sections to be equal we define a consensus variable
 
 	 \begin{align*}
-		 \bar{\theta}_{j,m} =&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
-		 =& \begin{cases} 
+	\bar{\theta}_{j, m}	=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
+	\bar{\theta}_{j, m}(t)	 =& \begin{cases} 
 			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
@@
 		 \label{eq:quantile_windows}
-		 &\sum_{m=1}^M\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
-		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
+		 &\underset{\theta_{1,m}, ..., \theta_{J,m}}{\min}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
+		 \lambda_j \lVert \mathbf{D}^{(k+1)} \theta_{j,m} \rVert_1 \right ] \\
 		 &\text{subject to: }\; \\
-		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
+		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le ... \le \theta_{J,m}(t) \text{ for all } m,t \\
 		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
 	 \end{align*}
-	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$ we can write the augmented Lagrangian for finding the trends in window $m$:
+	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$, the augmented Lagrangian for finding the trends in window $m$ is
 	 \begin{equation*}
-	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
-	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
+	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \mathbf{D}^{(k+1)}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
+	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2.
 	 \end{equation*}
-	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in algorithm 1.
+	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in Algorithm 1.
+The use of the Augmented Lagrangian converts the problem from a linear program into a quadratic program.
+The \texttt{gurobi} R package \citep{gurobi} can solve quadratic programs in addition to linear programs, but we can also use the free R package \texttt{quadprog} \citep{quadprog}.
 
@@
 			\State \textbf{initialize:} \\
-				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$. \\		
+				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < ...<\theta_{J,m}(t)$ for all $t$. \\		
 			 	 $\omega_{j,m}^{(0)} = \mathbf{0}$	
@@
 				$\theta_{j,m}^{(q)} = \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
-			 subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
+			 subject to $\theta_{1,m}(t) < ...<\theta_{J,m}(t)$ for all $t$.
 			\Until {convergence}
@@
 		\centering
+		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
 		\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.}
-		\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
 	\end{figure}
 
-	Timing experiments illustrate the advantages of using our ADMM algorithm even on datasets where solving the problem simultaneously is possible.
-For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below and trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using a $\lambda = n/5$.
+	Timing experiments illustrate the advantages of using our ADMM algorithm when the trend can be estimated with a single window.
+For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below.
+Trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using $\lambda = n/5$.
 We use from one to four windows for each data size with an overlap of 500.
 The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
-As is shown in \Fig{timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
-The timing experiments were conducted on
+As is shown in Fig.
+\ref{fig:timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
+The timing experiments were conducted on an Intel Xeon based Linux cluster using two processor cores.
 
@@
 We used this criteria with $\gamma = 1$, and $P=n-k-1$.
-We use a single dataset from our simulation study to illustrate the difference between the scaled, unscaled ($\sigma = 1$) and scaled extended BIC criteria in \Fig{BIC}.
+We use a single dataset from our simulation study to illustrate the difference between the scaled, unscaled ($\sigma = 1$) and scaled extended BIC criteria in Fig.
+\ref{fig:BIC}.
+In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
 
 	\begin{figure}[h!]
+		\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
+		\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
+		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png} 
 		\caption{(Left) Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.
@@
 
-	We compare the performance of our quantile trend filtering method with three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011, nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
+	We compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC (Eq.
+\ref{eq:SIC}) \citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria (Eq.
+\ref{eq:eBIC}).
+
+	We also compare the performance of our quantile trend filtering method with three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011, nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
 The regularization parameter $\lambda$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria as described in \cite{KoenkerNgPortnoy1994}, the regularization parameter for \texttt{qsreg} was chosen using generalized cross-validation based on the quantile criterion \cite{Oh2011}.
-
-	We also compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC \Eqn{SIC}\citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria \Eqn{eBIC}.
 
@@
 	To compare performance in estimating quantile trends, three simulation designs from \cite{Racine2017} were considered.
-For all designs $t = 1, \ldots, n$,  $x(t) = t/n$, and the response $y$ was generated as
+For all designs $t = 1, ..., n$,  $x(t) = t/n$, and the response $y$ was generated as
 	$$y(t) = sin(2\pi x(t)) + \epsilon(x(t))$$
@@
 	In all three designs the proposed detrend methods are either better than or comparable to existing methods.
-Overall the \texttt{detrend\_eBIC} performs best, and especially in the mixed normal design our methods have lower RMSEs for the 5\textsuperscript{th} and 95\textsubscript{th} quantiles.
-The \texttt{npqw} method performs particularly poorly in the mixed normal design due to the fact that it assumes the data comes from a scale-location model which is violated in this case.
+Overall the \texttt{detrend\_eBIC} performs best, and especially in the mixed normal design our methods have lower RMSEs for the 5\textsuperscript{th} and 95\textsuperscript{th} quantiles.
+The \texttt{npqw} method performs particularly poorly in the mixed normal design due to the fact that it assumes the data come from a scale-location model, which is violated in this case.
 
 	\begin{figure}
+		\includegraphics[width=\linewidth]{Figures/sim_metrics.png}	
 		\caption{RMSE by design, method, quantile and data size. Points and error bars represent mean RMSE $\pm$ twice the standard error.}
-		\includegraphics[width=\linewidth]{Figures/gaus_mse.png}	
-		\includegraphics[width=\linewidth]{Figures/shapebeta_mse.png}
-		\includegraphics[width=\linewidth]{Figures/mixednorm_mse.png}
 		\label{fig:quantile_mse}
@@
 %	\end{figure}
-	We use another simulation design based on the applied problem we aim to solve. We assume that the measured data can be represented by 
+	We use another simulation design based on our motivating application. We assume that the measured data can be represented by 
 	\begin{equation}
@@
 	\end{equation} 
-	with $t = 1, \ldots, n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, 0.25^2)$ is an error component.
+	with $t = 1, ..., n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, 0.25^2)$ is an error component.
 We generate $b(t)$ using cubic natural spline basis functions with degrees of freedom sampled from a Poisson distribution with mean parameter equal to $n/100$,  and coefficients drawn from an exponential distribution with rate 1.
@@
 One hundred datasets were generated for each $n=\{500, 1000, 2000, 4000\}$.
-We compare the ability of the methods to estimate the true quantiles of $Y(t)-s(t)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}).
+We compare the ability of the methods to estimate the true quantiles of $Y(t)-s(t)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (Fig.
+\ref{fig:peaks_rmse}).
 In this simulation study our \texttt{detrend\_eBIC} method outperforms the others substantially.
@@
 We compare three different quantiles for the baseline estimation and four different thresholds for classifying the signal after subtracting the estimated baseline from the observations.
-An illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal" is shown in \Fig{peaks_class_eg}.
+An illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal" is shown in Fig.
+\ref{fig:peaks_class_eg}.
 To compare the resulting signal classifications we calculate the class averaged accuracy (CAA).
@@
 	We use this metric because our classes tend to be very un-balanced with many more $0$s than $1$s.
-The CAA metric will always give a score of 0.5 for random guessing and also for trivial classifiers such as $\hat{s}(t) = 0$ for all $t$.
+The CAA metric will give a score of 0.5 for random guessing and also for trivial classifiers such as $\hat{s}(t) = 0$ for all $t$.
 
@@
 	\section{Application}
-
+	\label{sec:application}
 	The low-cost ``SPod" air quality sensors output a time series that includes a slowly varying baseline, high frequency random noise, and the sensor response to pollutants.
 A potential use for these sensors is to monitor pollutant concentrations at the fence lines of industrial facilities and detect time points when high concentrations are present.
-Ideally, three co-located and time aligned sensors (as shown in \Fig{raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
+Ideally, three co-located and time aligned sensors (as shown in Fig.
+\ref{fig:raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
 
 	\begin{figure}
+		\includegraphics[width = \linewidth]{Figures/short_trends.png}
 		\caption{Estimated 15\textsuperscript{th} quantile trends on subset of the data using \texttt{qsreg} and \texttt{detrend\_eBIC}.} 
-		\includegraphics[width = \linewidth]{Figures/short_trends.png}
 		\label{fig:short-trends}
@@
 We estimate the baseline trend using the 10\textsuperscript{th} and 15\textsuperscript{th} quantiles and compare three thresholds for classifying signal.
-The thresholds are calculated using the median plus a multiple of the median absolute deviation \Eqn{MAD} of the detrended series.
+The thresholds are calculated using the median plus a multiple of the median absolute deviation (Eq.
+\ref{eq:MAD}) of the detrended series.
 	\begin{equation}
 	\label{eq:MAD}
-	\mbox{MAD} = \mbox{median}(y-\tilde{y}),
+	\mbox{MAD} = \mbox{median}\lVert y-\tilde{y}\rVert,
 	\end{equation}
 	where $\tilde{y}$ is the median of $y$.
-Given a method, quantile level, and MAD multiple, we estimate the quantile trend for each of the sensor nodes and subtract it from the observations.
+Given a method, quantile level, and MAD multiple, we estimate the quantile trend for each of the sensors and subtract it from the observations.
 We then calculate the threshold using the median plus the MAD multiple of the corrected series and classify the corrected series based on the threshold.
-An example of the estimated baseline fit for each method is shown in \Fig{short-trends}, while \Fig{rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 3.
+An example of the estimated baseline fit for each method is shown in Fig.
+\ref{fig:short-trends}, while Fig.
+\ref{fig:rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 3.
 
  	\begin{figure}
+	 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
 	 	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile. Horizontal dashed lines represent the thresholds.}
-	 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
 	 	\label{fig:rugplot}
@@
 
-	Given the signal classifications for node 1, $s_1(t) \in \{0,1\}$ and node 2, $s_2(t)\in\{0,1\}$ we want to compare the similarity between the two classifications.
+	Given the signal classifications for SPod a, $s_a(t) \in \{0,1\}$ and SPod b, $s_b(t)\in\{0,1\}$ we want to compare the similarity between the two classifications.
 One metric for evaluating the distance between two classifications is the variation of information (VI):
 	\begin{align*}
-		r_{ij} =& \frac{1}{n}\sum_t \mathbf{I}(s_1(t) = i  \cap s_2(t) = j)\\
-		VI(s_1, s_2) =& -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_1(t) = i)}\right) + 
-		\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_2(t) = j)}\right) \right]
+		r_{ij} =& \frac{1}{n}\sum_t \mathbf{I}(s_a(t) = i  \cap s_b(t) = j)\\
+		VI(s_a, s_b) =& -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_a(t) = i)}\right) + 
+		\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_b(t) = j)}\right) \right]
 	\end{align*}	
 	The VI is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
-The VIs by method, quantile and trend are shown in \Fig{vi}.
+The VIs by method, quantile and trend are shown in Fig.
+\ref{fig:vi}.
 In all cases our \texttt{detrend} method results in classifications that are more similar than those from the \texttt{qsreg} method.
-As is illustrated in \Fig{short-trends} and \Fig{rugplot}, our method results in a smoother baseline estimate which improves signal classification.
+As is illustrated in Fig.
+\ref{fig:short-trends} and Fig.
+\ref{fig:rugplot}, our method results in a smoother baseline estimate which improves signal classification.
 
@@
 		\centering
-		\caption{Variation of Information between sensor nodes after trend removal by quantile and method and thresholding by factor of MAD.}
 		\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
+		\caption{Variation of Information between sensors by method (color), quantile (columns) and threshold MAD multiple (rows).}
 		\label{fig:vi}
@@
 	Our windowed \texttt{detrend\_eBIC} method was used to removed the baseline drift from the total dataset consisting of 86,401 observations per node.
-The VI scores for the full dataset were 0.36, 0.24, and 0.41 for nodes c and d, c and e, and d and e, respectively.
+The VI scores for the full dataset were 0.36, 0.24, and 0.41 for sensors a and b, a and c, and b and c, respectively.
 The complete confusion matrices for the nodes using a MAD multiple of 5 for the threshold is given in Table 1.

### 2019-02-08 Code/Fig_app_full.R

@@ -18,7 +18,8 @@ ggplot(spodRaw, aes(x=time, y=PID, col=node)) +
   geom_line(alpha=0.5) +
   theme_bw() +
   #facet_grid(node~., scales = "free")+
-  scale_color_brewer(palette = "Set1", labels = nodes)+
+  scale_color_brewer(palette = "Set1", labels = c("a", "b", "c"))+
+  labs(col="SPod")
   # xlim(c(as.POSIXct("2017-04-13 12:30:01 EST"), 
   #        as.POSIXct("2017-04-13 13:30:01 EST")))
 ggsave("../Manuscript/Figures/uncorrected_data.png", width = 7, height = 2.5)
@@ -30,8 +31,9 @@ spodLong <- spodPeaks %>% gather("node","PID", -time)
 
 ggplot(spodLong, aes(x=time, y=PID, col=node)) +
   geom_line(alpha=0.5) +
-  theme_bw() +
-  scale_color_brewer(palette = "Set1", labels = nodes)
+  theme_bw() +  
+  scale_color_brewer(palette = "Set1", labels = c("a", "b", "c"))+
+  labs(col="SPod")
 ggsave("../Manuscript/Figures/corrected_data.png", width = 7, height = 2.5)
 
 # ################################################################################
@@ -86,11 +88,11 @@ metric_df %>% filter(tau == 0.15, crit == 5, metric_type == "VI") %>% unnest()
 latex(conf,
       file = "../Manuscript/full_confusion_detrend.tex",
       rowlabel = "",
-      rowname = c("d = 0", "d = 1", "d=0", "d=1"),
-      cgroup = c("c = 0", "c = 1"),
-      colheads = rep(c("e = 0", "e = 1"),2),
+      rowname = c("b = 0", "b = 1", "b=0", "b=1"),
+      cgroup = c("a = 0", "a = 1"),
+      colheads = rep(c("c = 0", "c = 1"),2),
       n.cgroup = c(2,2),
-      caption = "Confusion matrices for 3 SPod nodes after baseline 
-      removal using 15th quantil and threshold of 3*MAD (n=52322).")
+      caption = "Confusion matrices for 3 SPods after baseline 
+      removal using 15th quantile and threshold of 5*MAD (n=86,401).")
 
 ################################################################################

### 2019-02-19 Manuscript/detrendify_JCGS.tex
Checking in edits to clarify the model and notation.

@@
 \RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
-\usepackage{graphicx}
+\usepackage{bbm}
+\usepackage{bm,graphicx,psfrag,epsf}
 %\usepackage{enumerate}
@@
 \addtolength{\topmargin}{-.8in}%
+
+\usepackage[usenames]{color}
+\usepackage{xspace}
+
+%% For comments to each other
+\definecolor{blue}{rgb}{0.2,0.5,0.7}
+\definecolor{green}{rgb}{0.3,0.68,0.29}
+\definecolor{purple}{rgb}{0.6,0.31,0.64}
+\newcommand{\Halley}[2]{{\bf {\color{purple}@Halley from #1: #2}}\xspace}
+\newcommand{\Joe}[2]{{\bf {\color{blue}@Joe from #1: #2}}\xspace}
+\newcommand{\Eric}[2]{{\bf {\color{green}@Eric from #1: #2}}\xspace}
+
+%% Turn on/off options for PDF
+\usepackage{ifpdf}
+
+%% For hyperlinks. Should always be the last package. Omit for DVI version.
+\ifpdf
+\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
+\else
+\newcommand{\href}[2]{{#2}}
+\usepackage{url}
+\fi
+
+%% ----------------------------------------------------------------------
+%% Hyper-linked References
+%% ----------------------------------------------------------------------
+
+\ifpdf
+\newcommand{\Sec}[1]{\hyperref[sec:#1]{Section~\ref*{sec:#1}}} %section
+\newcommand{\App}[1]{\hyperref[sec:#1]{Appendix~\ref*{sec:#1}}} %appendix
+\newcommand{\Eqn}[1]{\hyperref[eq:#1]{{\rm (\ref*{eq:#1})}}} %equation
+\newcommand{\Part}[1]{\hyperref[part:#1]{(\ref*{part:#1})}} %part of theorem
+\newcommand{\Fig}[1]{\hyperref[fig:#1]{Figure~\ref*{fig:#1}}} %figure
+\newcommand{\Tab}[1]{\hyperref[tab:#1]{Table~\ref*{tab:#1}}} %table
+\newcommand{\Thm}[1]{\hyperref[thm:#1]{Theorem~\ref*{thm:#1}}} %theorem
+\newcommand{\Lem}[1]{\hyperref[lem:#1]{Lemma~\ref*{lem:#1}}} %lemma
+\newcommand{\Prop}[1]{\hyperref[prop:#1]{Proposition~\ref*{prop:#1}}} %proposition
+\newcommand{\Cor}[1]{\hyperref[cor:#1]{Corollary~\ref*{cor:#1}}} %corollary
+\newcommand{\Def}[1]{\hyperref[def:#1]{Definition~\ref*{def:#1}}} %definition
+\newcommand{\Alg}[1]{\hyperref[alg:#1]{Algorithm~\ref*{alg:#1}}} %algorithm
+\newcommand{\Ex}[1]{\hyperref[ex:#1]{Example~\ref*{ex:#1}}} %example
+\newcommand{\As}[1]{\hyperref[as:#1]{Assumption~{\rm\ref*{as:#1}}}} %assumption
+\newcommand{\Reg}[1]{\hyperref[as:#1]{Condition~\ref*{reg:#1}}} %regularity condition
+\newcommand{\AlgLine}[2]{\hyperref[alg:#1]{line~\ref*{line:#2} of Algorithm~\ref*{alg:#1}}}
+\newcommand{\AlgLines}[3]{\hyperref[alg:#1]{lines~\ref*{line:#2}--\ref*{line:#3} of Algorithm~\ref*{alg:#1}}}
+\else
+\newcommand{\Sec}[1]{{Section~\ref{sec:#1}}} %section
+\newcommand{\App}[1]{{Appendix~\ref{sec:#1}}} %appendix
+\newcommand{\Eqn}[1]{{(\ref{eq:#1})}} %equation
+\newcommand{\Part}[1]{{(\ref{part:#1})}} %part of proof
+\newcommand{\Fig}[1]{{Figure~\ref{fig:#1}}} %figure
+\newcommand{\Tab}[1]{{Table~\ref{tab:#1}}} %table
+\newcommand{\Thm}[1]{{Theorem~\ref{thm:#1}}} %theorem
+\newcommand{\Lem}[1]{{Lemma~\ref{lem:#1}}} %lemma
+\newcommand{\Prop}[1]{{Property~\ref{prop:#1}}} %property
+\newcommand{\Cor}[1]{{Corollary~\ref{cor:#1}}} %corollary
+\newcommand{\Def}[1]{{Definition~\ref{def:#1}}} %definition
+\newcommand{\Alg}[1]{{Algorithm~\ref{alg:#1}}} %algorithm
+\newcommand{\Ex}[1]{{Example~\ref{ex:#1}}} %example
+\newcommand{\As}[1]{{Assumption~\ref{as:#1}}} %assumption
+\newcommand{\Reg}[1]{{R~\ref{reg:#1}}} %regularity condition
+\newcommand{\AlgLine}[2]{{line~\ref{line:#2} of Algorithm~\ref{alg:#1}}}
+\newcommand{\AlgLines}[3]{{lines~\ref{line:#2}--\ref{line:#3} of Algorithm~\ref{alg:#1}}}
+\fi
+
+%% ----------------------------------------------------------------------
+%% Definitions
+%% ----------------------------------------------------------------------
+
+%% Shorthand for Real
+\newcommand{\Real}{\mathbb{R}}
+
+%% Math notation
+\newcommand{\Tra}{^{\sf T}} % Transpose
+\newcommand{\Inv}{^{-1}} % Inverse
+\newcommand{\tr}{\operatorname{tr}} % Trace
+\newcommand{\Ker}{\operatorname{Ker}} % Kernel
+\def\vec{\mathop{\rm vec}\nolimits}
+\newcommand{\amp}{\mathop{\:\:\,}\nolimits}
+\newcommand{\bic}{\text{BIC}}
+\newcommand{\ebic}{\text{eBIC}}
+\newcommand{\sic}{\text{SIC}}
+\newcommand{\VI}{\text{VI}}
+\newcommand{\One}{\mathbbm{1}}
+
+%% Vectors
+\newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
+\newcommand{\Vbar}[1]{{\bm{\M{A}r \mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\Vn}[2]{\V{#1}^{(#2)}} % n-th vector
+\newcommand{\VnE}[3]{{#1}^{(#2)}_{#3}} % n-th vector
+\newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}} % vector element
+
+%% Matrices
+\newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} % matrix element
+\newcommand{\MC}[2]{\V{#1}_{#2}}
+\newcommand{\Mr}[2]{\V{#1}_{#2}} % matrix row
+\newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\MhatC}[2]{\Vhat{#1}_{#2}} % matrix column
+\newcommand{\Mbar}[1]{{\bm{\M{A}r \mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\MbarC}[2]{\Vbar{#1}_{#2}} % matrix column
+\newcommand{\Mn}[2]{\M{#1}^{(#2)}} % n-th matrix
+\newcommand{\Mbarn}[2]{\Mbar{#1}^{(#2)}} % n-th matrix
+\newcommand{\Mtilden}[2]{\Mtilde{#1}^{(#2)}} % n-th matrix
+\newcommand{\MnTra}[2]{\M{#1}^{(#2){{\sf T}}}} % n-th matrix transpose
+\newcommand{\MnE}[3]{\MakeLowercase{#1}^{(#2)}_{#3}} % n-th matrix element
+\newcommand{\MnC}[3]{\V{#1}^{(#2)}_{#3}} % n-th matrix column
+\newcommand{\MbarnC}[3]{\Vbar{#1}^{(#2)}_{#3}} % n-th matrix column
+\newcommand{\MnCTra}[3]{\V{#1}^{(#2){{\sf T}}}_{#3}} % n-th matrix column transpose
+\newcommand{\MCTra}[2]{\V{#1}^{{\sf T}}_{#2}} % n-th matrix column transpose
 
@@
 	\spacingset{1.5} % DON'T change the spacing!
+
+%% ----------------------------------------------------------------------
+%% Introduction
+%% ----------------------------------------------------------------------	
 \section{Introduction}
@@
 
-	In applications spanning chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar time series are observed and assumed to consist of a slowly varying trend and other more rapidly varying components.
-\cite{Kim2009} proposed \textit{$\ell_1$ trend filtering} to estimate trends that are piece-wise linear or polynomial, assuming that the series consists of a trend plus uncorrelated noise.
-\cite{Tib2014} later demonstrated empirically that the trend filtering estimates adapt to the local level of smoothness better than smoothing splines.
-In the trend filtering problem, given observations $y(t)$ with $t=1,...,N$, the trend, $\theta \in \mathbb{R}^n$, is estimated by solving the following convex problem:
-	\begin{eqnarray}
-	\underset{\theta}{\arg\min}\; \frac{1}{2} \lVert y - \theta \rVert_2^2 + \lambda \lVert \mathbf{D}^{(k+1)}\theta \rVert_1,
-	\end{eqnarray}
-	where $y \in \mathbb{R}^n$ is the vector of observations, $\lambda \geq 0$ is a regularization parameter, and the matrix $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
-To understand the purpose of penalizing $\mathbf{D}^{(k+1)}$ consider the difference operator when $k = 0$.
-	\begin{eqnarray}
-	\mathbf{D}^{(1)} = \begin{pmatrix}
+	In a wide range of applications that spans chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar functions of time $y(t)$ are observed and assumed to be a superposition of an underlying slowly varying trend $\theta(t)$, other more rapidly varying components $s(t)$, and noise.
+In practice, $y(t)$ is observed at discrete time points $t_1, \ldots, t_n$, and we model the vector of signal samples $\VE{y}{i} = y(t_i)$ as
+\begin{eqnarray*}
+\V{y} & = & \V{\theta} + \V{s} + \V{\varepsilon},
+\end{eqnarray*}
+where $\VE{\theta}{i} = \theta(t_i)$, $\VE{s}{i} = s(t_i)$, and $\V{\varepsilon} \in \Real^n$ is a vector of uncorrelated noise.
+For notational simplicity, for the rest of the paper, we assume that the time points take on the values $t_i = i$, but it is straightforward to generalize to an arbitrary grid of time points.
+
+In some applications, the slowly varying component $\V{\theta}$ is the signal of interest and the transient component $\V{s}$ is a vector of nuisance parameters.
+In other applications, the roles of $\V{\theta}$ and $\V{s}$ are reversed.
+The latter applies in the air quality applications motivating this work, as $\V{s}$ represents the signal of interest and $\V{\theta}$ represents a baseline drift that obscures the identification of the important transient events encoded in $\V{s}$.
+In either case, the estimation task is to demix the noisy observed vector $\V{y}$ into a slowly varying component $\V{\theta}$ and a transient component $\V{s}$.
+
+To tackle this demixing problem, we introduce a scalable baseline estimation framework by building on \textit{$\ell_1$-trend filtering}, a relatively new nonparametric estimation framework.
+\cite{Kim2009} originally proposed \textit{$\ell_1$-trend filtering} to estimate trends with piecewise polynomial functions, assuming that the observed time series $\V{y}$ consists of a trend $\V{\theta}$ plus uncorrelated noise $\V{\varepsilon}$, namely $\V{y} = \V{\theta} + \V{\varepsilon}$.
+The estimated trend is the solution to the following convex optimization problem
+\begin{eqnarray*}
+\underset{\V{\theta}}{\text{minimize}}\; \frac{1}{2} \lVert \V{y} - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
+\end{eqnarray*}
+	where $\lambda$ is a nonnegative regularization parameter, and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
+To understand the purpose of penalizing $\Mn{D}{k+1}$ consider the difference operator when $k = 0$.
+	\begin{eqnarray*}
+	\Mn{D}{1} & = & \begin{pmatrix}
+
 	-1 & 1 & 0 & \cdots & 0 & 0 \\
@@
 	0 & 0 & 0 & \cdots & -1 & 1 \\
-	\end{pmatrix}
-	\end{eqnarray}
-	Thus, $\lVert \mathbf{D}^{(1)}\theta \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_{i+1} - \theta_{i} \rvert$ which is just total variation in one dimension.
-The penalty incentivizes solutions that are piece-wise constant.
-For $k \geq 1$, the difference operator $\mathbf{D}^{(k+1)} \in \mathbb{R}^{(n-k-1) \times n}$ is defined recursively as follows
-	\begin{eqnarray}
-	\mathbf{D}^{(k+1)} & = & \mathbf{D}^{(1)}\mathbf{D}^{(k)}.
-	\end{eqnarray}
-	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piece-wise polynomials of order $k$.
 
-	\cite{Tib2014} showed that trend filtering is both fast and locally adaptive when the time series consists of only the trend and random noise.
-However, in some applications, such as the air quality monitoring problem considered in this paper, the data contain a rapidly varying signal in addition to the slowly varying trend and noise.
-Standard trend filtering struggles to distinguish between the slowly varying trend and the rapidly-varying signal.
-We propose quantile trend filtering as a more robust method for detrending in these situations (Fig.
-\ref{fig:trendfilter}).
-To estimate the trend in the $\tau$\textsuperscript{th} quantile, we minimize the criterion
+	\end{pmatrix}.
+	\end{eqnarray*}
+	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
+The penalty term incentivizes solutions which are piecewise constant.
+For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
+	\begin{eqnarray*}
+	\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
+	\end{eqnarray*}
+	Penalizing the 1-norm of the vector $\Mn{D}{k+1}\V{\theta}$ produces estimates of $\V{\theta}$ that are piecewise polynomials of order $k$.
+
+%	$k+1$ fold composition of the discrete difference operator produces estimates that are piecewise polynomials of order $k$.
+
+%\cite{Tib2014} later demonstrated empirically that the trend filtering estimates adapt to the local level of smoothness better than smoothing splines and also proved that with a judicious choice of tuning parameter the trend filtering estimate converges to the true underlying function at the minimax rate for functions whose $k$th derivative is of bounded variation.
+%	
+%	Given observations $y(t)$ with $t=1,...,N$ and defining $y \in \Real^n$ such that $y_i = y(t)$, in the trend filtering problem, the trend, $\V{\theta} \in \Real^n$, is estimated by solving the following convex problem:
+%	\begin{eqnarray}
+%	\underset{\V{\theta}}{\arg\min}\; \frac{1}{2} \lVert y - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
+%	\end{eqnarray}
+%	where $y \in \Real^n$ is the vector of observations, $\lambda \geq 0$ is a regularization parameter, and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
+To understand the purpose of penalizing $\Mn{D}{k+1}$ consider the difference operator when $k = 0$.
+%	\begin{eqnarray*}
+%	\Mn{D}{1} & = & \begin{pmatrix}
+%
+%	-1 & 1 & 0 & \cdots & 0 & 0 \\
+%	0 & -1 & 1 & \cdots & 0 & 0 \\
+%	\vdots & & & & & \\
+%	0 & 0 & 0 & \cdots & -1 & 1 \\
+%
+%	\end{pmatrix}.
+%	\end{eqnarray*}
+%	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
+The penalty term incentivizes solutions which are piecewise constant.
+For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
+%	\begin{eqnarray*}
+%	\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
+%	\end{eqnarray*}
+%	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
+
+	\cite{Tib2014} proved that with a judicious choice of $\lambda$ the trend filtering estimate converges to the true underlying function at the minimax rate for functions whose $k$th derivative is of bounded variation and showed that trend filtering is both fast and locally adaptive when the time series consists of only the trend and random noise, which is illustrated in \Fig{tpn}.
+As noted earlier though in some applications, such as the air quality monitoring problem considered in this paper, the data contain a rapidly varying signal in addition to the slowly varying trend and noise.
+Standard trend filtering struggles to distinguish between the slowly varying trend and the rapidly-varying signal, illustrated in \Fig{tpspn}.
+
+	To account for the presence of a transient component in the observed time series $\V{y}$, we propose quantile trend filtering \Fig{tpspn}.% as a more robust method for detrending in these situations \Fig{tpspn}.
+	To estimate the trend in the $\tau$\textsuperscript{th} quantile, we solve the convex optimization problem
 	\begin{eqnarray}
 	\label{eq:quantile_trend}
-		\underset{\theta}{\min}\; \rho_\tau(y - \theta) + \lambda \lVert \mathbf{D}^{(k+1)} \theta \rVert_1,
+	\underset{\V{\theta}}{\text{minimize}}\; \rho_\tau(\V{y} - \V{\theta}) + \lambda \lVert \Mn{D}{k+1} \V{\theta} \rVert_1,
 	\end{eqnarray}
-	where $\rho_\tau(z)$ is the check-loss function: 
-	\begin{equation}
-		\label{eq:checkloss}
-		\rho_{\tau}(z) = \sum_{i=1}^n z_i(\tau-\mathbf{I}(z_i<0)),
-	\end{equation}
-	 and $\mathbf{I}(\cdot) $ is the indicator function.
-This formulation was proposed by \cite{Kim2009} as a possible extension of $\ell_1$-trend filtering but not studied.
-Here, we extend the basic framework to model multiple quantiles simultaneously.
-We also implement non-crossing constraints to ensure valid trend estimates.
+	where $\rho_\tau(\V{r})$ is the check function
+	\begin{eqnarray}
+	\label{eq:check}
+	\rho_{\tau}(\V{r}) & = & \sum_{i=1}^n \VE{r}{i}(1-\One(\VE{r}{i}<0)),
+	\end{eqnarray}
+	and $\One(A)$ is 1 if its input $A$ is true and 0 otherwise.
+\cite{Kim2009} proposed using the check function as a possible extension of $\ell_1$-trend filtering but did not investigate it further.
+Here, we develop the basic $\ell_1$-quantile-trend-filtering framework and extend it to model multiple quantiles simultaneously with non-crossing constraints to ensure valid trend estimates.
 To reduce computation time and extend the method to larger series, we develop and a parallelizable ADMM algorithm.
@@
 		\centering
-		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}
-		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}
-		\caption{Examples of trend filtering solutions (red) and 15\textsuperscript{th} quantile trending filtering solution (blue).}
-		\label{fig:trendfilter}
+		\subfloat[Slow Trend + Noise]{\label{fig:tpn}
+		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg1.png}}
+		\subfloat[Slow Trend + Spikes + Noise]{\label{fig:tpspn}
+		\includegraphics[width = 0.45\linewidth]{Figures/trend_filter_eg2.png}}
+		\caption{Examples of trend filtering solutions (red) and 15\textsuperscript{th} quantile trending filtering solution (blue).
+Standard trend filtering performs well in the no-signal case (right) but struggles to distinguish between the slowly varying trend and the rapidly-varying signal (left).
+The quantile trend is not affected by the signal and provides an estimate of the baseline.}
 	\end{figure}
 
-	Our research is motivated by the analysis of low cost and portable air quality sensors, whose use has increased dramatically in the last decade.
+	This work was motivated by the analysis of signals measured by low cost and portable air quality sensors, whose use has increased dramatically in the last decade.
 These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
@@
 Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
-All of the sensors respond to the pollutant signal, but the amount of baseline drift varies from one to another.
+All of the sensors respond to the pollutant signal, the three sharp transient spikes at \Halley{Eric}{can you put the times at which the spikes occur here?}, but the amount of baseline drift varies from one sensor to another obscuring the detection of the peaks that herald the presence of pollutants.
+We show later that by estimating the baseline drift in each sensor using $\ell_1$-quantile-trend filtering and removing it from the observed signals, peaks can be reliably detected from concordant residual signals from a collection of SPods using a simple data-driven thresholding strategy.
+
 	\begin{figure}[t!]
@@
 A variety of methods for estimating quantile trends have already been proposed.
-\cite{Koenker1978} were the first to propose substituting the squared-loss penalty with the check-loss penalty (Eq.
-\ref{eq:checkloss}) to estimate a conditional quantile instead of the conditional mean.
-Later, \cite{KoenkerNgPortnoy1994} proposed quantile trend filtering with $k = 2$ producing quantile trends that are piecewise linear, but did not extent it to higher orders.
-Rather than using the $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the $\ell_2$-norm:
+\cite{Koenker1978} were the first to propose substituting the sum-of-squares term with the check function  \Eqn{check} to estimate a conditional quantile instead of the conditional mean.
+Later, \cite{KoenkerNgPortnoy1994} proposed quantile trend filtering with $k = 2$ producing quantile trends that are piecewise linear, but they did not consider extensions to higher order differences.
+Rather than using the $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the square of the $\ell_2$-norm:
 
@@
 	\label{eq:smoothingspline}
-	\rho_{\tau}(y - f) + \lambda\int (f''(t))^2 dt, 
+	\sum_{t=1}^n\rho_{\tau}(y(t) - \theta(t)) + \lambda\int \theta''(t)^2 dt, 
  	\end{equation*}	
-	where $f(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degree of smoothing.
-\cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check-loss function with a similar differentiable function.
+	where $\theta(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degree of smoothing.
+\cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check function with a similar differentiable function.
 \cite{Racine2017} take a different approach and constrain the response to follow a smooth location scale model of the form $y(t) = a(t) + b(t)\epsilon_i$ and estimate the $\tau$\textsuperscript{th} conditional quantile using a kernel estimator and local linear approach.
 
-	In the following sections we describe in detail our quantile trend filtering algorithms and regularization parameter criterion (Section \ref{sec:methods}).
+	In the rest of the paper, we detail our quantile trend filtering algorithms as well as how to choose the smoother parameter $\lambda$ (Section \ref{sec:methods}).
 We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods (Section \ref{sec:simluation}).
@@
 
+%% ----------------------------------------------------------------------
+%% Methods
+%% ----------------------------------------------------------------------
 \section{Methods}
 \label{sec:methods}
+%% ----------------------------------------------------------------------
+%% Quantile Trend Filtering
+%% ----------------------------------------------------------------------
 \subsection{Quantile Trend Filtering}
@@
 	We combine the ideas of quantile regression and trend filtering.
-For a single quantile level $\tau$, the quantile trend filtering problem is given in Eq.
-\ref{eq:quantile_trend}.
+For a single quantile level $\tau$, the quantile trend filtering problem is given in \Eqn{quantile_trend}.
 As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of methods.
 In many cases, including ours, we are interested in estimating multiple quantiles simultaneously.
-We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$.
-Given quantiles $\{\tau_1, ..., \tau_J\}$ such that $\tau_1 < \tau_2 < ...
-< \tau_J$, the optimization problem becomes
-
-	\begin{eqnarray}
+We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$ where $Q$ is the quantile function of $\V{y}$.
+Given quantiles $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
+	\begin{eqnarray*}
 	\label{eq:noncross_trend}
-	\underset{\theta_1, ..., \theta_J}{\min}\; \sum_{j=1}^J \left [\rho_{\tau_j}(y - \theta_{j}) + 
-	\lambda_j \lVert \mathbf{D}^{(k+1)} \theta_j \rVert_1 \right ] \\
-	 \text{subject to: }\; \theta_{1}(t) \le \theta_{2}(t) \le ... \le \theta_{J}(t) \text{ for all } t,
-	\end{eqnarray}	
-	where $\theta_j \in \mathcal{R}^n$.
+%	&&\underset{\V{\theta}_1, \ldots, \V{\theta}_J}{\text{minimize}}\; \sum_{j=1}^J \left [\rho_{\tau_j}(\V{y} - \V{\theta}_{j}) + 
+%	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\	
+	&&\underset{\M{\Theta} \in \mathcal{C}}{\text{minimize}}\; \sum_{j=1}^J \left [\rho_{\tau_j}(\V{y} - \V{\theta}_{j}) + 
+	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
+%	&&\text{subject to: }\; \V{\theta}_1 \leq \V{\theta}_2 \leq \ldots \leq \V{\theta}_J	
+%	\theta_{1}(t) \le \theta_{2}(t) \le \ldots \le \theta_{J}(t) \text{ for all } t, \nonumber
+	\end{eqnarray*}
+	where $\M{\Theta} = \begin{pmatrix} \V{\theta}_1 & \V{\theta}_2 & \cdots & \V{\theta}_J \end{pmatrix} \in \Real^{n \times J}$ is a matrix whose $j$th column corresponds to the $j$th quantile signal $\V{\theta}_j$ and the set $\mathcal{C} = \{\M{\Theta} \in \Real^{n \times J}: \ME{\theta}{ij} \leq \ME{\theta}{ij'} \;\text{for $j \leq j'$}\}$ encodes the non-crossing quantile constraints.
 The additional constraints are linear in the parameters, so the non-crossing quantile trends can still be estimated by a number of available solvers.
-In the rest of this paper we rely on the commercial solver Gurobi \citep{gurobi} and its R package implementation.
+In the rest of this paper, we report numerical results using the commercial solver Gurobi \citep{gurobi} and its R package implementation.
 However, we could easily substitute a free solver such as the Rglpk package by \cite{rglpk}.
 
+%% ----------------------------------------------------------------------
+%% ADMM for Big Data
+%% ----------------------------------------------------------------------
 	\subsection{ADMM for Big Data}
@@
 	The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
-As the size of the data and the number of quantiles grows, all solvers will eventually break.
+As the size of the data and the number of quantiles grows, all solvers will eventually break \Halley{Eric}{Can we be more specific about what we mean when we say ``break?"}.
 To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
-We propose an ADMM algorithm for solving large problems in a piece-wise fashion.
+We propose an ADMM algorithm for solving large problems in a piecewise fashion.
 The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is fully described by \cite{boyd2011distributed}.
-We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in Eq.
-(\ref{eq:quantile_trend}), by dividing our observed series $y(t)$ with $t = \{1, ..., n\}$ into overlapping windows
+Let $\mathcal{I}$ denote a subset of the indices $\{1, \ldots, n\}$ and $\V{y}_\mathcal{I} \in \Real^{\lvert I \rvert}$ denote the subvector obtained by stacking on top of each other the elements $\VE{y}{i}$ for $i \in \mathcal{I}$ in their canonical ordering.
+	We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in \Eqn{quantile_trend}, by dividing our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into overlapping $M$ windows of observations $\Vn{y}{m} = \V{y}_{\mathcal{I}_m}$ where $\mathcal{I}_m$ is the set of integers from $l_m$ to $u_m$ inclusive, namely $\mathcal{I}_m = \{l_m, l_m + 1, \ldots, u_m -1, u_m\}$ and $1 = l_{1} < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots < u_{M} = n$.
+\Fig{windows} shows an example of 1200 observations being mapped into three equally sized overlapping windows of observations.
 
-	\begin{align*}
-	y_1(t) = y(t) & \mbox{~~if~~} 1 \le t \le u_{1}\\
-	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
-	\cdots & \\
-	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n\\
-	\end{align*}
-	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< n$.
-An illustration is given in Fig.
-\ref{fig:windows}.
-Given quantiles $\tau_1 < ...
-< \tau_J$, we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$ at time $t$.
-In order to constrain the overlapping sections to be equal we define a consensus variable
+%	\begin{align*}
+%	y_1(t) = y(t) & \mbox{~~if~~} 1 \le t \le u_{1}\\
+%	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
+%	\vdots & \\
+%	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n,
+%	\end{align*}
+%
+%	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< n$. 
+	 Given quantiles $\tau_1 < \cdots < \tau_J$, we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$ at time $t$.
+In order to constrain the overlapping sections to be equal, we define a consensus variable
 
-	 \begin{align*}
-	\bar{\theta}_{j, m}	=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
-	\bar{\theta}_{j, m}(t)	 =& \begin{cases} 
+	 \begin{eqnarray*}
+	\bar{\theta}_{j, m}	&=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
+	\bar{\theta}_{j, m}(t)	 &=& \begin{cases} 
 			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
@@
 			 \end{cases},
-	\end{align*}
+	\end{eqnarray*}
+%	 
+%	 \begin{align*}
+%	\bar{\theta}_{j, m}	=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
+%	\bar{\theta}_{j, m}(t)	 =& \begin{cases} 
+%			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
+%			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
+%			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}  
+%			 \end{cases},
+%	\end{align*}
 	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
@@
 		 \label{eq:quantile_windows}
-		 &\underset{\theta_{1,m}, ..., \theta_{J,m}}{\min}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(y_m - \theta_{j,m}) + 
-		 \lambda_j \lVert \mathbf{D}^{(k+1)} \theta_{j,m} \rVert_1 \right ] \\
+		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\text{minimize}}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m}) + 
+		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
 		 &\text{subject to: }\; \\
-		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le ... \le \theta_{J,m}(t) \text{ for all } m,t \\
+		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
 		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
@@
 	 \begin{equation*}
-	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert \mathbf{D}^{(k+1)}\theta_{j,m}\rVert_1 +  \omega_{j,m}^T(\theta_{j,m} - \bar{\theta}_{j,m}) +
-	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2.
+	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \bar{\theta}_{j,m}) +
+	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
 	 \end{equation*}
@@
 		\begin{algorithmic}		
-			\State Define $D = D^{(k+1)}$. 
+			\State Define $\M{D} = \Mn{D}{k+1}$. 
 			\State \textbf{initialize:} \\
-				 $\theta_{j,m}^{(0)} = \arg \min \sum_{j=1}^J\rho_{\tau_j}(y_m - \theta_{j,m})+\lambda \lVert D\theta_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < ...<\theta_{J,m}(t)$ for all $t$. \\		
-			 	 $\omega_{j,m}^{(0)} = \mathbf{0}$	
+				 $\Vn{\theta}{0}_{j,m} \gets \arg \min \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{m} - \V{\theta}_{j,m})+\lambda \lVert \M{D}\V{\theta}_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
+\\
+			 	 $\Vn{\omega}{0}_{j,m} \gets \V{0}$	
 			\Repeat{}
 			\State  		
-			$\bar{\theta}_{j,m}^{(q)} = g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
+			$\bar{\theta}_{j,m}^{(q)} \gets g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
 			\State 
-			$\omega_{j,m}^{(q)} = \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \bar{\theta}_{j,m}^{(q)})$	
+			$\omega_{j,m}^{(q)} \gets \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \bar{\theta}_{j,m}^{(q)})$	
 			\State
-				$\theta_{j,m}^{(q)} = \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
-			 subject to $\theta_{1,m}(t) < ...<\theta_{J,m}(t)$ for all $t$.
+				$\theta_{j,m}^{(q)} \gets \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
+			 subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
 			\Until {convergence}
@@
 The primal residual,
-	\begin{equation}
-	r_p^{(q)} = \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
-	\end{equation}
+	\begin{eqnarray*}
+	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
+	\end{eqnarray*}
 	represents the difference between the trend values in the windows and the consensus trend value. The dual residual,
-	\begin{equation*}
-	r_d^{(q)} = \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
-	\end{equation*}
+	\begin{eqnarray*}
+	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
+	\end{eqnarray*}
 	represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when 
-	\begin{align}
-		&r_p^{(q)} < \epsilon_{abs}\sqrt{nJ} + \epsilon_{rel}\underset{m}{\max}\left[\max 
+	\begin{eqnarray*}
+		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{m}{\max}\left[\max 
 		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
-		&r_d^{(q)} < \epsilon_{abs}\sqrt{nJ} + \epsilon_{rel}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
-	\end{align}
+		r_d^{(q)} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
+	\end{eqnarray*}
 
@@
 The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
-As is shown in Fig.
-\ref{fig:timing}, using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
+\Fig{timing} shows that using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
 The timing experiments were conducted on an Intel Xeon based Linux cluster using two processor cores.
@@
 	\label{sec:lambda_choice}
-	An important problem in trend estimation is the choice of regularization parameter, or degree of smoothness.
+
+	An important practical issue in trend estimation is the choice of regularization parameter, or degree of smoothness.
 Our method can easily handle missing data by defining the check loss function to output 0 for missing values, allowing us to implement cross-validation.
@@
 \cite{KoenkerNgPortnoy1994} addressed the choice of regularization parameter by proposing the Schwarz criterion for the selection of $\lambda$
+	\begin{eqnarray*}
+	\label{eq:SIC}
+	\mbox{SIC}(p_{\lambda}) & = & \log\left[\frac{1}{n}\rho_{\tau}(y - \theta)\right] + \frac{1}{2n}p_{\lambda}\log n.
+	\end{eqnarray*}
+	where $p_{\lambda} = \sum_i \One(\VE{y}{i} = \widehat{\theta}_i)$ is the number of interpolated points, which can be thought of as active knots.
+The SIC is based on the traditional Bayesian Information Criterion (BIC) which is given by
 	\begin{equation}
-	\label{eq:SIC}
-	\mbox{SIC}(p_{\lambda}) = \log\left[\frac{1}{n}\rho_{\tau}(y - \theta)\right] + \frac{1}{2n}p_{\lambda}\log n.
+	\mbox{BIC}(\nu) = -2\log(L\{\Vhat{\theta}\}) + \nu\log n 
 	\end{equation}	
-	where $p_{\lambda} = \sum_t I(y(t) = \widehat{\theta}(t))$ is the number of interpolated points, which can be thought of as active knots.
-The SIC is based on the traditional Bayesian Information Criterion,
-	\begin{equation}
-	\mbox{BIC}(\nu) = -2\log(L\{\hat{\theta}\}) + \nu\log n 
-	\end{equation}	
-	where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\hat{\theta}$.
-If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and view minimizing the checkloss function as maximizing the asymmetric Laplace likelihood, \begin{equation}
-	L(y|\theta) = \left(\frac{\tau^n(1-\tau)}{\sigma}\right)^n\exp\left\{-\sum_t\rho_\tau\left(\frac{y(t) - \theta(t)}{\sigma}\right)\right\},
-	\end{equation} 
+	where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\Vhat{\theta}$.
+If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and view minimizing the check function as maximizing the asymmetric Laplace likelihood,
+
+	\begin{eqnarray*}
+	L(y \mid \theta) & = & \left[\frac{\tau^n(1-\tau)}{\sigma}\right]^n\exp\left\{-\sum_i\rho_\tau\left(\frac{\VE{y}{i} - \VE{\theta}{i}}{\sigma}\right)\right\},
+	\end{eqnarray*} 
 	we can compute the BIC as 
-	\begin{equation}
-	\mbox{BIC}(\nu) = 2\frac{1}{\sigma}\rho_{\tau}(y-\hat{\theta}) + \nu\log n
-	\end{equation} 
-	where $\nu$ is the number of non-zero elements of $\mathbf{D}^{(k+1)}\hat{\theta}$.
+	\begin{eqnarray*}
+	\mbox{BIC}(\nu) = 2\frac{1}{\sigma}\rho_{\tau}(\V{y}-\Vhat{\theta}) + \nu\log n
+	\end{eqnarray*} 
+	where $\Vhat{\theta}$ is the estimated trend, and $\nu$ is the number of non-zero elements of $\Mn{D}{k+1}\Vhat{\theta}$.
 We can choose any $\sigma>0$ and have found empirically that $\sigma =  \frac{1-|1-2\tau|}{2}$ produces stable estimates.
 
-	Another criteria, the extended Bayesian Information Criteria (eBIC), specifically designed for large parameter spaces was proposed by \cite{chen2008}. 
-	\begin{equation}
+	\cite{chen2008} proposed the extended Bayesian Information Criteria (eBIC), specifically designed for large parameter spaces. 
+	\begin{eqnarray*}
 	\label{eq:eBIC}
-	\mbox{eBIC}_{\gamma}(s) = -2\log(L\{\hat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
-	\end{equation}
+	\mbox{eBIC}_{\gamma}(s) & = & -2\log(L\{\Vhat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
+	\end{eqnarray*}
 	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given model.
 We used this criteria with $\gamma = 1$, and $P=n-k-1$.
-We use a single dataset from our simulation study to illustrate the difference between the scaled, unscaled ($\sigma = 1$) and scaled extended BIC criteria in Fig.
-\ref{fig:BIC}.
-In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
+\Fig{BIC} illustrates the difference among the scaled, unscaled ($\sigma = 1$), and scaled extended BIC criteria
+	when applied to a single dataset replicate from our simulation study. In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
 
@@
 		\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png} 
-		\caption{(Left) Degrees of freedom (number of non-zero elements of $D\theta$) by $\log(\lambda)$.
-(Middle) Estimated 10th quantile trend with regularization parameter chosen using various criterion.
-(Right) Criterion values by $\log(\lambda)$ vertical lines indicate locations of minima.}
+		\caption{(Left) Estimated degrees of freedom (number of non-zero elements of $\M{D}\V{\theta}$) as a function of $\log(\lambda)$.
+(Middle) Estimated 10th quantile trend with regularization parameter chosen using various criteria.
+(Right) Criteria values as a function of $\log(\lambda)$, vertical lines indicate locations of minima.}
 		\label{fig:BIC}
@@
 
+%% ----------------------------------------------------------------------
+%% Simulation Studies
+%% ----------------------------------------------------------------------
 \section{Simulation Studies}
 \label{sec:simluation}
+
 	We conduct two simulation studies to compare the performance of our quantile trend filtering method and regularization parameter selection criteria with previously published methods.
@@
 
-	We compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC (Eq.
-\ref{eq:SIC}) \citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria (Eq.
-\ref{eq:eBIC}).
+	We compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC \Eqn{SIC} \citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria \Eqn{eBIC}.
 
@@
 	To compare performance in estimating quantile trends, three simulation designs from \cite{Racine2017} were considered.
-For all designs $t = 1, ..., n$,  $x(t) = t/n$, and the response $y$ was generated as
-	$$y(t) = sin(2\pi x(t)) + \epsilon(x(t))$$
+For all designs $t = 1, \ldots, n$,  $x(t) = t/n$, and the response $\V{y}$ was generated as
+\begin{eqnarray*}
+y(t) & = & \sin(2\pi x(t)) + \epsilon(x(t))
+\end{eqnarray*}
 	The three error distributions considered were 
@@
 For each quantile trend and method the root mean squared error was calculated as $\mbox{RMSE} = \sqrt{\frac{1}{n}\sum_t (\hat{q}_{\tau}(t) - q_\tau(t))^2}$.
-The mean RMSE $\pm$ twice the standard error for each method, quantile level and sample size is shown in Figure \ref{fig:quantile_mse}.
+\Fig{quantile_mse} shows the mean RMSE $\pm$ twice the standard error for each method, quantile level, and sample size.
 	In all three designs the proposed detrend methods are either better than or comparable to existing methods.
@@
 	We use another simulation design based on our motivating application. We assume that the measured data can be represented by 
-	\begin{equation}
-	Y(t) = s(t) + b(t) + \epsilon(t)
-	\end{equation} 
-	with $t = 1, ..., n$, where $s(t)$ is the true signal at time $t$, $b(t)$ is the drift component that varies smoothly over time and $\epsilon(t) \sim N(0, 0.25^2)$ is an error component.
-We generate $b(t)$ using cubic natural spline basis functions with degrees of freedom sampled from a Poisson distribution with mean parameter equal to $n/100$,  and coefficients drawn from an exponential distribution with rate 1.
-The true signal function is assumed to be zero with peaks generated using the Gaussian density function.
+	\begin{eqnarray*}
+	y(t_i) & = & \theta(t_i) + s(t_i) + \VE{\varepsilon}{i},
+	\end{eqnarray*}  
+	where $t_i = i$ for $i = 1, ..., n$, $\theta(t)$ is the drift component that varies smoothly over time, $s(t)$ is the true signal at time $t$, and $\VE{\varepsilon}{i}$ are i.i.d.\@ errors distributed as $N(0, 0.25^2)$.
+We generate $\theta(t)$ using cubic natural spline basis functions with degrees of freedom sampled from a Poisson distribution with mean parameter equal to $n/100$,  and coefficients drawn from an exponential distribution with rate 1.
+The true signal function $s(t)$ is assumed to be zero with peaks generated using the Gaussian density function.
 The number of peaks is sampled from a binomial distribution with size equal to $n$ and probability equal to $0.005$ with location parameters uniformly distributed between $1$ and $n-1$ and bandwidths uniformly distributed between $2$ and $12$.
@@
 One hundred datasets were generated for each $n=\{500, 1000, 2000, 4000\}$.
-We compare the ability of the methods to estimate the true quantiles of $Y(t)-s(t)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (Fig.
-\ref{fig:peaks_rmse}).
-In this simulation study our \texttt{detrend\_eBIC} method outperforms the others substantially.
-The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets but its performance deteriorates as the data size grows.
+We compare the ability of the methods to estimate the true quantiles of $y(t_i)-s(t_i)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}).
+In this simulation study, our proposed method \texttt{detrend\_eBIC} method substantially outperforms the others.
+The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets, but its performance deteriorates as the data size grows.
 The \texttt{npqw} and \texttt{detrend\_valid} methods both perform poorly on this design.
@@
 		\includegraphics[width = \linewidth]{Figures/peaks_mse.png}	
-		\caption{RMSE by method, quantile and data size for peaks design.}
+		\caption{RMSE by method, quantile, and data size for peaks design.}
 		\label{fig:peaks_rmse}
@@
 
-	In our application, we want to accurately classify the observations into signal and no signal using a threshold.
+	While minimizing RMSE is desirable in general, in our application, the primary metric of success is accurately classifying observations $\VE{y}{i}$ as signal or no signal.
 To evaluate the accuracy of our method compared to other methods we define true signal as any time point when the simulated peak value is greater than 0.5.
 We compare three different quantiles for the baseline estimation and four different thresholds for classifying the signal after subtracting the estimated baseline from the observations.
-An illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal" is shown in Fig.
-\ref{fig:peaks_class_eg}.
-To compare the resulting signal classifications we calculate the class averaged accuracy (CAA).
-Defining $s(t) \in \{0,1\}$ as the vector of true signal classification and $\hat{s}(t) \in \{0,1\}$ as the estimated signal classification, the CAA is defined as
-	\begin{equation}
-	\mbox{CAA} = \frac{1}{2}\left(\frac{\sum_{t=1}^n \mathbf{I}[s(t) = 1 \cap \hat{s}(t)=1]}{\sum_{t=1}^n \mathbf{I}[s(t) = 1]} + \frac{\sum_{t=1}^n \mathbf{I}[s(t) = 0 \cap \hat{s}(t)=0]}{\sum_{t=1}^n \mathbf{I}[s(t) = 0]}\right).
-	\end{equation}  
-	We use this metric because our classes tend to be very un-balanced with many more $0$s than $1$s.
-The CAA metric will give a score of 0.5 for random guessing and also for trivial classifiers such as $\hat{s}(t) = 0$ for all $t$.
+\Fig{peaks_class_eg} shows an illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal." Let $\delta_i \in \{0,1\}$ be the vector of true signal classifications and $\hat{\delta}_i \in \{0,1\}$ be the vector of estimated signal classifications, namely $\hat{\delta}_i = \One(\VE{y}{i} - \hat{\theta}_{i} > 0.5)$.
+
+	To compare the resulting signal classifications, we calculate the class averaged accuracy (CAA), which is defined as
+	\begin{eqnarray*}
+	\mbox{CAA} & = & \frac{1}{2}\left[\frac{\sum_{i=1}^n \One[\delta_i = 1 \cap \hat{\delta}_i = 1]}{\sum_{i=1}^n \One[\delta_i = 1]} + \frac{\sum_{i=1}^n \One[\delta_i = 0 \cap \hat{\delta}_i=0]}{\sum_{t=i=1}^n \One[\delta_i = 0]}\right].
+	\end{eqnarray*}  
+	We use this metric because our classes tend to be very imbalanced with many more zeros than ones.
+The CAA metric will give a score of 0.5 for random guessing and also for trivial classifiers such as $\hat{\delta}_i = 0$ for all $i$.
 
@@
 
-	\section{Application}
+%% ----------------------------------------------------------------------
+%% Application
+%% ----------------------------------------------------------------------	
+	\section{Application in Air Quality Monitoring}
+
 	\label{sec:application}
-	The low-cost ``SPod" air quality sensors output a time series that includes a slowly varying baseline, high frequency random noise, and the sensor response to pollutants.
-A potential use for these sensors is to monitor pollutant concentrations at the fence lines of industrial facilities and detect time points when high concentrations are present.
-Ideally, three co-located and time aligned sensors (as shown in Fig.
-\ref{fig:raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
+	The low-cost ``SPod" air quality sensors output a time series that includes a slowly varying baseline, 
+	the sensor response to pollutants, and high frequency random noise.
+These sensors are used to monitor pollutant concentrations at the fence lines of industrial facilities and detect time points when high concentrations are present.
+Ideally, three co-located and time aligned sensors (as shown in \Fig{raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
 
@@
 We estimate the baseline trend using the 10\textsuperscript{th} and 15\textsuperscript{th} quantiles and compare three thresholds for classifying signal.
-The thresholds are calculated using the median plus a multiple of the median absolute deviation (Eq.
-\ref{eq:MAD}) of the detrended series.
-	\begin{equation}
+The thresholds are calculated using the median plus a multiple of the median absolute deviation \Eqn{MAD} of the detrended series.
+	\begin{eqnarray}
 	\label{eq:MAD}
 	\mbox{MAD} = \mbox{median}\lVert y-\tilde{y}\rVert,
-	\end{equation}
+	\end{eqnarray}
 	where $\tilde{y}$ is the median of $y$.
@@
 We then calculate the threshold using the median plus the MAD multiple of the corrected series and classify the corrected series based on the threshold.
-An example of the estimated baseline fit for each method is shown in Fig.
-\ref{fig:short-trends}, while Fig.
-\ref{fig:rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 3.
+\Fig{short-trends} shows an example of the estimated baseline fit for each method, while \Fig{rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 3.
 
@@
 
-	Given the signal classifications for SPod a, $s_a(t) \in \{0,1\}$ and SPod b, $s_b(t)\in\{0,1\}$ we want to compare the similarity between the two classifications.
+	Given the signal classifications for SPods a and b, $\delta^{(a)}_i \in \{0,1\}$ and $\delta^{(b)}_i\in\{0,1\}$, we want to compare the similarity between the two classifications.
 One metric for evaluating the distance between two classifications is the variation of information (VI):
-	\begin{align*}
-		r_{ij} =& \frac{1}{n}\sum_t \mathbf{I}(s_a(t) = i  \cap s_b(t) = j)\\
-		VI(s_a, s_b) =& -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_a(t) = i)}\right) + 
-		\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \mathbf{I}(s_b(t) = j)}\right) \right]
-	\end{align*}	
-	The VI is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
-The VIs by method, quantile and trend are shown in Fig.
-\ref{fig:vi}.
-In all cases our \texttt{detrend} method results in classifications that are more similar than those from the \texttt{qsreg} method.
-As is illustrated in Fig.
-\ref{fig:short-trends} and Fig.
-\ref{fig:rugplot}, our method results in a smoother baseline estimate which improves signal classification.
+	\begin{eqnarray*}
+	r_{ij} & = & \frac{1}{n}\sum_t \One\left(\delta^{(a)}_i = i  \cap \delta^{(b)}_i = j\right)\\
+	\VI(s_a, s_b) & = & -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \One(\delta^{(a)}_i = i)}\right) + 
+	\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \One(\delta^{(b)}_i = j)}\right) \right]
+	\end{eqnarray*}		
+	The $\VI$ is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
+\Fig{vi} shows plots the VIs by method, quantile, and trend.
+In all cases, our \texttt{detrend} method results in classifications that are more similar than those from the \texttt{qsreg} method.
+\Fig{short-trends} and \Fig{rugplot} illustrate that our method results in a smoother baseline estimate which improves signal classification.
 
@@
 
+%% ----------------------------------------------------------------------
+%% Conclusion and Discussion
+%% ----------------------------------------------------------------------	
 \section{Conclusion and Discussion}
 \label{sec:discussion}
+
 	We have expanded the quantile trend filtering method by implementing a non-crossing constraint and a new algorithm for processing large series, and proposing a modified criteria for smoothing parameter selection.

### 2019-02-22 Manuscript/detrendify_JCGS.tex
Editing consensus ADMM section.

@@
 The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is fully described by \cite{boyd2011distributed}.
+
 	Let $\mathcal{I}$ denote a subset of the indices $\{1, \ldots, n\}$ and $\V{y}_\mathcal{I} \in \Real^{\lvert I \rvert}$ denote the subvector obtained by stacking on top of each other the elements $\VE{y}{i}$ for $i \in \mathcal{I}$ in their canonical ordering.
-	We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in \Eqn{quantile_trend}, by dividing our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into overlapping $M$ windows of observations $\Vn{y}{m} = \V{y}_{\mathcal{I}_m}$ where $\mathcal{I}_m$ is the set of integers from $l_m$ to $u_m$ inclusive, namely $\mathcal{I}_m = \{l_m, l_m + 1, \ldots, u_m -1, u_m\}$ and $1 = l_{1} < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots < u_{M} = n$.
+	We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in \Eqn{quantile_trend}, by dividing our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into overlapping $M$ windows of observations $\Vn{y}{m} = \V{y}_{\mathcal{I}_m}$ where $\mathcal{I}_m$ is the set of integers from $l_m$ to $u_m$ inclusive, namely $\mathcal{I}_m = \{l_m, l_m + 1, \ldots, u_m -1, u_m\}$ where $1 = l_{1} < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots < u_{M} = n$.
 \Fig{windows} shows an example of 1200 observations being mapped into three equally sized overlapping windows of observations.
@@
 %	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< n$. 
-	 Given quantiles $\tau_1 < \cdots < \tau_J$, we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$ at time $t$.
+%	 Given quantiles $\tau_1 < \cdots < \tau_J$, we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$ at time $t$.
 In order to constrain the overlapping sections to be equal, we define a consensus variable
+	 Given quantiles $\tau_1 < \cdots < \tau_J$, we introduce dummy variables $\Vn{\theta}{m}_{j} \in \Real^{\lvert \mathcal{I}_m \rvert}$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$.
+We then ``stitch" together the $M$ quantile trend estimates into consensus over the overlapping regions by introducing the constraint $\VnE{\theta}{m}{ij} = \VE{\theta}{i + l_m - 1,j}$ for $i = 1, \ldots, u_m - l_m + 1$ and for all $j$.
+Let $\Mn{\Theta}{m}$ be the matrix whose $j$th column is $\Vn{\theta}{m}$.
+We can write these constraints more compactly as $\Mn{\Theta}{m} = \Mn{U}{m}\M{\Theta}$, where $\Mn{U}{m} \in \{0, 1\}^{\lvert \mathcal{I}_m \rvert \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $m$th window.
+Furthermore, let $\iota_{\mathcal{C}}$ denote the indicator function of the non-crossing quantile constraint, namely $\iota_{\mathcal{C}}(\M{\Theta})$ is zero if $\M{\Theta} \in \mathcal{C}$ and infinity otherwise.
 
+	Our windowed quantile trend optimization problem can then be written as 
+	 \begin{align*}
+		 \label{eq:quantile_windows}
+		 &\text{minimize}\; \sum_{m=1}^M \left \{\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{m} - \Vn{\theta}{m}_{j}) + 
+		 \lambda_j \lVert \Mn{D}{k+1} \Vn{\theta}{m}_{j} \rVert_1 \right ] + \iota_\mathcal{C}(\Mn{\Theta}{m}) \right\}\\
+		 &\text{subject to} \qquad \Mn{\Theta}{m} = \Mn{U}{m}\M{\Theta} \;\; \text{ for } m = 1, \ldots, M.
+	 \end{align*}
+
+	 	 Let $\Mn{\Omega}{m}$ denote the Lagrange multiplier matrix for the $m$th consensus constraint, namely $\Mn{\theta}{m} = \Mn{U}{m}\M{\theta}$, and let $\Vn{\omega}{m}_{j}$ denote its $j$th column.
+%		 
+%		  $\Vn{\omega}{m}_{j}$ denote Lagrange multipliers for the constraint $\Vn{\theta}{m}_j = \Mn{U}{m}\V{\theta}_j$, $\Mn{\Omega}{m}$ denote the matrix whose $j$th column is $\Vn{\omega}{m}_j$, and penalty parameter $\gamma > 0$,
+
+		  The augmented Lagrangian is given by
 	 \begin{eqnarray*}
-	\bar{\theta}_{j, m}	&=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
-	\bar{\theta}_{j, m}(t)	 &=& \begin{cases} 
-			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
-			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
-			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}  
-			 \end{cases},
+	 \mathcal{L}(\M{\Theta}, \{\Mn{\Theta}{m}\}, \{\Mn{\Omega}{m}\}) & = & \sum_{m=1}^M \biggl \{\sum_{j=1}^J \biggl [\rho_{\tau_j}(\Vn{y}{m} - \Vn{\theta}{m}_j)+\lambda \lVert \Mn{D}{k+1}\Vn{\theta}{m}_j\rVert_1 \\
+	 && + \; (\Vn{\theta}{m}_j - \Mn{U}{m}\V{\theta}_{j})\Tra\Vn{\omega}{m}_{j} + 
+	 \frac{\gamma}{2}\lVert \Vn{\theta}{m}_j - \Mn{U}{m}\V{\theta}_{j}\rVert_2^2 \biggr ] + \iota_\mathcal{C}(\Mn{\Theta}{m}) \biggr\},
 	 \end{eqnarray*}
+	 where $\gamma$ is a positive penalty parameter. 
+	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in Algorithm 1.
+The use of the Augmented Lagrangian converts the problem from a linear program into a quadratic program.
+The \texttt{gurobi} R package \citep{gurobi} can solve quadratic programs in addition to linear programs, but we can also use the free R package \texttt{quadprog} \citep{quadprog}.
+
+%	 \begin{eqnarray*}
+%	\bar{\theta}_{j, m}	&=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
+%	\bar{\theta}_{j, m}(t)	 &=& \begin{cases} 
+%			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
+%			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
+%			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}  
+%			 \end{cases},
+%	\end{eqnarray*}
 %	 
@@
 %	\end{align*}
-	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
-	 \begin{align*}
-		 \label{eq:quantile_windows}
-		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\text{minimize}}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m}) + 
-		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
-		 &\text{subject to: }\; \\
-		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
-		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
-	 \end{align*}
-	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$, the augmented Lagrangian for finding the trends in window $m$ is
-	 \begin{equation*}
-	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \bar{\theta}_{j,m}) +
-	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
-	 \end{equation*}
-	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in Algorithm 1.
+%	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as 
+%	 \begin{align*}
+%		 \label{eq:quantile_windows}
+%		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\text{minimize}}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m}) + 
+%		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
+%		 &\text{subject to: }\; \\
+%		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
+%		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
+%	 \end{align*}
+%	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$, the augmented Lagrangian for finding the trends in window $m$ is
+%	 \begin{equation*}
+%	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \bar{\theta}_{j,m}) +
+%	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
+%	 \end{equation*}
+%	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in Algorithm 1.
 The use of the Augmented Lagrangian converts the problem from a linear program into a quadratic program.
 The \texttt{gurobi} R package \citep{gurobi} can solve quadratic programs in addition to linear programs, but we can also use the free R package \texttt{quadprog} \citep{quadprog}.
-
+%	 
 	\begin{figure}[!h] 
@@
 The primal residual,
+%	\begin{eqnarray*}
+%	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
+%	\end{eqnarray*}
 	\begin{eqnarray*}
-	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
+	r_{\text{primal}}^{k} & = & \sqrt{\sum_{m=1}^M\lVert\Mn{\Theta}{m}_{k} - \M{\Theta}_k\rVert_{\text{F}}^2},
 	\end{eqnarray*}
 	represents the difference between the trend values in the windows and the consensus trend value. The dual residual,
+%	\begin{eqnarray*}
+%	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
+%	\end{eqnarray*}
 	\begin{eqnarray*}
-	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
+	r_{\text{dual}}^{k} & = & \gamma\sqrt{\sum_{m=1}^M \lVert\M{\Theta}_k - \M{\Theta}_{k-1}\rVert_{\text{F}}^2},
 	\end{eqnarray*}
 	represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when 
+%	\begin{eqnarray*}
+%		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{m}{\max}\left[\max 
+%		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
+%		r_d^{(q)} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
+%	\end{eqnarray*}
 	\begin{eqnarray*}
-		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{m}{\max}\left[\max 
-		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
-		r_d^{(q)} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
+		r_{\text{primal}}^{k} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\underset{m}{\max} \left[\max 
+		\left\{\lVert\Mn{\Theta}{m}_k \rVert_{\text{F}}, \lVert \M{\Theta}_{k} \rVert_{\text{F}} \right\}\right] \\%\lVert \M{\Theta}_{k} \rVert_{\text{F}}} \right )\right]\\
+		r_{\text{dual}}^{k} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\sqrt{\sum_{m=1}^M
+		\lVert \Mn{\Omega}{m}_k \rVert_{\text{F}}^2}.
 	\end{eqnarray*}

### 2019-03-06 Manuscript/detrendify_JCGS.tex

@@
 \usepackage{subfig} %% For \subfloat command
+
+\newtheorem{proposition}{Proposition}[section]
 
@@
 \newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
-\newcommand{\Vbar}[1]{{\bm{\M{A}r \mathbf{\MakeLowercase{#1}}}}} % vector
+\newcommand{\Vbar}[1]{{\bm{\bar \mathbf{\MakeLowercase{#1}}}}} % vector
 \newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
@@
 \newcommand{\MhatC}[2]{\Vhat{#1}_{#2}} % matrix column
-\newcommand{\Mbar}[1]{{\bm{\M{A}r \mathbf{\MakeUppercase{#1}}}}} % matrix
+\newcommand{\Mbar}[1]{{\bm{bar \mathbf{\MakeUppercase{#1}}}}} % matrix
 \newcommand{\MbarC}[2]{\Vbar{#1}_{#2}} % matrix column
@@
 
-	In a wide range of applications that spans chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar functions of time $y(t)$ are observed and assumed to be a superposition of an underlying slowly varying trend $\theta(t)$, other more rapidly varying components $s(t)$, and noise.
+	In a wide range of applications that spans chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar functions of time $y(t)$ are observed and assumed to be a superposition of an underlying slowly varying baseline trend $\theta(t)$, other more rapidly varying components $s(t)$, and noise.
 In practice, $y(t)$ is observed at discrete time points $t_1, \ldots, t_n$, and we model the vector of signal samples $\VE{y}{i} = y(t_i)$ as
@@
 The latter applies in the air quality applications motivating this work, as $\V{s}$ represents the signal of interest and $\V{\theta}$ represents a baseline drift that obscures the identification of the important transient events encoded in $\V{s}$.
-In either case, the estimation task is to demix the noisy observed vector $\V{y}$ into a slowly varying component $\V{\theta}$ and a transient component $\V{s}$.
+We describe this motivating problem further to illustrate the critical importance of being able to separate $\V{s}$ from $\V{\theta}$.
 
-To tackle this demixing problem, we introduce a scalable baseline estimation framework by building on \textit{$\ell_1$-trend filtering}, a relatively new nonparametric estimation framework.
+%	This work was motivated by the analysis of signals measured by low cost and portable air quality sensors, whose use has increased dramatically in the last decade. 
+In the last decade, low cost and portable air quality sensors have enjoyed dramatically increased usage.
+These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
+The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
+Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
+To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
+Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
+All of the sensors respond to the pollutant signal, the three sharp transient spikes at \Halley{Eric}{can you put the times at which the spikes occur here?}, but the baseline drift varies from one sensor to another obscuring the detection of the peaks that alert the intrusion of pollutants.
+We show later that by estimating the baseline drift in each sensor %using $\ell_1$-quantile-trend filtering
+ and  removing it from the observed signals, peaks can be reliably detected from concordant residual signals from a collection of SPods using a simple data-driven thresholding strategy.
+Thus, accurately demix a noisy observed vector $\V{y}$ into a slowly varying component $\V{\theta}$ and a transient component $\V{s}$ can lead to greatly improved and simplified downstream analysis.
+This is true for not just air quality applications, but universally across the many domains listed earlier.
+
+	\begin{figure}[t!]
+		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}		
+		\caption{Example of 3 co-located SPod PID sensor readings.}
+		\label{fig:raw_spod}
+	\end{figure}
+
+To tackle demixing problems, we introduce a scalable baseline estimation framework by building on \textit{$\ell_1$-trend filtering}, a relatively new nonparametric estimation framework.
+Our contributions are three-fold.
+\begin{itemize}
+\item \cite{Kim2009} proposed using the check function as a possible extension of $\ell_1$-trend filtering but did not investigate it further.
+Here, we develop the basic $\ell_1$-quantile-trend-filtering framework and extend it to model multiple quantiles simultaneously with non-crossing constraints to ensure valid trend estimates.
+\item To reduce computation time and extend the method to long time series, we develop a parallelizable ADMM algorithm. 
+\item Finally, we propose a modified criterion for performing model selection.
+\end{itemize}
+
+	In the rest of the paper, we detail our quantile trend filtering algorithms as well as how to choose the smoother parameter $\lambda$ (Section \ref{sec:methods}).
+We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods (Section \ref{sec:simluation}).
+We further show that quantile trend filtering is a more effective method of drift removal for low-cost air quality sensors and results in improved signal classification compared to quantile smoothing splines (Section \ref{sec:application}).
+Finally, we discuss potential extensions of quantile trend filtering (Section \ref{sec:discussion}).
+
+%% ----------------------------------------------------------------------
+%% Methods
+%% ----------------------------------------------------------------------
+\section{Baseline Trend Estimation}
+\label{sec:methods}
+
+%We review \textit{$\ell_1$-trend filtering} for estimating a mean trend before discussing how to adapt it to estimate a baseline trend.
+We then describe a new scalable algorithm for simultaneously computing multiple quantile trends over large times windows.
+
+%% ----------------------------------------------------------------------
+%% Background
+%% ----------------------------------------------------------------------
+\subsection{Background}
+
 \cite{Kim2009} originally proposed \textit{$\ell_1$-trend filtering} to estimate trends with piecewise polynomial functions, assuming that the observed time series $\V{y}$ consists of a trend $\V{\theta}$ plus uncorrelated noise $\V{\varepsilon}$, namely $\V{y} = \V{\theta} + \V{\varepsilon}$.
@@
 	where $\lambda$ is a nonnegative regularization parameter, and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
-To understand the purpose of penalizing $\Mn{D}{k+1}$ consider the difference operator when $k = 0$.
+To understand the purpose of penalizing the size of $\Mn{D}{k+1}\V{\theta}$ consider the difference operator when $k = 0$.
+
 	\begin{eqnarray*}
 	\Mn{D}{1} & = & \begin{pmatrix}
-
 	-1 & 1 & 0 & \cdots & 0 & 0 \\
 	0 & -1 & 1 & \cdots & 0 & 0 \\
-	\vdots & & & & & \\
+	\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
 	0 & 0 & 0 & \cdots & -1 & 1 \\
-
 	\end{pmatrix}.
 	\end{eqnarray*}
-	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
+
+	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$, which is known as total variation denoising penalty in one dimension in the signal processing literature \citep{Rudin1992} or the fused lasso penalty in the statistics literature \citep{Tibshirani2005}.
 The penalty term incentivizes solutions which are piecewise constant.
@@
 As noted earlier though in some applications, such as the air quality monitoring problem considered in this paper, the data contain a rapidly varying signal in addition to the slowly varying trend and noise.
-Standard trend filtering struggles to distinguish between the slowly varying trend and the rapidly-varying signal, illustrated in \Fig{tpspn}.
+\Fig{tpspn} shows that standard trend filtering is not designed to distinguish between the slowly varying trend and the rapidly-varying signal, as the smooth component estimate $\V{\theta}$ is biased towards the peak of the transient component.
 
-	To account for the presence of a transient component in the observed time series $\V{y}$, we propose quantile trend filtering \Fig{tpspn}.% as a more robust method for detrending in these situations \Fig{tpspn}.
+	To account for the presence of a transient component in the observed time series $\V{y}$, we propose quantile trend filtering \Fig{tpspn}.
+% as a more robust method for detrending in these situations \Fig{tpspn}.
 	To estimate the trend in the $\tau$\textsuperscript{th} quantile, we solve the convex optimization problem
@@
 	and $\One(A)$ is 1 if its input $A$ is true and 0 otherwise.
-\cite{Kim2009} proposed using the check function as a possible extension of $\ell_1$-trend filtering but did not investigate it further.
-Here, we develop the basic $\ell_1$-quantile-trend-filtering framework and extend it to model multiple quantiles simultaneously with non-crossing constraints to ensure valid trend estimates.
-To reduce computation time and extend the method to larger series, we develop and a parallelizable ADMM algorithm.
-Finally, we propose a modified criterion for choosing the smoothing parameter.
 
@@
 	\end{figure}
+	Note that we do not explicitly model $\V{s}$. Rather, we focus on estimating $\V{\theta}$. We then estimate $\V{s} + \V{\varepsilon}$ as the difference $\V{y} - \V{\theta}$.
 
-	This work was motivated by the analysis of signals measured by low cost and portable air quality sensors, whose use has increased dramatically in the last decade.
-These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
-The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
-Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
-To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
-Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
-All of the sensors respond to the pollutant signal, the three sharp transient spikes at \Halley{Eric}{can you put the times at which the spikes occur here?}, but the amount of baseline drift varies from one sensor to another obscuring the detection of the peaks that herald the presence of pollutants.
-We show later that by estimating the baseline drift in each sensor using $\ell_1$-quantile-trend filtering and removing it from the observed signals, peaks can be reliably detected from concordant residual signals from a collection of SPods using a simple data-driven thresholding strategy.
+	Before elaborating on how we compute our proposed $\ell_1$-quantile trend filtering estimator, we discuss similarities and differences between our proposed estimator and existing baseline estimators.
 
-	\begin{figure}[t!]
-		\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}		
-		\caption{Example of 3 co-located SPod PID sensor readings.}
-		\label{fig:raw_spod}
-	\end{figure}
+%% ----------------------------------------------------------------------
+%% Relationship to Prior Work
+%% ----------------------------------------------------------------------
+\subsubsection{Relationship to Prior Work}
 
@@
 	\label{eq:smoothingspline}
-	\sum_{t=1}^n\rho_{\tau}(y(t) - \theta(t)) + \lambda\int \theta''(t)^2 dt, 
+	\sum_{i=1}^n\rho_{\tau}(y(t_i) - \theta(t_i)) + \lambda\int \theta''(t)^2 dt, 
  	\end{equation*}	
 	where $\theta(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degree of smoothing.
-\cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check function with a similar differentiable function.
+\cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check function with a differentiable function.
 \cite{Racine2017} take a different approach and constrain the response to follow a smooth location scale model of the form $y(t) = a(t) + b(t)\epsilon_i$ and estimate the $\tau$\textsuperscript{th} conditional quantile using a kernel estimator and local linear approach.
+\Halley{Eric}{What are $a(t), b(t)$, and $\epsilon_i$?}
 
-	In the rest of the paper, we detail our quantile trend filtering algorithms as well as how to choose the smoother parameter $\lambda$ (Section \ref{sec:methods}).
-We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods (Section \ref{sec:simluation}).
-We further show that quantile trend filtering is a more effective method of drift removal for low-cost air quality sensors and results in improved signal classification compared to quantile smoothing splines (Section \ref{sec:application}).
-Finally, we discuss potential extensions of quantile trend filtering (Section \ref{sec:discussion}).
-
-%% ----------------------------------------------------------------------
-%% Methods
-%% ----------------------------------------------------------------------
-\section{Methods}
-\label{sec:methods}
 %% ----------------------------------------------------------------------
@@
 	&&\underset{\M{\Theta} \in \mathcal{C}}{\text{minimize}}\; \sum_{j=1}^J \left [\rho_{\tau_j}(\V{y} - \V{\theta}_{j}) + 
-	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
+	\lambda \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
 %	&&\text{subject to: }\; \V{\theta}_1 \leq \V{\theta}_2 \leq \ldots \leq \V{\theta}_J	
@@
 To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
-We propose an ADMM algorithm for solving large problems in a piecewise fashion.
-The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is fully described by \cite{boyd2011distributed}.
+We propose a divide-and-conquer approach via an ADMM algorithm for solving large problems in a piecewise fashion.
+The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is  described in greater detail by \cite{boyd2011distributed}, but we briefly review how it can be used to iteratively solve the following equality constrained optimization problem.
+\begin{equation}
+\label{eq:split_objective}
+\begin{split}
+&\text{minimize} \; f(\V{\theta}) + g(\Vtilde{\theta}) \\
+&\text{subject to} \; \M{A}\V{\theta} + \M{B}\Vtilde{\theta} = \V{c}.
+\end{split}
+\end{equation}
+Recall that finding the minimizer to an equality constrained optimization problem is equivalent to the identifying the saddle point of the Lagrangian function associated with the problem \Eqn{split_objective}.
+ADMM seeks the saddle point of a related function called the augmented Lagrangian,
+%The Lagrangian for the ALM problem
+\begin{eqnarray*}
+\mathcal{L}_{\gamma}(\V{\theta},\Vtilde{\theta},\V{\omega}) & = & f(\V{\theta}) + g(\Vtilde{\theta}) + \langle \V{\omega}, \V{c} - \M{A}\V{\theta} - \M{B}\Vtilde{\theta} \rangle
++ \frac{\gamma}{2} \lVert \V{c} - \M{A}\V{\theta} - \M{B}\Vtilde{\theta} \rVert_2^2,
+\end{eqnarray*}
+where the dual variable $\V{\omega}$ is a vector of Lagrange multipliers and $\gamma$ is a nonnegative tuning parameter.
+When $\gamma = 0$, the augmented Lagrangian coincides with the ordinary Lagrangian.
+
+ADMM minimizes the augmented Lagrangian one block of variables at a time before updating the dual variable $\V{\omega}$. This yields the following sequence of updates at the $m+1$th ADMM iteration
+\begin{equation}
+\label{eq:admm_updates}
+\begin{split}
+\V{\theta}_{m+1} & \amp = \amp \underset{\V{\theta}}{\arg\min}\; \mathcal{L}_\gamma(\V{\theta},\Vtilde{\theta}_{m}, \V{\omega}_{m}) \\
+\Vtilde{\theta}_{m+1} & \amp = \amp \underset{\Vtilde{\theta}}{\arg\min}\; \mathcal{L}_\gamma(\V{\theta}_{m+1},\Vtilde{\theta}, \V{\omega}_{m}) \\
+\V{\omega}_{m+1} & \amp = \amp \V{\omega}_{m} + \gamma(\V{c} - \M{A}\V{\theta}_{m+1} - \M{B}\Vtilde{\theta}_{m+1}).
+\end{split}
+\end{equation}
 
 	Let $\mathcal{I}$ denote a subset of the indices $\{1, \ldots, n\}$ and $\V{y}_\mathcal{I} \in \Real^{\lvert I \rvert}$ denote the subvector obtained by stacking on top of each other the elements $\VE{y}{i}$ for $i \in \mathcal{I}$ in their canonical ordering.
-	We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in \Eqn{quantile_trend}, by dividing our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into overlapping $M$ windows of observations $\Vn{y}{m} = \V{y}_{\mathcal{I}_m}$ where $\mathcal{I}_m$ is the set of integers from $l_m$ to $u_m$ inclusive, namely $\mathcal{I}_m = \{l_m, l_m + 1, \ldots, u_m -1, u_m\}$ where $1 = l_{1} < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots < u_{M} = n$.
+	We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in \Eqn{quantile_trend}, by dividing our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into overlapping $W$ windows of observations $\Vn{y}{w} = \V{y}_{\mathcal{I}_w}$ where $\mathcal{I}_w$ is the set of integers from $l_w$ to $u_w$ inclusive, namely $\mathcal{I}_w = \{l_w, l_w + 1, \ldots, u_w -1, u_w\}$ where
+\begin{eqnarray*}
+1 = l_{1} < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots < u_{W} = n.
+\end{eqnarray*}
 \Fig{windows} shows an example of 1200 observations being mapped into three equally sized overlapping windows of observations.
@@
 In order to constrain the overlapping sections to be equal, we define a consensus variable
-	 Given quantiles $\tau_1 < \cdots < \tau_J$, we introduce dummy variables $\Vn{\theta}{m}_{j} \in \Real^{\lvert \mathcal{I}_m \rvert}$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$.
-We then ``stitch" together the $M$ quantile trend estimates into consensus over the overlapping regions by introducing the constraint $\VnE{\theta}{m}{ij} = \VE{\theta}{i + l_m - 1,j}$ for $i = 1, \ldots, u_m - l_m + 1$ and for all $j$.
-Let $\Mn{\Theta}{m}$ be the matrix whose $j$th column is $\Vn{\theta}{m}$.
-We can write these constraints more compactly as $\Mn{\Theta}{m} = \Mn{U}{m}\M{\Theta}$, where $\Mn{U}{m} \in \{0, 1\}^{\lvert \mathcal{I}_m \rvert \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $m$th window.
+	 Given quantiles $\tau_1 < \cdots < \tau_J$, we introduce dummy variables $\Vn{\theta}{w}_{j} \in \Real^{\lvert \mathcal{I}_w \rvert}$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $w$.
+We then ``stitch" together the $W$ quantile trend estimates into consensus over the overlapping regions by introducing the constraint $\VnE{\theta}{w}{ij} = \VE{\theta}{i + l_w - 1,j}$ for $i = 1, \ldots, u_w - l_w + 1$ and for all $j$.
+Let $\Mn{\Theta}{w}$ be the matrix whose $j$th column is $\Vn{\theta}{w}$.
+Then we can write these constraints more compactly as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{\lvert \mathcal{I}_w \rvert \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $m$th window, namely
+\begin{eqnarray*}
+\Mn{U}{w} & = & \begin{pmatrix}
+\V{e}_{l_w}\Tra \\
+\vdots \\
+\V{e}_{u_w}\Tra
+\end{pmatrix},
+\end{eqnarray*}
+where $\V{e}_i \in \Real^{n}$ denotes the $i$th standard basis vector.
 Furthermore, let $\iota_{\mathcal{C}}$ denote the indicator function of the non-crossing quantile constraint, namely $\iota_{\mathcal{C}}(\M{\Theta})$ is zero if $\M{\Theta} \in \mathcal{C}$ and infinity otherwise.
@@
 	Our windowed quantile trend optimization problem can then be written as 
-	 \begin{align*}
+	 \begin{equation}
 		 \label{eq:quantile_windows}
-		 &\text{minimize}\; \sum_{m=1}^M \left \{\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{m} - \Vn{\theta}{m}_{j}) + 
-		 \lambda_j \lVert \Mn{D}{k+1} \Vn{\theta}{m}_{j} \rVert_1 \right ] + \iota_\mathcal{C}(\Mn{\Theta}{m}) \right\}\\
-		 &\text{subject to} \qquad \Mn{\Theta}{m} = \Mn{U}{m}\M{\Theta} \;\; \text{ for } m = 1, \ldots, M.
-	 \end{align*}
+		 \begin{split}
+		 &\text{minimize}\; \sum_{w=1}^W \left \{\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{w} - \Vn{\theta}{w}_{j}) + 
+		 \lambda \lVert \Mn{D}{k+1} \Vn{\theta}{w}_{j} \rVert_1 \right ] + \iota_\mathcal{C}(\Mn{\Theta}{w}) \right\}\\
+		 &\text{subject to} \qquad \Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta} \;\; \text{ for } w = 1, \ldots, W.
+		 \end{split}
+	 \end{equation}
 
-	 	 Let $\Mn{\Omega}{m}$ denote the Lagrange multiplier matrix for the $m$th consensus constraint, namely $\Mn{\theta}{m} = \Mn{U}{m}\M{\theta}$, and let $\Vn{\omega}{m}_{j}$ denote its $j$th column.
+	 	 Let $\Mn{\Omega}{w}$ denote the Lagrange multiplier matrix for the $m$th consensus constraint, namely $\Mn{\theta}{w} = \Mn{U}{w}\M{\theta}$, and let $\Vn{\omega}{w}_{j}$ denote its $j$th column.
 %		 
-%		  $\Vn{\omega}{m}_{j}$ denote Lagrange multipliers for the constraint $\Vn{\theta}{m}_j = \Mn{U}{m}\V{\theta}_j$, $\Mn{\Omega}{m}$ denote the matrix whose $j$th column is $\Vn{\omega}{m}_j$, and penalty parameter $\gamma > 0$,
+%		  $\V{\omega}_{m}_{j}$ denote Lagrange multipliers for the constraint $\Vn{\theta}{w}_j = \Mn{U}{m}\V{\theta}_j$, $\Mn{\Omega}{m}$ denote the matrix whose $j$th column is $\V{\omega}_{m}_j$, and penalty parameter $\gamma > 0$,
 
@@
 	 \begin{eqnarray*}
-	 \mathcal{L}(\M{\Theta}, \{\Mn{\Theta}{m}\}, \{\Mn{\Omega}{m}\}) & = & \sum_{m=1}^M \biggl \{\sum_{j=1}^J \biggl [\rho_{\tau_j}(\Vn{y}{m} - \Vn{\theta}{m}_j)+\lambda \lVert \Mn{D}{k+1}\Vn{\theta}{m}_j\rVert_1 \\
-	 && + \; (\Vn{\theta}{m}_j - \Mn{U}{m}\V{\theta}_{j})\Tra\Vn{\omega}{m}_{j} + 
-	 \frac{\gamma}{2}\lVert \Vn{\theta}{m}_j - \Mn{U}{m}\V{\theta}_{j}\rVert_2^2 \biggr ] + \iota_\mathcal{C}(\Mn{\Theta}{m}) \biggr\},
+	 \mathcal{L}(\M{\Theta}, \{\Mn{\Theta}{w}\}, \{\Mn{\Omega}{w}\}) & = & \sum_{w=1}^W \mathcal{L}_w(\M{\Theta}, \Mn{\Theta}{w}, \Mn{\Omega}{w}),
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+ \mathcal{L}_w(\M{\Theta}, \Mn{\Theta}{w}, \Mn{\Omega}{w}) & = & \sum_{j=1}^J \biggl [\rho_{\tau_j}(\Vn{y}{w} - \Vn{\theta}{w}_j)+\lambda \lVert \Mn{D}{k+1}\Vn{\theta}{w}_j\rVert_1 \\
+	 && + \; (\Vn{\theta}{w}_j - \Mn{U}{w}\V{\theta}_{j})\Tra\Vn{\omega}{w}_{j} + 
+	 \frac{\gamma}{2}\lVert \Vn{\theta}{w}_j - \Mn{U}{w}\V{\theta}_{j}\rVert_2^2 \biggr ] + \iota_\mathcal{C}(\Mn{\Theta}{w}),
 	 \end{eqnarray*}
 	 where $\gamma$ is a positive penalty parameter.
-	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in Algorithm 1.
+
+The ADMM algorithm alternates between updating the consensus variable $\M{\Theta}$, the window variables $\{\Mn{\Theta}{w}\}$, and the Lagrange multipliers $\{\Mn{\Omega}{w}\}$. 
+At the $m+1$th iteration, we perform the following sequence of updates
+
+\begin{eqnarray*}
+	\M{\Theta}_{m+1} & = & \underset{\M{\Theta}}{\arg\min}\; \mathcal{L}(\M{\Theta}, \{\Mn{\Theta}{w}_m\}, \{\Mn{\Omega}{w}_m\}) \\
+	\Mn{\Theta}{w}_{m+1} & = & \underset{\{\Mn{\Theta}{w}\}}{\arg\min}\; \mathcal{L}(\M{\Theta}_{m+1}, \{\Mn{\Theta}{w}\}, \{\Mn{\Omega}{w}_m\}) \\
+\end{eqnarray*}
+
+{\bf Updating $\M{\Theta}$: } Some algebra shows that updating the consensus variable step is computed as follows.
+
+\begin{eqnarray}
+\label{eq:update_consensus}
+\ME{\theta}{ij}	 &=& \begin{cases} 
+			 \frac{1}{2}\left (\MnE{\theta}{w-1}{ij} - \gamma\Inv \MnE{\omega}{w-1}{ij} \right)
+			 + 
+			 \frac{1}{2} \left (\MnE{\theta}{w}{ij} - \gamma\Inv \MnE{\omega}{w}{ij} \right)
+			  & \mbox{if~~} l_{w} \le i \le u_{w-1},  \\
+			 \MnE{\theta}{w}{ij} & \mbox{if~~} u_{w-1} < i \le l_{w+1}  \\
+			 \frac{1}{2} \left ( \MnE{\theta}{w}{ij} - \gamma\Inv \MnE{\omega}{w}{ij}\right )
+			 + 
+			 \frac{1}{2}	\left ( \MnE{\theta}{w+1}{ij} - \gamma\Inv \MnE{\omega}{w+1}{ij} \right )
+			 & \mbox{if~~} l_{m+1} < i \le u_{m}  
+			 \end{cases},
+\end{eqnarray}
+
+The consensus update \Eqn{update_consensus} is rather intuitive.
+We essentially average the trend estimates in overlapping sections of the windows, subject to some adjustment by the Lagrange multipliers, and leave the trend estimates in non-overlapping sections of the windows untouched.
+%This appears like we need to invert an underdetermined linear system, but we're actually just picking off the relevant entries.
+For notational ease, we write the consensus update \Eqn{update_consensus} compactly as $\M{\Theta} = \psi(\{\Mn{\Theta}{w}\},\{\Mn{\Omega}{w}\})$.
+
+{\bf Updating $\{\Mn{\Theta}{w}\}$: } We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in \Alg{admm}.
 The use of the Augmented Lagrangian converts the problem from a linear program into a quadratic program.
@@
 %	 \begin{eqnarray*}
-%	\bar{\theta}_{j, m}	&=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
-%	\bar{\theta}_{j, m}(t)	 &=& \begin{cases} 
+%	\M{A}r{\theta}_{j, m}	&=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
+%	\M{A}r{\theta}_{j, m}(t)	 &=& \begin{cases} 
 %			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
@@
 %	 \begin{align*}
-%	\bar{\theta}_{j, m}	=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
-%	\bar{\theta}_{j, m}(t)	 =& \begin{cases} 
+%	\M{A}r{\theta}_{j, m}	=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
+%	\M{A}r{\theta}_{j, m}(t)	 =& \begin{cases} 
 %			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
@@
 %		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
-%		 &\qquad \theta_{j,m}(t) = \bar{\theta}_{j,m}(t) \text{ for all } j, m, t
+%		 &\qquad \theta_{j,m}(t) = \M{A}r{\theta}_{j,m}(t) \text{ for all } j, m, t
 %	 \end{align*}
@@
 %	 \begin{equation*}
-%	 \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \bar{\theta}_{j,m}) +
-%	 \frac{\gamma}{2}||\theta_{j,m} - \bar{\theta}_{j,m}||_2^2
+%	 \mathcal{L}(\theta_{j,m}, \M{A}r{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \M{A}r{\theta}_{j,m}) +
+%	 \frac{\gamma}{2}||\theta_{j,m} - \M{A}r{\theta}_{j,m}||_2^2
 %	 \end{equation*}
@@
 %	 
-	\begin{figure}[!h] 
+	\begin{figure}[!t] 
 		\centering
@@
 	\begin{algorithm}
-		\caption{ADMM algorithm for quantile trend filtering with windows}\label{euclid}
+		\caption{ADMM algorithm for quantile trend filtering with windows}\label{alg:admm}
 		\begin{algorithmic}		
 			\State Define $\M{D} = \Mn{D}{k+1}$. 
-			\State \textbf{initialize:} \\
-				 $\Vn{\theta}{0}_{j,m} \gets \arg \min \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{m} - \V{\theta}_{j,m})+\lambda \lVert \M{D}\V{\theta}_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
+			\State \textbf{initialize:}
+			\For{$w = 1, \ldots, W$}
+			\State $\Mn{\Theta}{w}_0 \gets \underset{\Mn{\Theta}{w} \in \mathcal{C}}{\arg \min}\; 			
+			 \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{w} - \Vn{\theta}{w}_{j})+\lambda \lVert \M{D}\Vn{\theta}{w}_{j}\rVert_1$
+%				 $\Vn{\theta}{0}_{j,m} \gets \arg \min \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{w} - \V{\theta}_{j,m})+\lambda \lVert \M{D}\V{\theta}_{j,m}\rVert_1$ subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
 \\
-			 	 $\Vn{\omega}{0}_{j,m} \gets \V{0}$	
+			\State $\Mn{\Omega}{w}_0 \gets \M{0}$
+			\EndFor
+%			 	 $\Vn{\omega}{0}_{j,m} \gets \V{0}$	
+			\State $m \gets 0$
 			\Repeat{}
+			\State $\M{\Theta}_{m+1} \gets \psi(\{\Mn{\Theta}{w}_m\}, \{\Mn{\Omega}{w}_m\})$
+			\For{$w = 1, \ldots, W$}			
+			\State $\Mn{\Theta}{w}_{m+1} \gets \underset{\Mn{\Theta}{w}}{\arg\min}\; \mathcal{L}_w(\M{\Theta}_{m+1}, \Mn{\Theta}{w}, \Mn{\Omega}{w}_{m})$
 			\State 
-			$\bar{\theta}_{j,m}^{(q)} \gets g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
-			\State 
-			$\omega_{j,m}^{(q)} \gets \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \bar{\theta}_{j,m}^{(q)})$	
-			\State
-				$\theta_{j,m}^{(q)} \gets \arg\min \mathcal{L}(\theta_{j,m}, \bar{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
-			 subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
+			$\Mn{\Omega}{w}_{m+1} \gets \Mn{\Omega}{w}_m + \gamma(\Mn{\Theta}{w}_{m+1} - \Mn{U}{w}\M{\Theta}_{m+1})$			
+			\EndFor
+			\State $m \gets m + 1$
+%			\State  					
+%			$\M{A}r{\theta}_{j,m}^{(q)} \gets g(\theta_{j, m-1}^{(q-1)}, \theta_{j,m}^{(q-1)}, \theta_{j,m+1}^{(q-1)})$
+%			\State 
+%			$\omega_{j,m}^{(q)} \gets \omega_{j,m}^{(q-1)} + \gamma(\theta_{j,m}^{(q-1)} - \M{A}r{\theta}_{j,m}^{(q)})$			
+%			\State
+%				$\theta_{j,m}^{(q)} \gets \arg\min \mathcal{L}(\theta_{j,m}, \M{A}r{\theta}_{j,m}^{(q-1)}, \omega_{j,m}^{(q-1)})$			
+%			 subject to $\theta_{1,m}(t) < \ldots<\theta_{J,m}(t)$ for all $t$.
 			\Until {convergence}
-			\State \textbf{return} Non-overlapping sequence of $\bar{\theta}_{j,m}^{(q)}$ for all $j$, $m$.
-
+			\State \textbf{return} $\M{\Theta}_m$
+%			\State \textbf{return} Non-overlapping sequence of $\M{A}r{\theta}_{j,m}^{(q)}$ for all $j$, $m$.			
 		\end{algorithmic}
@@
 
-	We measure convergence using the stopping criteria described by \cite{boyd2011distributed}.
+	To terminate our algorithm, we use the stopping criteria described by \cite{boyd2011distributed}.
 The criteria are based on the primal and dual residuals, which represent the residuals for primal and dual feasibility, respectively.
-The primal residual,
+The primal residual at the $m$th iteration,
 %	\begin{eqnarray*}
-%	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q)}\rVert_2^2},
+%	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \M{A}r{\theta}_{j,m}^{(q)}\rVert_2^2},
 %	\end{eqnarray*}
 	\begin{eqnarray*}
-	r_{\text{primal}}^{k} & = & \sqrt{\sum_{m=1}^M\lVert\Mn{\Theta}{m}_{k} - \M{\Theta}_k\rVert_{\text{F}}^2},
+	r_{\text{primal}}^{m} & = & \sqrt{\sum_{w=1}^W\lVert\Mn{\Theta}{w}_{m} - \M{\Theta}_m\rVert_{\text{F}}^2},
 	\end{eqnarray*}
-	represents the difference between the trend values in the windows and the consensus trend value. The dual residual,
+	represents the difference between the trend values in the windows and the consensus trend value. The dual residual at the $m$th iteration,
 %	\begin{eqnarray*}
-%	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\bar{\theta}_{j,m}^{(q)} - \bar{\theta}_{j,m}^{(q-1)}\rVert_2^2},
+%	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\M{A}r{\theta}_{j,m}^{(q)} - \M{A}r{\theta}_{j,m}^{(q-1)}\rVert_2^2},
 %	\end{eqnarray*}
 	\begin{eqnarray*}
-	r_{\text{dual}}^{k} & = & \gamma\sqrt{\sum_{m=1}^M \lVert\M{\Theta}_k - \M{\Theta}_{k-1}\rVert_{\text{F}}^2},
+	r_{\text{dual}}^{m} & = & \gamma\sqrt{\sum_{w=1}^W \lVert\M{\Theta}_m - \M{\Theta}_{m-1}\rVert_{\text{F}}^2},
 	\end{eqnarray*}
@@
 %	\begin{eqnarray*}
-%		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{m}{\max}\left[\max 
-%		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \bar{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
+%		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{w}{\max}\left[\max 
+%		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \M{A}r{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
 %		r_d^{(q)} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
@@
 	\begin{eqnarray*}
-		r_{\text{primal}}^{k} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\underset{m}{\max} \left[\max 
-		\left\{\lVert\Mn{\Theta}{m}_k \rVert_{\text{F}}, \lVert \M{\Theta}_{k} \rVert_{\text{F}} \right\}\right] \\%\lVert \M{\Theta}_{k} \rVert_{\text{F}}} \right )\right]\\
-		r_{\text{dual}}^{k} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\sqrt{\sum_{m=1}^M
-		\lVert \Mn{\Omega}{m}_k \rVert_{\text{F}}^2}.
+		r_{\text{primal}}^{m} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\underset{w}{\max} \left[\max 
+		\left\{\lVert\Mn{\Theta}{w}_m \rVert_{\text{F}}, \lVert \M{\Theta}_{m} \rVert_{\text{F}} \right\}\right] \\%\lVert \M{\Theta}_{k} \rVert_{\text{F}}} \right )\right]\\
+		r_{\text{dual}}^{m} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\sqrt{\sum_{w=1}^W
+		\lVert \Mn{\Omega}{w}_m \rVert_{\text{F}}^2}.
 	\end{eqnarray*}
@@
 
-	\begin{figure}[!h] 
+	\begin{figure}[!t] 
 		\centering
@@
 
-	\subsection{Regularization Parameter Choice}
+%% ----------------------------------------------------------------------
+%% Computational Complexity and Convergence
+%% ----------------------------------------------------------------------
+\subsection{Computational Complexity and Convergence}
+
+\Eric{Eric}{Add computational complexity of quadratic program and discuss the savings compared with solving a single large LP}.
+
+\Alg{admm} has the following convergence guarantees.
+
+\begin{proposition}
+\label{prop:convergence}
+Let $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$ denote the $m$th collection of iterates generated by \Alg{admm}. Then (i) 
+$\lVert \Mn{\Theta}{w}_m - \Mn{U}{w}\M{\Theta}_m \rVert_{\text{F}} \rightarrow 0$ and (ii) $p_m \rightarrow p^\star$, where $p^\star$ is the optimal objective function value of 
+\Eqn{quantile_windows} and $p_m$ is the objective function value of \Eqn{quantile_windows} evaluated at $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$.
+\end{proposition}
+
+The proof of \Prop{convergence} is a straightforward application of the convergence result presented in Section 3.2 of \cite{boyd2011distributed}.
+
+%% ----------------------------------------------------------------------
+%% Model Selection
+%% ----------------------------------------------------------------------
+	\section{Model Selection}
 	\label{sec:lambda_choice}
 
-	An important practical issue in trend estimation is the choice of regularization parameter, or degree of smoothness.
-Our method can easily handle missing data by defining the check loss function to output 0 for missing values, allowing us to implement cross-validation.
-We also explore three information criteria.
+	An important practical issue in baseline estimation is the choice of regularization parameter $\lambda$, which controls the degree of smoothness in $\V{\theta}$.
+In this section, we introduce four methods for choosing $\lambda$.
+The first is validation based approach; the latter three are based on information criteria.
+
+\subsection{Validation}	
+	Our method can easily handle missing data by defining the check loss function to output 0 for missing values. %We also explore three information criteria.  
+	Specifically, we use the following modified function $\tilde{\rho}_\tau$ in place of the $\rho_\tau$ function given in \Eqn{check}
+	\begin{eqnarray}
+	\label{eq:modcheck}
+	\tilde{\rho}_{\tau}(\V{r}) & = & \sum_{i \not\in \mathcal{V}} \VE{r}{i}(1-\One(\VE{r}{i}<0)),
+	\end{eqnarray}
+	where $\mathcal{V}$ is a held-out validation subset of $\{1, \ldots, n\}$ and solve the problem
+
+	\begin{eqnarray}
+	\label{eq:validate}
+	&&\underset{\M{\Theta} \in \mathcal{C}}{\text{minimize}}\; \sum_{j=1}^J \left [\tilde{\rho}_{\tau_j}(\V{y} - \V{\theta}_{j}) + 
+	\lambda \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ],
+	\end{eqnarray}	
+	which can be solved via \Alg{admm} with trivial modification to the quadratic program subproblems.
+We then select the $\lambda$ that minimizes the hold-out prediction error quantified by $\breve{\rho}(\V{y} - \Vhat{\theta}(\lambda))$ where $\Vhat{\theta}(\lambda)$ is the solution to \Eqn{validate} and
+	\begin{eqnarray*}
+	\breve{\rho}_{\tau}(\V{r}) & = & \sum_{i \in \mathcal{V}} \VE{r}{i}(1-\One(\VE{r}{i}<0)).
+	\end{eqnarray*}
+
+\subsection{Information Criteria}
 \cite{KoenkerNgPortnoy1994} addressed the choice of regularization parameter by proposing the Schwarz criterion for the selection of $\lambda$
@@
 	\label{eq:SIC}
-	\mbox{SIC}(p_{\lambda}) & = & \log\left[\frac{1}{n}\rho_{\tau}(y - \theta)\right] + \frac{1}{2n}p_{\lambda}\log n.
+	\mbox{SIC}(p_{\lambda}) & = & \log\left[\frac{1}{n}\rho_{\tau}(\V{y} - \V{\theta})\right] + \frac{1}{2n}p_{\lambda}\log n.
 	\end{eqnarray*}
@@
 	\begin{eqnarray*}
-	L(y \mid \theta) & = & \left[\frac{\tau^n(1-\tau)}{\sigma}\right]^n\exp\left\{-\sum_i\rho_\tau\left(\frac{\VE{y}{i} - \VE{\theta}{i}}{\sigma}\right)\right\},
+	L(\V{y} \mid \V{\theta}) & = & \left[\frac{\tau^n(1-\tau)}{\sigma}\right]^n\exp\left\{-\sum_i\rho_\tau\left(\frac{\VE{y}{i} - \VE{\theta}{i}}{\sigma}\right)\right\},
 	\end{eqnarray*} 
@@
 	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given model.
-We used this criteria with $\gamma = 1$, and $P=n-k-1$.
+\Halley{Eric}{What is $s$?} We used this criteria with $\gamma = 1$, and $P=n-k-1$.
 \Fig{BIC} illustrates the difference among the scaled, unscaled ($\sigma = 1$), and scaled extended BIC criteria
@@
 
+%% ----------------------------------------------------------------------
+%% Estimating Quantiles
+%% ----------------------------------------------------------------------	 
 \subsection{Estimating Quantiles}
@@
 Only our detrend methods guarantee non-crossing quantiles.
-For each quantile trend and method the root mean squared error was calculated as $\mbox{RMSE} = \sqrt{\frac{1}{n}\sum_t (\hat{q}_{\tau}(t) - q_\tau(t))^2}$.
+For each quantile trend and method the root mean squared error was calculated as $\mbox{RMSE} = \sqrt{\frac{1}{n}\sum_i (\hat{q}_{\tau}(t_i) - q_\tau(t_i))^2}$.
 \Fig{quantile_mse} shows the mean RMSE $\pm$ twice the standard error for each method, quantile level, and sample size.
@@
 
+%% ----------------------------------------------------------------------
+%% Peak Detection
+%% ----------------------------------------------------------------------
 \subsection{Peak Detection}

### 2019-03-11 Manuscript/detrendify_JCGS.tex
small text edits to manuscript

@@
  and  removing it from the observed signals, peaks can be reliably detected from concordant residual signals from a collection of SPods using a simple data-driven thresholding strategy.
-Thus, accurately demix a noisy observed vector $\V{y}$ into a slowly varying component $\V{\theta}$ and a transient component $\V{s}$ can lead to greatly improved and simplified downstream analysis.
+Thus, accurately demixing a noisy observed vector $\V{y}$ into a slowly varying component $\V{\theta}$ and a transient component $\V{s}$ can lead to greatly improved and simplified downstream analysis.
 This is true for not just air quality applications, but universally across the many domains listed earlier.
@@
 Let $\Mn{\Theta}{w}$ be the matrix whose $j$th column is $\Vn{\theta}{w}$.
-Then we can write these constraints more compactly as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{\lvert \mathcal{I}_w \rvert \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $m$th window, namely
+Then we can write these constraints more compactly as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{\lvert \mathcal{I}_w \rvert \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $w$th window, namely
 \begin{eqnarray*}
@@
 
-	 	 Let $\Mn{\Omega}{w}$ denote the Lagrange multiplier matrix for the $m$th consensus constraint, namely $\Mn{\theta}{w} = \Mn{U}{w}\M{\theta}$, and let $\Vn{\omega}{w}_{j}$ denote its $j$th column.
+	 	 Let $\Mn{\Omega}{w}$ denote the Lagrange multiplier matrix for the $w$th consensus constraint, namely $\Mn{\theta}{w} = \Mn{U}{w}\M{\theta}$, and let $\Vn{\omega}{w}_{j}$ denote its $j$th column.
 %		 
@@
 In this section, we introduce four methods for choosing $\lambda$.
-The first is validation based approach; the latter three are based on information criteria.
+The first is a validation based approach; the latter three are based on information criteria.
 

### 2019-03-16 Manuscript/detrendify_JCGS.tex
joe's edits 3/16

@@
 	\begin{abstract}
-		We address the problem of estimating smoothly varying baseline trends in time series data using quantile trend filtering.
-We first extend the basic framework to ensure non-crossing while estimating multiple quantile trends simultaneously.
-We also implement a parallelizable alternating direction method of moments (ADMM) algorithm for estimating trends in longer series.
+		We address the problem of estimating smoothly varying baseline trends in time series data.
+This problem arises in several applications, including chemistry, macroeconomics, and medicine; however, our study is motivated by the analysis of data from low cost air quality sensors.
+Our methods extend the quantile trend filtering framework to allow the estimation of multiple quantile trends simultaneously while ensuring that the quantiles do not cross.
+To handle the computational challenge posed by very long time series, we propose a parallelizable alternating direction method of moments (ADMM) algorithm.
 The ADMM algorthim enables the estimation of trends in a piecewise manner, both reducing the computation time and extending the limits of the method to larger data sizes.
 We also address smoothing parameter selection and propose a modified criterion based on the extended Bayesian Information Criterion.
-Through simulation studies and an application to low cost air quality sensor data, we demonstrate that our model provides better quantile trend estimates than existing methods and improves signal classification of low-cost air quality sensor output.
+Through simulation studies and our motivating application to low cost air quality sensor data, we demonstrate that our model provides better quantile trend estimates than existing methods and improves signal classification of low-cost air quality sensor output.
 	\end{abstract}
@@
 
-In some applications, the slowly varying component $\V{\theta}$ is the signal of interest and the transient component $\V{s}$ is a vector of nuisance parameters.
-In other applications, the roles of $\V{\theta}$ and $\V{s}$ are reversed.
-The latter applies in the air quality applications motivating this work, as $\V{s}$ represents the signal of interest and $\V{\theta}$ represents a baseline drift that obscures the identification of the important transient events encoded in $\V{s}$.
+In some applications, the slowly varying component $\V{\theta}$ is the signal of interest, and the transient component $\V{s}$ is a vector of nuisance parameters.
+In our air quality application, the roles of $\V{\theta}$ and $\V{s}$ are reversed; $\V{s}$ represents the signal of interest and $\V{\theta}$ represents a baseline drift that obscures the identification of the important transient events encoded in $\V{s}$.
 We describe this motivating problem further to illustrate the critical importance of being able to separate $\V{s}$ from $\V{\theta}$.
@@
 Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
-All of the sensors respond to the pollutant signal, the three sharp transient spikes at \Halley{Eric}{can you put the times at which the spikes occur here?}, but the baseline drift varies from one sensor to another obscuring the detection of the peaks that alert the intrusion of pollutants.
+All of the sensors respond to the pollutant signal, the three sharp transient spikes at \Halley{Eric}{can you put the times at which the spikes occur here?}, but the baseline drift varies from one sensor to another, obscuring the detection of the peaks that alert the intrusion of pollutants.
 We show later that by estimating the baseline drift in each sensor %using $\ell_1$-quantile-trend filtering
@@
 Thus, accurately demixing a noisy observed vector $\V{y}$ into a slowly varying component $\V{\theta}$ and a transient component $\V{s}$ can lead to greatly improved and simplified downstream analysis.
-This is true for not just air quality applications, but universally across the many domains listed earlier.
+\Halley{Joe}{I would cut the next sentence, as we haven't shown this for other applications.} This is true for not just air quality applications, but universally across the many domains listed earlier.
 
@@
 \item To reduce computation time and extend the method to long time series, we develop a parallelizable ADMM algorithm.
+\Halley{Joe}{Let's say a bit more about what this does.
+something like ``...
+a parallelizable ADMM algorithm that splits the time domain into overlapping windows and reconciles estimates from the overlapped portions.''}
 \item Finally, we propose a modified criterion for performing model selection.
@@
 
-	In the rest of the paper, we detail our quantile trend filtering algorithms as well as how to choose the smoother parameter $\lambda$ (Section \ref{sec:methods}).
+	In the rest of the paper, we detail our quantile trend filtering algorithms as well as how to choose the smoothing parameter (Section \ref{sec:methods}).
 We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods (Section \ref{sec:simluation}).
@@
 \cite{Koenker1978} were the first to propose substituting the sum-of-squares term with the check function  \Eqn{check} to estimate a conditional quantile instead of the conditional mean.
-Later, \cite{KoenkerNgPortnoy1994} proposed quantile trend filtering with $k = 2$ producing quantile trends that are piecewise linear, but they did not consider extensions to higher order differences.
+Later, \cite{KoenkerNgPortnoy1994} proposed quantile trend filtering with $k = 2$ producing quantile trends that are piecewise linear \Halley{Joe}{Should this say $k=1$?
+If $k=0$ produces piecewise constant trends, wouldn't $k=1$ produce piecewise linear trends?}, but they did not consider extensions to higher order differences.
 Rather than using the $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the square of the $\ell_2$-norm:
@@
 \end{equation}
-Recall that finding the minimizer to an equality constrained optimization problem is equivalent to the identifying the saddle point of the Lagrangian function associated with the problem \Eqn{split_objective}.
+\Eric{Joe}{I'm already lost here.
+Can we give some context to what $\theta$ and $\widetilde{\theta}$ are in the context of our windowing algorithm?
+Maybe we should say something about the windowing first before we dive head first into Lagrangians?
+What are $A$ and $B$?
+Do these simplify to something in our specific application?
+The stuff below might seem obvious to you, but I'm totally lost.
+It might be better to talk about the notation for the windows first, and how different windows will give different thetas for the overlapping parts, and then talk about equality constrained optimization as a solution for the overlapping parts.} Recall that finding the minimizer to an equality constrained optimization problem is equivalent to the identifying the saddle point of the Lagrangian function associated with the problem \Eqn{split_objective}.
 ADMM seeks the saddle point of a related function called the augmented Lagrangian,
@@
 
-	We also compare the performance of our quantile trend filtering method with three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011, nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
+	We also compare the performance of our quantile trend filtering method with three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011} and \cite{nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
 The regularization parameter $\lambda$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria as described in \cite{KoenkerNgPortnoy1994}, the regularization parameter for \texttt{qsreg} was chosen using generalized cross-validation based on the quantile criterion \cite{Oh2011}.
@@
 \end{eqnarray*}
-	The three error distributions considered were 
+	The errors were independent and had distributions
 	\begin{itemize}
@@
 %	\end{figure}
-	We use another simulation design based on our motivating application. We assume that the measured data can be represented by 
+	The second simulation design is closely motivated by our air quality analysis problem. We assume that the measured data can be represented by
 	\begin{eqnarray*}
@@
 One hundred datasets were generated for each $n=\{500, 1000, 2000, 4000\}$.
-We compare the ability of the methods to estimate the true quantiles of $y(t_i)-s(t_i)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}).
+\Halley{Joe}{This is a lot to take in.
+Can you refer to a figure that shows a simulated dataset?} We compare the ability of the methods to estimate the true quantiles of $y(t_i)-s(t_i)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}).
 In this simulation study, our proposed method \texttt{detrend\_eBIC} method substantially outperforms the others.
@@
 %% ----------------------------------------------------------------------
-\section{Application in Air Quality Monitoring}
+\section{Analysis of Air Quality Data}
 

### 2019-03-18 Manuscript/detrendify_JCGS.tex
added simulation without non-crossing

--- 2019-03-11 Manuscript/detrendify_JCGS.tex
@@
 Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
-All of the sensors respond to the pollutant signal, the three sharp transient spikes at \Halley{Eric}{can you put the times at which the spikes occur here?}, but the baseline drift varies from one sensor to another obscuring the detection of the peaks that alert the intrusion of pollutants.
-We show later that by estimating the baseline drift in each sensor %using $\ell_1$-quantile-trend filtering
+All of the sensors respond to the pollutant signal, the three sharp transient spikes at 11:32, 14:10, and 16:10, but the baseline drift varies from one sensor to another obscuring the detection of the peaks that alert the intrusion of pollutants.
+We show later that by estimating the baseline drift in each sensor %using
  and  removing it from the observed signals, peaks can be reliably detected from concordant residual signals from a collection of SPods using a simple data-driven thresholding strategy.
@@
 The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
-As the size of the data and the number of quantiles grows, all solvers will eventually break \Halley{Eric}{Can we be more specific about what we mean when we say ``break?"}.
+As the size of the data and the number of quantiles grows, eventually the matrix of constraints will no longer fit into memory.
 To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
@@
 	\label{eq:eBIC}
-	\mbox{eBIC}_{\gamma}(s) & = & -2\log(L\{\Vhat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
+	\mbox{eBIC}_{\gamma}(\nu) & = & -2\log(L\{\Vhat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
 	\end{eqnarray*}
-	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given model.
-\Halley{Eric}{What is $s$?} We used this criteria with $\gamma = 1$, and $P=n-k-1$.
+	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod.
+We used this criteria with $\gamma = 1$, and $P=n-k-1$.
 \Fig{BIC} illustrates the difference among the scaled, unscaled ($\sigma = 1$), and scaled extended BIC criteria

### 2019-03-18 Code/Fig_app_full.R

@@ -36,6 +36,7 @@ ggplot(spodLong, aes(x=time, y=PID, col=node)) +
   labs(col="SPod")
 ggsave("../Manuscript/Figures/corrected_data.png", width = 7, height = 2.5)
 
+spodLong %>% filter(PID > 0.75)
 # ################################################################################
 detrendr_trends <- spod_trends
 methods <- c("detrendr")

### 2019-03-18 Manuscript/detrendify_JCGS.tex

--- 2019-03-16 Manuscript/detrendify_JCGS.tex
@@
 The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is  described in greater detail by \cite{boyd2011distributed}, but we briefly review how it can be used to iteratively solve the following equality constrained optimization problem.
+
 \begin{equation}
@@
 	\label{eq:eBIC}
-	\mbox{eBIC}_{\gamma}(s) & = & -2\log(L\{\Vhat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
+	\mbox{eBIC}_{\gamma}(\nu) & = & -2\log(L\{\Vhat{\theta}\}) + \nu\log n  + 2\gamma\log{P \choose \nu},~~\gamma \in [0,1]
 	\end{eqnarray*}
-	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given model.
-\Halley{Eric}{What is $s$?} We used this criteria with $\gamma = 1$, and $P=n-k-1$.
+
+	where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod.
+We used this criteria with $\gamma = 1$, and $P=n-k-1$.
 \Fig{BIC} illustrates the difference among the scaled, unscaled ($\sigma = 1$), and scaled extended BIC criteria

### 2019-03-24 Manuscript/detrendify_JCGS.tex
re-organized Manuscript folder

@@
 
-%	This work was motivated by the analysis of signals measured by low cost and portable air quality sensors, whose use has increased dramatically in the last decade.
 In the last decade, low cost and portable air quality sensors have enjoyed dramatically increased usage.
@@
 Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
-All of the sensors respond to the pollutant signal, the three sharp transient spikes at \Halley{Eric}{can you put the times at which the spikes occur here?}, but the baseline drift varies from one sensor to another, obscuring the detection of the peaks that alert the intrusion of pollutants.
+All of the sensors respond to the pollutant signal, the three sharp transient spikes at 11:32, 14:10, and 16:03, but the baseline drift varies from one sensor to another, obscuring the detection of the peaks that alert the intrusion of pollutants.
 We show later that by estimating the baseline drift in each sensor %using $\ell_1$-quantile-trend filtering
@@
 Thus, accurately demixing a noisy observed vector $\V{y}$ into a slowly varying component $\V{\theta}$ and a transient component $\V{s}$ can lead to greatly improved and simplified downstream analysis.
-\Halley{Joe}{I would cut the next sentence, as we haven't shown this for other applications.} This is true for not just air quality applications, but universally across the many domains listed earlier.
 
@@
 \item To reduce computation time and extend the method to long time series, we develop a parallelizable ADMM algorithm.
-\Halley{Joe}{Let's say a bit more about what this does.
-something like ``...
-a parallelizable ADMM algorithm that splits the time domain into overlapping windows and reconciles estimates from the overlapped portions.''}
+The algorithm proceeds by splitting the time domain into overlapping windows, fitting the model separately for each of the windows and reconciling estimates from the overlapping intervals.
 \item Finally, we propose a modified criterion for performing model selection.
@@
 
-	\input{full_confusion_detrend.tex}
+	\input{Tables/full_confusion_detrend.tex}
 

### 2019-03-24 Manuscript/detrendify_JCGS.tex
added manuscript in aoas template

@@
 	{
-		\title{\bf Baseline Drift Estimation for Time Series Data Using Quantile Trend Filtering}
+		\title{\bf Baseline Drift Estimation for Air Quality Data Using Quantile Trend Filtering}
 		\author{Halley Brantley\thanks{
@@
 	\noindent%
-	{\it Keywords:} quantile regression, non-parametric regression, trend estimation, smoothing splines
+	{\it Keywords:} air quality, quantile regression, non-parametric regression, trend estimation, smoothing splines
 	\vfill

### 2019-03-24 Manuscript/detrendify_aoas.tex
added manuscript in aoas template

--- 2019-03-24 Manuscript/detrendify_JCGS.tex
@@
-\documentclass[12pt]{article}
+% Template for the submission to:
+%   The Annals of Probability           [aop]
+%   The Annals of Applied Probability   [aap]
+%   The Annals of Statistics            [aos] 
+%   The Annals of Applied Statistics    [aoas]
+%   Stochastic Systems                  [ssy]
+%
+%Author: In this template, the places where you need to add information
+%        (or delete line) are indicated by {???}.  Mostly the information
+%        required is obvious, but some explanations are given in lines starting
+%Author:
+%All other lines should be ignored.  After editing, there should be
+%no instances of ??? after this line.
+
+% use option [preprint] to remove info line at bottom
+% journal options: aop,aap,aos,aoas,ssy
+% natbib option: authoryear
+\documentclass[aoas]{imsart}
 \RequirePackage{amsthm,amsmath,amsbsy,amsfonts}
@@
 \usepackage{bm,graphicx,psfrag,epsf}
-%\usepackage{enumerate}
 \usepackage{natbib}
@@
 \usepackage{subfig} %% For \subfloat command
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
+% provide arXiv number if available:
+%\arxiv{arXiv:0000.0000}
 
-\newtheorem{proposition}{Proposition}[section]
+% put your definitions there:
+\startlocaldefs
 
-%\pdfminorversion=4
-% NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{0}
-
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
-
+%% For comments to each other
 \usepackage[usenames]{color}
@@
 
-%% Turn on/off options for PDF
-\usepackage{ifpdf}
-
-%% For hyperlinks. Should always be the last package. Omit for DVI version.
-\ifpdf
-\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
-\else
-\newcommand{\href}[2]{{#2}}
-\usepackage{url}
-\fi
-
-%% ----------------------------------------------------------------------
-%% Hyper-linked References
-%% ----------------------------------------------------------------------
-
+\newtheorem{proposition}{Proposition}[section]
 \ifpdf
@@
 
+\endlocaldefs
+
 \begin{document}
 
-	%\bibliographystyle{natbib}
+\begin{frontmatter}
 
-	\def\spacingset#1{\renewcommand{\baselinestretch}%
-		{#1}\small\normalsize} \spacingset{1}
+% "Title of the paper"
+\title{Baseline Drift Estimation for Air Quality Data Using Quantile Trend Filtering}
+\runtitle{Quantile Trend Filtering}
 
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% indicate corresponding author with \corref{}
+% \author{\fnms{John} \snm{Smith}\corref{}\ead[label=e1]{smith@foo.com}\thanksref{t1}}
+% \thankstext{t1}{Thanks to somebody} 
+% \address{line 1\\ line 2\\ printead{e1}}
+% \affiliation{Some University}
 
-	\if0\blind
-	{
-		\title{\bf Baseline Drift Estimation for Time Series Data Using Quantile Trend Filtering}
-		\author{Halley Brantley\thanks{
-				Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: hlbrantl@ncsu.edu)} \,
-			Joseph Guinness\thanks{
-				Department of Statistics and Data Science, Cornell University, Ithaca, NY 14853 (E-mail: guinness@cornell.edu )} \,
-			and
-			%    and
-			Eric C. Chi\thanks{Department of Statistics, North Carolina State University, Raleigh, NC 27695 (E-mail: eric$\_$chi@ncsu.edu).}    \\}
-		\date{}
-		\maketitle
-	} \fi
+\author{\fnms{Halley L.} \snm{Brantley}\thanksref{m1}\ead[label=e1]{hlbrantl@ncsu.edu}},
+\author{\fnms{Joseph} \snm{Guinness}\thanksref{m2}\ead[label=e2]{guinness@cornell.edu}}
+\and
+\author{\fnms{Eric C.} \snm{Chi}\corref{}
+	\ead[label=e3]{eric$\_$chi@ncsu.edu}\thanksref{m1}}
+\affiliation{North Carolina State University\thanksmark{m1} and Cornell University\thanksmark{m2}}
+\address{Department of Statistics \\ 2311 Stinson Dr\\ Raleigh, NC 27606\\ \printead{e1}\\
+\printead{e3}}
 
-	\if1\blind
-	{
-		\bigskip
-		\bigskip
-		\bigskip
-		\begin{center}
-			{\LARGE\bf Title}
-		\end{center}
-		\medskip
-	} \fi
+\address{Department of Statistics and Data Science \\ 129 Garden Ave.\\ Ithaca, NY 14853\\ \printead{e2}}
 
-	\bigskip
+\runauthor{H. L. Brantley et al.}
+
 \begin{abstract}
@@
 
-	\noindent%
-	{\it Keywords:} quantile regression, non-parametric regression, trend estimation, smoothing splines
-	\vfill
+%\begin{keyword}[class=MSC]
+%\kwd[Primary ]{}
+%\kwd{}
+%\kwd[; secondary ]{}
+%\end{keyword}
 
-	\newpage
-	\spacingset{1.5} % DON'T change the spacing!
+\begin{keyword}
+\kwd{air quality}
+\kwd{non-parametric quantile regression}
+\kwd{trend estimation}
+\end{keyword}
 
-%% ----------------------------------------------------------------------
-%% Introduction
-%% ----------------------------------------------------------------------
+\end{frontmatter}
+
 \section{Introduction}
@@
 \cite{Koenker1978} were the first to propose substituting the sum-of-squares term with the check function  \Eqn{check} to estimate a conditional quantile instead of the conditional mean.
-Later, \cite{KoenkerNgPortnoy1994} proposed quantile trend filtering with $k = 2$ producing quantile trends that are piecewise linear \Halley{Joe}{Should this say $k=1$?
-If $k=0$ produces piecewise constant trends, wouldn't $k=1$ produce piecewise linear trends?}, but they did not consider extensions to higher order differences.
+Later, \cite{KoenkerNgPortnoy1994} proposed quantile trend filtering with $k = 1$ producing quantile trends that are piecewise linear, but they did not consider extensions to higher order differences.
 Rather than using the $\ell_1$-norm to penalize the discrete differences, \cite{nychka1995nonparametric} used the smoothing spline penalty based on the square of the $\ell_2$-norm:
@@
 \cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check function with a differentiable function.
-\cite{Racine2017} take a different approach and constrain the response to follow a smooth location scale model of the form $y(t) = a(t) + b(t)\epsilon_i$ and estimate the $\tau$\textsuperscript{th} conditional quantile using a kernel estimator and local linear approach.
+\cite{Racine2017} take a different approach and constrain the response to follow a location scale model and estimate the $\tau$\textsuperscript{th} conditional quantile using a kernel estimator and local linear approach.
 \Halley{Eric}{What are $a(t), b(t)$, and $\epsilon_i$?}
@@
 
+% AOS,AOAS: If there are supplements please fill:
+%\begin{supplement}[id=suppA]
+%  \sname{Supplement A}
+%  \stitle{Title}
+%  \slink[doi]{10.1214/00-AOASXXXXSUPP}
+%  \sdatatype{.pdf}" 
+%  \sdescription{Some text}
+%\end{supplement}
+
 \bigskip
@@
 
-	\bibliographystyle{asa}
+\bibliographystyle{imsart-nameyear}
 \bibliography{detrendify}

### 2019-03-25 Manuscript/detrendify_aoas.tex
Re-organized ADMM section, added discussion of detrendr_cross

@@
 Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
-To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) is not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
+To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) was not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
 Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
-All of the sensors respond to the pollutant signal, the three sharp transient spikes at 11:32, 14:10, and 16:03, but the baseline drift varies from one sensor to another, obscuring the detection of the peaks that alert the intrusion of pollutants.
+All of the sensors respond to the pollutant signal, which is illustrated by the three sharp transient spikes at 11:32, 14:10, and 16:03.
+However, the baseline drift varies from one sensor to another, obscuring the detection of the peaks that alert the intrusion of pollutants.
 We show later that by estimating the baseline drift in each sensor %using $\ell_1$-quantile-trend filtering
@@
 	\item \cite{Kim2009} proposed using the check function as a possible extension of $\ell_1$-trend filtering but did not investigate it further.
-Here, we develop the basic $\ell_1$-quantile-trend-filtering framework and extend it to model multiple quantiles simultaneously with non-crossing constraints to ensure valid trend estimates.
+Here, we develop the basic $\ell_1$-quantile-trend-filtering framework and extend it to model multiple quantiles simultaneously with non-crossing constraints to ensure validity and improve trend estimates.
 	\item To reduce computation time and extend the method to long time series, we develop a parallelizable ADMM algorithm.
@@
 \label{sec:methods}
-
-%We review \textit{$\ell_1$-trend filtering} for estimating a mean trend before discussing how to adapt it to estimate a baseline trend.
-We then describe a new scalable algorithm for simultaneously computing multiple quantile trends over large times windows.
 
@@
 
-%	$k+1$ fold composition of the discrete difference operator produces estimates that are piecewise polynomials of order $k$.
-
-%\cite{Tib2014} later demonstrated empirically that the trend filtering estimates adapt to the local level of smoothness better than smoothing splines and also proved that with a judicious choice of tuning parameter the trend filtering estimate converges to the true underlying function at the minimax rate for functions whose $k$th derivative is of bounded variation.
-%
-%	Given observations $y(t)$ with $t=1,...,N$ and defining $y \in \Real^n$ such that $y_i = y(t)$, in the trend filtering problem, the trend, $\V{\theta} \in \Real^n$, is estimated by solving the following convex problem:
-%	\begin{eqnarray}
-%	\underset{\V{\theta}}{\arg\min}\; \frac{1}{2} \lVert y - \V{\theta} \rVert_2^2 + \lambda \lVert \Mn{D}{k+1}\V{\theta} \rVert_1,
-%	\end{eqnarray}
-%	where $y \in \Real^n$ is the vector of observations, $\lambda \geq 0$ is a regularization parameter, and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
-To understand the purpose of penalizing $\Mn{D}{k+1}$ consider the difference operator when $k = 0$.
-%	\begin{eqnarray*}
-%	\Mn{D}{1} & = & \begin{pmatrix}
-%
-%	-1 & 1 & 0 & \cdots & 0 & 0 \\
-%	0 & -1 & 1 & \cdots & 0 & 0 \\
-%	\vdots & & & & & \\
-%	0 & 0 & 0 & \cdots & -1 & 1 \\
-%
-%	\end{pmatrix}.
-%	\end{eqnarray*}
-%	Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$ which is just total variation denoising in one dimension.
-The penalty term incentivizes solutions which are piecewise constant.
-For $k \geq 1$, the difference operator $\Mn{D}{k+1} \in \Real^{(n-k-1) \times n}$ is defined recursively as follows
-%	\begin{eqnarray*}
-%	\Mn{D}{k+1} & = & \Mn{D}{1}\Mn{D}{k}.
-%	\end{eqnarray*}
-%	By penalizing the $k+1$ fold composition of the discrete difference operator, we obtain solutions which are piecewise polynomials of order $k$.
-
 \cite{Tib2014} proved that with a judicious choice of $\lambda$ the trend filtering estimate converges to the true underlying function at the minimax rate for functions whose $k$th derivative is of bounded variation and showed that trend filtering is both fast and locally adaptive when the time series consists of only the trend and random noise, which is illustrated in \Fig{tpn}.
-As noted earlier though in some applications, such as the air quality monitoring problem considered in this paper, the data contain a rapidly varying signal in addition to the slowly varying trend and noise.
+As noted earlier, in some applications, such as the air quality monitoring problem considered in this paper, the data contain a rapidly varying signal in addition to the slowly varying trend and noise.
 \Fig{tpspn} shows that standard trend filtering is not designed to distinguish between the slowly varying trend and the rapidly-varying signal, as the smooth component estimate $\V{\theta}$ is biased towards the peak of the transient component.
@@
 \label{eq:check}
-\rho_{\tau}(\V{r}) & = & \sum_{i=1}^n \VE{r}{i}(1-\One(\VE{r}{i}<0)),
+\rho_{\tau}(\V{r}) & = & \sum_{i=1}^n \VE{r}{i}(\tau-\One(\VE{r}{i}<0)),
 \end{eqnarray}
 and $\One(A)$ is 1 if its input $A$ is true and 0 otherwise.
+Note that we do not explicitly model $\V{s}$.
+Rather, we focus on estimating $\V{\theta}$.
+We then estimate $\V{s} + \V{\varepsilon}$ as the difference $\V{y} - \V{\theta}$.
 
@@
 	\caption{Examples of trend filtering solutions (red) and 15\textsuperscript{th} quantile trending filtering solution (blue).
-Standard trend filtering performs well in the no-signal case (right) but struggles to distinguish between the slowly varying trend and the rapidly-varying signal (left).
+Standard trend filtering performs well in the no-signal case (a) but struggles to distinguish between the slowly varying trend and the rapidly-varying signal (b).
 The quantile trend is not affected by the signal and provides an estimate of the baseline.}
 \end{figure}
-Note that we do not explicitly model $\V{s}$. Rather, we focus on estimating $\V{\theta}$. We then estimate $\V{s} + \V{\varepsilon}$ as the difference $\V{y} - \V{\theta}$.
 
-Before elaborating on how we compute our proposed $\ell_1$-quantile trend filtering estimator, we discuss similarities and differences between our proposed estimator and existing baseline estimators.
+Before elaborating on how we compute our proposed $\ell_1$-quantile trend filtering estimator, we discuss similarities and differences between our proposed estimator and existing quantile trend estimators.
 
@@
 \cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check function with a differentiable function.
-\cite{Racine2017} take a different approach and constrain the response to follow a location scale model and estimate the $\tau$\textsuperscript{th} conditional quantile using a kernel estimator and local linear approach.
-\Halley{Eric}{What are $a(t), b(t)$, and $\epsilon_i$?}
+\cite{Racine2017} take a different approach and constrain the response to follow a location scale model and estimate the $\tau$\textsuperscript{th} conditional quantile by combining Gaussian quantile functions with a kernel smoother and solving a local-linear least squares problem.
 
@@
 As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of methods.
-In many cases, including ours, we are interested in estimating multiple quantiles simultaneously.
-We also want to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$ where $Q$ is the quantile function of $\V{y}$.
+We want to estimated multiple quantiles simultaneously and to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$ where $Q$ is the quantile function of $\V{y}$.
+Even if a single quantile is ultimately desired, ensuring non-crossing allows information from nearby quantiles to be used to improve the estimates.
 Given quantiles $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
@@
 	\label{eq:noncross_trend}
-	%	&&\underset{\V{\theta}_1, \ldots, \V{\theta}_J}{\text{minimize}}\; \sum_{j=1}^J \left [\rho_{\tau_j}(\V{y} - \V{\theta}_{j}) +
-	%	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
 	&&\underset{\M{\Theta} \in \mathcal{C}}{\text{minimize}}\; \sum_{j=1}^J \left [\rho_{\tau_j}(\V{y} - \V{\theta}_{j}) +
-	\lambda \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
-	%	&&\text{subject to: }\; \V{\theta}_1 \leq \V{\theta}_2 \leq \ldots \leq \V{\theta}_J
-	%	\theta_{1}(t) \le \theta_{2}(t) \le \ldots \le \theta_{J}(t) \text{ for all } t, \nonumber
+	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
 \end{eqnarray*}
@@
 The additional constraints are linear in the parameters, so the non-crossing quantile trends can still be estimated by a number of available solvers.
+We allow for the possibilty that the degree of smoothness in the trends varies by quantile and thus allow the smoothing parameter as well.
 In the rest of this paper, we report numerical results using the commercial solver Gurobi \citep{gurobi} and its R package implementation.
@@
 
-The number of parameters to be estimated in this problem is equal to the number of observations multiplied by the number of quantiles of interest.
-As the size of the data and the number of quantiles grows, eventually the matrix of constraints will no longer fit into memory.
+ As the size of the data increases, all methods require additional computation time and for our method eventually the matrix of constraints no longer fits into memory.
+In our application to air quality sensor data, measurements are recorded every second resulting in 86,400 observations per day.
+This number of observations is already too large to use with currently available R packages used for estimating quantile trends \citep{fields, quantreg}.
 To our knowledge, no one has addressed the problem of finding smooth quantile trends of series that are too large to be processed simultaneously.
 We propose a divide-and-conquer approach via an ADMM algorithm for solving large problems in a piecewise fashion.
-The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is  described in greater detail by \cite{boyd2011distributed}, but we briefly review how it can be used to iteratively solve the following equality constrained optimization problem.
+
+ \begin{figure}[!t]
+ 	\centering
+ 	\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
+ 	\caption{Window boundaries and trends fit separately in each window.}
+ 	\label{fig:windows}
+ \end{figure}
+
+ Let $\mathcal{I}$ denote a subset of the indices $\{1, \ldots, n\}$ and $\V{y}_\mathcal{I} \in \Real^{\lvert I \rvert}$ denote the subvector obtained by stacking on top of each other the elements $\VE{y}{i}$ for $i \in \mathcal{I}$ in their canonical ordering.
+ To decrease computation and extend our method to larger problems, we divide our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into $W$ overlapping windows of observations $\Vn{y}{w} = \V{y}_{\mathcal{I}_w}$ where $\mathcal{I}_w$ is the set of integers from $l_w$ to $u_w$ inclusive, namely $\mathcal{I}_w = \{l_w, l_w + 1, \ldots, u_w -1, u_w\}$ and
+ \begin{eqnarray*}
+ 	1 = l_{1} < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots < u_{W} = n.
+ \end{eqnarray*}
+ \Fig{windows} shows an example of 1200 observations being mapped into three equally sized overlapping windows of observations.
+While the overlapping trend estimates between $l_2$ and $u_2$ don't vary dramatically, the difference is more pronounced in the trend in the 5\textsuperscript{th} quantile between $l_3$ and $u_2$.
+
+ Given quantiles $\tau_1 < \cdots < \tau_J$, we introduce dummy variables $\Vn{\theta}{w}_{j} \in \Real^{\lvert \mathcal{I}_w \rvert}$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $w$.
+We then ``stitch" together the $W$ quantile trend estimates into consensus over the overlapping regions by introducing the constraint $\VnE{\theta}{w}{ij} = \VE{\theta}{i + l_w - 1,j}$ for $i = 1, \ldots, u_w - l_w + 1$ and for all $j$.
+Let $\Mn{\Theta}{w}$ be the matrix whose $j$th column is $\Vn{\theta}{w}_{j}$.
+Then we can write these constraints more compactly as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{\lvert \mathcal{I}_w \rvert \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $w$th window, namely
+ \begin{eqnarray*}
+ 	\Mn{U}{w} & = & \begin{pmatrix}
+ 		\V{e}_{l_w}\Tra \\
+ 		\vdots \\
+ 		\V{e}_{u_w}\Tra
+ 	\end{pmatrix},
+ \end{eqnarray*}
+ where $\V{e}_i \in \Real^{n}$ denotes the $i$th standard basis vector.
+Furthermore, let $\iota_{\mathcal{C}}$ denote the indicator function of the non-crossing quantile constraint, namely $\iota_{\mathcal{C}}(\M{\Theta})$ is zero if $\M{\Theta} \in \mathcal{C}$ and infinity otherwise.
+
+ Our windowed quantile trend optimization problem can then be written as
+ \begin{equation}
+ \label{eq:quantile_windows}
+ \begin{split}
+ &\text{minimize}\; \sum_{w=1}^W \left \{\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{w} - \Vn{\theta}{w}_{j}) +
+ \lambda_j \lVert \Mn{D}{k+1} \Vn{\theta}{w}_{j} \rVert_1 \right ] + \iota_\mathcal{C}(\Mn{\Theta}{w}) \right\}\\
+ &\text{subject to} \qquad \Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta} \;\; \text{ for } w = 1, \ldots, W.
+ \end{split}
+ \end{equation}
+
+ The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is  described in greater detail by \cite{boyd2011distributed}, but we briefly review how it can be used to iteratively solve the following equality constrained optimization problem which is a more general form of \Eqn{quantile_windows}.
 
@@
 \end{equation}
-\Eric{Joe}{I'm already lost here.
-Can we give some context to what $\theta$ and $\widetilde{\theta}$ are in the context of our windowing algorithm?
-Maybe we should say something about the windowing first before we dive head first into Lagrangians?
-What are $A$ and $B$?
-Do these simplify to something in our specific application?
-The stuff below might seem obvious to you, but I'm totally lost.
-It might be better to talk about the notation for the windows first, and how different windows will give different thetas for the overlapping parts, and then talk about equality constrained optimization as a solution for the overlapping parts.} Recall that finding the minimizer to an equality constrained optimization problem is equivalent to the identifying the saddle point of the Lagrangian function associated with the problem \Eqn{split_objective}.
+Recall that finding the minimizer to an equality constrained optimization problem is equivalent to the identifying the saddle point of the Lagrangian function associated with the problem \Eqn{split_objective}.
 ADMM seeks the saddle point of a related function called the augmented Lagrangian,
@@
 
-Let $\mathcal{I}$ denote a subset of the indices $\{1, \ldots, n\}$ and $\V{y}_\mathcal{I} \in \Real^{\lvert I \rvert}$ denote the subvector obtained by stacking on top of each other the elements $\VE{y}{i}$ for $i \in \mathcal{I}$ in their canonical ordering.
-We apply the consensus ADMM algorithm to the quantile regression trend filtering problem given in \Eqn{quantile_trend}, by dividing our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into overlapping $W$ windows of observations $\Vn{y}{w} = \V{y}_{\mathcal{I}_w}$ where $\mathcal{I}_w$ is the set of integers from $l_w$ to $u_w$ inclusive, namely $\mathcal{I}_w = \{l_w, l_w + 1, \ldots, u_w -1, u_w\}$ where
-\begin{eqnarray*}
-	1 = l_{1} < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < \cdots < u_{W} = n.
-\end{eqnarray*}
-\Fig{windows} shows an example of 1200 observations being mapped into three equally sized overlapping windows of observations.
-
-%	\begin{align*}
-%	y_1(t) = y(t) & \mbox{~~if~~} 1 \le t \le u_{1}\\
-%	y_2(t) = y(t) & \mbox{~~if~~} l_{2} \le t \le u_{2} \\
-%	\vdots & \\
-%	y_M(t) = y(t) & \mbox{~~if~~} l_{M} \le t \le  n,
-%	\end{align*}
-%
-%	 with boundaries $1 < l_{2} < u_{1} < l_{3} < u_{2} < l_{4} < u_{3} < ...< n$.
-%	 Given quantiles $\tau_1 < \cdots < \tau_J$, we define $\theta_{j,m}(t)$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $m$ at time $t$.
-In order to constrain the overlapping sections to be equal, we define a consensus variable
-Given quantiles $\tau_1 < \cdots < \tau_J$, we introduce dummy variables $\Vn{\theta}{w}_{j} \in \Real^{\lvert \mathcal{I}_w \rvert}$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $w$.
-We then ``stitch" together the $W$ quantile trend estimates into consensus over the overlapping regions by introducing the constraint $\VnE{\theta}{w}{ij} = \VE{\theta}{i + l_w - 1,j}$ for $i = 1, \ldots, u_w - l_w + 1$ and for all $j$.
-Let $\Mn{\Theta}{w}$ be the matrix whose $j$th column is $\Vn{\theta}{w}$.
-Then we can write these constraints more compactly as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{\lvert \mathcal{I}_w \rvert \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $w$th window, namely
-\begin{eqnarray*}
-	\Mn{U}{w} & = & \begin{pmatrix}
-		\V{e}_{l_w}\Tra \\
-		\vdots \\
-		\V{e}_{u_w}\Tra
-	\end{pmatrix},
-\end{eqnarray*}
-where $\V{e}_i \in \Real^{n}$ denotes the $i$th standard basis vector.
-Furthermore, let $\iota_{\mathcal{C}}$ denote the indicator function of the non-crossing quantile constraint, namely $\iota_{\mathcal{C}}(\M{\Theta})$ is zero if $\M{\Theta} \in \mathcal{C}$ and infinity otherwise.
-
-Our windowed quantile trend optimization problem can then be written as
-\begin{equation}
-\label{eq:quantile_windows}
-\begin{split}
-&\text{minimize}\; \sum_{w=1}^W \left \{\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{w} - \Vn{\theta}{w}_{j}) +
-\lambda \lVert \Mn{D}{k+1} \Vn{\theta}{w}_{j} \rVert_1 \right ] + \iota_\mathcal{C}(\Mn{\Theta}{w}) \right\}\\
-&\text{subject to} \qquad \Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta} \;\; \text{ for } w = 1, \ldots, W.
-\end{split}
-\end{equation}
-
-Let $\Mn{\Omega}{w}$ denote the Lagrange multiplier matrix for the $w$th consensus constraint, namely $\Mn{\theta}{w} = \Mn{U}{w}\M{\theta}$, and let $\Vn{\omega}{w}_{j}$ denote its $j$th column.
-%
-%		  $\V{\omega}_{m}_{j}$ denote Lagrange multipliers for the constraint $\Vn{\theta}{w}_j = \Mn{U}{m}\V{\theta}_j$, $\Mn{\Omega}{m}$ denote the matrix whose $j$th column is $\V{\omega}_{m}_j$, and penalty parameter $\gamma > 0$,
+Returning to our constrained windows problem giving in \Eqn{quantile_windows}, let $\Mn{\Omega}{w}$ denote the Lagrange multiplier matrix for the $w$th consensus constraint, namely $\Mn{\theta}{w} = \Mn{U}{w}\M{\theta}$, and let $\Vn{\omega}{w}_{j}$ denote its $j$th column.
 
@@
 \begin{eqnarray*}
-	\mathcal{L}_w(\M{\Theta}, \Mn{\Theta}{w}, \Mn{\Omega}{w}) & = & \sum_{j=1}^J \biggl [\rho_{\tau_j}(\Vn{y}{w} - \Vn{\theta}{w}_j)+\lambda \lVert \Mn{D}{k+1}\Vn{\theta}{w}_j\rVert_1 \\
+	\mathcal{L}_w(\M{\Theta}, \Mn{\Theta}{w}, \Mn{\Omega}{w}) & = & \sum_{j=1}^J \biggl [\rho_{\tau_j}(\Vn{y}{w} - \Vn{\theta}{w}_j)+\lambda_j \lVert \Mn{D}{k+1}\Vn{\theta}{w}_j\rVert_1 \\
 	&& + \; (\Vn{\theta}{w}_j - \Mn{U}{w}\V{\theta}_{j})\Tra\Vn{\omega}{w}_{j} +
@@
 The ADMM algorithm alternates between updating the consensus variable $\M{\Theta}$, the window variables $\{\Mn{\Theta}{w}\}$, and the Lagrange multipliers $\{\Mn{\Omega}{w}\}$.
-At the $m+1$th iteration, we perform the following sequence of updates
+At the $(m+1)$\textsuperscript{th} iteration, we perform the following sequence of updates
 
@@
 \label{eq:update_consensus}
-\ME{\theta}{ij}	 &=& \begin{cases}
-\frac{1}{2}\left (\MnE{\theta}{w-1}{ij} - \gamma\Inv \MnE{\omega}{w-1}{ij} \right)
-+
-\frac{1}{2} \left (\MnE{\theta}{w}{ij} - \gamma\Inv \MnE{\omega}{w}{ij} \right)
-& \mbox{if~~} l_{w} \le i \le u_{w-1},  \\
-\MnE{\theta}{w}{ij} & \mbox{if~~} u_{w-1} < i \le l_{w+1}  \\
-\frac{1}{2} \left ( \MnE{\theta}{w}{ij} - \gamma\Inv \MnE{\omega}{w}{ij}\right )
-+
-\frac{1}{2}	\left ( \MnE{\theta}{w+1}{ij} - \gamma\Inv \MnE{\omega}{w+1}{ij} \right )
-& \mbox{if~~} l_{m+1} < i \le u_{m}
+~~~~\ME{\theta}{ij}	 &=& \begin{cases}
+\frac{1}{2}\left (\MnE{\theta}{w-1}{ij} + \MnE{\theta}{w}{ij} \right)
+-
+\frac{1}{2\gamma} \left (\MnE{\omega}{w-1}{ij} + \MnE{\omega}{w}{ij} \right)
+& \mbox{if~} l_{w} \le i \le u_{w-1},  \\
+\MnE{\theta}{w}{ij} & \mbox{if~} u_{w-1} < i \le l_{w+1}  \\
+\frac{1}{2} \left ( \MnE{\theta}{w}{ij} + \MnE{\theta}{w+1}{ij} \right )
+-
+\frac{1}{2\gamma}	\left (\MnE{\omega}{w}{ij} + \MnE{\omega}{w+1}{ij} \right )
+& \mbox{if~} l_{m+1} < i \le u_{m}
 \end{cases},
@@
 The \texttt{gurobi} R package \citep{gurobi} can solve quadratic programs in addition to linear programs, but we can also use the free R package \texttt{quadprog} \citep{quadprog}.
-
-%	 \begin{eqnarray*}
-%	\M{A}r{\theta}_{j, m}	&=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
-%	\M{A}r{\theta}_{j, m}(t)	 &=& \begin{cases}
-%			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
-%			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
-%			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}
-%			 \end{cases},
-%	\end{eqnarray*}
-%
-%	 \begin{align*}
-%	\M{A}r{\theta}_{j, m}	=&  g(\theta_{j, m-1}, \theta_{j,m}, \theta_{j,m+1}) \\
-%	\M{A}r{\theta}_{j, m}(t)	 =& \begin{cases}
-%			 \frac{\theta_{j,m-1}(t)+\theta_{j,m}(t)}{2} & \mbox{if~~} l_{m} \le t \le u_{m-1}  \\
-%			 \theta_{j,m}(t) & \mbox{if~~} u_{m-1} \le t \le l_{m+1}  \\
-%			 \frac{\theta_{j,m}(t)+\theta_{j,m+1}(t)}{2} & \mbox{if~~} l_{m+1} \le t \le u_{m}
-%			 \end{cases},
-%	\end{align*}
-%	defining $\theta_{j,M+1} = \theta_{j,M}$ and $\theta_{j,0} = \theta_{j,1}$. Our windowed quantile trend optimization problem can then be written as
-%	 \begin{align*}
-%		 \label{eq:quantile_windows}
-%		 &\underset{\theta_{1,m}, \ldots, \theta_{J,m}}{\text{minimize}}\; \sum_{m=1}^M\sum_{j=1}^J \left [\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m}) +
-%		 \lambda_j \lVert \Mn{D}{k+1} \theta_{j,m} \rVert_1 \right ] \\
-%		 &\text{subject to: }\; \\
-%		 &\qquad \theta_{1,m}(t) \le \theta_{2,m}(t) \le \ldots \le \theta_{J,m}(t) \text{ for all } m,t \\
-%		 &\qquad \theta_{j,m}(t) = \M{A}r{\theta}_{j,m}(t) \text{ for all } j, m, t
-%	 \end{align*}
-%	 Defining Lagrange multipliers $\omega_{j,m}$, and penalty parameter $\gamma > 0$, the augmented Lagrangian for finding the trends in window $m$ is
-%	 \begin{equation*}
-%	 \mathcal{L}(\theta_{j,m}, \M{A}r{\theta}_{j,m}, \omega_{j,m}) = \sum_{j=1}^J\rho_{\tau_j}(\Vn{y}{m} - \theta_{j,m})+\lambda \lVert \Mn{D}{k+1}\theta_{j,m}\rVert_1 +  \omega_{j,m}\Tra(\theta_{j,m} - \M{A}r{\theta}_{j,m}) +
-%	 \frac{\gamma}{2}||\theta_{j,m} - \M{A}r{\theta}_{j,m}||_2^2
-%	 \end{equation*}
-%	 We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in Algorithm 1.
-The use of the Augmented Lagrangian converts the problem from a linear program into a quadratic program.
-The \texttt{gurobi} R package \citep{gurobi} can solve quadratic programs in addition to linear programs, but we can also use the free R package \texttt{quadprog} \citep{quadprog}.
-%
-\begin{figure}[!t]
-	\centering
-	\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
-	\caption{Window boundaries and trends fit separately in each window.}
-	\label{fig:windows}
-\end{figure}
 
@@
 \label{eq:modcheck}
-\tilde{\rho}_{\tau}(\V{r}) & = & \sum_{i \not\in \mathcal{V}} \VE{r}{i}(1-\One(\VE{r}{i}<0)),
+\tilde{\rho}_{\tau}(\V{r}) & = & \sum_{i \not\in \mathcal{V}} \VE{r}{i}(\tau-\One(\VE{r}{i}<0)),
 \end{eqnarray}
@@
 \end{eqnarray}
-which can be solved via \Alg{admm} with trivial modification to the quadratic program subproblems.
-We then select the $\lambda$ that minimizes the hold-out prediction error quantified by $\breve{\rho}(\V{y} - \Vhat{\theta}(\lambda))$ where $\Vhat{\theta}(\lambda)$ is the solution to \Eqn{validate} and
+which can be solved via \Alg{admm} with trivial modification to the quadratic program sub-problems.
+We fit a sequence of values for $\lambda$ setting $\lambda_j = \lambda$ for all $j$.
+Then for each quantile level $j$ we select the $\lambda_j$ that minimizes the hold-out prediction error quantified by $\breve{\rho}_{\tau_j}(\V{y} - \Vhat{\theta}_j(\lambda))$ where $\Vhat{\theta}_j(\lambda)$ is the solution to \Eqn{validate} and
 \begin{eqnarray*}
-	\breve{\rho}_{\tau}(\V{r}) & = & \sum_{i \in \mathcal{V}} \VE{r}{i}(1-\One(\VE{r}{i}<0)).
+	\breve{\rho}_{\tau}(\V{r}) & = & \sum_{i \in \mathcal{V}} \VE{r}{i}(\tau-\One(\VE{r}{i}<0)).
 \end{eqnarray*}
@@
 where $p_{\lambda} = \sum_i \One(\VE{y}{i} = \widehat{\theta}_i)$ is the number of interpolated points, which can be thought of as active knots.
+Equivalently, $p_{\lambda}$ can be substituted with the number of non-zero components in $\M{D}^{(k+1)}\V{\theta}$ which we denote $\nu$ and have found to be more numerically stable.
 The SIC is based on the traditional Bayesian Information Criterion (BIC) which is given by
@@
 \end{equation}
-where $L$ is the likelihood function and $\nu$ is the number of non-zero components in $\Vhat{\theta}$.
+where $L$ is the likelihood function.
 If we take the approach used in Bayesian quantile regression \citep{yu2001bayesian}, and view minimizing the check function as maximizing the asymmetric Laplace likelihood,
@@
 
-where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod.
-We used this criteria with $\gamma = 1$, and $P=n-k-1$.
-\Fig{BIC} illustrates the difference among the scaled, unscaled ($\sigma = 1$), and scaled extended BIC criteria
-when applied to a single dataset replicate from our simulation study. In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
+where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod. We used this criteria with $\gamma = 1$, and $P=n-k-1$.  
+As in the validation method, to chose $\V{\lambda}$ using the information criteria we fit a sequence of values for $\lambda$ setting $\lambda_j = \lambda$ for all $j$.
+For each quantile level $j$, we select the $\lambda_j$ that minimizes the SIC, BIC, or eBIC value evaluated using $\tau_j$ and $\V{\theta}_j$.
+In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
 
-\begin{figure}[h!]
-	\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
-	\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
-	\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
-	\caption{(Left) Estimated degrees of freedom (number of non-zero elements of $\M{D}\V{\theta}$) as a function of $\log(\lambda)$.
+%\begin{figure}[h!]
+%	\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
+%	\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
+%	\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
+%	\caption{(Left) Estimated degrees of freedom (number of non-zero elements of $\M{D}\V{\theta}$) as a function of $\log(\lambda)$.
 (Middle) Estimated 10th quantile trend with regularization parameter chosen using various criteria.
 (Right) Criteria values as a function of $\log(\lambda)$, vertical lines indicate locations of minima.}
-	\label{fig:BIC}
-\end{figure}
+%	\label{fig:BIC}
+%\end{figure}
 
@@
 
-We compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window: \texttt{detrendr\_SIC} $\lambda$ chosen using SIC \Eqn{SIC} \citep{KoenkerNgPortnoy1994}; \texttt{detrendr\_valid}: $\lambda$ is chosen by leaving out every 5th observation as a validation data set and minimizing the check loss function evaluated at the validation data; \texttt{detrendr\_eBIC}:  the proposed scaled eBIC criteria \Eqn{eBIC}.
+We compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window:  $\V{\lambda}$ chosen using SIC \Eqn{SIC} \citep{KoenkerNgPortnoy1994} (\texttt{detrendr\_SIC}); $\V{\lambda}$ chosen using the validation method with the validation set consisting of every 5th observation (\texttt{detrendr\_valid}); and $\V{\lambda}$ chosen using the proposed scaled eBIC criteria \Eqn{eBIC} \texttt{detrendr\_eBIC}.
+For the second study we also examine the effect of the non-crossing quantile constraint by estimating the quantile trends separately and choosing $\V{\lambda}$ using the proposed scaled eBIC criteria (\texttt{detrendr\_cross}).
+We do not include \texttt{detrendr\_cross} in the first study because the difference in quantiles is large enough that we would not expect the non-crossing constraint to make a difference.
 
 We also compare the performance of our quantile trend filtering method with three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011} and \cite{nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
-The regularization parameter $\lambda$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria as described in \cite{KoenkerNgPortnoy1994}, the regularization parameter for \texttt{qsreg} was chosen using generalized cross-validation based on the quantile criterion \cite{Oh2011}.
+Only our detrend methods guarantee non-crossing quantiles.
+The regularization parameter $\V{\lambda}$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria separately for each quantile as described in \cite{KoenkerNgPortnoy1994}, the regularization parameter for \texttt{qsreg} was chosen using generalized cross-validation separately for each quantile based on the quantile criterion \cite{Oh2011}.
 
@@
 \begin{figure}
-	\includegraphics[width=.28\linewidth]{Figures/gaus.png}
-	\includegraphics[width=.3\linewidth]{Figures/shapebeta.png}
-	\includegraphics[width=.45\linewidth]{Figures/mixednorm.png}
+	\includegraphics[width=.27\linewidth]{Figures/gaus.png}
+	\includegraphics[width=.29\linewidth]{Figures/shapebeta.png}
+	\includegraphics[width=.44\linewidth]{Figures/mixednorm.png}
 	\caption{Simulated data with true quantile trends.}
+	\label{fig:sim1}
 \end{figure}
 
-To compare performance in estimating quantile trends, three simulation designs from \cite{Racine2017} were considered.
+To compare performance in estimating quantile trends in the absence of a signal component, three simulation designs from \cite{Racine2017} were considered.
 For all designs $t = 1, \ldots, n$,  $x(t) = t/n$, and the response $\V{y}$ was generated as
@@
 \end{eqnarray*}
-The errors were independent and had distributions
+The errors were independent and had distributions:
 \begin{itemize}
@@
 \end{itemize}
+The true quantile trends and an example simulated data set is show in \Fig{sim1}. One hundred datasets were generated of sizes 300, 500 and 1000.
 
-One hundred datasets were generated of sizes 300, 500 and 1000.
 For each method quantile trends were estimated for $\tau = \{0.05, 0.25, 0.5, 0.75, 0.95\}$.
-Only our detrend methods guarantee non-crossing quantiles.
-For each quantile trend and method the root mean squared error was calculated as $\mbox{RMSE} = \sqrt{\frac{1}{n}\sum_i (\hat{q}_{\tau}(t_i) - q_\tau(t_i))^2}$.
-\Fig{quantile_mse} shows the mean RMSE $\pm$ twice the standard error for each method, quantile level, and sample size.
+For each quantile trend and method the root mean squared error was calculated as $\mbox{RMSE} = \sqrt{\frac{1}{n}\sum_{i=1}^n (\hat{q}_{\tau}(t_i) - q_\tau(t_i))^2}$.
+\Fig{quantile_mse} shows the mean RMSE plus or minus twice the standard error for each method, quantile level, and sample size.
 In all three designs the proposed detrend methods are either better than or comparable to existing methods.
@@
 The simulated peaks were multiplied by a factor that was randomly drawn from a normal distribution with mean 20 and standard deviation of 4.
+An example dataset with 4 signal peaks is shown in \Fig{peaks_class_eg}.
 One hundred datasets were generated for each $n=\{500, 1000, 2000, 4000\}$.
-\Halley{Joe}{This is a lot to take in.
-Can you refer to a figure that shows a simulated dataset?} We compare the ability of the methods to estimate the true quantiles of $y(t_i)-s(t_i)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}).
+
+We compare the ability of the methods to estimate the true quantiles of $y(t_i)-s(t_i)$  for $\tau \in \{0.01, 0.05, 0.1\}$ and calculate the RMSE (\Fig{peaks_rmse}).
 In this simulation study, our proposed method \texttt{detrend\_eBIC} method substantially outperforms the others.
+The \texttt{detrend\_cross} method, which is the \texttt{detrend\_eBIC} method fit without the non-crossing constraints, performs similarly for larger quantiles and larger datasets.
+However, \texttt{detrend\_cross} produces significantly worse estimates for more extreme quantiles ($\tau = 0.01$) and smaller data sets ($\tau = 0.05$ and $n = 500$).
+These results indicate that even when a single quantile is of interest, simultaneously fitting nearby quantiles and utilizing the non-crossing constraint can improve estimates when data is sparse by using information from nearby quantiles.
 The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets, but its performance deteriorates as the data size grows.
@@
 We use this metric because our classes tend to be very imbalanced with many more zeros than ones.
-The CAA metric will give a score of 0.5 for random guessing and also for trivial classifiers such as $\hat{\delta}_i = 0$ for all $i$.
+The CAA metric should give a score close to 0.5 both for random guessing and also for trivial classifiers such as $\hat{\delta}_i = 0$ for all $i$.
 

### 2019-03-26 Manuscript/detrendify_aoas.tex
application week edits and manuscript edits

@@
  Let $\mathcal{I}$ denote a subset of the indices $\{1, \ldots, n\}$ and $\V{y}_\mathcal{I} \in \Real^{\lvert I \rvert}$ denote the subvector obtained by stacking on top of each other the elements $\VE{y}{i}$ for $i \in \mathcal{I}$ in their canonical ordering.
- To decrease computation and extend our method to larger problems, we divide our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into $W$ overlapping windows of observations $\Vn{y}{w} = \V{y}_{\mathcal{I}_w}$ where $\mathcal{I}_w$ is the set of integers from $l_w$ to $u_w$ inclusive, namely $\mathcal{I}_w = \{l_w, l_w + 1, \ldots, u_w -1, u_w\}$ and
+ To decrease computation time and extend our method to larger problems, we divide our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into $W$ overlapping windows of observations $\Vn{y}{w} = \V{y}_{\mathcal{I}_w}$ where $\mathcal{I}_w$ is the set of integers from $l_w$ to $u_w$ inclusive, namely $\mathcal{I}_w = \{l_w, l_w + 1, \ldots, u_w -1, u_w\}$ and
  \begin{eqnarray*}
@@
 \begin{split}
-&\text{minimize} \; f(\V{\theta}) + g(\Vtilde{\theta}) \\
-&\text{subject to} \; \M{A}\V{\theta} + \M{B}\Vtilde{\theta} = \V{c}.
+&\text{minimize} \; f(\V{\phi}) + g(\Vtilde{\phi}) \\ 
+&\text{subject to} \; \M{A}\V{\phi} + \M{B}\Vtilde{\phi} = \V{c}.
 \end{split}
@@
 \begin{eqnarray*}
-	\mathcal{L}_{\gamma}(\V{\theta},\Vtilde{\theta},\V{\omega}) & = & f(\V{\theta}) + g(\Vtilde{\theta}) + \langle \V{\omega}, \V{c} - \M{A}\V{\theta} - \M{B}\Vtilde{\theta} \rangle
-	+ \frac{\gamma}{2} \lVert \V{c} - \M{A}\V{\theta} - \M{B}\Vtilde{\theta} \rVert_2^2,
+	\mathcal{L}_{\gamma}(\V{\phi},\Vtilde{\phi},\V{\omega}) & = & f(\V{\phi}) + g(\Vtilde{\phi}) + \langle \V{\omega}, \V{c} - \M{A}\V{\phi} - \M{B}\Vtilde{\phi} \rangle
+	+ \frac{\gamma}{2} \lVert \V{c} - \M{A}\V{\phi} - \M{B}\Vtilde{\phi} \rVert_2^2,
 \end{eqnarray*}
@@
 
-ADMM minimizes the augmented Lagrangian one block of variables at a time before updating the dual variable $\V{\omega}$. This yields the following sequence of updates at the $m+1$th ADMM iteration
+ADMM minimizes the augmented Lagrangian one block of variables at a time before updating the dual variable $\V{\omega}$.
+This yields the following sequence of updates at the $(m+1)$\textsuperscript{th} ADMM iteration
 \begin{equation}
@@
 \begin{split}
-\V{\theta}_{m+1} & \amp = \amp \underset{\V{\theta}}{\arg\min}\; \mathcal{L}_\gamma(\V{\theta},\Vtilde{\theta}_{m}, \V{\omega}_{m}) \\
-\Vtilde{\theta}_{m+1} & \amp = \amp \underset{\Vtilde{\theta}}{\arg\min}\; \mathcal{L}_\gamma(\V{\theta}_{m+1},\Vtilde{\theta}, \V{\omega}_{m}) \\
-\V{\omega}_{m+1} & \amp = \amp \V{\omega}_{m} + \gamma(\V{c} - \M{A}\V{\theta}_{m+1} - \M{B}\Vtilde{\theta}_{m+1}).
+\V{\phi}_{m+1} & \amp = \amp \underset{\V{\phi}}{\arg\min}\; \mathcal{L}_\gamma(\V{\phi},\Vtilde{\phi}_{m}, \V{\omega}_{m}) \\
+\Vtilde{\phi}_{m+1} & \amp = \amp \underset{\Vtilde{\phi}}{\arg\min}\; \mathcal{L}_\gamma(\V{\phi}_{m+1},\Vtilde{\phi}, \V{\omega}_{m}) \\
+\V{\omega}_{m+1} & \amp = \amp \V{\omega}_{m} + \gamma(\V{c} - \M{A}\V{\phi}_{m+1} - \M{B}\Vtilde{\phi}_{m+1}).
 \end{split}
@@
 the sensor response to pollutants, and high frequency random noise.
-These sensors are used to monitor pollutant concentrations at the fence lines of industrial facilities and detect time points when high concentrations are present.
+These sensors record measurements every second are used to monitor pollutant concentrations at the fence lines of industrial facilities and detect time points when high concentrations are present.
 Ideally, three co-located and time aligned sensors (as shown in \Fig{raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
+We first illustrate the difference between our \texttt{detrend\_eBIC} and the \texttt{qsreg} using just 2 hours of measurements so that the methods can be easily visualized and the \texttt{qsreg} method can be fit to the entire dataset \Sec{short-app}.
+We then compare the methods on an entire week of measurements, fitting the \texttt{qsreg} method to non-overlapping windows of the data \Sec{long-app}.
 
+\subsection{Short series of air quality measurements}
+\label{sec:short-app}
 \begin{figure}
@@
 
-We compare our \texttt{detrend\_eBIC} method with the \texttt{qsreg} method on a subset of the SPod data (n=6000) since the \texttt{qsreg} method cannot handle all 24 hours simultaneously.
+We compare our \texttt{detrend\_eBIC} method with the \texttt{qsreg} method on a subset of the SPod data (n=7200) since the \texttt{qsreg} method cannot handle all 24 hours simultaneously.
 We estimate the baseline trend using the 10\textsuperscript{th} and 15\textsuperscript{th} quantiles and compare three thresholds for classifying signal.
@@
 We then calculate the threshold using the median plus the MAD multiple of the corrected series and classify the corrected series based on the threshold.
-\Fig{short-trends} shows an example of the estimated baseline fit for each method, while \Fig{rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 3.
+\Fig{short-trends} shows an example of the estimated baseline fit for each method.
+The \texttt{qsreg} method absorbs more of the peaks while the \texttt{detrendr} method results in a smoother baseline estimate.
+\Fig{rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 5.
+The largest peaks at 14:10 are easily identified as signal, but good baseline estimates also enable proper classification of the smaller peaks like the one at 15:12.
 
@@
 	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
-	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile. Horizontal dashed lines represent the thresholds.}
+	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile. Horizontal dashed lines represent the thresholds using a MAD multiple of 5.}
 	\label{fig:rugplot}
@@
 
+\subsection{Long series of air quality measurements}
+\label{sec:long-app}
 Our windowed \texttt{detrend\_eBIC} method was used to removed the baseline drift from the total dataset consisting of 86,401 observations per node.

### 2019-03-26 Code/application_week.R

@@ -11,7 +11,7 @@ library(zoo)
 load_all("detrendr")
 rm(list=ls())
 
-i = 0 
+i = 0
 spod <- read.csv(sprintf("../SPod/SPod_week/SENTINEL Data_2017-03-0%d.csv",i), 
                  header=TRUE,  na.strings = "N/A")
 spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
@@ -26,18 +26,20 @@ spodPIDs$c <- na.locf(spodPIDs$c)
 spodPIDs$e <- na.locf(spodPIDs$e)
 
 spod_trends <- data.frame(time = spod$time)
-window_size <- 5000
-overlap <- 1000
+window_size <- 3600
+overlap <- 600
 max_iter <- 5
-tau <- c(0.1, 0.15)
+tau <- c(0.01, 0.05, 0.1)
 
 
 for (node in nodes){
-  result <- get_windows_BIC(y=spodPIDs[,node], tau, k=3, window_size, overlap,
-                            lambdaSeq = exp(seq(8,14,1)),
+  result0 <- get_windows_BIC(y=spodPIDs, tau, k=3, window_size, overlap,
+                            lambdaSeq = c(400, 800, 1000, 1600, 3200, 5000),
                             df_tol = 1e-9,
                             gamma = 1,
                             plot_lambda = TRUE,
+                            eps_abs = 0.02, 
+                            eps_rel = 1e-3,
                             solver = NULL,
                             criteria = "eBIC", 
                             max_iter = max_iter, 
@@ -47,7 +49,7 @@ for (node in nodes){
   names(spod_trends)[(ncol(spod_trends)-length(tau)+1):ncol(spod_trends)] <-
     paste(node, tau, sep = "_")
   save(spod_trends,  spodPIDs, 
-       file = sprintf("../SPod/SPod_week/trends_2017-03-0%d.RData",i))
+       file = sprintf("../SPod/SPod_week/trends_%s_2017-03-0%d.RData",node,i))
 }
 
 

### 2019-03-26 Code/plot_week.R

@@ -1,0 +1,24 @@
+library(devtools)
+library(Hmisc)
+library(fields)
+library(gurobi)
+library(zoo)
+load_all("detrendr")
+rm(list=ls())
+spodAll <- {}
+for (i in 3:9){
+  spod <- read.csv(sprintf("../SPod/SPod_week/SENTINEL Data_2017-03-0%d.csv",i), 
+                 header=TRUE,  na.strings = "N/A")
+  spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
+                                   format= "%m/%d/%Y %H:%M:%S")) 
+  nodes <- c("c", "e")
+  spodPIDs <- as.data.frame(spod[, paste(nodes, "SPOD.PID..V.", sep=".")])
+  names(spodPIDs) <- nodes
+  spodAll <- bind_rows(spodAll, spodPIDs)
+}
+
+plot(spodAll$c, type="l")
+plot(spodAll$e, type="l")
+
+min(spodAll$c, na.rm=TRUE)
+min(spodAll$e, na.rm=TRUE)

### 2019-03-28 Manuscript/detrendify_aoas.tex
working on week-long application

--- 2019-03-26 Manuscript/detrendify_aoas.tex
@@
 The primal residual at the $m$th iteration,
-%	\begin{eqnarray*}
-%	r_p^{(q)} & = & \sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert\theta_{j,m}^{(q)} - \M{A}r{\theta}_{j,m}^{(q)}\rVert_2^2},
-%	\end{eqnarray*}
+
 \begin{eqnarray*}
@@
 represents the difference between the trend values in the windows and the consensus trend value. The dual residual at the $m$th iteration,
-%	\begin{eqnarray*}
-%	r_d^{(q)} & = & \gamma\sqrt{\sum_{m=1}^M \sum_{j=1}^J\lVert\M{A}r{\theta}_{j,m}^{(q)} - \M{A}r{\theta}_{j,m}^{(q-1)}\rVert_2^2},
-%	\end{eqnarray*}
+
 \begin{eqnarray*}
@@
 represents the change in the consensus variable from one iterate to the next. The algorithm is stopped when
-%	\begin{eqnarray*}
-%		r_p^{(q)} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\underset{w}{\max}\left[\max
-%		\left(\sqrt{\sum_{j=1}^J \lVert\theta_{j,m}^{(q)}\rVert_2^2}, \sqrt{\sum_{j=1}^J \lVert \M{A}r{\theta}_{j,m}^{(q)} \rVert_2^2} \right )\right]\\
-%		r_d^{(q)} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\sqrt{\sum_{m=1}^M\sum_{j=1}^J\lVert \omega_{j,m}^{(q)}\rVert_2^2}
-%	\end{eqnarray*}
+
 \begin{eqnarray*}
@@
 In this simulation study, our proposed method \texttt{detrend\_eBIC} method substantially outperforms the others.
-The \texttt{detrend\_cross} method, which is the \texttt{detrend\_eBIC} method fit without the non-crossing constraints, performs similarly for larger quantiles and larger datasets.
-However, \texttt{detrend\_cross} produces significantly worse estimates for more extreme quantiles ($\tau = 0.01$) and smaller data sets ($\tau = 0.05$ and $n = 500$).
+The \texttt{detrend\_Xing} method, which is the \texttt{detrend\_eBIC} method fit without the non-crossing constraints, performs similarly for larger quantiles and larger datasets.
+However, \texttt{detrend\_Xing} produces significantly worse estimates for more extreme quantiles ($\tau = 0.01$) and smaller data sets ($\tau = 0.05$ and $n = 500$).
 These results indicate that even when a single quantile is of interest, simultaneously fitting nearby quantiles and utilizing the non-crossing constraint can improve estimates when data is sparse by using information from nearby quantiles.

### 2019-03-28 Code/Fig_peaks_metrics.R

@@ -24,10 +24,10 @@ trend_detrend$qsreg2 <- trend_qsreg[,2]
 
 ggplot(df, aes(x=x, y=y)) + 
   geom_line(col="grey") +
-  geom_line(data=trend_detrend, aes(y=q, x=x, col = "true 0.1"), lwd = 1.5) +
+  #geom_line(data=trend_detrend, aes(y=q, x=x, col = "true 0.1"), lwd = 1.5) +
   geom_line(data=trend_detrend, aes(y=q2, x=x, col = "true 0.05"), lwd = 1.5) +
-  geom_line(data=trend_detrend, aes(y=V3, x=x, col = "detrendr 0.1")) +
-  geom_line(data=trend_detrend, aes(y=qsreg, x=x, col = "qsreg 0.1")) + 
+  #geom_line(data=trend_detrend, aes(y=V3, x=x, col = "detrendr 0.1")) +
+  #geom_line(data=trend_detrend, aes(y=qsreg, x=x, col = "qsreg 0.1")) + 
   geom_line(data=trend_detrend, aes(y=V2, x=x, col = "detrendr 0.05")) +
   geom_line(data=trend_detrend, aes(y=qsreg2, x=x, col = "qsreg 0.05")) +
   theme_bw() + 
@@ -41,7 +41,7 @@ tau <- c(0.01, 0.05, 0.1)
 nSim <- 100
 colPal <- rev(c('#762a83','#9970ab','#c2a5cf', '#d9f0d3',
                 '#a6dba0','#5aae61','#1b7837'))
-methods <- c("detrend_eBIC", "detrend_SIC", "detrend_valid", "detrend_cross",
+methods <- c("detrend_eBIC", "detrend_SIC", "detrend_valid", "detrend_Xing",
              "rqss", "npqw", "qsreg") 
 MSEs <- as.data.frame(matrix(NA, nrow = nSim*length(methods), 
                              ncol = length(tau)+3))

### 2019-03-28 Code/Sim_qsreg.R

@@ -4,17 +4,18 @@ tau <- c(0.01, 0.05, 0.25, 0.5, .75, 0.95, 0.99)
 tau <- c(0.01, 0.05, 0.1)
 simDesigns <- c("gaus", "mixednorm", "shapebeta", "peaks")
 simDesigns <- "peaks"
-i = 0
+for (i in 1:100){
 for (n in c(500,1000, 2000, 4000)){
 for(simDesign in simDesigns){                
     load(sprintf("../SimData/%s_n_%i_sim%03.0f.RData", simDesign, n, i))
     trend <- matrix(NA, n, length(tau))
     for (j in 1:length(tau)){
-      fit_qsreg <- qsreg(df$x, df$y, maxit.cv = 50, alpha=tau[j], hmin = -28)
+      fit_qsreg <- qsreg(df$x, df$y, maxit.cv = 50, alpha=tau[j], hmin = -38)
       trend[,j] <- predict(fit_qsreg)    
     }
     save(trend, 
          file = sprintf("../SimResults/qsreg/%s_n_%i_sim%03.0f.RData", 
                         simDesign, n, i))
+}
 }
 }

### 2019-03-28 Code/application_413.R

@@ -1,0 +1,86 @@
+library(fields)
+library(tidyverse)
+library(devtools)
+library(Hmisc)
+library(caret)
+library(aricode)
+library(mcclust)
+library(lubridate)
+load_all("detrendr")
+rm(list=ls())
+source("application_functions.R")
+colPal <- c('#1b7837', '#762a83')
+
+
+load(sprintf("../SPod/SPod_week/trends_e_2017-04-13.RData"))
+load(sprintf("../SPod/SPod_week/qsreg_trends_2017-04-13.RData"))
+
+# spodFig <- data.frame(time = spodPIDs$time,
+#                      # detrendr = spod_trends$e_0.1,
+#                       qsreg = qsreg_trends$e_0.01) %>%
+#   gather("type","value", -time)
+# 
+# ggplot(spodFig, aes(x=time, y=value)) +
+#   geom_line(data=spodPIDs, aes(y=e), col="darkgrey")+
+#   geom_line(aes(col=type, group=type)) +
+#   theme_bw() +
+#   scale_color_manual(breaks = c("detrendr", "qsreg"),
+#                      values = c(colPal)) +
+#   labs(col="", x="", y="PID") +
+# xlim(c(spodPIDs$time[1], spodPIDs$time[10000])) +
+#   ylim(c(1, 2))
+
+################################################################################
+  detrendr_trends <- spod_trends 
+  methods <- "qsreg" #c("detrendr", "qsreg")
+  metric_df <- tibble(metric=NA, method=NA, tau=NA, crit=NA, metric_type=NA, 
+                      h=NA)
+  i <- 1
+  metrics <- c("VI", "confusion")
+  for (h in 0:1){
+    start_ind <- h*43200+1
+    end_ind <- min((h+1)*43200, nrow(spodPIDs))
+    for (method in methods){
+      trends <- get(paste(method, "trends", sep = "_"))[start_ind:end_ind, ]
+      for (j in 1:length(tau)){
+        for (crit in c(5, 6)){
+          signal <- get_spod_signal(tau[j], trends, 
+                                    spodPIDs[start_ind:end_ind, ], 
+                                    crit)
+          
+          for (metric in metrics){
+            if (any(signal > 0)){
+              if (metric == "confusion"){
+                metric_df$metric[i] <- I(list(get_confusion(signal, nodes)))
+              } else if (metric == "VI") {
+                metric_df$metric[i] <- I(list(get_VI(signal, nodes)))
[…]

### 2019-03-28 Code/week_stats.R

@@ -11,7 +11,7 @@ rm(list=ls())
 source("application_functions.R")
 colPal <- c('#1b7837', '#762a83')
 
-tau <- c(0.1, 0.15)
+tau <- c(0.01, 0.05, .1)
 metric_all <- {}
 for (d in 2:8){
   print(d)
@@ -50,14 +50,14 @@ ggplot(spodFig, aes(x=time, y=value)) +
   metric_df <- tibble(metric=NA, method=NA, tau=NA, crit=NA, metric_type=NA, 
                       h=NA, d=NA)
   i <- 1
-  metrics <- c("confusion",  "VI", "pos")
-for (h in 0:2){
-  start_ind <- h*28800+1
-  end_ind <- min((h+1)*28800, nrow(spodPIDs))
+  metrics <- c("VI", "pos")
+for (h in 0:1){
+  start_ind <- h*43200+1
+  end_ind <- min((h+1)*43200, nrow(spodPIDs))
   for (method in methods){
     trends <- get(paste(method, "trends", sep = "_"))[start_ind:end_ind, ]
     for (j in 1:length(tau)){
-      for (crit in c(3, 4, 5)){
+      for (crit in c(5, 6)){
         signal <- get_spod_signal(tau[j], trends, 
                                   spodPIDs[start_ind:end_ind, ], 
                                   crit)

### 2019-03-30 Manuscript/detrendify_aoas.tex
new threshold choice

@@
 Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
-To reduce cost and power consumption of the SPod, the temperature and relative humidity of the air presented to the photoionization detectors (PIDs) was not controlled and as a result the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
+Due to changes in temperature and relative humidity the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
 Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
@@
 
+\Alg{admm} has the following convergence guarantees.
+
+\begin{proposition}
+	\label{prop:convergence}
+	Let $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$ denote the $m$th collection of iterates generated by \Alg{admm}. Then (i)
+	$\lVert \Mn{\Theta}{w}_m - \Mn{U}{w}\M{\Theta}_m \rVert_{\text{F}} \rightarrow 0$ and (ii) $p_m \rightarrow p^\star$, where $p^\star$ is the optimal objective function value of
+	\Eqn{quantile_windows} and $p_m$ is the objective function value of \Eqn{quantile_windows} evaluated at $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$.
+\end{proposition}
+
+The proof of \Prop{convergence} is a straightforward application of the convergence result presented in Section 3.2 of \cite{boyd2011distributed}.
+
+\Eric{Eric}{Add computational complexity of quadratic program and discuss the savings compared with solving a single large LP}.
+
 \begin{figure}
@@
 %% ----------------------------------------------------------------------
-%% Computational Complexity and Convergence
-%% ----------------------------------------------------------------------
-\subsection{Computational Complexity and Convergence}
-
-\Eric{Eric}{Add computational complexity of quadratic program and discuss the savings compared with solving a single large LP}.
-
-\Alg{admm} has the following convergence guarantees.
-
-\begin{proposition}
-	\label{prop:convergence}
-	Let $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$ denote the $m$th collection of iterates generated by \Alg{admm}. Then (i)
-	$\lVert \Mn{\Theta}{w}_m - \Mn{U}{w}\M{\Theta}_m \rVert_{\text{F}} \rightarrow 0$ and (ii) $p_m \rightarrow p^\star$, where $p^\star$ is the optimal objective function value of
-	\Eqn{quantile_windows} and $p_m$ is the objective function value of \Eqn{quantile_windows} evaluated at $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$.
-\end{proposition}
-
-The proof of \Prop{convergence} is a straightforward application of the convergence result presented in Section 3.2 of \cite{boyd2011distributed}.
-
-%% ----------------------------------------------------------------------
 %% Model Selection
@@
 
-An important practical issue in baseline estimation is the choice of regularization parameter $\lambda$, which controls the degree of smoothness in $\V{\theta}$.
+An important practical issue in baseline estimation is the choice of the regularization parameter $\lambda$, which controls the degree of smoothness in $\V{\theta}$.
 In this section, we introduce four methods for choosing $\lambda$.
 The first is a validation based approach; the latter three are based on information criteria.
+Each of the criteria we compare is calculated for a single quantile ($\tau_j$).
+Rather than combine results over quantiles, we allow the value of $\lambda$ to vary by quantile resulting in $\V{\lambda} = \{\lambda_1, ..., \lambda_J\}$.
+To choose the best value for each $\lambda_j$, we first estimate all of the quantile trends using $\V{\lambda} = \lambda\mathbf{1}$ over a grid of values for $\lambda$.
+We then determine the $\lambda_j$ that maximizes the criteria chosen evaluated using $\tau_j$.
+Finally, we re-estimate the non-crossing trends with the optimal values for $\lambda_j$.
+A more thorough approach would involve fitting the model on a J dimensional grid of values for $\V{\lambda}$ but this is computationally infeasible.
 
@@
 &&\underset{\M{\Theta} \in \mathcal{C}}{\text{minimize}}\; \sum_{j=1}^J \left [\tilde{\rho}_{\tau_j}(\V{y} - \V{\theta}_{j}) +
-\lambda \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ],
+\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ],
 \end{eqnarray}
 which can be solved via \Alg{admm} with trivial modification to the quadratic program sub-problems.
-We fit a sequence of values for $\lambda$ setting $\lambda_j = \lambda$ for all $j$.
-Then for each quantile level $j$ we select the $\lambda_j$ that minimizes the hold-out prediction error quantified by $\breve{\rho}_{\tau_j}(\V{y} - \Vhat{\theta}_j(\lambda))$ where $\Vhat{\theta}_j(\lambda)$ is the solution to \Eqn{validate} and
+For each quantile level $j$, we select the $\lambda_j$ that minimizes the hold-out prediction error quantified by $\breve{\rho}_{\tau_j}(\V{y} - \Vhat{\theta}_j(\lambda_j))$ where $\Vhat{\theta}_j(\lambda_j)$ is the solution to \Eqn{validate} and
 \begin{eqnarray*}
@@
 	\label{eq:SIC}
-	\mbox{SIC}(p_{\lambda}) & = & \log\left[\frac{1}{n}\rho_{\tau}(\V{y} - \V{\theta})\right] + \frac{1}{2n}p_{\lambda}\log n.
+	\mbox{SIC}(p_{\lambda}, \tau_j) & = & \log\left[\frac{1}{n}\rho_{\tau_j}(\V{y} - \V{\theta}_j)\right] + \frac{1}{2n}p_{\lambda}\log n.
 \end{eqnarray*}
 where $p_{\lambda} = \sum_i \One(\VE{y}{i} = \widehat{\theta}_i)$ is the number of interpolated points, which can be thought of as active knots.
-Equivalently, $p_{\lambda}$ can be substituted with the number of non-zero components in $\M{D}^{(k+1)}\V{\theta}$ which we denote $\nu$ and have found to be more numerically stable.
+Equivalently, $p_{\lambda}$ can be substituted with the number of non-zero components in $\M{D}^{(k+1)}\Vhat{\theta}_j$ which we denote $\nu$ and have found to be more numerically stable.
 The SIC is based on the traditional Bayesian Information Criterion (BIC) which is given by
@@
 \begin{eqnarray*}
-	\mbox{BIC}(\nu) = 2\frac{1}{\sigma}\rho_{\tau}(\V{y}-\Vhat{\theta}) + \nu\log n
+	\mbox{BIC}(\nu, \tau_j) = 2\frac{1}{\sigma}\rho_{\tau_j}(\V{y}-\Vhat{\theta}_j) + \nu\log n
 \end{eqnarray*}
@@
 
-where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod. We used this criteria with $\gamma = 1$, and $P=n-k-1$.  
-As in the validation method, to chose $\V{\lambda}$ using the information criteria we fit a sequence of values for $\lambda$ setting $\lambda_j = \lambda$ for all $j$.
-For each quantile level $j$, we select the $\lambda_j$ that minimizes the SIC, BIC, or eBIC value evaluated using $\tau_j$ and $\V{\theta}_j$.
+where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod.
+We used this criteria with $\gamma = 1$, and $P=n-k-1$.
 In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
@@
 We first illustrate the difference between our \texttt{detrend\_eBIC} and the \texttt{qsreg} using just 2 hours of measurements so that the methods can be easily visualized and the \texttt{qsreg} method can be fit to the entire dataset \Sec{short-app}.
-We then compare the methods on an entire week of measurements, fitting the \texttt{qsreg} method to non-overlapping windows of the data \Sec{long-app}.
+We then compare the methods on the complete day shown in \Fig{raw_spod}, fitting the \texttt{qsreg} method to 2 hour non-overlapping windows of the data.
+We focus on this day because all three sensors were functioning properly.
+Finally, we examine an entire week of measurements for the two sensors for which data passed was available \Sec{long-app}.
 

### 2019-03-30 Code/fit_0302.R

@@ -1,0 +1,107 @@
+################################################################################
+# Fit quantile trends on week of data
+# Halley Brantley
+################################################################################
+library(tidyverse)
+library(devtools)
+library(Hmisc)
+library(fields)
+library(gurobi)
+library(zoo)
+load_all("detrendr")
+source("application_functions.R")
+rm(list=ls())
+
+spod <- read.csv("../SPod/SPod_week/SENTINEL Data_2017-03-02.csv",
+                 header=TRUE,  na.strings = "N/A")
+load(sprintf("../SPod/SPod_week/trends_e_2017-03-02.RData"))
+load(sprintf("../SPod/SPod_week/qsreg_trends_2017-03-02.RData"))
+
+spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
+                                 format= "%m/%d/%Y %H:%M:%S")) 
+nodes <- c("c", "e")
+spodPIDs <- as.data.frame(spod[, paste(nodes, "SPOD.PID..V.", sep=".")])
+names(spodPIDs) <- nodes
+spodPIDs[,nodes] <- spodPIDs[,nodes]/1000
+spodPIDs$time <- spod$time
+spodPIDs[, nodes[1]] <- na.locf(spodPIDs[,nodes[1]])
+spodPIDs[, nodes[2]] <- na.locf(spodPIDs[,nodes[2]])
+spodPIDs <- spodPIDs[1:nrow(qsreg_trends), ]
+spod_trends <- spod_trends[1:nrow(qsreg_trends), ]
+tau <- c(0.01, 0.05, 0.1)
+
+plot(spodPIDs$c, type="l")
+lines(spod_trends$c_0.05, col="red")
+lines(qsreg_trends$c_0.05, col="blue")
+
+plot(spodPIDs$e, type="l")
+lines(spod_trends$e_0.05, col="red")
+lines(qsreg_trends$e_0.05, col="blue")
+
+peaks_qsreg <- select(spodPIDs, -time) - 
+  select(qsreg_trends, ends_with(paste(0.05)))
+
+peaks_detrend <- select(spodPIDs, -time) - 
+  select(spod_trends, ends_with(paste(0.05)))
+
+plot(c~e, peaks_qsreg)
+plot(c~e, peaks_detrend)
+cor(peaks_detrend[,c("c", "e")], method="spearman")
[…]

### 2019-03-30 Code/fit_0413.R

@@ -1,0 +1,168 @@
+################################################################################
+# Fit quantile trends on week of data
+# Halley Brantley
+################################################################################
+library(tidyverse)
+library(devtools)
+library(Hmisc)
+library(fields)
+library(gurobi)
+library(zoo)
+load_all("detrendr")
+rm(list=ls())
+
+spod <- read.csv("../SPod/SPod_week/SENTINEL Data_2017-04-13.csv",
+                 header=TRUE,  na.strings = "N/A")
+
+spod$time <- as.POSIXct(strptime(as.character(spod$TimeStamp), 
+                                 format= "%m/%d/%Y %H:%M:%S")) 
+nodes <- c("c", "e")
+spodPIDs <- as.data.frame(spod[, paste(nodes, "SPOD.PID..V.", sep=".")])
+names(spodPIDs) <- nodes
+spodPIDs[,nodes] <- spodPIDs[,nodes]/1000
+spodPIDs$time <- spod$time
+spodPIDs[, nodes[1]] <- na.locf(spodPIDs[,nodes[1]])
+spodPIDs[, nodes[2]] <- na.locf(spodPIDs[,nodes[2]])
+
+plot(c~e, spodPIDs[1:7200,])
+tau <- c(0.01, 0.05, 0.1)
+j <- 3
+ind_start <- (j-1)*7200 + 1
+ind_end <- min(nrow(spodPIDs), j*7200)
+x <- seq(ind_start, ind_end, 1)
+n <- length(x)
+
+qsreg_trends <- data.frame(time = spodPIDs$time)
+
+for (j in 1:12){
+  ind_start <- (j-1)*7200 + 1
+  ind_end <- min(nrow(spodPIDs), j*7200)
+  x <- seq(ind_start, ind_end, 1)
+  n <- length(x)
+  trends <- data.frame(time = spodPIDs$time[ind_start:ind_end])
+  for (node in nodes){
+    trend <- matrix(NA, n, length(tau))
+    for (i in 1:length(tau)){
+      fit_qsreg <- qsreg(x, spodPIDs[ind_start:ind_end,node], 
+                         maxit.cv = 50, 
+                         alpha=tau[i], hmin = -12, hmax = NA)
+      trend[,i] <- predict(fit_qsreg)   
+    }
+    trends <- cbind(trends, as.data.frame(trend))
+    names(trends)[(ncol(trends)-(length(tau)-1)):ncol(trends)] <-
[…]

### 2019-03-30 Code/week_stats.R

@@ -10,37 +10,49 @@ load_all("detrendr")
 rm(list=ls())
 source("application_functions.R")
 colPal <- c('#1b7837', '#762a83')
-
+nodes <- c("c", "e")
 tau <- c(0.01, 0.05, .1)
 metric_all <- {}
 for (d in 2:8){
   print(d)
-  load(sprintf("../SPod/SPod_week/trends_2017-03-0%d.RData",d))
+  load(sprintf("../SPod/SPod_week/trends_e_2017-03-0%d.RData",d))
   load(sprintf("../SPod/SPod_week/qsreg_trends_2017-03-0%d.RData",d))
-# spodPeaks <- select(spodPIDs, -time) - select(qsreg_trends,
-#                                               contains(paste(0.15)))
+# spodPeaks <- dplyr::select(spodPIDs, -time) - dplyr::select(spod_trends,
+#                                               contains(paste(0.05)))
 spodFig <- data.frame(time = spodPIDs$time,
-                      detrendr = spod_trends$e_0.1,
-                      qsreg = qsreg_trends$e_0.1) %>%
+                      detrendr = spod_trends$c_0.1,
+                      qsreg = qsreg_trends$c_0.1) %>%
   gather("type","value", -time)
 # 
-# spodFig2 <- data.frame(time = spodPIDs$time,
-#                       detrendr_e = spodPIDs$e - spod_trends$e_0.1,
-#                       qsreg_e = spodPIDs$e - qsreg_trends$e_0.1, 
-#                       detrendr_c = spodPIDs$c - spod_trends$c_0.1,
-#                       qsreg_c = spodPIDs$c - qsreg_trends$c_0.1) 
-#   gather("type","value", -time)
+spodFig2 <- data.frame(time = spodPIDs$time,
+                      detrendr_e = spodPIDs$e - spod_trends$e_0.05,
+                      qsreg_e = spodPIDs$e - qsreg_trends$e_0.05,
+                      detrendr_c = spodPIDs$c - spod_trends$c_0.05,
+                      qsreg_c = spodPIDs$c - qsreg_trends$c_0.05)
+%>%
+  gather("type","value", -time)
 # 
-# 
+#
+ggplot(spodFig2, aes(x=time, y=value)) +
+    geom_line(aes(col=type, group=type)) +
+    theme_bw() +
[…]

### 2019-04-01 Manuscript/detrendify_aoas.tex
added some discussion of complexity, working on application

@@
 
- As the size of the data increases, all methods require additional computation time and for our method eventually the matrix of constraints no longer fits into memory.
+ As the size of the data increases, all methods require additional computation time and at some point the penalty matrices will no longer fit into memory.
 In our application to air quality sensor data, measurements are recorded every second resulting in 86,400 observations per day.
@@
 
-\Eric{Eric}{Add computational complexity of quadratic program and discuss the savings compared with solving a single large LP}.
-
 \begin{figure}
@@
 
-Timing experiments illustrate the advantages of using our ADMM algorithm when the trend can be estimated with a single window.
+The quantile trend filtering problem for a single window is a linear program with $N\times J$ parameters (number of observations by number of quantiles), which can be solved in computational time proportional to $(NJ)^3$.
+By dividing the problem into four smaller problems we can theoretically solve each of the 4 sub-problems 4 times and still reduce the computation time required by a factor of 4.
+Consequently, solving a large problem using \Alg{admm} should require less computational time than simply using a linear program solver given small enough windows and convergence in few enough iterations.
+We demonstrate the advantages of \Alg{admm} empirically using timing experiments.
 For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below.

### 2019-04-01 Code/fit_0302.R

@@ -59,21 +59,6 @@ lines(peaks_qsreg$c, col="red")
 node <- "c"
 
 
-getThresh <- function(x){mean(x, na.rm=T)+4*sd(x, na.rm=T)}
-
-getCutoff <- function(peaks, node){
-  thresh0 <- getThresh(peaks[,node])
-  print(thresh0)
-  peaks[which(peaks[,node] > thresh0), node] <- NA
-  thresh <- getThresh(peaks[,node])
-  while(thresh0 - thresh > 0.001){
-    thresh0 <- thresh
-    print(thresh0)
-    peaks[which(peaks[,node] > thresh), node] <- NA
-    thresh <- getThresh(peaks[,node])
-  }
-  thresh
-}
 
 getCutoff(peaks_detrend, "e")
 

### 2019-04-02 Manuscript/detrendify_aoas.tex
new figures and discussion for application section

@@
 the sensor response to pollutants, and high frequency random noise.
-These sensors record measurements every second are used to monitor pollutant concentrations at the fence lines of industrial facilities and detect time points when high concentrations are present.
+These sensors record measurements every second and are used to monitor pollutant concentrations at the fence lines of industrial facilities.
+Time points with high concentrations are identified and compared with concurrent wind direction and speed.
 Ideally, three co-located and time aligned sensors (as shown in \Fig{raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
-We first illustrate the difference between our \texttt{detrend\_eBIC} and the \texttt{qsreg} using just 2 hours of measurements so that the methods can be easily visualized and the \texttt{qsreg} method can be fit to the entire dataset \Sec{short-app}.
-We then compare the methods on the complete day shown in \Fig{raw_spod}, fitting the \texttt{qsreg} method to 2 hour non-overlapping windows of the data.
-We focus on this day because all three sensors were functioning properly.
-Finally, we examine an entire week of measurements for the two sensors for which data passed was available \Sec{long-app}.
+We first illustrate the difference between our \texttt{detrend\_eBIC} method,  hereafter referred to as \texttt{detrendr}, and \texttt{qsreg} using a small subset of measurements from \Fig{raw_spod} (\Sec{short-app}).
+We then compare the methods on the complete day shown in \Fig{raw_spod}, estimating trends by applying the \texttt{qsreg} method to 2 hour non-overlapping windows of the data and \Alg{admm} to the entire day.
+We focus on this day because data from three sensors was available.
+Finally, we examine an entire week of measurements from two co-located sensors (\Sec{long-app}).
 
@@
 	\includegraphics[width = \linewidth]{Figures/short_trends.png}
-	\caption{Estimated 15\textsuperscript{th} quantile trends on subset of the data using \texttt{qsreg} and \texttt{detrend\_eBIC}.}
+	\caption{Estimated 5\textsuperscript{th} quantile trends for SPods a, b, and c, using \texttt{qsreg} and \texttt{detrendr}.
+SPod c contains some missing values that were interpolated before the trends were estimated.
+\texttt{qsreg} is more influenced by the signal component apparent on both nodes resulting in under-smoothing.}
 	\label{fig:short-trends}
@@
 
-We compare our \texttt{detrend\_eBIC} method with the \texttt{qsreg} method on a subset of the SPod data (n=7200) since the \texttt{qsreg} method cannot handle all 24 hours simultaneously.
-We estimate the baseline trend using the 10\textsuperscript{th} and 15\textsuperscript{th} quantiles and compare three thresholds for classifying signal.
-The thresholds are calculated using the median plus a multiple of the median absolute deviation \Eqn{MAD} of the detrended series.
-\begin{eqnarray}
-\label{eq:MAD}
-\mbox{MAD} = \mbox{median}\lVert y-\tilde{y}\rVert,
-\end{eqnarray}
-where $\tilde{y}$ is the median of $y$.
-Given a method, quantile level, and MAD multiple, we estimate the quantile trend for each of the sensors and subtract it from the observations.
-We then calculate the threshold using the median plus the MAD multiple of the corrected series and classify the corrected series based on the threshold.
-\Fig{short-trends} shows an example of the estimated baseline fit for each method.
-The \texttt{qsreg} method absorbs more of the peaks while the \texttt{detrendr} method results in a smoother baseline estimate.
-\Fig{rugplot} shows the series after subtracting the \texttt{detrend\_eBIC} estimate of the 15\textsuperscript{th} quantile and classifying the signal using a MAD multiple of 5.
-The largest peaks at 14:10 are easily identified as signal, but good baseline estimates also enable proper classification of the smaller peaks like the one at 15:12.
-
-\begin{figure}
-	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
-	\caption{Rugplot showing locations of signal after baseline removal using detrendr estimate of 15th quantile. Horizontal dashed lines represent the thresholds using a MAD multiple of 5.}
-	\label{fig:rugplot}
-\end{figure}
-
-Given the signal classifications for SPods a and b, $\delta^{(a)}_i \in \{0,1\}$ and $\delta^{(b)}_i\in\{0,1\}$, we want to compare the similarity between the two classifications.
-One metric for evaluating the distance between two classifications is the variation of information (VI):
+We compare our \texttt{detrendr} method with the \texttt{qsreg} method on a subset of the SPod data (n=7200) both to facilitate visualization and because the \texttt{qsreg} method cannot handle all 24 hours simultaneously.
+We estimate the baseline trend using $\tau = \{0.01, 0.05, 0.1\}$ and compare three thresholds for classifying the signal.
+The thresholds are calculated as the 90\textsuperscript{th}, 95\textsuperscript{th}, and 99\textsuperscript{th} quantiles of the de-trended series for each SPod.
+If there is signal present in the dataset, values above these thresholds should occur simultaneously on all three SPods.
+We do not use class-averaged accuracy to compare the signal classifications because we do not have a reference value to define as the ``true" signal.
+Instead, we compute the variation of information (VI) which compares the similarity between the two classifications.
+Given the signal classifications for SPods a and b, $\delta^{(a)}_i \in \{0,1\}$ and $\delta^{(b)}_i\in\{0,1\}$, the VI is defined as:
 \begin{eqnarray*}
@@
 The $\VI$ is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
-\Fig{vi} shows plots the VIs by method, quantile, and trend.
-In all cases, our \texttt{detrend} method results in classifications that are more similar than those from the \texttt{qsreg} method.
-\Fig{short-trends} and \Fig{rugplot} illustrate that our method results in a smoother baseline estimate which improves signal classification.
+
+\Fig{short-trends} shows the estimated 5\textsuperscript{th} quantile trends from each method for each SPod.
+The \texttt{detrendr} method results in a smoother baseline estimate while the \texttt{qsreg} method absorbs more of the peaks obscuring some of the signal.
+\Fig{rugplot} shows the series after subtracting the \texttt{detrendr} estimate of the 5\textsuperscript{th} quantile and classifying the signal using the 95\textsuperscript{th} quantile of the detrended data.
+The 90\textsuperscript{th} and 99\textsuperscript{th} thresholds are also shown for comparison in blue and orange, respectively.
+The largest peaks at 14:10 are easily identified as signal, but good baseline estimates also enable proper classification of the smaller peaks like the one at 15:12.
+The under-smoothing of the \texttt{qsreg} method results in less similar signal classifications and higher VI values for the 90\textsuperscript{th} and 95\textsuperscript{th} quantile thresholds (\Fig{vi}).
+However, when the 99\textsuperscript{th} threshold is used only the highest observations are classified as signal and the baseline estimation method isn't as important (\Fig{vi}).
+
+\begin{figure}
+	\includegraphics[width = \linewidth]{Figures/corrected_rugplot.png}
+	\caption{Rugplot showing locations of signal after baseline removal using the \texttt{detrendr} estimate of 5th quantile.
+Horizontal dashed lines represent the thresholds calculated using the 90\textsuperscript{th}, 95\textsuperscript{th}, and 99\textsuperscript{th} quantiles of the detrended data.
+The 95\textsuperscript{th} quantile (black) was used to classify the signal shown as vertical lines at the bottom of the plot.}
+	\label{fig:rugplot}
+\end{figure}
 
@@
 	\centering
-	\includegraphics[width = .9\linewidth]{Figures/VI_app_short.png}
-	\caption{Variation of Information between sensors by method (color), quantile (columns) and threshold MAD multiple (rows).}
+	\includegraphics[width = 1\linewidth]{Figures/VI_app_short.png}
+	\caption{Variation of Information between sensors by method (color), quantile (columns) and threshold (shape) for two hour time period. }
 	\label{fig:vi}
@@
 \label{sec:long-app}
-Our windowed \texttt{detrend\_eBIC} method was used to removed the baseline drift from the total dataset consisting of 86,401 observations per node.
-The VI scores for the full dataset were 0.36, 0.24, and 0.41 for sensors a and b, a and c, and b and c, respectively.
-The complete confusion matrices for the nodes using a MAD multiple of 5 for the threshold is given in Table 1.
+\Alg{admm} was used to removed the baseline drift from the full day of data  (\Fig{raw_spod}) consisting of 86,401 observations per SPod and compared to the series detrended using the \texttt{qsreg} trends estimated using non-overlapping 2 hour windows.
+As in the shorter illustration, the \texttt{detrendr} method results in generally higher VI scores than the \texttt{qsreg} method (\Fig{vi_day}).
+The \texttt{detrendr} method also results in better correlation in the detrended series as is illustrated in (\Fig{scatter_day}).
 
 \begin{figure}
-	\includegraphics[width = \linewidth]{Figures/corrected_data.png}
-	\caption{Low cost sensor data after drift removal using windowed detrend with eBIC.}
+	\centering
+	\includegraphics[width = 1\linewidth]{Figures/VI_app_long.png}
+	\caption{Variation of Information between sensors by method (color), quantile (columns) and threshold (shape) for full day.}
+	\label{fig:vi_day}
 \end{figure}
 
-\input{Tables/full_confusion_detrend.tex}
+\begin{figure}
+	\centering
+	\includegraphics[width = 0.3\linewidth]{Figures/scatter_day_raw.png}
+	\includegraphics[width = 0.3\linewidth]{Figures/scatter_day_qsreg.png}
+	\includegraphics[width = 0.3\linewidth]{Figures/scatter_day_detrendr.png}
+	\caption{SPod a versus SPod c before and after de-trending. }
+	\label{fig:scatter_day}
+\end{figure}
+
+Finally we estimated the quantile trends for 7 days of measurements of two co-located SPods.
+\Fig{vi_week} demonstrates the improvement in classification similarity when using \texttt{detrendr}.
+Each point represents a day of measurements and all points that fall below the dashed line have more similar classifications using \texttt{detrendr} compared to \texttt{qsreg}.
+
+\begin{figure}
+	\includegraphics[width = 0.7\linewidth]{Figures/VI_by_day.png}
+	\caption{Variation of Information (VI) for \texttt{detrendr} and \texttt{qsreg} by quantile trend and threshold.
+Each point represents a full day of data.
+The dashed line represents y=x.
+In most cases \texttt{detrendr} results in a lower VI than \texttt{qsreg}.}
+	\label{fig:vi_week}
+\end{figure}
 

### 2019-04-02 Code/Fig_app_day.R

@@ -1,0 +1,146 @@
+################################################################################
+# Figures for full day of data
+# Halley Brantley
+################################################################################
+library(tidyverse)
+library(devtools)
+library(Hmisc)
+rm(list=ls())
+source("application_functions.R")
+colPal <- c('#1b7837', '#762a83')
+tau <- c(0.01, 0.05, 0.1)
+nodes <- c("c", "d", "e")
+
+load(sprintf("../SPod/Results/trends_e_2017-04-13.RData"))
+load(sprintf("../SPod/Results/qsreg_trends_2017-04-13.RData"))
+
+spodRaw <- spodPIDs %>%
+  gather("node", "PID", -time)
+
+ggplot(spodRaw, aes(x=time, y=PID, col=node)) +
+  geom_line(alpha=0.5) +
+  theme_bw() +
+  scale_color_brewer(palette = "Set1", labels = c("a", "b", "c"))+
+  labs(col="SPod")
+
+ggsave("../Manuscript/Figures/uncorrected_data.png", width = 7, height = 2.5)
+################################################################################
+peaks_qsreg <- select(spodPIDs, -time) - 
+  select(qsreg_trends, ends_with(paste(tau[2])))
+
+peaks_detrend <- select(spodPIDs, -time) - 
+  select(spod_trends, ends_with(paste(tau[2])))
+
+ggplot(spodPIDs, aes(x=c, y=e)) + 
+  geom_point(alpha = 0.1) + 
+  theme_bw() + 
+  labs(x = "SPod a", y = "SPod c", title = "Raw Data")
+ggsave("../Manuscript/Figures/scatter_day_raw.png", width = 3, height = 3)
+
+ggplot(peaks_detrend, aes(x=c, y=e)) + 
+  geom_point(alpha = 0.1) + 
+  theme_bw() + 
+  labs(x = "SPod a", y = "SPod c", title = "detrendr")
+ggsave("../Manuscript/Figures/scatter_day_detrendr.png", width = 3, height = 3)
+
+ggplot(peaks_qsreg, aes(x=c, y=e)) + 
+  geom_point(alpha = 0.1) + 
+  theme_bw() + 
+  labs(x = "SPod a", y = "SPod c", title = "qsreg")
+ggsave("../Manuscript/Figures/scatter_day_qsreg.png", width = 3, height = 3)
[…]

### 2019-04-02 Code/Fig_app_short.R

@@ -10,37 +10,48 @@ load_all("detrendr")
 rm(list=ls())
 load("../SPod/trends_short.RData")
 source("application_functions.R")
-colPal <- c('#1b7837', '#762a83')
-nodes <- c("c", "d", "e")
-
+colPal <- c('#1b7837', '#762a83', "darkgrey")
+nodes <- c("d", "e")
+missInd <- which(is.na(spodPIDs$d))
+tau <- c(0.01, 0.05, 0.1)
+detrendr_trends[missInd, paste("d", tau, sep="_")] <- NA
+qsreg_trends[missInd, paste("d", tau, sep="_")] <- NA
 spodPeaks <- select(spodPIDs, -time) - select(detrendr_trends,
                                               contains(paste(0.05)))
-spodFig <- data.frame(time = spodPIDs$time, 
-                      detrendr = detrendr_trends$c_0.05, 
-                      qsreg = qsreg_trends$c_0.05) %>%
-  gather("type","value", -time)
+spodFig <- data.frame(time = c(spodPIDs$time, spodPIDs$time, spodPIDs$time), 
+                      raw = c(spodPIDs$c, spodPIDs$d, spodPIDs$e),
+                      detrendr = c(detrendr_trends$c_0.05, 
+                                   detrendr_trends$d_0.05, 
+                                   detrendr_trends$e_0.05), 
+                      qsreg = c(qsreg_trends$c_0.05, 
+                                qsreg_trends$d_0.05, 
+                                qsreg_trends$e_0.05), 
+                      node = c(rep("a", nrow(spodPIDs)), 
+                               rep("b", nrow(spodPIDs)), 
+                               rep("c", nrow(spodPIDs)))) %>%
+  gather("type","value", -c(time, node)) %>%
+  arrange(time, rev(type))
 
 ggplot(spodFig, aes(x=time, y=value)) + 
-  geom_line(data=spodPIDs, aes(y=c), col="darkgrey")+
-  geom_line(aes(col=type, group=type)) + 
+  geom_line(aes(col=type, group=type), alpha=0.5) + 
   theme_bw() +
-  scale_color_manual(breaks = c("detrendr", "qsreg"), 
+  facet_grid(node~., scales="free")+
+  scale_color_manual(breaks = c("raw", "detrendr", "qsreg"), 
[…]

### 2019-04-02 Code/application_week_qsreg.R

@@ -12,7 +12,7 @@ load_all("detrendr")
 rm(list=ls())
 tau <- c(0.01, 0.05, 0.1)
 
-for (d in 3:5){
+for (d in 6:8){
   if (d == 1){
     nodes <- c("c", "d", "e")
     spod <- read.csv("../SPod/SPod_week/SENTINEL Data_2017-04-13.csv", 
@@ -61,7 +61,7 @@ for (d in 3:5){
   }
   if (d == 1){
     save(qsreg_trends, tau, nodes, spodPIDs,
-        file = "../SPod/SPod_week/qsreg_trends_2017-04-13.RData")
+        file = "../SPod/Results/qsreg_trends_2017-04-13.RData")
   } else {
     save(qsreg_trends, tau, nodes, spodPIDs,
       file = sprintf("../SPod/Results/qsreg_trends_2018-06-%d.RData",d+13))

### 2019-04-02 Code/detrendr/R/get_trend_BIC.R

@@ -24,7 +24,7 @@
 #' plot(trend$theta[,1]~x, type="l")  
 #' @export
 get_trend_BIC <- function(y, tau, k,
-                       lambdaSeq = exp(seq(0, 1.5, length.out=20)), 
+                       lambdaSeq = exp(seq(0, 14, 1)), 
                        df_tol = 1e-9, 
                        gamma = 1,
                        plot_lambda = FALSE, 

### 2019-04-02 Code/Fig_app_short.R
Manuscript edits and updated figures, draft ready for comments

@@ -11,7 +11,7 @@ rm(list=ls())
 load("../SPod/trends_short.RData")
 source("application_functions.R")
 colPal <- c('#1b7837', '#762a83', "darkgrey")
-nodes <- c("d", "e")
+nodes <- c("c", "d", "e")
 missInd <- which(is.na(spodPIDs$d))
 tau <- c(0.01, 0.05, 0.1)
 detrendr_trends[missInd, paste("d", tau, sep="_")] <- NA

### 2019-04-02 Code/application_day_qsreg.R

@@ -12,9 +12,9 @@ load("../SPod/spodPIDs.RData")
 nodes <- c("c", "d", "e")
 for (node in nodes){
   missInd <- which(is.na(spodPIDs[,node]))
-  spodPIDs[,node] <- na.locf(spodPIDs[,node]) 
+  spodPIDs[,node] <- na.approx(spodPIDs[,node]) 
   spodPIDs[missInd, node] <- spodPIDs[missInd, node] + 
-    rnorm(length(missInd), 0, .001)
+    rnorm(length(missInd), 0, .002)
 }  
 qsreg_trends <- data.frame(time = spodPIDs$time)
 
@@ -29,7 +29,7 @@ for (j in 1:12){
     for (i in 1:length(tau)){
       fit_qsreg <- qsreg(x, spodPIDs[ind_start:ind_end,node], 
                          maxit.cv = 50, 
-                         alpha=tau[i], hmin = -12, hmax = NA)
+                         alpha=tau[i], hmin = -14, hmax = NA)
       trend[,i] <- predict(fit_qsreg)   
     }
     trends <- cbind(trends, as.data.frame(trend))

### 2019-04-02 Code/application_short.R

@@ -28,9 +28,11 @@ x <- seq(1, nrow(spodPIDs), 1)
 for (node in nodes){
   missID <- which(is.na(spodPIDs[, node]))
   spodPIDs[,node] <- na.approx(spodPIDs[,node], na.rm=FALSE)
-  spodPIDs[missID, node] <- spodPIDs[missID, node]  + rnorm(length(missID), 0, .002)
+  spodPIDs[missID, node] <- spodPIDs[missID, node]  +
+    rnorm(length(missID), 0, .002)
   result <- get_trend_BIC(spodPIDs[, node], tau, k,
                             df_tol = 1e-9,
+                            lambdaSeq= exp(seq(5, 10, 1)),
                             gamma = 1,
                             plot_lambda = TRUE,
                             solver = "gurobi",
@@ -42,9 +44,9 @@ for (node in nodes){
 
   trend <- matrix(NA, n, length(tau))
   for (j in 1:length(tau)){
-    fit_qsreg <- qsreg(x, spodPIDs[,node], maxit.cv = 50, 
-                       alpha=tau[j], hmin = -12, hmax = NA)
-    trend[,j] <- predict(fit_qsreg)    
+    fit_qsreg <- qsreg(x, spodPIDs[,node], maxit.cv = 50,
+                       alpha=tau[j], hmin = -16, hmax = NA)
+    trend[,j] <- predict(fit_qsreg)
   }
   qsreg_trends <- cbind(qsreg_trends, as.data.frame(trend))
   names(qsreg_trends)[(ncol(qsreg_trends)-(length(tau)-1)):ncol(qsreg_trends)] <-
@@ -55,3 +57,4 @@ for (node in nodes){
 save(spodPIDs, detrendr_trends, qsreg_trends, tau,
      file = "../SPod/trends_short.RData")
 
+load("../SPod/trends_short.RData")

### 2019-04-03 Code/application_day.R
update plots for full day

@@ -37,7 +37,7 @@ tau <- c(0.01, 0.05, 0.1)
 k <- 3
 spod_trends <- data.frame(time = spod$time)
 
-for (node in c("c")){
+for (node in c("c", "d", "e")){
   missID <- which(is.na(spodPIDs[, node]))
   spodPIDs[,node] <- na.approx(spodPIDs[,node], na.rm=FALSE)
   spodPIDs[missID, node] <- spodPIDs[missID, node]  + 
@@ -46,7 +46,7 @@ for (node in c("c")){
   names(spodNode)[2] <- c("pid")
   result <- get_windows_BIC(spodNode$pid, tau, k, window_size, overlap,
                           df_tol = 1e-9,
-                          lambdaSeq= exp(seq(8, 12, 1)),
+                          lambdaSeq= exp(seq(8, 13, 1)),
                           gamma = 1,
                           plot_lambda = TRUE,
                           solver = NULL,

### 2019-04-04 Manuscript/detrendify_aoas.tex
Manuscript edits and updated figures, draft ready for comments
update plots for full day
added another sentance to application section

@@
 The \texttt{detrendr} method also results in better correlation in the detrended series as is illustrated in (\Fig{scatter_day}).
+The Spearman correlation coefficients for SPods a and b, SPods a and c, and SPods b and c after removing the 5\textsuperscript{th} quantile trend using \texttt{detrendr} were 0.37, 0.75, and 0.43, compared with 0.07, 0.24, and 0.16 using \texttt{qsreg}.
+The noise variance was higher for SPod b than for SPods a and c resulting in lower correlation and higher VI values for the ab and bc metrics compared with ac.
 

### 2019-04-08 Manuscript/detrendify_aoas.tex

@@
 As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of methods.
-We want to estimated multiple quantiles simultaneously and to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$ where $Q$ is the quantile function of $\V{y}$.
+We want to estimate multiple quantiles simultaneously and to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$ where $Q$ is the quantile function of $\V{y}$.
 Even if a single quantile is ultimately desired, ensuring non-crossing allows information from nearby quantiles to be used to improve the estimates.
@@
 {\bf Updating $\{\Mn{\Theta}{w}\}$: } We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in \Alg{admm}.
-The use of the Augmented Lagrangian converts the problem from a linear program into a quadratic program.
+The use of the Augmented Lagrangian converts the problem of solving a potentially large linear program into a solving a collection of smaller quadratic programs.
 The \texttt{gurobi} R package \citep{gurobi} can solve quadratic programs in addition to linear programs, but we can also use the free R package \texttt{quadprog} \citep{quadprog}.
@@
 \end{eqnarray*}
-
 where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod.
@@
 \label{sec:long-app}
-\Alg{admm} was used to removed the baseline drift from the full day of data  (\Fig{raw_spod}) consisting of 86,401 observations per SPod and compared to the series detrended using the \texttt{qsreg} trends estimated using non-overlapping 2 hour windows.
+\Alg{admm} was used to remove the baseline drift from the full day of data  (\Fig{raw_spod}) consisting of 86,401 observations per SPod and compared to the series detrended using the \texttt{qsreg} trends estimated using non-overlapping 2 hour windows.
 As in the shorter illustration, the \texttt{detrendr} method results in generally higher VI scores than the \texttt{qsreg} method (\Fig{vi_day}).

### 2019-04-09 Manuscript/detrendify_aoas.tex
added comment to Halley.

@@
 These results indicate that even when a single quantile is of interest, simultaneously fitting nearby quantiles and utilizing the non-crossing constraint can improve estimates when data is sparse by using information from nearby quantiles.
-The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets, but its performance deteriorates as the data size grows.
+\Halley{Eric}{Should this be reversed?
+qsreg looks like it does as well as detrend-eBIC for large $n$ and worse for small $n$.}The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets, but its performance deteriorates as the data size grows.
 The \texttt{npqw} and \texttt{detrend\_valid} methods both perform poorly on this design.

### 2019-04-09 Manuscript/detrendify_aoas.tex
adjusted LB for bw parameter npqw

--- 2019-04-08 Manuscript/detrendify_aoas.tex
@@
 These results indicate that even when a single quantile is of interest, simultaneously fitting nearby quantiles and utilizing the non-crossing constraint can improve estimates when data is sparse by using information from nearby quantiles.
-The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the smaller datasets, but its performance deteriorates as the data size grows.
+The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the larger datasets, but its performance deteriorates as the data size shrinks.
 The \texttt{npqw} and \texttt{detrend\_valid} methods both perform poorly on this design.

### 2019-04-09 Manuscript/detrendify_aoas.tex
responding to Joe's comments, updated npqw for Fig 8 and 9

@@
 \label{sec:intro}
+In the last decade, low cost and portable air quality sensors have enjoyed dramatically increased usage.
+These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
+The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
+Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
+Due to changes in temperature and relative humidity the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
+Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
+All of the sensors respond to the pollutant signal, which is illustrated by the three sharp transient spikes at 11:32, 14:10, and 16:03.
+However, the baseline drift varies from one sensor to another, obscuring the detection of the peaks that alert the intrusion of pollutants.
+We show later that by estimating the baseline drift in each sensor and  removing it from the observed signals, peaks can be reliably detected from concordant residual signals from a collection of SPods using a simple data-driven thresholding strategy.
+Thus, accurately demixing a noisy observed vector into a slowly varying component and a transient component can lead to greatly improved and simplified downstream analysis.
 
@@
 We describe this motivating problem further to illustrate the critical importance of being able to separate $\V{s}$ from $\V{\theta}$.
-
-In the last decade, low cost and portable air quality sensors have enjoyed dramatically increased usage.
-These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
-The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
-Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
-Due to changes in temperature and relative humidity the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
-Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
-All of the sensors respond to the pollutant signal, which is illustrated by the three sharp transient spikes at 11:32, 14:10, and 16:03.
-However, the baseline drift varies from one sensor to another, obscuring the detection of the peaks that alert the intrusion of pollutants.
-We show later that by estimating the baseline drift in each sensor %using $\ell_1$-quantile-trend filtering
-and  removing it from the observed signals, peaks can be reliably detected from concordant residual signals from a collection of SPods using a simple data-driven thresholding strategy.
-Thus, accurately demixing a noisy observed vector $\V{y}$ into a slowly varying component $\V{\theta}$ and a transient component $\V{s}$ can lead to greatly improved and simplified downstream analysis.
 
@@
 The additional constraints are linear in the parameters, so the non-crossing quantile trends can still be estimated by a number of available solvers.
-We allow for the possibilty that the degree of smoothness in the trends varies by quantile and thus allow the smoothing parameter as well.
+We allow for the possibility that the degree of smoothness in the trends varies by quantile by allowing the smoothing parameter to vary with quantile as well.
 In the rest of this paper, we report numerical results using the commercial solver Gurobi \citep{gurobi} and its R package implementation.
@@
 
- As the size of the data increases, all methods require additional computation time and at some point the penalty matrices will no longer fit into memory.
+ As the size of the data increases, computation time becomes prohibitive \Fig{timing}.
 In our application to air quality sensor data, measurements are recorded every second resulting in 86,400 observations per day.
@@
 
-Returning to our constrained windows problem giving in \Eqn{quantile_windows}, let $\Mn{\Omega}{w}$ denote the Lagrange multiplier matrix for the $w$th consensus constraint, namely $\Mn{\theta}{w} = \Mn{U}{w}\M{\theta}$, and let $\Vn{\omega}{w}_{j}$ denote its $j$th column.
+Returning to our constrained windows problem giving in \Eqn{quantile_windows}, let $\Mn{\Omega}{w}$ denote the Lagrange multiplier matrix for the $w$th consensus constraint, namely $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, and let $\Vn{\omega}{w}_{j}$ denote its $j$th column.
 
@@
 the sensor response to pollutants, and high frequency random noise.
-These sensors record measurements every second and are used to monitor pollutant concentrations at the fence lines of industrial facilities.
+These sensors record measurements every second and are used to monitor pollutant concentrations at the perimeter of industrial facilities.
 Time points with high concentrations are identified and compared with concurrent wind direction and speed.
 Ideally, three co-located and time aligned sensors (as shown in \Fig{raw_spod}) responding to a pollutant plume would result in the same signal classification after baseline trend removal and proper threshold choice.
-We first illustrate the difference between our \texttt{detrend\_eBIC} method,  hereafter referred to as \texttt{detrendr}, and \texttt{qsreg} using a small subset of measurements from \Fig{raw_spod} (\Sec{short-app}).
+We first illustrate the difference between our \texttt{detrend\_eBIC} method,  hereafter referred to as \texttt{detrendr}, and \texttt{qsreg} using data from 13:10 to 15:10 from \Fig{raw_spod} (\Sec{short-app}).
 We then compare the methods on the complete day shown in \Fig{raw_spod}, estimating trends by applying the \texttt{qsreg} method to 2 hour non-overlapping windows of the data and \Alg{admm} to the entire day.
@@
 
-We compare our \texttt{detrendr} method with the \texttt{qsreg} method on a subset of the SPod data (n=7200) both to facilitate visualization and because the \texttt{qsreg} method cannot handle all 24 hours simultaneously.
+We compare our \texttt{detrendr} method with the \texttt{qsreg} method on a two-hour subset of one-second  SPod data (n=7200) both to facilitate visualization and because the \texttt{qsreg} method cannot handle all 24 hours simultaneously.
 We estimate the baseline trend using $\tau = \{0.01, 0.05, 0.1\}$ and compare three thresholds for classifying the signal.
@@
 \Alg{admm} was used to remove the baseline drift from the full day of data  (\Fig{raw_spod}) consisting of 86,401 observations per SPod and compared to the series detrended using the \texttt{qsreg} trends estimated using non-overlapping 2 hour windows.
-As in the shorter illustration, the \texttt{detrendr} method results in generally higher VI scores than the \texttt{qsreg} method (\Fig{vi_day}).
+As in the shorter illustration, the \texttt{detrendr} method results in generally lower VI scores than the \texttt{qsreg} method (\Fig{vi_day}).
 The \texttt{detrendr} method also results in better correlation in the detrended series as is illustrated in (\Fig{scatter_day}).
@@
 Each point represents a day of measurements and all points that fall below the dashed line have more similar classifications using \texttt{detrendr} compared to \texttt{qsreg}.
+The improvement of \texttt{detrendr} over \texttt{qsreg} is more severe at lower thresholds.
+This indicates that \texttt{detrendr} gives greater agreement on signal classification when the methods are tuned to deliver positive classifications more frequently.
 
 \begin{figure}
-	\includegraphics[width = 0.7\linewidth]{Figures/VI_by_day.png}
+	\includegraphics[width = 0.6\linewidth]{Figures/VI_by_day.png}
 	\caption{Variation of Information (VI) for \texttt{detrendr} and \texttt{qsreg} by quantile trend and threshold.
@@
 Our ADMM algorithm for large series both reduces the computing time and allows trends to be estimated on series that cannot be estimated simultaneously while our scaled extended BIC criteria was shown to provided better estimated of quantile trends in series with and without a signal component.
-We have also shown that the baseline drift in low cost air quality sensors can be removed through estimating quantile trends.
+
+In our application to low cost air quality sensor data, we have shown that the baseline drift in low cost air quality sensors can be removed through estimating quantile trends, but the data size was too large for existing methods to be computationally feasible.
+While \texttt{qsreg} cannot feasibly handle more than a few hour windows of data, our new methods were able to process 24 hours simultaneously and deliver signal classifications that were more consistent between the two sensors for a week of data (168 hours).
 
@@
 
-\label{sec:conc}
+\section{Acknowledgments}
+The authors greatly appreciate Eben Thoma at the US Environmental Protection Agency for providing the SPod datasets.
+The authors are also thankful for the Oak Ridge Institute of Science Fellowship that partially supported this work.
+Plots were created using ggplot2 \citep{ggplot2}.
 

### 2019-04-09 Code/Fig_peaks_metrics.R

@@ -39,9 +39,9 @@ ggsave("../Manuscript/Figures/ex_baseline.png", width = 7, height = 4)
 simDesign <- "peaks"
 tau <- c(0.01, 0.05, 0.1)
 nSim <- 100
-colPal <- rev(c('#006d2c', '#2ca25f', '#66c2a4', '#b2e2e2',
-                #'#762a83','#9970ab','#c2a5cf', '#d9f0d3',
-                '#a6dba0','#5aae61','#1b7837'))
+colPal <- c('#006d2c', '#2ca25f', '#66c2a4', '#1c9099', 
+            "#c2a5cf", "#9970ab", "#762a83")
+
 methods <- c("detrend_eBIC", "detrend_SIC", "detrend_valid", "detrend_Xing",
              "rqss", "npqw", "qsreg") 
 MSEs <- as.data.frame(matrix(NA, nrow = nSim*length(methods), 
@@ -86,7 +86,7 @@ text_size <- 14
 
 summary_peaks %>% 
   ggplot( aes(x = factor(n), y = mean_mse, col = Method)) + 
-  geom_point(position = position_dodge(width = 0.5)) +
+  geom_point(position = position_dodge(width = 0.5), size=2) +
   geom_linerange(aes(ymin = mean_mse - 2*sd_mse, ymax = mean_mse + 2*sd_mse), 
                  position = position_dodge(width = 0.5))+
   facet_grid(.~factor(tau), scales = "free")+

### 2019-04-09 Code/Fig_sim_misclass.R

@@ -6,8 +6,8 @@ library(tidyverse)
 library(devtools)
 rm(list=ls())
 load_all("detrendr")
-colPal <- rev(c('#762a83','#9970ab','#c2a5cf', '#d9f0d3',
-                '#a6dba0','#5aae61','#1b7837'))
+colPal <- c('#006d2c', '#2ca25f', '#66c2a4', '#1c9099', 
+            "#c2a5cf", "#9970ab", "#762a83")
 
 tau <- c(0.01, 0.05, 0.1)
 nSim <- 100

### 2019-04-11 Manuscript/detrendify_aoas.tex
More minor edits.

@@
 	We address the problem of estimating smoothly varying baseline trends in time series data.
-This problem arises in several applications, including chemistry, macroeconomics, and medicine; however, our study is motivated by the analysis of data from low cost air quality sensors.
-Our methods extend the quantile trend filtering framework to allow the estimation of multiple quantile trends simultaneously while ensuring that the quantiles do not cross.
+This problem arises in a wide range of fields, including chemistry, macroeconomics, and medicine; however, our study is motivated by the analysis of data from low cost air quality sensors.
+Our methods extend the quantile trend filtering framework to enable the estimation of multiple quantile trends simultaneously while ensuring that the quantiles do not cross.
 To handle the computational challenge posed by very long time series, we propose a parallelizable alternating direction method of moments (ADMM) algorithm.
@@
 We show later that by estimating the baseline drift in each sensor and  removing it from the observed signals, peaks can be reliably detected from concordant residual signals from a collection of SPods using a simple data-driven thresholding strategy.
-Thus, accurately demixing a noisy observed vector into a slowly varying component and a transient component can lead to greatly improved and simplified downstream analysis.
+Thus, accurately demixing a noisy observed time series into a slowly varying component and a transient component can lead to greatly improved and simplified downstream analysis.
 
+While this work is motivated by the analysis of data from low cost air quality sensors, the problem of demixing nosy time series into trends and transients is ubiquitous across many fields of study.
 In a wide range of applications that spans chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar functions of time $y(t)$ are observed and assumed to be a superposition of an underlying slowly varying baseline trend $\theta(t)$, other more rapidly varying components $s(t)$, and noise.
@@
 In our air quality application, the roles of $\V{\theta}$ and $\V{s}$ are reversed; $\V{s}$ represents the signal of interest and $\V{\theta}$ represents a baseline drift that obscures the identification of the important transient events encoded in $\V{s}$.
-We describe this motivating problem further to illustrate the critical importance of being able to separate $\V{s}$ from $\V{\theta}$.
 
@@
 	\item \cite{Kim2009} proposed using the check function as a possible extension of $\ell_1$-trend filtering but did not investigate it further.
-Here, we develop the basic $\ell_1$-quantile-trend-filtering framework and extend it to model multiple quantiles simultaneously with non-crossing constraints to ensure validity and improve trend estimates.
+Here, we develop the basic $\ell_1$-quantile-trend-filtering framework and extend it to model {\em multiple quantiles simultaneously with non-crossing constraints to ensure validity and improve trend estimates}.
 	\item To reduce computation time and extend the method to long time series, we develop a parallelizable ADMM algorithm.
@@
 
- As the size of the data increases, computation time becomes prohibitive \Fig{timing}.
+ As the size of the data increases, computation time becomes prohibitive (\Fig{timing}).
 In our application to air quality sensor data, measurements are recorded every second resulting in 86,400 observations per day.

### 2019-04-11 Manuscript/detrendify_aoas.tex
removed \mathcal{I} notation, added note on differences between objectives, and corrected the theta update statement

@@
 Given quantiles $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
-\begin{eqnarray*}
+\begin{eqnarray}
 	\label{eq:noncross_trend}
 	&&\underset{\M{\Theta} \in \mathcal{C}}{\text{minimize}}\; \sum_{j=1}^J \left [\rho_{\tau_j}(\V{y} - \V{\theta}_{j}) +
-	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] \\
-\end{eqnarray*}
+	\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ] 
+\end{eqnarray}
 where $\M{\Theta} = \begin{pmatrix} \V{\theta}_1 & \V{\theta}_2 & \cdots & \V{\theta}_J \end{pmatrix} \in \Real^{n \times J}$ is a matrix whose $j$th column corresponds to the $j$th quantile signal $\V{\theta}_j$ and the set $\mathcal{C} = \{\M{\Theta} \in \Real^{n \times J}: \ME{\theta}{ij} \leq \ME{\theta}{ij'} \;\text{for $j \leq j'$}\}$ encodes the non-crossing quantile constraints.
@@
 
- Let $\mathcal{I}$ denote a subset of the indices $\{1, \ldots, n\}$ and $\V{y}_\mathcal{I} \in \Real^{\lvert I \rvert}$ denote the subvector obtained by stacking on top of each other the elements $\VE{y}{i}$ for $i \in \mathcal{I}$ in their canonical ordering.
- To decrease computation time and extend our method to larger problems, we divide our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into $W$ overlapping windows of observations $\Vn{y}{w} = \V{y}_{\mathcal{I}_w}$ where $\mathcal{I}_w$ is the set of integers from $l_w$ to $u_w$ inclusive, namely $\mathcal{I}_w = \{l_w, l_w + 1, \ldots, u_w -1, u_w\}$ and
+ To decrease computation time and extend our method to larger problems, we divide our observed series $\VE{y}{i}$ with $i = 1, \ldots, n$ into $W$ overlapping windows of observations, defining the vector of sequential elements indexed from $l_w$ to $u_w$ as $\Vn{y}{w} = \{y_{l_w}, y_{l_w + 1}, \ldots, y_{u_w -1}, y_{u_w}\}$, with
  \begin{eqnarray*}
@@
  \end{eqnarray*}
+ We define $n_w = u_w-l_w+1$ so that $\Vn{y}{w} \in \Real^{n_w}$.
 \Fig{windows} shows an example of 1200 observations being mapped into three equally sized overlapping windows of observations.
@@
 
- Given quantiles $\tau_1 < \cdots < \tau_J$, we introduce dummy variables $\Vn{\theta}{w}_{j} \in \Real^{\lvert \mathcal{I}_w \rvert}$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $w$.
+ Given quantiles $\tau_1 < \cdots < \tau_J$, we introduce dummy variables $\Vn{\theta}{w}_{j} \in \Real^{n_w}$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $w$.
 We then ``stitch" together the $W$ quantile trend estimates into consensus over the overlapping regions by introducing the constraint $\VnE{\theta}{w}{ij} = \VE{\theta}{i + l_w - 1,j}$ for $i = 1, \ldots, u_w - l_w + 1$ and for all $j$.
 Let $\Mn{\Theta}{w}$ be the matrix whose $j$th column is $\Vn{\theta}{w}_{j}$.
-Then we can write these constraints more compactly as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{\lvert \mathcal{I}_w \rvert \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $w$th window, namely
+Then we can write these constraints more compactly as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{n_w \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $w$th window, namely
  \begin{eqnarray*}
@@
 
+ The solution to \Eqn{quantile_windows} is {\em not} identical to the solution to \Eqn{noncross_trend} because of double counting of the overlapping sections.
+The solutions are very close, however, and the differences are essentially immaterial concerning downstream analysis.
 The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is  described in greater detail by \cite{boyd2011distributed}, but we briefly review how it can be used to iteratively solve the following equality constrained optimization problem which is a more general form of \Eqn{quantile_windows}.
@@
 
-{\bf Updating $\{\Mn{\Theta}{w}\}$: } We then estimate the trend separately in each window, which can be done in parallel, while constraining the overlapping pieces of the trends to be equal as outlined in \Alg{admm}.
+{\bf Updating $\{\Mn{\Theta}{w}\}$: } We then estimate the trend separately in each window, which can be done in parallel, while penalizing the differences in the overlapping pieces of the trends  as outlined in \Alg{admm}.
 The use of the Augmented Lagrangian converts the problem of solving a potentially large linear program into a solving a collection of smaller quadratic programs.

### 2019-04-17 Manuscript/detrendify_aoas.tex
more updates

@@
 We compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window:  $\V{\lambda}$ chosen using SIC \Eqn{SIC} \citep{KoenkerNgPortnoy1994} (\texttt{detrendr\_SIC}); $\V{\lambda}$ chosen using the validation method with the validation set consisting of every 5th observation (\texttt{detrendr\_valid}); and $\V{\lambda}$ chosen using the proposed scaled eBIC criteria \Eqn{eBIC} \texttt{detrendr\_eBIC}.
-For the second study we also examine the effect of the non-crossing quantile constraint by estimating the quantile trends separately and choosing $\V{\lambda}$ using the proposed scaled eBIC criteria (\texttt{detrendr\_cross}).
-We do not include \texttt{detrendr\_cross} in the first study because the difference in quantiles is large enough that we would not expect the non-crossing constraint to make a difference.
+For the second study we also examine the effect of the non-crossing quantile constraint by estimating the quantile trends separately and choosing $\V{\lambda}$ using the proposed scaled eBIC criteria (\texttt{detrendr\_Xing}).
+We do not include \texttt{detrendr\_Xing} in the first study because the difference in quantiles is large enough that we would not expect the non-crossing constraint to make a difference.
 

### 2019-04-17 Code/Sim_windows.R

@@ -6,7 +6,7 @@ tau <- c(0.01, 0.05, 0.1)
 simDesigns <- c("gaus", "mixednorm", "shapebeta", "peaks")
 simDesigns <- "peaks" 
 
-i = 1
+for (i in c(24, 19, 58, 91, 78, 72, 71, 31, 76, 86)){ 
 for (n in c(500,1000,2000,4000)){
   lambdaSeq = n^seq(0,1.4,length.out=15) 
   overlap <- n/10
@@ -16,12 +16,16 @@ for (n in c(500,1000,2000,4000)){
     trend_BIC <- get_windows_BIC(df$y, tau, k=3,
                           lambdaSeq = lambdaSeq,
                           window_size = window_size, 
-                          overlap = overlap,
+                          overlap = overlap, 
+                          rho = 1,
+                          max_iter = 20,
                           plot_lambda = FALSE,
                           criteria = "eBIC")
+    
     trend <- trend_BIC$trend
     save(trend, trend_BIC,
          file = sprintf("../SimResults/windows/%s_n_%i_sim%03.0f.RData",
                         simDesign, n, i))
   }
+}
 }

### 2019-04-17 Manuscript/detrendify_aoas.tex
more edits.

@@
 
+\begin{figure}[t!]
+	\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
+	\caption{Example of 3 co-located SPod PID sensor readings over a 24 hour period.}
+	\label{fig:raw_spod}
+\end{figure}s
+
 While this work is motivated by the analysis of data from low cost air quality sensors, the problem of demixing nosy time series into trends and transients is ubiquitous across many fields of study.
@@
 In our air quality application, the roles of $\V{\theta}$ and $\V{s}$ are reversed; $\V{s}$ represents the signal of interest and $\V{\theta}$ represents a baseline drift that obscures the identification of the important transient events encoded in $\V{s}$.
-
-\begin{figure}[t!]
-	\includegraphics[width = \linewidth]{Figures/uncorrected_data.png}
-	\caption{Example of 3 co-located SPod PID sensor readings.}
-	\label{fig:raw_spod}
-\end{figure}
 
@@
 where $\lambda$ is a nonnegative regularization parameter, and the matrix $\Mn{D}{k+1} \in \Real^{(n - k -1) \times n}$ is the discrete difference operator of order $k+1$.
-To understand the purpose of penalizing the size of $\Mn{D}{k+1}\V{\theta}$ consider the difference operator when $k = 0$.
+To understand the purpose of penalizing the 1-norm of $\Mn{D}{k+1}\V{\theta}$ consider the difference operator when $k = 0$.
 
@@
 We want to estimate multiple quantiles simultaneously and to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$ where $Q$ is the quantile function of $\V{y}$.
-Even if a single quantile is ultimately desired, ensuring non-crossing allows information from nearby quantiles to be used to improve the estimates.
+Even if a single quantile is ultimately desired, ensuring non-crossing allows information from nearby quantiles to be used to improve the estimates as we will see in the peak detection experiments in \Sec{peak_detection}.
 Given quantiles $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
@@
 where $\M{\Theta} = \begin{pmatrix} \V{\theta}_1 & \V{\theta}_2 & \cdots & \V{\theta}_J \end{pmatrix} \in \Real^{n \times J}$ is a matrix whose $j$th column corresponds to the $j$th quantile signal $\V{\theta}_j$ and the set $\mathcal{C} = \{\M{\Theta} \in \Real^{n \times J}: \ME{\theta}{ij} \leq \ME{\theta}{ij'} \;\text{for $j \leq j'$}\}$ encodes the non-crossing quantile constraints.
-The additional constraints are linear in the parameters, so the non-crossing quantile trends can still be estimated by a number of available solvers.
+The additional non-crossing constraints are linear inequalities involving the parameters, so the non-crossing quantile trends can still be estimated by a number of available linear programming solvers.
 We allow for the possibility that the degree of smoothness in the trends varies by quantile by allowing the smoothing parameter to vary with quantile as well.
@@
  	\includegraphics[width = 0.8\linewidth]{Figures/overlapping_windows.png}
- 	\caption{Window boundaries and trends fit separately in each window.}
+ 	\caption{Window boundaries and trends fit {\em separately} in each window. Each window's trend estimate is plotted in a different line type.}
  	\label{fig:windows}
  \end{figure}
+
+\subsubsection{Formulation}
 
@@
 \Fig{windows} shows an example of 1200 observations being mapped into three equally sized overlapping windows of observations.
-While the overlapping trend estimates between $l_2$ and $u_2$ don't vary dramatically, the difference is more pronounced in the trend in the 5\textsuperscript{th} quantile between $l_3$ and $u_2$.
+While the overlapping trend estimates between $l_2$ and $u_2$ do not vary dramatically, the difference is more pronounced in the trend in the 5\textsuperscript{th} quantile between $l_3$ and $u_2$.
+Thus, we need a way of enforcing estimates to be identical in the overlapping regions.
 
@@
 Let $\Mn{\Theta}{w}$ be the matrix whose $j$th column is $\Vn{\theta}{w}_{j}$.
-Then we can write these constraints more compactly as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{n_w \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $w$th window, namely
+Then we can write these constraints more concisely as $\Mn{\Theta}{w} = \Mn{U}{w}\M{\Theta}$, where $\Mn{U}{w} \in \{0, 1\}^{n_w \times n}$ is a matrix that selects rows of $\M{\Theta}$ corresponding to the $w$th window, namely
  \begin{eqnarray*}
@@
 The solutions are very close, however, and the differences are essentially immaterial concerning downstream analysis.
+
+\subsubsection{Algorithm}
 The ADMM algorithm \citep{gabay1975dual, glowinski1975approximation} is  described in greater detail by \cite{boyd2011distributed}, but we briefly review how it can be used to iteratively solve the following equality constrained optimization problem which is a more general form of \Eqn{quantile_windows}.
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 	\caption{ADMM algorithm for quantile trend filtering with windows}\label{alg:admm}
@@
 	\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
-	\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.}
+	\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit. \Halley{Eric}{Need to reference in the text.}}
 \end{figure}
@@
 \subsection{Peak Detection}
+\label{sec:peak_detection}
 %	\begin{figure}[h]
@@
 	\caption{Example signal classification using threshold.
-Red indicates true signal $>0.5$, blue indicates observations classified as signal after baseline removal using \texttt{detrend\_eBIC} and a threshold of 1.2.}
+Red indicates true signal $(\VE{y}{i} - \VE{\theta}{i} > 0.5)$, blue indicates observations classified as signal after baseline removal using \texttt{detrend\_eBIC} and a threshold of 1.2.
+\Halley{Eric}{Need to explain what the threshold of 1.2 is.}}
 	\label{fig:peaks_class_eg}
@@
 	\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
-	\caption{Class averaged accuracy by threshold, data size, and method (1 is best 0.5 is worst).}
+	\caption{Class averaged accuracy by threshold, data size, and method (1 is best 0.5 is worst). \Halley{Eric}{Need to reference in the text.}}
 \end{figure}
@@
 	\includegraphics[width = 0.3\linewidth]{Figures/scatter_day_detrendr.png}
-	\caption{SPod a versus SPod c before and after de-trending. }
+	\caption{SPod a versus SPod c before and after de-trending with \texttt{qsreg} and \texttt{detrendr}. }
 	\label{fig:scatter_day}

### 2019-04-17 Manuscript/detrendify_aoas.tex
changed to eBIC
added references to figures

@@
 The solutions are very close, however, and the differences are essentially immaterial concerning downstream analysis.
+\Fig{windowsfit} provides an illustration of trends estimated using \Alg{admm} and multiple windows compared with the trends estimated using a single window.
+
+ \begin{figure}
+ 	\centering
+ 	\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
+ 	\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.} 
+ 	\label{fig:windowsfit}
+ \end{figure}
 
@@
 The proof of \Prop{convergence} is a straightforward application of the convergence result presented in Section 3.2 of \cite{boyd2011distributed}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
-	\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit. \Halley{Eric}{Need to reference in the text.}}
-\end{figure}
 
@@
 
-Our \texttt{detrend\_BIC} method performs the best overall in terms of both RMSE and CAA.
+Our \texttt{detrend\_eBIC} method performs the best overall in terms of both RMSE (\Fig{peaks_rmse}) and CAA (\Fig{CAA}).
 While \texttt{qsreg} was competitive with our method in some cases, in the majority of cases the largest CAA values for each threshold were produced using the \texttt{detrend\_eBIC} method with the 1\textsuperscript{st} or 5\textsuperscript{th} quantiles.
@@
 	\caption{Example signal classification using threshold.
-Red indicates true signal $(\VE{y}{i} - \VE{\theta}{i} > 0.5)$, blue indicates observations classified as signal after baseline removal using \texttt{detrend\_eBIC} and a threshold of 1.2.
-\Halley{Eric}{Need to explain what the threshold of 1.2 is.}}
+Red indicates true signal $(\VE{y}{i} - \VE{\theta}{i} > 0.5)$, blue indicates observations classified as signal, i.e.
+values greater than 1.2 after baseline removal using \texttt{detrend\_eBIC}.}
 	\label{fig:peaks_class_eg}
@@
 	\includegraphics[width = \linewidth]{Figures/peaks_CAA.png}
-	\caption{Class averaged accuracy by threshold, data size, and method (1 is best 0.5 is worst). \Halley{Eric}{Need to reference in the text.}}
+	\caption{Class averaged accuracy by threshold, data size, and method (1 is best 0.5 is worst).}
+	\label{fig:CAA}
 \end{figure}

### 2019-04-18 Manuscript/detrendify_aoas.tex
organizing code, manuscript edits

@@
 	\label{fig:raw_spod}
-\end{figure}s
+\end{figure}
 
@@
 In a wide range of applications that spans chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar functions of time $y(t)$ are observed and assumed to be a superposition of an underlying slowly varying baseline trend $\theta(t)$, other more rapidly varying components $s(t)$, and noise.
-In practice, $y(t)$ is observed at discrete time points $t_1, \ldots, t_n$, and we model the vector of signal samples $\VE{y}{i} = y(t_i)$ as
+In practice, $y(t)$ is observed at discrete time points $t_1, \ldots, t_n$, and we model the vector of samples $\VE{y}{i} = y(t_i)$ as
 \begin{eqnarray*}
@@
 
-In the rest of the paper, we detail our quantile trend filtering algorithms as well as how to choose the smoothing parameter (Section \ref{sec:methods}).
+In the rest of the paper, we detail our quantile trend filtering algorithms (Section \ref{sec:methods}) as well as how to choose the smoothing parameter (Section \ref{sec:lambda_choice}).
 We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods (Section \ref{sec:simluation}).
@@
 
-Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$, which is known as total variation denoising penalty in one dimension in the signal processing literature \citep{Rudin1992} or the fused lasso penalty in the statistics literature \citep{Tibshirani2005}.
+Thus, $\lVert \Mn{D}{1}\V{\theta} \rVert_1 = \sum_{i=1}^{n-1} \lvert \theta_i - \theta_{i+1} \rvert$, which is known as the total variation denoising penalty in one dimension in the signal processing literature \citep{Rudin1992} or the fused lasso penalty in the statistics literature \citep{Tibshirani2005}.
 The penalty term incentivizes solutions which are piecewise constant.
@@
 
-\cite{Tib2014} proved that with a judicious choice of $\lambda$ the trend filtering estimate converges to the true underlying function at the minimax rate for functions whose $k$th derivative is of bounded variation and showed that trend filtering is both fast and locally adaptive when the time series consists of only the trend and random noise, which is illustrated in \Fig{tpn}.
+\cite{Tib2014} proved that with a judicious choice of $\lambda$ the trend filtering estimate converges to the true underlying function at the minimax rate for functions whose $k$th derivative is of bounded variation and showed that trend filtering is locally adaptive when the time series consists of only the trend and random noise, which is illustrated in \Fig{tpn}.
 As noted earlier, in some applications, such as the air quality monitoring problem considered in this paper, the data contain a rapidly varying signal in addition to the slowly varying trend and noise.
-\Fig{tpspn} shows that standard trend filtering is not designed to distinguish between the slowly varying trend and the rapidly-varying signal, as the smooth component estimate $\V{\theta}$ is biased towards the peak of the transient component.
+\Fig{tpspn} shows that standard trend filtering is not designed to distinguish between the slowly varying trend and the rapidly-varying signal, as the smooth component estimate $\V{\theta}$ is biased towards the peaks of the transient components.
 
-To account for the presence of a transient component in the observed time series $\V{y}$, we propose quantile trend filtering \Fig{tpspn}.
+To account for the presence of transient components in the observed time series $\V{y}$, we propose quantile trend filtering \Fig{tpspn}.
 % as a more robust method for detrending in these situations \Fig{tpspn}.
@@
 \cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check function with a differentiable function.
-\cite{Racine2017} take a different approach and constrain the response to follow a location scale model and estimate the $\tau$\textsuperscript{th} conditional quantile by combining Gaussian quantile functions with a kernel smoother and solving a local-linear least squares problem.
+\cite{Racine2017} propose a method for estimating quantile trends that does employ the check function.
+In their method, the response is constrained to follow a location scale model and the conditional quantiles are estimated by combining Gaussian quantile functions with a kernel smoother and solving a local-linear least squares problem.
 
@@
 We allow for the possibility that the degree of smoothness in the trends varies by quantile by allowing the smoothing parameter to vary with quantile as well.
-In the rest of this paper, we report numerical results using the commercial solver Gurobi \citep{gurobi} and its R package implementation.
+In the rest of this paper, we use $k=2$ to produce piecewise quadratic polynomials and report numerical results using the commercial solver Gurobi \citep{gurobi} and its R package implementation.
 However, we could easily substitute a free solver such as the Rglpk package by \cite{rglpk}.
@@
 
- As the size of the data increases, computation time becomes prohibitive (\Fig{timing}).
+ As the size of the data increases, computation time becomes prohibitive.
 In our application to air quality sensor data, measurements are recorded every second resulting in 86,400 observations per day.
@@
  Given quantiles $\tau_1 < \cdots < \tau_J$, we introduce dummy variables $\Vn{\theta}{w}_{j} \in \Real^{n_w}$ as the value of the $\tau_j$\textsuperscript{th} quantile trend in window $w$.
-We then ``stitch" together the $W$ quantile trend estimates into consensus over the overlapping regions by introducing the constraint $\VnE{\theta}{w}{ij} = \VE{\theta}{i + l_w - 1,j}$ for $i = 1, \ldots, u_w - l_w + 1$ and for all $j$.
+We then ``stitch" together the $W$ quantile trend estimates into consensus over the overlapping regions by introducing the constraint $\VnE{\theta}{w}{ij} = \VE{\theta}{i + l_w - 1,j}$ for $i = 1, \ldots, n_w$ and for all $j$.
 Let $\Mn{\Theta}{w}$ be the matrix whose $j$th column is $\Vn{\theta}{w}_{j}$.
@@
 Furthermore, let $\iota_{\mathcal{C}}$ denote the indicator function of the non-crossing quantile constraint, namely $\iota_{\mathcal{C}}(\M{\Theta})$ is zero if $\M{\Theta} \in \mathcal{C}$ and infinity otherwise.
-
 Our windowed quantile trend optimization problem can then be written as
@@
 The solutions are very close, however, and the differences are essentially immaterial concerning downstream analysis.
-\Fig{windowsfit} provides an illustration of trends estimated using \Alg{admm} and multiple windows compared with the trends estimated using a single window.
+\Fig{windowsfit} provides an illustration of the trends estimated using multiple windows compared with the trends estimated using a single window.
 
@@
  	\includegraphics[width = 0.8\linewidth]{Figures/admm_windows.png}
- 	\caption{Trend fit with our ADMM algorithm with 3 windows which converged in 7 iterations compared to trend from simultaneous fit.} 
+ 	\caption{Trend fit with our ADMM algorithm with 3 windows (converged in 7 iterations), compared to trend from simultaneous fit.} 
  	\label{fig:windowsfit}

### 2019-04-18 Code/Fig3-4_admm.R

@@ -1,0 +1,84 @@
+################################################################################
+# Figure 3 and 4 - ADMM algorithm examples 
+# Halley Brantley
+################################################################################
+library(tidyverse)
+library(devtools)
+devtools::install_github("const-ae/ggsignif")
+load_all("detrendr")
+
+rm(list=ls())
+source("sim_generating_functions.R")
+set.seed(987652)
+overlap <- 150
+window_size <- 500
+n <- window_size*3 - overlap*2
+df.data <- generate_peaks_design(n)
+df.data$x <- seq(1, n, 1)
+y <- df.data$y
+x <- df.data$x
+k <- 3
+tau <- c(0.05, 0.1)
+
+lambda <- length(y)
+max_iter <- 25
+result0 <- get_trend(y, tau, lambda, k)
+result1 <- get_trend(y[1:window_size], tau, lambda, k)
+result2 <- get_trend(y[(window_size - overlap + 1):(2*window_size - overlap)], tau, lambda, k)
+result3 <- get_trend(y[(2*window_size - 2*overlap + 1):(3*window_size - 2*overlap)], tau, lambda, k)
+
+df.sep.no <- data.frame(rbind(#data.frame(x = x, result0, method="Single Fit"), 
+                              data.frame(x = x[1:window_size], result1, method = "Window 1"),
+                              data.frame(x = x[(window_size - overlap + 1):(2*window_size - overlap)], result2, method = "Window 2"),
+                              data.frame(x = x[(2*window_size - 2*overlap + 1):(3*window_size - 2*overlap)], result3, method = "Window 3"))) 
+
+
+ggplot(df.data, aes(x=x, y=y)) +
+  geom_line(col="grey") +
+  geom_line(data = df.sep.no, aes(y=X1, col = "0.05", linetype = method))+
+  geom_line(data = df.sep.no, aes(y=X2, col = "0.10", linetype = method))+
+  scale_color_brewer(palette = "Set1")+
+  labs(col="Quantile", linetype = "", x = "") + 
+  guides(linetype = "none") +
+  theme_bw() +
+  geom_segment(aes(x = 1, xend = 1, y=0, yend=1.5, linetype = "dotted")) +
[…]

### 2019-04-18 Code/Fig_app_day.R

@@ -21,16 +21,6 @@ for (node in nodes){
   names(spod_trends)[(ncol(spod_trends)-length(tau)+1):ncol(spod_trends)] <-
     paste(node, tau, sep = "_")
 }
-spodRaw <- spodPIDs %>%
-  gather("node", "PID", -time)
-
-ggplot(spodRaw, aes(x=time, y=PID, col=node)) +
-  geom_line(alpha=0.5) +
-  theme_bw() +
-  scale_color_brewer(palette = "Set1", labels = c("a", "b", "c"))+
-  labs(col="SPod")
-
-ggsave("../Manuscript/Figures/uncorrected_data.png", width = 7, height = 2.5)
 ################################################################################
 peaks_qsreg <- select(spodPIDs, -time) - 
   select(qsreg_trends, ends_with(paste(tau[2])))

### 2019-04-18 Code/Fig_consensus.R

@@ -1,195 +1,0 @@
-library(tidyverse)
-library(devtools)
-devtools::install_github("const-ae/ggsignif")
-load_all("detrendr")
-
-rm(list=ls())
-source("sim_generating_functions.R")
-set.seed(987652)
-overlap <- 150
-window_size <- 500
-n <- window_size*3 - overlap*2
-df.data <- generate_peaks_design(n)
-df.data$x <- seq(1, n, 1)
-y <- df.data$y
-x <- df.data$x
-k <- 3
-tau <- c(0.05, 0.1)
-
-lambda <- length(y)
-max_iter <- 25
-result0 <- get_trend(y, tau, lambda, k)
-result1 <- get_trend(y[1:window_size], tau, lambda, k)
-result2 <- get_trend(y[(window_size - overlap + 1):(2*window_size - overlap)], tau, lambda, k)
-result3 <- get_trend(y[(2*window_size - 2*overlap + 1):(3*window_size - 2*overlap)], tau, lambda, k)
-
-df.sep.no <- data.frame(rbind(#data.frame(x = x, result0, method="Single Fit"), 
-                              data.frame(x = x[1:window_size], result1, method = "Window 1"),
-                              data.frame(x = x[(window_size - overlap + 1):(2*window_size - overlap)], result2, method = "Window 2"),
-                              data.frame(x = x[(2*window_size - 2*overlap + 1):(3*window_size - 2*overlap)], result3, method = "Window 3"))) 
-
-
-ggplot(df.data, aes(x=x, y=y)) +
-  geom_line(col="grey") +
-  geom_line(data = df.sep.no, aes(y=X1, col = "0.05", linetype = method))+
-  geom_line(data = df.sep.no, aes(y=X2, col = "0.10", linetype = method))+
-  scale_color_brewer(palette = "Set1")+
-  labs(col="Quantile", linetype = "", x = "") + 
-  guides(linetype = "none") +
-  theme_bw() +
-  geom_segment(aes(x = 1, xend = 1, y=0, yend=1.5, linetype = "dotted")) +
-  geom_segment(aes(x = x[350], xend=x[350], y=0, yend=1.5, linetype = "dotted")) +
-  geom_segment(aes(x = x[500], xend=x[500], y=0, yend=1.5, linetype = "dotted")) +
-  geom_segment(aes(x = x[700], xend=x[700], y=0, yend=1.5, linetype = "dotted")) +
[…]

### 2019-04-18 Manuscript/detrendify_aoas.tex
updated notation for consensus update

@@
 \end{eqnarray*}
-where $\gamma$ is a positive penalty parameter.
+where $\gamma$ is a positive tuning parameter.
 
@@
 
-{\bf Updating $\M{\Theta}$: } Some algebra shows that updating the consensus variable step is computed as follows.
+{\bf Updating $\M{\Theta}$: } Some algebra shows that,  defining $i_w = i-l_w+1$, updating the consensus variable step is computed as follows.
 
-\begin{eqnarray}
+\begin{align}
 \label{eq:update_consensus}
 ~~~~\ME{\theta}{ij}	 &=& \begin{cases}
-\frac{1}{2}\left (\MnE{\theta}{w-1}{ij} + \MnE{\theta}{w}{ij} \right)
+\frac{1}{2}\left (\MnE{\theta}{w-1}{i_{w-1}j} + \MnE{\theta}{w}{i_wj} \right)
 -
-\frac{1}{2\gamma} \left (\MnE{\omega}{w-1}{ij} + \MnE{\omega}{w}{ij} \right)
+\frac{1}{2\gamma} \left (\MnE{\omega}{w-1}{i_{w-1}j} + \MnE{\omega}{w}{i_{w}j} \right)
 & \mbox{if~} l_{w} \le i \le u_{w-1},  \\
-\MnE{\theta}{w}{ij} & \mbox{if~} u_{w-1} < i \le l_{w+1}  \\
-\frac{1}{2} \left ( \MnE{\theta}{w}{ij} + \MnE{\theta}{w+1}{ij} \right )
+\MnE{\theta}{w}{i_{w}j} & \mbox{if~} u_{w-1} < i \le l_{w+1}  \\
+\frac{1}{2} \left ( \MnE{\theta}{w}{i_{w}j} + \MnE{\theta}{w+1}{i_{w+1}j} \right )
 -
-\frac{1}{2\gamma}	\left (\MnE{\omega}{w}{ij} + \MnE{\omega}{w+1}{ij} \right )
-& \mbox{if~} l_{m+1} < i \le u_{m}
+\frac{1}{2\gamma}	\left (\MnE{\omega}{w}{i_{w}j} + \MnE{\omega}{w+1}{i_{w+1}j} \right )
+& \mbox{if~} l_{w+1} < i \le u_{w}
 \end{cases},
-\end{eqnarray}
+\end{align}
 

### 2019-04-18 Manuscript/detrendify_aoas.tex
fixed primal residual
edited timing section

@@
 
+\Alg{admm} has the following convergence guarantees.
+\begin{proposition}
+	\label{prop:convergence}
+	Let $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$ denote the $m$th collection of iterates generated by \Alg{admm}. Then (i)
+	$\lVert \Mn{\Theta}{w}_m - \Mn{U}{w}\M{\Theta}_m \rVert_{\text{F}} \rightarrow 0$ and (ii) $p_m \rightarrow p^\star$, where $p^\star$ is the optimal objective function value of
+	\Eqn{quantile_windows} and $p_m$ is the objective function value of \Eqn{quantile_windows} evaluated at $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$.
+\end{proposition}
+The proof of \Prop{convergence} is a straightforward application of the convergence result presented in Section 3.2 of \cite{boyd2011distributed}.
+
 To terminate our algorithm, we use the stopping criteria described by \cite{boyd2011distributed}.
@@
 \begin{eqnarray*}
-	r_{\text{primal}}^{m} & = & \sqrt{\sum_{w=1}^W\lVert\Mn{\Theta}{w}_{m} - \M{\Theta}_m\rVert_{\text{F}}^2},
+	r_{\text{primal}}^{m} & = & \sqrt{\sum_{w=1}^W\lVert\Mn{\Theta}{w}_{m} - \Mn{U}{w}\M{\Theta}_m\rVert_{\text{F}}^2},
 \end{eqnarray*}
@@
 
-\Alg{admm} has the following convergence guarantees.
-
-\begin{proposition}
-	\label{prop:convergence}
-	Let $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$ denote the $m$th collection of iterates generated by \Alg{admm}. Then (i)
-	$\lVert \Mn{\Theta}{w}_m - \Mn{U}{w}\M{\Theta}_m \rVert_{\text{F}} \rightarrow 0$ and (ii) $p_m \rightarrow p^\star$, where $p^\star$ is the optimal objective function value of
-	\Eqn{quantile_windows} and $p_m$ is the objective function value of \Eqn{quantile_windows} evaluated at $\{\{\Mn{\Theta}{w}_m\}, \M{\Theta}_m\}$.
-\end{proposition}
-
-The proof of \Prop{convergence} is a straightforward application of the convergence result presented in Section 3.2 of \cite{boyd2011distributed}.
-
 The quantile trend filtering problem for a single window is a linear program with $N\times J$ parameters (number of observations by number of quantiles), which can be solved in computational time proportional to $(NJ)^3$.
-By dividing the problem into four smaller problems we can theoretically solve each of the 4 sub-problems 4 times and still reduce the computation time required by a factor of 4.
-Consequently, solving a large problem using \Alg{admm} should require less computational time than simply using a linear program solver given small enough windows and convergence in few enough iterations.
-We demonstrate the advantages of \Alg{admm} empirically using timing experiments.
-For each data size, $n$, 25 datasets were simulated using the peaks simulation design described below.
-Trends for three quantiles were fit simultaneously: 0.05, 0.1, and 0.15 using $\lambda = n/5$.
+Consequently, solving a large problem using \Alg{admm} should require less computational time than solving \Eqn{noncross_trend}, even if the sub-problems are solved sequentially.
+We demonstrate the advantages of \Alg{admm} through timing experiments (\Fig{timing}).
+For each data size, 25 datasets were simulated using the peaks simulation design described below.
+Trends for the fifth, tenth, and fifteenth quantiles were fit simultaneously: using $\lambda = n/5$.
 We use from one to four windows for each data size with an overlap of 500.
 The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
-\Fig{timing} shows that using 4 windows instead of one on data sizes of 55000 provides a factor of 3 decrease in computation time.
+\Fig{timing} shows that using 4 windows instead of one on data sizes of 55,000 provides a factor of 3 decrease in computation time.
 The timing experiments were conducted on an Intel Xeon based Linux cluster using two processor cores.

### 2019-04-18 Code/Fig3-4_admm.R

@@ -47,13 +47,10 @@ ggplot(df.data, aes(x=x, y=y)) +
   geom_segment(aes(x = x[700], xend=x[700], y=0, yend=1.5, linetype = "dotted")) +
   geom_segment(aes(x = x[850], xend=x[850], y=0, yend=1.5, linetype = "dotted")) +
   geom_segment(aes(x = n, xend = n, y=0, yend=1.5, linetype = "dotted")) +
-  
   geom_segment(aes(x = x[1], xend=x[500], y=2, yend=2)) +
   geom_text(aes(x=250, y=2.15, label = "Window 1")) +
-  
   geom_segment(aes(x = 350, xend=850, y=2.4, yend=2.4)) +
   geom_text(aes(x=600, y=2.55, label = "Window 2")) +
-  
   geom_segment(aes(x = 700, xend=1200, y=2.7, yend=2.7)) +
   geom_text(aes(x=950, y=2.85, label = "Window 3")) +
   geom_label(aes(x=1, y=-.2, label = "1")) +

### 2019-04-18 Manuscript/detrendify_aoas.tex
edits on model selection

@@
 
-\begin{eqnarray*}
+\begin{eqnarray}
 	r_{\text{primal}}^{m} & < &\epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\underset{w}{\max} \left[\max
-	\left\{\lVert\Mn{\Theta}{w}_m \rVert_{\text{F}}, \lVert \M{\Theta}_{m} \rVert_{\text{F}} \right\}\right] \\%\lVert \M{\Theta}_{k} \rVert_{\text{F}}} \right )\right]\\
+	\left\{\lVert\Mn{\Theta}{w}_m \rVert_{\text{F}}, \lVert \M{\Theta}_{m} \rVert_{\text{F}} \right\}\right] \nonumber \\%\lVert \M{\Theta}_{k} \rVert_{\text{F}}} \right )\right]\\
 	r_{\text{dual}}^{m} & < & \epsilon_{\text{abs}}\sqrt{nJ} + \epsilon_{\text{rel}}\,\sqrt{\sum_{w=1}^W
 		\lVert \Mn{\Omega}{w}_m \rVert_{\text{F}}^2}.
-\end{eqnarray*}
+\label{eq:stopping}
+\end{eqnarray}
 
@@
 For each data size, 25 datasets were simulated using the peaks simulation design described below.
-Trends for the fifth, tenth, and fifteenth quantiles were fit simultaneously: using $\lambda = n/5$.
+Trends for the fifth, tenth, and fifteenth quantiles were fit simultaneously using $\lambda_j = n/5$ for all $j$.
 We use from one to four windows for each data size with an overlap of 500.
-The windows algorithm was run until the stopping criteria were met using $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
+\Alg{admm} was the stopped when \Eqn{stopping} was satisfied, defining $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
 \Fig{timing} shows that using 4 windows instead of one on data sizes of 55,000 provides a factor of 3 decrease in computation time.
@@
 Rather than combine results over quantiles, we allow the value of $\lambda$ to vary by quantile resulting in $\V{\lambda} = \{\lambda_1, ..., \lambda_J\}$.
-To choose the best value for each $\lambda_j$, we first estimate all of the quantile trends using $\V{\lambda} = \lambda\mathbf{1}$ over a grid of values for $\lambda$.
+To choose the best value for each $\lambda_j$, we first estimate all of the quantile trends using $\lambda_j = \lambda$ for all $j$ over a grid of values for $\lambda$.
 We then determine the $\lambda_j$ that maximizes the criteria chosen evaluated using $\tau_j$.
@@
 \end{eqnarray*}
-where $p_{\lambda} = \sum_i \One(\VE{y}{i} = \widehat{\theta}_i)$ is the number of interpolated points, which can be thought of as active knots.
+where $p_{\lambda} = \sum_i \One(\VE{y}{i} = \widehat{\theta}_i)$ is the number of non-interpolated points, which can be thought of as active knots.
 Equivalently, $p_{\lambda}$ can be substituted with the number of non-zero components in $\M{D}^{(k+1)}\Vhat{\theta}_j$ which we denote $\nu$ and have found to be more numerically stable.
@@
 \begin{eqnarray*}
-	L(\V{y} \mid \V{\theta}) & = & \left[\frac{\tau^n(1-\tau)}{\sigma}\right]^n\exp\left\{-\sum_i\rho_\tau\left(\frac{\VE{y}{i} - \VE{\theta}{i}}{\sigma}\right)\right\},
+	L(\V{y} \mid \V{\theta}) & = & \left[\frac{\tau^n(1-\tau)}{\sigma}\right]^n\exp\left\{-\rho_\tau\left(\frac{\V{y} - \V{\theta}}{\sigma}\right)\right\},
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-	\mbox{BIC}(\nu, \tau_j) = 2\frac{1}{\sigma}\rho_{\tau_j}(\V{y}-\Vhat{\theta}_j) + \nu\log n
+	\mbox{BIC}(\nu, \tau_j) = \frac{2}{\sigma}\rho_{\tau_j}(\V{y}-\Vhat{\theta}_j) + \nu\log n
 \end{eqnarray*}
@@
 We used this criteria with $\gamma = 1$, and $P=n-k-1$.
-In the simulation study, we compare the performance of the SIC, scaled eBIC, and validation methods.
-
-%\begin{figure}[h!]
-%	\includegraphics[width = 0.25\linewidth]{Figures/df_by_lambda.png}
-%	\includegraphics[width = 0.25\linewidth]{Figures/BIC_data.png}
-%	\includegraphics[width = 0.5\linewidth]{Figures/BIC_by_lambda.png}
-%	\caption{(Left) Estimated degrees of freedom (number of non-zero elements of $\M{D}\V{\theta}$) as a function of $\log(\lambda)$.
-(Middle) Estimated 10th quantile trend with regularization parameter chosen using various criteria.
-(Right) Criteria values as a function of $\log(\lambda)$, vertical lines indicate locations of minima.}
-%	\label{fig:BIC}
-%\end{figure}
+In the simulation study, we compare the performance of the SIC, scaled eBIC (with $\sigma$ defined above), and validation methods.
 
@@
 We conduct two simulation studies to compare the performance of our quantile trend filtering method and regularization parameter selection criteria with previously published methods.
-The first study compares the method's ability to estimate quantiles when the only components of the observed series are a smooth trend and a random component.
-The second study is based on our application and compares the method's ability to estimate baseline trends and enable peak detection when the time series contains a non-negative signal component in addition to the trend and random component.
+The first study compares the method's ability to estimate quantiles when the observed series consists of a smooth trend plus independent error, but does not contain transient components.
+The second study is based on our application and compares the method's ability to estimate baseline trends and enable peak detection when the time series contains a non-negative, transient signal in addition to the trend and random component.
 
-We compare three criteria for choosing the smoothing parameter with our quantile trend filtering method with a single window:  $\V{\lambda}$ chosen using SIC \Eqn{SIC} \citep{KoenkerNgPortnoy1994} (\texttt{detrendr\_SIC}); $\V{\lambda}$ chosen using the validation method with the validation set consisting of every 5th observation (\texttt{detrendr\_valid}); and $\V{\lambda}$ chosen using the proposed scaled eBIC criteria \Eqn{eBIC} \texttt{detrendr\_eBIC}.
+We compare three criteria for choosing the smoothing parameter for quantile trend filtering:  $\V{\lambda}$ chosen using SIC \Eqn{SIC} (\texttt{detrendr\_SIC}); $\V{\lambda}$ chosen using the validation method with the validation set consisting of every 5th observation (\texttt{detrendr\_valid}); and $\V{\lambda}$ chosen using the proposed scaled eBIC criteria \Eqn{eBIC} (\texttt{detrendr\_eBIC}).
 For the second study we also examine the effect of the non-crossing quantile constraint by estimating the quantile trends separately and choosing $\V{\lambda}$ using the proposed scaled eBIC criteria (\texttt{detrendr\_Xing}).

### 2019-04-18 Manuscript/detrendify_aoas.tex
simulation text edits

@@
 
-We compare three criteria for choosing the smoothing parameter for quantile trend filtering:  $\V{\lambda}$ chosen using SIC \Eqn{SIC} (\texttt{detrendr\_SIC}); $\V{\lambda}$ chosen using the validation method with the validation set consisting of every 5th observation (\texttt{detrendr\_valid}); and $\V{\lambda}$ chosen using the proposed scaled eBIC criteria \Eqn{eBIC} (\texttt{detrendr\_eBIC}).
-For the second study we also examine the effect of the non-crossing quantile constraint by estimating the quantile trends separately and choosing $\V{\lambda}$ using the proposed scaled eBIC criteria (\texttt{detrendr\_Xing}).
+We compare three criteria for choosing the smoothing parameter for quantile trend filtering:  $\V{\lambda}$ chosen using SIC \Eqn{SIC} (\texttt{detrendr\_SIC}); $\V{\lambda}$ chosen using the validation method with the validation set consisting of every 5th observation (\texttt{detrendr\_valid}); and $\V{\lambda}$ chosen using the proposed eBIC criterion \Eqn{eBIC} (\texttt{detrendr\_eBIC}).
+For the second study we also examine the effect of the non-crossing quantile constraint by estimating the quantile trends separately and choosing $\V{\lambda}$ using eBIC (\texttt{detrendr\_Xing}).
 We do not include \texttt{detrendr\_Xing} in the first study because the difference in quantiles is large enough that we would not expect the non-crossing constraint to make a difference.
 
-We also compare the performance of our quantile trend filtering method with three previously published methods: \texttt{npqw} which is the quantile-ll method described in \cite{Racine2017}, code was obtained from the author; \texttt{qsreg} in the \texttt{fields} R package and described in \cite{Oh2011} and \cite{nychka1995nonparametric}; \texttt{rqss} available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
-Only our detrend methods guarantee non-crossing quantiles.
-The regularization parameter $\V{\lambda}$ for the \texttt{rqss} method is chosen using a grid search and minimizing the SIC criteria separately for each quantile as described in \cite{KoenkerNgPortnoy1994}, the regularization parameter for \texttt{qsreg} was chosen using generalized cross-validation separately for each quantile based on the quantile criterion \cite{Oh2011}.
+We also compare the performance of our quantile trend filtering method with three previously published methods, none of which guarantee non-crossing quantiles:
+
+\begin{itemize}
+	\item \texttt{npqw}: The local linear quantile method (quantile-ll) described in \cite{Racine2017}. Code was obtained from the author.
+	\item \texttt{qsreg}: Quantile smoothing splines described in \cite{Oh2004period} and \cite{nychka1995nonparametric} and implemented in the \texttt{fields} R package \citep{fields}.
+The regularization parameter was chosen using generalized cross-validation.
+	\item \texttt{rqss}: Quantile trend filtering with $k=1$ available in the \texttt{quantreg} package and described in \cite{KoenkerNgPortnoy1994}.
+The regularization parameter is chosen using a grid search and minimizing the SIC \Eqn{SIC} as described in \cite{KoenkerNgPortnoy1994}.
+\end{itemize}
 
@@
 \end{eqnarray*}
-The errors were independent and had distributions:
+The errors were simulated as independent draws from the following distributions:
 \begin{itemize}
@@
 
-For each method quantile trends were estimated for $\tau = \{0.05, 0.25, 0.5, 0.75, 0.95\}$.
-For each quantile trend and method the root mean squared error was calculated as $\mbox{RMSE} = \sqrt{\frac{1}{n}\sum_{i=1}^n (\hat{q}_{\tau}(t_i) - q_\tau(t_i))^2}$.
+Quantile trends were estimated for $\tau = \{0.05, 0.25, 0.5, 0.75, 0.95\}$ and the root mean squared error was calculated as $\mbox{RMSE}(\tau_j) = \sqrt{\frac{1}{n}\sum_{i=1}^n (\hat{\theta}_{ij} - \theta_{ij})^2}$, where $\theta_{ij}$ is the true value of the $\tau_j$th quantile of $y$ at $t=i$.
 \Fig{quantile_mse} shows the mean RMSE plus or minus twice the standard error for each method, quantile level, and sample size.
-In all three designs the proposed detrend methods are either better than or comparable to existing methods.
-Overall the \texttt{detrend\_eBIC} performs best, and especially in the mixed normal design our methods have lower RMSEs for the 5\textsuperscript{th} and 95\textsuperscript{th} quantiles.
-The \texttt{npqw} method performs particularly poorly in the mixed normal design due to the fact that it assumes the data come from a scale-location model, which is violated in this case.
+In all three designs the proposed \texttt{detrend} methods are either better than or comparable to existing methods.
+Overall the \texttt{detrend\_eBIC} performs best.
+In the mixed normal design, specifically, our methods have lower RMSEs for the 5\textsuperscript{th} and 95\textsuperscript{th} quantiles.
+The \texttt{npqw} method performs particularly poorly on the mixed normal design due to the violation of the assumption that the data come from a scale-location model.
 
@@
 An example dataset with 4 signal peaks is shown in \Fig{peaks_class_eg}.
-One hundred datasets were generated for each $n=\{500, 1000, 2000, 4000\}$.
+One hundred datasets were generated for each $n\in \{500, 1000, 2000, 4000\}$.
 

### 2019-04-19 Manuscript/detrendify_aoas.tex
moved R package to its own repository

@@
 The \texttt{qsreg} method is comparable to the \texttt{detrend\_eBIC} method on the larger datasets, but its performance deteriorates as the data size shrinks.
-The \texttt{npqw} and \texttt{detrend\_valid} methods both perform poorly on this design.
+The \texttt{npqw} and \texttt{rqss} methods both perform poorly on this design.
 
@@
 
-While minimizing RMSE is desirable in general, in our application, the primary metric of success is accurately classifying observations $\VE{y}{i}$ as signal or no signal.
+While minimizing RMSE is desirable in general, in our application, the primary metric of success is accurately classifying observations $\VE{y}{i}$ into signal present or absent.
 To evaluate the accuracy of our method compared to other methods we define true signal as any time point when the simulated peak value is greater than 0.5.
 We compare three different quantiles for the baseline estimation and four different thresholds for classifying the signal after subtracting the estimated baseline from the observations.
-\Fig{peaks_class_eg} shows an illustration of the observations classified as signal after subtracting the baseline trend compared to the ``true signal." Let $\delta_i \in \{0,1\}$ be the vector of true signal classifications and $\hat{\delta}_i \in \{0,1\}$ be the vector of estimated signal classifications, namely $\hat{\delta}_i = \One(\VE{y}{i} - \hat{\theta}_{i} > 0.5)$.
+\Fig{peaks_class_eg} illustrates the observations classified as signal after subtracting the baseline trend compared to the ``true signal".
 
@@
 \end{eqnarray*}
+where $\delta_i \in \{0,1\}$ is the vector of true signal classifications and $\hat{\delta}_i \in \{0,1\}$ is the vector of estimated signal classifications, namely $\hat{\delta}_i = \One(\VE{y}{i} - \hat{\theta}_{i} > 0.5)$.
 We use this metric because our classes tend to be very imbalanced with many more zeros than ones.
@@
 
-Our \texttt{detrend\_eBIC} method performs the best overall in terms of both RMSE (\Fig{peaks_rmse}) and CAA (\Fig{CAA}).
+Our \texttt{detrend\_eBIC} method results in the largest CAA values (\Fig{CAA}) in addition to the smallest RMSE values (\Fig{peaks_rmse}).
 While \texttt{qsreg} was competitive with our method in some cases, in the majority of cases the largest CAA values for each threshold were produced using the \texttt{detrend\_eBIC} method with the 1\textsuperscript{st} or 5\textsuperscript{th} quantiles.
@@
 We do not use class-averaged accuracy to compare the signal classifications because we do not have a reference value to define as the ``true" signal.
-Instead, we compute the variation of information (VI) which compares the similarity between the two classifications.
-Given the signal classifications for SPods a and b, $\delta^{(a)}_i \in \{0,1\}$ and $\delta^{(b)}_i\in\{0,1\}$, the VI is defined as:
+Instead, we compute the variation of information (VI) which compares the similarity between two classifications.
+Given the signal classifications for SPods a and b, $\delta^{(a)}_i \in \{0,1\}$ and $\delta^{(b)}_i\in\{0,1\}$, for $i \in \{1, ..., n\}$ the VI is defined as:
 \begin{eqnarray*}
-	r_{ij} & = & \frac{1}{n}\sum_t \One\left(\delta^{(a)}_i = i  \cap \delta^{(b)}_i = j\right)\\
-	\VI(s_a, s_b) & = & -\sum_{i,j} r_{ij} \left[ \log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \One(\delta^{(a)}_i = i)}\right) +
-	\log \left(\frac{r_{ij}}{\frac{1}{n}\sum_t \One(\delta^{(b)}_i = j)}\right) \right]
+	r_{jk} & = & \frac{1}{n}\sum_i \One\left(\delta^{(a)}_i = j  \cap \delta^{(b)}_i = k\right)\\
+	\VI(a, b) & = & -\sum_{j,k} r_{jk} \left[ \log \left(\frac{r_{jk}}{\frac{1}{n}\sum_i \One(\delta^{(a)}_i = j)}\right) +
+	\log \left(\frac{r_{jk}}{\frac{1}{n}\sum_i \One(\delta^{(b)}_i = k)}\right) \right]
 \end{eqnarray*}
+where $(j,k) \in \{(0,0), (0,1), (1,0), (1,1)\}$.
 The $\VI$ is a distance metric for measuring similarity of classifications and will be 0 if the classifications are identical and increase as the classifications become more different.
@@
 \label{sec:long-app}
-\Alg{admm} was used to remove the baseline drift from the full day of data  (\Fig{raw_spod}) consisting of 86,401 observations per SPod and compared to the series detrended using the \texttt{qsreg} trends estimated using non-overlapping 2 hour windows.
+\Alg{admm} was used to remove the baseline drift from the full day of data  (\Fig{raw_spod}) consisting of 86,400 observations per SPod and compared to the series detrended using the \texttt{qsreg} trends estimated using non-overlapping 2 hour windows.
 As in the shorter illustration, the \texttt{detrendr} method results in generally lower VI scores than the \texttt{qsreg} method (\Fig{vi_day}).
@@
 
-Finally we estimated the quantile trends for 7 days of measurements of two co-located SPods.
+Finally, we estimated the quantile trends for 7 days of measurements of two co-located SPods.
 \Fig{vi_week} demonstrates the improvement in classification similarity when using \texttt{detrendr}.
@@
 
-We have expanded the quantile trend filtering method by implementing a non-crossing constraint and a new algorithm for processing large series, and proposing a modified criteria for smoothing parameter selection.
-Furthermore we have demonstrated the utility of quantile trend filtering in both simulations and applied settings.
-Our ADMM algorithm for large series both reduces the computing time and allows trends to be estimated on series that cannot be estimated simultaneously while our scaled extended BIC criteria was shown to provided better estimated of quantile trends in series with and without a signal component.
+We have expanded the quantile trend filtering method by implementing a non-crossing constraint, a new algorithm for processing large series, and proposing a modified criteria for smoothing parameter selection.
+Furthermore, we have demonstrated the utility of quantile trend filtering in both simulations and applied settings.
+Our ADMM algorithm for large series both reduces the computing time and allows trends to be estimated on series that cannot be estimated simultaneously while our scaled extended BIC criterion was shown to provide better estimates of quantile trends in series with and without a signal component.
 
@@
 
-	\item[R-package for detrend routine:] R-package detrendr containing code to perform the diagnostic methods described in the article. (GNU zipped tar file)
+	\item[R-package for detrend routine:] R-package detrendr containing code to perform the methods described in the article.
+(GNU zipped tar file also available at https://github.com/halleybrantley/detrendr)
 

### 2019-04-19 Code/Fig_app_short.R

@@ -32,8 +32,11 @@ spodFig <- data.frame(time = c(spodPIDs$time, spodPIDs$time, spodPIDs$time),
   gather("type","value", -c(time, node)) %>%
   arrange(time, rev(type))
 
-ggplot(spodFig, aes(x=time, y=value)) + 
-  geom_line(aes(col=type, group=type), alpha=0.5) + 
+ggplot() + 
+  geom_line(data=subset(spodFig, type=="raw"), 
+            aes(col=type, group=type, x=time, y=value)) + 
+  geom_line(data=subset(spodFig, type!="raw"), 
+            aes(col=type, group=type, x=time, y=value), alpha=.8) +
   theme_bw() +
   facet_grid(node~., scales="free")+
   scale_color_manual(breaks = c("raw", "detrendr", "qsreg"), 

### 2019-04-19 Code/detrendr/R/get_criteria.R

@@ -1,57 +1,0 @@
-
-# Criteria for evaluating the smoothness parameter
-#'
-#' \code{get_criteria}
-#'
-#' @param criteria label of criteria to be used, must be one of "eBIC", "SIC", "valid"
-#' @param f_trend matrix of fitted quantile trend(s)
-#' @param y observed data vector
-#' @param tau quantile vector
-#' @param D discrete differencing matrix (for SIC and eBIC)
-#' @param df_tol tolerance for determining degrees of freedom (Dtheta > df_tol)
-#' @param gamma parameter for eBIC
-#' @param validID index of data to be used for validation (for valid method) 
-#' @param yValid validation data (for valid method) 
-#' @export
-get_criteria <- function(criteria, f_trend, y, tau, 
-                         D = NULL, df_tol=1e-9, gamma=1, 
-                         validID = NULL, yValid = NULL){
-  
-  if (criteria == "eBIC" || criteria == "SIC") {
-    n <- length(y)
-    missInd <- which(is.na(y))
-    if (sum(missInd) > 0){
-      resid_trend <- checkloss(y[-missInd]-f_trend[-missInd,,drop=FALSE], tau)
-    } else {
-      resid_trend <- checkloss(y-f_trend, tau)
-    }
-    df <- Matrix::colSums(abs(D%*%f_trend) > df_tol) 
-    if (criteria == "eBIC"){
-      scale_param <- 0.5 - abs(0.5-tau)
-      BIC <- 2*colSums(resid_trend, na.rm=T)/scale_param + log(n)*df +
-        2*gamma*log(choose(nrow(D), df))
-      if (any(sapply(BIC, is.infinite))) {
-        H <- function(x) {
-          return(x*log(1/x) + (1-x)*log(1/(1-x)))
-        }
-        BIC <- 2*colSums(resid_trend, na.rm=T)/scale_param + log(n)*df +
-          2*gamma*nrow(D)*H(df/nrow(D))
-      }
-    } else {
-      BIC <- log(colMeans(resid_trend, na.rm=T)) + log(n)*df/(2*n)
-    }
-  } else if (criteria == "valid"){
-    missInd <- is.na(yValid)
-    if (sum(missInd) > 0){
-      BIC <- colMeans(checkloss(
-        yValid[-missInd]-f_trend[validID[-missInd],,drop=FALSE], tau))
-    } else {
-      BIC <- colMeans(checkloss(
[…]

### 2019-04-19 Code/detrendr/R/get_metric.R

@@ -1,37 +1,0 @@
-#' Get evaluation metrics for classification
-#'
-#' \code{get_metric} Returns classification metric requested
-#'
-#' @param trend estimated trend vector
-#' @param y vector of observations
-#' @param signal vector of true signal classification
-#' @param threshold value for threshold defining signal
-#' @param metric choice of "missclass", "recall", "CAA", "precision" or "F1"
-#' @export
-get_metric <- function(trend, y, signal, threshold, metric){
-  y_adj <- y - trend
-  signal_hat <- as.numeric(y_adj > threshold)
-  if (metric == "missclass"){
-    return(mean(abs(signal-signal_hat), na.rm=T))
-  } else {
-    true_pos <- sum(signal_hat==1 & signal==1, na.rm=T)
-    recall <- true_pos/sum(signal == 1, na.rm=T)
-    if (metric == "recall"){
-      return(recall)
-    } else if (metric == "CAA"){
-      recall2 <- sum(signal_hat==0 & signal==0, na.rm=T) / 
-        sum(signal == 0, na.rm=T)
-      caa <- (recall + recall2)/2
-      return(caa)
-    }
-    precision <- true_pos/sum(signal_hat == 1, na.rm=T)
-    if (metric == "precision"){
-      return(precision)
-    } else if (metric == "F1"){
-      F1 <- 2 * (precision*recall)/(precision + recall)
-      return(F1)
-    } else {
-      stop("Metric must be in missclass, recall, precistion, F1, CAA")
-    }
-  }
-}

### 2019-04-19 Code/detrendr/R/get_trend_spingarn.R

@@ -1,25 +1,0 @@
-#' Get quantile trend using Spingarn's Method
-#'
-#' \code{get_trend_spingarn} Returns the quantile trend vector
-#'
-#' @param y observed data, should be equally spaced, may contain NA
-#' @param tau quantile level at which to evaluate trend
-#' @param lambda penalty paramter controlling smoothness
-#' @param k order of differencing
-#' @param numIter number of iterations to run algorithm 
-#' @export
-get_trend_spingarn <- function(y, tau, lambda, k, numIter = 10000){
-  if (length(tau)!=1){
-    stop("Can only evaluate single quantile")
-  }
-  
-  theta0 <- y
-  n <- length(y)
-  D <- get_Dk(length(y), k)
-  eta0 <- as.numeric(D%*%theta0)
-  M <- Matrix::chol(Matrix::Diagonal(n) + Matrix::crossprod(D))
-  spign_const <- spingarn_multi_step(theta0, eta0, y, D, M, lambda, 
-                                     tau, 1, numberIter=numIter, k)
-  return(spign_const[["theta"]])
-}
-  

### 2019-04-19 Code/detrendr/R/solve_functions.R

@@ -1,110 +1,0 @@
-#' Solves the model using chosen solver
-#'
-#' \code{solve_model} Returns the trend
-#'
-#' @param model  model object containing coefficients and constraints
-#' @param solver solver to be used, options are "gurobi", "Rglpk", and "lpSolve"
-#' @param y vector of observations, only needed if trend = TRUE
-#' @param trend if TRUE returns trend, if FALSE returns residuals
-#' @export
-solve_model <- function(model, solver, y=NULL, trend = TRUE){
-  if (trend & is.null(y)){
-    stop("y required to get trend")
-  }
-  
-  if (solver == "gurobi"){
-    x <- solve_gurobi(model)
-  } else if (solver == "Rglpk") {
-    x <- solve_glpk(model)
-  } else if (solver == "lpSolve"){
-    x <- solve_lp(model)
-  } else if (solver == "ipop"){
-    x <- solve_ipop(model)
-  } else if (solver == "quadprog"){
-    x <- solve_quadprog(model)
-  } else {
-    stop("Solver must be one of 'gurobi', 'Rglpk', 'lpSolve', 'ipop', 'quadprog'")
-  }
-  
-  nT <- model$nT
-  n <- model$n
-  np <- model$np
-  phi <- matrix(0, nrow=n, ncol = nT)
-  
-  if (!is.null(x)){
-    for (i in 1:nT){
-      phi[,i] <- x[(1+np*(i-1)):(n+np*(i-1))] - 
-        x[(n+1 + np*(i-1)):(2*n + np*(i-1))]
-    }
-  }
-
-  if (trend){
-    y[which(is.na(y))] <- 0
-    return(y-phi)
-  } else {
-    return(phi)
-  }
-}  
-
-
-# Wrappers for different solvers
-
-solve_gurobi <- function(model){
-  params <- list(OutputFlag=0)
-  result <- gurobi::gurobi(model, params)
-  if (result$status == "NUMERIC"){
-    print("Optimization was terminated due to unrecoverable numerical difficulties.")
-    result <- gurobi::gurobi(model, params)
-  }
-  return(result$x)
-}
-
-solve_glpk <- function(model_list){
-  result <- Rglpk::Rglpk_solve_LP(obj = model_list$obj, 
-                           mat = model_list$A, 
-                           dir = paste0(model_list$sense, "="), 
-                           rhs = model_list$rhs)
-  return(result$solution)
-}
[…]

### 2019-04-19 Code/detrendr/R/update_windows.R

@@ -1,130 +1,0 @@
-#' Update windows
-#'
-#' \code{update_windows} Minimizes the Lagrangian for each window
-#'
-#' @param w_list Dual variable
-#' @param phiBar_list Consensus variable for phi
-#' @param etaBar_list Consensus variable for eta
-#' @param model_list List of model objects for solver
-#' @param rho ADMM parameter
-#' @param nT number of quantiles
-#' @param solver must be "gurobi" or "quadprog"
-#' @export
-update_windows <- function(w_list, phiBar_list, etaBar_list, 
-                           model_list, rho, nT, solver="gurobi"){
- 
-  model_list <- mapply(update_model, model_list, w_list, phiBar_list,
-                       etaBar_list,
-                       rho=rho, nT=nT, SIMPLIFY = FALSE)
-
-  phi_list <- lapply(model_list, solve_model, solver=solver, trend = FALSE)
-  # If solver didn't succeed use value of phiBar
-  phi_list0 <- mapply(function(x,y) {if(!any(x != 0)){return(y)}else{return(x)}}, 
-                     phi_list, phiBar_list)
-  return(phi_list)
-}
-
-#' Update consensus variable (overlap)
-#'
-#' \code{update_consensus}
-#'
-#' @param phi_list Primal variables
-#' @param windows Matrix indicating window group
-#' @param overlapInd Indices of overlap between windows
-#' @export
-update_consensus <- function(phi_list, windows, overlapInd){
-  nT <- dim(phi_list[[1]])[2]
-  phiBar <- matrix(0, nrow = nrow(windows), ncol = nT)
-  n_windows <- length(phi_list)
-  
-  for (i in 1:n_windows){
-    phiBar[windows[,i],] <- phiBar[windows[,i],] + phi_list[[i]]
-  }
-  phiBar[overlapInd, ] <- phiBar[overlapInd, ]/2
-
-  phiBar_list <- list()
-  for (i in 1:n_windows){
-    phiBar_list[[i]] <- phiBar[windows[,i],,drop=FALSE]
-  }
-  return(phiBar_list)
-}
-
-#' Update consensus variable (overlap)
-#'
-#' \code{get_phiBar}
-#'
-#' @param phiBar_list List of consensus variable
-#' @param windows Matrix indicating window group
-#' @export
[…]

### 2019-04-23 Manuscript/detrendify_aoas.tex
Minor edits.

@@
 The solutions are very close, however, and the differences are essentially immaterial concerning downstream analysis.
-\Fig{windowsfit} provides an illustration of the trends estimated using multiple windows compared with the trends estimated using a single window.
+\Fig{windowsfit} provides an illustration of the trends estimated using multiple windows compared with the trends estimated using a single window; estimates using multiple and single windows are nearly indistinguishable.
 
@@
 We compare three different quantiles for the baseline estimation and four different thresholds for classifying the signal after subtracting the estimated baseline from the observations.
-\Fig{peaks_class_eg} illustrates the observations classified as signal after subtracting the baseline trend compared to the ``true signal".
+\Fig{peaks_class_eg} illustrates the observations classified as signal after subtracting the baseline trend compared to the ``true signal."
 

### 2019-05-08 Manuscript/detrendify_aoas.tex
editing figures for defense presentation

@@
 \section{Acknowledgments}
-The authors greatly appreciate Eben Thoma at the US Environmental Protection Agency for providing the SPod datasets.
-The authors are also thankful for the Oak Ridge Institute of Science Fellowship that partially supported this work.
+The authors greatly appreciate Eben Thoma at the U.S.
+Environmental Protection Agency for providing the SPod datasets.
+This work was supported by the National Science Foundation under grant No.
+1613219 and the National Institutes of Health under grant No.
+R01ES027892.
+The authors are also thankful for the Oak Ridge Institute of Science Fellowship that supported this work.
 Plots were created using ggplot2 \citep{ggplot2}.
+
+\noindent {\large\bf Disclaimer}
+The views expressed in this publication are those of the authors and do not represent the views or policies of the U.S. Environmental Protection Agency.
 
@@
 
-	\item[R-package for detrend routine:] R-package detrendr containing code to perform the methods described in the article.
-(GNU zipped tar file also available at https://github.com/halleybrantley/detrendr)
+	\item[R-package for detrend routine:] R-package detrendr containing code to perform the methods described in the article (GNU zipped tar file).
+Package is also available at \url{https://github.com/halleybrantley/detrendr}.
 

### 2019-05-08 Code/BIC_example.R

@@ -3,7 +3,7 @@
 ################################################################################
 library(tidyverse)
 library(devtools)
-load_all("detrendr")
+load_all("../../detrendr")
 rm(list=ls())
 
 set.seed(9856882)
@@ -40,10 +40,10 @@ for (i in 1:length(lambdaSeq)){
 }
 
 plot.df <- data.frame(df = df_trend[,1])
-plot.df$eBIC <- SIC
-plot.df$BIC_noScale <- SIC2
-plot.df$BIC_scale <- SIC3
-plot.df$SIC <- n*(SIC4+5)
+plot.df$eBIC <- SIC-min(SIC)
+plot.df$BIC_noScale <- SIC2-min(SIC2)
+plot.df$BIC_scale <- SIC3 - min(SIC3)
+plot.df$SIC <- n*(SIC4 - min(SIC4))
 plot.df$lambda <- lambdaSeq
 
 
@@ -55,25 +55,31 @@ ggplot(plot.df, aes(x=log(lambda), y=df)) +
   labs(x=bquote(log(lambda)))
   ggsave("../Manuscript/Figures/df_by_lambda.png", width = 3, height = 3)
 
-plot.df %>% gather("criteria", "value", -c(df, lambda)) %>%
-ggplot(aes(x=log(lambda), y=value, col = criteria)) +
-  geom_line() + 
+colPal <- c('#006d2c', '#2ca25f', '#66c2a4', '#1c9099', '#d8b365',
+              "#c2a5cf", "#9970ab", "#762a83")
+
+BIC.df <- plot.df %>% gather("criteria", "value", -c(df, lambda)) %>%
+  mutate(criteria = factor(criteria, 
+                           levels = c("BIC_noScale", "BIC_scale", "eBIC", "SIC")))
+
+ggplot(BIC.df, aes(x=log(lambda), y=value, col = criteria)) +
+  geom_line(size=1.2) + 
   theme_bw() +
-  scale_color_brewer(palette = "Set1", 
-                     labels = c("BIC no scale", "BIC with scale", "SIC",
-                                "eBIC")) +
-  labs(y = "BIC", col="") +
+  scale_color_manual(values = c('#fd8d3c', '#df65b0', '#006d2c', '#2ca25f'), 
+                     labels = c("BIC no scale", "BIC with scale",
+                                "eBIC with scale", "SIC")) +
+  labs(y = "BIC (shifted)", col="") +
   geom_vline(aes(xintercept = log(lambdaSeq[which.min(SIC)]), 
[…]

### 2019-11-20 Manuscript/response/response.tex
Submitted revision

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amsfonts,amssymb,graphicx}
\usepackage[mathscr]{eucal}
\usepackage{stmaryrd}
\usepackage{amsbsy}
\usepackage{bm}
\usepackage[usenames]{color}
\usepackage[dvipsnames]{xcolor}
\usepackage{parskip} 
\usepackage{natbib}
\usepackage{comment}
%% For \subfloat command
\usepackage{subfig}

\newcommand{\fixme}[1]{{\color{red}\textbf{#1}}}
\newcommand{\T}[1]{\boldsymbol{\mathscr{\MakeUppercase{#1}}}} %tensor

%% Shorthand for Real
\newcommand{\Real}{\mathbb{R}}

\newcommand{\tr}{\operatorname{tr}} % Trace
\newcommand{\Kron}{\otimes} %Kronecker

\newcommand{\E}{\mathrm{E}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\Corr}{\mathrm{Corr}}
\newcommand{\Ehat}{\hat{\E}}
\newcommand{\Covhat}{\hat{\Cov}}
\newcommand{\vect}{\mathrm{vec}}
\newcommand{\vech}{\mathrm{vech}}
\newcommand{\indep}{\;\, \rule[0em]{.03em}{.6em} \hspace{-.25em}
\rule[0em]{.65em}{.03em} \hspace{-.25em}
\rule[0em]{.03em}{.6em}\;\,}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\real}[1]{\Real \mathit{^{#1}}}
\newcommand{\trans}{^{\mbox{\tiny {\sf T}}}}
%\newcommand{\itrans}{^{\mbox{-\tiny {\sf T}}}}
\newcommand{\spn}{\mbox{span}}
\newcommand{\spc}{{\mathcal S}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\cs}{\spc_{Y|\Xbf}}
\newcommand{\Var}{\mathrm{Var}}

%% Vectors
\newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} % vector element
\newcommand{\Vbar}[1]{{\bm{\M{A}r \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}} % vector
\newcommand{\Vn}[2]{\V{#1}^{(#2)}} % n-th vector
\newcommand{\VnE}[3]{{#1}^{(#2)}_{#3}} % n-th vector
\newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}} % vector element

%% Matrices
\newcommand{\Tra}{^{\sf T}} % Transpose
\newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} % matrix element
\newcommand{\MC}[2]{\V{#1}_{#2}}
\newcommand{\Mr}[2]{\V{#1}_{#2}} % matrix row
\newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\MhatC}[2]{\Vhat{#1}_{#2}} % matrix column
\newcommand{\Mbar}[1]{{\bm{\bar \mathbf{\MakeUppercase{#1}}}}} % matrix
\newcommand{\MbarC}[2]{\Vbar{#1}_{#2}} % matrix column
\newcommand{\Mn}[2]{\M{#1}^{(#2)}} % n-th matrix
\newcommand{\Mbarn}[2]{\Mbar{#1}^{(#2)}} % n-th matrix
\newcommand{\Mtilden}[2]{\Mtilde{#1}^{(#2)}} % n-th matrix
\newcommand{\MnTra}[2]{\M{#1}^{(#2){{\sf T}}}} % n-th matrix transpose
\newcommand{\MnE}[3]{\MakeLowercase{#1}^{(#2)}_{#3}} % n-th matrix element
\newcommand{\MnC}[3]{\V{#1}^{(#2)}_{#3}} % n-th matrix column
\newcommand{\MbarnC}[3]{\Vbar{#1}^{(#2)}_{#3}} % n-th matrix column
\newcommand{\MnCTra}[3]{\V{#1}^{(#2){{\sf T}}}_{#3}} % n-th matrix column transpose
\newcommand{\MCTra}[2]{\V{#1}^{{\sf T}}_{#2}} % n-th matrix column transpose
\newcommand{\prox}{\mathop{\rm prox}\nolimits}
\newcommand{\Pcal}{{\mathcal{P}}}
\newcommand{\greekbold}[1]{\mbox{\boldmath $#1$}}
\newcommand{\Sigmabf}{\greekbold{\Sigma}}

%% For comments to each other
\newcommand{\Remark}[2]{{\color{red}Remark from #1: #2}}
\newcommand{\Note}[1]{{{\color{red}{#1}}}}

\newcommand{\Eric}[2]{{\bf {\color{MidnightBlue}@Eric from #1: #2}}\xspace}
\newcommand{\Joe}[2]{{\bf {\color{Plum}@Joe from #1: #2}}\xspace}
\newcommand{\Halley}[2]{{\bf {\color{OliveGreen}@Halley from #1: #2}}\xspace}
\newcommand{\All}[2]{{\bf {\color{Sepia}@All from #1: #2}}\xspace}

\begin{document}
\title{Baseline Drift Estimation for Air Quality Data Using Quantile Trend Filtering}
\author{}
\date{}
\maketitle

\section*{General Remarks}

We thank the referees for their useful feedback and believe the manuscript is improved as a result.

In our revision we have denoted in \color{red}red \color{black} substantial newly added text. For the sake of readability we did not highlight minor edits in the revision.
Finally, we give point-by-point responses to referee comments below.
%\section*{Editor}
%
%\begin{quote}

%\Note{Should we mention here that we address reviewer 2's comments in the file?}
%
%\end{quote}

%\section*{Associate Editor}

\section*{Reviewer 1}

\begin{comment}
Authors present a method for quantile trend filtering with a non-crossing constraint. 
The use the ADMM algorithm to reduce the computing time. 
They compare their method with other approaches by numerical experiments. 
Finally, their approach is used on real data.

I really enjoy the paper. The method is interesting and well introduced. The approach requires different tuning parameters and I think that the choice of these parameters should be more discussed (see my comments below).

Major comments:
- The approach implies different tuning parameters: the penalty $\lambda$, the number of windows $W$, the bounds of the windows (l_w, u_w) and the degree of the polynomials (k). The selection of $\lambda$ is well described in Section 3. However, how the number of windows and the bounds of the windows are selected? In the paper, the degree of polynomials is k=2. Can the information criteria introduced in Section 3 be used to estimate k?

- In Section 3.1, authors said that the method can easily handle missing data. What are the assumption on the process of randomness (MAR, MCAR,...)?

- In Section 3.2, the BIC is introduced. I'm agree that for Bayesian quantile regression, the asymmetric Laplace likelihood is generally used. Does the choice of $\sigma$ impact the results? Moreover, in a Bayesian framework, a penalized regression can often be defined with a specific prior distribution on the parameters (L1-type regression can be defined by using a Laplace distribution as prior). Is it possible to define such prior for the problem defined by Equation (10)?

Minor comments:

- At the end of Section 3.2, maybe more details should be given about the properties of the eBIC for the large parameter spaces (is it more efficient than the BIC)?

- Figure 7,8 and 10 are not really clear. May be they can be replaced by tables.

\end{comment}

1. The approach implies different tuning parameters: the penalty $\lambda$, the number of windows $W$, the bounds of the windows $(l_w, u_w)$ and the degree of the polynomials $(k)$. The selection of $\lambda$ is well described in Section 3. However, how the number of windows and the bounds of the windows are selected?

\begin{quote}
The referee raises a good question. Our motivation for casting the problem as a linked set of estimations over overlapping windows is computational. Ideally there would be just one window to estimate the trend over, rather than $W > 1$ but as the trends in the timing experiments in Figure~5 indicate, using a single window will become computationally impractical if one wishes to fit a model over a very large set of time points. Our numerical experiments show a little less than a linear speed up in the number of windows used. Thus, if one wished to cut the overall running time to fit a family quantile trends by a roughly factor of 10, one would choose in the ballpark of 11 or 12 windows. The bounds of the windows and amount of overlap between windows depends on the application and should be larger than the expected time duration of transient events. We have added additional sentences addressing these points in the revised manuscript at the end of Section 2.
\end{quote}

2. In the paper, the degree of polynomials is $k=2$. Can the information criteria introduced in Section 3 be used to estimate $k$?

\begin{quote}
The expected degrees of freedom in the standard trend-filtering problem is given by the number of knots in $\V{\theta}$ plus $k+1$ \citep{Tib2014}. Consequently, a natural strategy to estimate $k$ is to use the eBIC criterion introduced in Section 3 but with $\nu$ equal to the number of knots in $\V{\theta}$ plus $k+1$. We have added additional sentences addressing these points in the revised manuscript at the end of Section 3.
\end{quote}

3. In Section 3.1, authors said that the method can easily handle missing data. What are the assumption on the process of randomness (MAR, MCAR,...)?

\begin{quote}
This is a good question. It turns out the penalized negative log-likelihood in Section 3.1 corresponds to making a MCAR assumption. We've added a short discussion of this in Section 3.1.
\end{quote}

4. In Section 3.2, the BIC is introduced. I'm agree that for Bayesian quantile regression, the asymmetric Laplace likelihood is generally used. Does the choice of $\sigma$ impact the results?

\begin{quote}
We did not find that the choice of $\sigma$ affected the results. The following figure plots the eBIC criteria against $\sigma$ for an example dataset. The value of $\lambda$ chosen is the same for all values of $\sigma$ compared, but the $\sigma$ used in the analysis (0.1) results in a eBIC value that rises more rapidly after achieving the minimum.

\begin{figure}[h!]
\centering
\includegraphics[width = 0.7\linewidth]{BIC_by_lambda_and_sigma.png}
\end{figure}
\end{quote}

5. Moreover, in a Bayesian framework, a penalized regression can often be defined with a specific prior distribution on the parameters (L1-type regression can be defined by using a Laplace distribution as prior). Is it possible to define such prior for the problem defined by Equation (10)?

\begin{quote}
Yes it is possible to define such a prior for the probelm defined by Equation (10). \cite{faulkner2018locally} proposed a few fully Bayesian variations of the trend filter utilizing shrinkage priors on the $k$th-order differences in values of the unknown target function. They used a Gaussian likelihood for the data corresponding to the squared loss used by \cite{Kim2009}. A fully Bayesian variation of the quantile trend filter could be implemented by substituting the Laplace likelihood for the Gaussian likelihood. %We have added these points to the discussion of prior work. 
We have added these points in the revised discussion.
\end{quote}

\subsection*{Minor comments:}

6. At the end of Section 3.2, maybe more details should be given about the properties of the eBIC for the large parameter spaces (is it more efficient than the BIC)?

\begin{quote}
Thank you for the suggestion. The eBIC has been proven to be model selection consistent in cases where the BIC can be model selection inconsistent. This difference between the eBIC and BIC manifests itself practically in the trend filtering problem in that the BIC selected models tend to be undersmoothed in comparison to the eBIC selected models. We have added sentence highlighting these advantageous properties of the eBIC for large parameter spaces in Section 3.2.

%``A limitation of the BIC, however, is that it may be inconsistent in the high-dimensional setting. This issue arises as a consequence of an implicit assumption in deriving the BIC that all models under consideration are equally likely. Consequently, in the context of variable selection methods such as the lasso, models that include more covariates (up to half of all possible covariates) will have a larger prior probability of being selected under the BIC criterion. In the context of this work, using the BIC may have a tendency to favor the selection of undersmoothed signals, as signals $\V{\theta}$ with more non-zero entries in the vector  $\Mn{D}{k+1}\V{\theta}$ will be assigned higher prior probabilities under the BIC.

%To address this issue, \cite{chen2008} modified the prior probabilities to dampen the prior weight on larger models, or in the context of this work, undersmoothed signals, and proposed 
%... They prove that the eBIC is model selection consistent under mild regularity conditions."
\end{quote}

7. Figure 7,8 and 10 are not really clear. May be they can be replaced by tables.

We have added tables with the results to the supplemental materials but feel that they are too long to include in the manuscript itself.

\section*{Reviewer 2}

\begin{comment}
The paper address the issue of estimating multiple quantile trends simultaneously in time series data. This problem is common in several fields from economics to medicine. The authors focused on the application to low cost air quality sensor. Thus the paper fits with AOAS policy. However, to improve the suitability, I suggest the authors to add more on how to employ their model and algorithm to data beyond to the application presented and which specific question this algorithm can address in different disciplines. While the problem to be addressed is clearly explained, the introduction miss a broader presentation of the topic to attract Statistician readers from differ fields. The authors claim that demixing problems is common to different applications with some references, thus they should comment more on this and reinforce the interest of the paper. Part of the technical material on the ADMM algorithm should be moved to appendices or online supplements.

The paper is correct. The algorithm is tested on simulated and real data supporting the validity of the results. The methodology and the results are well supported by figures and comments all over the test. Authors should check some typos spread in the manuscript.

Overall, the paper is sounds and interesting. It will benefit the community of Statistician working on low cost sensor data, an extremely increasing community. It will be of inspiration for scientists working with time series.
\end{comment}

The paper address the issue of estimating multiple quantile trends simultaneously in time series data. This problem is common in several fields from economics to medicine. The authors focused on the application to low cost air quality sensor. Thus the paper fits with AOAS policy.

8. However, to improve the suitability, I suggest the authors to add more on how to employ their model and algorithm to data beyond to the application presented and which specific question this algorithm can address in different disciplines. While the problem to be addressed is clearly explained, the introduction miss a broader presentation of the topic to attract Statistician readers from differ fields. The authors claim that demixing problems is common to different applications with some references, thus they should comment more on this and reinforce the interest of the paper.

\begin{quote}
We have expanded our discussion of baseline trend filtering applications to include electrocardiogram data, electrooculographic data, mass spectrometry, chromatography, macroeconomics, environmental science, laser-induced breakdown spectroscopy, and astronomy. We have also included one or more references for each example that requires baseline removal. These additions have been made on page 2 of the revision. We have also added a short paragraph at the end of the Section 6 on how our proposed method could potentially be used more generally for robust smoothing of time series data.
%\All{Eric}{Joe and Eric take lead in addressing, but please add any thoughts, Halley.} More applications: mass spec \cite{Du2006}. We don't explore it in this paper but we could use the framework for performing robust trend filtering by focusing on the quantiles close to the median. This might be useful for separating financial time series data into smooth trends and removing transient ups and downs. Maybe add this last observation to the discussion.
\end{quote}

9. Part of the technical material on the ADMM algorithm should be moved to appendices or online supplements.

\begin{quote}
We have moved part of the technical details of the ADMM algorithm to the supplementary materials.
\end{quote}

10. Authors should check some typos spread in the manuscript.

\begin{quote}
Thank you. We have checked and corrected for typos.
\end{quote}

\bibliographystyle{plainnat}
\bibliography{detrendify}
\end{document}

### 2019-11-20 Manuscript/response/main.tex
Submitted revision

--- 2019-05-08 Manuscript/detrendify_aoas.tex
@@
 \usepackage{algorithm}
+\usepackage{comment}
 \usepackage{algpseudocode}
@@
 These sensors can provide an un-calibrated measure of a variety of pollutants in near real time, but deriving meaningful information from sensor data remains a challenge \citep{snyder2013changing}.
-The ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
+For example, the ``SPod" is a low-cost sensor currently being investigated by researchers at the U.S.
 Environmental Protection Agency to detect volatile organic compound (VOC) emissions from industrial facilities \citep{thoma2016south}.
-Due to changes in temperature and relative humidity the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
+Due to changes in temperature and relative humidity, the output signal exhibits a slowly varying baseline drift on the order of minutes to hours.
 Figure \ref{fig:raw_spod} provides an example of measurements from three SPod sensors co-located at the border of an industrial facility.
@@
 
-While this work is motivated by the analysis of data from low cost air quality sensors, the problem of demixing nosy time series into trends and transients is ubiquitous across many fields of study.
-In a wide range of applications that spans chemistry \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{brantley2014mobile}, and medical sciences \citep{pettersson2013algorithm, marandi2015qualitative}, scalar functions of time $y(t)$ are observed and assumed to be a superposition of an underlying slowly varying baseline trend $\theta(t)$, other more rapidly varying components $s(t)$, and noise.
+While this work is motivated by the analysis of data from low cost air quality sensors, the problem of demixing noisy time series into trends and transients is ubiquitous across many applications.
+\color{red} This includes electrocardiogram data \citep{luo2013hierarchical}, electrooculographic data (recording of eye movements) \citep{pettersson2013algorithm, marandi2015qualitative}, mass spectrometry \citep{Du2006,coombes2007pre}, chromatography \citep{Ning2014}, macroeconomics \citep{yamada2017estimating}, environmental science \citep{mecozzi2014polynomial,brantley2014mobile}, laser-induced breakdown spectroscopy \citep{franco2018comparison}, and astronomy \citep{Oh2004period}.
+In most instances, scalar functions of time $y(t)$ are observed and assumed to be a superposition of (i) an underlying slowly varying baseline trend $\theta(t)$, (ii) other more rapidly varying components $s(t)$, and (iii) noise.
+For example, in electrocardiogram data, one wishes to separate a train of transient waveforms $s(t)$ reflecting  the electrical activity of the heart from a nuisance ``baseline wander" $\theta(t)$ due a patient's respiration and other movements.\color{black}
+
 In practice, $y(t)$ is observed at discrete time points $t_1, \ldots, t_n$, and we model the vector of samples $\VE{y}{i} = y(t_i)$ as
@@
 For notational simplicity, for the rest of the paper, we assume that the time points take on the values $t_i = i$, but it is straightforward to generalize to an arbitrary grid of time points.
-
 In some applications, the slowly varying component $\V{\theta}$ is the signal of interest, and the transient component $\V{s}$ is a vector of nuisance parameters.
@@
 
-In the rest of the paper, we detail our quantile trend filtering algorithms (Section \ref{sec:methods}) as well as how to choose the smoothing parameter (Section \ref{sec:lambda_choice}).
-We demonstrate through simulation studies that our proposed model provides better or comparable estimates of non-parametric quantile trends than existing methods (Section \ref{sec:simluation}).
+In the rest of the paper, we detail our quantile trend filtering algorithms (Section \ref{sec:methods}) and investigate how to choose the smoothing parameter (Section \ref{sec:lambda_choice}).
+We demonstrate through simulation studies that our proposed methods provide better or comparable estimates of non-parametric quantile trends than existing methods (Section \ref{sec:simluation}).
 We further show that quantile trend filtering is a more effective method of drift removal for low-cost air quality sensors and results in improved signal classification compared to quantile smoothing splines (Section \ref{sec:application}).
@@
 
-In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, the goal is to estimate the trend in the baseline not the mean.
+In this application, as well as those described in \cite{Ning2014}, \cite{marandi2015qualitative}, and \cite{pettersson2013algorithm}, the goal is to estimate the trend in the baseline, not the mean.
 We can define the trend in the baseline as the trend in a low quantile of the data.
@@
 \end{equation*}
-where $\theta(t)$ is a smooth function of time and $\lambda$ is a tuning parameter that controls the degree of smoothing.
+where $\theta(t)$ is a smooth function of time, and $\lambda$ is a tuning parameter that controls the degree of smoothing.
 \cite{Oh2011} proposed an algorithm for solving the quantile smoothing spline problem by approximating the check function with a differentiable function.
-\cite{Racine2017} propose a method for estimating quantile trends that does employ the check function.
+\cite{Racine2017} propose a method for estimating quantile trends that does not employ the check function.
 In their method, the response is constrained to follow a location scale model and the conditional quantiles are estimated by combining Gaussian quantile functions with a kernel smoother and solving a local-linear least squares problem.
@@
 As with classic quantile regression, the quantile trend filtering problem is a linear program which can be solved by a number of methods.
-We want to estimate multiple quantiles simultaneously and to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$ where $Q$ is the quantile function of $\V{y}$.
-Even if a single quantile is ultimately desired, ensuring non-crossing allows information from nearby quantiles to be used to improve the estimates as we will see in the peak detection experiments in \Sec{peak_detection}.
+We want to estimate multiple quantiles simultaneously and to ensure that our quantile estimates are valid by enforcing the constraint that if $\tau_2 > \tau_1$ then $Q(\tau_2) \ge Q(\tau_1)$, where $Q$ is the quantile function of $\V{y}$.
+Even if a single quantile is ultimately desired, ensuring non-crossing allows information from nearby quantiles to be used to improve the estimates, as we will see in the peak detection experiments in \Sec{peak_detection}.
 Given quantiles $\tau_1 < \tau_2 < \ldots < \tau_J$, the optimization problem becomes
@@
 The additional non-crossing constraints are linear inequalities involving the parameters, so the non-crossing quantile trends can still be estimated by a number of available linear programming solvers.
-We allow for the possibility that the degree of smoothness in the trends varies by quantile by allowing the smoothing parameter to vary with quantile as well.
+We allow for the possibility that the degree of smoothness in the trends varies by quantile by allowing the smoothing parameter to vary with quantile.
 In the rest of this paper, we use $k=2$ to produce piecewise quadratic polynomials and report numerical results using the commercial solver Gurobi \citep{gurobi} and its R package implementation.
@@
 
+\begin{comment}
 \Alg{admm} has the following convergence guarantees.
@@
 \end{eqnarray}
+\end{comment}
 
+The windowed quantile trend filtering optimization problem \Eqn{quantile_windows} opens the door to significant computational gains over solving the original quantile trend filtering problem \Eqn{noncross_trend}.
 The quantile trend filtering problem for a single window is a linear program with $N\times J$ parameters (number of observations by number of quantiles), which can be solved in computational time proportional to $(NJ)^3$.
@@
 We use from one to four windows for each data size with an overlap of 500.
-\Alg{admm} was the stopped when \Eqn{stopping} was satisfied, defining $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
+%\Alg{admm} was terminated using a standard stopping rule for ADMM, which is described in the Supplementary Materials.
+%when \Eqn{stopping} was satisfied, defining $\epsilon_{abs} = 0.01$ and $\epsilon_{rel} = 0.001$.
 \Fig{timing} shows that using 4 windows instead of one on data sizes of 55,000 provides a factor of 3 decrease in computation time.
 The timing experiments were conducted on an Intel Xeon based Linux cluster using two processor cores.
+
+%\color{red} We solve \Eqn{quantile_windows} via an ADMM algorithm \citep{gabay1975dual, glowinski1975approximation, boyd2011distributed}.
+We defer details on the algorithm to the supplementary materials and instead discuss at a high level its advantages.
+The windowed quantile trend filtering optimization problem \Eqn{quantile_windows} opens the door to significant computational gains over solving the original quantile trend filtering problem \Eqn{noncross_trend}.
+%The quantile trend filtering problem for a single window is a linear program with $N\times J$ parameters (number of observations by number of quantiles), which can be solved in computational time proportional to $(NJ)^3$.
+Our proposed ADMM algorithm converts the problem of solving a potentially large linear program into a solving a collection of smaller quadratic programs.
+Consequently, solving a large problem via \Eqn{quantile_windows} should require less computational time than solving \Eqn{noncross_trend}, even if the sub-problems are solved sequentially.
+We demonstrate the advantages of solving \Eqn{quantile_windows} through timing experiments (\Fig{timing}).
+For each data size, 25 datasets were simulated using the peaks simulation design described below.
+Trends for the fifth, tenth, and fifteenth quantiles were fit simultaneously using $\lambda_j = n/5$ for all $j$.
+We use from one to four windows for each data size with an overlap of 500.
+\Fig{timing} shows that using 4 windows instead of one on data sizes of 55,000 provides a factor of 3 decrease in computation time.
+The timing experiments were conducted on an Intel Xeon based Linux cluster using two processor cores.
+Additional details on the timing experiment can be found in the supplementary materials.
+
+\color{red} We close the discussion of \Alg{admm} by addressing how to decide the number of windows and their degree of overlap.
+As the windowed and original quantile trend filtering problems provide nearly indistinguishable estimates, a simple rule of thumb for choosing $W$ is to choose as few windows as one's computational budget permits.
+Regarding the overlap between adjacent windows, the overlap is application dependent and should be larger than the expected time duration of transient events.
+
+Additional details regarding the convergence properties and stopping criteria for \Alg{admm} are in the supplementary materials.
+\color{black}
 
@@
 Each of the criteria we compare is calculated for a single quantile ($\tau_j$).
-Rather than combine results over quantiles, we allow the value of $\lambda$ to vary by quantile resulting in $\V{\lambda} = \{\lambda_1, ..., \lambda_J\}$.
+Rather than combine results over quantiles, we allow the value of $\lambda$ to vary by quantile, resulting in $\V{\lambda} = \{\lambda_1, ..., \lambda_J\}$.
 To choose the best value for each $\lambda_j$, we first estimate all of the quantile trends using $\lambda_j = \lambda$ for all $j$ over a grid of values for $\lambda$.
@@
 &&\underset{\M{\Theta} \in \mathcal{C}}{\text{minimize}}\; \sum_{j=1}^J \left [\tilde{\rho}_{\tau_j}(\V{y} - \V{\theta}_{j}) +
-\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ],
+\lambda_j \lVert \Mn{D}{k+1} \V{\theta}_j \rVert_1 \right ].
 \end{eqnarray}
-which can be solved via \Alg{admm} with trivial modification to the quadratic program sub-problems.
-For each quantile level $j$, we select the $\lambda_j$ that minimizes the hold-out prediction error quantified by $\breve{\rho}_{\tau_j}(\V{y} - \Vhat{\theta}_j(\lambda_j))$ where $\Vhat{\theta}_j(\lambda_j)$ is the solution to \Eqn{validate} and
+\color{red} The modified penalized negative log-likelihood in \Eqn{validate} corresponds to making a missing completely at random (MCAR) assumption and can be solved via \Alg{admm} with a trivial modification to the quadratic program sub-problems.
+Although it is normally considered a strong assumption in general, the MCAR assumption does not present any disadvantages for model selection purposes as the validation subset can be selected by taking a completely random sample from among the observed entries $\VE{y}{i}$.
+\color{black}  For each quantile level $j$, we select the $\lambda_j$ that minimizes the hold-out prediction error quantified by $\breve{\rho}_{\tau_j}(\V{y} - \Vhat{\theta}_j(\lambda_j))$ where $\Vhat{\theta}_j(\lambda_j)$ is the solution to \Eqn{validate} and
 \begin{eqnarray*}
@@
 
-\cite{chen2008} proposed the extended Bayesian Information Criteria (eBIC), specifically designed for large parameter spaces.
+\color{red}
+A limitation of the BIC, however, is that it may be inconsistent in the high-dimensional setting.
+This issue arises as a consequence of an implicit assumption in deriving the BIC that all models under consideration are equally likely.
+Consequently, in the context of variable selection methods such as the lasso, models that include more covariates (up to half of all possible covariates) will have a larger prior probability of being selected under the BIC criterion.
+In the context of this work, using the BIC may have a tendency to favor the selection of undersmoothed signals, as signals $\V{\theta}$ with more non-zero entries in the vector  $\Mn{D}{k+1}\V{\theta}$ will be assigned higher prior probabilities under the BIC.
+
+To address this issue, \cite{chen2008} modified the prior probabilities to dampen the prior weight on larger models, or in the context of this work, undersmoothed signals, and proposed the extended Bayesian Information Criteria (eBIC), specifically designed for large parameter spaces.
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in given mod.
-We used this criteria with $\gamma = 1$, and $P=n-k-1$.
-In the simulation study, we compare the performance of the SIC, scaled eBIC (with $\sigma$ defined above), and validation methods.
+where $P$ is the total number of possible parameters and $\nu$ is the number of non-zero parameters included in a given model.
+\cite{chen2008} prove that the eBIC is model selection consistent under mild regularity conditions.
+We used the eBIC criterion with $\gamma = 1$, and $P=n-k-1$.
+We note that the eBIC could be used to select not only $\lambda$ but also $k$ by plugging in for $\nu$ the expected degrees of freedom in the regular trend filtering model \citep{Tib2014}.
+
+We compare the performance of the SIC, scaled eBIC (with $\sigma$ defined above), and validation methods in our simulation studies below. \color{black}
 
@@
 \end{itemize}
-The true quantile trends and an example simulated data set is show in \Fig{sim1}. One hundred datasets were generated of sizes 300, 500 and 1000.
+The true quantile trends and example simulated data sets are shown in \Fig{sim1}.
+One hundred datasets were generated for each design and size (300, 500, and 1000) considered for a total of 900 datasets.
 
@@
 	\includegraphics[width=\linewidth]{Figures/sim_metrics.png}
-	\caption{RMSE by design, method, quantile and data size. Points and error bars represent mean RMSE $\pm$ twice the standard error.}
+	\caption{RMSE by design, method, quantile and data size for designs without a transient component. Points and error bars represent mean RMSE $\pm$ twice the standard error.}
 	\label{fig:quantile_mse}
@@
 	\includegraphics[width = \linewidth]{Figures/peaks_mse.png}
-	\caption{RMSE by method, quantile, and data size for peaks design.}
+	\caption{RMSE by method, quantile, and data size for peaks design. Points and error bars represent mean RMSE $\pm$ twice the standard error.}
 	\label{fig:peaks_rmse}
@@
 \end{eqnarray*}
-where $\delta_i \in \{0,1\}$ is the vector of true signal classifications and $\hat{\delta}_i \in \{0,1\}$ is the vector of estimated signal classifications, namely $\hat{\delta}_i = \One(\VE{y}{i} - \hat{\theta}_{i} > 0.5)$.
+where $\delta_i \in \{0,1\}$ is the vector of true signal classifications and $\hat{\delta}_i \in \{0,1\}$ is the vector of estimated signal classifications, namely $\hat{\delta}_i = \One(\VE{y}{i} - \hat{\theta}_{i} > h)$, where $h$ is the chosen threshold.
 We use this metric because our classes tend to be very imbalanced with many more zeros than ones.
@@
 SPod c contains some missing values that were interpolated before the trends were estimated.
-\texttt{qsreg} is more influenced by the signal component apparent on both nodes resulting in under-smoothing.}
+\texttt{qsreg} is more influenced by the signal component apparent on all nodes resulting in under-smoothing.}
 	\label{fig:short-trends}
@@
 
-We have expanded the quantile trend filtering method by implementing a non-crossing constraint, a new algorithm for processing large series, and proposing a modified criteria for smoothing parameter selection.
+We have expanded the quantile trend filtering method by implementing a non-crossing constraint, introduced a new algorithm for processing big time  series data, and proposing a modified criteria for smoothing parameter selection.
 Furthermore, we have demonstrated the utility of quantile trend filtering in both simulations and applied settings.
-Our ADMM algorithm for large series both reduces the computing time and allows trends to be estimated on series that cannot be estimated simultaneously while our scaled extended BIC criterion was shown to provide better estimates of quantile trends in series with and without a signal component.
+Our ADMM algorithm for large series both reduces the computing time and allows trends to be estimated on series that cannot be estimated simultaneously, while our scaled extended BIC criterion was shown to provide better estimates of quantile trends in series with and without a signal component.
 
 In our application to low cost air quality sensor data, we have shown that the baseline drift in low cost air quality sensors can be removed through estimating quantile trends, but the data size was too large for existing methods to be computationally feasible.
-While \texttt{qsreg} cannot feasibly handle more than a few hour windows of data, our new methods were able to process 24 hours simultaneously and deliver signal classifications that were more consistent between the two sensors for a week of data (168 hours).
-
+While \texttt{qsreg} cannot feasibly handle more than a few hours data, our new methods were able to process 24 hours simultaneously and deliver signal classifications that were more consistent between the two sensors for a week of data (168 hours).
 In the future, quantile trend filtering could be extended to observations measured at non-uniform spacing by incorporating the distance in covariate spacing into the differencing matrix.
 It could also be extended to estimate smooth spatial trends by a similar adjustment to the differencing matrix based on spatial distances between observations.
+
+\color{red} While we focus on obtaining a quantile trend filtering estimator by solving an optimization problem in this work, another potential extension would be to investigate a fully Bayesian formulation of quantile trend filtering.
+\cite{faulkner2018locally} proposed several fully Bayesian variations of the trend filter by utilizing Laplace and Horseshoe shrinkage priors on the $k$\textsuperscript{th}-order differences in values of the unknown target function.
+They used a Gaussian likelihood for the data corresponding to the squared loss used by \cite{Kim2009}.
+A fully Bayesian variation of the quantile trend filter could be implemented by building on the work of \cite{faulkner2018locally} by substituting an asymmetric Laplace likelihood for the Gaussian likelihood.
+\color{black}
+
+\color{red} Finally, we note that quantile regression has often been used to perform robust estimation in the presence of outliers.
+Indeed, one could pose our peak detection problem as an outlier detection problem where outliers could only occur as extreme positive and sudden shifts from a smooth and slowly varying baseline.
+Consequently, an important potential use case of our quantile trend filtering method is in robust smoothing of time series data.
+We leave this and the other extensions discussed above for future work.\color{black}
+
+An R-package \texttt{detrendr} containing code to perform the methods described in the article is available at \url{https://github.com/halleybrantley/detrendr}.
 
@@
 Environmental Protection Agency for providing the SPod datasets.
-This work was supported by the National Science Foundation under grant No.
-1613219 and the National Institutes of Health under grant No.
-R01ES027892.
+The authors also thank Chandni Malhotra for help with preliminary numerical experiments.
+This work was supported by the National Science Foundation under grant No.\@ 1613219 and the National Institutes of Health under grant No.\@ R01ES027892.
 The authors are also thankful for the Oak Ridge Institute of Science Fellowship that supported this work.
@@
 
+\begin{comment}
 \bigskip
@@
 \end{description}
+\end{comment}
+
+\begin{supplement}[id=suppA]
+  \sname{Supplement A}
+  \stitle{ADMM Algorithm}
+  \slink[doi]{COMPLETED BY THE TYPESETTER}
+  \sdatatype{.pdf}
+  \sdescription{The supplementary materials contain additional details on the ADMM algorithm as well as tables summarizing the results plotted in Figures 7, 8, and 10.}
+\end{supplement}
 
